### 2020-10-23 draft_rsfmri.Rmd
rename R project

---
title: "Manuscript_rsfMRI"
author:
- xx [1][2][3]
- xx [2]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
  Day Clinic for Cognitive Neurology, University Clinic Leipzig

output:
  bookdown::pdf_book: 
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: yes
bibliography: "bibliography.bib" 
---

```{r "Package loading", include=FALSE}
library(plyr)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(haven)
library(psych)
library(patchwork)
library(lsr) 
library(wesanderson)
library(cowplot)
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4)
options(knitr.table.format = "latex") 
# put the path to this file here (working directory for Rmarkdown):
opts_knit$set(root.dir = "/data/pt_02161/Publications/Papers/manuscript_rsfmri/")
```
\newpage
# Abstract

\newpage
# Introduction

<!-- BARS & weight loss --> Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018]. On the other hand, dietary interventions such as caloric restriction and related metabolic improvements and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018].
Bariatric surgery is one option to treat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions and co-morbidities, such as diabetes. 
<!-- BARS --> Yet, little is known about the potential beneficial effects of bariatric surgery on brain structure and function in this population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
<!-- Reward network --> In lean participants, food valuation processes in decision-making have been linked to brain activity and activity within the reward network, including ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
<!-- Default mode network -->

<!-- H1 Reward network --> In this study, we aim to investigate possible dynamics of within-reward network connectivity before and after bariatric surgery. Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity.
<!-- H2 Default mode network --> In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN). Functional connectivity of this network has been shown to be linked to obesity [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].

<!-- biomarker --> Furthermore, we aim to explore potential underlying mechanisms using blood- and stool-derived biomarkers of possible mediating metabolic pathways. For example, changes in leptin and insulin levels are hypothesized to mediate responses in brain motivation-reward areas [@Jastreboff_2014; @Kullmann_2017; @Kullmann_2011; @Kullmann_2017; @Tiedemann_2017]. Therefore, we aim to investigate whether changes in central leptin or insulin sensitivity (measured using peripheral leptin, glucose and insulin levels in combination with metabolic status) or other metabolic parameters could mediate changes in functional connectivity within the brain reward network.

\newpage
# Methods

```{r Load data, include=FALSE}
source("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /adi_analysis/get_subject_ID_group_tp_forSwE.R")
final = save_txt_for_swe(IG_only = FALSE, only2tp = FALSE)
opts_knit$set(root.dir = "/data/pt_02161/Publications/Papers/manuscript_rsfmri/")

# sample size per group
subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each Subject
table(final[subj_idx,'condition'])

# sex of participants
subj_df <- final[subj_idx,]
subj_df$Sex <- plyr::mapvalues(subj_df$Sex, from = c("-1", "1"), to = c("f", "m"))

# time points count per participant
freqtable <- as.data.frame(table(final$subj.ID))

dfw <- tidyr::pivot_wider(data = final,
                          id_cols = c("subj.ID","condition","Sex","Age_BL"),
                          names_from = "tp",
                          values_from = c("BMI","logmFD")
)
```

## Study design

## Sample and study design

The ADIPOSITAS-Study conducted in the Charité Berlin was an observational, prospective and longitudinal design described in detail by @Prehn_2020. The sample is identical with subsample of this analysis that was replenished with participants taking part in a second follow-up fMRI session 12 months post-surgery/ after baseline. Participants with at least one data point were included in the analysis.
<!-- Descriptive Statistics --> The analysis was done only on morbidly obese individuals with fMRI data (n=`r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; mean age `r round(mean(subj_df$Age_BL),2)` years $\pm$ SD `r round(sd(subj_df$Age_BL),2)`, that either underwent surgery (n = `r nrow(filter(subj_df,condition == "IG"))`, `r nrow(filter(subj_df,condition == "IG" & Sex == "f"))` female) or were waiting list control (n = `r nrow(filter(subj_df,condition == "KG"))`, `r nrow(filter(subj_df,condition == "KG" & Sex == "f"))` female).
Measures were taken at baseline (BL), 6 (FU1) and 12 (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tab4descrptives)). `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.

```{r tab4descrptives, echo=FALSE}
# table for descriptive statistics
descr <- psych::describeBy(dfw[,5:ncol(dfw)], group=dfw$condition,mat=T)
descr$predictor <- c(rep("BMI",6),rep("logmFD",6))
descr$tp <- car::recode(descr$vars, "c(1,4)='fu2';c(2,5)='fu';c(3,6)='bl'")
descr$group <- descr$group1
descr <- descr[,c("predictor","tp","group","n","mean","sd")]
rownames(descr) <- NULL

descr <- kableExtra::kbl(descr[with(descr, order(predictor, tp)), ],
                         format = "latex",
                         booktabs = T,
                         caption = "BMI and mean FD of participants acquired in each condition and time point",)
descr
```

## fMRI data acquisition

MRI images were retrieved with a 12-channel head coil of a 3 Tesla Siemens Trio system.
<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, TR = 1900 ms, TE = 2.52 ms, a = 9, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes. The total acquisition time was 5:45 minutes.

## Data preproceessing

Data preprocessing was conducted with FSL and Freesurfer.
<!-- minimal preprocessing --> Preprocessing pipeline included removal of first four volumes and a transformation of the blood oxygen level dependent (BOLD) time series into the subject's anatomical space. This transform combines parameters from motion correction (FSL's MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL's FUGUE) and coregistration to the subject's anatomical template space (using Freesurfer's bbregister and the transform from individual to longitudinal template space in Freesurfer's longitudinal processing stream).
<!-- Aroma + CC/ + GSR --> Volumes were further preprocessed with ICA-AROMA and compCor (CC) and, optionally, with a subsequent global signal regression (GSR) [@Ciric_2017; @Parkes_2018] to reduce the effect of physiological noise and head motion on the connectivity estimates.
As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but is known to introduces spatial dependency and spurious correlations, we performed GSR after ICA-AROMA in a separate sensitivity analysis.
<!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC). Subsequently, r-to-Z transform was applied to the correlation (usin nilearn, nipype and python). 
Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).
During analyses, beyond an implicit mask, the MNI resampled brain mask of the MNI ICBM 2009 atlases was used as explicit mask.

### Quality Assessment

<!-- exclusion criteria --> Participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g. caused Freesurfer to fail or perform very poorly, i.e. rated "Failed" according to @Klapwijk_2019) were excluded because a subject's motion is highly similar across scans and we thus suspect the fMRI data to also suffer from motion artifacts in these subjects. 
As the expected average head motion is high [@Hodgson_2016] we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample. 
<!-- motion correction --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018]. As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].

### Outcomes

<!-- Computation of whole brain FC --> Main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
<!-- computation of FD --> Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012 and log-transformed.

## Analysis

```{r calculate VIF, include=FALSE}
# generate random vector 
final$randny <- rnorm(nrow(final), mean = 0, sd = 1)
m_grouptime <- lm("randny ~ group + tp + Sex + Age_BL + logmFD", data = final)
m_bmi <- lm("randny ~ avgBMIc + cgnBMI + Sex + Age_BL + logmFD", data = final)
m_fdbmi <- lm("randny ~ avgBMIc + cgnBMI + avgFDc + cgnFD", data = final)

# variance inflation factor (VIF) - i.e.loss of precision implied by the observed level of collinearity/multicollinearity: all_vif = 1/tol
car::vif(m_grouptime)
car::vif(m_bmi)
car::vif(m_fdbmi)

# tolerance - take each predictor in turn and regress that variable (treating it as the outcome variable) on  theremaining q-1 predictors, then calculate the proportion of variance each predictor shares with other predictors in the full model all_tol = 1 - R²(Xi)
1/car::vif(m_grouptime)
1/car::vif(m_bmi)
1/car::vif(m_fdbmi)

psych::corr.test(final[,c("avgBMIc", "cgnBMI", "avgFDc", "cgnFD")], use = "pairwise")
```

Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].

<!-- Model setup --> Two models were examined. The first model was aimed at testing the interaction of time and group: $FC = group + time + group*time$. The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$. These two models were run for both, the PCC and the NAcc, resulting four different models.
The marginal model set up with the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. It therefore accomodates any repeated measurement covariance structures. We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size and therefore accommodating for heteroscedasticity.
<!-- Design matrix --> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends. The resulting flexible factorial design contained one regressor for each time point per group. For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
With a model containing average BMI and longitudinal change per subject, we investigated potential effects .............. on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
<!-- Nuisance covariates --> Age, sex, and mean framewise displacement (FD) were entered into the first and second model as nuisance variables. Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points. However, as FD seems to share variance with BMI [@Beyer_2019], its inclusion may be disputed.
<!-- Post hoch analyses  --> For that reason, we performed each analysis seperately one with mean FD and once without. Differences in results will be reported. For a better understanding of the unique contribution of average and longitudinal change in BMI and FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject BMI + within-subject BMI + between-subject FD + within-subject FD$.

<!-- Inference --> All parametric inferences on activation were drawn based on False Discovery Rate (FDR) adjusted p-values $\leq 0.05$. As collinearity may cause problems we calculated an approximation of the variance inflation factor (VIF) which was shown to be minimal for all models except the third exploratory. Here, there was minor variance inflation for longitudinal change in both BMI and FD was moderately positively correlated (MLM HERE $r$ = `r cor(final$cgnFD,final$cgnBMI)`, $VIF$ = `r round(car::vif(m_fdbmi),2)`). <!-- Problem(?): correlation not does not consider multilevel - compute linear mixed model -->
To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference. 
In contrast to the parametric estimation, bootstrap is adapted to use family-wise error correction, and unlike permutation based framework appropriate for longitudinal data.

\newpage
# Results
## Results for brain mask
<!-- Group*Time Interaction -->
There was no interaction of group and time point regarding the resting state activity in neither the default mode network nor the reward center, nor was there a significant main effect for any of the covariates of interest in voxelwise inference with FDR correction.
<!-- average and individual BMI -->

<!-- FD -->
Surprisingly,

## Results for grey matter mask
<!-- t-tests to check for network presense --> We ran seperate one-sample t-test for each time point to check whether the assumed reward and default mode network was at all identifyable.
<!-- individual statistical maps -->

\newpage
# Discussion

<!-- Multicollinearity --> Prior to the analysis we assessed multicollinearity of our predictors and correlations between our individual change in BMI and FD. As described in @Monti_2011 in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled. Confidence in $\beta$ estimates will suffers and estimated can be erratic, depending on which regressors included in the model. However, correlation between both longitudinal changes were expected as they share ...............

<!-- Control for satiety? --> A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus. After the intervention, the functional connectivity was higher in bariatric patients. It decreased with after a meal while dietary patients showed an increase [@Lepping_2015].

### 2020-10-26 draft_rsfmri.Rmd
added absolute path import

@@
 # put the path to this file here (working directory for Rmarkdown):
-opts_knit$set(root.dir = "/data/pt_02161/Publications/Papers/manuscript_rsfmri/")
+opts_knit$set(root.dir = getwd())
 ```
@@
 ```{r Load data, include=FALSE}
-source("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /adi_analysis/get_subject_ID_group_tp_forSwE.R")
+source("get_subject_ID_group_tp_forSwE.R")
 final = save_txt_for_swe(IG_only = FALSE, only2tp = FALSE)
-opts_knit$set(root.dir = "/data/pt_02161/Publications/Papers/manuscript_rsfmri/")
 
@@
                           names_from = "tp",
-                          values_from = c("BMI","logmFD")
-)
+                          values_from = c("BMI","logmFD"))
 ```

### 2020-10-26 get_subject_ID_group_tp_forSwE.R

@@ -2,10 +2,18 @@ library(dplyr)
 library(tidyr)
 
 save_txt_for_swe <- function(IG_only, only2tp){
-  parentdir <- "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files"
+  # save original working directory 
+  original_wd <- getwd()
+  
+  # import absolute path
+  abs_path <- read.csv("abs_path.csv", header=FALSE, stringsAsFactors=FALSE)
+  
+  # set working directory
+  swe_path <- "Analysis/Project2_resting_state/seed-based/Second_level /SwE_files"
+  parentdir <- file.path(abs_path, swe_path, fsep = .Platform$file.sep)[1]
   setwd(parentdir)
   
-  files=read.table("scans_PCC_CC_z.txt")
+  files=read.table("../SwE_files/scans_PCC_CC_z.txt")
   #split info from path in txt
   for (i in 1:nrow(files)){
     #print(strsplit(toString(rs_QA[i,1]),'_')[[1]][1])
@@ -17,8 +25,8 @@ save_txt_for_swe <- function(IG_only, only2tp){
   }
   files$V1 <- NULL
   
-  #load group info and merge with path info
-  info_file=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
+  #load group info and merge with path info 
+  info_file=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv") # CAVE: info file elsewhere !!!
   group_info=info_file[,c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI")]
   condition=merge(files, group_info[!is.na(group_info$condition),], by=c("subj.ID","tp"))
   
@@ -215,6 +223,9 @@ save_txt_for_swe <- function(IG_only, only2tp){
               file='cgnFD.txt')
   write.table(final$meanFD, col.names=FALSE, row.names=FALSE,quote=FALSE,
               file='meanFD.txt')
+  # reset to original working directory
+  setwd(original_wd)
+  
   return(final)
 }
 
@@ -223,5 +234,3 @@ save_txt_for_swe <- function(IG_only, only2tp){
 save_txt_for_swe(IG_only = FALSE, only2tp = FALSE)
 # save txt first only IG (under different directory as specified in the function)
 # save_txt_for_swe(IG_only = TRUE)
-
-

### 2020-10-26 run_or_display.m

@@ -61,10 +61,11 @@
 % results will only then change, if OVERWRITE is set on true.
 % define relevant input and output directories
 %% ========================================================================
-
-% create a struct, with all important parameters
+% import absolute path for project
 ABS_DIR = readcell("abs_path.txt");
 ABS_DIR = ABS_DIR{1};
+
+% create a struct, with all important parameters
 param.OUT_DIR = fullfile(ABS_DIR,'/Results/Project2_resting_state/connectivity/Analysis/'); %'/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/preliminary_analysis/'; 
 param.INFO_DIR = fullfile(ABS_DIR,'/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files/');
 param.MASK_DIR = fullfile(ABS_DIR, '/Analysis/Project2_resting_state/seed-based/Brain_masks/');

### 2020-10-27 draft_rsfmri.Rmd
Results for GroupTime and BMI

@@
 
+<!-- to read in literature -->
+... 
+group effect because direct effect of intervention surgery -> rsfmri (cares less, less preoccupation with hunger, less/more reward sensitivity) vs.
+sugery -> BMI -> rsfmri (neural signalling, fatty tissue)
+...
+
 <!-- biomarker --> Furthermore, we aim to explore potential underlying mechanisms using blood- and stool-derived biomarkers of possible mediating metabolic pathways.
@@
 Participants with at least one data point were included in the analysis.
-<!-- Descriptive Statistics --> The analysis was done only on morbidly obese individuals with fMRI data (n=`r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; mean age `r round(mean(subj_df$Age_BL),2)` years $\pm$ SD `r round(sd(subj_df$Age_BL),2)`, that either underwent surgery (n = `r nrow(filter(subj_df,condition == "IG"))`, `r nrow(filter(subj_df,condition == "IG" & Sex == "f"))` female) or were waiting list control (n = `r nrow(filter(subj_df,condition == "KG"))`, `r nrow(filter(subj_df,condition == "KG" & Sex == "f"))` female).
-Measures were taken at baseline (BL), 6 (FU1) and 12 (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tab4descrptives)).
+<!-- Descriptive Statistics --> The analysis was done only on morbidly obese individuals with fMRI data (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; mean age `r round(mean(subj_df$Age_BL),2)` years $\pm$ SD `r round(sd(subj_df$Age_BL),2)`, that either underwent surgery (n = `r nrow(filter(subj_df,condition == "IG"))`, `r nrow(filter(subj_df,condition == "IG" & Sex == "f"))` female) or were waiting list control (n = `r nrow(filter(subj_df,condition == "KG"))`, `r nrow(filter(subj_df,condition == "KG" & Sex == "f"))` female).
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tab4descrptives)).
 `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
@@
 <!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, TR = 1900 ms, TE = 2.52 ms, a = 9, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
-<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes.
+<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes<!-- dimension = [60;72;60] ???
+-->.
 The total acquisition time was 5:45 minutes.
@@
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-However, as FD seems to share variance with BMI [@Beyer_2019], its inclusion may be disputed.
-<!-- Post hoch analyses  --> For that reason, we performed each analysis seperately one with mean FD and once without.
+However, as change in FD and change in BMI share variance [@Beyer_2019].
+Therefore, inclusion of mean FD may be disputed.
+<!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
 Differences in results will be reported.
 For a better understanding of the unique contribution of average and longitudinal change in BMI and FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject BMI + within-subject BMI + between-subject FD + within-subject FD$.
+These post hoc analyses deviate from the preregistration and should be considered to be exploratory.
 
@@
 <!-- Problem(?): correlation not does not consider multilevel - compute linear mixed model -->
-To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference. 
+To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
 In contrast to the parametric estimation, bootstrap is adapted to use family-wise error correction, and unlike permutation based framework appropriate for longitudinal data.
@@
 # Results
-## Results for brain mask
-<!-- Group*Time Interaction -->
-There was no interaction of group and time point regarding the resting state activity in neither the default mode network nor the reward center, nor was there a significant main effect for any of the covariates of interest in voxelwise inference with FDR correction.
-<!-- average and individual BMI -->
 
-<!-- FD -->
-Surprisingly,
+The same analyses as described a priori were carried out with a grey matter mask replacing the whole brain mask.
+Focusing only on grey matter instead if the whole brain reduced the number of tests (voxels of interest) from $52378$ to $51796$.
+Computation with the grey matter mask were executed post hoc and deviate from the preregistration.
 
-## Results for grey matter mask
-<!-- t-tests to check for network presense --> We ran seperate one-sample t-test for each time point to check whether the assumed reward and default mode network was at all identifyable.
+## Manipulation check
+
+<!-- t-tests to check for network presence --> We ran separate one-sample t-test for each time point to check whether the assumed reward and default mode network was at all identifiable.
 <!-- individual statistical maps -->
+
+## Interaction effect of intervention and Time
+
+<!-- Brain mask: Group*Time Interaction -->
+There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
+Nor was there a significant main effect for any of the covariates of interest (time, group) in voxelwise inference with FDR correction.
+Results obtained for parametric estimation and non-parametric wild bootstrap did not differ.
+When reducing the number of nuisance covariates by leaving out mean FD, connectivity was found between the PCC with one other voxel.
+In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm]) yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
+However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{FWEcorr} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!
+-->.
+There was no other significant co-activation for any other region of interest for neither of the preprocessing variants.
+
+<!-- GM mask: Group*Time Interaction -->
+Parametric and non-parametric significance test for the model testing for group-time-interaction did not show any different results when using a grey matter mask.
+Again, parametric estimation and non-parametric estimation differed in regard to the group-time-interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
+While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{FDRcorr} = 0.998$)<!-- also FWE-corr n.s -->.
+
+## Effect of BMI
+
+<!-- Brain mask: average and individual BMI -->
+<!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
+There was no significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
+Omitting mean FD as nuisance covariate did not change these results. 
+<!-- BMIonlyIG --> Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was a significant effect of change in BMI on co-activation between the NAcc and a voxel ($-18,-42,-9$ [mm]) belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in connectivity maps without GSR).
+However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)<!-- but FWE-corr sig?!
+-->.
+The effect also disappeared when running those tests without mean FD as nuisance covariate.
+<!-- BMI2tp --> Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the regions of interest for neither of the preprocessing variants.
+Omitting mean FD did not change these results.
+
+<!-- GM mask: average and individual BMI -->
+Results obtained from analyses with the grey matter masks were similar to those reported for the grey matter masks; test statistics for the significant longitudinal effect differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
+In regard to the exploratory expeditions on subsamples, consistent with previous results, no significant co-activation was found.
+
+## Further investigation of the effect of FD
+
+<!-- Brain mask: FD --> Surprisingly, 
+<!-- GM mask: FD -->
+
+## Target regions of the default mode and reward network
+
+exploratory expeditions
+deviated from the preregistration 
+were executed post hoc
 
@@
 # Discussion
+<!-- Description of analyses carried out --> Marginal models were used to adequately capture longitudinal trends in resting state connectivity of in an intervention group and the waiting control group.
+
+With the model on group-time-interaction, we investigated average effects of the intervention as well as average changes over time while controlling for individual variance in resting-state connectivity over time and at baseline.
+<!-- correct interpretation of marginal models?
+--> The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI
+
+To summarize, there were few significant co-activation emerged from the marginal models.
+Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
+Ventures to verify effects by recalculating with non-parametric Wild bootstrap revealed that effects were not robust.
+This suggests that significant results are likely to be false positive.
 
@@
 It decreased with after a meal while dietary patients showed an increase [@Lepping_2015].
+
+<!-- Incomplete data --> how many had three/ two/ one time point?

### 2020-11-03 draft_rsfmri.Rmd
result section FD model

@@
 
-<!-- Brain mask: FD --> Surprisingly, 
+<!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
+Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only<!-- or for  the first two time points only -->.
+
 <!-- GM mask: FD -->
+
+<!-- Neurovault --> 
+Contrasts images are published on NeuroVault: ......
 
@@
 <!-- Incomplete data --> how many had three/ two/ one time point?
+\newpage

### 2020-11-13 draft_rsfmri.Rmd
more on methds, BMI figure, sample distribution as table

@@
   Day Clinic for Cognitive Neurology, University Clinic Leipzig
-
 output:
@@
 bibliography: "bibliography.bib" 
+editor_options: 
+  chunk_output_type: console
 ---
@@
 library(kableExtra)
-library(haven)
+library(ggplot2)
+library(RColorBrewer)
+# library(haven)
 library(psych)
@@
 library(wesanderson)
-library(cowplot)
+#library(cowplot)
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4)
@@
 On the other hand, dietary interventions such as caloric restriction and related metabolic improvements and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018].
-Bariatric surgery is one option to treat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions and co-morbidities, such as diabetes. 
-<!-- BARS --> Yet, little is known about the potential beneficial effects of bariatric surgery on brain structure and function in this population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+
+<!-- BARS --> Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions and co-morbidities, such as diabetes. 
+Yet, little is known about the potential beneficial effects of bariatric surgery on brain structure and function in this population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+
 <!-- Reward network --> In lean participants, food valuation processes in decision-making have been linked to brain activity and activity within the reward network, including ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
@@
 <!-- to read in literature -->
+<!-- something about surgeries -->
+Roux-en-Y gastric bypass (RYGB) is the most common procedure, as it not only restricts the food intake by reducing stomach volume, but also affects the absorption of nutrients and hormonal profile [@].
+Other common procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption.
+
 ... 
-group effect because direct effect of intervention surgery -> rsfmri (cares less, less preoccupation with hunger, less/more reward sensitivity) vs.
+group effect because direct effect of intervention surgery -> rsfmri (cares lessfreq, less preoccupation with hunger, less/more reward sensitivity) vs.
 sugery -> BMI -> rsfmri (neural signalling, fatty tissue)
 ...
-
-<!-- biomarker --> Furthermore, we aim to explore potential underlying mechanisms using blood- and stool-derived biomarkers of possible mediating metabolic pathways.
-For example, changes in leptin and insulin levels are hypothesized to mediate responses in brain motivation-reward areas [@Jastreboff_2014; @Kullmann_2017; @Kullmann_2011; @Kullmann_2017; @Tiedemann_2017].
-Therefore, we aim to investigate whether changes in central leptin or insulin sensitivity (measured using peripheral leptin, glucose and insulin levels in combination with metabolic status) or other metabolic parameters could mediate changes in functional connectivity within the brain reward network.
+possibly leading to 
+Adipose tissue, gut hormones and gut microbiota may play a mediating role effects for the interaction between body weight and (brian)
 
@@
 source("get_subject_ID_group_tp_forSwE.R")
-final = save_txt_for_swe(IG_only = FALSE, only2tp = FALSE)
+final = save_txt_for_swe(group = "all", tp = "all")
 
-# sample size per group
-subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each Subject
-table(final[subj_idx,'condition'])
+# sample size
+#nrow(subj_df)
+subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each subject
 
-# sex of participants
+# completed time points
+freqtable <- as.data.frame(table(final$subj.ID))
+tpfreq <- table(freqtable$Freq)
+
+# overview distribution of data points
+tp1 <- as.character(freqtable$Var1[freqtable$Freq == 1])
+tp2 <- as.character(freqtable$Var1[freqtable$Freq == 2])
+tp3 <- as.character(freqtable$Var1[freqtable$Freq == 3])
+tp1_tab <- table(subset(final, subj.ID %in% tp1)$tp, subset(final, subj.ID %in% tp1)$condition)
+tp2_tab <- table(subset(final, subj.ID %in% tp2)$tp, subset(final, subj.ID %in% tp2)$condition)[c(1,3),]
+tp3_tab <- table(subset(final, subj.ID %in% tp3)$tp, subset(final, subj.ID %in% tp3)$condition)[1,]
+tab_sample <- data.frame(rbind(tp1_tab, tp2_tab, tp3_tab))
+rownames(tab_sample) <- c("only 0","only 6","only 12","0 and 6","6 and 12","complete data")
+
+# distribution of condition and sex
 subj_df <- final[subj_idx,]
 subj_df$Sex <- plyr::mapvalues(subj_df$Sex, from = c("-1", "1"), to = c("f", "m"))
-
-# time points count per participant
-freqtable <- as.data.frame(table(final$subj.ID))
+cs <- addmargins(table(subj_df$Sex,subj_df$condition))
 
@@
 ```
-
+```{r tabsample, echo=FALSE}
+tab_sample <-
+  knitr::kable(
+    tab_sample,
+    row.names = TRUE,
+    format = "latex",
+    booktabs = T,
+    caption = "Distribution of data points at months after intervention")
+tab_sample
+```
 ## Study design
@@
 
-The ADIPOSITAS-Study conducted in the Charité Berlin was an observational, prospective and longitudinal design described in detail by @Prehn_2020.
-The sample is identical with subsample of this analysis that was replenished with participants taking part in a second follow-up fMRI session 12 months post-surgery/ after baseline.
-Participants with at least one data point were included in the analysis.
-<!-- Descriptive Statistics --> The analysis was done only on morbidly obese individuals with fMRI data (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; mean age `r round(mean(subj_df$Age_BL),2)` years $\pm$ SD `r round(sd(subj_df$Age_BL),2)`, that either underwent surgery (n = `r nrow(filter(subj_df,condition == "IG"))`, `r nrow(filter(subj_df,condition == "IG" & Sex == "f"))` female) or were waiting list control (n = `r nrow(filter(subj_df,condition == "KG"))`, `r nrow(filter(subj_df,condition == "KG" & Sex == "f"))` female).
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tab4descrptives)).
+<!-- STUDY DESIGN --> The ADIPOSITAS-Study conducted in the Charité Berlin is a observational, prospective and longitudinal design described in detail by @Prehn_2020.
+Data collection is still ongoing but only data available in April 2019 is used.
+the procedures followed were in accordance with the Helsinki Declaration.
+This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/mzyxv](https://osf.io/mzyxv).
+
+<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
+These were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
+type-2 diabetes, hypertension, non-alcoholic fatty liver disease).
+<!-- Exclusion criteria --> Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
+<!-- Frage: Ausschluss bei head movement?
+ADI009_fu2 nicht enthalten (siehe Readme preprocessing) -->
+
+<!-- DESCRIPTIVE STATISTICS --> The analysis was done only on morbidly obese individuals with fMRI data (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; mean age = `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range: `r min(subj_df$Age_BL)` to `r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list control (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
+The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing all subjects that have received MRI.
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tabdescr)).
 `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
+Analyses were performed on all participants that provided at least one data point (detailed information on data availability in \@ref(tab:tabsample)).
+```{r baseline characteristics, include=FALSE}
+dfBL <- save_txt_for_swe(group = "all", tp = "BL")
 
-```{r tab4descrptives, echo=FALSE}
+# distributions per group
+# BMI 
+p <- ggplot(data=dfBL, aes(group = condition, x=BMI, fill=condition)) +
+  geom_density(alpha=0.6) + 
+  scale_fill_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")) +
+  theme_bw()
+# log mean FD
+p <- ggplot(data=dfBL, aes(group = condition, x=logmFD, fill=condition)) +
+  geom_density(alpha=0.6) + 
+  scale_fill_manual(
+    values = c("#00305E55", "#C0004555"),
+    labels = c("intervention", "control")) +
+  theme_bw()
+# Age
+p <- ggplot(data=dfBL, aes(group = condition, x=Age_BL, fill=condition)) +
+  geom_density(alpha=0.6) + 
+  scale_fill_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")) +
+  theme_bw()
+tab_age <- describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
+# Sex
+p <- ggplot(data = dfBL, aes(x = condition, fill = Sex)) + 
+  geom_bar(position = "dodge") +
+  scale_fill_manual(labels = c("female", "male"),
+                    values = c("#00305E55", "#C0004555"),
+  )
+tab_sex <- table(dfBL$condition, dfBL$Sex)
+```
+Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group, except for the distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
+`r round(tab_age$mean[tab_age$group1=="IG"],2)`).
+<!-- Details about surgeries --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
+```{r tabdescr, echo=FALSE}
+
 # table for descriptive statistics
@@
 descr <- descr[,c("predictor","tp","group","n","mean","sd")]
+descr <- descr[with(descr, order(predictor, tp)),]
+descr[,c("mean","sd")] <- round(descr[,c("mean","sd")],2)
+descr[c(2:6,8:12),"predictor"] <- ""
+descr[c(2,4,6,8,10,12),"tp"] <- ""
 rownames(descr) <- NULL
 
-descr <- kableExtra::kbl(descr[with(descr, order(predictor, tp)), ],
+tab_descr <-
+  knitr::kable(
+    descr,
+    row.names = FALSE,
     format = "latex",
     booktabs = T,
-                         caption = "BMI and mean FD of participants acquired in each condition and time point",)
-descr
+    caption = "BMI and mean FD of participants acquired in each condition and time point")
+
+tab_descr
+
+```
+
+<!-- Study procedure -->
+Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for STANDARDIZED??
+breakfast.
+MRI scanning was done after further psychological tests (for details see @Prehn_2020).
+
+<!-- Figures zu BMI über die Zeit in den Gruppen. -->
+```{r fig_BMIdescr, fig.asp = 1, fig.width = 5}
+# BMI over time each group
+fig_BMIdescr <-
+  ggplot(final, aes(
+    x = tp,
+    y = BMI,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition) + xlab("time points")
+fig_BMIdescr +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
 ```
@@
 <!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, TR = 1900 ms, TE = 2.52 ms, a = 9, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
-<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes<!-- dimension = [60;72;60] ???
+<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3 s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes<!-- dimension = [60;72;60] ???
 -->.
 The total acquisition time was 5:45 minutes.
+Field map or phase-encode reversed images were not acquired.
+Participants were instructed to close their eyes but to remain awake during scanning.
 
@@
 
-Data preprocessing was conducted with FSL and Freesurfer.
-<!-- minimal preprocessing --> Preprocessing pipeline included removal of first four volumes and a transformation of the blood oxygen level dependent (BOLD) time series into the subject's anatomical space.
+Data preprocessing was conducted with FSL and Freesurfer using a nipype (version 14.0.2) workflow implemented in python 2.7. 
+The preprocessing workflow included anatomical, functional and diffusion-weighted imaging preprocessing and wrapped modules from FSL version 6.0.1<!-- ?!
+-->, Freesurfer version 6.0.0p1, ANTS <!-- version?!
+and so on-->.
+Details have been published on [github]()
+<!-- anatomical preprocessing write something about anatomical preprocessing (Freesurfer etc.) here -->
+
+The T1-weighted image was segmented to remove non-brain tissue and together with the functional images normalized to a default brain template from the Monteal Neurological Institute (MNI) space.
+Functional images were smoothed with an $0$ mm FWHM Gaussian kernel.
+
+<!-- minimal functional preprocessing --> Preprocessing pipeline included removal of first four volumes and a transformation of the blood oxygen level dependent (BOLD) time series into the subject's anatomical space.
 This transform combines parameters from motion correction (FSL's MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL's FUGUE) and coregistration to the subject's anatomical template space (using Freesurfer's bbregister and the transform from individual to longitudinal template space in Freesurfer's longitudinal processing stream).
+
 <!-- Aroma + CC/ + GSR --> Volumes were further preprocessed with ICA-AROMA and compCor (CC) and, optionally, with a subsequent global signal regression (GSR) [@Ciric_2017; @Parkes_2018] to reduce the effect of physiological noise and head motion on the connectivity estimates.
 As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but is known to introduces spatial dependency and spurious correlations, we performed GSR after ICA-AROMA in a separate sensitivity analysis.
-<!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC).
-Subsequently, r-to-Z transform was applied to the correlation (usin nilearn, nipype and python).
+<!-- Regions of interest (definition and transformation into freesurfer long space) -->
+<!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint (more details)) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC).
+Subsequently, r-to-z transform was applied to the correlation to obtain normally distributed functional connectivity (using nilearn, nipype and python).
 Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).
-During analyses, beyond an implicit mask, the MNI resampled brain mask of the MNI ICBM 2009 atlases was used as explicit mask.
 
@@
 As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].
+Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012 and log-transformed and examined within and separate from the main models.
 
 ### Outcomes
-
-<!-- Computation of whole brain FC --> Main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
-<!-- computation of FD --> Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012 and log-transformed.
+<!-- MEASURES AND COVARIATES -->
+<!-- Computation of whole brain FC --> The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
 
@@
 
-```{r calculate VIF, include=FALSE}
-# generate random vector 
-final$randny <- rnorm(nrow(final), mean = 0, sd = 1)
-m_grouptime <- lm("randny ~ group + tp + Sex + Age_BL + logmFD", data = final)
-m_bmi <- lm("randny ~ avgBMIc + cgnBMI + Sex + Age_BL + logmFD", data = final)
-m_fdbmi <- lm("randny ~ avgBMIc + cgnBMI + avgFDc + cgnFD", data = final)
+Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
+During analyses, beyond an implicit mask, the the MNI ICBM 2009a Nonlinear Symmetric brain masks [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled t0 3 x 3 x 3 mm$^3$, then binarized at threshold $0.5$ and used as explicit mask.
 
-# variance inflation factor (VIF) - i.e.loss of precision implied by the observed level of collinearity/multicollinearity: all_vif = 1/tol
-car::vif(m_grouptime)
-car::vif(m_bmi)
-car::vif(m_fdbmi)
+<!-- es waren ja mehr als 2, nämlich 2 (group vs bmi) * 2 (mit/ohne logFD) * 2 (ACC/PC) *2 (CC + GSR) *2 (brain mask + GM mask), versuch das alles hier zu nennen, damit man sich später besser zurechtfindet -->
 
-# tolerance - take each predictor in turn and regress that variable (treating it as the outcome variable) on  theremaining q-1 predictors, then calculate the proportion of variance each predictor shares with other predictors in the full model all_tol = 1 - R²(Xi)
-1/car::vif(m_grouptime)
-1/car::vif(m_bmi)
-1/car::vif(m_fdbmi)
-
-psych::corr.test(final[,c("avgBMIc", "cgnBMI", "avgFDc", "cgnFD")], use = "pairwise")
-```
-
-Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
-
-<!-- Model setup --> Two models were examined.
+<!-- Model setup --> Two theoretically derived models were set up.
 The first model was aimed at testing the interaction of time and group: $FC = group + time + group*time$.
 The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.
-These two models were run for both, the PCC and the NAcc, resulting four different models.
-The marginal model set up with the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+These were examined each for correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $16$ a priori planned analyses.
+Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$.
+As a result a total of $32$ different analyses were performed.
+Computation with the grey matter mask were exploratory and deviate from the preregistration.
+
+<!-- Description marginal model --> The marginal model set up with the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
 It therefore accomodates any repeated measurement covariance structures.
 We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size and therefore accommodating for heteroscedasticity.
-<!-- Design matrix --> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
+
+<!-- Design matrix, vielleicht beide auch als Abbildung kurz zeigen?
+--> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
 The resulting flexible factorial design contained one regressor for each time point per group.
@@
 on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- Nuisance covariates --> Age, sex, and mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+<!-- Nuisance covariates --> Age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
 However, as change in FD and change in BMI share variance [@Beyer_2019].
-Therefore, inclusion of mean FD may be disputed.
+Hence, inclusion of log mean FD may be disputed and done optionally.
 <!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
 Differences in results will be reported.
-For a better understanding of the unique contribution of average and longitudinal change in BMI and FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject BMI + within-subject BMI + between-subject FD + within-subject FD$.
-These post hoc analyses deviate from the preregistration and should be considered to be exploratory.
+For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
+These post-hoc analyses deviate from the preregistration and should be considered to be exploratory.
 
@@
 As collinearity may cause problems we calculated an approximation of the variance inflation factor (VIF) which was shown to be minimal for all models except the third exploratory.
-Here, there was minor variance inflation for longitudinal change in both BMI and FD was moderately positively correlated (MLM HERE $r$ = `r cor(final$cgnFD,final$cgnBMI)`, $VIF$ = `r round(car::vif(m_fdbmi),2)`).
+Here, there was minor variance inflation for longitudinal change in both BMI and FD was moderately positively correlated.
 <!-- Problem(?): correlation not does not consider multilevel - compute linear mixed model -->
-To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
-In contrast to the parametric estimation, bootstrap is adapted to use family-wise error correction, and unlike permutation based framework appropriate for longitudinal data.
+<!-- Ensure robustness of inference --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
+In contrast to the parametric estimation, bootstrap is adapted to use family-wise error correction, and unlike permutation based framework appropriate <!-- ??? clarify why--> for longitudinal data.
 
@@
 
-The same analyses as described a priori were carried out with a grey matter mask replacing the whole brain mask.
-Focusing only on grey matter instead if the whole brain reduced the number of tests (voxels of interest) from $52378$ to $51796$.
-Computation with the grey matter mask were executed post hoc and deviate from the preregistration.
+## Manipulation check 
+<!-- check network presence -->
 
-## Manipulation check
+<!-- singletp (t-tests), nur diesen Check für BL reporten (gib dann oben noch an, dass du dafür SPM verwendet hast) --> We ran separate one-sample t-test for network activation at baseline controlling for log mean FD, age, and sex for each time point to check whether the assumed reward and default mode network was at all identifiable.
 
-<!-- t-tests to check for network presence --> We ran separate one-sample t-test for each time point to check whether the assumed reward and default mode network was at all identifiable.
-<!-- individual statistical maps -->
-
-## Interaction effect of intervention and Time
+## Interaction effect of intervention and time
 
@@
 In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm]) yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
-However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{FWEcorr} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!
+However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{FWEcorr} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
 -->.
@@
 Omitting mean FD as nuisance covariate did not change these results. 
-<!-- BMIonlyIG --> Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was a significant effect of change in BMI on co-activation between the NAcc and a voxel ($-18,-42,-9$ [mm]) belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in connectivity maps without GSR).
+<!-- BMIonlyIG --> Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ --> with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps).
 However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)<!-- but FWE-corr sig?!
@@
 
-<!-- GM mask: average and individual BMI -->
-Results obtained from analyses with the grey matter masks were similar to those reported for the grey matter masks; test statistics for the significant longitudinal effect differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no significant co-activation was found.
+<!-- GM mask: average and individual BMI, damit die Reihenfolge so ist wie oben. -->
+Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
+In regard to the exploratory expeditions on subsamples, consistent with previous results, no significant co-activation was found when using the grey matter mask.
 
@@
 <!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
-Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only<!-- or for  the first two time points only -->.
-
-<!-- GM mask: FD -->
+Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
+<!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
 
 <!-- Neurovault --> 
-Contrasts images are published on NeuroVault: ......
-
-## Target regions of the default mode and reward network
-
-exploratory expeditions
-deviated from the preregistration 
-were executed post hoc
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 
@@
 # Discussion
+
+<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
+<!-- SIMILARITY OF RESULTS -->
+<!-- INTERPRETATION -->
+<!-- GENERALIZABILITY -->
+<!-- IMPLICATIONS -->
+
 <!-- Description of analyses carried out --> Marginal models were used to adequately capture longitudinal trends in resting state connectivity of in an intervention group and the waiting control group.
 
-With the model on group-time-interaction, we investigated average effects of the intervention as well as average changes over time while controlling for individual variance in resting-state connectivity over time and at baseline.
+With the model on group-time-interaction, we primarily investiage whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
+Beyond that, also main effects of the intervention and of the time were examined.
 <!-- correct interpretation of marginal models?
---> The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI
+-->
+The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
 
-To summarize, there were few significant co-activation emerged from the marginal models.
-Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
-Ventures to verify effects by recalculating with non-parametric Wild bootstrap revealed that effects were not robust.
-This suggests that significant results are likely to be false positive.
-
-<!-- Multicollinearity --> Prior to the analysis we assessed multicollinearity of our predictors and correlations between our individual change in BMI and FD.
+<!-- Problem head motion --> Based on concerns expressed in the literature, we tried to closely inspect and rigorously control for the influence of head motion.
+The presence of multicollinearity of our predictors has been established.
 As described in @Monti_2011 in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
 Confidence in $\beta$ estimates will suffers and estimated can be erratic, depending on which regressors included in the model.
-However, correlation between both longitudinal changes were expected as they share ...............
+
+To summarize, there were few significant co-activation that emerged from the marginal models.
+Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
+Ventures to verify effects by recalculating with non-parametric Wild bootstrap revealed that effects were not robust across estimation procedures.
+This suggests that significant results are likely to be false positive.
 
@@
 
-<!-- Incomplete data --> how many had three/ two/ one time point?
+# Code availability
+
+All code is available in the github repository [https://github.com/hsx1/adi2_rsfmri](https://github.com/hsx1/adi2_rsfmri).
+The code includes script for data preparation and analysis of different models.
+
 \newpage
+# References
+
+```{r Exploratory plots, eval=FALSE}
+# BMI over time per group
+tspag <-
+  ggplot(final, aes(x = tp, y = BMI)) +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  geom_line(aes(group = subj.ID, color = condition)) +
+  geom_point(aes(color = condition))
+tspag +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "bottom",
+    legend.background = element_rect(colour = "grey")
+  )
+
+# BMI over time per subject
+# group x time interaction for BMI
+tspag <- ggplot(final, aes(
+  x = tp,
+  y = BMI,
+  color = condition,
+  group = condition
+)) +
+  geom_point() +
+
+  stat_summary(fun = mean, geom = "point") +
+  stat_summary(fun = mean, geom = "line")
+tspag  +
+  scale_color_manual(values = c("#00305E", "#C00045")) +
+  theme_bw() +
+  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.position = "top"
+)
+```
+```{r expample plots, eval=FALSE}
+# example plots for interaction effect (arbitrary variables for illustration purposes)
+ggplot(final, aes(x=tp, y=BMI, group=condition, color=condition)) +
+  #geom_smooth(aes(group=condition)) +
+  stat_summary(fun = mean, geom = "point", size=3, shape=15) +
+  stat_summary(fun = mean, geom = "line", size=1) +
+  #scale_color_brewer(values = "Set1") +
+  scale_x_discrete(labels=c("bl"= "0", "fu"= "6", "fu2" = "12")) +
+  scale_color_manual(values=c("#00305E","#C00045"),labels = c("intervention", "control")) +
+  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10), axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12), legend.text = element_text(size = 10),
+    legend.position = "bottom", legend.background = element_rect(colour = "grey"))
+
+# example plot for change effect (purely descriptive)
+getPalette = colorRampPalette(brewer.pal(9, "Set1"))
+cgnplot <- ggplot(final, aes(x=a,y=logmFD, group=subj.ID)) + 
+  geom_line(aes(x=BMI, color=subj.ID))+ 
+  geom_point(aes(x=BMI, color=subj.ID, shape=tp), size=3, alpha = 0.5)+ 
+  facet_grid(. ~ condition) 
+cgnplot + 
+  labs(shape="Time after intervention [months]",labels=c("0","6","12")) +
+  guides(color = FALSE, alpha=FALSE, size=FALSE) +
+  scale_color_manual(values = getPalette(length(unique(final$subj.ID)))) +
+  ylab("log(mean FD) (mm)") + xlab("BMI [kg/m²]") + xlim(30,55) +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10), axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12), legend.text = element_text(size = 10),
+    legend.position = "bottom", legend.background = element_rect(colour = "grey"))
+
+# example plot for change effect (purely descriptive)
+tspag=ggplot(data=final, aes(y=cgnFD, x=cgnBMI, color=condition))+
+  geom_point(size=2)+
+  #geom_line(aes(color=subj.ID))+ 
+  #geom_smooth(method = "lm", colour = "black", linetype = 1, fill = "mistyrose3" ) +
+  xlab("within-subject BMI change [kg/m²]")+
+  ylab("within-subject log mean FD change [mm]")+ 
+  guides(alpha=FALSE, size=FALSE,
+         color = guide_legend(override.aes = list(size=4))) +
+  facet_grid(. ~ condition) 
+tspag + scale_color_manual(values=c("#00305E","#C00045")) +
+  theme_bw() + theme(axis.text=element_text(colour = "black",size=10),
+              axis.title=element_text(size=14),
+              axis.text.x = element_text(size=12),
+              axis.text.y = element_text(size=12),
+              strip.text = element_text(size=14),
+              legend.position="top",
+              legend.title = element_text(size = 14),
+              legend.text = element_text(size = 14))
+```

### 2020-11-16 draft_rsfmri.Rmd
function to create scan list added

@@
 Data collection is still ongoing but only data available in April 2019 is used.
-the procedures followed were in accordance with the Helsinki Declaration.
+The procedures followed were in accordance with the Helsinki Declaration.
 This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/mzyxv](https://osf.io/mzyxv).
-
 <!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
@@
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease).
-<!-- Exclusion criteria --> Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
+Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
 <!-- Frage: Ausschluss bei head movement?
 ADI009_fu2 nicht enthalten (siehe Readme preprocessing) -->
-
-<!-- DESCRIPTIVE STATISTICS --> The analysis was done only on morbidly obese individuals with fMRI data (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; mean age = `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range: `r min(subj_df$Age_BL)` to `r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list control (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
-The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing all subjects that have received MRI.
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tabdescr)).
 `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
-Analyses were performed on all participants that provided at least one data point (detailed information on data availability in \@ref(tab:tabsample)).
+The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing only subjects that have received MRI.
+Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tabsample)).
+<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; ages `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (`r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
+The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
 ```{r baseline characteristics, include=FALSE}
@@
 ```
-Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group, except for the distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
@@
 ```
-
-<!-- Study procedure -->
+<!-- STUDY PROCEDURE -->
 Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
@@
 
-<!-- Figures zu BMI über die Zeit in den Gruppen. -->
 ```{r fig_BMIdescr, fig.asp = 1, fig.width = 5}
@@
 
-<!-- exclusion criteria --> Participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
+<!-- EXCLUSION CRITERIA --> Participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
 caused Freesurfer to fail or perform very poorly, i.e.
 rated "Failed" according to @Klapwijk_2019) were excluded because a subject's motion is highly similar across scans and we thus suspect the fMRI data to also suffer from motion artifacts in these subjects.
-As the expected average head motion is high [@Hodgson_2016] we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample.
-<!-- motion correction --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
+As the expected average head motion is high [@Hodgson_2016], we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample.
+<!-- MOTION CORRECTION --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
 As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].
-Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012 and log-transformed and examined within and separate from the main models.
+Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012, log-transformed and examined within and separate from the main models.
 
 ### Outcomes
-<!-- MEASURES AND COVARIATES -->
-<!-- Computation of whole brain FC --> The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
+<!-- MEASURES AND COVARIATES --> The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
 
@@
 During analyses, beyond an implicit mask, the the MNI ICBM 2009a Nonlinear Symmetric brain masks [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled t0 3 x 3 x 3 mm$^3$, then binarized at threshold $0.5$ and used as explicit mask.
-
-<!-- es waren ja mehr als 2, nämlich 2 (group vs bmi) * 2 (mit/ohne logFD) * 2 (ACC/PC) *2 (CC + GSR) *2 (brain mask + GM mask), versuch das alles hier zu nennen, damit man sich später besser zurechtfindet -->
 
@@
 
-<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
-<!-- SIMILARITY OF RESULTS -->
-<!-- INTERPRETATION -->
-<!-- GENERALIZABILITY -->
-<!-- IMPLICATIONS -->
-
-<!-- Description of analyses carried out --> Marginal models were used to adequately capture longitudinal trends in resting state connectivity of in an intervention group and the waiting control group.
-
-With the model on group-time-interaction, we primarily investiage whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
+<!-- summary -->
+With the model on group-time-interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
 Beyond that, also main effects of the intervention and of the time were examined.
@@
 
-<!-- Problem head motion --> Based on concerns expressed in the literature, we tried to closely inspect and rigorously control for the influence of head motion.
-The presence of multicollinearity of our predictors has been established.
+<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
+
+<!-- SIMILARITY OF RESULTS -->
+<!-- Discuss mixed effect model -->
+
+<!-- INTERPRETATION -->
+
+<!-- GENERALIZABILITY -->
+
+<!-- IMPLICATIONS -->
+<!-- Relevance of rsfMRI as biomarker -->
+<!-- Consideration of head motion as covariate -->
+
+Marginal models were used to adequately capture longitudinal trends in resting state connectivity of in an intervention group and the waiting control group.
+
+<!-- Problem head motion --> Based on concerns expressed in the literature [@], we tried to closely inspect and rigorously control for the influence of head motion.
+The presence of multicollinearity of our predictors has been established before.
 As described in @Monti_2011 in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
-Confidence in $\beta$ estimates will suffers and estimated can be erratic, depending on which regressors included in the model.
+Confidence in $\beta$ estimates will suffers and estimates can behave erratic depending on which regressors included in the model.
 
@@
 It decreased with after a meal while dietary patients showed an increase [@Lepping_2015].
+
+<!-- For all patients included, conservative obesity treatment had failed, e.g.
+behavioral diet intervention.
+As shown before, patients who were successful in behavioral diet increased connectivity in the default mode network.
+This might be indicative for self-referent information processing.
+-->
 

### 2020-11-16 AllTPEval.m

@@ -15,7 +15,7 @@ for i = 1:length(param.MODEL)
         end
     end
 end
-spm('Quit')
+%spm('Quit')
 
 %% ========================================================================
 function display_message(COVARIATES)
@@ -78,9 +78,13 @@ for crun = 1:nrun
             matlabbatch = DisplayResults(location_SwE_mat);
             spm_jobman('run', matlabbatch);
             pause(param.VIEWSEC)
-            save("xSwE.mat")
         elseif not(exist_already) || param.OVERWRITE
             fprintf('Estimate model...\n')
+            % delete existing files in folder if existent
+            if exist_already 
+                rmdir(out_folder,'s'); % delete former dir
+                mkdir(out_folder); % create new empty one
+            end
             spm('defaults', 'FMRI'); 
             matlabbatch = RunModel(matlabbatch, location_SwE_mat);
             spm_jobman('run', matlabbatch);

### 2020-11-16 RunModelBMI.m

@@ -12,6 +12,7 @@ for i = 1:length(param.MODEL)
         end
     end
 end
+fprintf("Enter > spm('Quit') < to exit.")
 %spm('Quit')
 
 %% ========================================================================
@@ -76,7 +77,6 @@ for crun = 1:nrun
             matlabbatch = DisplayResults(location_SwE_mat);
             spm_jobman('run', matlabbatch);
             pause(param.VIEWSEC)
-            save("xSwE.mat")
         elseif not(exist_already) || param.OVERWRITE
             fprintf('Estimate model...\n')
             % delete existing files in folder if existent
@@ -108,8 +108,13 @@ for crun = 1:nrun
                 spm_jobman('run', matlabbatch);
                 pause(param.VIEWSEC)
                 save("xSwE.mat")
-            elseif not(exist_already)
+            elseif not(exist_already)|| param.OVERWRITE
                 fprintf('Estimate model...\n')
+                % delete existing files in folder if existent
+                if exist_already 
+                    rmdir(out_folder,'s'); % delete former dir
+                    mkdir(out_folder); % create new empty one
+                end
                 spm('defaults', 'FMRI');
                 matlabbatch = RunModel(matlabbatch, location_SwE_mat);
                 spm_jobman('run', matlabbatch);
@@ -140,25 +145,16 @@ cd(param.INFO_DIR)
 % save directory of (new) folder
 matlabbatch{1}.spm.tools.swe.smodel.dir = {out_folder};
 
+% cifti + gifti additional information (SwE 2.2.1)
+smodel.ciftiAdditionalInfo.ciftiGeomFile = struct('brainStructureLabel', {}, 'geomFile', {}, 'areaFile', {});
+smodel.ciftiAdditionalInfo.volRoiConstraint = 0;
+smodel.giftiAdditionalInfo.areaFileForGiftiInputs = {};
+
 %% Load Scans -------------------------------------------------------------
 % constructing cell array with the scans (FC maps per subject and time point)
 scans_dir = readcell(fullfile(param.INFO_DIR,'scans.txt'), 'Delimiter',' ','Whitespace',"'");
 roi_prep = param.ROI_PREP{crun};
-if contains(roi_prep, 'gsr')
[…]

### 2020-11-16 RunModelFD.m

@@ -12,7 +12,8 @@ for i = 1:length(param.MODEL)
         end
     end
 end
-spm('Quit')
+fprintf("Enter > spm('Quit') < to exit.")
+%spm('Quit')
 
 %% ========================================================================
 function display_message(COVARIATES)
@@ -77,7 +78,6 @@ for crun = 1:nrun
             matlabbatch = DisplayResults(location_SwE_mat);
             spm_jobman('run', matlabbatch);
             pause(param.VIEWSEC)
-            save("xSwE.mat")
         elseif not(exist_already) || param.OVERWRITE
             fprintf('Estimate model...\n')
             % delete existing files in folder if existent
@@ -140,25 +140,16 @@ cd(param.INFO_DIR)
 % save directory of (new) folder
 matlabbatch{1}.spm.tools.swe.smodel.dir = {out_folder};
 
+% cifti + gifti additional information (SwE 2.2.1)
+smodel.ciftiAdditionalInfo.ciftiGeomFile = struct('brainStructureLabel', {}, 'geomFile', {}, 'areaFile', {});
+smodel.ciftiAdditionalInfo.volRoiConstraint = 0;
+smodel.giftiAdditionalInfo.areaFileForGiftiInputs = {};
+
 %% Load Scans -------------------------------------------------------------
 % constructing cell array with the scans (FC maps per subject and time point)
 scans_dir = readcell(fullfile(param.INFO_DIR,'scans.txt'), 'Delimiter',' ','Whitespace',"'");
 roi_prep = param.ROI_PREP{crun};
-if contains(roi_prep, 'gsr')
-    % special case if preprocessing is 'gsr'
-    if endsWith(roi_prep,'_z')
-        FC_files = strcat(lower(roi_prep(1:end-5)),'cc_gsr_seed_correlation_z_trans.nii');
-    else
-        FC_files = strcat(lower(roi_prep(1:end-3)),'cc_gsr_seed_correlation_trans.nii');
-    end
-else
-    if endsWith(param.ROI_PREP{crun},'_z')
-        FC_files = strcat(lower(roi_prep(1:end-2)),'_seed_correlation_z_trans.nii');
-    else
-        FC_files = strcat(lower(roi_prep),'_seed_correlation_trans.nii');
-    end
-end
-scans_of_roi = fullfile(scans_dir, roi_prep, FC_files);
+scans_of_roi = create_scans_list(scans_dir, roi_prep);
 
 % load scans to matlabbatch
[…]

### 2020-11-16 RunModelGroupTime.m

@@ -12,7 +12,8 @@ for i = 1:length(param.MODEL)
         end
     end
 end
-spm('Quit')
+fprintf("Enter > spm('Quit') < to exit.")
+%spm('Quit')
 
 %% ========================================================================
 function display_message(param)
@@ -107,8 +108,7 @@ for crun = 1:nrun
             spm('defaults', 'FMRI'); 
             matlabbatch = DisplayResults(location_SwE_mat);
             spm_jobman('run', matlabbatch);
-            pause(param.VIEWSEC)
-            save("xSwE.mat")
+            pause(param.VIEWSEC) 
         elseif not(exist_already) || param.OVERWRITE
             fprintf('Estimate model...\n')
             % delete existing files in folder if existent
@@ -140,8 +140,13 @@ for crun = 1:nrun
                 spm_jobman('run', matlabbatch);
                 pause(param.VIEWSEC)
                 save("xSwE.mat")
-            elseif not(exist_already)
+            elseif not(exist_already) || param.OVERWRITE
                 fprintf('Estimate model...\n')
+                % delete existing files in folder if existent
+                if exist_already
+                    rmdir(out_folder,'s'); % delete former dir
+                    mkdir(out_folder); % create new empty one
+                end
                 spm('defaults', 'FMRI');
                 matlabbatch = RunModel(matlabbatch, location_SwE_mat);
                 spm_jobman('run', matlabbatch);
@@ -172,25 +177,16 @@ cd(param.INFO_DIR)
 % save directory of (new) folder
 matlabbatch{1}.spm.tools.swe.smodel.dir = {out_folder};
 
+% cifti + gifti additional information (SwE 2.2.1)
+smodel.ciftiAdditionalInfo.ciftiGeomFile = struct('brainStructureLabel', {}, 'geomFile', {}, 'areaFile', {});
+smodel.ciftiAdditionalInfo.volRoiConstraint = 0;
+smodel.giftiAdditionalInfo.areaFileForGiftiInputs = {};
+
 %% Load Scans -------------------------------------------------------------
 % constructing cell array with the scans (FC maps per subject and time point)
 scans_dir = readcell(fullfile(param.INFO_DIR,'scans.txt'), 'Delimiter',' ','Whitespace',"'");
 roi_prep = param.ROI_PREP{crun};
[…]

### 2020-11-16 SingleTPEval.m

@@ -103,8 +103,7 @@ if param.ONLY_DISPLAY == false
     fmri_est.write_residuals = 0;
     fmri_est.method.Classical = 1;
     matlabbatch{2}.spm.stats.fmri_est = fmri_est;
-
-elseif param.ONLY_DISPLAY == true
+    
     %% Contrast manager
     con.spmmat = {fullfile(out_folder,'SPM.mat')};
     con.consess{1}.tcon.name = 'mean';
@@ -112,6 +111,8 @@ elseif param.ONLY_DISPLAY == true
     con.consess{1}.tcon.sessrep = 'none';
     con.delete = 0;
     matlabbatch{3}.spm.stats.con = con;
+
+elseif param.ONLY_DISPLAY == true
 
     %% Results
     results.spmmat = {fullfile(out_folder,'SPM.mat')};
@@ -124,7 +125,8 @@ elseif param.ONLY_DISPLAY == true
     results.conspec.mask.none = 1;
     results.units = 1;
     results.export{1}.ps = true;
-    matlabbatch{4}.spm.stats.results = results;
+    matlabbatch{1}.spm.stats.results = results; % matlabbatch{4} if done in one step with analysis
+
 end
 
 end

### 2020-11-16 get_subject_ID_group_tp_forSwE.R

@@ -1,7 +1,7 @@
 library(dplyr)
 library(tidyr)
 
-save_txt_for_swe <- function(group = "BL", tp = "all"){
+save_txt_for_swe <- function(group = "all", tp = "all"){
   # group = IG/KG/both
   # tp = BL/FU/FU2/BLFU/FUFU2/all
   
@@ -85,10 +85,10 @@ save_txt_for_swe <- function(group = "BL", tp = "all"){
 # Modify Design matrix for subsamples -------------------------------------
   
   # Model specification incl. design matrix for different options
+  output_dir <- parentdir
   # selection of groups
   if (group == "both") {
     # total: both groups
-    output_dir <- parentdir
   }else if (group == "IG") {
     output_dir <- file.path(parentdir, "IG_only")
     if (!dir.exists(output_dir)) {dir.create(output_dir)}

### 2020-11-16 run_or_display.m

@@ -70,8 +70,10 @@
 % Especially for results of non-parametric estimation, you need to specify
 % how long you want to look at the results in VIEWSEC. Alternatively you
 % can also set breakpoints in the funtion scripts.
+
 %% ========================================================================
 % import absolute path for project
+% SET PATH TO code_and_manuscript FOLDER AS CURRENT DIRECTORY !!
 ABS_DIR = readcell("abs_path.csv");
 ABS_DIR = ABS_DIR{1};
 % create a struct, with all important parameters
@@ -84,50 +86,42 @@ param.MASK_B = 'MNI_resampled_brain_mask.nii,1';
 
 % define ROI
 roi_prep = readcell(fullfile(param.INFO_DIR,'ROIs.txt'), 'Delimiter',' ','Whitespace',"'");
-param.ROI_PREP = {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z'}; % {roi_prep{[4, 6, 12, 14]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z'}
+param.ROI_PREP = {roi_prep{[12,14]}}; % {roi_prep{[4, 6, 12, 14]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z'}
 
 % Model definition
 % All three models have unique options for covariate definition, the
 % association to a model is indicated by the tens digit (GroupTime_: 1_; 
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
 
-param.MODEL = {'grouptime','grouptime2tp'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [11, 12];        % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
+param.MODEL = {'singletp'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
+param.COVARIATES = [42];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
 
 % define masking and type of inference
-param.MASK = 'gm';                  % 'brain' or 'gm'
-param.WILD_BOOT = false;            % false
[…]

### 2020-11-24 draft_rsfmri.Rmd
added text, corrected tables

@@
 ---
-title: "Manuscript_rsfMRI"
+title: "Effects of bariatric surger and weight loss on resting-state functional connectivity"
 author:
-- xx [1][2][3]
-- xx [2]
+- xxx [1]
+- xxx [1][2]
+- xxx [1][2][3]
+- xxx [1][2]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
@@
   Day Clinic for Cognitive Neurology, University Clinic Leipzig
+
 output:
@@
 
-```{r "Package loading", include=FALSE}
+```{r "environment setup", include=FALSE}
+
+# load packages
 library(plyr)
@@
 library(tidyr)
-library(knitr)
 library(kableExtra)
 library(ggplot2)
+library(patchwork) # API for sequentially building up a plot (similar to gridExtra and cowplot)
+#library(cowplot)
 library(RColorBrewer)
-# library(haven)
 library(psych)
-library(patchwork)
-library(lsr) 
-library(wesanderson)
-#library(cowplot)
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4)
+#library(lsr) #
+
+#library(wesanderson) # colour palette
+#library(haven)
+
+library(knitr)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
 options(knitr.table.format = "latex") 
@@
 opts_knit$set(root.dir = getwd())
+
 ```
@@
 <!-- H1 Reward network --> In this study, we aim to investigate possible dynamics of within-reward network connectivity before and after bariatric surgery.
-Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity.
+Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
 <!-- H2 Default mode network --> In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
@@
 ...
-possibly leading to 
-Adipose tissue, gut hormones and gut microbiota may play a mediating role effects for the interaction between body weight and (brian)
 
@@
 subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each subject
+subj_df <- final[subj_idx,] # df with unique subjects
 
@@
 tp3_tab <- table(subset(final, subj.ID %in% tp3)$tp, subset(final, subj.ID %in% tp3)$condition)[1,]
-tab_sample <- data.frame(rbind(tp1_tab, tp2_tab, tp3_tab))
-rownames(tab_sample) <- c("only 0","only 6","only 12","0 and 6","6 and 12","complete data")
+total_datapoints <- c(sum(final$group==1),sum(final$group==2))
+total_subjects <- plyr::count(final[subj_idx,"group"])$freq
+tab_sample <- data.frame(rbind(tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints))
+rm(tp1,tp2,tp3,tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints)
+rownames(tab_sample) <- c("count: only 0","count: only 6","count: only 12","count: 0 and 6","count: 6 and 12","count: complete data","total number of subjects","total data points")
+colnames(tab_sample) <- c("BARS","NBARS")
 
 # distribution of condition and sex
-subj_df <- final[subj_idx,]
 subj_df$Sex <- plyr::mapvalues(subj_df$Sex, from = c("-1", "1"), to = c("f", "m"))
@@
 ```
-## Study design
 
@@
 Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tabsample)).
-<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; ages `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (`r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
+<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; ages `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
@@
 `r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-<!-- Details about surgeries --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
+<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
 ```{r tabdescr, echo=FALSE}
@@
 
-```{r fig_BMIdescr, fig.asp = 1, fig.width = 5}
+```{r fig_BMIdescr, fig.align='center', out.width="90%"}
+# eval = TRUE, fig.asp = 0.7, fig.width = 6
+condition.labs <- c("BARS","NBARS") 
+names(condition.labs) <- c("IG","KG")
 # BMI over time each group
@@
   ) +
-  facet_grid(. ~ condition) + xlab("time points")
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
 fig_BMIdescr +
@@
 
-## fMRI data acquisition
+## fMRI acquisition
 
@@
 
-## Data preproceessing
+## Pre-proceessing
 
@@
 ### Outcomes
+
 <!-- MEASURES AND COVARIATES --> The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
@@
 Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
-During analyses, beyond an implicit mask, the the MNI ICBM 2009a Nonlinear Symmetric brain masks [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled t0 3 x 3 x 3 mm$^3$, then binarized at threshold $0.5$ and used as explicit mask.
+During analyses an implicit mask and an explicit mask were used.
+Therefore, the MNI ICBM 2009a Nonlinear Symmetric brain mask [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
 
-<!-- Model setup --> Two theoretically derived models were set up.
+<!-- MODEL SETIP --> Two theoretically derived models were set up.
 The first model was aimed at testing the interaction of time and group: $FC = group + time + group*time$.
@@
 
-<!-- Description marginal model --> The marginal model set up with the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
-It therefore accomodates any repeated measurement covariance structures.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size and therefore accommodating for heteroscedasticity.
+<!-- Description marginal model --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+It therefore accommodates any repeated measurement covariance structures.
+We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
 
@@
 --> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
-The resulting flexible factorial design contained one regressor for each time point per group.
+The resulting flexible factorial design <!-- check: difference from full factorial or other?
+--> contained one regressor for each time point per group.
 For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
@@
 on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- Nuisance covariates --> Age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+<!-- Nuisance covariates --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+Age and sex were between-subject covariates while FD is time-varying.
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-However, as change in FD and change in BMI share variance [@Beyer_2019].
-Hence, inclusion of log mean FD may be disputed and done optionally.
+As change in FD and change in BMI share variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
 <!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
@@
 <!-- Inference --> All parametric inferences on activation were drawn based on False Discovery Rate (FDR) adjusted p-values $\leq 0.05$.
-As collinearity may cause problems we calculated an approximation of the variance inflation factor (VIF) which was shown to be minimal for all models except the third exploratory.
-Here, there was minor variance inflation for longitudinal change in both BMI and FD was moderately positively correlated.
-<!-- Problem(?): correlation not does not consider multilevel - compute linear mixed model -->
-<!-- Ensure robustness of inference --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
-In contrast to the parametric estimation, bootstrap is adapted to use family-wise error correction, and unlike permutation based framework appropriate <!-- ??? clarify why--> for longitudinal data.
+To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
+In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
+random field theory), and is unlike permutation based framework appropriate <!-- ???
+clarify why--> for longitudinal data.
 
-\newpage
 # Results
@@
 
-\newpage
 # Discussion
 
-<!-- summary -->
-With the model on group-time-interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
+<!-- SUMMARY
+-->
+<!-- general --> Building on the evidence for post-surgery changes in functional activity, this study explored the longitudinal relationship between surgery-induced weight loss and functional integrity in target networks indicated by resting-state functional connectivity.
+Over the course of one year we collected neuroimaging and anthropometric data from `r cs["Sum","IG"]` morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of  `r cs["Sum","KG"]` patients on the waiting list.
+<!-- model specific --> With the model on group-time-interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
 Beyond that, also main effects of the intervention and of the time were examined.
@@
 
-<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
+<!-- SUPPORT OF ORIGINAL HYPOTHESIS 
+-->
 
-<!-- SIMILARITY OF RESULTS -->
-<!-- Discuss mixed effect model -->
+<!-- 
+SIMILARITY OF RESULTS -->
+Our results are not consistent with the present literature.
+Although there are few studies investigating the impact of bariatric surgery on functional connectivity at resting state, the present articles do significant changes in network activation [@;@;@].
+<!-- publication bias -->
 
-<!-- INTERPRETATION -->
+<!-- 
+INTERPRETATION -->
+<!-- absence of evidence --> However, hypotheses about the plasticity functional networks following bariatric surgery could not be validated.
+Due to the shortcomings of the inferential paradigm used, our null findings may either put conclusions drawn from previous research into question or may neither provide evidence for the presence nor the absence of the effect.
+<!-- Control for satiety?
+--> A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
+After the intervention, the functional connectivity was higher in bariatric patients.
+It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
 
-<!-- GENERALIZABILITY -->
+<!-- 
+GENERALIZABILITY -->
 
-<!-- IMPLICATIONS -->
+<!-- advantage:  -->
+
+<!-- no RCT --> Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
+<!-- different conclusion from other studies --> To isolate the effect of the surgical intervention to a maximum possible extent, the intervention group has to be comparable to a control group in regard  to predisposition.
+Studies comparing the sample of bariatric patients to lean individuals or behavioral dieters may have investigated risk or contributing factors to the condition or other dispositional difference between groups, e.g.
+in cognitive control, openness or responsiveness to behavioral interventions [@;@].
+However, they provide insufficient evidence to directly evaluate impact of the surgical procedure.
+<!-- prospective intervention control study --> Prospective non-randomization trial study design are secondary to randomized control trials in regard to internal control.
+By establishing temporality, longitudinal measurements support the validity of conclusions about the effects of the surgery.
+Contrasting pre- and post-surgery measures change while controlling for individual variance.
+The waiting list control group equivalent in BMI, treatment history and  treatment recommendation allows controlling for concurrent factors [@Thiese_2014].
+Incorporating these consideration, our study aimed at a higher internal validity for conclusions about the effects of bariatric surgery on functional connectivity.
+
+<!-- advantage: representative sample --> Size and composition of our sample did not allow separate analyses for the sexes.
+The disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
+
+<!-- mention larger sample size -->
+
+<!-- mixed effect model -->
+
+We re-ran analyses with a flexible factorial model as implemented in SPM to estimate the group-time-interaction while controlling for variance on subject-level as done exemplarily by @McFarquhar_2019.
+Thereby,
+<!-- demonstrated / we ruled out gross missspecification in retrieving estimates from the Sandwich estimator. -->
+
+@McFarquhar_2019 demonstrated that specifying complex multi-level models with standard neuroimaging software packages such as SPM and FSL can be quite difficult if done appropriately. 
+From the necessity to include subject effect, the necessity of using a overparamaterized design arises.
+
+Not only do  but modeling between-subjects and within-subject nuisance covariates demands caution and becomes even more complex when using multiple within-subject covariates.
+....
+In both models, the variation between subjects was modeled and therefore ...
+
+Opposed to flexible factorial models for analysis of variance, where each individual is modelled separately, models are parsimonious and preserver more degrees of freedom.
+
+<!-- 
+IMPLICATIONS -->
+
 <!-- Relevance of rsfMRI as biomarker -->
 <!-- Consideration of head motion as covariate -->
-
-Marginal models were used to adequately capture longitudinal trends in resting state connectivity of in an intervention group and the waiting control group.
 
@@
 This suggests that significant results are likely to be false positive.
-
-<!-- Control for satiety?
---> A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
-After the intervention, the functional connectivity was higher in bariatric patients.
-It decreased with after a meal while dietary patients showed an increase [@Lepping_2015].
 
@@
 ```
+
 ```{r expample plots, eval=FALSE}

### 2020-11-27 draft_rsfmri.Rmd
updated introduction

@@
 ---
+```{r "To do", include=FALSE}
+
+# [ ] Introduction
+# [ ] Discussion
+# [ ] hervorheben, was von Präregistrierung abweicht
+# [ ] brain images mit nilearn plotting tools einbinden
+# [x] Figure zu weight loss
+# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
+
+```
 
@@
 
-<!-- BARS & weight loss --> Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018].
-On the other hand, dietary interventions such as caloric restriction and related metabolic improvements and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018].
+Over the past decades, the global burden of obesity has increasingly been confronted as a matter of urgency by research.
 
-<!-- BARS --> Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions and co-morbidities, such as diabetes. 
-Yet, little is known about the potential beneficial effects of bariatric surgery on brain structure and function in this population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018].
+On the other hand, dietary interventions such as caloric restriction and related metabolic improvements and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018; @Veronese_2017].
+Behavioral intervention require ... and often fail to yield the desired success regarding weight loss, especially in in morbid obesity [@]. There are alternative interventions at hand.
+<!-- bars -->
+Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions and co-morbidities, such as diabetes.
+Roux-en-Y gastric bypass (RYGB) is the most common procedure, as it not only restricts the food intake by reducing stomach volume, but also affects the absorption of macronutrients and the hormonal profile [@].
+Other common procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption.
+<!-- Potential mechanisms? -->
+Precise mechanisms are yet to be elucidated, but as a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed to better infer potential mechanisms and support interventions proven to be effective.
+<!-- / immitate interventions that have been proven effective with less invasive alternatives -->
 
-<!-- Reward network --> In lean participants, food valuation processes in decision-making have been linked to brain activity and activity within the reward network, including ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+<!-- Reward network -->
+Little is known about the potential beneficial effects of bariatric surgery on brain function in this population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+<!-- mechanism reward and appetite, overating -->
+Besides malabsorption and restricted food ingestion, patients report post-surgical changes in food preferences and eating behavior.
+Those psychophysiological and behavioral effects may be related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
+Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which a sustainable weight loss [@Mulla_2017, @Nudel_2019].
+
+Several studies have demonstrated differences in resting-state activity in reward regions depending on weight status.
+
+Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
+These differences seemed to normalize, as trends reversed when looking at the resting state of obese individuals after compared to before the surgery [@Duan_2020].
+
+<!-- saliency and reward -->
+In lean participants, food valuation processes in decision-making have been linked to brain activity and activity within the reward network, including ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+
+Studies concerning such processes following a surgical intervention are inconclusive.
+
+<!-- important differences in the response to visual food-related cues between thin and reduced-obese individuals that place the latter group at an increased risk of  -->
+
+<!-- Though there exists strong evidence for important differences in the ..
+between lean or normal-weight individuals and obese individuals, few studies have investigated the ...
+and even fewer compared reduced-obese individuals reduced-obese individuals -->
+<!-- [@Garcia_Garcia_2012] -->
+
 <!-- Default mode network -->
 
-<!-- H1 Reward network --> In this study, we aim to investigate possible dynamics of within-reward network connectivity before and after bariatric surgery.
+Weight regulation and energy expenditure is dependent on a complex interplay of homeostatic signal and internal sensory processing.
+Obesity may be cause or effect of a decompensation of the psychophysiological system that ensures homeostasis.
+The default mode network is commonly associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
+<!-- parietal activation correlated with appetite measures after an overnight fast [@Tregellas_2011].
+--> It is mainly composed of the posterior cingulate cortex (PCC)/ Precuneus (PC), the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
+BA 39) [@Raichle_2015].
+The right ventral anterior cingulate cortex (VACC), superior frontal cortex (SFC), and inferior temporal cortex (ITC) show weaker but also reliable association with the PCC over the course of one year.
+Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
+<!-- difference lean obese --> 
+Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].
+Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
+A longitudinal study that confirmed the stronger connectivity (of PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG).
+However, default mode connectivity in these regions seemed weakened over the course of one year.
+Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable 1 year after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
+This is consistent with other findings of reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC [@Li_2018] <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery.
+
+<!-- argument: FC ~ BMI --> The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
+Among patients undergoing surgery, change in BMI was negatively correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
+
+On the other hand, connectivity also served as a predictor of weight loss after surgery.
+On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011].
+
+<!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
+Yet, to our knowledge there is no quasi-experimental study, comparing patients that only differs in regard to the intervention, i.e.
+undergoing bariatric surgery, over an extended time period.
+<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery.
 Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
-<!-- H2 Default mode network --> In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
-Functional connectivity of this network has been shown to be linked to obesity [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].
+In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
+Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
 
-<!-- to read in literature -->
-<!-- something about surgeries -->
-Roux-en-Y gastric bypass (RYGB) is the most common procedure, as it not only restricts the food intake by reducing stomach volume, but also affects the absorption of nutrients and hormonal profile [@].
-Other common procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption.
-
-... 
-group effect because direct effect of intervention surgery -> rsfmri (cares lessfreq, less preoccupation with hunger, less/more reward sensitivity) vs.
-sugery -> BMI -> rsfmri (neural signalling, fatty tissue)
-...
+<!-- group effect because direct effect of intervention surgery -> rsfmri (cares lessfreq, less preoccupation with hunger, less/more reward sensitivity) vs.
+sugery -> BMI -> rsfmri (neural signalling, fatty tissue) -->
 
@@
 These were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
-type-2 diabetes, hypertension, non-alcoholic fatty liver disease).
+type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
 Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
@@
 Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tabsample)).
+
 <!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; ages `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
@@
 and so on-->.
-Details have been published on [github]()
+Details have been published on github [https//:.???.com]()
 <!-- anatomical preprocessing write something about anatomical preprocessing (Freesurfer etc.) here --> 
-
 The T1-weighted image was segmented to remove non-brain tissue and together with the functional images normalized to a default brain template from the Monteal Neurological Institute (MNI) space.
@@
 As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but is known to introduces spatial dependency and spurious correlations, we performed GSR after ICA-AROMA in a separate sensitivity analysis.
+
+### Seed region of interest (ROI)
 <!-- Regions of interest (definition and transformation into freesurfer long space) -->
+
+### Connectivity maps
 <!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint (more details)) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC).
@@
 
-<!-- MEASURES AND COVARIATES --> The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
+The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
 
@@
 
-<!-- MODEL SETIP --> Two theoretically derived models were set up.
+<!-- MODEL SETUP --> Two theoretically derived models were set up.
 The first model was aimed at testing the interaction of time and group: $FC = group + time + group*time$.
@@
 When reducing the number of nuisance covariates by leaving out mean FD, connectivity was found between the PCC with one other voxel.
-In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm]) yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
+In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm] yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
 However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{FWEcorr} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
@@
 
-<!-- SUMMARY
--->
+<!-- SUMMARY -->
+Utilizing resting-state functional MRI on morbidly obese individuals that either underwent or awaited bariatric surgery over the course of one year.
 <!-- general --> Building on the evidence for post-surgery changes in functional activity, this study explored the longitudinal relationship between surgery-induced weight loss and functional integrity in target networks indicated by resting-state functional connectivity.
@@
 
-<!-- SUPPORT OF ORIGINAL HYPOTHESIS 
--->
+<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
+<!-- false positive -->
+To summarize, there were few significant co-activation that emerged from the marginal models.
+Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
+Given the multiplicity of of models resulting from variations of inspected networks and sensitivity analysis for different preprocessing and correction approached, the obtained significant correlations between seed regions and voxels are likely to be false positive.
+Ventures to verify effects by re-calculating with non-parametric Wild bootstrap confirmed supported presumption as effects were not robust across estimation procedures.
 
-<!-- 
-SIMILARITY OF RESULTS -->
+<!-- SIMILARITY OF RESULTS -->
 Our results are not consistent with the present literature.
@@
 
-<!-- 
-INTERPRETATION -->
+<!-- INTERPRETATION -->
 <!-- absence of evidence --> However, hypotheses about the plasticity functional networks following bariatric surgery could not be validated.
@@
 
-<!-- 
-GENERALIZABILITY -->
-
-<!-- advantage:  -->
-
-<!-- no RCT --> Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
-<!-- different conclusion from other studies --> To isolate the effect of the surgical intervention to a maximum possible extent, the intervention group has to be comparable to a control group in regard  to predisposition.
+<!-- GENERALIZABILITY -->
+<!-- advantage: prospective intervention control study --> 
+Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
+To isolate the effect of the surgical intervention to a maximum possible extent, the intervention group has to be comparable to a control group in regard  to predisposition.
 Studies comparing the sample of bariatric patients to lean individuals or behavioral dieters may have investigated risk or contributing factors to the condition or other dispositional difference between groups, e.g.
@@
 However, they provide insufficient evidence to directly evaluate impact of the surgical procedure.
-<!-- prospective intervention control study --> Prospective non-randomization trial study design are secondary to randomized control trials in regard to internal control.
+Prospective non-randomization trial study design are secondary to randomized control trials in regard to internal control.
 By establishing temporality, longitudinal measurements support the validity of conclusions about the effects of the surgery.
@@
 
-<!-- advantage: representative sample --> Size and composition of our sample did not allow separate analyses for the sexes.
+<!-- advantage: representative sample --> 
+Size and composition of our sample did not allow separate analyses for the sexes.
 The disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
@@
 <!-- mixed effect model -->
-
 We re-ran analyses with a flexible factorial model as implemented in SPM to estimate the group-time-interaction while controlling for variance on subject-level as done exemplarily by @McFarquhar_2019.
@@
 <!-- demonstrated / we ruled out gross missspecification in retrieving estimates from the Sandwich estimator. -->
-
 @McFarquhar_2019 demonstrated that specifying complex multi-level models with standard neuroimaging software packages such as SPM and FSL can be quite difficult if done appropriately. 
@@
 In both models, the variation between subjects was modeled and therefore ...
-
 Opposed to flexible factorial models for analysis of variance, where each individual is modelled separately, models are parsimonious and preserver more degrees of freedom.
 
-<!-- 
-IMPLICATIONS -->
-
+<!-- IMPLICATIONS -->
 <!-- Relevance of rsfMRI as biomarker -->
 <!-- Consideration of head motion as covariate -->
-
-<!-- Problem head motion --> Based on concerns expressed in the literature [@], we tried to closely inspect and rigorously control for the influence of head motion.
+<!-- Problem head motion --> 
+Based on concerns expressed in the literature [@], we tried to closely inspect and rigorously control for the influence of head motion.
 The presence of multicollinearity of our predictors has been established before.
@@
 Confidence in $\beta$ estimates will suffers and estimates can behave erratic depending on which regressors included in the model.
-
-To summarize, there were few significant co-activation that emerged from the marginal models.
-Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
-Ventures to verify effects by recalculating with non-parametric Wild bootstrap revealed that effects were not robust across estimation procedures.
-This suggests that significant results are likely to be false positive.
 
@@
 The code includes script for data preparation and analysis of different models.
+
+# Declaration of interest
+# Funding
+# Acknowledgments
+# Author Contributions
+
+Stud conception and design
+Aqcuisiiton of data:
+Analysis andinterpretation of data: 
+Frafting of manuscript: 
+Ciritical revision:
 

### 2020-11-29 draft_rsfmri.Rmd
update introduction

@@
 
+<!-- obesity and interventions -->
 Over the past decades, the global burden of obesity has increasingly been confronted as a matter of urgency by research.
+_something about comorbidities_
+Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018]. 
+Luckily, dietary interventions such as caloric restriction and related metabolic improvements, and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018; @Veronese_2017].
+In some cases behavioral intervention fail to yield the desired success regarding weight loss. There are non-behavioral alternatives at hand.
+<!-- bars most invasive (brain metabolism), last resort -->
+Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes.
+Roux-en-Y gastric bypass (RYGB) is the most common procedure as it not only restricts the food intake by reducing stomach volume but also affects the absorption of macronutrients and the hormonal profile [@].
+<!-- ref for "common", ref for procedure -->
+Other widely used procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption. <!-- mehr zu Veränderung in Gasto-intestinaler Physiologie -->
 
-Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018].
-On the other hand, dietary interventions such as caloric restriction and related metabolic improvements and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018; @Veronese_2017].
-Behavioral intervention require ... and often fail to yield the desired success regarding weight loss, especially in in morbid obesity [@]. There are alternative interventions at hand.
-<!-- bars -->
-Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions and co-morbidities, such as diabetes.
-Roux-en-Y gastric bypass (RYGB) is the most common procedure, as it not only restricts the food intake by reducing stomach volume, but also affects the absorption of macronutrients and the hormonal profile [@].
-Other common procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption.
 <!-- Potential mechanisms? -->
-Precise mechanisms are yet to be elucidated, but as a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed to better infer potential mechanisms and support interventions proven to be effective.
-<!-- / immitate interventions that have been proven effective with less invasive alternatives -->
+Precise mechanisms are yet to be elucidated.
+As a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed.
+This will inform inferences about potential mechanisms and either support interventions proven to be effective or ideally allow imitating them will less invasive alternatives.
+
+Recently, research focuses have slowly shifted to investigate brain metabolism structural and functional changes after surgery.
+<!-- severe nutritional stress [@Crujeiras_2017] -->
 
 <!-- Reward network -->
-Little is known about the potential beneficial effects of bariatric surgery on brain function in this population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
 <!-- mechanism reward and appetite, overating -->
@@
 
-Weight regulation and energy expenditure is dependent on a complex interplay of homeostatic signal and internal sensory processing.
+Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
 Obesity may be cause or effect of a decompensation of the psychophysiological system that ensures homeostasis.
@@
 <!-- parietal activation correlated with appetite measures after an overnight fast [@Tregellas_2011].
---> It is mainly composed of the posterior cingulate cortex (PCC)/ Precuneus (PC), the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
+--> It is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
 BA 39) [@Raichle_2015].
@@
 On the other hand, connectivity also served as a predictor of weight loss after surgery.
+
 On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011].
@@
   scale_fill_manual(labels = c("female", "male"),
-                    values = c("#00305E55", "#C0004555"),
+                    values = c("#00305E55", "#C0004555")
   )

### 2020-11-30 draft_rsfmri.Rmd
change in reward network section

@@
 
-<!-- Potential mechanisms? -->
+<!-- Neural mechanisms/concurrent plasticity? -->
 Precise mechanisms are yet to be elucidated.
@@
 Those psychophysiological and behavioral effects may be related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
-Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which a sustainable weight loss [@Mulla_2017, @Nudel_2019].
-
-Several studies have demonstrated differences in resting-state activity in reward regions depending on weight status.
-
+Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017, @Nudel_2019].
+The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in deciision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+<!-- study results -->
+Accordingly, several studies have found differences in resting-state activity in reward regions depending on weight status.
 Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
@@
 
+On the other hand, BMI prior to the surgery was associated with changes in connectivity [@Duan_2020].
+
 <!-- saliency and reward -->
-In lean participants, food valuation processes in decision-making have been linked to brain activity and activity within the reward network, including ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
 

### 2020-11-30 draft_rsfmri.Rmd
discussion of marginal model/ SPM vs. SwE

@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
+# [ ] SPM model comparison
 
@@
 Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017, @Nudel_2019].
-The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in deciision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
-<!-- study results -->
-Accordingly, several studies have found differences in resting-state activity in reward regions depending on weight status.
+<!-- results in lean -->
+The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+<!-- results in obese longitudinal -->
+Similarly, several studies have found differences in resting-state activity in reward regions depending on weight status.
+
 Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
@@
 ### Seed region of interest (ROI)
-<!-- Regions of interest (definition and transformation into freesurfer long space) -->
+<!-- Regions of interest (definition and transformation into freesurfer long space) --> Seed regions were ...
+mm spheres in the precuneus (MNI: x,y,z = x,x,x) associated with the default mode network, and the Nucleus Accumbens (MNI: x,y,z = x,x,x) by averaging the time series across voxels.
 
@@
 It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
+prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016]. putamine, insula
 
@@
 
-<!-- advantage: representative sample --> 
+<!-- advantage: Nucleus Accumbensrepresentative sample --> 
 Size and composition of our sample did not allow separate analyses for the sexes.
@@
 <!-- mixed effect model -->
-We re-ran analyses with a flexible factorial model as implemented in SPM to estimate the group-time-interaction while controlling for variance on subject-level as done exemplarily by @McFarquhar_2019.
-Thereby,
-<!-- demonstrated / we ruled out gross missspecification in retrieving estimates from the Sandwich estimator. -->
-@McFarquhar_2019 demonstrated that specifying complex multi-level models with standard neuroimaging software packages such as SPM and FSL can be quite difficult if done appropriately. 
-From the necessity to include subject effect, the necessity of using a overparamaterized design arises.
+Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), often more modern models such as mixed-effect models and marginal models are recommended [@Chen_2013;@Guillaume_2014;@McFarquhar_2019].
+@McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
+However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the group-time-interaction with age and sex as nuisance covariates.
+<!-- (FD is time-varying...
+schwierig) with a simplified model without nuisance covariates in SPM12.
+We opted for a simplified model because different nuisance covariates, time-invariant or time-varying, create additional complications.
+-->
+An advantage of the group-time interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
+The model was set up as flexible factorial design with factors "subject", "group", and "time".
 
-Not only do  but modeling between-subjects and within-subject nuisance covariates demands caution and becomes even more complex when using multiple within-subject covariates.
-....
-In both models, the variation between subjects was modeled and therefore ...
-Opposed to flexible factorial models for analysis of variance, where each individual is modelled separately, models are parsimonious and preserver more degrees of freedom.
+Differences may arise due to adjustments to non-spherity, but to rule out any further differences due to correction procedure, both models were evaluated with an uncorrected $p < 0.01$.
+Estimation was carried out via restricted maximum likelihood for $166$ voxels to correct for non-spherity.
+Concernes about comparability have been expressed with other toolboxed have been expressed elsewhere [@McFarquhar_2019].
+
+Estimates retrieved with the Sandwich estimator ... comparable with 
+In both models, the retrieved estimates between subjects was modeled and therefore ...
+
+Opposed to flexible factorial models for analysis of variance, where each individual is modelled separately, models are parsimonious and preserve more degrees of freedom. Furthermore,
 

### 2020-11-30 draft_rsfmri.Rmd
few updates

@@
 
+# [ ] Evaluate SPM model comparison (more activation in SPM..)
 # [ ] Introduction
@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] SPM model comparison
 
@@
 library(psych)
-#library(lsr) #
-
+#library(lsr) 
 #library(wesanderson) # colour palette
@@
 
-<!-- saliency and reward -->
-
-Studies concerning such processes following a surgical intervention are inconclusive.
+s concerning such processes following a surgical intervention are inconclusive.
 
@@
 <!-- mixed effect model -->
-Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), often more modern models such as mixed-effect models and marginal models are recommended [@Chen_2013;@Guillaume_2014;@McFarquhar_2019].
+Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), it is often recommended to use modern models such as mixed-effect models and marginal models [@Chen_2013;@Guillaume_2014;@McFarquhar_2019].
 @McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
@@
 An advantage of the group-time interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
-The model was set up as flexible factorial design with factors "subject", "group", and "time".
-
-Differences may arise due to adjustments to non-spherity, but to rule out any further differences due to correction procedure, both models were evaluated with an uncorrected $p < 0.01$.
-Estimation was carried out via restricted maximum likelihood for $166$ voxels to correct for non-spherity.
-Concernes about comparability have been expressed with other toolboxed have been expressed elsewhere [@McFarquhar_2019].
-
-Estimates retrieved with the Sandwich estimator ... comparable with 
-In both models, the retrieved estimates between subjects was modeled and therefore ...
-
-Opposed to flexible factorial models for analysis of variance, where each individual is modelled separately, models are parsimonious and preserve more degrees of freedom. Furthermore,
+The model was set up as flexible factorial design with factors "subject", "group", and "time" assuming unequal group variances.
+<!-- vastly different - different implicit masks? -->
+Differences could be attributed to different corrections for non-spherity and assumptions on voxel covariance structures.
+Detailed comparisons between SPM12, SwE, and other software packages have been described elsewhere [@McFarquhar_2016].
+<!-- conclusion??? -->
+Opposed to flexible factorial models, where each individual is modelled separately, models are parsimonious and preserve more degrees of freedom.
 

### 2020-11-30 draft_rsfmri.Rmd
fix p-value format

@@
 In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm] yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
-However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{FWEcorr} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
+However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{\mathrm{FWEcorr}} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
 -->.

### 2020-12-01 draft_rsfmri.Rmd
update novelty statement

--- 2020-11-30 draft_rsfmri.Rmd
@@
 <!-- bars most invasive (brain metabolism), last resort -->
-Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes.
+Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes [@].
 Roux-en-Y gastric bypass (RYGB) is the most common procedure as it not only restricts the food intake by reducing stomach volume but also affects the absorption of macronutrients and the hormonal profile [@].
@@
 <!-- Default mode network --> 
-
 Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
@@
 <!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
-Yet, to our knowledge there is no quasi-experimental study, comparing patients that only differs in regard to the intervention, i.e.
-undergoing bariatric surgery, over an extended time period.
-<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery.
+Yet, to our knowledge there is no quasi-experimental study, comparing a cohort of patients to a cohort with similar characteristics but unexposed to the intervention, i.e.
+bariatric surgery, over an extended time period as would be demanded to draw more substantiated inference [@George_2016].
+<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery by comparison with a contemporaneously measured waiting-control group.
 Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.

### 2020-12-01 draft_rsfmri.Rmd
specified package versions

@@
 # load packages
-library(plyr)
-library(dplyr)
-library(tidyr)
-library(kableExtra)
-library(ggplot2)
-library(patchwork) # API for sequentially building up a plot (similar to gridExtra and cowplot)
+library(car) # version 3.0.9
+library(plyr) # version 1.8.6
+library(dplyr) # version 1.0.2
+library(tidyr) # version 1.1.2
+library(kableExtra) # version 1.2.1
+library(ggplot2) # version 3.3.2
+library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
 #library(cowplot)
-library(RColorBrewer)
-library(psych)
+library(RColorBrewer) # version 1.1.2
+library(psych) # version 2.0.8
 #library(lsr) 
@@
 
-library(knitr)
+library(knitr) # version 1.30
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
@@
   theme_bw()
-tab_age <- describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
+tab_age <- psych::describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
 # Sex

### 2020-12-02 draft_rsfmri.Rmd
minor changes
no changes
modified Introduction of Rmd

--- 2020-12-01 draft_rsfmri.Rmd
@@
 author:
-- xxx [1]
-- xxx [1][2]
-- xxx [1][2][3]
-- xxx [1][2]
+- Hannah Heinrichs* [1]
+- Frauke Beyer* [1][2]
+- Kerstin Prehn 
+- Agnes Floel 
+- Arno Villringer [1][2][3]
+- A. Veronica Witte [1][2][3]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
@@
 
+#set wd (change depending on user)
+setwd("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/")
+
 # load packages
@@
 # Introduction
+<!--Suggested Order of Introduction
+1.
+Obesity is a major health concern.
+[...
+some more details about metabolic and mortality risk, especially related to morbid obesity ...] More and more recognized, midlie obesity also impacts later cognitive decline and dementia risk.
+Obesity has been associated with differences in brain structure and function.
+Regarding structural alterations, longitudinal studies have indicated that higher body mass index is related to increased cortial thinning in midlife (path through life study), potentially mediated by inflammation or glucose dysregulation...
+Similarly, white matter microstructure and may be related to adverse metabolic effects (lampe)
+Differences in brain function in obesity have been reported in the reward network during task-based fMRI, highlighting the role of reward processing in obesity development and maintainance.
+Resting state fMRI studies have also shown differences in connectivity in diverse brain networks, including reward and salience network, visual networks and the default mode network.
+In part, these differences have been related to metabolic and behavioral differences, yet there may be an important confounding due to head motion.
+In this study, we aim to investigate the impact of bariatric surery on..
+-->
 
@@
 <!-- bars most invasive (brain metabolism), last resort -->
-Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes [@].
+Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes.
 Roux-en-Y gastric bypass (RYGB) is the most common procedure as it not only restricts the food intake by reducing stomach volume but also affects the absorption of macronutrients and the hormonal profile [@].
@@
 <!-- Default mode network -->
+
 Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
@@
 <!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
-Yet, to our knowledge there is no quasi-experimental study, comparing a cohort of patients to a cohort with similar characteristics but unexposed to the intervention, i.e.
-bariatric surgery, over an extended time period as would be demanded to draw more substantiated inference [@George_2016].
-<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery by comparison with a contemporaneously measured waiting-control group.
+Yet, to our knowledge there is no quasi-experimental study, comparing patients that only differs in regard to the intervention, i.e.
+undergoing bariatric surgery, over an extended time period.
+<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery.
 Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
@@
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
+```{r test wd}
+print(getwd())
+```
+
 ```{r baseline characteristics, include=FALSE}
@@
   theme_bw()
-tab_age <- describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
+tab_age <- describeBy(x=dfBL$Age_BL, group=dfBL$condition, mat=T)
 # Sex

### 2020-12-02 get_subject_ID_group_tp_forSwE.R

@@ -16,6 +16,7 @@ save_txt_for_swe <- function(group = "all", tp = "all"){
   parentdir <- file.path(abs_path, swe_path, fsep = .Platform$file.sep)[1]
   setwd(parentdir)
   
+
   files=read.table("../SwE_files/scans_PCC_CC_z.txt")
   #split info from path in txt
   for (i in 1:nrow(files)){

### 2020-12-02 draft_rsfmri.Rmd

--- 2020-12-01 draft_rsfmri.Rmd
@@
 ```{r "environment setup", include=FALSE}
+
+#set wd (change depending on user)
+setwd("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/")
 
@@
 
-<!-- argument: FC ~ BMI --> The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
+The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
 Among patients undergoing surgery, change in BMI was negatively correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
@@
   theme_bw()
+
 tab_age <- psych::describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
+
 # Sex

### 2020-12-02 intro_rsfmri.Rmd
introduction to manuscript saved and edited separately

--- 2020-12-01 draft_rsfmri.Rmd
@@
 ```
-\newpage
 # Abstract
@@
 Over the past decades, the global burden of obesity has increasingly been confronted as a matter of urgency by research.
-_something about comorbidities_
 Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018]. 
@@
 Other widely used procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption. <!-- mehr zu Veränderung in Gasto-intestinaler Physiologie -->
+<!-- Neural mechanisms/concurrent plasticity? -->
+Precise mechanisms as to how bariatric surgery leads to recovery are yet to be elucidated.
+As a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed.
+This will inform inferences about potential mechanisms and either support interventions proven to be effective or ideally allow imitating them with less invasive alternatives.
 
-<!-- Neural mechanisms/concurrent plasticity? -->
-Precise mechanisms are yet to be elucidated.
-As a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed.
-This will inform inferences about potential mechanisms and either support interventions proven to be effective or ideally allow imitating them will less invasive alternatives.
-
-Recently, research focuses have slowly shifted to investigate brain metabolism structural and functional changes after surgery.
 <!-- severe nutritional stress [@Crujeiras_2017] -->
@@
 <!-- Reward network -->
-Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
-<!-- mechanism reward and appetite, overating -->
-Besides malabsorption and restricted food ingestion, patients report post-surgical changes in food preferences and eating behavior.
-Those psychophysiological and behavioral effects may be related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
-Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017, @Nudel_2019].
+Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population.
+Thus, one recently emerging research focus is how brain metabolism and functional connectivity changes in response to surgery, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+The reward network comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+These regions also seem to be altered after bariatric surgery [@Lips_2014; @Hogenkamp_2016; @Garcia_2012].
+<!-- Zitationen checken -->
+<!-- Normalization of connectivity --> Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
+But these differences seemed to normalize, as trends reversed when looking at the resting state connectivity after compared to before the surgery [@Duan_2020].
+Resting state activitiy increased in subcortical regions, such as the insula and striatum<!-- specifically putamen --> [@Wiemerslage_2016].
+<!-- mechanism reward and appetite, overeating --> Interestingly, patients report changes in food preferences and eating behavior after the surgery that may influence weight loss beyond malabsorption of nutrients and restricted food ingestion.
+Those psychophysiological and behavioral effects are presumably related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
+Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017; @Nudel_2019].
 <!-- results in lean -->
-The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+
 <!-- results in obese longitudinal -->
-Similarly, several studies have found differences in resting-state activity in reward regions depending on weight status.
+<!-- Similarly, several studies have found differences in resting-state activity in reward regions depending on weight status.  -->
 
-Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
-These differences seemed to normalize, as trends reversed when looking at the resting state of obese individuals after compared to before the surgery [@Duan_2020].
+<!-- On the other hand, BMI prior to the surgery was associated with changes in connectivity [@Duan_2020]. -->
 
-On the other hand, BMI prior to the surgery was associated with changes in connectivity [@Duan_2020].
-
-s concerning such processes following a surgical intervention are inconclusive.
+<!-- s concerning such processes following a surgical intervention are inconclusive. -->
 
@@
 <!-- Default mode network --> 
-Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
-Obesity may be cause or effect of a decompensation of the psychophysiological system that ensures homeostasis.
-The default mode network is commonly associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
+Because multiple resting state networks are relevant to obesity and associated with abnormal neurocircuitry interactions [@Ding_2020], extensive research has been devoted to investigating the the relationship between the control network and the reward network.
+However, the dynamics within the default network may also be of importance because it is associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
 <!-- parietal activation correlated with appetite measures after an overnight fast [@Tregellas_2011].
---> It is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
+-->
+Weight regulation and energy expenditure, on the other hand, are highly dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation. 
+<!-- Obesity = decompensation of the psychophysiological system that ensures homeostasis. -->
+
+The default mdoe network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
 BA 39) [@Raichle_2015].
@@
 Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
-<!-- difference lean obese --> 
-Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].
+<!-- study results --> 
+Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2017].
 Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
-A longitudinal study that confirmed the stronger connectivity (of PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG).
-However, default mode connectivity in these regions seemed weakened over the course of one year.
-Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable 1 year after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
-This is consistent with other findings of reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC [@Li_2018] <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery.
+Comparing spontaneous activitiy before and after surgery shows a decrease in areas including superior frontal regions<!-- gehört zu SFC -->, and a decrease superior parietal lobule when in sated state [@Wiemerslage_2016].
+Concerning spontaneuous functional connectivity in these regions, patient having undergone bariatric surgery resemble normal-weight individuals rather than obese individual in having a weaker connectivity within the default mode network [@Frank_2013].
+A longitudinal study investigating confirmed that while there was stronger connectivity in the default model network <!-- (PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG) --> in obese patients before the surgery, it seemed weakened over the course of one year.
+Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
+Consistent with these findings, other studies reported a reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery [@Li_2018].
 
-<!-- argument: FC ~ BMI --> The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
-Among patients undergoing surgery, change in BMI was negatively correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
+<!-- The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated and especially the role of weight and weight change. -->
+... 
+It has been stated that among patients undergoing surgery, change in BMI was<!-- negatively --> correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
+On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011]. ...
 
-On the other hand, connectivity also served as a predictor of weight loss after surgery.
+<!-- On the other hand, connectivity also served as a predictor of weight loss after surgery. -->
 
-On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011].
+<!--  -->
 
@@
 Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
+Notably, we will consider that different courses in increase and decrease of functional connectivity that are not necessarily linear are possible over the course of one year, depending on the phase of weight management [@Olivo_2017].
 
@@
 \newpage
-# Methods
-
-```{r Load data, include=FALSE}
-source("get_subject_ID_group_tp_forSwE.R")
-final = save_txt_for_swe(group = "all", tp = "all")
-
-# sample size
-#nrow(subj_df)
-subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each subject
-subj_df <- final[subj_idx,] # df with unique subjects
-
-# completed time points
-freqtable <- as.data.frame(table(final$subj.ID))
-tpfreq <- table(freqtable$Freq)
-
-# overview distribution of data points
-tp1 <- as.character(freqtable$Var1[freqtable$Freq == 1])
-tp2 <- as.character(freqtable$Var1[freqtable$Freq == 2])
-tp3 <- as.character(freqtable$Var1[freqtable$Freq == 3])
-tp1_tab <- table(subset(final, subj.ID %in% tp1)$tp, subset(final, subj.ID %in% tp1)$condition)
-tp2_tab <- table(subset(final, subj.ID %in% tp2)$tp, subset(final, subj.ID %in% tp2)$condition)[c(1,3),]
-tp3_tab <- table(subset(final, subj.ID %in% tp3)$tp, subset(final, subj.ID %in% tp3)$condition)[1,]
-total_datapoints <- c(sum(final$group==1),sum(final$group==2))
-total_subjects <- plyr::count(final[subj_idx,"group"])$freq
-tab_sample <- data.frame(rbind(tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints))
-rm(tp1,tp2,tp3,tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints)
-rownames(tab_sample) <- c("count: only 0","count: only 6","count: only 12","count: 0 and 6","count: 6 and 12","count: complete data","total number of subjects","total data points")
-colnames(tab_sample) <- c("BARS","NBARS")
-
-# distribution of condition and sex
-subj_df$Sex <- plyr::mapvalues(subj_df$Sex, from = c("-1", "1"), to = c("f", "m"))
-cs <- addmargins(table(subj_df$Sex,subj_df$condition))
-
-dfw <- tidyr::pivot_wider(data = final,
-                          id_cols = c("subj.ID","condition","Sex","Age_BL"),
-                          names_from = "tp",
-                          values_from = c("BMI","logmFD"))
-```
-```{r tabsample, echo=FALSE}
-tab_sample <-
-  knitr::kable(
-    tab_sample,
-    row.names = TRUE,
-    format = "latex",
-    booktabs = T,
-    caption = "Distribution of data points at months after intervention")
-tab_sample
-```
-
-## Sample and study design
-
-<!-- STUDY DESIGN --> The ADIPOSITAS-Study conducted in the Charité Berlin is a observational, prospective and longitudinal design described in detail by @Prehn_2020.
-Data collection is still ongoing but only data available in April 2019 is used.
-The procedures followed were in accordance with the Helsinki Declaration.
-This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/mzyxv](https://osf.io/mzyxv).
-<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
-These were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
-type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
-<!-- Frage: Ausschluss bei head movement?
-ADI009_fu2 nicht enthalten (siehe Readme preprocessing) -->
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tabdescr)).
-`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
-The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing only subjects that have received MRI.
-Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tabsample)).
-
-<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; ages `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
-The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
-```{r baseline characteristics, include=FALSE}
-dfBL <- save_txt_for_swe(group = "all", tp = "BL")
-
-# distributions per group
-# BMI 
-p <- ggplot(data=dfBL, aes(group = condition, x=BMI, fill=condition)) +
-  geom_density(alpha=0.6) + 
-  scale_fill_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")) +
-  theme_bw()
-# log mean FD
-p <- ggplot(data=dfBL, aes(group = condition, x=logmFD, fill=condition)) +
-  geom_density(alpha=0.6) + 
-  scale_fill_manual(
-    values = c("#00305E55", "#C0004555"),
-    labels = c("intervention", "control")) +
-  theme_bw()
-# Age
-p <- ggplot(data=dfBL, aes(group = condition, x=Age_BL, fill=condition)) +
-  geom_density(alpha=0.6) + 
-  scale_fill_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")) +
-  theme_bw()
-tab_age <- psych::describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
-# Sex
-p <- ggplot(data = dfBL, aes(x = condition, fill = Sex)) + 
-  geom_bar(position = "dodge") +
-  scale_fill_manual(labels = c("female", "male"),
-                    values = c("#00305E55", "#C0004555")
-  )
-tab_sex <- table(dfBL$condition, dfBL$Sex)
-```
-Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
-n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
-`r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
-```{r tabdescr, echo=FALSE}
-
-# table for descriptive statistics
-descr <- psych::describeBy(dfw[,5:ncol(dfw)], group=dfw$condition,mat=T)
-descr$predictor <- c(rep("BMI",6),rep("logmFD",6))
-descr$tp <- car::recode(descr$vars, "c(1,4)='fu2';c(2,5)='fu';c(3,6)='bl'")
-descr$group <- descr$group1
-descr <- descr[,c("predictor","tp","group","n","mean","sd")]
-descr <- descr[with(descr, order(predictor, tp)),]
-descr[,c("mean","sd")] <- round(descr[,c("mean","sd")],2)
-descr[c(2:6,8:12),"predictor"] <- ""
-descr[c(2,4,6,8,10,12),"tp"] <- ""
-rownames(descr) <- NULL
-
-tab_descr <-
-  knitr::kable(
-    descr,
-    row.names = FALSE,
-    format = "latex",
-    booktabs = T,
-    caption = "BMI and mean FD of participants acquired in each condition and time point")
-
-tab_descr
-
-```
-<!-- STUDY PROCEDURE -->
-Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for STANDARDIZED??
-breakfast.
-MRI scanning was done after further psychological tests (for details see @Prehn_2020).
-
-```{r fig_BMIdescr, fig.align='center', out.width="90%"}
-# eval = TRUE, fig.asp = 0.7, fig.width = 6
-condition.labs <- c("BARS","NBARS") 
-names(condition.labs) <- c("IG","KG")
-# BMI over time each group
-fig_BMIdescr <-
-  ggplot(final, aes(
-    x = tp,
-    y = BMI,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig_BMIdescr +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
-```
-
-## fMRI acquisition
-
-MRI images were retrieved with a 12-channel head coil of a 3 Tesla Siemens Trio system.
-<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, TR = 1900 ms, TE = 2.52 ms, a = 9, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
-<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3 s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes<!-- dimension = [60;72;60] ???
--->.
-The total acquisition time was 5:45 minutes.
-Field map or phase-encode reversed images were not acquired.
-Participants were instructed to close their eyes but to remain awake during scanning.
-
-## Pre-proceessing
-
-Data preprocessing was conducted with FSL and Freesurfer using a nipype (version 14.0.2) workflow implemented in python 2.7. 
-The preprocessing workflow included anatomical, functional and diffusion-weighted imaging preprocessing and wrapped modules from FSL version 6.0.1<!-- ?!
--->, Freesurfer version 6.0.0p1, ANTS <!-- version?!
-and so on-->.
-Details have been published on github [https//:.???.com]()
-<!-- anatomical preprocessing write something about anatomical preprocessing (Freesurfer etc.) here --> 
-The T1-weighted image was segmented to remove non-brain tissue and together with the functional images normalized to a default brain template from the Monteal Neurological Institute (MNI) space.
-Functional images were smoothed with an $0$ mm FWHM Gaussian kernel.
-
-<!-- minimal functional preprocessing --> Preprocessing pipeline included removal of first four volumes and a transformation of the blood oxygen level dependent (BOLD) time series into the subject's anatomical space.
-This transform combines parameters from motion correction (FSL's MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL's FUGUE) and coregistration to the subject's anatomical template space (using Freesurfer's bbregister and the transform from individual to longitudinal template space in Freesurfer's longitudinal processing stream).
-
-<!-- Aroma + CC/ + GSR --> Volumes were further preprocessed with ICA-AROMA and compCor (CC) and, optionally, with a subsequent global signal regression (GSR) [@Ciric_2017; @Parkes_2018] to reduce the effect of physiological noise and head motion on the connectivity estimates.
-As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but is known to introduces spatial dependency and spurious correlations, we performed GSR after ICA-AROMA in a separate sensitivity analysis.
-
-### Seed region of interest (ROI)
-<!-- Regions of interest (definition and transformation into freesurfer long space) --> Seed regions were ...
-mm spheres in the precuneus (MNI: x,y,z = x,x,x) associated with the default mode network, and the Nucleus Accumbens (MNI: x,y,z = x,x,x) by averaging the time series across voxels.
-
-### Connectivity maps
-<!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint (more details)) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC).
-Subsequently, r-to-z transform was applied to the correlation to obtain normally distributed functional connectivity (using nilearn, nipype and python).
-Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).
-
-### Quality Assessment
-
-<!-- EXCLUSION CRITERIA --> Participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
-caused Freesurfer to fail or perform very poorly, i.e.
-rated "Failed" according to @Klapwijk_2019) were excluded because a subject's motion is highly similar across scans and we thus suspect the fMRI data to also suffer from motion artifacts in these subjects.
-As the expected average head motion is high [@Hodgson_2016], we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample.
-<!-- MOTION CORRECTION --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].
-Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012, log-transformed and examined within and separate from the main models.
-
-### Outcomes
-
-The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
-
-## Analysis
-
-Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
-During analyses an implicit mask and an explicit mask were used.
-Therefore, the MNI ICBM 2009a Nonlinear Symmetric brain mask [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
-
-<!-- MODEL SETUP --> Two theoretically derived models were set up.
-The first model was aimed at testing the interaction of time and group: $FC = group + time + group*time$.
-The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.
-These were examined each for correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $16$ a priori planned analyses.
-Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$.
-As a result a total of $32$ different analyses were performed.
-Computation with the grey matter mask were exploratory and deviate from the preregistration.
-
-<!-- Description marginal model --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
-It therefore accommodates any repeated measurement covariance structures.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
-
-<!-- Design matrix, vielleicht beide auch als Abbildung kurz zeigen?
---> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
-The resulting flexible factorial design <!-- check: difference from full factorial or other?
---> contained one regressor for each time point per group.
-For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
-With a model containing average BMI and longitudinal change per subject, we investigated potential effects ..............
-on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- Nuisance covariates --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
-Age and sex were between-subject covariates while FD is time-varying.
-Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-As change in FD and change in BMI share variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
-<!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
-Differences in results will be reported.
-For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
-These post-hoc analyses deviate from the preregistration and should be considered to be exploratory.
-
-<!-- Inference --> All parametric inferences on activation were drawn based on False Discovery Rate (FDR) adjusted p-values $\leq 0.05$.
-To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
-In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and is unlike permutation based framework appropriate <!-- ???
-clarify why--> for longitudinal data.
-
-# Results
-
-## Manipulation check 
-<!-- check network presence -->
-
-<!-- singletp (t-tests), nur diesen Check für BL reporten (gib dann oben noch an, dass du dafür SPM verwendet hast) --> We ran separate one-sample t-test for network activation at baseline controlling for log mean FD, age, and sex for each time point to check whether the assumed reward and default mode network was at all identifiable.
-
-## Interaction effect of intervention and time
-
-<!-- Brain mask: Group*Time Interaction -->
-There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
-Nor was there a significant main effect for any of the covariates of interest (time, group) in voxelwise inference with FDR correction.
-Results obtained for parametric estimation and non-parametric wild bootstrap did not differ.
-When reducing the number of nuisance covariates by leaving out mean FD, connectivity was found between the PCC with one other voxel.
-In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm] yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
-However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{\mathrm{FWEcorr}} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
--->.
-There was no other significant co-activation for any other region of interest for neither of the preprocessing variants.
-
-<!-- GM mask: Group*Time Interaction -->
-Parametric and non-parametric significance test for the model testing for group-time-interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the group-time-interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{FDRcorr} = 0.998$)<!-- also FWE-corr n.s -->.
-
-## Effect of BMI
-
-<!-- Brain mask: average and individual BMI -->
-<!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
-There was no significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
-Omitting mean FD as nuisance covariate did not change these results. 
-<!-- BMIonlyIG --> Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ --> with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps).
-However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)<!-- but FWE-corr sig?!
--->.
-The effect also disappeared when running those tests without mean FD as nuisance covariate.
-<!-- BMI2tp --> Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the regions of interest for neither of the preprocessing variants.
-Omitting mean FD did not change these results.
-
-<!-- GM mask: average and individual BMI, damit die Reihenfolge so ist wie oben. -->
-Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no significant co-activation was found when using the grey matter mask.
-
-## Further investigation of the effect of FD
-
-<!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
-Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
-<!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
-
-<!-- Neurovault --> 
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
-
-# Discussion
-
-<!-- SUMMARY -->
-Utilizing resting-state functional MRI on morbidly obese individuals that either underwent or awaited bariatric surgery over the course of one year.
-<!-- general --> Building on the evidence for post-surgery changes in functional activity, this study explored the longitudinal relationship between surgery-induced weight loss and functional integrity in target networks indicated by resting-state functional connectivity.
-Over the course of one year we collected neuroimaging and anthropometric data from `r cs["Sum","IG"]` morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of  `r cs["Sum","KG"]` patients on the waiting list.
-<!-- model specific --> With the model on group-time-interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
-Beyond that, also main effects of the intervention and of the time were examined.
-<!-- correct interpretation of marginal models?
--->
-The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
-
-<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
-<!-- false positive -->
-To summarize, there were few significant co-activation that emerged from the marginal models.
-Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
-Given the multiplicity of of models resulting from variations of inspected networks and sensitivity analysis for different preprocessing and correction approached, the obtained significant correlations between seed regions and voxels are likely to be false positive.
-Ventures to verify effects by re-calculating with non-parametric Wild bootstrap confirmed supported presumption as effects were not robust across estimation procedures.
-
-<!-- SIMILARITY OF RESULTS -->
-Our results are not consistent with the present literature.
-Although there are few studies investigating the impact of bariatric surgery on functional connectivity at resting state, the present articles do significant changes in network activation [@;@;@].
-<!-- publication bias -->
-
-<!-- INTERPRETATION -->
-<!-- absence of evidence --> However, hypotheses about the plasticity functional networks following bariatric surgery could not be validated.
-Due to the shortcomings of the inferential paradigm used, our null findings may either put conclusions drawn from previous research into question or may neither provide evidence for the presence nor the absence of the effect.
-<!-- Control for satiety?
---> A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
-After the intervention, the functional connectivity was higher in bariatric patients.
-It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
-prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016]. putamine, insula
-
-<!-- GENERALIZABILITY -->
-<!-- advantage: prospective intervention control study --> 
-Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
-To isolate the effect of the surgical intervention to a maximum possible extent, the intervention group has to be comparable to a control group in regard  to predisposition.
-Studies comparing the sample of bariatric patients to lean individuals or behavioral dieters may have investigated risk or contributing factors to the condition or other dispositional difference between groups, e.g.
-in cognitive control, openness or responsiveness to behavioral interventions [@;@].
-However, they provide insufficient evidence to directly evaluate impact of the surgical procedure.
-Prospective non-randomization trial study design are secondary to randomized control trials in regard to internal control.
-By establishing temporality, longitudinal measurements support the validity of conclusions about the effects of the surgery.
-Contrasting pre- and post-surgery measures change while controlling for individual variance.
-The waiting list control group equivalent in BMI, treatment history and  treatment recommendation allows controlling for concurrent factors [@Thiese_2014].
-Incorporating these consideration, our study aimed at a higher internal validity for conclusions about the effects of bariatric surgery on functional connectivity.
-
-<!-- advantage: Nucleus Accumbensrepresentative sample --> 
-Size and composition of our sample did not allow separate analyses for the sexes.
-The disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
-
-<!-- mention larger sample size -->
-
-<!-- mixed effect model -->
-Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), it is often recommended to use modern models such as mixed-effect models and marginal models [@Chen_2013;@Guillaume_2014;@McFarquhar_2019].
-@McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
-However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the group-time-interaction with age and sex as nuisance covariates.
-<!-- (FD is time-varying...
-schwierig) with a simplified model without nuisance covariates in SPM12.
-We opted for a simplified model because different nuisance covariates, time-invariant or time-varying, create additional complications.
--->
-An advantage of the group-time interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
-The model was set up as flexible factorial design with factors "subject", "group", and "time" assuming unequal group variances.
-<!-- vastly different - different implicit masks? -->
-Differences could be attributed to different corrections for non-spherity and assumptions on voxel covariance structures.
-Detailed comparisons between SPM12, SwE, and other software packages have been described elsewhere [@McFarquhar_2016].
-<!-- conclusion??? -->
-Opposed to flexible factorial models, where each individual is modelled separately, models are parsimonious and preserve more degrees of freedom.
-
-<!-- IMPLICATIONS -->
-<!-- Relevance of rsfMRI as biomarker -->
-<!-- Consideration of head motion as covariate -->
-<!-- Problem head motion --> 
-Based on concerns expressed in the literature [@], we tried to closely inspect and rigorously control for the influence of head motion.
-The presence of multicollinearity of our predictors has been established before.
-As described in @Monti_2011 in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
-Confidence in $\beta$ estimates will suffers and estimates can behave erratic depending on which regressors included in the model.
-
-<!-- For all patients included, conservative obesity treatment had failed, e.g.
-behavioral diet intervention.
-As shown before, patients who were successful in behavioral diet increased connectivity in the default mode network.
-This might be indicative for self-referent information processing.
--->
-
-# Code availability
-
-All code is available in the github repository [https://github.com/hsx1/adi2_rsfmri](https://github.com/hsx1/adi2_rsfmri).
-The code includes script for data preparation and analysis of different models.
-
-# Declaration of interest
-# Funding
-# Acknowledgments
-# Author Contributions
-
-Stud conception and design
-Aqcuisiiton of data:
-Analysis andinterpretation of data: 
-Frafting of manuscript: 
-Ciritical revision:
-
-\newpage
 # References
-
-```{r Exploratory plots, eval=FALSE}
-# BMI over time per group
-tspag <-
-  ggplot(final, aes(x = tp, y = BMI)) +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  geom_line(aes(group = subj.ID, color = condition)) +
-  geom_point(aes(color = condition))
-tspag +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "bottom",
-    legend.background = element_rect(colour = "grey")
-  )
-
-# BMI over time per subject
-# group x time interaction for BMI
-tspag <- ggplot(final, aes(
-  x = tp,
-  y = BMI,
-  color = condition,
-  group = condition
-)) +
-  geom_point() +
-
-  stat_summary(fun = mean, geom = "point") +
-  stat_summary(fun = mean, geom = "line")
-tspag  +
-  scale_color_manual(values = c("#00305E", "#C00045")) +
-  theme_bw() +
-  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.position = "top"
-)
-```
-
-```{r expample plots, eval=FALSE}
-# example plots for interaction effect (arbitrary variables for illustration purposes)
-ggplot(final, aes(x=tp, y=BMI, group=condition, color=condition)) +
-  #geom_smooth(aes(group=condition)) +
-  stat_summary(fun = mean, geom = "point", size=3, shape=15) +
-  stat_summary(fun = mean, geom = "line", size=1) +
-  #scale_color_brewer(values = "Set1") +
-  scale_x_discrete(labels=c("bl"= "0", "fu"= "6", "fu2" = "12")) +
-  scale_color_manual(values=c("#00305E","#C00045"),labels = c("intervention", "control")) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10), axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12), legend.text = element_text(size = 10),
-    legend.position = "bottom", legend.background = element_rect(colour = "grey"))
-
-# example plot for change effect (purely descriptive)
-getPalette = colorRampPalette(brewer.pal(9, "Set1"))
-cgnplot <- ggplot(final, aes(x=a,y=logmFD, group=subj.ID)) + 
-  geom_line(aes(x=BMI, color=subj.ID))+ 
-  geom_point(aes(x=BMI, color=subj.ID, shape=tp), size=3, alpha = 0.5)+ 
-  facet_grid(. ~ condition) 
-cgnplot + 
-  labs(shape="Time after intervention [months]",labels=c("0","6","12")) +
-  guides(color = FALSE, alpha=FALSE, size=FALSE) +
-  scale_color_manual(values = getPalette(length(unique(final$subj.ID)))) +
-  ylab("log(mean FD) (mm)") + xlab("BMI [kg/m²]") + xlim(30,55) +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10), axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12), legend.text = element_text(size = 10),
-    legend.position = "bottom", legend.background = element_rect(colour = "grey"))
-
-# example plot for change effect (purely descriptive)
-tspag=ggplot(data=final, aes(y=cgnFD, x=cgnBMI, color=condition))+
-  geom_point(size=2)+
-  #geom_line(aes(color=subj.ID))+ 
-  #geom_smooth(method = "lm", colour = "black", linetype = 1, fill = "mistyrose3" ) +
-  xlab("within-subject BMI change [kg/m²]")+
-  ylab("within-subject log mean FD change [mm]")+ 
-  guides(alpha=FALSE, size=FALSE,
-         color = guide_legend(override.aes = list(size=4))) +
-  facet_grid(. ~ condition) 
-tspag + scale_color_manual(values=c("#00305E","#C00045")) +
-  theme_bw() + theme(axis.text=element_text(colour = "black",size=10),
-              axis.title=element_text(size=14),
-              axis.text.x = element_text(size=12),
-              axis.text.y = element_text(size=12),
-              strip.text = element_text(size=14),
-              legend.position="top",
-              legend.title = element_text(size = 14),
-              legend.text = element_text(size = 14))
-```

### 2020-12-02 intro_rsfmri.Rmd
deleted abbreviations

@@
 Thus, one recently emerging research focus is how brain metabolism and functional connectivity changes in response to surgery, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
-The reward network comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+The reward network comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antINS) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
 These regions also seem to be altered after bariatric surgery [@Lips_2014; @Hogenkamp_2016; @Garcia_2012].
@@
 
-The default mdoe network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
+The default mdoe network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and inferior and lateral parietal cortex (approx.
 BA 39) [@Raichle_2015].
-The right ventral anterior cingulate cortex (VACC), superior frontal cortex (SFC), and inferior temporal cortex (ITC) show weaker but also reliable association with the PCC over the course of one year.
-Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
+The right ventral anterior cingulate cortex (vACC), superior frontal cortex, and inferior temporal cortex show weaker but also reliable association with the PCC over the course of one year.
+Even broader conceptions count left VACC, middle temporal gyrus, and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
 <!-- study results --> 
@@
 Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
-Consistent with these findings, other studies reported a reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery [@Li_2018].
+Consistent with these findings, other studies reported a reduction of functional connectivity of these and other regions of the default mode network, such as the vmPFC <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery [@Li_2018].
 
@@
 It has been stated that among patients undergoing surgery, change in BMI was<!-- negatively --> correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
-On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011]. ...
+On the other hand, there are also indicators of reduced functional integrity, e.g. for the ACC at resting state compared to lean individuals [@Kullmann_2011]. ...
 

### 2020-12-02 intro_rsfmri.Rmd
edited output parameter

--- 2020-12-02 intro_rsfmri.Rmd
@@
   Day Clinic for Cognitive Neurology, University Clinic Leipzig
-output:
-  bookdown::pdf_book: 
-    fig_caption: yes
-    always_allow_html: true
-    number_sections: yes
-    toc: yes
-bibliography: "bibliography.bib"
+output: word_document
 editor_options:
   chunk_output_type: console
+bibliography: bibliography.bib
 ---
-```{r "To do", include=FALSE}
-
-# [ ] Evaluate SPM model comparison (more activation in SPM..)
-# [ ] Introduction
-# [ ] Discussion
-# [ ] hervorheben, was von Präregistrierung abweicht
-# [ ] brain images mit nilearn plotting tools einbinden
-# [x] Figure zu weight loss
-# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-
-```
 

### 2020-12-03 manuscript/draft_rsfmri.Rmd
Moved files"

--- 2020-12-01 draft_rsfmri.Rmd
@@
 ```
-\newpage
 # Abstract
 
-\newpage
 # Introduction
-
-<!-- obesity and interventions -->
-Over the past decades, the global burden of obesity has increasingly been confronted as a matter of urgency by research.
-_something about comorbidities_
-Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018]. 
-Luckily, dietary interventions such as caloric restriction and related metabolic improvements, and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018; @Veronese_2017].
-In some cases behavioral intervention fail to yield the desired success regarding weight loss. There are non-behavioral alternatives at hand.
-<!-- bars most invasive (brain metabolism), last resort -->
-Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes [@].
-Roux-en-Y gastric bypass (RYGB) is the most common procedure as it not only restricts the food intake by reducing stomach volume but also affects the absorption of macronutrients and the hormonal profile [@].
-<!-- ref for "common", ref for procedure -->
-Other widely used procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption. <!-- mehr zu Veränderung in Gasto-intestinaler Physiologie -->
-
-<!-- Neural mechanisms/concurrent plasticity? -->
-Precise mechanisms are yet to be elucidated.
-As a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed.
-This will inform inferences about potential mechanisms and either support interventions proven to be effective or ideally allow imitating them will less invasive alternatives.
-
-Recently, research focuses have slowly shifted to investigate brain metabolism structural and functional changes after surgery.
-<!-- severe nutritional stress [@Crujeiras_2017] -->
-
-<!-- Reward network -->
-Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
-<!-- mechanism reward and appetite, overating -->
-Besides malabsorption and restricted food ingestion, patients report post-surgical changes in food preferences and eating behavior.
-Those psychophysiological and behavioral effects may be related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
-Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017, @Nudel_2019].
-<!-- results in lean -->
-The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
-<!-- results in obese longitudinal -->
-Similarly, several studies have found differences in resting-state activity in reward regions depending on weight status.
-
-Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
-These differences seemed to normalize, as trends reversed when looking at the resting state of obese individuals after compared to before the surgery [@Duan_2020].
-
-On the other hand, BMI prior to the surgery was associated with changes in connectivity [@Duan_2020].
-
-s concerning such processes following a surgical intervention are inconclusive.
-
-<!-- important differences in the response to visual food-related cues between thin and reduced-obese individuals that place the latter group at an increased risk of  -->
-
-<!-- Though there exists strong evidence for important differences in the ..
-between lean or normal-weight individuals and obese individuals, few studies have investigated the ...
-and even fewer compared reduced-obese individuals reduced-obese individuals -->
-<!-- [@Garcia_Garcia_2012] -->
-
-<!-- Default mode network --> 
-Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
-Obesity may be cause or effect of a decompensation of the psychophysiological system that ensures homeostasis.
-The default mode network is commonly associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
-<!-- parietal activation correlated with appetite measures after an overnight fast [@Tregellas_2011].
---> It is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
-BA 39) [@Raichle_2015].
-The right ventral anterior cingulate cortex (VACC), superior frontal cortex (SFC), and inferior temporal cortex (ITC) show weaker but also reliable association with the PCC over the course of one year.
-Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
-<!-- difference lean obese --> 
-Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].
-Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
-A longitudinal study that confirmed the stronger connectivity (of PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG).
-However, default mode connectivity in these regions seemed weakened over the course of one year.
-Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable 1 year after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
-This is consistent with other findings of reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC [@Li_2018] <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery.
-
-<!-- argument: FC ~ BMI --> The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
-Among patients undergoing surgery, change in BMI was negatively correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
-
-On the other hand, connectivity also served as a predictor of weight loss after surgery.
-
-On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011].
-
-<!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
-Yet, to our knowledge there is no quasi-experimental study, comparing a cohort of patients to a cohort with similar characteristics but unexposed to the intervention, i.e.
-bariatric surgery, over an extended time period as would be demanded to draw more substantiated inference [@George_2016].
-<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery by comparison with a contemporaneously measured waiting-control group.
-Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
-In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
-Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
-
-<!-- group effect because direct effect of intervention surgery -> rsfmri (cares lessfreq, less preoccupation with hunger, less/more reward sensitivity) vs.
-sugery -> BMI -> rsfmri (neural signalling, fatty tissue) -->
 

### 2020-12-03 manuscript/intro_rsfmri.Rmd
corrected mistakes

@@
 These regions also seem to be altered after bariatric surgery [@Lips_2014; @Hogenkamp_2016; @Garcia_2012].
-<!-- Zitationen checken -->
+<!-- mehr zu Zitationen -->
 <!-- Normalization of connectivity --> Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
@@
 <!-- Default mode network --> 
-Because multiple resting state networks are relevant to obesity and associated with abnormal neurocircuitry interactions [@Ding_2020], extensive research has been devoted to investigating the the relationship between the control network and the reward network.
+Because multiple resting state networks are relevant to obesity and associated with abnormal neurocircuitry interactions [@Ding_2020], extensive research has been devoted to investigating the relationship between the control network and the reward network.
 However, the dynamics within the default network may also be of importance because it is associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
@@
 
-The default mdoe network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and inferior and lateral parietal cortex (approx.
+The default mode network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and inferior and lateral parietal cortex (approx.
 BA 39) [@Raichle_2015].
 The right ventral anterior cingulate cortex (vACC), superior frontal cortex, and inferior temporal cortex show weaker but also reliable association with the PCC over the course of one year.
-Even broader conceptions count left VACC, middle temporal gyrus, and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
+Even broader conceptions count left vACC, middle temporal gyrus, and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
 <!-- study results --> 
@@
 
-<!--  -->
-
 <!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
@@
 Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
-In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
+In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network.
 Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].

### 2020-12-03 manuscript/draft_rsfmri.Rmd
fix path reference

@@
 ---
+
 ```{r "To do", include=FALSE}
@@
 opts_knit$set(root.dir = getwd())
-
 ```
@@
 ```{r Load data, include=FALSE}
-source("get_subject_ID_group_tp_forSwE.R")
-final = save_txt_for_swe(group = "all", tp = "all")
+source("code/get_subject_ID_group_tp_forSwE.R")
+final = txt_for_swe(group = "all", tp = "all")
 
@@
 ```{r baseline characteristics, include=FALSE}
-dfBL <- save_txt_for_swe(group = "all", tp = "BL")
+dfBL <- txt_for_swe(group = "all", tp = "BL")
 

### 2020-12-03 code/get_subject_ID_group_tp_forSwE.R

@@ -9,7 +9,7 @@ txt_for_swe <- function(group = "all", tp = "all"){
   original_wd <- getwd()
   
   # import absolute path
-  abs_path <- read.csv("../code/abs_path.csv", header=FALSE, stringsAsFactors=FALSE)
+  abs_path <- read.csv("abs_path.csv", header=FALSE, stringsAsFactors=FALSE)
   
   # set working directory
   swe_path <- "Analysis/Project2_resting_state/seed-based/Second_level /SwE_files"

### 2020-12-03 draft_rsfmri.Rmd
modified methods section

@@
 ---
-title: "Effects of bariatric surger and weight loss on resting-state functional connectivity"
+title: "Effects of bariatric surgery and weight loss on resting state functional connectivity"
 author:
@@
 
-MRI images were retrieved with a 12-channel head coil of a 3 Tesla Siemens Trio system.
-<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, TR = 1900 ms, TE = 2.52 ms, a = 9, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
-<!-- rsfMRI --> Resting state gradient echo were acquired in 34 slices at an 90$^\circ$ flip angel with a repetition time (TR) of 2.3 s and echo time (TE) of 30 ms; image matrix = 64 x 64 voxel with in-plane resolution of 3 mm x 3 mm and a slice thickness of 4 mm, 150 volumes<!-- dimension = [60;72;60] ???
--->.
-The total acquisition time was 5:45 minutes.
-Field map or phase-encode reversed images were not acquired.
+MRI images were retrieved with a 12-channel head coil of a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
+<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
+<!-- rsfMRI --> Resting state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms.
+The image matrix was 64 x 64 with an in-plane resolution of 3 mm x 3 mm and 34 slices with a slice thickness of 4 mm.
+150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes.
+Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneties.
 Participants were instructed to close their eyes but to remain awake during scanning.
@@
 
-Data preprocessing was conducted with FSL and Freesurfer using a nipype (version 14.0.2) workflow implemented in python 2.7. 
-The preprocessing workflow included anatomical, functional and diffusion-weighted imaging preprocessing and wrapped modules from FSL version 6.0.1<!-- ?!
--->, Freesurfer version 6.0.0p1, ANTS <!-- version?!
-and so on-->.
-Details have been published on github [https//:.???.com]()
-<!-- anatomical preprocessing write something about anatomical preprocessing (Freesurfer etc.) here --> 
-The T1-weighted image was segmented to remove non-brain tissue and together with the functional images normalized to a default brain template from the Monteal Neurological Institute (MNI) space.
-Functional images were smoothed with an $0$ mm FWHM Gaussian kernel.
+Imaging data was conducted using AFNI 19.1.05, ANTS '2.3.1', FSL '6.0.1' and FreeSurfer '6.0.0p1', wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15.
+The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel.
+The workflow can be found on [github](https://github.com/fBeyer89/ADI_preproc/).
+### Anatomical preprocessing
+Within the module for anatomical preprocessing, the T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross].
+Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long].
+Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rs preprocessing.
+The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@antsref].
+The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
+In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.
 
-<!-- minimal functional preprocessing --> Preprocessing pipeline included removal of first four volumes and a transformation of the blood oxygen level dependent (BOLD) time series into the subject's anatomical space.
-This transform combines parameters from motion correction (FSL's MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL's FUGUE) and coregistration to the subject's anatomical template space (using Freesurfer's bbregister and the transform from individual to longitudinal template space in Freesurfer's longitudinal processing stream).
+##Functional preprocessing
+### Minimal functional preprocessing 
+The function preprocessing module included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
+In more detail, the transforms derived from the latter three steps were combined into one transformation.
+For further analysis and ICA-AROMA processing, the data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
+This part of the analysis was described like "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated." in the preregistration.
+The resulting fMRI data is termed "minimally preprocessed" in the following.
+### ICA-AROMA preprocessing
+We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion- FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
+The minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+Finally, the timeseries of the  nuisance components were regressed from the overall timeseries.
+As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result.
+We used CompCor to estimate 5 variance components from a combined WM and CSF mask, which was further eroded using `fslmaths -nan -thr 0.99 -ero -bin` [@Behzadi_2007;@Muschelli_2014].
+After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
+This analysis was described in the preregistration as "Then we will employ state-of-the art nuisance regression (ICA-AROMA and CompCor regression of white matter and CSF signal to reduce the effect of physiological noise and head motion on the connectivity estimates."
+### Global Signal regression
+[@Ciric_2017] showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+Yet, it introduces spatial dependency and spurious correlations.
+Previously, we reported on head motion differences between groups over time in this study [@Beyer_2020], and thus it was very important to minimize the motion-FC correlation.
+The global signal was derived from the average of all voxels in the brain mask.
+Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA fMRI data and high-pass filtered the resulting file as above.
+In the preregistration we stated " As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but introduces spatial dependency and spurious correlations, we perform GSR after ICA-AROMA in a separate sensitivity analysis."
 
-<!-- Aroma + CC/ + GSR --> Volumes were further preprocessed with ICA-AROMA and compCor (CC) and, optionally, with a subsequent global signal regression (GSR) [@Ciric_2017; @Parkes_2018] to reduce the effect of physiological noise and head motion on the connectivity estimates.
-As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but is known to introduces spatial dependency and spurious correlations, we performed GSR after ICA-AROMA in a separate sensitivity analysis.
-
-### Seed region of interest (ROI)
-<!-- Regions of interest (definition and transformation into freesurfer long space) --> Seed regions were ...
-mm spheres in the precuneus (MNI: x,y,z = x,x,x) associated with the default mode network, and the Nucleus Accumbens (MNI: x,y,z = x,x,x) by averaging the time series across voxels.
-
-### Connectivity maps
-<!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint (more details)) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC).
-Subsequently, r-to-z transform was applied to the correlation to obtain normally distributed functional connectivity (using nilearn, nipype and python).
-Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).
+### Functional connectivity
+To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed ROIs, respectively.
+Based on FreeSurfer's `aseg.mgz` and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
+Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
+We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (min, aroma+cc, aroma+gsr).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived during anatomical preprocessing.
 
 ### Quality Assessment
+<!-- from Prereg -->
+
+We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
+
+Whole brain FC maps will be calculated using Pearson’s correlation between average seed time series and whole brain timeseries on the subject level (using nilearn).
+After transforming from r-to-Z, we will transform the images into Freesurfer longitudinal space, using /mri/transforms/_to_.lta).
+Then, will calculate difference maps between BL-FU1 (confirmatory hypothesis), BL-FU2, FU1-FU2 (exploratory hypotheses).
+These images will be transformed into MNI space (using ANTS non-linear registration from the longitudinal template to MNI space) for statistical analysis.
+Aggregated FC values will be calculated using Pearson’s correlation of average seed time series from pre-defined in the longitudinal template space.
+Aggregated values will be extracted for the reward and the default mode network.
+
+Quality control:
+- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
+- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
+-  We will first perfom the analysis with all participants who were correctly preprocessed.
+In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
+
+We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
+
+-	Anthropometry: body mass index, % body fat from bioelectrical impedance measurement 
+-	Grouping variable is intervention versus control group
 

### 2020-12-03 manuscript/draft_rsfmri.Rmd
fix working directory

@@
 
-```{r "environment setup", include=FALSE}
+```{r "load packages", include=FALSE}
 
@@
 #library(haven)
+library(knitr) # version 1.30
+```
 
-library(knitr) # version 1.30
+```{r "knit setup"}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
 options(knitr.table.format = "latex") 
-# put the path to this file here (working directory for Rmarkdown):
-opts_knit$set(root.dir = getwd())
+# set working directory of knit (RMarkdown) to directory of Rproj
+opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+setwd(knitr::opts_knit$get("root.dir"))
 ```
-# Abstract
-
-# Introduction
 

### 2020-12-03 manuscript/draft_rsfmri.Rmd
format MNI space coordinates

--- 2020-12-03 manuscript/draft_rsfmri.Rmd
@@
 <!-- Regions of interest (definition and transformation into freesurfer long space) --> Seed regions were ...
-mm spheres in the precuneus (MNI: x,y,z = x,x,x) associated with the default mode network, and the Nucleus Accumbens (MNI: x,y,z = x,x,x) by averaging the time series across voxels.
+mm spheres in the precuneus ((MNI)X,Y,Z = x,x,x) associated with the default mode network, and the Nucleus Accumbens (x,x,x) by averaging the time series across voxels.
 
@@
 Again, parametric estimation and non-parametric estimation differed in regard to the group-time-interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{FDRcorr} = 0.998$)<!-- also FWE-corr n.s -->.
+While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->.
 

### 2020-12-04 draft_rsfmri.Rmd
modified introduction

@@
 # Introduction
-<!--Suggested Order of Introduction
-1.
-Obesity is a major health concern.
-[...
-some more details about metabolic and mortality risk, especially related to morbid obesity ...] More and more recognized, midlie obesity also impacts later cognitive decline and dementia risk.
-Obesity has been associated with differences in brain structure and function.
-Regarding structural alterations, longitudinal studies have indicated that higher body mass index is related to increased cortial thinning in midlife (path through life study), potentially mediated by inflammation or glucose dysregulation...
-Similarly, white matter microstructure and may be related to adverse metabolic effects (lampe)
-Differences in brain function in obesity have been reported in the reward network during task-based fMRI, highlighting the role of reward processing in obesity development and maintainance.
-Resting state fMRI studies have also shown differences in connectivity in diverse brain networks, including reward and salience network, visual networks and the default mode network.
-In part, these differences have been related to metabolic and behavioral differences, yet there may be an important confounding due to head motion.
-In this study, we aim to investigate the impact of bariatric surery on..
--->
 
@@
 Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (min, aroma+cc, aroma+gsr).
-Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived during anatomical preprocessing.
+We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
 
@@
 <!-- from Prereg --> 
-
+#### Anatomical Quality control:
+- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
+- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
+-  We will first perfom the analysis with all participants who were correctly preprocessed.
+In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
 We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
@@
 Aggregated values will be extracted for the reward and the default mode network.
-
-Quality control:
-- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
-- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
--  We will first perfom the analysis with all participants who were correctly preprocessed.
-In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
 

### 2020-12-07 manuscript/draft_rsfmri.Rmd
added Lei_2019, Muschelli_2014 and Behzadi_2007

@@
 ---
-title: "Effects of bariatric surger and weight loss on resting-state functional connectivity"
+title: "Effects of bariatric surgery and weight loss on resting state functional connectivity"
 author:
-- xxx [1]
-- xxx [1][2]
-- xxx [1][2][3]
-- xxx [1][2]
+- Hannah Heinrichs* [1]
+- Frauke Beyer* [1][2]
+- Kerstin Prehn
+- Agnes Floel
+- Arno Villringer [1][2][3]
+- A. Veronica Witte [1][2][3]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
@@
 ```
+\newpage
+# Abstract
+
+\newpage
+# Introduction
+<!--Suggested Order of Introduction
+1.
+Obesity is a major health concern.
+[...
+some more details about metabolic and mortality risk, especially related to morbid obesity ...] More and more recognized, midlie obesity also impacts later cognitive decline and dementia risk.
+Obesity has been associated with differences in brain structure and function.
+Regarding structural alterations, longitudinal studies have indicated that higher body mass index is related to increased cortial thinning in midlife (path through life study), potentially mediated by inflammation or glucose dysregulation...
+Similarly, white matter microstructure and may be related to adverse metabolic effects (lampe)
+Differences in brain function in obesity have been reported in the reward network during task-based fMRI, highlighting the role of reward processing in obesity development and maintainance.
+Resting state fMRI studies have also shown differences in connectivity in diverse brain networks, including reward and salience network, visual networks and the default mode network.
+In part, these differences have been related to metabolic and behavioral differences, yet there may be an important confounding due to head motion.
+In this study, we aim to investigate the impact of bariatric surery on..
+-->
+
+<!-- obesity and interventions -->
+Over the past decades, the global burden of obesity has increasingly been confronted as a matter of urgency by research.
+_something about comorbidities_
+Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018].
+Luckily, dietary interventions such as caloric restriction and related metabolic improvements, and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018; @Veronese_2017].
+In some cases behavioral intervention fail to yield the desired success regarding weight loss. There are non-behavioral alternatives at hand.
+<!-- bars most invasive (brain metabolism), last resort -->
+Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes [@].
+Roux-en-Y gastric bypass (RYGB) is the most common procedure as it not only restricts the food intake by reducing stomach volume but also affects the absorption of macronutrients and the hormonal profile [@].
+<!-- ref for "common", ref for procedure -->
+Other widely used procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption. <!-- mehr zu Veränderung in Gasto-intestinaler Physiologie -->
+
+<!-- Neural mechanisms/concurrent plasticity? -->
+Precise mechanisms are yet to be elucidated.
+As a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed.
+This will inform inferences about potential mechanisms and either support interventions proven to be effective or ideally allow imitating them will less invasive alternatives.
+
+Recently, research focuses have slowly shifted to investigate brain metabolism structural and functional changes after surgery.
+<!-- severe nutritional stress [@Crujeiras_2017] -->
+
+<!-- Reward network -->
+Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
+<!-- mechanism reward and appetite, overating -->
+Besides malabsorption and restricted food ingestion, patients report post-surgical changes in food preferences and eating behavior.
+Those psychophysiological and behavioral effects may be related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
+Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017, @Nudel_2019].
+<!-- results in lean -->
+The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
+<!-- results in obese longitudinal -->
+Similarly, several studies have found differences in resting-state activity in reward regions depending on weight status.
+
+Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
+These differences seemed to normalize, as trends reversed when looking at the resting state of obese individuals after compared to before the surgery [@Duan_2020].
+
+On the other hand, BMI prior to the surgery was associated with changes in connectivity [@Duan_2020].
+
+s concerning such processes following a surgical intervention are inconclusive.
+
+<!-- important differences in the response to visual food-related cues between thin and reduced-obese individuals that place the latter group at an increased risk of  -->
+
+<!-- Though there exists strong evidence for important differences in the ..
+between lean or normal-weight individuals and obese individuals, few studies have investigated the ...
+and even fewer compared reduced-obese individuals reduced-obese individuals -->
+<!-- [@Garcia_Garcia_2012] -->
+
+<!-- Default mode network -->
+Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
+Obesity may be cause or effect of a decompensation of the psychophysiological system that ensures homeostasis.
+The default mode network is commonly associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
+<!-- parietal activation correlated with appetite measures after an overnight fast [@Tregellas_2011].
+--> It is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
+BA 39) [@Raichle_2015].
+The right ventral anterior cingulate cortex (VACC), superior frontal cortex (SFC), and inferior temporal cortex (ITC) show weaker but also reliable association with the PCC over the course of one year.
+Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
+<!-- difference lean obese -->
+Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].
+Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
+A longitudinal study that confirmed the stronger connectivity (of PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG).
+However, default mode connectivity in these regions seemed weakened over the course of one year.
+Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable 1 year after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
+This is consistent with other findings of reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC [@Li_2018] <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery.
+
+The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
+Among patients undergoing surgery, change in BMI was negatively correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
+
+On the other hand, connectivity also served as a predictor of weight loss after surgery.
+
+On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011].
+
+<!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
+Yet, to our knowledge there is no quasi-experimental study, comparing a cohort of patients to a cohort with similar characteristics but unexposed to the intervention, i.e.
+bariatric surgery, over an extended time period as would be demanded to draw more substantiated inference [@George_2016].
+<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery by comparison with a contemporaneously measured waiting-control group.
+Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
+In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
+Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
+
+<!-- group effect because direct effect of intervention surgery -> rsfmri (cares lessfreq, less preoccupation with hunger, less/more reward sensitivity) vs.
+sugery -> BMI -> rsfmri (neural signalling, fatty tissue) -->
 
@@
 
-Data preprocessing was conducted with FSL and Freesurfer using a nipype (version 14.0.2) workflow implemented in python 2.7. 
-The preprocessing workflow included anatomical, functional and diffusion-weighted imaging preprocessing and wrapped modules from FSL version 6.0.1<!-- ?!
--->, Freesurfer version 6.0.0p1, ANTS <!-- version?!
-and so on-->.
-Details have been published on github [https//:.???.com]()
-<!-- anatomical preprocessing write something about anatomical preprocessing (Freesurfer etc.) here --> 
-The T1-weighted image was segmented to remove non-brain tissue and together with the functional images normalized to a default brain template from the Monteal Neurological Institute (MNI) space.
-Functional images were smoothed with an $0$ mm FWHM Gaussian kernel.
+Imaging data was conducted using AFNI 19.1.05, ANTS '2.3.1', FSL '6.0.1' and FreeSurfer '6.0.0p1', wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15.
+The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel.
+The workflow can be found on [github](https://github.com/fBeyer89/ADI_preproc/).
+### Anatomical preprocessing
+Within the module for anatomical preprocessing, the T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross].
+Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long].
+Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rs preprocessing.
+The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@antsref].
+The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
+In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.
 
-<!-- minimal functional preprocessing --> Preprocessing pipeline included removal of first four volumes and a transformation of the blood oxygen level dependent (BOLD) time series into the subject's anatomical space.
-This transform combines parameters from motion correction (FSL's MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL's FUGUE) and coregistration to the subject's anatomical template space (using Freesurfer's bbregister and the transform from individual to longitudinal template space in Freesurfer's longitudinal processing stream).
+##Functional preprocessing
+### Minimal functional preprocessing
+The function preprocessing module included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
+In more detail, the transforms derived from the latter three steps were combined into one transformation.
+For further analysis and ICA-AROMA processing, the data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
+This part of the analysis was described like "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated." in the preregistration.
+The resulting fMRI data is termed "minimally preprocessed" in the following.
+### ICA-AROMA preprocessing
+We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion- FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
+The minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+Finally, the timeseries of the  nuisance components were regressed from the overall timeseries.
+As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result.
+We used CompCor to estimate 5 variance components from a combined WM and CSF mask, which was further eroded using `fslmaths -nan -thr 0.99 -ero -bin` [@Behzadi_2007;@Muschelli_2014].
+After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
+This analysis was described in the preregistration as "Then we will employ state-of-the art nuisance regression (ICA-AROMA and CompCor regression of white matter and CSF signal to reduce the effect of physiological noise and head motion on the connectivity estimates."
+### Global Signal regression
+[@Ciric_2017] showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+Yet, it introduces spatial dependency and spurious correlations.
+Previously, we reported on head motion differences between groups over time in this study [@Beyer_2020], and thus it was very important to minimize the motion-FC correlation.
+The global signal was derived from the average of all voxels in the brain mask.
+Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA fMRI data and high-pass filtered the resulting file as above.
+In the preregistration we stated " As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but introduces spatial dependency and spurious correlations, we perform GSR after ICA-AROMA in a separate sensitivity analysis."
 
-<!-- Aroma + CC/ + GSR --> Volumes were further preprocessed with ICA-AROMA and compCor (CC) and, optionally, with a subsequent global signal regression (GSR) [@Ciric_2017; @Parkes_2018] to reduce the effect of physiological noise and head motion on the connectivity estimates.
-As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but is known to introduces spatial dependency and spurious correlations, we performed GSR after ICA-AROMA in a separate sensitivity analysis.
-
-### Seed region of interest (ROI)
-<!-- Regions of interest (definition and transformation into freesurfer long space) --> Seed regions were ...
-mm spheres in the precuneus ((MNI)X,Y,Z = x,x,x) associated with the default mode network, and the Nucleus Accumbens (x,x,x) by averaging the time series across voxels.
-
-### Connectivity maps
-<!-- coregistration + normalization --> Voxel-wise seed-based connectivity maps in the individual subject's space (which was coregistered to the Freesurfer longitudinally preprocessed timepoint (more details)) were calculated using Pearson's correlation between average seed time series from pre-defined ROIs within the reward network (Nucleus accumbens, NAcc) and the DMN (precuneus, PCC).
-Subsequently, r-to-z transform was applied to the correlation to obtain normally distributed functional connectivity (using nilearn, nipype and python).
-Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).
+### Functional connectivity
+To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed ROIs, respectively.
+Based on FreeSurfer's `aseg.mgz` and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
+Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
+We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (min, aroma+cc, aroma+gsr).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived during anatomical preprocessing.
 
 ### Quality Assessment
+<!-- from Prereg -->
+
+We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
+
+Whole brain FC maps will be calculated using Pearson’s correlation between average seed time series and whole brain timeseries on the subject level (using nilearn).
+After transforming from r-to-Z, we will transform the images into Freesurfer longitudinal space, using /mri/transforms/_to_.lta).
+Then, will calculate difference maps between BL-FU1 (confirmatory hypothesis), BL-FU2, FU1-FU2 (exploratory hypotheses).
+These images will be transformed into MNI space (using ANTS non-linear registration from the longitudinal template to MNI space) for statistical analysis.
+Aggregated FC values will be calculated using Pearson’s correlation of average seed time series from pre-defined in the longitudinal template space.
+Aggregated values will be extracted for the reward and the default mode network.
+
+Quality control:
+- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
+- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
+-  We will first perfom the analysis with all participants who were correctly preprocessed.
+In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
+
+We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
+
+-	Anthropometry: body mass index, % body fat from bioelectrical impedance measurement
+-	Grouping variable is intervention versus control group
 

### 2020-12-07 manuscript/draft_rsfmri.Rmd

@@
 Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (min, aroma+cc, aroma+gsr).
-Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived during anatomical preprocessing.
+We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
 
 ### Quality Assessment
-<!-- from Prereg -->
-
+#### Anatomical Quality control
+FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
+We excluded 5 datasets from 3 participants rated "Exclude" because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration .
+<!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
+Participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
+caused Freesurfer to fail or perform very poorly, i.e.
+rated "Failed" according to ) were excluded because a subject's motion is highly similar across scans and we thus suspect the fMRI data to also suffer from motion artifacts in these subjects.
+- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
+- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
+-  We will first perfom the analysis with all participants who were correctly preprocessed.
+In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
 We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.

### 2020-12-07 manuscript/intro_rsfmri.Rmd

@@
 ---
-
-```{r "environment setup", include=FALSE}
+```{r "load packages", include=FALSE}
 
@@
 #library(haven)
+library(knitr) # version 1.30
+```
 
-library(knitr) # version 1.30
+```{r "knit setup"}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
 options(knitr.table.format = "latex") 
-# put the path to this file here (working directory for Rmarkdown):
-opts_knit$set(root.dir = getwd())
+# set working directory of knit (RMarkdown) to directory of Rproj
+opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+setwd(knitr::opts_knit$get("root.dir"))
+```
 
-```
 # Abstract

### 2020-12-07 manuscript/draft_rsfmri.Rmd

@@
 
-Whole brain FC maps will be calculated using Pearson’s correlation between average seed time series and whole brain timeseries on the subject level (using nilearn).
-After transforming from r-to-Z, we will transform the images into Freesurfer longitudinal space, using /mri/transforms/_to_.lta).
-Then, will calculate difference maps between BL-FU1 (confirmatory hypothesis), BL-FU2, FU1-FU2 (exploratory hypotheses).
-These images will be transformed into MNI space (using ANTS non-linear registration from the longitudinal template to MNI space) for statistical analysis.
-Aggregated FC values will be calculated using Pearson’s correlation of average seed time series from pre-defined in the longitudinal template space.
-Aggregated values will be extracted for the reward and the default mode network.
+Quality control:
+- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
+- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
+-  We will first perfom the analysis with all participants who were correctly preprocessed.
+In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
 

### 2020-12-08 manuscript/draft_rsfmri.Rmd
changed directory to txt files

--- 2020-12-07 manuscript/draft_rsfmri.Rmd
@@
 ```{r Load data, include=FALSE}
-source("code/get_subject_ID_group_tp_forSwE.R")
-final = txt_for_swe(group = "all", tp = "all")
+source("code/create_sample_df.R")
+final = create_sample_df(group = "all", tp = "all")
 
@@
 ```{r baseline characteristics, include=FALSE}
-dfBL <- txt_for_swe(group = "all", tp = "BL")
+dfBL <- create_sample_df(group = "all", tp = "BL")
 

### 2020-12-08 code/RunModelBMI.m

@@ -52,11 +52,11 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) %strcmp(param.ACTION,'display')
 end
 
 if strcmp(param.MODEL,'bmiIG')
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'IG');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'IG/total');
 elseif strcmp(param.MODEL,'bmi2tp')
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'BLFU');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'both/BLFU');
 else
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'total');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'both/total');
 end
 
 nrun = length(param.ROI_PREP);

### 2020-12-08 code/RunModelFD.m

@@ -55,9 +55,9 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) %strcmp(param.ACTION,'display')
 end
 
 if strcmp(param.MODEL,'fdIG')
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'IG');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'IG/total');
 else
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'total');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'both/total');
 end
 
 nrun = length(param.ROI_PREP);

### 2020-12-08 code/RunModelGroupTime.m

@@ -86,9 +86,9 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) % strcmp(param.ACTION, 'display')
 end
 
 if strcmp(param.MODEL,'grouptime2tp')
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'BLFU');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'both/BLFU');
 else
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'total');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'both/total');
 end
 
 nrun = length(param.ROI_PREP);

### 2020-12-08 code/get_subject_ID_group_tp_forSwE.R

@@ -9,8 +9,9 @@ get_txt_for_swe <- function(group = "all", tp = "all"){
   # import absolute path
   abs_path=read.csv("abs_path.csv", header=FALSE, stringsAsFactors=FALSE)
   # set working directory
-  parentdir=file.path(getwd(), "../SwE_files/", fsep = .Platform$file.sep)[1]
+  parentdir=file.path(getwd(), "../SwE_files", fsep = .Platform$file.sep)[1]
   setwd(parentdir)
+  parentdir=getwd()
   
   # participants with mri data (txt file with path to .nii file) ---------------
   
@@ -85,47 +86,50 @@ get_txt_for_swe <- function(group = "all", tp = "all"){
   output_dir <- parentdir
   # selection of groups
   if (group == "both") {
-    # total: both groups
+    output_dir <- file.path(parentdir, "both")
+    if (!dir.exists(output_dir)) {dir.create(output_dir)}
+    setwd(output_dir)
   }else if (group == "IG") {
-    output_dir <- file.path(parentdir, "IG_only")
+    output_dir <- file.path(parentdir, "IG")
     if (!dir.exists(output_dir)) {dir.create(output_dir)}
     setwd(output_dir)
     final <- final[final$condition=="IG",]
   }else if (group == "KG") {
-    output_dir <- file.path(parentdir, "KG_only")
+    output_dir <- file.path(parentdir, "KG")
     if (!dir.exists(output_dir)) {dir.create(output_dir)}
     setwd(output_dir)
     final <- final[final$condition=="KG",]
   }
   
   # selection of time points
+  
   if (tp == "all"){
     output_dir <- file.path(output_dir, "total")
     if (!dir.exists(output_dir)) {dir.create(output_dir)}
     setwd(output_dir)
     final <- final
   }else if (tp == "BL"){
-    output_dir <- file.path(output_dir, "onlyBL")
+    output_dir <- file.path(output_dir, "BL")
     if (!dir.exists(output_dir)) {dir.create(output_dir)}
     setwd(output_dir)
     final <- final[final$tp == "bl",]
   }else if (tp == "FU"){
-    output_dir <- file.path(output_dir, "onlyFU")
+    output_dir <- file.path(output_dir, "FU")
     if (!dir.exists(output_dir)) {dir.create(output_dir)}
     setwd(output_dir)
     final <- final[final$tp == "fu",]
[…]

### 2020-12-08 manuscript/draft_rsfmri.Rmd
implemented exlusion criteria: (1) preprocessing failure, (2) extensive head motion; text in Methods

--- 2020-12-08 manuscript/draft_rsfmri.Rmd
@@
 
+# [ ] include exclusion criteria according to "create_sample_df.R"
 # [ ] Evaluate SPM model comparison (more activation in SPM..)
-# [ ] Introduction
 # [ ] Discussion
@@
 # [ ] brain images mit nilearn plotting tools einbinden
-# [x] Figure zu weight loss
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
@@
 
-```{r "knit setup"}
+```{r "knit setup", include=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
@@
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
-```{r test wd}
-print(getwd())
-```
 
@@
 rated "Failed" according to @Klapwijk_2019) were excluded because a subject's motion is highly similar across scans and we thus suspect the fMRI data to also suffer from motion artifacts in these subjects.
-As the expected average head motion is high [@Hodgson_2016], we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample.
+<!-- As the expected average head motion is high [@Hodgson_2016], we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample.
+-->
+Five data points were excluded due to failure of correct preprocessing attributed to T1-artifacts, and three additional data points were excluded due to extensive head motion (greater then $2 SD$ above the average mean FD).
 <!-- MOTION CORRECTION --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].

### 2020-12-08 code/create_sample_df.R

@@ -18,15 +18,31 @@ create_sample_df <- function(group = "all", tp = "all"){
     mri_files[i,"subj.Nr"]=as.numeric(substr(mri_files$subj.ID[i],4,6))
   }
   mri_files$V1 <- NULL
-  
-  
+
   # load info from full sample and merge with path info (fmri only)-------------
   
   full_sample=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv") # CAVE: info file elsewhere !!!
-  full_sample=full_sample[,c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI")]
-  condition=merge(mri_files, full_sample[!is.na(full_sample$condition),], by=c("subj.ID","tp"))
-  df=plyr::join(mri_files,condition)
-  rm(mri_files, full_sample, condition, tmp, i)
+  full_sample=full_sample[,c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI","Exclude")]
+  
+  # Exclusion of participants --------------------------------------------------
+  # exclusion due to problems in preprocessing
+  full_sample$exclude_prep=full_sample$Exclude
+  full_sample$exclude_prep[is.na(full_sample$exclude_prep)]=FALSE
+  mri_sample=merge(mri_files, full_sample[!is.na(full_sample$condition),c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI","exclude_prep")], by=c("subj.ID","tp"))
+  
+  mri_sample[mri_sample$exclude_prep==TRUE,] # ADI063_bl ADI063_fu ADI116_bl ADI116_fu ADI116_fu2
+  cat("Note that",nrow(mri_sample[mri_sample$exclude_prep==TRUE,]),"data points were excluded due to problems in preprocessing.\n")
+  sample_after_1stExcl=mri_sample[mri_sample$exclude_prep==FALSE,]
+  
+  # exclusion if mean FD > average meanFD + 2 SD
+  sample_after_1stExcl$exclude_fd = (sample_after_1stExcl$meanFD > mean(sample_after_1stExcl$meanFD)+ 2*sd(sample_after_1stExcl$meanFD))
+  cat("Note that",nrow(sample_after_1stExcl[sample_after_1stExcl$exclude_fd==TRUE,]),"data points were excluded due to extensive head motion.\n")
+  
+  final_sample=sample_after_1stExcl[sample_after_1stExcl$exclude_fd==FALSE,]
[…]

### 2020-12-08 manuscript/draft_rsfmri.Rmd
implemented post-hoc analysis exclusion criterion 10% of data points with worst FD; ordering of df/ design matrix; "Extra analysis" moved to draft_rsfmri.Rmd

@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
+# [ ] redo analysis with new sample 
+# [ ] repeat analysis with exclusion of 10% of worst data
 
@@
 ```{r expample plots, eval=FALSE}
+
+df_wide <- tidyr::pivot_wider(
+  data = final,
+  id_cols = "subj.ID",
+  names_from = "tp",
+  values_from = c("tp", "BMI", "logmFD")
+)
+
+if (tp == "BLFU") {
+  s = 2
+} else {
+  s = 1
+}
+
+BMI_vector <- c("BMI_fu2", "BMI_fu", "BMI_bl")
+BMI_vector <- c("BMI_bl", "BMI_fu", "BMI_fu2")
+FD_vector <- c("logmFD_fu2", "logmFD_fu", "logmFD_bl")
+df_wide$avgBMI <- apply(df_wide[, BMI_vector[s:3]],
+                        MARGIN = 1,
+                        FUN = mean,
+                        na.rm = TRUE)
+df_wide$avgFD <- apply(df_wide[, FD_vector[s:3]],
+                       MARGIN = 1,
+                       FUN = mean,
+                       na.rm = TRUE)
+df_long <- tidyr::pivot_longer(
+  data = df_wide,
+  cols = all_of(order_vector),
+  # bl, fu, fu2 (maintain original order)
+  # CAREFUL: Order according to dataframe df
+  names_to = "tp_tp",
+  values_to = "tp",
+  values_drop_na = TRUE
+)
+all(df_long$tp == final$tp) # check if order correct
+
+head(final[, c("subj.ID", "tp")])
+head(df_long[, c("subj.ID", "tp")])
+
+df_long$cgnBMI <- final$BMI - df_long$avgBMI
+df_long$avgBMIc <-
+  df_long$avgBMI - mean(df_long$avgBMI, na.rm = TRUE)
+df$avgBMIc <- df_long$avgBMIc
+df$cgnBMI <- df_long$cgnBMI
+
+df_long$cgnFD <- final$logmFD - df_long$avgFD
+df_long$avgFDc <-
+  df_long$avgFD - mean(df_long$avgFD, na.rm = TRUE)
+final$avgFDc <- df_long$avgFDc
+final$cgnFD <- df_long$cgnFD
+
 # example plots for interaction effect (arbitrary variables for illustration purposes)

### 2020-12-08 code/RunModelBMI.m

@@ -51,6 +51,12 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) %strcmp(param.ACTION,'display')
     pause(7)
 end
 
+% if param.EXCLFD==true
+%     param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
+% else
+%     param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
+% end
+
 if strcmp(param.MODEL,'bmiIG')
     param.INFO_DIR = fullfile(param.INFO_DIR, 'IG/total');
 elseif strcmp(param.MODEL,'bmi2tp')
@@ -206,7 +212,7 @@ else
     % .. Small sample adjustments for WB resampling (4 = type C2)
     smodel.WB.WB_yes.WB_ss = 4;
     % .. Number of bootstraps
-    smodel.WB.WB_yes.WB_nB = 999;
+    smodel.WB.WB_yes.WB_nB = 1000;
     % .. Type of SwE (0 = U-SwE (recommended))
     smodel.WB.WB_yes.WB_SwE = 0;
     % ... T or F contrast (CAVE: only one contrast at a time)
@@ -268,6 +274,12 @@ else
     parent_folder = 'BMI_total';
 end
 
+if param.EXCLFD==true
+    excl = 'ExclFD';
+else
+    excl = 'noExclFD';
+end
+
 if strcmp(param.MASK, 'gm')
     mask_def = 'gm';
 else
@@ -293,7 +305,7 @@ else % Parametric Estimation
 end
 
 % Create folder
-out_folder = fullfile(param.OUT_DIR, parent_folder, mask_def, param.ROI_PREP{crun}, model_name);
+out_folder = fullfile(param.OUT_DIR, parent_folder, excl, mask_def, param.ROI_PREP{crun}, model_name);
 if ~exist(out_folder, 'dir')
     exist_already = false;
     mkdir(out_folder)

### 2020-12-08 code/RunModelFD.m

@@ -54,6 +54,12 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) %strcmp(param.ACTION,'display')
     pause(7)
 end
 
+if param.EXCLFD==true
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
+else
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
+end
+
 if strcmp(param.MODEL,'fdIG')
     param.INFO_DIR = fullfile(param.INFO_DIR, 'IG/total');
 else
@@ -202,7 +208,7 @@ else
     % .. Small sample adjustments for WB resampling (4 = type C2)
     smodel.WB.WB_yes.WB_ss = 4;
     % .. Number of bootstraps
-    smodel.WB.WB_yes.WB_nB = 999;
+    smodel.WB.WB_yes.WB_nB = 1000;
     % .. Type of SwE (0 = U-SwE (recommended))
     smodel.WB.WB_yes.WB_SwE = 0;
     % ... T or F contrast (CAVE: only one contrast at a time)
@@ -266,6 +272,12 @@ else
     parent_folder = 'FD_total';
 end
 
+if param.EXCLFD==true
+    excl = 'ExclFD';
+else
+    excl = 'noExclFD';
+end
+
 if strcmp(param.MASK, 'gm')
     mask_def = 'gm';
 else
@@ -291,7 +303,7 @@ else % Parametric Estimation
 end
 
 % Create folder
-out_folder = fullfile(param.OUT_DIR, parent_folder, mask_def, param.ROI_PREP{crun}, model_name);
+out_folder = fullfile(param.OUT_DIR, parent_folder, excl, mask_def, param.ROI_PREP{crun}, model_name);
 if ~exist(out_folder, 'dir')
     exist_already = false;
     mkdir(out_folder)

### 2020-12-08 code/RunModelGroupTime.m

@@ -85,6 +85,12 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) % strcmp(param.ACTION, 'display')
     pause(7)
 end
 
+if param.EXCLFD==true
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
+else
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
+end
+
 if strcmp(param.MODEL,'grouptime2tp')
     param.INFO_DIR = fullfile(param.INFO_DIR, 'both/BLFU');
 else
@@ -239,7 +245,7 @@ else
     % .. Small sample adjustments for WB resampling (4 = type C2)
     smodel.WB.WB_yes.WB_ss = 4;
     % .. Number of bootstraps
-    smodel.WB.WB_yes.WB_nB = 999;
+    smodel.WB.WB_yes.WB_nB = 1000;
     % .. Type of SwE (0 = U-SwE (recommended))
     smodel.WB.WB_yes.WB_SwE = 0;
     % ... T or F contrast (CAVE: only one contrast at a time)
@@ -306,6 +312,11 @@ else
     parent_folder = 'GroupTime_total';
 end
 
+if param.EXCLFD==true
+    excl = 'ExclFD';
+else
+    excl = 'noExclFD';
+end
 
 if strcmp(param.MASK, 'gm')
     mask_def = 'gm';
@@ -336,7 +347,7 @@ else % Parametric Estimation
 end
 
 % Create folder
-out_folder = fullfile(param.OUT_DIR, parent_folder, mask_def, param.ROI_PREP{crun}, model_name);
+out_folder = fullfile(param.OUT_DIR, parent_folder, excl, mask_def, param.ROI_PREP{crun}, model_name);
 if ~exist(out_folder, 'dir')
     exist_already = false;
     mkdir(out_folder)

### 2020-12-08 code/get_subject_ID_group_tp_forSwE.R

@@ -1,9 +1,13 @@
 library(dplyr) # version 1.0.2
 library(tidyr) # version 1.1.2
 
-get_txt_for_swe <- function(group = "all", tp = "all"){
+get_txt_for_swe <- function(group = "both", tp = "all",exclFD = FALSE){
   # group = IG/KG/both
   # tp = BL/FU/FU2/BLFU/FUFU2/all
+  
+  # ----------------------------------------------------------------------------
+  # primary definitions
+  # ----------------------------------------------------------------------------
   
   original_wd=getwd()
   # import absolute path
@@ -12,6 +16,39 @@ get_txt_for_swe <- function(group = "all", tp = "all"){
   parentdir=file.path(getwd(), "../SwE_files", fsep = .Platform$file.sep)[1]
   setwd(parentdir)
   parentdir=getwd()
+  
+  
+  # ----------------------------------------------------------------------------
+  # second level functions
+  # ----------------------------------------------------------------------------
+
+  apply_excl_criteria <- function(data = full_sample, exclFD = FALSE){
+    # exclusion due to problems in preprocessing
+    full_sample$exclude_prep <- full_sample$Exclude
+    full_sample$exclude_prep[is.na(full_sample$exclude_prep)]=FALSE
+    mri_sample <- merge(mri_files, full_sample[!is.na(full_sample$condition),c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI","exclude_prep")], by=c("subj.ID","tp"))
+    
+    mri_sample[mri_sample$exclude_prep==TRUE,] # ADI063_bl ADI063_fu ADI116_bl ADI116_fu ADI116_fu2
+    cat("Note that",nrow(mri_sample[mri_sample$exclude_prep==TRUE,]),"data points were excluded due to problems in preprocessing.\n")
+    sample_after_1stExcl <- mri_sample[mri_sample$exclude_prep==FALSE,]
+    
+    if (exclFD == TRUE){
+      # exclusion of 10% of the worst mean FD
+      thres10=sort(sample_after_1stExcl$meanFD, decreasing = TRUE)[round(nrow(sample_after_1stExcl)/10)]
[…]

### 2020-12-08 code/run_or_display.m

@@ -87,7 +87,7 @@ param.MASK_B = 'MNI_resampled_brain_mask.nii,1';
 
 % define ROI
 roi_prep = readcell(fullfile(param.INFO_DIR,'ROIs.txt'), 'Delimiter',' ','Whitespace',"'");
-param.ROI_PREP = {roi_prep{[14]}}; % {roi_prep{[4, 6, 12, 14]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z'}
+param.ROI_PREP = {roi_prep{[12, 14]}}; % {roi_prep{[4, 6, 12, 14]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z'}
 
 % Model definition
 % All three models have unique options for covariate definition, the
@@ -95,22 +95,23 @@ param.ROI_PREP = {roi_prep{[14]}}; % {roi_prep{[4, 6, 12, 14]}} or {'Nacc_cc_z',
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
 
 param.MODEL = {'grouptime'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [12];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
+param.COVARIATES = [11];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
 
 % define masking and type of inference
 param.MASK = 'brain';               % 'brain' or 'gm'
-param.WILD_BOOT = true;             % false
-param.INFERENCE_TYPE = {'voxel'};   % {'voxel','cluster','tfce'};
+param.EXCLFD = false;                % false
+param.WILD_BOOT = false;             % false
+param.INFERENCE_TYPE = {'cluster'};   % {'voxel','cluster','tfce'};
 % analysis parameter (estimate or display?)
-param.ONLY_DISPLAY = true;         % false
-param.OVERWRITE = false;            % false
-param.VIEWSEC = 2; % for ONLY_DISPLAY: seconds you want to view the results
+param.ONLY_DISPLAY = false;         % false
+param.OVERWRITE = true;            % false
+param.VIEWSEC = 1; % for ONLY_DISPLAY: seconds you want to view the results
[…]

### 2020-12-09 manuscript/draft_rsfmri.Rmd
bugfix to work with SwE 2.2.1 (development version, date: 12/9/2020); fix coherence with preregistration (Wild bootstrap, clusterwise inference)

@@
 
-# [ ] include exclusion criteria according to "create_sample_df.R"
 # [ ] Evaluate SPM model comparison (more activation in SPM..)
@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] redo analysis with new sample 
+# [ ] re-do analysis with new sample 
 # [ ] repeat analysis with exclusion of 10% of worst data
@@
 
-Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
-During analyses an implicit mask and an explicit mask were used.
-Therefore, the MNI ICBM 2009a Nonlinear Symmetric brain mask [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in SPM12.7770<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+During analyses an implicit mask and an explicit mask were used. Therefore, the brain mask was used.
+<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+-->
 
@@
 
-<!-- Description marginal model --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
 It therefore accommodates any repeated measurement covariance structures.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+<!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
+Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.
+-->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
 
-<!-- Design matrix, vielleicht beide auch als Abbildung kurz zeigen?
+<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
 --> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
@@
 on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- Nuisance covariates --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+<!-- NUISANCE COVARIATES --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Age and sex were between-subject covariates while FD is time-varying.
@@
 As change in FD and change in BMI share variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
-<!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
+<!-- POST HOC  --> For that reason, we performed each analysis separately one with mean FD and without.
 Differences in results will be reported.
@@
 
-<!-- Inference --> All parametric inferences on activation were drawn based on False Discovery Rate (FDR) adjusted p-values $\leq 0.05$.
-To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
+<!-- INFERENCE --> <!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
+Significant clusters are defined as family-wise error - corrected p<0.05.
+-->
+To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
 In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and is unlike permutation based framework appropriate <!-- ???
-clarify why--> for longitudinal data.
+random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
+Unlike the description in the preregistrastion, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment and $1000$ bootstraps due to required computation time.
+Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
 

### 2020-12-09 code/RunModelBMI.m

@@ -236,8 +236,8 @@ else
         smodel.WB.WB_yes.WB_infType.WB_voxelwise = 0;
     elseif strcmp(param.INFERENCE_TYPE,'cluster')
         % cluster-forming threshold (default)
-        smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_inputType.WB_img = 0;
         smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_clusThresh = 0.001;
+        smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_inputType.WB_img = 0;
     elseif strcmp(param.INFERENCE_TYPE,'tfce')
         % E and H values as default (strongly recommended)
         smodel.WB.WB_yes.WB_infType.WB_TFCE.WB_TFCE_E = 0.5;

### 2020-12-09 code/RunModelFD.m

@@ -235,8 +235,8 @@ else
         smodel.WB.WB_yes.WB_infType.WB_voxelwise = 0;
     elseif strcmp(param.INFERENCE_TYPE,'cluster')
         % cluster-forming threshold (default)
-        smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_inputType.WB_img = 0;
         smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_clusThresh = 0.001;
+        smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_inputType.WB_img = 0;
     elseif strcmp(INFERENCE_TYPE,'tfce')
         % E and H values as default (strongly recommended)
         smodel.WB.WB_yes.WB_infType.WB_TFCE.WB_TFCE_E = 0.5;

### 2020-12-09 code/RunModelGroupTime.m

@@ -275,8 +275,8 @@ else
         smodel.WB.WB_yes.WB_infType.WB_voxelwise = 0;
     elseif strcmp(param.INFERENCE_TYPE,'cluster')
         % cluster-forming threshold (default)
-        smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_inputType.WB_img = 0;
         smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_clusThresh = 0.001;
+        smodel.WB.WB_yes.WB_infType.WB_clusterwise.WB_inputType.WB_img = 0;
     elseif strcmp(param.INFERENCE_TYPE,'tfce')
         % E and H values as default (strongly recommended)
         smodel.WB.WB_yes.WB_infType.WB_TFCE.WB_TFCE_E = 0.5;
@@ -286,10 +286,10 @@ end
 
 %% Other ------------------------------------------------------------------
 % Global calculation - Omit
-smodel.globalc.g_omit = 1;
+smodel.globalc.g_omit = 1; 
 % Global normalisation
 %.Overall grand mean scaling - No
-smodel.globalm.gmsca.gmsca_no = 1;
+smodel.globalm.gmsca.gmsca_no = 1; 
 % . Normalisation - None
 smodel.globalm.glonorm = 1;
 

### 2020-12-09 code/run_or_display.m

@@ -73,14 +73,17 @@
 
 %% ========================================================================
 % set path for spm and path with my functions
-addpath('/data/u_heinrichs_software/MATLAB/spm12/') % swe version 2.2.1
-addpath('/data/u_heinrichs_software/MATLAB/spm12/toolbox/SwE-toolbox-2.2.1/')
-%addpath(genpath('/data/pt_life/data_fbeyer/spm-fbeyer')) % swe version
-%2.1.1 (instead of both above mentioned links)
+% swe version 2.2.1 download of development 
+addpath('/data/u_heinrichs_software/MATLAB/spm12/')
+addpath('/data/u_heinrichs_software/MATLAB/spm12/toolbox/SwE-toolbox-2.2.1-1292020')
+% swe version 2.1.1
+% addpath(genpath('/data/pt_life/data_fbeyer/spm-fbeyer'))
+
 
 % import absolute path for project
 % SET PATH TO code_and_manuscript FOLDER AS CURRENT DIRECTORY !!
-ABS_DIR = readcell("abs_path.csv");
+cd("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code")
+ABS_DIR = readcell("../abs_path.csv");
 ABS_DIR = ABS_DIR{1};
 addpath(fullfile(ABS_DIR,'/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript')) % for functions
 % create a struct, with all important parameters
@@ -99,14 +102,13 @@ param.ROI_PREP = {roi_prep{[4, 6, 12, 14]}}; % {roi_prep{[4, 6, 12, 14]}} or {'N
 % All three models have unique options for covariate definition, the
 % association to a model is indicated by the tens digit (GroupTime_: 1_; 
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
-
 param.MODEL = {'fd','fdIG'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [31, 32];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
+param.COVARIATES = [31,32];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
 
 % define masking and type of inference
 param.MASK = 'brain';               % 'brain' or 'gm'
[…]

### 2020-12-09 manuscript/draft_rsfmri.Rmd
design matrices plot in draft_rsfmri.Rmd; save txt file for scans per region in .m

@@
 library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
-#library(cowplot)
+library(cowplot)
 library(RColorBrewer) # version 1.1.2
@@
 
-##Functional preprocessing
+## Functional preprocessing
 ### Minimal functional preprocessing
@@
 ### Functional connectivity
-To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed ROIs, respectively.
+To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed regions of interest (ROI), respectively.
 Based on FreeSurfer's `aseg.mgz` and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
@@
 
-Whole brain FC maps will be calculated using Pearson’s correlation between average seed time series and whole brain timeseries on the subject level (using nilearn).
+Whole brain FC maps will be calculated using Pearson's correlation between average seed time series and whole brain timeseries on the subject level (using nilearn).
 After transforming from r-to-Z, we will transform the images into Freesurfer longitudinal space, using /mri/transforms/_to_.lta).
@@
 Quality control:
+
 - Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
@@
 -->
-Five data points were excluded due to failure of correct preprocessing attributed to T1-artifacts, and three additional data points were excluded due to extensive head motion (greater then $2 SD$ above the average mean FD).
+Five data points were excluded due to failure of correct preprocessing attributed to T1-artifacts.
+For significant group-time
 <!-- MOTION CORRECTION --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
@@
 Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in SPM12.7770<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
-During analyses an implicit mask and an explicit mask were used. Therefore, the brain mask was used.
-<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+During analyses an implicit mask and an explicit mask were used.
+For this, FSL's standard space the brain mask that corresponds th "152 nonlinear 6th generation" atlas was re-sampled to 3 x 3 x 3 mm$^3$.<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
 -->
@@
 <!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
---> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
-The resulting flexible factorial design <!-- check: difference from full factorial or other?
+--> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and therefore refrained from modelling time as continous variable.
+The resulting factorial design <!-- check: flexible factorial?
 --> contained one regressor for each time point per group.
+```{r "Plot design matrices", fig.align='center', out.width="90%"}
+# design matrix time as factor
+txt_path <- "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files/noExclFD/both/total/"
+
+ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
+condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
+KG_bl <- read.delim(paste0(txt_path,"KG_bl.txt"), head=FALSE)
+KG_fu <- read.delim(paste0(txt_path,"KG_fu.txt"), head=FALSE)
+KG_fu2 <- read.delim(paste0(txt_path,"KG_fu2.txt"), head=FALSE)
+IG_bl <- read.delim(paste0(txt_path,"IG_bl.txt"), head=FALSE)
+IG_fu <- read.delim(paste0(txt_path,"IG_fu.txt"), head=FALSE)
+IG_fu2 <- read.delim(paste0(txt_path,"IG_fu2.txt"), head=FALSE)
+dmf <- data.frame(ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
+colnames(dmf) <- c("ID","condition","KG_bl","KG_fu","KG_fu2","IG_bl","IG_fu","IG_fu2")
+dmf$measurement <- c(1:nrow(dmf))
+
+# design matrix time as continuous variable
+ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
+condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
+group_IG <- read.delim(paste0(txt_path,"group_IG.txt"), head=FALSE)
+group_KG <- read.delim(paste0(txt_path,"group_KG.txt"), head=FALSE)
+tp_IG <- read.delim(paste0(txt_path,"tp_IG.txt"), head=FALSE)
+tp_KG <- read.delim(paste0(txt_path,"tp_KG.txt"), head=FALSE)
+dmc <- data.frame(ID,condition,group_IG,group_KG,tp_IG,tp_KG)
+colnames(dmc) <- c("ID","condition","group_IG", "group_KG","tp_IG","tp_KG")
+dmc$measurement <- c(1:nrow(dmc))
+
+# plot dmf
+input_dmf <- dmf %>%
+  pivot_longer(
+    cols= c("KG_bl","KG_fu","KG_fu2","IG_bl","IG_fu","IG_fu2"),
+    names_to = "regressor")
+input_dmc <- dmc %>%
+  pivot_longer(
+    cols= c("group_IG", "group_KG","tp_IG","tp_KG"),
+    names_to = "regressor")
+
+plotA <- ggplot(input_dmf, aes(x = regressor, y = measurement, fill = value)) + 
+  geom_tile() + 
+  scale_fill_gradient(low="black", high="white") +
+  labs(x="regressors", y="scans", title="(A)") +
+  # design
+  scale_x_discrete(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  theme_bw() + theme(legend.position = "none")
+# + theme(panel.grid = element_blank(),
+#         panel.border = element_rect(colour = "black", size=1),
+#         axis.text.x=element_text(size=9, angle=0, vjust=0.3),
+#         axis.text.y=element_text(size=9),
+#         plot.title=element_text(size=11))
+plotB <- ggplot(input_dmc, aes(x = regressor, y = measurement, fill = value)) + 
+  geom_tile() + 
+  scale_fill_gradient(low="black", high="white") +
+  labs(x="regressors", y="scans", title="(B)") +
+  # design
+  scale_x_discrete(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  theme_bw() + theme(legend.position = "none")
+# + theme(panel.grid = element_blank(),
+#         panel.border = element_rect(colour = "black", size=1),
+#         axis.text.x=element_text(size=9, angle=0, vjust=0.3),
+#         axis.text.y=element_text(size=9),
+#         plot.title=element_text(size=11))
+
+gridExtra::grid.arrange(plotA, plotB, nrow=1)
+```
 For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
-With a model containing average BMI and longitudinal change per subject, we investigated potential effects ..............
-on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
+With a model containing average BMI and longitudinal change per subject, we investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
 <!-- NUISANCE COVARIATES --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
@@
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-As change in FD and change in BMI share variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
-<!-- POST HOC  --> For that reason, we performed each analysis separately one with mean FD and without.
+<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
+As change in FD and change in BMI share common variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
+For that reason, we performed each analysis separately one with mean FD (three and without.
 Differences in results will be reported.
-For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
+We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).
+<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
 These post-hoc analyses deviate from the preregistration and should be considered to be exploratory.
@@
 To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
 In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
 random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Unlike the description in the preregistrastion, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment and $1000$ bootstraps due to required computation time.
+Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment and $1000$ bootstraps due to required computation time.
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
@@
 The effect also disappeared when running those tests without mean FD as nuisance covariate.
-<!-- BMI2tp --> Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the regions of interest for neither of the preprocessing variants.
+<!-- BMI2tp --> Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants.
 Omitting mean FD did not change these results.

### 2020-12-09 code/RunModelBMI.m

@@ -161,7 +161,7 @@ smodel.giftiAdditionalInfo.areaFileForGiftiInputs = {};
 scans_dir = readcell(fullfile(param.INFO_DIR,'scans.txt'), 'Delimiter',' ','Whitespace',"'");
 roi_prep = param.ROI_PREP{crun};
 scans_of_roi = create_scans_list(scans_dir, roi_prep);
-
+writecell(scans_of_roi,fullfile(param.INFO_DIR,strcat("scans_",roi_prep,".txt")))
 % load scans to matlabbatch
 smodel.scans = scans_of_roi;
                                          

### 2020-12-09 code/RunModelFD.m

@@ -155,7 +155,7 @@ smodel.giftiAdditionalInfo.areaFileForGiftiInputs = {};
 scans_dir = readcell(fullfile(param.INFO_DIR,'scans.txt'), 'Delimiter',' ','Whitespace',"'");
 roi_prep = param.ROI_PREP{crun};
 scans_of_roi = create_scans_list(scans_dir, roi_prep);
-
+writecell(scans_of_roi,fullfile(param.INFO_DIR,strcat("scans_",roi_prep,".txt")))
 % load scans to matlabbatch
 smodel.scans = scans_of_roi;
                                          

### 2020-12-09 code/RunModelGroupTime.m

@@ -193,7 +193,7 @@ smodel.giftiAdditionalInfo.areaFileForGiftiInputs = {};
 scans_dir = readcell(fullfile(param.INFO_DIR,'scans.txt'), 'Delimiter',' ','Whitespace',"'");
 roi_prep = param.ROI_PREP{crun};
 scans_of_roi = create_scans_list(scans_dir, roi_prep);
-
+writecell(scans_of_roi,fullfile(param.INFO_DIR,strcat("scans_",roi_prep,".txt")))
 % load scans to matlabbatch
 smodel.scans = scans_of_roi;
                                          

### 2020-12-09 code/run_or_display.m

@@ -102,13 +102,13 @@ param.ROI_PREP = {roi_prep{[4, 6, 12, 14]}}; % {roi_prep{[4, 6, 12, 14]}} or {'N
 % All three models have unique options for covariate definition, the
 % association to a model is indicated by the tens digit (GroupTime_: 1_; 
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
-param.MODEL = {'fd','fdIG'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [31,32];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
+param.MODEL = {'grouptime'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
+param.COVARIATES = [11];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % []
 
 % define masking and type of inference
 param.MASK = 'brain';               % 'brain' or 'gm'
 param.EXCLFD = false;                % false
-param.WILD_BOOT = true;             % false
+param.WILD_BOOT = false;             % false
 param.INFERENCE_TYPE = {'cluster'};   % {'voxel','cluster','tfce'};
 % analysis parameter (estimate or display?)
 param.ONLY_DISPLAY = false;         % false

### 2020-12-10 manuscript/draft_rsfmri.Rmd
optimize figure of design matrices

@@
 author:
-- Hannah Heinrichs* [1]
+- Hannah Sophie Heinrichs* [1]
 - Frauke Beyer* [1][2]
@@
 `r round(tab_age$mean[tab_age$group1=="IG"],2)`).
+Change in BMI throuout the study is depicted in \@ref(fig:figBMIdescr).
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
@@
 
-```{r fig_BMIdescr, fig.align='center', out.width="90%"}
+```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
@@
 # BMI over time each group
-fig_BMIdescr <-
+figBMIdescr <-
   ggplot(final, aes(
@@
   facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig_BMIdescr +
+figBMIdescr +
   scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
@@
 The resulting fMRI data is termed "minimally preprocessed" in the following.
+
 ### ICA-AROMA preprocessing
@@
 This analysis was described in the preregistration as "Then we will employ state-of-the art nuisance regression (ICA-AROMA and CompCor regression of white matter and CSF signal to reduce the effect of physiological noise and head motion on the connectivity estimates."
+
 ### Global Signal regression
@@
 <!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
---> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and therefore refrained from modelling time as continous variable.
+--> For the first model, time was introduced as factor because we did not assume strict linearity for the increase in connectivity and therefore refrained from modeling time as continuous variable (see Figure \@ref(fig:figdesignmatrix)).
 The resulting factorial design <!-- check: flexible factorial?
 --> contained one regressor for each time point per group.
-```{r "Plot design matrices", fig.align='center', out.width="90%"}
+```{r figdesignmatrix,  out.width="80%", fig.align = 'center', fig.cap="Possible design matrices of group-time model.", echo=FALSE}
+# fig.show="hold",  out.width="40%", fig.height = 8, 
 # design matrix time as factor
@@
 IG_fu2 <- read.delim(paste0(txt_path,"IG_fu2.txt"), head=FALSE)
-dmf <- data.frame(ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
-colnames(dmf) <- c("ID","condition","KG_bl","KG_fu","KG_fu2","IG_bl","IG_fu","IG_fu2")
-dmf$measurement <- c(1:nrow(dmf))
+measurement <- c(1:nrow(ID))
+dmf <- data.frame(measurement,ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
+colnames(dmf) <- c("measurement","ID","condition","Nbl","Nfu","Nfu2","Bbl","Bfu","Bfu2")
 
@@
 tp_KG <- read.delim(paste0(txt_path,"tp_KG.txt"), head=FALSE)
-dmc <- data.frame(ID,condition,group_IG,group_KG,tp_IG,tp_KG)
-colnames(dmc) <- c("ID","condition","group_IG", "group_KG","tp_IG","tp_KG")
-dmc$measurement <- c(1:nrow(dmc))
+measurement <- c(1:nrow(ID))
+dmc <- data.frame(measurement,ID,condition,group_IG,group_KG,tp_IG,tp_KG)
+colnames(dmc) <- c("measurement","ID","condition","groupB", "groupN","timeB","timeN")
 
@@
   pivot_longer(
-    cols= c("KG_bl","KG_fu","KG_fu2","IG_bl","IG_fu","IG_fu2"),
+    cols= colnames(dmf[,c(4:ncol(dmf))]),
     names_to = "regressor")
@@
   pivot_longer(
-    cols= c("group_IG", "group_KG","tp_IG","tp_KG"),
+    cols= colnames(dmc[,c(4:ncol(dmc))]),
     names_to = "regressor")
@@
   # design
-  scale_x_discrete(expand = c(0,0)) +
+  scale_x_discrete(expand = c(0,0),labels=colnames(dmf)[4:ncol(dmf)]) +
   scale_y_continuous(expand = c(0,0)) +
-  theme_bw() + theme(legend.position = "none")
+  theme_bw() + theme(legend.position = "none",
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12))
 # + theme(panel.grid = element_blank(),
 #         panel.border = element_rect(colour = "black", size=1),
-#         axis.text.x=element_text(size=9, angle=0, vjust=0.3),
-#         axis.text.y=element_text(size=9),
+#    legend.text = element_text(size = 10)
 #         plot.title=element_text(size=11))
@@
   # design
-  scale_x_discrete(expand = c(0,0)) +
+  scale_x_discrete(expand = c(0,0),labels=colnames(dmc)[4:ncol(dmc)]) +
   scale_y_continuous(expand = c(0,0)) +
-  theme_bw() + theme(legend.position = "none")
+  theme_bw() + theme(legend.position = "none",
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12))
 # + theme(panel.grid = element_blank(),
 #         panel.border = element_rect(colour = "black", size=1),
-#         axis.text.x=element_text(size=9, angle=0, vjust=0.3),
-#         axis.text.y=element_text(size=9),
+#    legend.text = element_text(size = 10)
 #         plot.title=element_text(size=11))

### 2020-12-14 manuscript/draft_rsfmri.Rmd
restructured manuscript and added QA section

--- 2020-12-09 manuscript/draft_rsfmri.Rmd
@@
 # [ ] Evaluate SPM model comparison (more activation in SPM..)
+# [ ] Introduction
 # [ ] Discussion
@@
 # [ ] brain images mit nilearn plotting tools einbinden
+# [x] Figure zu weight loss
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] re-do analysis with new sample 
-# [ ] repeat analysis with exclusion of 10% of worst data
 
@@
 
-```{r "knit setup", include=FALSE}
+```{r "knit setup"}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
@@
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
+```{r test wd}
+print(getwd())
+```
 
@@
 The workflow can be found on [github](https://github.com/fBeyer89/ADI_preproc/).
+
 ### Anatomical preprocessing
@@
 
-##Functional preprocessing
+## Functional preprocessing
 ### Minimal functional preprocessing
 The function preprocessing module included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
-In more detail, the transforms derived from the latter three steps were combined into one transformation.
-For further analysis and ICA-AROMA processing, the data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
-This part of the analysis was described like "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated." in the preregistration.
-The resulting fMRI data is termed "minimally preprocessed" in the following.
+In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
+For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
+*This part of the analysis was described in the preregistration as "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated."*
+
 ### ICA-AROMA preprocessing   
-We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion- FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
-The minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
-Finally, the timeseries of the  nuisance components were regressed from the overall timeseries.
+We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
+As described in [@Pruim2015], the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+Finally, the timeseries of the  nuisance components were regressed from the raw fMRI timeseries.
 As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result.
@@
 After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
-This analysis was described in the preregistration as "Then we will employ state-of-the art nuisance regression (ICA-AROMA and CompCor regression of white matter and CSF signal to reduce the effect of physiological noise and head motion on the connectivity estimates."
+*This analysis was described in the preregistration as "Then we will employ state-of-the art nuisance regression (i.e.
+ICA-AROMA with compCor regression (CC) and, optionally global signal regression (GSR) (Ciric et al., 2017; Parkes et al., 2018) to reduce the effect of physiological noise and head motion on the connectivity estimates."*
+
 ### Global Signal regression   
@@
 Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA fMRI data and high-pass filtered the resulting file as above.
-In the preregistration we stated " As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but introduces spatial dependency and spurious correlations, we perform GSR after ICA-AROMA in a separate sensitivity analysis."
+*In the preregistration we stated " As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but introduces spatial dependency and spurious correlations, we perform GSR after ICA-AROMA in a separate sensitivity analysis."*
 
@@
 Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (min, aroma+cc, aroma+gsr).
-Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived during anatomical preprocessing.
+We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
+*In the preregistration we stated "Resting-state functional connectivity (FC) using seed-based connectivity analysis extracted with in-house pre-processing pipelines (see (Beyer et al., 2019)): Whole-brain FC maps for a priori defined seed regions (defined based on the results of the subcortical segmentation and Desikan-Killiany cortical parcellation in the Freesurfer version 6.0.0 longitudinal processing stream).
+For the reward network a ROI combined from bilateral Nucleus accumbens from the subcortical segmentation (26, Left-Accumbens-area, 58, Right-Accumbens-area) was selected.
+Ventro-medialprefrontal cortex was not selected as seed region due to low signal-to-noise ratio in this part of the brain in many participants (noticed during quality checking).
+For the default mode network, the seed region was combined from bilateral precuneus from the Desikan-Killiany atlas (1025 ctx-lh-precuneus, 2025 ctx-rh-precuneus).
+We will calculate  voxel-wise seed-based connectivity mapsin the individual subject’s space (which is coregistered to the Freesurfer longitudinally preprocessed timepoint) and apply r-to-Z-transform to these images.
+These steps will be performed using nilearn, nipype and python.
+Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).*
 
 ### Quality Assessment
-<!-- from Prereg -->
+#### Anatomical Quality control
+FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
+We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration .
+<!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
+*In the preregistration we stated that "Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary", and that "We will exclude participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
+caused Freesurfer to fail or perform very poorly (rated “Failed” according to (Klapwijk et al., 2019)".*
 
-We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
+#### Functional Quality control
+Resting state fMRI data quality control was performed along the protocol of @Ciric_2018.
+As the expected average head motion is high in this sample [@Beyer_2020], it is difficult to determine an appropriate exclusion threshold which does not lead to the exclusion of a high number of individuals.
+Thus, we did not exclude any participants based on mean framewise displacement estimates beyond those who failed the anatomical preprocessing.
+We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
+In addition, for models showing a significant group-by-time interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
+On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
+Figure \@ref{carpet_plot} shows an exemplary carpet plot with (from top to bottom): DVARS (spatial root mean square of the data after temporal differencing), mFD, percent outliers defined by AFNI `OutlierCount`, voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data.
+All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
+We evaluated mFD-FC relationships by computing a functional connectome based on (Power et al., 2011).
+In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
+Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{example_FC}.
+We reported the mFD-FC correlation per preprocessing scheme.
+Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.
 
-Whole brain FC maps will be calculated using Pearson’s correlation between average seed time series and whole brain timeseries on the subject level (using nilearn).
-After transforming from r-to-Z, we will transform the images into Freesurfer longitudinal space, using /mri/transforms/_to_.lta).
-Then, will calculate difference maps between BL-FU1 (confirmatory hypothesis), BL-FU2, FU1-FU2 (exploratory hypotheses).
-These images will be transformed into MNI space (using ANTS non-linear registration from the longitudinal template to MNI space) for statistical analysis.
-Aggregated FC values will be calculated using Pearson’s correlation of average seed time series from pre-defined in the longitudinal template space.
-Aggregated values will be extracted for the reward and the default mode network.
+### Aggregated FC values
+To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
+We used the modified SwE option and used global signal regressed data as input.
+Network masks were formed from all voxels with a clusterwise p_FWE < 0.05 and we extracted the average connectivity from these masks for all data points.
 
-Quality control:
-- Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary.
-- Participants who cannot be correctly preprocessed due to (uncorrectable) artifacts in the T1-weighted or resting-state fMRI will be excluded.
--  We will first perfom the analysis with all participants who were correctly preprocessed.
-In a sensitivity analysis, participants with mean FD larger than the average mean FD + 2 standard deviations will be excluded.
+<!-- param.ROI_PREP = {roi_prep{[6,14]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'} -->
+<!-- % Model definition -->
+<!-- % All three models have unique options for covariate definition, the -->
+<!-- % association to a model is indicated by the tens digit (GroupTime_: 1_;  -->
+<!-- % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit -->
+<!-- param.MODEL = {'alltp'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp}  -->
+<!-- param.COVARIATES = [41];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % [41, 42, 43] -->
 
-We will also extract mean and maximal framewise displacement (FD) from 6 translational/rotational motion parameters of the whole timeseries according to (Power et al., 2012) and calculate mean FD as a summary measure of head motion.
+<!-- % define masking and type of inference -->
+<!-- param.MASK = 'brain';               % 'brain' -->
+<!-- param.EXCLFD = false;               % false -->
+<!-- param.WILD_BOOT = false;            % false -->
+<!-- param.INFERENCE_TYPE = {'cluster'}; % {'voxel','cluster','tfce'}; -->
 
--	Anthropometry: body mass index, % body fat from bioelectrical impedance measurement
--	Grouping variable is intervention versus control group
-
-<!-- EXCLUSION CRITERIA --> Participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
-caused Freesurfer to fail or perform very poorly, i.e.
-rated "Failed" according to @Klapwijk_2019) were excluded because a subject's motion is highly similar across scans and we thus suspect the fMRI data to also suffer from motion artifacts in these subjects.
-<!-- As the expected average head motion is high [@Hodgson_2016], we did not exclude any other participants based on head motion estimates because it is difficult to determine a critical threshold in this sample.
--->
-Five data points were excluded due to failure of correct preprocessing attributed to T1-artifacts, and three additional data points were excluded due to extensive head motion (greater then $2 SD$ above the average mean FD).
-<!-- MOTION CORRECTION --> We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].
-Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012, log-transformed and examined within and separate from the main models.
-
-### Outcomes
-
-The main outcome measure was the aggregated functional connectivity (FC) values for the two ROI NAcc and PCC.
+*In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
+The mask will include all voxels surviving FWE-correction of p<0.05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"*
 
@@
 
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in SPM12.7770<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
-During analyses an implicit mask and an explicit mask were used. Therefore, the brain mask was used.
-<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
--->
+Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
+During analyses an implicit mask and an explicit mask were used.
+Therefore, the MNI ICBM 2009a Nonlinear Symmetric brain mask [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
 
-<!-- MODEL SETUP --> Two theoretically derived models were set up.
-The first model was aimed at testing the interaction of time and group: $FC = group + time + group*time$.
-The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.
+<!-- MODEL SETUP --> As specified in the preregistration, we tested the interaction of time and group: $FC = group + time + group*time$.
+Unlike prespecified, we did not include the model without covariates (as this is deprecated \@ref{nuisance_paper}).
+Instead, we calculated a model ajusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2).
+We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
 These were examined each for correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $16$ a priori planned analyses.
@@
 Computation with the grey matter mask were exploratory and deviate from the preregistration.
+The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.
 
-<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+<!-- Description marginal model --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
 It therefore accommodates any repeated measurement covariance structures.
-<!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
-Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.
--->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
 
-<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
+<!-- Design matrix, vielleicht beide auch als Abbildung kurz zeigen?
 --> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
@@
 on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- NUISANCE COVARIATES --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+<!-- Nuisance covariates --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Age and sex were between-subject covariates while FD is time-varying.
@@
 As change in FD and change in BMI share variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
-<!-- POST HOC  --> For that reason, we performed each analysis separately one with mean FD and without.
+<!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
 Differences in results will be reported.
@@
 
-<!-- INFERENCE --> <!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
-We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
-Significant clusters are defined as family-wise error - corrected p<0.05.
--->
-To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+<!-- Inference --> All parametric inferences on activation were drawn based on False Discovery Rate (FDR) adjusted p-values $\leq 0.05$.
+To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
 In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Unlike the description in the preregistrastion, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment and $1000$ bootstraps due to required computation time.
-Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
+random field theory), and is unlike permutation based framework appropriate <!-- ???
+clarify why--> for longitudinal data.
+
+<!-- R analysis of aggregated values --> 
+We analyzed the aggregate connectivity values using linear mixed models in the R package lme4.
+Following the preregistration, we first investigated a time-by-group interaction between baseline and followup.
+We used a model comparison of R1 and R0 models with `anova`:
+`R1= lmer(agg FC ~ timepoint*group + (1|subj))`
+`R0 = lmer(agg FC ~ timepoint + group + (1|subj))`
+We performed a sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI.
+Again, the group-by-time point interaction was considered significant if a model comparison between R1 and Null models using anova-function yielded a p<0.05.
+Finally, we estimated the above-mentioned interaction model for all three time points.
+Again, the interaction was considered significant if a model comparison between R1 and R0 using `anova` had p<0.05.
 
@@
 
-## Manipulation check
-<!-- check network presence -->
+## Quality control
+We performed different quality controls for the rsfMRI data.
+As described in \@ref[Beyer_2020], there was a significant interaction of group and time on mFD.
+In particular, the intervention group reduced in head motion in the follow-up assessments compared to the control group.
+We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data.
+We saw that DVARS did not qualitatively differ between groups and timepoints, and that the correlation of DVARS and mFD remained similar over time points.
+This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
+```{r summary table of QA measures}
+for (i in 1:nrow(final))
+  {tmp=strsplit(toString(final[i,"scan_dir"]),'/')[[1]][8]
+   conf=read.csv(paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",tmp,"/confounds.csv"))
+   tmp=cor.test(conf$DVARS,conf$FD)
+   final[i,"corr_DVARS_FD"]=tmp$estimate
+   final[i,"mean_DVARS"]=mean(conf$DVARS,na.rm = TRUE)
+   final[i,"max_DVARS"]=max(conf$DVARS,na.rm = TRUE)
+   final[i,"mean_outliers"]=mean(conf$outliers,na.rm = TRUE)
+  }
 
-<!-- singletp (t-tests), nur diesen Check für BL reporten (gib dann oben noch an, dass du dafür SPM verwendet hast) --> We ran separate one-sample t-test for network activation at baseline controlling for log mean FD, age, and sex for each time point to check whether the assumed reward and default mode network was at all identifiable.
+```
 
-## Interaction effect of intervention and time
+```{r fig_summary_dvars_mFD, fig.align='center', out.width="100%"}
+# eval = TRUE, fig.asp = 0.7, fig.width = 6
+condition.labs <- c("BARS","NBARS")
+names(condition.labs) <- c("IG","KG")
+# BMI over time each group
+fig_DVARSdescr <-
+  ggplot(final, aes(
+    x = tp,
+    y = mean_DVARS,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+fig1=fig_DVARSdescr +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time [months]") + ylab ("Mean DVARS") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
 
+fig_mFDdescr <-
+  ggplot(final, aes(
+    x = tp,
+    y = meanFD,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+fig2=fig_mFDdescr +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time [months]") + ylab ("mFD in mm") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
+fig_mFDDVARSdescr <-
+  ggplot(final, aes(
+    x = tp,
+    y = corr_DVARS_FD,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+fig3=fig_mFDDVARSdescr +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time [months]") + ylab ("DVARS-FD correlation") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
+
+(fig1 | fig2 ) /
+      fig3
+```
+
+```{r FDFC correlation, include=FALSE}
+source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
+
+res=calc_qc_fc(final)
+```
+Further, we evaluated the impact of our denoising pipelines.
+Table \@ref(tab:table_qc_fc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
+Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
+AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
+Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
+```{r table_qc_fc, echo=FALSE}
+tab_qc_fc <-
+  knitr::kable(
+    round(res,3),
+    #row.names = c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR"),
+    #col.names = c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval'),
+    format = "latex",
+    booktabs = T,
+    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
+
+tab_qc_fc
+```
+
+## Interaction effect of intervention and time (preregistered analysis)
+
+### DMN/reward network with whole brain mask (preregistered analysis)
 <!-- Brain mask: Group*Time Interaction -->
+#### without adjustment
 There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
@@
 Results obtained for parametric estimation and non-parametric wild bootstrap did not differ.
+#### adjusting for age, sex, meanFD
 When reducing the number of nuisance covariates by leaving out mean FD, connectivity was found between the PCC with one other voxel.
@@
 
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+
+### DMN/reward network with GM mask (exploratory analysis)
 <!-- GM mask: Group*Time Interaction -->
@@
 
-## Effect of BMI
+## Interaction effect group by time on aggregated FC (preregistered analysis)
+
+### DMN for two timepoints
+```{r calc DMN two tp, include=FALSE}
+library(lme4)
+library(sjPlot)
+R1= lmer(DMN_FC~ timepoint*group + (1|subj))
+R0 = lmer(DMN_FC ~ timepoint + group + (1|subj))
+```
+### DMN for two timepoints, adjusted for age, sex, mean FD
+```{r calc DMN two tp adjusted, include=FALSE}
+library(lme4)
+library(sjPlot)
+R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+```
+### DMN for three timepoints, adjusted for age, sex, mean FD
+```{r calc DMN three tp, include=FALSE}
+library(lme4)
+library(sjPlot)
+R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lmer(DMN_FC ~ timepoint + group ++ age + sex + meanFD (1|subj))
+```
+
+### reward
+```{r calc rew two tp, include=FALSE}
+R1= lmer(rew_FC ~ timepoint*group + (1|subj))
+R0 = lmer(rew_FC ~ timepoint + group + (1|subj))
+```
+### reward for two timepoints, adjusted for age, sex, mean FD
+```{r calc rew two tp adjusted, include=FALSE}
+library(lme4)
+library(sjPlot)
+R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
+```
+### reward for three timepoints, adjusted for age, sex, mean FD
+```{r calc rew three tp, include=FALSE}
+library(lme4)
+library(sjPlot)
+R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+```
+
+## Effect of mean and change in BMI (exploratory analysis)
 
@@
 
-## Further investigation of the effect of FD
+## Effect of FD (exploratory analysis)
 
@@
 
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+## Effect on hypothalamic connectivity (exploratory analysis)
+
+<!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
+Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
+<!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
 
@@
 ```{r expample plots, eval=FALSE}
-
-df_wide <- tidyr::pivot_wider(
-  data = final,
-  id_cols = "subj.ID",
-  names_from = "tp",
-  values_from = c("tp", "BMI", "logmFD")
-)
-
-if (tp == "BLFU") {
-  s = 2
-} else {
-  s = 1
-}
-
-BMI_vector <- c("BMI_fu2", "BMI_fu", "BMI_bl")
-BMI_vector <- c("BMI_bl", "BMI_fu", "BMI_fu2")
-FD_vector <- c("logmFD_fu2", "logmFD_fu", "logmFD_bl")
-df_wide$avgBMI <- apply(df_wide[, BMI_vector[s:3]],
-                        MARGIN = 1,
-                        FUN = mean,
-                        na.rm = TRUE)
-df_wide$avgFD <- apply(df_wide[, FD_vector[s:3]],
-                       MARGIN = 1,
-                       FUN = mean,
-                       na.rm = TRUE)
-df_long <- tidyr::pivot_longer(
-  data = df_wide,
-  cols = all_of(order_vector),
-  # bl, fu, fu2 (maintain original order)
-  # CAREFUL: Order according to dataframe df
-  names_to = "tp_tp",
-  values_to = "tp",
-  values_drop_na = TRUE
-)
-all(df_long$tp == final$tp) # check if order correct
-
-head(final[, c("subj.ID", "tp")])
-head(df_long[, c("subj.ID", "tp")])
-
-df_long$cgnBMI <- final$BMI - df_long$avgBMI
-df_long$avgBMIc <-
-  df_long$avgBMI - mean(df_long$avgBMI, na.rm = TRUE)
-df$avgBMIc <- df_long$avgBMIc
-df$cgnBMI <- df_long$cgnBMI
-
-df_long$cgnFD <- final$logmFD - df_long$avgFD
-df_long$avgFDc <-
-  df_long$avgFD - mean(df_long$avgFD, na.rm = TRUE)
-final$avgFDc <- df_long$avgFDc
-final$cgnFD <- df_long$cgnFD
-
 # example plots for interaction effect (arbitrary variables for illustration purposes)

### 2020-12-14 code/create_sample_df.R

@@ -8,7 +8,7 @@ create_sample_df <- function(group = "all", tp = "all"){
   
   # participants with mri data (txt file with path to .nii file) ---------------
   
-  mri_files=read.table("../SwE_files/scans.txt")
+  mri_files=read.table("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level //SwE_files/scans.txt")
   # split info from path in txt
   for (i in 1:nrow(mri_files)){
     tmp=strsplit(toString(mri_files[i,1]),'/')[[1]][8]
@@ -23,13 +23,16 @@ create_sample_df <- function(group = "all", tp = "all"){
   # load info from full sample and merge with path info (fmri only)-------------
   
   full_sample=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv") # CAVE: info file elsewhere !!!
-  full_sample=full_sample[,c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI")]
+  full_sample=full_sample[,c("subj.ID","condition","tp","Age_BL","Sex","meanFD","maxFD", "BMI", "Final_Score",
+                             "Excessive_motion", "Exclude", "Check_comments")]
   condition=merge(mri_files, full_sample[!is.na(full_sample$condition),], by=c("subj.ID","tp"))
   df=plyr::join(mri_files,condition)
   rm(mri_files, full_sample, condition, tmp, i)
   
-  
-  # preparation of variables ---------------------------------------------------
+  # exclude participants who failed in FreeSurfer
+  df$Exclude=is.na(df$Exclude)
+  df=df[df$Exclude==TRUE,]
+  # preparation of variables --------------------------
   
   df$logmFD <- log10(df$meanFD)
   
@@ -76,7 +79,7 @@ create_sample_df <- function(group = "all", tp = "all"){
   }
 
   # ----------------------------------------------------------------------------
-  # Additional computations (also better to implement in Matlab)
+  # Preparation of within/between-participant variables 
   # ----------------------------------------------------------------------------
   
   if ((group == "all" | group == "IG") & (tp == "all" | tp == "BLFU")) {

### 2020-12-14 manuscript/draft_rsfmri.Rmd
adopted to new folder structure and Wild bootstrapping

--- 2020-12-10 manuscript/draft_rsfmri.Rmd
@@
 # [ ] repeat analysis with exclusion of 10% of worst data
+# [ ] Network tests
 
@@
                           names_from = "tp",
-                          values_from = c("BMI","logmFD"))
+                          values_from = c("BMI","meanFD"))
 ```
@@
 descr <- psych::describeBy(dfw[,5:ncol(dfw)], group=dfw$condition,mat=T)
-descr$predictor <- c(rep("BMI",6),rep("logmFD",6))
+descr$predictor <- c(rep("BMI",6),rep("meanFD",6))
 descr$tp <- car::recode(descr$vars, "c(1,4)='fu2';c(2,5)='fu';c(3,6)='bl'")
@@
 dmf <- data.frame(measurement,ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
-colnames(dmf) <- c("measurement","ID","condition","Nbl","Nfu","Nfu2","Bbl","Bfu","Bfu2")
+colnames(dmf) <- c("measurement","ID","condition","N0","N6","N12","B0","B6","B12")
 
@@
 
-<!-- Brain mask: Group*Time Interaction -->
+<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
 There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
-Nor was there a significant main effect for any of the covariates of interest (time, group) in voxelwise inference with FDR correction.
-Results obtained for parametric estimation and non-parametric wild bootstrap did not differ.
-When reducing the number of nuisance covariates by leaving out mean FD, connectivity was found between the PCC with one other voxel.
-In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm] yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
-However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{\mathrm{FWEcorr}} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
--->.
-There was no other significant co-activation for any other region of interest for neither of the preprocessing variants.
+Nor was there a significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD, no connectivity was found between the PCC with one other voxel.
+Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
+<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
 
-<!-- GM mask: Group*Time Interaction -->
+<!-- GM mask: Group*Time Interaction [OLD] -->
 Parametric and non-parametric significance test for the model testing for group-time-interaction did not show any different results when using a grey matter mask.
@@
 Ventures to verify effects by re-calculating with non-parametric Wild bootstrap confirmed supported presumption as effects were not robust across estimation procedures.
+<!-- limitations -->
+Without an a priori power analysis it is not possible to determine if there is no effect or 
+To clarify, extended considerations should be made to use a priori power analysis (e.g. with fMRIpower) or check evidence of absence with alternative
 

### 2020-12-14 code/AllTPEval.m

@@ -20,20 +20,14 @@ end
 %% ========================================================================
 function display_message(COVARIATES)
 % displays message according to the COVARIATES, that define the model
-if COVARIATES == 31
+if COVARIATES == 41
     fprintf('%s\n',...
-    'Model1_bmi_fd_cage_sex: In case of parametric estimation',...
-    'please type in the following contrasts for all runs',...
-    'avgBMI [0 1 0 0 0 0 0]',...
-    'BMIcgn [0 0 1 0 0 0 0]',...
-    'avgFD [0 0 0 1 0 0 0]',...
-    'FDcgn [0 0 0 0 1 0 0]')
-elseif COVARIATES == 32
+    'fsthdbt',...
+    '...')
+elseif COVARIATES == 42
     fprintf('%s\n',...
-    'Model2_fd_cage_sex: In case of parametric estimation',...
-    'please type in the following contrasts for all runs',...
-    'avgFD [0 1 0 0 0]',...
-    'FDcgn [0 0 1 0 0]')
+    'jzdizd',...
+    '...')
 end
 
 end
@@ -49,15 +43,20 @@ param.COVARIATES = param.COVARIATES(j);
 param.INFERENCE_TYPE = param.INFERENCE_TYPE(k);
 
 
-
 if param.ONLY_DISPLAY && not(param.WILD_BOOT) %strcmp(param.ACTION,'display')
     display_message(param.COVARIATES)
     % wait so displayed information can be read
     pause(3)
 end
 
+if param.EXCLFD==true
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
+else
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
+end
+
 if strcmp(param.MODEL,'alltp')
-    param.INFO_DIR = fullfile(param.INFO_DIR, 'total');
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'both/total/');
 end
 
 nrun = length(param.ROI_PREP);
@@ -139,6 +138,11 @@ cd(param.INFO_DIR)
 % save directory of (new) folder
 smodel.dir = {out_folder};
 
+% cifti + gifti additional information (SwE 2.2.1)
+smodel.ciftiAdditionalInfo.ciftiGeomFile = struct('brainStructureLabel', {}, 'geomFile', {}, 'areaFile', {});
[…]

### 2020-12-14 code/RunModelBMI.m

@@ -73,7 +73,7 @@ for crun = 1:nrun
         param.wild_con = false;
         % check if analysis has been conducted already
         [out_folder, exist_already] = create_out_folder(param, crun);
-        fprintf('If necessary create output folder \n %s ...\n',out_folder)
+        fprintf('Output folder: \n %s ...\n',out_folder)
         clear matlabbatch
         fprintf('Specify Model...\n')
         [matlabbatch,location_SwE_mat] = SpecifyModel(param, crun, out_folder);
@@ -83,9 +83,9 @@ for crun = 1:nrun
             matlabbatch = DisplayResults(location_SwE_mat);
             spm_jobman('run', matlabbatch);
             pause(param.VIEWSEC)
-        elseif not(exist_already) || param.OVERWRITE %strcmp(param.ACTION,'overwrite')
+        elseif not(exist_already) || param.OVERWRITE 
             fprintf('Estimate model...\n')
-            % delete existing files in folder if existent
+            % delete existing files in folder if existent 
             if exist_already 
                 rmdir(out_folder,'s'); % delete former dir
                 mkdir(out_folder); % create new empty one

### 2020-12-14 code/RunModelFD.m

@@ -73,7 +73,7 @@ for crun = 1:nrun
         param.wild_con = false;
         % check if analysis has been conducted already
         [out_folder, exist_already] = create_out_folder(param, crun);
-        fprintf('If necessary create output folder\n %s ...\n',out_folder)
+        fprintf('Output folder: \n %s ...\n',out_folder)
         clear matlabbatch
         fprintf('Specify Model...\n')
         [matlabbatch,location_SwE_mat] = SpecifyModel(param, crun, out_folder);

### 2020-12-14 code/RunModelGroupTime.m

@@ -135,7 +135,7 @@ for crun = 1:nrun
             param.wild_con = wild_con;
             % check if analysis has run already
             [out_folder, exist_already] = create_out_folder(param, crun);
-            fprintf('If necessary create output folder \n %s ...\n',out_folder)
+            fprintf('Output folder: \n %s ...\n',out_folder)
             clear matlabbatch
             fprintf('Specify Model...\n')
             [matlabbatch, location_SwE_mat] = SpecifyModel(param, crun, out_folder); 
@@ -146,7 +146,7 @@ for crun = 1:nrun
                 spm_jobman('run', matlabbatch);
                 pause(param.VIEWSEC)
                 save("xSwE.mat")
-            elseif not(exist_already) || param.OVERWRITE % strcmp(param.ACTION,'overwrite')
+            elseif not(exist_already) || param.OVERWRITE % else
                 fprintf('Estimate model...\n')
                 % delete existing files in folder if existent
                 if exist_already

### 2020-12-14 code/SingleTPEval.m

@@ -2,11 +2,17 @@ function SingleTPEval(param)
 % t-test for network activity seperate for each time point.
 % Normal SPM analysis - results can be viewed as is common for SPM.
 
-% if not otherwise specified
-param.tp = readmatrix(fullfile(param.INFO_DIR,'total','tp.txt'));
+if param.EXCLFD==true
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
+else
+    param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
+end
+
+
+param.INFO_DIR = fullfile(param.INFO_DIR,'both/total/');
+param.tp = readmatrix(fullfile(param.INFO_DIR,'tp.txt'));
 
 nrun = length(param.ROI_PREP);
-ntp = length(unique(param.tp));
 for j = 1:length(param.COVARIATES)
     for crun = 1:nrun % for every roi_prep
         for ctp = 1 %:ntp % for every time point
@@ -31,14 +37,13 @@ elseif ctp == 3
     param.wave = 'fu2'; 
 end
 
-[out_folder, exist_already] = create_out_folder(param, crun);
-if ~exist(out_folder, 'dir')
-    mkdir(out_folder)
-end
+[out_folder, already_exist] = create_out_folder(param, crun);
+fprintf('Output folder: \n %s ...\n',out_folder)
 clear matlabbatch
 fprintf('Specify model...\n')
 [matlabbatch] = SpecifyModel(param, crun, ctp, out_folder);
 fprintf('Run model and display results...\n')
+%fprintf('Evaluate network test with covariates %i...\n',param.COVARIATES(j))
 % run SPM 
 spm('defaults', 'FMRI'); 
 spm_jobman('run', matlabbatch);
@@ -73,8 +78,8 @@ if param.ONLY_DISPLAY == false
         end
     end
     scans_of_roi = fullfile(scans_dir, roi_prep, FC_files);
-    % filter by current time point
-    scans_of_roi(param.tp == ctp)
+    % filter by current time point (check)
+    % scans_of_roi(param.tp == ctp)
 
     %% Design matrix
     [cov, des] = ConstructDesignMatrix(param, scans_of_roi);
@@ -134,10 +139,6 @@ end
 %% ------------------------------------------------------------------------
 function [cov, des] = ConstructDesignMatrix(param, scans_of_roi)
     if param.COVARIATES == 41
[…]

### 2020-12-14 code/run_or_display.m

@@ -8,7 +8,9 @@
 % Use string for stating *ROI and preprocessing step* and z-transform in
 % in the following format, e.g. ROI_PREP = 'PCC_min_z' or 'NAcc_aroma'.
 % ROI_PREP must be in a cell array of multiple or a single cell!
-% For roi enter either 'PCC' or 'Nacc'. For prep enter either 'min', 
+% For roi enter either 'PCC', 'Nacc', 'MH' or 'LH', corresponding to DMN,
+% reward network, medial and lateral hypothalamus, respectively.
+% For prep enter either 'min', 
 % 'aroma', 'cc', or 'gsr'. If desired, append 'z'.
 % To evaluate all specific ROI/ preprocessing combinations type
 % ROI_PREP = readcell(fullfile(INFO_DIR,'ROIs.txt'), 'Delimiter',' ','Whitespace',"'");
@@ -32,8 +34,9 @@
 % 22: bmi-age-sex
 % 31: bmi-fd-age-sex
 % 32: fd-age-sex
-% 41: network-age-sex
-% 42: network-meanFD-age-sex
+% 41: network-meanFD-age-sex
+% 42: network-age-sex
+% 43: network (= one-sample t-test to check for network topography)
 %
 % Specify directories relativ to project folder. The absolute path to the
 % project folder should be specified in a file called "abs_path.csv".
@@ -50,14 +53,15 @@
 % WILD_BOOT if set on true will use non-parametri Wild Bootstrap instead 
 % of parametric estimation. It will automatically run all relevant contrast 
 % of the model (as defined in the script).
-% NOTE: the contrast implemented in the script are optimized for 
-% 'bmi-age-sex-meanFD'.
-% 
+% Note: Bootstrap is not applicable to 'alltp' and 'singletp'
 % When using WILD_BOOT = true, you must specify which kind of inference
 % is used for bootstrapping:
 % 'voxel': voxelwise 
 % 'cluster': clusterwise
 % 'tfce': TFCE
+% 
+% EXCLFD specifies application of 2nd exclusion criterion (exclusion of 10% 
+% of datapoints with worst mean FD); EXCLFD = false means "not applied".
 % 
 % *** IMPORTANT ***
 % If DISPLAY_ONLY is set on false and the model has not yet been estimated,
@@ -74,11 +78,10 @@
[…]

### 2020-12-14 manuscript/draft_rsfmri.Rmd

--- 2020-12-14 manuscript/draft_rsfmri.Rmd
@@
 # [ ] Evaluate SPM model comparison (more activation in SPM..)
-# [ ] Introduction
 # [ ] Discussion
@@
 # [ ] brain images mit nilearn plotting tools einbinden
-# [x] Figure zu weight loss
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
+# [ ] re-do analysis with new sample
+# [ ] repeat analysis with exclusion of 10% of worst data
+# [ ] Network tests
 
@@
 We used the modified SwE option and used global signal regressed data as input.
-Network masks were formed from all voxels with a clusterwise p_FWE < 0.05 and we extracted the average connectivity from these masks for all data points.
+Network masks were formed from all voxels with a clusterwise $p_FWE < 0.05$ and we extracted the average connectivity from these masks for all data points.
 
@@
 
+<!-- MOTION CORRECTION We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
+As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].
+Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012, log-transformed and examined within and separate from the main models.
+-->
+
 ## Analysis
 
-Statistical analysis were performed in MATLAB using the sandwich estimator (SwE) toolbox 2.2.1 [@Guillaume_2014] as implemented in SPM12 and R version 3.6.1 [@team2013r].
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in SPM12.7770<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
 During analyses an implicit mask and an explicit mask were used.
-Therefore, the MNI ICBM 2009a Nonlinear Symmetric brain mask [http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+For this, FSL's standard space the brain mask that corresponds th "152 nonlinear 6th generation" atlas was re-sampled to 3 x 3 x 3 mm$^3$.<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+-->
 
@@
 Unlike prespecified, we did not include the model without covariates (as this is deprecated \@ref{nuisance_paper}).
-Instead, we calculated a model ajusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2).
+Instead, we calculated a model adjusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2).
 We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
@@
 
-<!-- Description marginal model --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
 It therefore accommodates any repeated measurement covariance structures.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
-
-<!-- Design matrix, vielleicht beide auch als Abbildung kurz zeigen?
---> For the first model, time was introduced as factor because did not assume strict linearity for the increase in connectivity and sought to increase power to detect specific trends.
-The resulting flexible factorial design <!-- check: difference from full factorial or other?
+<!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
+Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.
+-->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+
+<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
+--> For the first model, time was introduced as factor because we did not assume strict linearity for the increase in connectivity and therefore refrained from modeling time as continuous variable (see Figure \@ref(fig:figdesignmatrix)).
+The resulting factorial design <!-- check: flexible factorial?
 --> contained one regressor for each time point per group.
+```{r figdesignmatrix,  out.width="80%", fig.align = 'center', fig.cap="Possible design matrices of group-time model.", echo=FALSE}
+# fig.show="hold",  out.width="40%", fig.height = 8,
+# design matrix time as factor
+txt_path <- "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files/noExclFD/both/total/"
+
+ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
+condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
+KG_bl <- read.delim(paste0(txt_path,"KG_bl.txt"), head=FALSE)
+KG_fu <- read.delim(paste0(txt_path,"KG_fu.txt"), head=FALSE)
+KG_fu2 <- read.delim(paste0(txt_path,"KG_fu2.txt"), head=FALSE)
+IG_bl <- read.delim(paste0(txt_path,"IG_bl.txt"), head=FALSE)
+IG_fu <- read.delim(paste0(txt_path,"IG_fu.txt"), head=FALSE)
+IG_fu2 <- read.delim(paste0(txt_path,"IG_fu2.txt"), head=FALSE)
+measurement <- c(1:nrow(ID))
+dmf <- data.frame(measurement,ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
+colnames(dmf) <- c("measurement","ID","condition","N0","N6","N12","B0","B6","B12")
+
+# design matrix time as continuous variable
+ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
+condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
+group_IG <- read.delim(paste0(txt_path,"group_IG.txt"), head=FALSE)
+group_KG <- read.delim(paste0(txt_path,"group_KG.txt"), head=FALSE)
+tp_IG <- read.delim(paste0(txt_path,"tp_IG.txt"), head=FALSE)
+tp_KG <- read.delim(paste0(txt_path,"tp_KG.txt"), head=FALSE)
+measurement <- c(1:nrow(ID))
+dmc <- data.frame(measurement,ID,condition,group_IG,group_KG,tp_IG,tp_KG)
+colnames(dmc) <- c("measurement","ID","condition","groupB", "groupN","timeB","timeN")
+
+# plot dmf
+input_dmf <- dmf %>%
+  pivot_longer(
+    cols= colnames(dmf[,c(4:ncol(dmf))]),
+    names_to = "regressor")
+input_dmc <- dmc %>%
+  pivot_longer(
+    cols= colnames(dmc[,c(4:ncol(dmc))]),
+    names_to = "regressor")
+
+plotA <- ggplot(input_dmf, aes(x = regressor, y = measurement, fill = value)) +
+  geom_tile() +
+  scale_fill_gradient(low="black", high="white") +
+  labs(x="regressors", y="scans", title="(A)") +
+  # design
+  scale_x_discrete(expand = c(0,0),labels=colnames(dmf)[4:ncol(dmf)]) +
+  scale_y_continuous(expand = c(0,0)) +
+  theme_bw() + theme(legend.position = "none",
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12))
+# + theme(panel.grid = element_blank(),
+#         panel.border = element_rect(colour = "black", size=1),
+#    legend.text = element_text(size = 10)
+#         plot.title=element_text(size=11))
+plotB <- ggplot(input_dmc, aes(x = regressor, y = measurement, fill = value)) +
+  geom_tile() +
+  scale_fill_gradient(low="black", high="white") +
+  labs(x="regressors", y="scans", title="(B)") +
+  # design
+  scale_x_discrete(expand = c(0,0),labels=colnames(dmc)[4:ncol(dmc)]) +
+  scale_y_continuous(expand = c(0,0)) +
+  theme_bw() + theme(legend.position = "none",
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12))
+# + theme(panel.grid = element_blank(),
+#         panel.border = element_rect(colour = "black", size=1),
+#    legend.text = element_text(size = 10)
+#         plot.title=element_text(size=11))
+
+gridExtra::grid.arrange(plotA, plotB, nrow=1)
+```
 For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
-With a model containing average BMI and longitudinal change per subject, we investigated potential effects ..............
-on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- Nuisance covariates --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+With a model containing average BMI and longitudinal change per subject, we investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
+<!-- NUISANCE COVARIATES --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Age and sex were between-subject covariates while FD is time-varying.
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-As change in FD and change in BMI share variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
-<!-- Post hoc analyses  --> For that reason, we performed each analysis separately one with mean FD and without.
+<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
+As change in FD and change in BMI share common variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
+For that reason, we performed each analysis separately one with mean FD (three and without.
 Differences in results will be reported.
-For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
+We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).
+<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
 These post-hoc analyses deviate from the preregistration and should be considered to be exploratory.
-
-<!-- Inference --> All parametric inferences on activation were drawn based on False Discovery Rate (FDR) adjusted p-values $\leq 0.05$.
-To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for voxelwise inference, p-values were FDR-corrected for $\alpha = 5%$.
+<!-- INFERENCE --> <!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
+Significant clusters are defined as family-wise error - corrected p<0.05.
+-->
+To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
 In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and is unlike permutation based framework appropriate <!-- ???
-clarify why--> for longitudinal data.
+random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
+Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment (instead of $p < 0.01), and $1000$ bootstraps due to required computation time (instead of 5000).
+Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
 
@@
 ### DMN/reward network with whole brain mask (preregistered analysis)
-<!-- Brain mask: Group*Time Interaction -->
-#### without adjustment
+
+#### with adjustment for age, sex, meanFD
+<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
 There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
-Nor was there a significant main effect for any of the covariates of interest (time, group) in voxelwise inference with FDR correction.
-Results obtained for parametric estimation and non-parametric wild bootstrap did not differ.
-#### adjusting for age, sex, meanFD
-When reducing the number of nuisance covariates by leaving out mean FD, connectivity was found between the PCC with one other voxel.
-In the connectivity maps only corrected with AROMA-CC co-activation of voxel $x,y,z = 18,42,30$ [mm] yielded $Z = 4.72$ ($q_{FDRcorr} = 0.040$).
-However, verification of the effect with non-parametric bootstrapping failed to yield significant ($p_{\mathrm{FWEcorr}} = 0.061$; $q_{FDRcorr} = 0.999$)<!-- but FWE-corr sig?!, ich bin mir nicht ganz sicher, ob wir das überhaupt erwähnen wollen ...
--->.
-There was no other significant co-activation for any other region of interest for neither of the preprocessing variants.
+Nor was there a significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+
+#### with adjustment for age and sex
+Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD, no connectivity was found between the PCC with one other voxel.
+Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
+<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
 
@@
 ### DMN/reward network with GM mask (exploratory analysis)
-<!-- GM mask: Group*Time Interaction -->
+<!-- GM mask: Group*Time Interaction [OLD]-->
 Parametric and non-parametric significance test for the model testing for group-time-interaction did not show any different results when using a grey matter mask.
@@
 
-<!-- Brain mask: average and individual BMI -->
+<!-- Brain mask: average and individual BMI [OLD] -->
 <!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
@@
 Ventures to verify effects by re-calculating with non-parametric Wild bootstrap confirmed supported presumption as effects were not robust across estimation procedures.
+<!-- limitations -->
+Without an a priori power analysis it is not possible to determine if there is no effect or
+To clarify, extended considerations should be made to use a priori power analysis (e.g. with fMRIpower) or check evidence of absence with alternative
 

### 2020-12-14 manuscript/draft_rsfmri.Rmd
merged changes by Frauke

@@
     toc: yes
+    extra_dependencies: "subfig"
 bibliography: "bibliography.bib"
@@
 #library(haven)
+
+library(lme4)
+library(sjPlot)
+
 library(knitr) # version 1.30
@@
 ```
-```{r tabsample, echo=FALSE}
+
+```{r tableSample, echo=FALSE}
 tab_sample <-
@@
 ADI009_fu2 nicht enthalten (siehe Readme preprocessing) -->
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tabdescr)).
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tableDescr)).
 `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
 The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing only subjects that have received MRI.
-Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tabsample)).
+Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tableSample)).
 
@@
 
-```{r baseline characteristics, include=FALSE}
+```{r tableBaselineCharacteristics, include=FALSE}
 dfBL <- create_sample_df(group = "all", tp = "BL")
@@
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
-```{r tabdescr, echo=FALSE}
+```{r tableDescr, echo=FALSE}
 
@@
 ### Quality Assessment
+
 #### Anatomical Quality control
+
 FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
+To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and time points, adjusted for age and sex.
 We used the modified SwE option and used global signal regressed data as input.
-Network masks were formed from all voxels with a clusterwise $p_FWE < 0.05$ and we extracted the average connectivity from these masks for all data points.
+Network masks were formed from all voxels with a clusterwise $p_{FWE} < 0.05$ and we extracted the average connectivity from these masks for all data points.
 
@@
 <!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
---> For the first model, time was introduced as factor because we did not assume strict linearity for the increase in connectivity and therefore refrained from modeling time as continuous variable (see Figure \@ref(fig:figdesignmatrix)).
+--> For the first model, time was introduced as factor because we did not assume strict linearity for the increase in connectivity and therefore refrained from modeling time as continuous variable (see Figure \@ref(fig:figDesignmatrix)).
 The resulting factorial design <!-- check: flexible factorial?
 --> contained one regressor for each time point per group.
-```{r figdesignmatrix,  out.width="80%", fig.align = 'center', fig.cap="Possible design matrices of group-time model.", echo=FALSE}
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of group-time model.", echo=FALSE}
 # fig.show="hold",  out.width="40%", fig.height = 8,
+
 # design matrix time as factor
@@
   scale_fill_gradient(low="black", high="white") +
-  labs(x="regressors", y="scans", title="(A)") +
+  labs(x="regressors", y="scans") +
   # design
@@
   scale_fill_gradient(low="black", high="white") +
-  labs(x="regressors", y="scans", title="(B)") +
+  labs(x="regressors", y="scans") +
   # design
@@
 
-gridExtra::grid.arrange(plotA, plotB, nrow=1)
+plotA
+plotB
+#gridExtra::grid.arrange(plotA, plotB, nrow=1)
 ```
@@
 random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment (instead of $p < 0.01), and $1000$ bootstraps due to required computation time (instead of 5000).
+Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
@@
 We performed different quality controls for the rsfMRI data.
-As described in \@ref[Beyer_2020], there was a significant interaction of group and time on mFD.
+As described in @Beyer_2020, there was a significant interaction of group and time on mFD.
 In particular, the intervention group reduced in head motion in the follow-up assessments compared to the control group.
@@
 This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
-```{r summary table of QA measures}
-for (i in 1:nrow(final))
-  {tmp=strsplit(toString(final[i,"scan_dir"]),'/')[[1]][8]
-   conf=read.csv(paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",tmp,"/confounds.csv"))
-   tmp=cor.test(conf$DVARS,conf$FD)
-   final[i,"corr_DVARS_FD"]=tmp$estimate
-   final[i,"mean_DVARS"]=mean(conf$DVARS,na.rm = TRUE)
-   final[i,"max_DVARS"]=max(conf$DVARS,na.rm = TRUE)
-   final[i,"mean_outliers"]=mean(conf$outliers,na.rm = TRUE)
+```{r tableQA, fig.cap="Summary table of QA measures"}
+for (i in 1:nrow(final)) {
+  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
+  conf = read.csv(
+    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",tmp,"/confounds.csv")
+  )
+  tmp = cor.test(conf$DVARS, conf$FD)
+  final[i, "corr_DVARS_FD"] = tmp$estimate
+  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
+  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
+  final[i, "mean_outliers"] = mean(conf$outliers, na.rm = TRUE)
 }
@@
 
-```{r fig_summary_dvars_mFD, fig.align='center', out.width="100%"}
+```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD'}
+# ,  fig.align='center', }
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
@@
 names(condition.labs) <- c("IG","KG")
-# BMI over time each group
+
+# DVARS over time each group
 fig_DVARSdescr <-
@@
   )
-
+# mFD over time each group
 fig_mFDdescr <-
@@
   )
+# DVARS-FD over time each group
 fig_mFDDVARSdescr <-
@@
 
-(fig1 | fig2 ) /
-      fig3
-```
-
-```{r FDFC correlation, include=FALSE}
+(fig1 | fig2) / fig3
+
+```
+
+```{r FDFC correlation, eval=FALSE, echo=FALSE}
 source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
 
-res=calc_qc_fc(final)
-```
+FDFC=calc_qc_fc(final)
+write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
+```
+
+```{r tableQcfc, echo=FALSE}
+res <- read.csv("./report/FDFCcorrelations.csv")
+rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
+colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
+
+tab_qc_fc <-
+  knitr::kable(
+    round(res,3),
+    col.names = colnames(res),
+    row.names = TRUE,
+    format = "latex",
+    booktabs = T,
+    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
+
+tab_qc_fc
+```
+
 Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:table_qc_fc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
@@
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
-```{r table_qc_fc, echo=FALSE}
-tab_qc_fc <-
-  knitr::kable(
-    round(res,3),
-    #row.names = c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR"),
-    #col.names = c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval'),
-    format = "latex",
-    booktabs = T,
-    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
-
-tab_qc_fc
-```
 
@@
 ### DMN for two timepoints
-```{r calc DMN two tp, include=FALSE}
-library(lme4)
-library(sjPlot)
-R1= lmer(DMN_FC~ timepoint*group + (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + (1|subj))
+```{r calc DMN two tp, eval = FALSE, echo = FALSE}
+R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
 ```
 ### DMN for two timepoints, adjusted for age, sex, mean FD
-```{r calc DMN two tp adjusted, include=FALSE}
-library(lme4)
-library(sjPlot)
-R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
+R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
 ```
 ### DMN for three timepoints, adjusted for age, sex, mean FD
-```{r calc DMN three tp, include=FALSE}
-library(lme4)
-library(sjPlot)
-R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group ++ age + sex + meanFD (1|subj))
+```{r calc DMN three tp, eval = FALSE, echo = FALSE}
+R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group ++ age + sex + meanFD (1|subj))
 ```
@@
 ### reward
-```{r calc rew two tp, include=FALSE}
-R1= lmer(rew_FC ~ timepoint*group + (1|subj))
-R0 = lmer(rew_FC ~ timepoint + group + (1|subj))
+```{r calc rew two tp, eval = FALSE, echo = FALSE}
+R1 = lme4::lmer(rew_FC ~ timepoint*group + (1|subj))
+R0 = lme4::lmer(rew_FC ~ timepoint + group + (1|subj))
 ```
 ### reward for two timepoints, adjusted for age, sex, mean FD
-```{r calc rew two tp adjusted, include=FALSE}
-library(lme4)
-library(sjPlot)
-R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
+```{r calc rew two tp adjusted, eval = FALSE, echo = FALSE}
+R1 = lme4::lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
 ```
 ### reward for three timepoints, adjusted for age, sex, mean FD
-```{r calc rew three tp, include=FALSE}
-library(lme4)
-library(sjPlot)
-R1= lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+```{r calc rew three tp, eval = FALSE, echo = FALSE}
+R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
 R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))

### 2020-12-14 code/create_sample_df.R

@@ -11,12 +11,13 @@ create_sample_df <- function(group = "all", tp = "all", exclFD = FALSE){
   # ----------------------------------------------------------------------------
 
   apply_excl_criteria <- function(data = full_sample, exclFD = FALSE){
-    # exclusion due to problems in preprocessing
-    cat("N=",nrow(full_sample),"\n")
+    # exclusion of participants who failed preprocessing in FreeSurfer
+    cat("N =",nrow(full_sample),"\n")
     full_sample$exclude_prep <- full_sample$Exclude
     full_sample$exclude_prep[is.na(full_sample$exclude_prep)]=FALSE
+    
     mri_sample <- merge(mri_files, full_sample[!is.na(full_sample$condition),c("subj.ID","condition","tp","Age_BL","Sex","meanFD", "BMI","exclude_prep")], by=c("subj.ID","tp"))
-    cat("n=",nrow(mri_sample),"with fMRI data points.\n")
+    cat("n =",nrow(mri_sample),"with fMRI data points.\n")
 
     mri_sample[mri_sample$exclude_prep==TRUE,] # ADI063_bl ADI063_fu ADI116_bl ADI116_fu ADI116_fu2
     cat("Note that",nrow(mri_sample[mri_sample$exclude_prep==TRUE,]),"data points were excluded due to problems in preprocessing.\n")
@@ -34,7 +35,7 @@ create_sample_df <- function(group = "all", tp = "all", exclFD = FALSE){
     }else if (exclFD ==FALSE){
       final_sample=sample_after_1stExcl
     }
-    cat("The remaining sample comprises",nrow(final_sample),"data points across across time for all groups.\n")
+    cat("The remaining sample comprises n =",nrow(final_sample),"data points across across time for all groups.\n")
     # all remaining data points should be included
     final_sample$include <- TRUE
     return(final_sample)
@@ -70,9 +71,6 @@ create_sample_df <- function(group = "all", tp = "all", exclFD = FALSE){
   df=df[!is.na(df$include),]
   rm(apply_excl_criteria, mri_files, full_sample, final_sample, tmp, i)
 
-  # exclude participants who failed in FreeSurfer (Frauke)
-  df$Exclude=is.na(df$Exclude)
-  df=df[df$Exclude==TRUE,]
   # preparation of variables --------------------------
 
   df$logmFD <- log10(df$meanFD)

### 2020-12-14 code/run_or_display.m

@@ -100,14 +100,14 @@ roi_prep = readcell(fullfile(param.INFO_DIR,'ROIs.txt'), 'Delimiter',' ','Whites
 
 %% ------------------------------------------------------------------------
 % define ROI
-param.ROI_PREP = {roi_prep{[6,14]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'}
+param.ROI_PREP = {roi_prep{[14]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'}
 
 % Model definition
 % All three models have unique options for covariate definition, the
 % association to a model is indicated by the tens digit (GroupTime_: 1_; 
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
-param.MODEL = {'alltp'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [42];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % [41, 42, 43]
+param.MODEL = {'bmi'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
+param.COVARIATES = [21];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % [41, 42, 43]
 
 % define masking and type of inference
 param.MASK = 'brain';               % 'brain'
@@ -117,7 +117,7 @@ param.INFERENCE_TYPE = {'cluster'}; % {'voxel','cluster','tfce'};
 
 % analysis parameter (estimate or display?)
 param.ONLY_DISPLAY = true;         % false
-param.OVERWRITE = false;             % false
+param.OVERWRITE = true;             % false
 param.VIEWSEC = 1; % for ONLY_DISPLAY: seconds you want to view the results
 % param.ACTION = 'estimate' % 'display' or 'overwrite'
 

### 2020-12-15 manuscript/draft_rsfmri.Rmd
minor changes in manuscript

@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and time points, adjusted for age and sex.
-We used the modified SwE option and used global signal regressed data as input.
-Network masks were formed from all voxels with a clusterwise $p_{FWE} < 0.05$ and we extracted the average connectivity from these masks for all data points.
+To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
+We used global signal regressed data as input and clusterwise bootstrapping with N=999.
+Network masks were formed from all voxels with a clusterwise p_FWE < 0.05 and we extracted the average connectivity from these masks for all data points.
 
@@
 We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
-These were examined each for correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $16$ a priori planned analyses.
+The model was examined for the correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $8$ a priori planned analyses.
+
 Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$.
@@
 write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
+
 ```
@@
 <!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction will be published on NeuroVault.
 
@@
 ### DMN for two timepoints
+
 ```{r calc DMN two tp, eval = FALSE, echo = FALSE}
@@
 R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
 ```

### 2020-12-15 manuscript/draft_rsfmri.Rmd
minor changes in manuscript

@@
 # [ ] Discussion
-# [ ] hervorheben, was von Präregistrierung abweicht
 # [ ] brain images mit nilearn plotting tools einbinden
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] re-do analysis with new sample
+# [x] re-do analysis with new sample
 # [ ] repeat analysis with exclusion of 10% of worst data
@@
 options(knitr.table.format = "latex") 
+options(knitr.kable.NA = '') # hide NA in tables
 # set working directory of knit (RMarkdown) to directory of Rproj
@@
 
-```{r Load data, include=FALSE}
-source("code/create_sample_df.R")
+```{r "Load data", include=FALSE}
+source("report/create_sample_df.R")
 final = create_sample_df(group = "all", tp = "all")
@@
 
-# overview distribution of data points
-tp1 <- as.character(freqtable$Var1[freqtable$Freq == 1])
-tp2 <- as.character(freqtable$Var1[freqtable$Freq == 2])
-tp3 <- as.character(freqtable$Var1[freqtable$Freq == 3])
-tp1_tab <- table(subset(final, subj.ID %in% tp1)$tp, subset(final, subj.ID %in% tp1)$condition)
-tp2_tab <- table(subset(final, subj.ID %in% tp2)$tp, subset(final, subj.ID %in% tp2)$condition)[c(1,3),]
-tp3_tab <- table(subset(final, subj.ID %in% tp3)$tp, subset(final, subj.ID %in% tp3)$condition)[1,]
-total_datapoints <- c(sum(final$group==1),sum(final$group==2))
-total_subjects <- plyr::count(final[subj_idx,"group"])$freq
-tab_sample <- data.frame(rbind(tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints))
-rm(tp1,tp2,tp3,tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints)
-rownames(tab_sample) <- c("count: only 0","count: only 6","count: only 12","count: 0 and 6","count: 6 and 12","count: complete data","total number of subjects","total data points")
-colnames(tab_sample) <- c("BARS","NBARS")
-
 # distribution of condition and sex
@@
 
-```{r tableSample, echo=FALSE}
+```{r "tableSample", echo=FALSE}
+source("./report/tables.R")
+tab_sample <- mk_SampleTable(final)
+
 tab_sample <-
@@
     booktabs = T,
+    linesep = "",
     caption = "Distribution of data points at months after intervention")
@@
 Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
-<!-- Frage: Ausschluss bei head movement?
-ADI009_fu2 nicht enthalten (siehe Readme preprocessing) -->
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tableDescr)).
@@
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
-
 ```{r tableBaselineCharacteristics, include=FALSE}
@@
 ```
-Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
 `r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-Change in BMI throuout the study is depicted in \@ref(fig:figBMIdescr).
+Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
@@
 
+# table_descr <- mk_tableDescr()
 # table for descriptive statistics
@@
     booktabs = T,
+    linesep = "",
     caption = "BMI and mean FD of participants acquired in each condition and time point")
@@
 ```
-<!-- STUDY PROCEDURE -->
-Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for STANDARDIZED??
-breakfast.
+<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for <!-- standadized??
+-->breakfast.
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
-
 ```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-condition.labs <- c("BARS","NBARS")
-names(condition.labs) <- c("IG","KG")
-# BMI over time each group
-figBMIdescr <-
-  ggplot(final, aes(
-    x = tp,
-    y = BMI,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-figBMIdescr +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
+source("report/figures.R")
+figBMIdescr <- mk_figBMIdescr(final)
+
 ```
@@
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-In addition, for models showing a significant group-by-time interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
+In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
 On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
@@
 All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
-We evaluated mFD-FC relationships by computing a functional connectome based on (Power et al., 2011).
+We evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012].
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
@@
 
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in SPM12.7770<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
 During analyses an implicit mask and an explicit mask were used.
@@
 --> contained one regressor for each time point per group.
-```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of group-time model.", echo=FALSE}
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
 # fig.show="hold",  out.width="40%", fig.height = 8,
+source("report/figures.R")
+PlotList <- mk_figDesignMatrix()
 
-# design matrix time as factor
-txt_path <- "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files/noExclFD/both/total/"
-
-ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
-condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
-KG_bl <- read.delim(paste0(txt_path,"KG_bl.txt"), head=FALSE)
-KG_fu <- read.delim(paste0(txt_path,"KG_fu.txt"), head=FALSE)
-KG_fu2 <- read.delim(paste0(txt_path,"KG_fu2.txt"), head=FALSE)
-IG_bl <- read.delim(paste0(txt_path,"IG_bl.txt"), head=FALSE)
-IG_fu <- read.delim(paste0(txt_path,"IG_fu.txt"), head=FALSE)
-IG_fu2 <- read.delim(paste0(txt_path,"IG_fu2.txt"), head=FALSE)
-measurement <- c(1:nrow(ID))
-dmf <- data.frame(measurement,ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
-colnames(dmf) <- c("measurement","ID","condition","N0","N6","N12","B0","B6","B12")
-
-# design matrix time as continuous variable
-ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
-condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
-group_IG <- read.delim(paste0(txt_path,"group_IG.txt"), head=FALSE)
-group_KG <- read.delim(paste0(txt_path,"group_KG.txt"), head=FALSE)
-tp_IG <- read.delim(paste0(txt_path,"tp_IG.txt"), head=FALSE)
-tp_KG <- read.delim(paste0(txt_path,"tp_KG.txt"), head=FALSE)
-measurement <- c(1:nrow(ID))
-dmc <- data.frame(measurement,ID,condition,group_IG,group_KG,tp_IG,tp_KG)
-colnames(dmc) <- c("measurement","ID","condition","groupB", "groupN","timeB","timeN")
-
-# plot dmf
-input_dmf <- dmf %>%
-  pivot_longer(
-    cols= colnames(dmf[,c(4:ncol(dmf))]),
-    names_to = "regressor")
-input_dmc <- dmc %>%
-  pivot_longer(
-    cols= colnames(dmc[,c(4:ncol(dmc))]),
-    names_to = "regressor")
-
-plotA <- ggplot(input_dmf, aes(x = regressor, y = measurement, fill = value)) +
-  geom_tile() +
-  scale_fill_gradient(low="black", high="white") +
-  labs(x="regressors", y="scans") +
-  # design
-  scale_x_discrete(expand = c(0,0),labels=colnames(dmf)[4:ncol(dmf)]) +
-  scale_y_continuous(expand = c(0,0)) +
-  theme_bw() + theme(legend.position = "none",
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12))
-# + theme(panel.grid = element_blank(),
-#         panel.border = element_rect(colour = "black", size=1),
-#    legend.text = element_text(size = 10)
-#         plot.title=element_text(size=11))
-plotB <- ggplot(input_dmc, aes(x = regressor, y = measurement, fill = value)) +
-  geom_tile() +
-  scale_fill_gradient(low="black", high="white") +
-  labs(x="regressors", y="scans") +
-  # design
-  scale_x_discrete(expand = c(0,0),labels=colnames(dmc)[4:ncol(dmc)]) +
-  scale_y_continuous(expand = c(0,0)) +
-  theme_bw() + theme(legend.position = "none",
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12))
-# + theme(panel.grid = element_blank(),
-#         panel.border = element_rect(colour = "black", size=1),
-#    legend.text = element_text(size = 10)
-#         plot.title=element_text(size=11))
-
-plotA
-plotB
+PlotList$plotA
+PlotList$plotB
 #gridExtra::grid.arrange(plotA, plotB, nrow=1)
@@
 With a model containing average BMI and longitudinal change per subject, we investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- NUISANCE COVARIATES --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+<!-- NUISANCE COVARIATES --> Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Age and sex were between-subject covariates while FD is time-varying.
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
 As change in FD and change in BMI share common variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
@@
 Differences in results will be reported.
-We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).
-<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
-These post-hoc analyses deviate from the preregistration and should be considered to be exploratory.
-<!-- INFERENCE --> <!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
-We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
-Significant clusters are defined as family-wise error - corrected p<0.05.
--->
-To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
+<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
+As these post-hoc analyses were not preregistered, they should be considered to be exploratory.
+
+<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
 Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual. 
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
+Significant clusters are defined as family-wise error - corrected p<0.05.
+-->
 
 <!-- R analysis of aggregated values -->
-We analyzed the aggregate connectivity values using linear mixed models in the R package lme4.
-Following the preregistration, we first investigated a time-by-group interaction between baseline and followup.
+We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
+Following the preregistration, we first investigated a time-by-group interaction between baseline and follow-up.
 We used a model comparison of R1 and R0 models with `anova`:
@@
 We performed a sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI.
-Again, the group-by-time point interaction was considered significant if a model comparison between R1 and Null models using anova-function yielded a p<0.05.
+Again, the time-by-group point interaction was considered significant if a model comparison between R1 and Null models using anova-function yielded a p<0.05.
 Finally, we estimated the above-mentioned interaction model for all three time points.
@@
 We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data.
-We saw that DVARS did not qualitatively differ between groups and timepoints, and that the correlation of DVARS and mFD remained similar over time points.
+We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
 This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
 ```{r tableQA, fig.cap="Summary table of QA measures"}
+
 for (i in 1:nrow(final)) {
@@
   conf = read.csv(
-    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",tmp,"/confounds.csv")
-  )
+    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",
+      tmp, "/confounds.csv"))
   tmp = cor.test(conf$DVARS, conf$FD)
@@
   final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
-  final[i, "mean_outliers"] = mean(conf$outliers, na.rm = TRUE)
+  final
 }
@@
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-condition.labs <- c("BARS","NBARS")
-names(condition.labs) <- c("IG","KG")
 
-# DVARS over time each group
-fig_DVARSdescr <-
-  ggplot(final, aes(
-    x = tp,
-    y = mean_DVARS,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig1=fig_DVARSdescr +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time [months]") + ylab ("Mean DVARS") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
-# mFD over time each group
-fig_mFDdescr <-
-  ggplot(final, aes(
-    x = tp,
-    y = meanFD,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig2=fig_mFDdescr +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time [months]") + ylab ("mFD in mm") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
-# DVARS-FD over time each group
-fig_mFDDVARSdescr <-
-  ggplot(final, aes(
-    x = tp,
-    y = corr_DVARS_FD,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig3=fig_mFDDVARSdescr +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time [months]") + ylab ("DVARS-FD correlation") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
-
-(fig1 | fig2) / fig3
+FigureList <- figDvarsmFD(final)
+(FigureList$fig1 | FigureList$fig2) / FigureList$fig3
 
@@
 
-tab_qc_fc <-
-  knitr::kable(
-    round(res,3),
-    col.names = colnames(res),
-    row.names = TRUE,
-    format = "latex",
-    booktabs = T,
-    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
-
+tab_qc_fc <- mk_tableQcfc(res)
 tab_qc_fc
@@
 Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's $r = $`r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's $r = $`r round(res[1,5],2)`).
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
-AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
+AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r = $`r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r = $`r round(res[4,5],2)`).
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
@@
 #### with adjustment for age and sex
-Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD, no connectivity was found between the PCC with one other voxel.
+Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
 Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
@@
 <!-- GM mask: Group*Time Interaction [OLD]-->
-Parametric and non-parametric significance test for the model testing for group-time-interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the group-time-interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
+Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
+Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
 While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->.
@@
 R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group ++ age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
 ```
@@
 
+```{r tableBMImodel}
+source("report/tables.R")
+TableList <- mk_FCTables()
+tab_BMImodel <- TableList[[1]]
+tab_BMImodel
+```
+
+```{r tableFDmodel, echo=FALSE}
+tab_FDmodel <- TableList[[2]]
+tab_FDmodel
+```
+
+```{r tableBMIFDmodel, echo=FALSE}
+tab_BMIFDmodel <- TableList[[3]]
+tab_BMIFDmodel
+```
+
 <!-- Brain mask: average and individual BMI [OLD] -->
 <!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
-There was no significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
+We did not find a significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
 Omitting mean FD as nuisance covariate did not change these results.
@@
 Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no significant co-activation was found when using the grey matter mask.
+In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
 
@@
 <!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
-Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
+Surprisingly, we did not find a significant effect neither for the total sample, nor for the intervention group only.
 <!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
@@
 ## Effect on hypothalamic connectivity (exploratory analysis)
-
-<!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
-Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
-<!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
 
@@
 Over the course of one year we collected neuroimaging and anthropometric data from `r cs["Sum","IG"]` morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of  `r cs["Sum","KG"]` patients on the waiting list.
-<!-- model specific --> With the model on group-time-interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
+<!-- model specific --> With the model on time-by-group--interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
 Beyond that, also main effects of the intervention and of the time were examined.
@@
 @McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
-However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the group-time-interaction with age and sex as nuisance covariates.
+However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the time-by-group--interaction with age and sex as nuisance covariates.
 <!-- (FD is time-varying...
@@
 -->
-An advantage of the group-time interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
+An advantage of the time-by-group interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
 The model was set up as flexible factorial design with factors "subject", "group", and "time" assuming unequal group variances.
@@
 # References
-
-```{r Exploratory plots, eval=FALSE}
-# BMI over time per group
-tspag <-
-  ggplot(final, aes(x = tp, y = BMI)) +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  geom_line(aes(group = subj.ID, color = condition)) +
-  geom_point(aes(color = condition))
-tspag +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "bottom",
-    legend.background = element_rect(colour = "grey")
-  )
-
-# BMI over time per subject
-# group x time interaction for BMI
-tspag <- ggplot(final, aes(
-  x = tp,
-  y = BMI,
-  color = condition,
-  group = condition
-)) +
-  geom_point() +
-
-  stat_summary(fun = mean, geom = "point") +
-  stat_summary(fun = mean, geom = "line")
-tspag  +
-  scale_color_manual(values = c("#00305E", "#C00045")) +
-  theme_bw() +
-  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.position = "top"
-)
-```
-
-```{r expample plots, eval=FALSE}
-
-df_wide <- tidyr::pivot_wider(
-  data = final,
-  id_cols = "subj.ID",
-  names_from = "tp",
-  values_from = c("tp", "BMI", "logmFD")
-)
-
-if (tp == "BLFU") {
-  s = 2
-} else {
-  s = 1
-}
-
-BMI_vector <- c("BMI_fu2", "BMI_fu", "BMI_bl")
-BMI_vector <- c("BMI_bl", "BMI_fu", "BMI_fu2")
-FD_vector <- c("logmFD_fu2", "logmFD_fu", "logmFD_bl")
-df_wide$avgBMI <- apply(df_wide[, BMI_vector[s:3]],
-                        MARGIN = 1,
-                        FUN = mean,
-                        na.rm = TRUE)
-df_wide$avgFD <- apply(df_wide[, FD_vector[s:3]],
-                       MARGIN = 1,
-                       FUN = mean,
-                       na.rm = TRUE)
-df_long <- tidyr::pivot_longer(
-  data = df_wide,
-  cols = all_of(order_vector),
-  # bl, fu, fu2 (maintain original order)
-  # CAREFUL: Order according to dataframe df
-  names_to = "tp_tp",
-  values_to = "tp",
-  values_drop_na = TRUE
-)
-all(df_long$tp == final$tp) # check if order correct
-
-head(final[, c("subj.ID", "tp")])
-head(df_long[, c("subj.ID", "tp")])
-
-df_long$cgnBMI <- final$BMI - df_long$avgBMI
-df_long$avgBMIc <-
-  df_long$avgBMI - mean(df_long$avgBMI, na.rm = TRUE)
-df$avgBMIc <- df_long$avgBMIc
-df$cgnBMI <- df_long$cgnBMI
-
-df_long$cgnFD <- final$logmFD - df_long$avgFD
-df_long$avgFDc <-
-  df_long$avgFD - mean(df_long$avgFD, na.rm = TRUE)
-final$avgFDc <- df_long$avgFDc
-final$cgnFD <- df_long$cgnFD
-
-# example plots for interaction effect (arbitrary variables for illustration purposes)
-ggplot(final, aes(x=tp, y=BMI, group=condition, color=condition)) +
-  #geom_smooth(aes(group=condition)) +
-  stat_summary(fun = mean, geom = "point", size=3, shape=15) +
-  stat_summary(fun = mean, geom = "line", size=1) +
-  #scale_color_brewer(values = "Set1") +
-  scale_x_discrete(labels=c("bl"= "0", "fu"= "6", "fu2" = "12")) +
-  scale_color_manual(values=c("#00305E","#C00045"),labels = c("intervention", "control")) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10), axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12), legend.text = element_text(size = 10),
-    legend.position = "bottom", legend.background = element_rect(colour = "grey"))
-
-# example plot for change effect (purely descriptive)
-getPalette = colorRampPalette(brewer.pal(9, "Set1"))
-cgnplot <- ggplot(final, aes(x=a,y=logmFD, group=subj.ID)) +
-  geom_line(aes(x=BMI, color=subj.ID))+
-  geom_point(aes(x=BMI, color=subj.ID, shape=tp), size=3, alpha = 0.5)+
-  facet_grid(. ~ condition)
-cgnplot +
-  labs(shape="Time after intervention [months]",labels=c("0","6","12")) +
-  guides(color = FALSE, alpha=FALSE, size=FALSE) +
-  scale_color_manual(values = getPalette(length(unique(final$subj.ID)))) +
-  ylab("log(mean FD) (mm)") + xlab("BMI [kg/m²]") + xlim(30,55) +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10), axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12), legend.text = element_text(size = 10),
-    legend.position = "bottom", legend.background = element_rect(colour = "grey"))
-
-# example plot for change effect (purely descriptive)
-tspag=ggplot(data=final, aes(y=cgnFD, x=cgnBMI, color=condition))+
-  geom_point(size=2)+
-  #geom_line(aes(color=subj.ID))+
-  #geom_smooth(method = "lm", colour = "black", linetype = 1, fill = "mistyrose3" ) +
-  xlab("within-subject BMI change [kg/m²]")+
-  ylab("within-subject log mean FD change [mm]")+
-  guides(alpha=FALSE, size=FALSE,
-         color = guide_legend(override.aes = list(size=4))) +
-  facet_grid(. ~ condition)
-tspag + scale_color_manual(values=c("#00305E","#C00045")) +
-  theme_bw() + theme(axis.text=element_text(colour = "black",size=10),
-              axis.title=element_text(size=14),
-              axis.text.x = element_text(size=12),
-              axis.text.y = element_text(size=12),
-              strip.text = element_text(size=14),
-              legend.position="top",
-              legend.title = element_text(size = 14),
-              legend.text = element_text(size = 14))
-```

### 2020-12-17 manuscript/draft_rsfmri.Rmd
script to save results; remove unnessecary code from RunModel.. scripts

--- 2020-12-14 manuscript/draft_rsfmri.Rmd
@@
 ```
-<!-- STUDY PROCEDURE -->
-Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for STANDARDIZED??
-breakfast.
+<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for <!-- standadized??
+-->breakfast.
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
-
 ```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
@@
 --> contained one regressor for each time point per group.
-```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of group-time model.", echo=FALSE}
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
 # fig.show="hold",  out.width="40%", fig.height = 8,
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
+<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
+Significant clusters are defined as family-wise error - corrected p<0.05.
+-->
 
 <!-- R analysis of aggregated values -->
-We analyzed the aggregate connectivity values using linear mixed models in the R package lme4.
-Following the preregistration, we first investigated a time-by-group interaction between baseline and followup.
+We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
+Following the preregistration, we first investigated a time-by-group interaction between baseline and follow-up.
 We used a model comparison of R1 and R0 models with `anova`:
@@
 <!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
-There was no significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
+We did not find a significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
 Omitting mean FD as nuisance covariate did not change these results.

### 2020-12-17 code/RunModelBMI.m

@@ -66,76 +66,54 @@ else
 end
 
 nrun = length(param.ROI_PREP);
+% estimate any of the models for every roi_prep in ROI_PREP
 for crun = 1:nrun
-    % estimate any of the six models for every roi_prep in ROI_PREP
-    if not(param.WILD_BOOT)
+    if param.WILD_BOOT
+        % wild bootstrap
+        param.wild_con = 1%:2;
+    else
         % parametric estimation
         param.wild_con = false;
-        % check if analysis has been conducted already
+    end
+    for i = param.wild_con
+        param.wild_con = i;
+        % check if analysis has run already
         [out_folder, exist_already] = create_out_folder(param, crun);
-        fprintf('Output folder: \n %s ...\n',out_folder)
-        clear matlabbatch
+        fprintf('If necessary create output folder \n %s ...\n', out_folder)
+          clear matlabbatch
         fprintf('Specify Model...\n')
-        [matlabbatch,location_SwE_mat] = SpecifyModel(param, crun, out_folder);
+          [matlabbatch, location_SwE_mat] = SpecifyModel(param, crun, out_folder);
         if param.ONLY_DISPLAY %strcmp(param.ACTION,'display')
             fprintf('Display Results...\n')
-            spm('defaults', 'FMRI'); 
-            matlabbatch = DisplayResults(location_SwE_mat);
-            spm_jobman('run', matlabbatch);
+            %spm('defaults', 'FMRI');
+            %matlabbatch = DisplayResults(location_SwE_mat);
+            %spm_jobman('run', matlabbatch);
+            cd(out_folder)
+            [hReg,xSwE,SwE] = swe_results_ui("Setup");
+            spm2csv(hReg,xSwE);
             pause(param.VIEWSEC)
-        elseif not(exist_already) || param.OVERWRITE 
+        elseif not(exist_already) || param.OVERWRITE %strcmp(param.ACTION,'overwrite')
             fprintf('Estimate model...\n')
-            % delete existing files in folder if existent 
-            if exist_already 
-                rmdir(out_folder,'s'); % delete former dir
+              % delete existing files in folder if existent
+            if exist_already
+                rmdir(out_folder, 's'); % delete former dir
                 mkdir(out_folder); % create new empty one
             end
-            spm('defaults', 'FMRI'); 
+            spm('defaults', 'FMRI');
             matlabbatch = RunModel(matlabbatch, location_SwE_mat);
             spm_jobman('run', matlabbatch);
         else
[…]

### 2020-12-17 code/RunModelFD.m

@@ -67,69 +67,57 @@ end
 
 nrun = length(param.ROI_PREP);
 for crun = 1:nrun
-    % estimate any of the six models for every roi_prep in ROI_PREP
-    if not(param.WILD_BOOT)
+    if param.WILD_BOOT
+        % wild bootstrap
+        param.wild_con = 1:4;
+    else
         % parametric estimation
         param.wild_con = false;
-        % check if analysis has been conducted already
+    end
+    for i = param.wild_con
+        param.wild_con = i;
+        if (param.wild_con > 2 && param.COVARIATES == 32)
+            continue;
+        end
+        % check if analysis has run already
         [out_folder, exist_already] = create_out_folder(param, crun);
-        fprintf('Output folder: \n %s ...\n',out_folder)
+        fprintf('If necessary create output folder\n %s...\n', out_folder)
         clear matlabbatch
         fprintf('Specify Model...\n')
-        [matlabbatch,location_SwE_mat] = SpecifyModel(param, crun, out_folder);
+          [matlabbatch, location_SwE_mat] = SpecifyModel(param, crun, out_folder);
         if param.ONLY_DISPLAY %strcmp(param.ACTION,'display')
             fprintf('Display Results...\n')
-            spm('defaults', 'FMRI'); 
-            matlabbatch = DisplayResults(location_SwE_mat);
-            spm_jobman('run', matlabbatch);
+            %spm('defaults', 'FMRI');
+            %matlabbatch = DisplayResults(location_SwE_mat);
+            %spm_jobman('run', matlabbatch);
+            cd(out_folder)
+            [hReg,xSwE,SwE] = swe_results_ui("Setup");
+            spm2csv(hReg,xSwE);
             pause(param.VIEWSEC)
         elseif not(exist_already) || param.OVERWRITE %strcmp(param.ACTION,'overwrite')
             fprintf('Estimate model...\n')
             % delete existing files in folder if existent
-            if exist_already 
-                rmdir(out_folder,'s'); % delete former dir
+            if exist_already
+                rmdir(out_folder, 's'); % delete former dir
                 mkdir(out_folder); % create new empty one
             end
-            spm('defaults', 'FMRI'); 
+            spm('defaults', 'FMRI');
             matlabbatch = RunModel(matlabbatch, location_SwE_mat);
             spm_jobman('run', matlabbatch);
-        end
-        % manually type in contrasts (issue: see below)
-    else
-        % non-parametric estimation with Wild-Bootstrap
[…]

### 2020-12-17 code/RunModelGroupTime.m

@@ -98,76 +98,55 @@ else
 end
 
 nrun = length(param.ROI_PREP);
+% estimate any of the models for every roi_prep in ROI_PREP
 for crun = 1:nrun
-    % estimate any of the six models for every roi_prep in ROI_PREP
-    if not(param.WILD_BOOT)
+    if param.WILD_BOOT
+        % wild bootstrap
+        param.wild_con = 1:3;
+    else
         % parametric estimation
         param.wild_con = false;
-        % check if analysis has been conducted already
+    end
+    for i = param.wild_con
+        param.wild_con = i;
+        % check if analysis has run already
         [out_folder, exist_already] = create_out_folder(param, crun);
-        fprintf('If necessary create output folder \n %s ...\n',out_folder)
+        fprintf('Output folder: \n %s ...\n', out_folder)
         clear matlabbatch
         fprintf('Specify Model...\n')
-        [matlabbatch,location_SwE_mat] = SpecifyModel(param, crun, out_folder);
+          [matlabbatch, location_SwE_mat] = SpecifyModel(param, crun, out_folder);
         if param.ONLY_DISPLAY % strcmp(param.ACTION,'display')
             fprintf('Display Results...\n')
-            spm('defaults', 'FMRI'); 
-            matlabbatch = DisplayResults(location_SwE_mat);
-            spm_jobman('run', matlabbatch);
-            pause(param.VIEWSEC) 
-        elseif not(exist_already) || param.OVERWRITE % strcmp(param.ACTION,'overwrite')
+            %spm('defaults', 'FMRI');
+            %matlabbatch = DisplayResults(location_SwE_mat);
+            %spm_jobman('run', matlabbatch);
+            cd(out_folder)
+            [hReg,xSwE,SwE] = swe_results_ui("Setup"); % calls SwE Gui directly
+            spm2csv(hReg,xSwE);
+            pause(param.VIEWSEC)
+        elseif not(exist_already) || param.OVERWRITE % else
             fprintf('Estimate model...\n')
             % delete existing files in folder if existent
             if exist_already
-                rmdir(out_folder,'s'); % delete former dir
+                rmdir(out_folder, 's'); % delete former dir
                 mkdir(out_folder); % create new empty one
             end
-            spm('defaults', 'FMRI'); 
+            spm('defaults', 'FMRI');
             matlabbatch = RunModel(matlabbatch, location_SwE_mat);
             spm_jobman('run', matlabbatch);
         else
             fprintf('... model already estimated or error in parameter definitions.\n')
[…]

### 2020-12-17 code/run_or_display.m

@@ -83,6 +83,7 @@ addpath('/data/pt_02161/Analysis/Software/spm12/toolbox/SwE-toolbox-2.2.1-129202
 % swe version 2.1.1
 %addpath(genpath('/data/pt_life/data_fbeyer/spm-fbeyer'))
 
+clear variables
 % import absolute path for project
 % SET PATH TO code_and_manuscript FOLDER AS CURRENT DIRECTORY !!
 cd("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code")
@@ -100,25 +101,25 @@ roi_prep = readcell(fullfile(param.INFO_DIR,'ROIs.txt'), 'Delimiter',' ','Whites
 
 %% ------------------------------------------------------------------------
 % define ROI
-param.ROI_PREP = {roi_prep{[14]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'}
+param.ROI_PREP = {roi_prep{[12]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'}
 
 % Model definition
 % All three models have unique options for covariate definition, the
 % association to a model is indicated by the tens digit (GroupTime_: 1_; 
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
 param.MODEL = {'bmi'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [21];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % [41, 42, 43]
+param.COVARIATES = [22];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
 
 % define masking and type of inference
 param.MASK = 'brain';               % 'brain'
 param.EXCLFD = false;               % false
-param.WILD_BOOT = false;            % false
+param.WILD_BOOT = true;            % false
[…]

### 2020-12-17 code/spm2csv.m

@@ -1,0 +1,47 @@
+% script that prints the results of an SPM contrats (obtained via the GUI on SPM)
+% to a csv file
+function [result_dir] = spm2csv(hReg,xSwE)
+TabDat = swe_list('List',xSwE,hReg);
+
+%% Print a csv files of the results
+% if there are any results
+if size(TabDat.dat,1) > 0
+    myfile = 'results.csv';
+    fid = fopen (myfile, 'w');
+    %%
+    % print header
+    for j=1:size(TabDat.hdr,2)
+        fprintf(fid, '%s', TabDat.hdr{1,j});
+        if strcmp(TabDat.hdr{1,11},' ') fprintf(fid, '_'); end
+        fprintf(fid, '%s,', TabDat.hdr{2,j});
+    end
+    fprintf(fid,'\n');
+    %%
+    % print data
+    for i=1:size(TabDat.dat,1)
+        for j=1:size(TabDat.dat,2)
+            fprintf(fid, '%6.3f,', TabDat.dat{i,j});
+        end
+        fprintf(fid, '\n');
+    end
+    fclose(fid);
+end
+%%
+myfile2 = 'info.csv';
+fid2 = fopen (myfile2, 'w');
+% print info text
+currentfolder = pwd;
+fprintf(fid2, '%s,', currentfolder(69:end));
+fprintf(fid2, '%s,', xSwE.title);
+fprintf(fid2, '%s', TabDat.tit);
+fprintf(fid2, '\n');
+for i = 1:size(TabDat.ftr,1)
+    fprintf(fid2, TabDat.ftr{i,1},TabDat.ftr{i,2});
+    fprintf(fid2, '\n\n');
+end
+
+print('...done.')
+fclose(fid2);
+
+result_dir = currentfolder;
+end

### 2020-12-18 manuscript/draft_rsfmri.Rmd
FC result tables; two design options

@@
 # [ ] Discussion
-# [ ] hervorheben, was von Präregistrierung abweicht
 # [ ] brain images mit nilearn plotting tools einbinden
@@
 }
-
 ```
@@
     booktabs = T,
+    linesep = "",
     caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
@@
 #### with adjustment for age and sex
-Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD, no connectivity was found between the PCC with one other voxel.
+Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
 Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
 <!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 
@@
 ## Effect of mean and change in BMI (exploratory analysis)
+```{r tableResBMI, include=FALSE}
+result_dir <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
+BMImodel_dirs <- 
+  c("noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex-meanFD_WB-c01cl",# deact
+    "noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex_WB-c01cl")# deact
+FDBMImodel_dirs <- 
+  c("noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c02cl", # deact
+    "noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c03cl", # act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/bmi-fd-age-sex_WB-c03cl", # act
+    "noExclFD/FD_total/brain/PCC_cc_z/bmi-fd-age-sex_WB-c01cl") # deact
+FDmodel_dirs <- 
+  c("noExclFD/FD_total/brain/Nacc_cc_z/fd-age-sex_WB-c01cl", # act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/fd-age-sex_WB-c01cl") # act
+deact_file <- "results_deact.csv"
+act_file <- "results_act.csv"
+```
+
+```{r}
+# noExclFD, total sample, all tp
+### first option
+# BMIagesexfd_PCCcc_avgBMI_deact
+df1 <- read.csv(file.path(result_dir,BMImodel_dirs[1],deact_file))
+colnr <- ncol(df1)
+df1$seed <- NULL
+df1 <- rbind(rep(NA, colnr),df1)
+df1[1:2,"seed"] <- c("\\textit{average BMI (decrease)}","PCC (cc)")
+df1$covariates <- NA
+df1$covariates[2] <- "age, sex, log mFD"
+# BMIagesex_PCCcc_avgBMI_deact 
+df2 <- read.csv(file.path(result_dir,BMImodel_dirs[2],deact_file))
+df2$seed <- NULL
+df2 <- rbind(rep(NA, colnr),df2)
+df2[1:2,"seed"] <- c("\\textit{average BMI (decrease)}","PCC (cc)")
+df2$covariates <- NA
+df2$covariates[2] <- "age, sex"
+
+### alternative (with group_row())
+df1 <- read.csv(file.path(result_dir,BMImodel_dirs[1],deact_file))
+colnr <- ncol(df1)
+df1$seed <- NULL
+df1[1,"seed"] <- c("PCC (cc)")
+df1$covariates <- NA
+df1$covariates[1] <- "age, sex, log mFD"
+# BMIagesex_PCCcc_avgBMI_deact 
+df2 <- read.csv(file.path(result_dir,BMImodel_dirs[2],deact_file))
+df2$seed <- NULL
+df2[1,"seed"] <- c("PCC (cc)")
+df2$covariates <- NA
+df2$covariates[1] <- "age, sex"
+
+# BMIFDagesex_NACCcc_cggBMI_deact 
+df3 <- read.csv(file.path(result_dir,FDBMImodel_dirs[1],deact_file))
+df3$seed <- NA
+ df3 <- rbind(rep(NA, colnr),df3)
+df3[1:2,"seed"] <- c("\\textit{change in BMI (decrease)}","NAcc (cc)")
+df3$covariates <- NA
+df3$covariates[2] <- "age, sex"
+# BMIFDagesex_NACCcc_avgFD_act
+ df4 <- read.csv(file.path(result_dir,FDBMImodel_dirs[2],act_file))
+df4$seed <- NA
+df4 <- rbind(rep(NA, colnr),df4)
+df4[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}","NAcc (cc)")
+df4$covariates <- "age, sex"
+# BMIFDagesex_NACCgsr_avgFD_act 
+df5 <- read.csv(file.path(result_dir,FDBMImodel_dirs[3],act_file))
+df5$seed <- NA
+df5 <- rbind(rep(NA, colnr),df5)
+df5[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}","NAcc (gsr)")
+df5$covariates <- NA
+df5$covariates[2] <- "age, sex"
+# BMIFDagesex_PCCcc_avgBMI-deact 
+df6 <- read.csv(file.path(result_dir,FDBMImodel_dirs[4],deact_file))
+df6$seed <- NA
+df6 <- rbind(rep(NA, colnr),df6)
+df6[1:2,"seed"] <- c("\\textit{average BMI (decrease)}","PCC (cc)")
+df6$covariates <- NA
+df6$covariates[2] <- "age, sex"
+
+# BMIFDagesex_NACCcc_avgFD_act 
+df7 <- read.csv(file.path(result_dir,FDmodel_dirs[1],act_file))
+df7$seed <- NA
+df7 <- rbind(rep(NA, colnr),df7)
+df7[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}","NAcc (cc)")
+df7$covariates <- NA
+df7$covariates[2] <- "age,sex"
+# BMIFDagesex_NACCgsr_avgFD_act <
+df8 <- read.csv(file.path(result_dir,FDmodel_dirs[2],act_file))
+df8$seed <- NA
+df8 <- rbind(rep(NA, colnr),df8)
+df8[1:2,"seed"] <- c("\\textit{average log mean FD (increase)}", "NAcc (gsr)")
+df8$covariates <- NA
+df8$covariates[2] <- "age, sex"
+
+BMImodel <- rbind(df1,df2)
+FDmodel <- rbind(df7,df8)
+BMIFDmodel <- rbind(df3,df4,df5,df6)
+```
+
+```{r include = FALSE}
+model_list <- list(BMImodel,FDmodel,BMIFDmodel)
+
+for (i in 1:length(model_list)){
+  model_list[[i]]$anat <- NA
+  model_list[[i]]$hemisphere <- NA
+  # select columns
+  model_list[[i]] <- model_list[[i]][,c("seed","covariates","anat","hemisphere","cluster_p.FWE.corr.","cluster_equivk","cluster_equivkZ","peak_Z","X_x","y","z..mm.")]
+  # name columns
+  colnames(model_list[[i]]) <- c("Seed","Covariates","Anatomical region","L/R","clusterwise FWE-corr. P","cluster size","Z score of cluster size","Z score in local maximum","X","Y","Z") 
+}
+
+for (i in 1:length(model_list)){
+  # only keep cluster statistics and peak local maximum
+  model_list[[i]] <- model_list[[i]][!is.na(model_list[[i]]$Seed) | !is.na(model_list[[i]]$`clusterwise FWE-corr. P`),]
+}
+
+BMImodel <- model_list[[1]]
+FDmodel <- model_list[[2]]
+BMIFDmodel <- model_list[[3]]
+#rm(df1,df2,df3,df4,df5,df6,df7,df8)
+
+```
+
+```{r tabModel}
+# hide NA in tables
+fn1 <- "R, right hemisphere;L, left hemisphere; MNI (Montreal Neurological Institute) coordinates of primary peak location; x = sagittal, y = coronal,"
+fn2 <- "z = axial. To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000 samples."
+title_vec <- c("BMI", "FD", "BMI-FD")
+
+table_list <- list()
+for (i in 1:length(model_list)){
+  table_list[[i]] <-
+    model_list[[i]] %>%
+    knitr::kable(
+      escape = FALSE,     # use font spec of latex with kableExtra
+      col.names = colnames(model_list[[i]]),
+      row.names = FALSE,
+      format = "latex",
+      booktabs = T,
+      linesep = "",      # disable "\\addlinespace" at every 5th line
+      caption = sprintf("Changes in functional connectivity in whole brain analysis for %s model",title_vec[i])
+    ) %>%
+    kable_styling(latex_options = c('hold_position')) %>% # "scale_down"
+    column_spec(c(3, 5, 7, 8), width = "2.4cm") %>%
+    add_footnote(c(fn1, fn2), notation = "none") %>%
+    kableExtra::landscape()
+}
+tab_BMImodel <- table_list[[1]]
+tab_FDmodel <- table_list[[2]]
+tab_BMIFDmodel <- table_list[[3]]
+```
+
+```{r}
+group <- c(1, NA, NA, 2, NA, NA)
+quest <- c("How is your mood today?", "good or very good", "bad or very bad","What colour is your hair?", "brown", "other")
+percent <- c(NA, 80, 20, NA, 50, 50)
+df <- tibble(group, quest, percent)
+df <- data.frame(quest, percent)
+
+tab_BMImodel %>%
+  group_rows("average BMI (decrease)", 1, 3) %>%
+  group_rows("average BMI (decrease)", 4, 7)
+
+```
+
+```{r tabBMImodel}
+tab_BMImodel
+```
+
+```{r tabFDmodel}
+tab_FDmodel
+```
+
+```{r tabBMIFDmodel, echo = FALSE}
+tab_BMIFDmodel
+```
 
@@
 ## Effect on hypothalamic connectivity (exploratory analysis)
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 

### 2020-12-18 code/RunModelBMI.m

@@ -51,12 +51,12 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) %strcmp(param.ACTION,'display')
     pause(7)
 end
 
+% specify path to get info txt for SwE model
 if param.EXCLFD==true
     param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
 else
     param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
 end
-
 if strcmp(param.MODEL,'bmiIG')
     param.INFO_DIR = fullfile(param.INFO_DIR, 'IG/total');
 elseif strcmp(param.MODEL,'bmi2tp')
@@ -65,16 +65,18 @@ else
     param.INFO_DIR = fullfile(param.INFO_DIR, 'both/total');
 end
 
+% specify number pf ROI + estimation procedure
 nrun = length(param.ROI_PREP);
-% estimate any of the models for every roi_prep in ROI_PREP
+if param.WILD_BOOT
+    % wild bootstrap
+    param.wild_con = 1:2;
+else
+    % parametric estimation
+    param.wild_con = false;
+end
+
+% run swe
 for crun = 1:nrun
-    if param.WILD_BOOT
-        % wild bootstrap
-        param.wild_con = 1%:2;
-    else
-        % parametric estimation
-        param.wild_con = false;
-    end
     for i = param.wild_con
         param.wild_con = i;
         % check if analysis has run already
@@ -91,7 +93,7 @@ for crun = 1:nrun
             pause(param.VIEWSEC)
         elseif not(exist_already) || param.OVERWRITE %strcmp(param.ACTION,'overwrite')
             fprintf('Estimate model...\n')
-              % delete existing files in folder if existent
+            % delete existing files in folder if existent
             if exist_already
                 rmdir(out_folder, 's'); % delete former dir
                 mkdir(out_folder); % create new empty one
@@ -101,7 +103,7 @@ for crun = 1:nrun
             spm_jobman('run', matlabbatch);
         else
             fprintf('... model already estimated or error in parameter definitions.\n')
-          end
+        end
     end
  
     % save results of contrasts in folder roi_prep and subfolder modelx
@@ -109,6 +111,7 @@ for crun = 1:nrun
     % https://github.com/NISOx-BDI/SwE-toolbox/issues/135
 end
 fprintf('... done.')
+
 end
 
 

### 2020-12-18 code/RunModelGroupTime.m

@@ -85,28 +85,30 @@ if param.ONLY_DISPLAY && not(param.WILD_BOOT) % strcmp(param.ACTION, 'display')
     pause(7)
 end
 
+% specify path to get info txt for SwE model
 if param.EXCLFD==true
     param.INFO_DIR = fullfile(param.INFO_DIR, 'ExclFD');
 else
     param.INFO_DIR = fullfile(param.INFO_DIR, 'noExclFD');
 end
-
 if strcmp(param.MODEL,'grouptime2tp')
     param.INFO_DIR = fullfile(param.INFO_DIR, 'both/BLFU');
 else
     param.INFO_DIR = fullfile(param.INFO_DIR, 'both/total');
 end
 
+% specify number pf ROI + estimation procedure
 nrun = length(param.ROI_PREP);
-% estimate any of the models for every roi_prep in ROI_PREP
+if param.WILD_BOOT
+    % wild bootstrap
+    param.wild_con = 1:3;
+else
+    % parametric estimation
+    param.wild_con = false;
+end
+
+% run swe
 for crun = 1:nrun
-    if param.WILD_BOOT
-        % wild bootstrap
-        param.wild_con = 1:3;
-    else
-        % parametric estimation
-        param.wild_con = false;
-    end
     for i = param.wild_con
         param.wild_con = i;
         % check if analysis has run already

### 2020-12-18 code/run_or_display.m

@@ -108,7 +108,7 @@ param.ROI_PREP = {roi_prep{[12]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} 
 % association to a model is indicated by the tens digit (GroupTime_: 1_; 
 % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
 param.MODEL = {'bmi'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp} 
-param.COVARIATES = [22];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
+param.COVARIATES = [21];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
 
 % define masking and type of inference
 param.MASK = 'brain';               % 'brain'

### 2020-12-20 manuscript/draft_rsfmri.Rmd
prepare extractions of anatomical regions, create tables

--- 2020-12-17 manuscript/draft_rsfmri.Rmd
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 3 [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
 We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
@@
     booktabs = T,
-    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
+    linesep = "",
+    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points") %>%
+    column_spec(c(2:6), width = "1.8cm")
 
@@
 ## Effect of mean and change in BMI (exploratory analysis)
+
+```{r tableBMImodel}
+source("code/construct_FCtables.R")
+TableList <- construct_FCtables()
+tableBMImodel <- TableList[[1]]
+tableBMImodel
+```
+
+```{r tableFDmodel, echo=FALSE}
+tableFDmodel <- TableList[[2]]
+tableFDmodel
+```
+
+```{r tableBMIFDmodel, echo=FALSE}
+tableBMIFDmodel <- TableList[[3]]
+tableBMIFDmodel
+```
 

### 2020-12-20 code/construct_FCtables.R

@@ -1,0 +1,299 @@
+
+construct_FCtables <- function() {
+  
+# ------------------------------------------------------------------------------
+# Directories
+# ------------------------------------------------------------------------------
+
+res_fc <- list()
+res_fc$result_dir <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
+res_fc$BMImodel_dirs <- 
+  c("noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex-meanFD_WB-c01cl",# deact
+    "noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex_WB-c01cl")# deact
+res_fc$FDBMImodel_dirs <- 
+  c("noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c02cl", # deact
+    "noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c03cl", # act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/bmi-fd-age-sex_WB-c03cl", # act
+    "noExclFD/FD_total/brain/PCC_cc_z/bmi-fd-age-sex_WB-c01cl") # deact
+res_fc$FDmodel_dirs <- 
+  c("noExclFD/FD_total/brain/Nacc_cc_z/fd-age-sex_WB-c01cl", # act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/fd-age-sex_WB-c01cl") # act
+res_fc$deact_csv <- "results_deact.csv"
+res_fc$act_csv <- "results_act.csv"
+
+
+# ------------------------------------------------------------------------------
+# Anatomical Labelling
+# ------------------------------------------------------------------------------
+
+f <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD/result_report.txt"
+result_report <- read.table(f, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)
+
+ResultFCList <- list()
+for (r in 1:nrow(result_report)) {
+  f = result_report[r,"dir"]
+  txt <- read.table(trimws(f), sep="\t", fill = TRUE)
+  name_model <- result_report$model[r]
+  
+  # construct list of clusters for model
+  # ClusterList <- list()
[…]

### 2020-12-20 code/resample_nii.m

@@ -1,0 +1,21 @@
+result_dir = "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis";
+BMImodel_dirs =  [
+    "noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex-meanFD_WB-c01cl",% deact
+    "noExclFD/BMI_total/brain/PCC_cc_z/bmi-age-sex_WB-c01cl" % deact
+    ]; 
+FDBMImodel_dirs = [
+    "noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c02cl", % deact
+    "noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c03cl", % act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/bmi-fd-age-sex_WB-c03cl", % act
+    "noExclFD/FD_total/brain/PCC_cc_z/bmi-fd-age-sex_WB-c01cl" % deact
+    ]; 
+FDmodel_dirs = [
+    "noExclFD/FD_total/brain/Nacc_cc_z/fd-age-sex_WB-c01cl", % act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/fd-age-sex_WB-c01cl" % act
+    ];
+
+folder = fullfile(result_dir, BMImodel_dirs);
+nii = "swe_vox_zTstat_c01.nii";
+V = niftiread(fullfile(folder, nii));
+Vres = imresize3(V, 3, "nearest");
+niftiwrite(Vres,fullfile(folder, strcat(nii{1}(1:end-4),"_res.nii")))

### 2020-12-21 manuscript/draft_rsfmri.Rmd
rename files, update README with pipeline to render manuscipt

@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] re-do analysis with new sample
+# [x] re-do analysis with new sample
 # [ ] repeat analysis with exclusion of 10% of worst data
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 3 [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
 We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
@@
 ```{r tableBMImodel}
-source("code/construct_FCtables.R")
+source("code/tables_FC.R")
 TableList <- construct_FCtables()
-tableBMImodel <- TableList[[1]]
-tableBMImodel
+tab_BMImodel <- TableList[[1]]
+tab_BMImodel
 ```
@@
 ```{r tableFDmodel, echo=FALSE}
-tableFDmodel <- TableList[[2]]
-tableFDmodel
+tab_FDmodel <- TableList[[2]]
+tab_FDmodel
 ```
@@
 ```{r tableBMIFDmodel, echo=FALSE}
-tableBMIFDmodel <- TableList[[3]]
-tableBMIFDmodel
+tab_BMIFDmodel <- TableList[[3]]
+tab_BMIFDmodel
 ```

### 2020-12-21 manuscript/draft_rsfmri.Rmd
clean up: figures + tables are created in other scripts

@@
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-condition.labs <- c("BARS","NBARS")
-names(condition.labs) <- c("IG","KG")
-# BMI over time each group
-figBMIdescr <-
-  ggplot(final, aes(
-    x = tp,
-    y = BMI,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-figBMIdescr +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
+source("report/figures.R")
+figBMIdescr <- mk_figBMIdescr(final)
+
 ```
@@
 # fig.show="hold",  out.width="40%", fig.height = 8,
+source("report/figures.R")
+PlotList <- mk_figDesignMatrix()
 
-# design matrix time as factor
-txt_path <- "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files/noExclFD/both/total/"
-
-ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
-condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
-KG_bl <- read.delim(paste0(txt_path,"KG_bl.txt"), head=FALSE)
-KG_fu <- read.delim(paste0(txt_path,"KG_fu.txt"), head=FALSE)
-KG_fu2 <- read.delim(paste0(txt_path,"KG_fu2.txt"), head=FALSE)
-IG_bl <- read.delim(paste0(txt_path,"IG_bl.txt"), head=FALSE)
-IG_fu <- read.delim(paste0(txt_path,"IG_fu.txt"), head=FALSE)
-IG_fu2 <- read.delim(paste0(txt_path,"IG_fu2.txt"), head=FALSE)
-measurement <- c(1:nrow(ID))
-dmf <- data.frame(measurement,ID,condition,KG_bl,KG_fu,KG_fu2,IG_bl,IG_fu,IG_fu2)
-colnames(dmf) <- c("measurement","ID","condition","N0","N6","N12","B0","B6","B12")
-
-# design matrix time as continuous variable
-ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
-condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
-group_IG <- read.delim(paste0(txt_path,"group_IG.txt"), head=FALSE)
-group_KG <- read.delim(paste0(txt_path,"group_KG.txt"), head=FALSE)
-tp_IG <- read.delim(paste0(txt_path,"tp_IG.txt"), head=FALSE)
-tp_KG <- read.delim(paste0(txt_path,"tp_KG.txt"), head=FALSE)
-measurement <- c(1:nrow(ID))
-dmc <- data.frame(measurement,ID,condition,group_IG,group_KG,tp_IG,tp_KG)
-colnames(dmc) <- c("measurement","ID","condition","groupB", "groupN","timeB","timeN")
-
-# plot dmf
-input_dmf <- dmf %>%
-  pivot_longer(
-    cols= colnames(dmf[,c(4:ncol(dmf))]),
-    names_to = "regressor")
-input_dmc <- dmc %>%
-  pivot_longer(
-    cols= colnames(dmc[,c(4:ncol(dmc))]),
-    names_to = "regressor")
-
-plotA <- ggplot(input_dmf, aes(x = regressor, y = measurement, fill = value)) +
-  geom_tile() +
-  scale_fill_gradient(low="black", high="white") +
-  labs(x="regressors", y="scans") +
-  # design
-  scale_x_discrete(expand = c(0,0),labels=colnames(dmf)[4:ncol(dmf)]) +
-  scale_y_continuous(expand = c(0,0)) +
-  theme_bw() + theme(legend.position = "none",
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12))
-# + theme(panel.grid = element_blank(),
-#         panel.border = element_rect(colour = "black", size=1),
-#    legend.text = element_text(size = 10)
-#         plot.title=element_text(size=11))
-plotB <- ggplot(input_dmc, aes(x = regressor, y = measurement, fill = value)) +
-  geom_tile() +
-  scale_fill_gradient(low="black", high="white") +
-  labs(x="regressors", y="scans") +
-  # design
-  scale_x_discrete(expand = c(0,0),labels=colnames(dmc)[4:ncol(dmc)]) +
-  scale_y_continuous(expand = c(0,0)) +
-  theme_bw() + theme(legend.position = "none",
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12))
-# + theme(panel.grid = element_blank(),
-#         panel.border = element_rect(colour = "black", size=1),
-#    legend.text = element_text(size = 10)
-#         plot.title=element_text(size=11))
-
-plotA
-plotB
+PlotList$plotA
+PlotList$plotB
 #gridExtra::grid.arrange(plotA, plotB, nrow=1)
@@
   final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
-  final[i, "mean_outliers"] = mean(conf$outliers, na.rm = TRUE)
+  final
 }
+
 ```
@@
 
-tab_qc_fc <-
-  knitr::kable(
-    round(res,3),
-    col.names = colnames(res),
-    row.names = TRUE,
-    format = "latex",
-    booktabs = T,
-    linesep = "",
-    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points") %>%
-    column_spec(c(2:6), width = "1.8cm")
-
+tab_qc_fc <- mk_tableQcfc(res)
 tab_qc_fc
@@
 ```{r tableBMImodel}
-source("code/tables_FC.R")
-TableList <- construct_FCtables()
+source("report/tables.R")
+TableList <- mk_FCTables()
 tab_BMImodel <- TableList[[1]]

### 2020-12-21 report/figures.R

@@ -1,0 +1,378 @@
+
+mk_figBMIdescr <- function(final) {
+  
+  condition.labs <- c("BARS","NBARS")
+  names(condition.labs) <- c("IG","KG")
+  # BMI over time each group
+  figBMIdescr <-
+    ggplot(final, aes(
+      x = tp,
+      y = BMI,
+      group = subj.ID,
+      color = condition
+    )) +
+    geom_line() + geom_point() +
+    stat_summary(
+      aes(group = condition),
+      fun = mean,
+      geom  = "line",
+      size = 2,
+      color = c(rep("#00305E", 3), rep("#C00045", 3))
+    ) +
+    facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+  figBMIdescr +
+    scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+    scale_color_manual(
+      values = c("#00305E77", "#C0004577"),
+      labels = c("intervention", "control")
+    ) +
+    xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
+    theme_bw() +  theme(
+      axis.text = element_text(size = 10),
+      axis.title = element_text(size = 12),
+      strip.text = element_text(size = 12),
+      legend.text = element_text(size = 10),
+      legend.position = "none"
+    )
+  return(figBMIdescr)
+}
+
+mk_figDesignMatrix <- function() {
+  
+  # design matrix time as factor
+  txt_path <- "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /SwE_files/noExclFD/both/total/"
+  
+  ID <- read.delim(paste0(txt_path,"subjID.txt"), head=FALSE)
+  condition <- read.delim(paste0(txt_path,"group.txt"), head=FALSE)
+  KG_bl <- read.delim(paste0(txt_path,"KG_bl.txt"), head=FALSE)
+  KG_fu <- read.delim(paste0(txt_path,"KG_fu.txt"), head=FALSE)
+  KG_fu2 <- read.delim(paste0(txt_path,"KG_fu2.txt"), head=FALSE)
+  IG_bl <- read.delim(paste0(txt_path,"IG_bl.txt"), head=FALSE)
+  IG_fu <- read.delim(paste0(txt_path,"IG_fu.txt"), head=FALSE)
[…]

### 2020-12-21 report/tables.R

@@ -1,5 +1,5 @@
 
-construct_FCtables <- function() {
+mk_FCTables <- function() {
   
 # ------------------------------------------------------------------------------
 # Directories
@@ -296,4 +296,42 @@ TableList[[3]] <- TableList[[3]] %>%
 
 rm(df1,df2,df3,df4,df5,df6,df7,df8)
 return(TableList)
+}
+
+mk_SampleTable <- function(final) {
+  # completed time points
+  freqtable <- as.data.frame(table(final$subj.ID))
+  tpfreq <- table(freqtable$Freq)
+  
+  # overview distribution of data points
+  tp1 <- as.character(freqtable$Var1[freqtable$Freq == 1])
+  tp2 <- as.character(freqtable$Var1[freqtable$Freq == 2])
+  tp3 <- as.character(freqtable$Var1[freqtable$Freq == 3])
+  tp1_tab <- table(subset(final, subj.ID %in% tp1)$tp, subset(final, subj.ID %in% tp1)$condition)
+  tp2_tab <- table(subset(final, subj.ID %in% tp2)$tp, subset(final, subj.ID %in% tp2)$condition)[c(1,3),]
+  tp3_tab <- table(subset(final, subj.ID %in% tp3)$tp, subset(final, subj.ID %in% tp3)$condition)[1,]
+  total_datapoints <- c(sum(final$group==1),sum(final$group==2))
+  total_subjects <- plyr::count(final[subj_idx,"group"])$freq
+  tab_sample <- data.frame(rbind(tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints))
+  rm(tp1,tp2,tp3,tp1_tab, tp2_tab, tp3_tab, total_subjects, total_datapoints)
+  rownames(tab_sample) <- c("count: only 0","count: only 6","count: only 12","count: 0 and 6","count: 6 and 12","count: complete data","total number of subjects","total data points")
+  colnames(tab_sample) <- c("BARS","NBARS")
+  return(tab_sample)
+}
+
+mk_tableQcfc <- function(res) {
+  
+  tab_qc_fc <-
+    knitr::kable(
+      round(res,3),
+      col.names = colnames(res),
+      row.names = TRUE,
+      format = "latex",
+      booktabs = T,
+      linesep = "",
+      caption = "Summary of quality metrics for different denoising pipelines across conditions and time points") %>%
+    column_spec(c(2:6), width = "1.8cm") 
+  
+  
+  return(tab_qc_fc)
 }

### 2020-12-21 manuscript/draft_rsfmri.Rmd
include glass brain of networks and LLMMs into draft

--- 2020-12-20 manuscript/draft_rsfmri.Rmd
@@
 # [ ] Discussion
+# [ ] hervorheben, was von Präregistrierung abweicht
 # [ ] brain images mit nilearn plotting tools einbinden
@@
 library(psych) # version 2.0.8
+
 #library(lsr)
-#library(wesanderson) # colour palette
+library(wesanderson) # colour palette
 #library(haven)
@@
 options(knitr.table.format = "latex")
-options(knitr.kable.NA = '') # hide NA in tables
 # set working directory of knit (RMarkdown) to directory of Rproj
@@
     booktabs = T,
-    linesep = "",
     caption = "Distribution of data points at months after intervention")
@@
 Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
+<!-- Frage: Ausschluss bei head movement?
+ADI009_fu2 nicht enthalten (siehe Readme preprocessing) -->
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tableDescr)).
@@
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
+
 ```{r tableBaselineCharacteristics, include=FALSE}
@@
 ```
-<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
 `r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
+Change in BMI throuout the study is depicted in \@ref(fig:figBMIdescr).
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
@@
     booktabs = T,
-    linesep = "",
     caption = "BMI and mean FD of participants acquired in each condition and time point")
@@
 ```
-<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for <!-- standadized??
--->breakfast.
+<!-- STUDY PROCEDURE -->
+Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for STANDARDIZED??
+breakfast.
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
+
 ```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
@@
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
+In addition, for models showing a significant group-by-time interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
 On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
@@
 All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
-We evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012].
+We evaluated mFD-FC relationships by computing a functional connectome based on (Power et al., 2011).
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and time points, adjusted for age and sex.
-We used the modified SwE option and used global signal regressed data as input.
-Network masks were formed from all voxels with a clusterwise $p_{FWE} < 0.05$ and we extracted the average connectivity from these masks for all data points.
+To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
+We used global signal regressed data as input and clusterwise bootstrapping with N=999.
+Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of p_FWE < 0.05.
+We extracted the average connectivity from these masks for all data points.
 
@@
 
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in SPM12.7770<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
 During analyses an implicit mask and an explicit mask were used.
@@
 --> contained one regressor for each time point per group.
-```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of group-time model.", echo=FALSE}
 # fig.show="hold",  out.width="40%", fig.height = 8,
@@
 With a model containing average BMI and longitudinal change per subject, we investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- NUISANCE COVARIATES --> Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+<!-- NUISANCE COVARIATES --> Age at baseline, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Age and sex were between-subject covariates while FD is time-varying.
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
+<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
 As change in FD and change in BMI share common variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
@@
 Differences in results will be reported.
-We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
-<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
-As these post-hoc analyses were not preregistered, they should be considered to be exploratory.
-
-<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).
+<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$.
+These post-hoc analyses deviate from the preregistration and should be considered to be exploratory.
+<!-- INFERENCE --> <!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
+Significant clusters are defined as family-wise error - corrected p<0.05.
+-->
+To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
 Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 3 [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
-We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
-Significant clusters are defined as family-wise error - corrected p<0.05.
--->
 
 <!-- R analysis of aggregated values -->
-We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
-Following the preregistration, we first investigated a time-by-group interaction between baseline and follow-up.
+We analyzed the aggregate connectivity values using linear mixed models in the R package lme4.
+Following the preregistration, we first investigated a time-by-group interaction between baseline and followup.
 We used a model comparison of R1 and R0 models with `anova`:
@@
 We performed a sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI.
-Again, the time-by-group point interaction was considered significant if a model comparison between R1 and Null models using anova-function yielded a p<0.05.
+Again, the group-by-time point interaction was considered significant if a model comparison between R1 and Null models using anova-function yielded a p<0.05.
 Finally, we estimated the above-mentioned interaction model for all three time points.
@@
 We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data.
-We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
+We saw that DVARS did not qualitatively differ between groups and timepoints, and that the correlation of DVARS and mFD remained similar over time points.
 This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
 ```{r tableQA, fig.cap="Summary table of QA measures"}
-
 for (i in 1:nrow(final)) {
@@
 }
+
 ```
@@
     booktabs = T,
-    linesep = "",
-    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points") %>%
-    column_spec(c(2:6), width = "1.8cm")
+    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points")
 
@@
 Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's $r = $`r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's $r = $`r round(res[1,5],2)`).
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
-AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r = $`r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r = $`r round(res[4,5],2)`).
+AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
+Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC valued.
+Both networks include previously described brain regions (i.e.
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulare, right amygdala and ventromedial-prefrontal cortex for the reward network).
+Yet, the reward network is less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to SNR differences in the seed brain regions (see Figure \@{ref:fig:figtSNR}.
+
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p_FWE<0.05)",fig.subcap=c('DMN (seeded in bilateral precuneus)','Reward network (seeded in bilateral Nucleus Accumbens'),fig.show='hold',fig.align='center'}
+networks=c("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/DMN_thresholded_for_clusterFWE0.05.png","/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/Rew_thresholded_for_clusterFWE0.05.png")
+knitr::include_graphics(networks)
+```
+```{r prepare tsnr ROI plot, echo=FALSE}
+final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
+
+tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
+tsnr=tsnr[tsnr$V1!="subj",]
+colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+levels(tsnr$roi)=droplevels(tsnr$roi)
+levels(tsnr$roi)=c("NAcc", "precuneus")
+tsnr$mean=as.double(tsnr$mean)
+
+final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
+
+p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
+  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
+  theme(axis.text=element_text(size=10),
+  axis.title=element_text(size=12),
+  strip.text = element_text(size=12),
+  legend.position="")
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", fig.width=6, fig.height=4,  echo=FALSE}
+p
+```
 
 ## Interaction effect of intervention and time (preregistered analysis)
-
 ### DMN/reward network with whole brain mask (preregistered analysis)
@@
 #### with adjustment for age and sex
-Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
+Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD, no connectivity was found between the PCC with one other voxel.
 Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
@@
 
+<!-- Neurovault -->
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction will be published on NeuroVault.
+
 ### DMN/reward network with GM mask (exploratory analysis)
 <!-- GM mask: Group*Time Interaction [OLD]-->
-Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
+Parametric and non-parametric significance test for the model testing for group-time-interaction did not show any different results when using a grey matter mask.
+Again, parametric estimation and non-parametric estimation differed in regard to the group-time-interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
 While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->.
@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
+```{r load agg FC data and prepare, echo=FALSE}
+agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
+agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
+agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
+agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
+for (i in c(1:nrow(agg_FC_DMN_ID))){
+  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
+  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
+}
+
+colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
+colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
+final_FC$subj.ID=as.factor(final_FC$subj.ID)
+
+final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
+final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
+
+#calculate mean FD over timepoints
+final_FC.meanFD=tapply(X=final_FC$logmFD,
+                        INDEX=final_FC$subj.Nr, FUN=mean, na.rm=TRUE)
+final_FC$mean.logmFD=
+  final_FC.meanFD[as.numeric(final_FC$subj.Nr)]
+final_FC$within.logmFD=
+  final_FC$logmFD-final_FC$mean.logmFD
+```
 
 ### DMN for two timepoints
+```{r vif for models, echo=FALSE}
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
+vif(test_vif)
+```
+
 ```{r calc DMN two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
-```
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anova(R1,R0)
+tab_model(R1,R0)
+```
+
 ### DMN for two timepoints, adjusted for age, sex, mean FD
 ```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anova(R1,R0)
+tab_model(R1,R0)
 ```
@@
 ```{r calc DMN three tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
+anova(R1,R0)
+```
+
+```{r DMN conn over three timepoints, echo=FALSE}
+fig_DMNconn <-
+  ggplot(final_FC, aes(
+    x = tp,
+    y = mean_DMN_conn,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+fig1=fig_DMNconn +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time [months]") + ylab ("mean DMN connectivity") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
+fig1
 ```
@@
 ```{r calc rew two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(rew_FC ~ timepoint*group + (1|subj))
-R0 = lme4::lmer(rew_FC ~ timepoint + group + (1|subj))
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anova(R1,R0)
 ```
@@
 ```{r calc rew two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anova(R1,R0)
 ```
@@
 ```{r calc rew three tp, eval = FALSE, echo = FALSE}
-R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
-```
-
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
+anova(R1,R0)
+```
+
+```{r Rew conn over three timepoints, echo=FALSE}
+fig_Rewconn <-
+  ggplot(final_FC, aes(
+    x = tp,
+    y = mean_Rew_conn,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+fig2=fig_Rewconn +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time [months]") + ylab ("mean reward network connectivity") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
+fig2
+```
 ## Effect of mean and change in BMI (exploratory analysis)
-
-```{r tableBMImodel}
-source("code/construct_FCtables.R")
-TableList <- construct_FCtables()
-tableBMImodel <- TableList[[1]]
-tableBMImodel
-```
-
-```{r tableFDmodel, echo=FALSE}
-tableFDmodel <- TableList[[2]]
-tableFDmodel
-```
-
-```{r tableBMIFDmodel, echo=FALSE}
-tableBMIFDmodel <- TableList[[3]]
-tableBMIFDmodel
-```
 
@@
 <!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
-We did not find a significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
+There was no significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
 Omitting mean FD as nuisance covariate did not change these results.
@@
 Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
+In regard to the exploratory expeditions on subsamples, consistent with previous results, no significant co-activation was found when using the grey matter mask.
 
@@
 <!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
-Surprisingly, we did not find a significant effect neither for the total sample, nor for the intervention group only.
+Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
 <!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
@@
 
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+<!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
+Surprisingly, there was no significant effect neither for the total sample, nor for the intervention group only.
+<!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
 
@@
 Over the course of one year we collected neuroimaging and anthropometric data from `r cs["Sum","IG"]` morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of  `r cs["Sum","KG"]` patients on the waiting list.
-<!-- model specific --> With the model on time-by-group--interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
+<!-- model specific --> With the model on group-time-interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
 Beyond that, also main effects of the intervention and of the time were examined.
@@
 @McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
-However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the time-by-group--interaction with age and sex as nuisance covariates.
+However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the group-time-interaction with age and sex as nuisance covariates.
 <!-- (FD is time-varying...
@@
 -->
-An advantage of the time-by-group interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
+An advantage of the group-time interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
 The model was set up as flexible factorial design with factors "subject", "group", and "time" assuming unequal group variances.

### 2020-12-21 manuscript/draft_rsfmri.Rmd
include glass brain of networks and LLMMs into draft

--- 2020-12-21 manuscript/draft_rsfmri.Rmd
@@
 Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's $r = $`r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's $r = $`r round(res[1,5],2)`).
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
-AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r = $`r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r = $`r round(res[4,5],2)`).
+AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
+Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC valued.
+Both networks include previously described brain regions (i.e.
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulare, right amygdala and ventromedial-prefrontal cortex for the reward network).
+Yet, the reward network is less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to SNR differences in the seed brain regions (see Figure \@{ref:fig:figtSNR}.
+>>>>>>> include glass brain of networks and LLMMs into draft
+
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p_FWE<0.05)",fig.subcap=c('DMN (seeded in bilateral precuneus)','Reward network (seeded in bilateral Nucleus Accumbens'),fig.show='hold',fig.align='center'}
+networks=c("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/DMN_thresholded_for_clusterFWE0.05.png","/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/Rew_thresholded_for_clusterFWE0.05.png")
+knitr::include_graphics(networks)
+```
+```{r prepare tsnr ROI plot, echo=FALSE}
+final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
+
+tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
+tsnr=tsnr[tsnr$V1!="subj",]
+colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+levels(tsnr$roi)=droplevels(tsnr$roi)
+levels(tsnr$roi)=c("NAcc", "precuneus")
+tsnr$mean=as.double(tsnr$mean)
+
+final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
+
+p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
+  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
+  theme(axis.text=element_text(size=10),
+  axis.title=element_text(size=12),
+  strip.text = element_text(size=12),
+  legend.position="")
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", fig.width=6, fig.height=4,  echo=FALSE}
+p
+```
 
 ## Interaction effect of intervention and time (preregistered analysis)
-
 ### DMN/reward network with whole brain mask (preregistered analysis)
@@
 
+<!-- Neurovault -->
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction will be published on NeuroVault.
+
 ### DMN/reward network with GM mask (exploratory analysis)
@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
+```{r load agg FC data and prepare, echo=FALSE}
+agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
+agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
+agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
+agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
+for (i in c(1:nrow(agg_FC_DMN_ID))){
+  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
+  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
+}
+
+colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
+colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
+final_FC$subj.ID=as.factor(final_FC$subj.ID)
+
+final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
+final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
+
+#calculate mean FD over timepoints
+final_FC.meanFD=tapply(X=final_FC$logmFD,
+                        INDEX=final_FC$subj.Nr, FUN=mean, na.rm=TRUE)
+final_FC$mean.logmFD=
+  final_FC.meanFD[as.numeric(final_FC$subj.Nr)]
+final_FC$within.logmFD=
+  final_FC$logmFD-final_FC$mean.logmFD
+```
 
 ### DMN for two timepoints
+```{r vif for models, echo=FALSE}
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
+vif(test_vif)
+```
+
 ```{r calc DMN two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anova(R1,R0)
+tab_model(R1,R0)
 ```
+
 ### DMN for two timepoints, adjusted for age, sex, mean FD
 ```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anova(R1,R0)
+tab_model(R1,R0)
 ```
@@
 ```{r calc rew three tp, eval = FALSE, echo = FALSE}
-R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
+anova(R1,R0)
 ```
 
+```{r Rew conn over three timepoints, echo=FALSE}
+fig_Rewconn <-
+  ggplot(final_FC, aes(
+    x = tp,
+    y = mean_Rew_conn,
+    group = subj.ID,
+    color = condition
+  )) +
+  geom_line() + geom_point() +
+  stat_summary(
+    aes(group = condition),
+    fun = mean,
+    geom  = "line",
+    size = 2,
+    color = c(rep("#00305E", 3), rep("#C00045", 3))
+  ) +
+  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+fig2=fig_Rewconn +
+  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+  scale_color_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")
+  ) +
+  xlab("Time [months]") + ylab ("mean reward network connectivity") +
+  theme_bw() +  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    strip.text = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.position = "none"
+  )
+fig2
+```
 ## Effect of mean and change in BMI (exploratory analysis)
@@
 ## Effect on hypothalamic connectivity (exploratory analysis)
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 

### 2020-12-21 manuscript/draft_rsfmri.Rmd

@@
 Yet, the reward network is less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to SNR differences in the seed brain regions (see Figure \@{ref:fig:figtSNR}.
->>>>>>> include glass brain of networks and LLMMs into draft
 
@@
 
+=======
+
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p_FWE<0.05)",fig.subcap=c('DMN (seeded in bilateral precuneus)','Reward network (seeded in bilateral Nucleus Accumbens'),fig.show='hold',fig.align='center'}
+networks=c("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/DMN_thresholded_for_clusterFWE0.05.png","/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/Rew_thresholded_for_clusterFWE0.05.png")
+knitr::include_graphics(networks)
+```
+```{r prepare tsnr ROI plot, echo=FALSE}
+final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
+
+tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
+tsnr=tsnr[tsnr$V1!="subj",]
+colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+levels(tsnr$roi)=droplevels(tsnr$roi)
+levels(tsnr$roi)=c("NAcc", "precuneus")
+tsnr$mean=as.double(tsnr$mean)
+
+final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
+
+p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
+  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
+  theme(axis.text=element_text(size=10),
+  axis.title=element_text(size=12),
+  strip.text = element_text(size=12),
+  legend.position="")
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", fig.width=6, fig.height=4,  echo=FALSE}
+p
+```
+
+>>>>>>> 30e451a04dedc0054f54c9f8b5bb0ffb11922e68
 ## Interaction effect of intervention and time (preregistered analysis)

### 2020-12-21 manuscript/draft_rsfmri.Rmd
remove intro text (now in intro_rsfmri)

@@
 library(psych) # version 2.0.8
-
 #library(lsr)
-library(wesanderson) # colour palette
+#library(wesanderson) # colour palette
 #library(haven)
@@
 ```
-\newpage
-# Abstract
-
-\newpage
-# Introduction
-
-<!-- obesity and interventions -->
-Over the past decades, the global burden of obesity has increasingly been confronted as a matter of urgency by research.
-_something about comorbidities_
-Besides metabolic dysfunctions, obesity has been consistently associated with reduced gray matter volume, less intact white matter structure and reduced cognitive function [@Beyer_2019; @Zhang_2018].
-Luckily, dietary interventions such as caloric restriction and related metabolic improvements, and weight loss have been suggested to improve neuronal plasticity and cognitive performance [@Witte_2009; @Zhang_2018; @Veronese_2017].
-In some cases behavioral intervention fail to yield the desired success regarding weight loss. There are non-behavioral alternatives at hand.
-<!-- bars most invasive (brain metabolism), last resort -->
-Bariatric surgery is one promising option to combat morbid obesity, as it rapidly improves weight status, metabolic dysfunctions, and co-morbidities, such as diabetes [@].
-Roux-en-Y gastric bypass (RYGB) is the most common procedure as it not only restricts the food intake by reducing stomach volume but also affects the absorption of macronutrients and the hormonal profile [@].
-<!-- ref for "common", ref for procedure -->
-Other widely used procedures are vertical sleeve gastrectomy (VSG) and gastric banding (GB) which only restrict food consumption. <!-- mehr zu Veränderung in Gasto-intestinaler Physiologie -->
-
-<!-- Neural mechanisms/concurrent plasticity? -->
-Precise mechanisms are yet to be elucidated.
-As a first step, a comprehensive understanding of the concurrent physiological changes underlying weight loss and recovery is needed.
-This will inform inferences about potential mechanisms and either support interventions proven to be effective or ideally allow imitating them will less invasive alternatives.
-
-Recently, research focuses have slowly shifted to investigate brain metabolism structural and functional changes after surgery.
-<!-- severe nutritional stress [@Crujeiras_2017] -->
-
-<!-- Reward network -->
-Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
-<!-- mechanism reward and appetite, overating -->
-Besides malabsorption and restricted food ingestion, patients report post-surgical changes in food preferences and eating behavior.
-Those psychophysiological and behavioral effects may be related to post-surgical changes in visceral sensory perceptions and endocrinological signaling.
-Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017, @Nudel_2019].
-<!-- results in lean -->
-The reward network entailing the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016].
-<!-- results in obese longitudinal -->
-Similarly, several studies have found differences in resting-state activity in reward regions depending on weight status.
-
-Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
-These differences seemed to normalize, as trends reversed when looking at the resting state of obese individuals after compared to before the surgery [@Duan_2020].
-
-On the other hand, BMI prior to the surgery was associated with changes in connectivity [@Duan_2020].
-
-s concerning such processes following a surgical intervention are inconclusive.
-
-<!-- important differences in the response to visual food-related cues between thin and reduced-obese individuals that place the latter group at an increased risk of  -->
-
-<!-- Though there exists strong evidence for important differences in the ..
-between lean or normal-weight individuals and obese individuals, few studies have investigated the ...
-and even fewer compared reduced-obese individuals reduced-obese individuals -->
-<!-- [@Garcia_Garcia_2012] -->
-
-<!-- Default mode network -->
-Weight regulation and energy expenditure is dependent on a complex interplay of external and internal sensory processing, as well as molecular homeostasis regulation.
-Obesity may be cause or effect of a decompensation of the psychophysiological system that ensures homeostasis.
-The default mode network is commonly associated with processing of self-referent as well as introspective information, such as hunger sensation [@;@] and appetite [@Tregellas_2011].
-<!-- parietal activation correlated with appetite measures after an overnight fast [@Tregellas_2011].
---> It is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx.
-BA 39) [@Raichle_2015].
-The right ventral anterior cingulate cortex (VACC), superior frontal cortex (SFC), and inferior temporal cortex (ITC) show weaker but also reliable association with the PCC over the course of one year.
-Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
-<!-- difference lean obese -->
-Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2016].
-Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
-A longitudinal study that confirmed the stronger connectivity (of PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG).
-However, default mode connectivity in these regions seemed weakened over the course of one year.
-Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable 1 year after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
-This is consistent with other findings of reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC [@Li_2018] <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery.
-
-The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated.
-Among patients undergoing surgery, change in BMI was negatively correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
-
-On the other hand, connectivity also served as a predictor of weight loss after surgery.
-
-On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011].
-
-<!-- novelty statement --> As described previously, several studies have investigated the effects of bariatric surgery on the intrinsic functional connectivity in patients.
-Yet, to our knowledge there is no quasi-experimental study, comparing a cohort of patients to a cohort with similar characteristics but unexposed to the intervention, i.e.
-bariatric surgery, over an extended time period as would be demanded to draw more substantiated inference [@George_2016].
-<!-- hypotheses --> In this study, we aim to bridge this gap and investigate possible dynamics of within-reward network connectivity before and after bariatric surgery by comparison with a contemporaneously measured waiting-control group.
-Moreover, we will explore whether the changes in functional connectivity are associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential gradient link.
-In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network (DMN).
-Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
-
-<!-- group effect because direct effect of intervention surgery -> rsfmri (cares lessfreq, less preoccupation with hunger, less/more reward sensitivity) vs.
-sugery -> BMI -> rsfmri (neural signalling, fatty tissue) -->
 
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
-We used global signal regressed data as input and clusterwise bootstrapping with N=999.
-Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of p_FWE < 0.05.
-We extracted the average connectivity from these masks for all data points.
+To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and time points, adjusted for age and sex.
+We used the modified SwE option and used global signal regressed data as input.
+Network masks were formed from all voxels with a clusterwise $p_{FWE} < 0.05$ and we extracted the average connectivity from these masks for all data points.
 
@@
 We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
-The model was examined for the correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $8$ a priori planned analyses.
-
+These were examined each for correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $16$ a priori planned analyses.
 Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$.
@@
 write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
-
 ```
@@
 Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's r = `r round(res[1,5],2)`).
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's $r = $`r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's $r = $`r round(res[1,5],2)`).
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
-AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
+AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r = $`r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r = $`r round(res[4,5],2)`).
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
-Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC valued.
-Both networks include previously described brain regions (i.e.
-medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulare, right amygdala and ventromedial-prefrontal cortex for the reward network).
-Yet, the reward network is less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to SNR differences in the seed brain regions (see Figure \@{ref:fig:figtSNR}.
 
-```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p_FWE<0.05)",fig.subcap=c('DMN (seeded in bilateral precuneus)','Reward network (seeded in bilateral Nucleus Accumbens'),fig.show='hold',fig.align='center'}
-networks=c("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/DMN_thresholded_for_clusterFWE0.05.png","/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/Rew_thresholded_for_clusterFWE0.05.png")
-knitr::include_graphics(networks)
-```
-```{r prepare tsnr ROI plot, echo=FALSE}
-final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
+## Interaction effect of intervention and time (preregistered analysis)
 
-tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
-tsnr=tsnr[tsnr$V1!="subj",]
-colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
-levels(tsnr$roi)=droplevels(tsnr$roi)
-levels(tsnr$roi)=c("NAcc", "precuneus")
-tsnr$mean=as.double(tsnr$mean)
-
-final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
-
-p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
-  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
-  theme(axis.text=element_text(size=10),
-  axis.title=element_text(size=12),
-  strip.text = element_text(size=12),
-  legend.position="")
-```
-
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", fig.width=6, fig.height=4,  echo=FALSE}
-p
-```
-
-=======
-
-```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p_FWE<0.05)",fig.subcap=c('DMN (seeded in bilateral precuneus)','Reward network (seeded in bilateral Nucleus Accumbens'),fig.show='hold',fig.align='center'}
-networks=c("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/DMN_thresholded_for_clusterFWE0.05.png","/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/Rew_thresholded_for_clusterFWE0.05.png")
-knitr::include_graphics(networks)
-```
-```{r prepare tsnr ROI plot, echo=FALSE}
-final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
-
-tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
-tsnr=tsnr[tsnr$V1!="subj",]
-colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
-levels(tsnr$roi)=droplevels(tsnr$roi)
-levels(tsnr$roi)=c("NAcc", "precuneus")
-tsnr$mean=as.double(tsnr$mean)
-
-final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
-
-p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
-  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
-  theme(axis.text=element_text(size=10),
-  axis.title=element_text(size=12),
-  strip.text = element_text(size=12),
-  legend.position="")
-```
-
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", fig.width=6, fig.height=4,  echo=FALSE}
-p
-```
-
->>>>>>> 30e451a04dedc0054f54c9f8b5bb0ffb11922e68
-## Interaction effect of intervention and time (preregistered analysis)
 ### DMN/reward network with whole brain mask (preregistered analysis)
@@
 
-<!-- Neurovault -->
-Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction will be published on NeuroVault.
-
 ### DMN/reward network with GM mask (exploratory analysis)
@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
-```{r load agg FC data and prepare, echo=FALSE}
-agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
-agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
-agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
-agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
-for (i in c(1:nrow(agg_FC_DMN_ID))){
-  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
-  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
-}
-
-colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
-colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
-final_FC$subj.ID=as.factor(final_FC$subj.ID)
-
-final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
-final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
-
-#calculate mean FD over timepoints
-final_FC.meanFD=tapply(X=final_FC$logmFD,
-                        INDEX=final_FC$subj.Nr, FUN=mean, na.rm=TRUE)
-final_FC$mean.logmFD=
-  final_FC.meanFD[as.numeric(final_FC$subj.Nr)]
-final_FC$within.logmFD=
-  final_FC$logmFD-final_FC$mean.logmFD
-```
 
 ### DMN for two timepoints
-```{r vif for models, echo=FALSE}
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
-vif(test_vif)
+```{r calc DMN two tp, eval = FALSE, echo = FALSE}
+R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
 ```
-
-```{r calc DMN two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anova(R1,R0)
-tab_model(R1,R0)
-```
-
 ### DMN for two timepoints, adjusted for age, sex, mean FD
 ```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anova(R1,R0)
-tab_model(R1,R0)
+R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
 ```
@@
 ```{r calc DMN three tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
-anova(R1,R0)
-```
-
-```{r DMN conn over three timepoints, echo=FALSE}
-fig_DMNconn <-
-  ggplot(final_FC, aes(
-    x = tp,
-    y = mean_DMN_conn,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig1=fig_DMNconn +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time [months]") + ylab ("mean DMN connectivity") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
-fig1
+R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
 ```
@@
 ```{r calc rew two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anova(R1,R0)
+R1 = lme4::lmer(rew_FC ~ timepoint*group + (1|subj))
+R0 = lme4::lmer(rew_FC ~ timepoint + group + (1|subj))
 ```
@@
 ```{r calc rew two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anova(R1,R0)
+R1 = lme4::lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
 ```
@@
 ```{r calc rew three tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
-anova(R1,R0)
+R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
+R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
 ```
 
-```{r Rew conn over three timepoints, echo=FALSE}
-fig_Rewconn <-
-  ggplot(final_FC, aes(
-    x = tp,
-    y = mean_Rew_conn,
-    group = subj.ID,
-    color = condition
-  )) +
-  geom_line() + geom_point() +
-  stat_summary(
-    aes(group = condition),
-    fun = mean,
-    geom  = "line",
-    size = 2,
-    color = c(rep("#00305E", 3), rep("#C00045", 3))
-  ) +
-  facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-fig2=fig_Rewconn +
-  scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
-  scale_color_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")
-  ) +
-  xlab("Time [months]") + ylab ("mean reward network connectivity") +
-  theme_bw() +  theme(
-    axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    strip.text = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.position = "none"
-  )
-fig2
-```
 ## Effect of mean and change in BMI (exploratory analysis)

### 2021-01-06 manuscript/draft_rsfmri.Rmd
corrected results

@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [x] re-do analysis with new sample
-# [ ] repeat analysis with exclusion of 10% of worst data
+# [ ] repeat analysis with exclusion of 10% of worst data ??
 # [ ] Network tests
+# [ ] Swap mFD with logmFD ??
 
@@
 source("report/figures.R")
-figBMIdescr <- mk_figBMIdescr(final)
-
+fig_BMIdescr <- mk_figBMIdescr(final)
+fig_BMIdescr
 ```
@@
 
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the development version of the SwE toolbox [@Guillaume_2014] ([https://github.com/NISOx-BDI/SwE-toolbox](https://github.com/NISOx-BDI/SwE-toolbox) as of commit #384d2d7944e0c1919a3d22a0282a513916c25f59 was used because of bugs in most recent release) as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
 During analyses an implicit mask and an explicit mask were used.
@@
 #### with adjustment for age, sex, meanFD
-<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
 There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
@@
 Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
-Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
+Results did not differ between analyses of connectivity maps from different preprocessing pipelines.
 <!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
@@
 <!-- GM mask: Group*Time Interaction [OLD]-->
-Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
+<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
 Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->.
+While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
 
@@
 
+Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
+<!-- ### DMN -->
+<!-- #### DMN for three timepoints -->
+<!-- cc -->Using the images preprocessed with AROMA+CC, we did find a significant co-activation with the PCC in three clusters associated with average BMI.
+Omitting mean FD as nuisance covariate yielded similar results but four significant clusters.
+However, in both, there is significant peak activation within the largest cluster.
+<!-- gsr -->In analyses on images that were additionally preprocessed with GSR, no cluster survived statistical thresholds for either ROI.
+All results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
+<!-- ### DMN for two timepoints
+BMIONLYIG 
+Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps).
+However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
+<!-- BMI2TP 
+Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants.
+Omitting mean FD did not change these results.
+-->
+<!-- GM mask 
+Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
+In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
+ -->
+
+## Effect of FD (exploratory analysis)
+
+To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models. 
+First, BMI measures are excluded from the design matrix.
+Covariance of the mFD and outcome meausres is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor.
+Age and sex were nuisance covariates.
+<!-- #### Reward -->
+As shown in Table \@ref(tab:tableFDmodel), there is a significant co-activation of one cluster with the NAcc.
+Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
+<!-- ### DMN -->
+We did not find a any clusters for functional connectivity with the PCC.
+
+## Effect of BMI and FD (exploratory analysis)
+
+Average BMI and change in BMI remain covariates of interest.
+While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as.
+Age and sex remain nuisance covariates.
+
+<!-- ### Reward -->
+<!-- cc -->There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
+Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
+<!-- gsr -->When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds. 
+On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
+
+<!-- ### DMN -->
+<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
+There is a significant association between average BMI and the connecitivity between PCC and three other clusters.
+Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
+<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
+
+## Effect on hypothalamic connectivity (exploratory analysis)
+...
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+
 ```{r tableBMImodel}
@@
 
-```{r tableFDmodel, echo=FALSE}
-tab_FDmodel <- TableList[[2]]
+```{r tableFDmodel}
+tab_FDmodel <- mk_FCTables()[[2]]
 tab_FDmodel
@@
 
-```{r tableBMIFDmodel, echo=FALSE}
-tab_BMIFDmodel <- TableList[[3]]
+```{r tableBMIFDmodel}
+tab_BMIFDmodel <- mk_FCTables()[[3]]
 tab_BMIFDmodel
@@
 
-<!-- Brain mask: average and individual BMI [OLD] -->
-<!-- BMItotal --> Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
-We did not find a significant co-activation for neither the PCC nor the NAcc associated with average BMI nor change in BMI.
-Omitting mean FD as nuisance covariate did not change these results.
-<!-- BMIonlyIG --> Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ --> with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps).
-However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)<!-- but FWE-corr sig?!
--->.
-The effect also disappeared when running those tests without mean FD as nuisance covariate.
-<!-- BMI2tp --> Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants.
-Omitting mean FD did not change these results.
-
-<!-- GM mask: average and individual BMI, damit die Reihenfolge so ist wie oben. -->
-Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
-
-## Effect of FD (exploratory analysis)
-
-<!-- Brain mask: FD --> To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed a model that entailed both of the predictors as covariates of interest and only age and sex as nuisance covariates.
-Surprisingly, we did not find a significant effect neither for the total sample, nor for the intervention group only.
-<!-- GM mask: FD --> Similarly, analysis with the grey matter mask did not reveal any findings either.
-
-## Effect on hypothalamic connectivity (exploratory analysis)
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
-
 # Discussion
 
-<!-- SUMMARY -->
+### Summary
 Utilizing resting-state functional MRI on morbidly obese individuals that either underwent or awaited bariatric surgery over the course of one year.
@@
 Over the course of one year we collected neuroimaging and anthropometric data from `r cs["Sum","IG"]` morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of  `r cs["Sum","KG"]` patients on the waiting list.
-<!-- model specific --> With the model on time-by-group--interaction, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
+<!-- model specific --> Looking at time-by-group interactions, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
 Beyond that, also main effects of the intervention and of the time were examined.
-<!-- correct interpretation of marginal models?
+The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
+<!-- sensitivity analyses -->To get a better understanding of our results, we conducted multiverse exploratory analyses representing different choices regarding preprocessing pipelines, inclusion of covariates and subsets of the data relevant for interpretation.
+
+### Support of original hypotheses
+<!-- false positive -->
+To summarize, while our primary hypothesis on group time interactions could not be confirmed, there was a significant relationship between functional activation 
+and the continuous BMI measures for some but not all analyses.
+
+- DMN: average BMI but only with lenient AROMA-CC but then for both covariate structures age/sex/mFD and age/sex; effect stable when separating avg mFD and cgn FD into two regressors 
+- Reward: effect of change BMI only when separating mFD, bit high peak activation
+The individuals BMI averaged across measures taken throughout the one-year period negatively correlated with functional connectivity between PCC and [regions from clusters].
+Results indicate a decrease a functional connectivity within the default mode network.
+<!-- discuss consistency with previous literatur -->
+... not? in accordance with our hypothesis regarding network activity changes.
+
+Regarding the reward network, we could not predict changes in functional connectivity with the preregistered models. 
+However, when investigating the unique contributions to variance of head motion measures and BMI measures, we detected a significant relationship between the individuals' change in BMI throughout the year.
+This may be preliminary evidence for changes in the functional connectivity between reward and frontal regions induced by weight-loss.
+<!-- discuss consistency with previous literatur -->
+
+In our sample, extend of micromovements of the head averaged across all measurements is associated with increased functional connectivity between reward and frontal regions.
+<!-- explanation?
+overlap with region from change in BMI?
 -->
-The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
 
-<!-- SUPPORT OF ORIGINAL HYPOTHESIS -->
-<!-- false positive -->
-To summarize, there were few significant co-activation that emerged from the marginal models.
-Co-activation was locally highly specific and (NOT???) in accordance with our hypothesis regarding network activity changes.
-Given the multiplicity of of models resulting from variations of inspected networks and sensitivity analysis for different preprocessing and correction approached, the obtained significant correlations between seed regions and voxels are likely to be false positive.
+<!-- Head motion as covariate, multicollinearity -->
+Based on concerns expressed in the literature, we tried to closely inspect and rigorously control for the influence of head motion.
+The presence of multi-collinearity of our predictors has been established before [@Beyer_2019].
+As described by @Monti_2011, in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
+Surprisingly, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region is only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
+However, such unexpected findings would be in accordance with @Monti_2011 stating that <!-- not only will confidence in $\beta$ estimates suffer but  --> estimates can behave erratic depending on which regressors are included in the model.
+
+Notwithstanding the difficulties of quantitatively disentangling multicollinearity between head motion and BMI, preliminary conclusions of our analyses may be...
+<!-- Given the multiplicity of of models resulting from variations of inspected networks and sensitivity analysis for different preprocessing and correction approached, the obtained significant correlations between seed regions and voxels are likely to be false positive.
 Ventures to verify effects by re-calculating with non-parametric Wild bootstrap confirmed supported presumption as effects were not robust across estimation procedures.
+-->
+
 <!-- limitations -->
 Without an a priori power analysis it is not possible to determine if there is no effect or
-To clarify, extended considerations should be made to use a priori power analysis (e.g. with fMRIpower) or check evidence of absence with alternative
+To clarify, extended considerations should be made to use a priori power analysis (e.g. with fMRIpower) or check evidence of absence with alternative statistical procedures.
 
-<!-- SIMILARITY OF RESULTS -->
+#### Similarity of results with literature (random notes)
 Our results are not consistent with the present literature.
@@
 
-<!-- INTERPRETATION -->
+#### Interpretation (random notes)
 <!-- absence of evidence --> However, hypotheses about the plasticity functional networks following bariatric surgery could not be validated.
@@
 --> A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
-After the intervention, the functional connectivity was higher in bariatric patients.
+After the intervention, functional connectivity was higher in bariatric patients.
 It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
@@
 <!-- GENERALIZABILITY -->
-<!-- advantage: prospective intervention control study -->
-Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
-To isolate the effect of the surgical intervention to a maximum possible extent, the intervention group has to be comparable to a control group in regard  to predisposition.
-Studies comparing the sample of bariatric patients to lean individuals or behavioral dieters may have investigated risk or contributing factors to the condition or other dispositional difference between groups, e.g.
-in cognitive control, openness or responsiveness to behavioral interventions [@;@].
-However, they provide insufficient evidence to directly evaluate impact of the surgical procedure.
-Prospective non-randomization trial study design are secondary to randomized control trials in regard to internal control.
-By establishing temporality, longitudinal measurements support the validity of conclusions about the effects of the surgery.
-Contrasting pre- and post-surgery measures change while controlling for individual variance.
-The waiting list control group equivalent in BMI, treatment history and  treatment recommendation allows controlling for concurrent factors [@Thiese_2014].
-Incorporating these consideration, our study aimed at a higher internal validity for conclusions about the effects of bariatric surgery on functional connectivity.
+#### Internal validity
+<!-- advantage: prospective intervention control study -->Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
+Previous studies have compared the sample of bariatric patients to lean individuals, behavioral dieters, or obese individuals without a recommendation for bariatric surgery.
+Despite the accumulated evidence, effects may still be attributable to baseline differences, contributing factors to the condition or other psychophysiological and dispositional difference between groups [@].
+In order to build upon and complement previous evidence, we aimed to isolate the impact of the surgical procedure on changes in resting-state connectivity to a maximum possible extent by conducting a prospective non-randomization trial study.
+Such a study design is secondary to randomized control trials in regard to internal control by means of temporality and controllability.
+We established temporality by contrasting pre- and post-surgery measures change while controlling for individual variance.
+<!-- Such longitudinal analyses support the validity of conclusions about the effects of the surgery.
+-->
+The waiting list control group is equivalent in BMI, treatment history and  treatment recommendation and therefore allows controlling for concurrent factors [@Thiese_2014]. 
+Incorporating these consideration, our study aimed at a higher internal validity for conclusions about the changes in functional connectivity specific to surgery-induced changes.
 
-<!-- advantage: Nucleus Accumbensrepresentative sample -->
+#### Study sample
+<!-- advantage: representative sample -->
 Size and composition of our sample did not allow separate analyses for the sexes.
-The disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
+However, the disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
 
@@
 
-<!-- mixed effect model -->
+#### Marginal model
+<!-- advantage: controlling for individual variance -->
 Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), it is often recommended to use modern models such as mixed-effect models and marginal models [@Chen_2013;@Guillaume_2014;@McFarquhar_2019].
 @McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
-However, <!-- to rule out gross missspecification in the SwE model,  -->for the purpose of comparison, we re-ran the analysis for the time-by-group--interaction with age and sex as nuisance covariates.
-<!-- (FD is time-varying...
-schwierig) with a simplified model without nuisance covariates in SPM12.
-We opted for a simplified model because different nuisance covariates, time-invariant or time-varying, create additional complications.
--->
+<!-- However, for the purpose of comparison, we re-ran the analysis for the time-by-group--interaction with age and sex as nuisance covariates.
+We opted for a simplified model investigating the time-by group interaction in ...
+because different nuisance covariates, time-invariant or time-varying, create additional complications.
 An advantage of the time-by-group interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
 The model was set up as flexible factorial design with factors "subject", "group", and "time" assuming unequal group variances.
-<!-- vastly different - different implicit masks? -->
-Differences could be attributed to different corrections for non-spherity and assumptions on voxel covariance structures.
+-->
+<!-- Differences could be attributed to different corrections for non-spherity and assumptions on voxel covariance structures.
 Detailed comparisons between SPM12, SwE, and other software packages have been described elsewhere [@McFarquhar_2016].
+-->
 <!-- conclusion??? -->
-Opposed to flexible factorial models, where each individual is modelled separately, models are parsimonious and preserve more degrees of freedom.
+Opposed to flexible factorial models, where each individual is modelled separately, marginal models are parsimonious and preserve more degrees of freedom.
 
@@
 <!-- Relevance of rsfMRI as biomarker -->
-<!-- Consideration of head motion as covariate -->
-<!-- Problem head motion -->
-Based on concerns expressed in the literature [@], we tried to closely inspect and rigorously control for the influence of head motion.
-The presence of multicollinearity of our predictors has been established before.
-As described in @Monti_2011 in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
-Confidence in $\beta$ estimates will suffers and estimates can behave erratic depending on which regressors included in the model.
 

### 2021-01-07 manuscript/draft_rsfmri.Rmd
add figures and QA measures (pre-christmas commit possibly not ready)

--- 2020-12-21 manuscript/draft_rsfmri.Rmd
@@
 library(lme4)
-library(sjPlot)
 
+library(sjPlot) #only works for HTML output
+library(texreg) #works for Rmarkdown
 library(knitr) # version 1.30
@@
 ```
+\newpage
+# Abstract
+
+\newpage
+# Introduction
 
@@
 figBMIdescr <- mk_figBMIdescr(final)
-
+figBMIdescr
 ```
@@
 
+\newpage
 # Results
@@
 We performed different quality controls for the rsfMRI data.
-As described in @Beyer_2020, there was a significant interaction of group and time on mFD.
-In particular, the intervention group reduced in head motion in the follow-up assessments compared to the control group.
-We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data.
+As described in @Beyer_2020, group and time show an interaction effect for mFD where the intervention group reduced in head motion in the follow-up assessments compared to the control group.
+We also inspected DVARS, a measure of whole-brain signal variation which is reflective of noise in the data.
 We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
 This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
-```{r tableQA, fig.cap="Summary table of QA measures"}
+```{r tableQcfc, echo=FALSE}
+res <- read.csv("./report/FDFCcorrelations.csv")
+rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
+colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
 
+tab_qc_fc <- mk_tableQcfc(res)
+tab_qc_fc
+```
+Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018].
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`).
+This relationship was negatively associated with distance between nodes (mean Spearman's r = `r round(res[1,5],2)`).
+Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
+Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency.
+Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
+Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
+
+```{r calculate summary of QA, eval=FALSE, echo=FALSE}
 for (i in 1:nrow(final)) {
@@
 }
-
 ```
 
-```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD'}
-# ,  fig.align='center', }
-# eval = TRUE, fig.asp = 0.7, fig.width = 6
+```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
+final=read.csv(file = "./report/all_QA_measures.csv")
 
@@
 source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
-
 FDFC=calc_qc_fc(final)
@@
 
-```{r tableQcfc, echo=FALSE}
-res <- read.csv("./report/FDFCcorrelations.csv")
-rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
-colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
-
-tab_qc_fc <- mk_tableQcfc(res)
-tab_qc_fc
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
+networks=c("../report/DMN_thresholded_for_clusterFWE005.png","../report/Rew_thresholded_for_clusterFWE005.png")
+knitr::include_graphics(networks)
 ```
 
-Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's $r = $`r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's $r = $`r round(res[1,5],2)`).
-Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
-AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r = $`r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r = $`r round(res[4,5],2)`).
-Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
+Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC values.
+Both networks included the typically described brain regions (i.e.
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulare, right amygdala and ventromedial-prefrontal cortex for the reward network).
+
+```{r prepare tsnr ROI plot, echo=FALSE}
+final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
+
+tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
+tsnr=tsnr[tsnr$V1!="subj",]
+colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+levels(tsnr$roi)=droplevels(tsnr$roi)
+levels(tsnr$roi)=c("NAcc", "precuneus")
+tsnr$mean=as.double(tsnr$mean)
+
+final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
+
+p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
+  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
+  theme(axis.text=element_text(size=10),
+  axis.title=element_text(size=12),
+  strip.text = element_text(size=12),
+  legend.position="")
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%",  echo=FALSE}
+p
+```
+Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref{fig:figtSNR}.
 
@@
 <!-- Brain mask: Group*Time Interaction [CORRECTED] -->
-There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
-Nor was there a significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc seeds against our initial hypothesis.
+There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
 
 #### with adjustment for age and sex
-Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
-Results did not differ between analyses of connetivity maps from different preprocessing pipelines.
+Similarly, there were no significant clusters when leaving out mean FD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 <!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
 
@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
+```{r load agg FC data and prepare, echo=FALSE}
+agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
+agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
+agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
+agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
+for (i in c(1:nrow(agg_FC_DMN_ID))){
+  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
+  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
+}
+
+colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
+colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
+
+final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
+final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
+final_FC$subj.ID=as.factor(final_FC$subj.ID)
+
+#calculate mean FD over timepoints
+final_FC.meanFD=tapply(X=final_FC$logmFD,
+                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
+final_FC$mean.logmFD=
+  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
+final_FC$within.logmFD=
+  final_FC$logmFD-final_FC$mean.logmFD
+```
 
 ### DMN for two timepoints
-```{r calc DMN two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
-```
-### DMN for two timepoints, adjusted for age, sex, mean FD
-```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
-```
-### DMN for three timepoints, adjusted for age, sex, mean FD
-```{r calc DMN three tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+```{r vif for models, echo=FALSE}
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
+vif_res=vif(test_vif)
 ```
 
-### reward
-```{r calc rew two tp, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(rew_FC ~ timepoint*group + (1|subj))
-R0 = lme4::lmer(rew_FC ~ timepoint + group + (1|subj))
+```{r calc DMN two tp,results = 'asis', echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tp=anova(R1,R0)
 ```
-### reward for two timepoints, adjusted for age, sex, mean FD
-```{r calc rew two tp adjusted, eval = FALSE, echo = FALSE}
-R1 = lme4::lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
+
+```{r DMN2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
-### reward for three timepoints, adjusted for age, sex, mean FD
-```{r calc rew three tp, eval = FALSE, echo = FALSE}
-R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
-R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
+
+### DMN for two timepoints, adjusted for age, sex, mean FD
+```{r calc DMN two tp adjusted, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tpadj=anova(R1,R0)
+res=summary(R1)
+#tab_model(R0,R1)
+```
+
+```{r DMN2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","mean FD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
+### DMN for three timepoints, adjusted for age, sex, mean FD
+```{r calc DMN three tp, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
+anvDMN3tp=anova(R1,R0)
+```
+
+```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","mean FD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
+```{r DMN conn over three timepoints, echo=FALSE}
+fig_DMNconn<-mk_figDMNdescr(final_FC)
+fig_DMNconn
+```
+
+### Reward network for two time points
+```{r calc rew two tp, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvRew2tp=anova(R1,R0)
+```
+
+```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
+### Reward network for two timepoints, adjusted for age, sex, mean FD
+```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvRew2tpadj=anova(R1,R0)
+```
+```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","mean FD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
+### Reward network for three timepoints, adjusted for age, sex, mean FD
+```{r calc rew three tp, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
+anvRew3tp=anova(R1,R0)
+```
+```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","mean FD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+```{r Rew conn over three timepoints, echo=FALSE}
+fig_Rewconn<-mk_figRewdescr(final_FC)
+fig_Rewconn
 ```

### 2021-01-07 report/figures.R

@@ -20,10 +20,10 @@ mk_figBMIdescr <- function(final) {
       color = c(rep("#00305E", 3), rep("#C00045", 3))
     ) +
     facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-  figBMIdescr +
+  figBMIdescr= figBMIdescr +
     scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
     scale_color_manual(
-      values = c("#00305E77", "#C0004577"),
+      values = c("#00305E", "#C00045"),
       labels = c("intervention", "control")
     ) +
     xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
@@ -35,6 +35,82 @@ mk_figBMIdescr <- function(final) {
       legend.position = "none"
     )
   return(figBMIdescr)
+}
+
+mk_figRewdescr <- function(final_FC) {
+  
+  condition.labs <- c("BARS","NBARS")
+  names(condition.labs) <- c("IG","KG")
+  # BMI over time each group
+  figRewdescr <-
+    ggplot(final_FC, aes(
+      x = tp,
+      y = mean_Rew_conn,
+      group = subj.ID,
+      color = condition
+    )) +
+    geom_line() + geom_point() +
+    stat_summary(
+      aes(group = condition),
+      fun = mean,
+      geom  = "line",
+      size = 2,
+      color = c(rep("#00305E", 3), rep("#C00045", 3))
+    ) +
+    facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
+  figRewdescr=figRewdescr +
+    scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+    scale_color_manual(
+      values = c("#00305E", "#C00045"),
+      labels = c("intervention", "control")
+    ) +
+    xlab("Time after intervention [months]") + ylab ("mean reward connectivity") +
+    theme_bw() +  theme(
+      axis.text = element_text(size = 10),
+      axis.title = element_text(size = 12),
+      strip.text = element_text(size = 12),
+      legend.text = element_text(size = 10),
+      legend.position = "none"
+    )
+  return(figRewdescr)
+}
+
+mk_figDMNdescr <- function(final_FC) {
+  
+  condition.labs <- c("BARS","NBARS")
+  names(condition.labs) <- c("IG","KG")
[…]

### 2021-01-07 manuscript/draft_rsfmri.Rmd
fix figure design

--- 2021-01-06 manuscript/draft_rsfmri.Rmd
@@
 ## Functional preprocessing
+
 ### Minimal functional preprocessing
@@
 We performed a sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI.
-Again, the time-by-group point interaction was considered significant if a model comparison between R1 and Null models using anova-function yielded a p<0.05.
+Again, the time-by-group point interaction was considered significant if a model comparison between R1 and Null models using `anova` function yielded a $p < 0.05$.
 Finally, we estimated the above-mentioned interaction model for all three time points.
-Again, the interaction was considered significant if a model comparison between R1 and R0 using `anova` had p<0.05.
+Again, the interaction was considered significant if a model comparison between R1 and R0 using `anova` had $p < 0.05$.
 
@@
 (FigureList$fig1 | FigureList$fig2) / FigureList$fig3
-
 ```
@@
 
-### Summary
+#### Summary
 Utilizing resting-state functional MRI on morbidly obese individuals that either underwent or awaited bariatric surgery over the course of one year.
@@
 <!-- model specific --> Looking at time-by-group interactions, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
-Beyond that, also main effects of the intervention and of the time were examined.
+Beyond that, main effects of the intervention and of the time were examined.
 The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
@@
 
-### Support of original hypotheses
+#### Support of original hypotheses
 <!-- false positive -->

### 2021-01-07 report/figures.R

@@ -19,8 +19,7 @@ mk_figBMIdescr <- function(final) {
       size = 2,
       color = c(rep("#00305E", 3), rep("#C00045", 3))
     ) +
-    facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
-  figBMIdescr +
+    facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points") +
     scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
     scale_color_manual(
       values = c("#00305E77", "#C0004577"),

### 2021-01-07 manuscript/draft_rsfmri.Rmd

@@
 
-```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD'}
-# ,  fig.align='center', }
-# eval = TRUE, fig.asp = 0.7, fig.width = 6
+```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
+final=read.csv(file = "./report/all_QA_measures.csv")
 
@@
 
-```{r tableQcfc, echo=FALSE}
-res <- read.csv("./report/FDFCcorrelations.csv")
-rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
-colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
-
-tab_qc_fc <- mk_tableQcfc(res)
-tab_qc_fc
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
+networks=c("../report/DMN_thresholded_for_clusterFWE005.png","../report/Rew_thresholded_for_clusterFWE005.png")
+knitr::include_graphics(networks)
 ```
 
-Further, we evaluated the impact of our denoising pipelines.
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and connectivity was significantly positively correlated (mean Spearman's $r = $`r round(res[1,1],2)`), with a strong distance dependence (mean Spearman's $r = $`r round(res[1,5],2)`).
-Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and also alleviated the distance dependency.
-AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r = $`r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r = $`r round(res[4,5],2)`).
-Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
+Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC values.
+Both networks included the typically described brain regions (i.e.
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulare, right amygdala and ventromedial-prefrontal cortex for the reward network).
+
+```{r prepare tsnr ROI plot, echo=FALSE}
+final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
+
+tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
+tsnr=tsnr[tsnr$V1!="subj",]
+colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+levels(tsnr$roi)=droplevels(tsnr$roi)
+levels(tsnr$roi)=c("NAcc", "precuneus")
+tsnr$mean=as.double(tsnr$mean)
+
+final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
+
+p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") +
+  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
+  theme(axis.text=element_text(size=10),
+  axis.title=element_text(size=12),
+  strip.text = element_text(size=12),
+  legend.position="")
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%",  echo=FALSE}
+p
+```
+Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref{fig:figtSNR}.
 
@@
 
+<!-- Neurovault -->
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
+
 ### DMN/reward network with GM mask (exploratory analysis)

### 2021-01-07 manuscript/draft_rsfmri.Rmd

@@
 ```
+
+\newpage
+# Abstract
+
+\newpage
+# Introduction
 
@@
 source("report/figures.R")
+
 fig_BMIdescr <- mk_figBMIdescr(final)
 fig_BMIdescr
+
 ```
@@
 #### with adjustment for age, sex, meanFD
-There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
-Nor was there a significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+
+<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
+There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc seeds against our initial hypothesis.
+There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
 
 #### with adjustment for age and sex
-Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
-Results did not differ between analyses of connectivity maps from different preprocessing pipelines.
+Similarly, there were no significant clusters when leaving out mean FD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+
 <!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
@@
 
+# Discussion
 In our sample, extend of micromovements of the head averaged across all measurements is associated with increased functional connectivity between reward and frontal regions.

### 2021-01-07 manuscript/draft_rsfmri.Rmd
merged changes

--- 2021-01-07 manuscript/draft_rsfmri.Rmd
@@
 ```{r FDFC correlation, eval=FALSE, echo=FALSE}
+# do not evaluate if saved already
 source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
@@
 ```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
-networks=c("../report/DMN_thresholded_for_clusterFWE005.png","../report/Rew_thresholded_for_clusterFWE005.png")
+#setwd(knitr::opts_knit$get("root.dir"))
+networks = c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png",
+             "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png")
 knitr::include_graphics(networks)
@@
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%",  echo=FALSE}
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%", fig.align="center", echo=FALSE}
 p

### 2021-01-07 manuscript/draft_rsfmri.Rmd

@@
 library(knitr) # version 1.30
+
 ```
@@
 setwd(knitr::opts_knit$get("root.dir"))
+
+source("report/figures.R")
+source("report/tables.R")
+source("report/create_sample_df.R")
 ```
@@
 ```{r "Load data", include=FALSE}
-source("report/create_sample_df.R")
 final = create_sample_df(group = "all", tp = "all")
@@
 ```{r "tableSample", echo=FALSE}
-source("./report/tables.R")
 tab_sample <- mk_SampleTable(final)
@@
 The procedures followed were in accordance with the Helsinki Declaration.
-This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/mzyxv](https://osf.io/mzyxv).
+This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/yp42s](https://osf.io/yp42s).
+<!-- previous version: https://osf.io/f8tpn, update: https://osf.io/59bh7/ -->
 <!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
@@
 
-<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals (n = `r nrow(subj_df)`, `r nrow(filter(subj_df,Sex == "f"))` female; ages `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), that either underwent surgery (n = `r cs["Sum","KG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female).
+<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals 
+(n = `r nrow(subj_df)`, 
+`r nrow(filter(subj_df,Sex == "f"))` female; 
+`r round(mean(subj_df$Age_BL),2)` 
+$\pm$ `r round(sd(subj_df$Age_BL),2)` SD years old
+(range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), 
+that either underwent surgery 
+(n = `r cs["Sum","IG"]`, 
+`r cs["f","IG"]` female, 
+`r round(mean(subj_df$Age_BL[subj_df$condition == "IG"]),2)` 
+$\pm$ `r round(sd(subj_df$Age_BL[subj_df$condition == "IG"]),2)` SD years old) 
+or were waiting list controls 
+(n = `r cs["Sum","KG"]`, 
+`r cs["f","KG"]` female,
+`r round(mean(subj_df$Age_BL[subj_df$condition == "KG"]),2)`
+$\pm$ `r round(sd(subj_df$Age_BL[subj_df$condition == "KG"]),2)` SD years old) 
 The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
@@
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-source("report/figures.R")
 fig_BMIdescr <- mk_figBMIdescr(final)
@@
 ### Global Signal regression   
-[@Ciric_2017] showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+@Ciric_2017 showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
 Yet, it introduces spatial dependency and spurious correlations.
@@
 # fig.show="hold",  out.width="40%", fig.height = 8,
-source("report/figures.R")
 PlotList <- mk_figDesignMatrix()
@@
 Both networks included the typically described brain regions (i.e.
-medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulare, right amygdala and ventromedial-prefrontal cortex for the reward network).
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
 
 ```{r prepare tsnr ROI plot, echo=FALSE}
-final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
-
-tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
-tsnr=tsnr[tsnr$V1!="subj",]
-colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
-levels(tsnr$roi)=droplevels(tsnr$roi)
-levels(tsnr$roi)=c("NAcc", "precuneus")
-tsnr$mean=as.double(tsnr$mean)
-
-final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
-
-p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") +
-  scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
-  theme(axis.text=element_text(size=10),
-  axis.title=element_text(size=12),
-  strip.text = element_text(size=12),
-  legend.position="")
+fig_tSNR <- mk_figtSNR(final)
 ```
@@
 ```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%", fig.align="center", echo=FALSE}
-p
+
+fig_tSNR
 ```
@@
 ```{r tableBMImodel}
-source("report/tables.R")
 TableList <- mk_FCTables()

### 2021-01-07 manuscript/intro_rsfmri.Rmd

@@
   Day Clinic for Cognitive Neurology, University Clinic Leipzig
-output: word_document
+output:
+  pdf_document: default
+  word_document: default
 editor_options:
@@
 
-```{r "knit setup"}
+```{r "knit setup", echo = FALSE}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
@@
 In addition, we will investigate potential changes in within-network functional connectivity after surgery in other resting-state networks such as default mode network.
-Participants are measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
+Participants are measured at baseline, at 6 months and at 12 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
 Notably, we will consider that different courses in increase and decrease of functional connectivity that are not necessarily linear are possible over the course of one year, depending on the phase of weight management [@Olivo_2017].

### 2021-01-14 manuscript/draft_rsfmri.Rmd
add pause option to inspect contrasts

@@
 
+```{r "Descriptive results", include=FALSE}
+m = lme4::lmer(BMI ~ group*visit + (visit+1|subj.ID), data=final)
+summary(m)
+confint(m, method='profile', oldNames=FALSE)
+```
 ## Interaction effect of intervention and time (preregistered analysis)

### 2021-01-14 manuscript/intro_rsfmri.Rmd
add pause option to inspect contrasts

@@
 <!-- Normalization of connectivity --> Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions).
-But these differences seemed to normalize, as trends reversed when looking at the resting state connectivity after compared to before the surgery [@Duan_2020].
+These abnormalities in resting-state connectivity in obese compared to normal-weight individuals seemed to normalize, as trends reversed after compared to before the surgery [@Duan_2020].
 Resting state activitiy increased in subcortical regions, such as the insula and striatum<!-- specifically putamen --> [@Wiemerslage_2016].

### 2021-01-14 code/run_or_display.m

@@ -102,7 +102,7 @@ roi_prep = readcell(fullfile(param.INFO_DIR,'ROIs.txt'), 'Delimiter',' ','Whites
 
 %% ------------------------------------------------------------------------
 % define ROI
-param.ROI_PREP = {roi_prep{[6]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'}
+param.ROI_PREP = {roi_prep{[4]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'}
 
 % Model definition
 % All three models have unique options for covariate definition, the
@@ -120,7 +120,7 @@ param.INFERENCE_TYPE = {'cluster'}; % {'voxel','cluster','tfce'};
 % analysis parameter (estimate or display?)
 param.ONLY_DISPLAY = true;         % false
 param.OVERWRITE = false;             % false
-param.VIEWSEC = 5; % for ONLY_DISPLAY: seconds you want to view the results
+%param.VIEWSEC = 5; % for ONLY_DISPLAY: seconds you want to view the results
 % param.ACTION = 'estimate' % 'display' or 'overwrite'
 
 %% ------------------------------------------------------------------------

### 2021-01-14 report/figures.R

@@ -1,6 +1,8 @@
 
 pink_blue = c("#00305E", "#C00045")
 blue_organge = c("#046C9A", "#D69C4E")
+
+# BMI --------------------------------------------------------------------------
 mk_figBMIdescr <- function(final) {
 
   condition.labs <- c("BARS","NBARS")
@@ -36,9 +38,11 @@ mk_figBMIdescr <- function(final) {
       legend.text = element_text(size = 10),
       legend.position = "none"
     )
+  # ggsave("/data/pt_02161/Publications/Abstracts/Heinrichs_OHBM2021/bmi.pdf", units = "cm", width=15, height=8)
   return(figBMIdescr)
 }
 
+# Reward -----------------------------------------------------------------------
 mk_figRewdescr <- function(final_FC) {
 
   condition.labs <- c("BARS","NBARS")
@@ -77,6 +81,7 @@ mk_figRewdescr <- function(final_FC) {
   return(figRewdescr)
 }
 
+# DMN --------------------------------------------------------------------------
 mk_figDMNdescr <- function(final_FC) {
 
   condition.labs <- c("BARS","NBARS")
@@ -115,6 +120,7 @@ mk_figDMNdescr <- function(final_FC) {
   return(figDMNdescr)
 }
 
+# Design matrix ----------------------------------------------------------------
 mk_figDesignMatrix <- function() {
 
   # design matrix time as factor
@@ -210,8 +216,8 @@ mk_figtSNR <- function (final) {
   return(p)
 }
 
-# ------------------------------------------------------------------------------
-# exploratory and example plots
+# exploratory and example plots ------------------------------------------------
+
 mk_otherplots <- function(final){
   # BMI over time per group
   tspag <-

### 2021-01-21 report/figures.R
correct missspecification of  time points for one subject

@@ -42,6 +42,108 @@ mk_figBMIdescr <- function(final) {
   return(figBMIdescr)
 }
 
+mk_figBMIdescr_for_OHBM <- function(final) {
+  
+  condition.labs <- c("BS","NBS")
+  names(condition.labs) <- c("IG","KG")
+  # BMI over time each group
+  figBMIdescrOHBM <-
+    ggplot(final, aes(
+      x = tp,
+      y = BMI,
+      group = subj.ID,
+      color = condition
+    )) +
+    geom_line(position=position_nudge(x = -0.02)) + geom_point(position=position_nudge(x = -0.02)) +
+    stat_summary(
+      aes(group = condition),
+      fun = mean,
+      geom  = "line",
+      size = 1,
+      position=position_nudge(x = 0.02, y = 0),
+      color = c(rep("#046C9A", 3), rep("#D69C4E", 3))
+    ) +
+    stat_summary(
+      aes(group = condition),
+      fun= mean,
+      fun.min = function(x) mean(x) - sd(x), 
+      fun.max = function(x) mean(x) + sd(x),
+      geom  = "errorbar",
+      size = 1, width = 0.1,
+      position=position_nudge(x = 0.02, y = 0),
+      color = c(rep("#046C9A", 3), rep("#D69C4E", 3))
+    ) +
+    facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) +
+    # geom_errorbar(aes(ymin=BMI-sd(BMI),ymax=BMI+sd(BMI))) +
+    xlab("time points") +
+    scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
+    scale_color_manual(
+      values = c("#046C9A70", "#D69C4E70"),
+      labels = c("intervention", "control")
+    ) +
+    xlab("Time after intervention [months]") + ylab ("BMI [kg/m²]") +
+    theme_bw() +  theme(
+      axis.text = element_text(size = 10),
+      axis.title = element_text(size = 12),
+      strip.text = element_text(size = 12),
+      legend.text = element_text(size = 10),
+      legend.position = "none"
+    )
+  #ggsave("/data/pt_02161/Publications/Abstracts/HeinrichsBeyer_OHBM2021/bmi.pdf", units = "cm", width=15, height=8)
+  return(figBMIdescrOHBM)
+}
+
+mk_figFDdescr_for_OHBM <- function(final) {
+  
+  condition.labs <- c("BS","NBS")
[…]

### 2021-01-26 manuscript/draft_rsfmri.Rmd
correct missspecification of  time points for one subject
enable parallel processing for evaluation of models; script for glass brain; fix wrong time point allocation of subject 61

--- 2021-01-07 manuscript/draft_rsfmri.Rmd
@@
 ```{r "Load data", include=FALSE}
-final = create_sample_df(group = "all", tp = "all")
+final <- create_sample_df(group = "both", tp = "all")
 
@@
 ```{r tableQcfc, echo=FALSE}
-res <- read.csv("./report/FDFCcorrelations.csv")
+res <- read.csv("report/FDFCcorrelations.csv")
 rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
@@
 ```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
-final=read.csv(file = "./report/all_QA_measures.csv")
+final=read.csv(file = "report/all_QA_measures.csv")
 
@@
 FDFC=calc_qc_fc(final)
-write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
+write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
 ```
@@
  -->
+```{r "BMIagesexfd_PCCcc_avgBMI_deact", out.width = "50%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
+```
+
+```{r "BMIagesex_PCCcc_avgBMI_deact", out.width = "85%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf")
+```
 
@@
 We did not find a any clusters for functional connectivity with the PCC.
+```{r "FDagesex_NACCcc_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCcc_avgFD_act.pdf")
+```
+
+```{r "FDagesex_NACCgsr_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCgsr_avgFD_act.pdf")
+```
 
@@
 <!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
+```{r "BMIFDagesex_NACCcc_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACCcc_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
+```
+
+```{r "BMIFDagesex_NACCgsr_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACCgsr_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
+```
+
+```{r "BMIFDagesex_PCCcc_avgBMI_deact", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
 

### 2021-01-26 code/build_runs.m

@@ -16,13 +16,18 @@ for l = 1:length(param.MODEL)
             ctype = "BMI";
         elseif startsWith(cmodel,"fd") 
             ctype = "FD";
+        elseif startsWith(cmodel,"all") 
+            ctype = "ALL";
+        elseif startsWith(cmodel,"single") 
+            ctype = "SINGLE";
         else
             error("invalide model name")
         end
 
         if ((strcmp(ctype,"GT") && (ccov < 10 || ccov >= 20) ) || ...
                 (strcmp(ctype,"BMI") && (ccov < 20 || ccov >= 30))|| ...
-                (strcmp(ctype,"FD") && (ccov < 30 || ccov >= 40)) ) 
+                (strcmp(ctype,"FD") && (ccov < 30 || ccov >= 40)) || ...
+                ((strcmp(ctype,"ALL") || strcmp(ctype,"SINGLE")) && (ccov < 40 || ccov >= 50)) ) 
             continue;
         end
 
@@ -30,6 +35,7 @@ for l = 1:length(param.MODEL)
             for m = 1:length(param.ROI_PREP)
                 
                 crun.ONLY_DISPLAY = param.ONLY_DISPLAY;
+                crun.VIEW = param.VIEW;
                 crun.OVERWRITE = param.OVERWRITE;
                 crun.OUT_DIR = param.OUT_DIR;
                 crun.INFO_DIR = param.INFO_DIR;
@@ -39,7 +45,11 @@ for l = 1:length(param.MODEL)
                 crun.MASK = param.MASK; 
                 crun.EXCLFD = param.EXCLFD;
                 crun.WILD_BOOT = param.WILD_BOOT;
-                crun.INFERENCE_TYPE = param.INFERENCE_TYPE;
+                if ~crun.WILD_BOOT 
+                    crun.INFERENCE_TYPE = "parametric"; 
+                else
+                    crun.INFERENCE_TYPE = param.INFERENCE_TYPE;
+                end
                 crun.TYPE = ctype;
                 crun.MODEL = param.MODEL(l);
                 crun.COVARIATES = param.COVARIATES(j);

### 2021-01-26 code/process_run.m

@@ -1,5 +1,6 @@
 function process_run(crun)
-% one estimation at a time
+% calls model specific function to perform one estimation at a time.
+% called by run_or_display.m
 
 if crun.TYPE == "GT"
     RunModelGroupTime(crun)
@@ -7,6 +8,10 @@ elseif crun.TYPE == "BMI"
     RunModelBMI(crun)
 elseif crun.TYPE == "FD"
     RunModelFD(crun)
+elseif crun.TYPE == "ALL"
+    RunModelAllTp(crun)
+elseif crun.TYPE == "SINGLE"
+    RunModelSingleTp(crun)
 else
     error("Invalid specification!")
 

### 2021-01-26 code/run_or_display.m

@@ -71,9 +71,8 @@
 % If DISPLAY_ONLY is set on false and the model has been estimated, the
 % results will only then change, if OVERWRITE is set on true.
 % define relevant input and output directories
-% Especially for results of non-parametric estimation, you need to specify
-% how long you want to look at the results in VIEWSEC. Alternatively you
-% can also set breakpoints in the funtion scripts.
+% Especially for results of non-parametric estimation, enable *VIEW" in 
+% order to look at the results. 
 
 %% ========================================================================
 % MATLAB R2017b not compatible
@@ -81,7 +80,7 @@
 % set path for spm and path with my functions
 % swe version 2.2.1 download of development 
 addpath("/data/pt_02161/Analysis/Software/spm12/")
-addpath(genpath("/data/pt_02161/Analysis/Software/spm12/toolbox"))
+addpath(genpath("/data/pt_02161/Analysis/Software/spm12/toolbox/SwE-toolbox-2.2.2"))
 addpath("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code")
 % swe version 2.1.1
 %addpath(genpath("/data/pt_life/data_fbeyer/spm-fbeyer"))
@@ -103,10 +102,11 @@ param.MASK_B = "MNI_resampled_brain_mask.nii,1";
 roi_prep = convertCharsToStrings(readcell(fullfile(param.INFO_DIR,"ROIs.txt"), "Delimiter"," ","Whitespace","'"));
 %% ------------------------------------------------------------------------
 
-param.PRESET = "bmi";
-param.ONLY_DISPLAY = false;         
-param.OVERWRITE = true; 
+param.PRESET = "manual";
+param.ONLY_DISPLAY = true;         
+param.OVERWRITE = false; 
 param.WILD_BOOT = true;   
+param.parallel = false;
 
 if param.PRESET == "standard"
     param.MODEL = ["grouptime","grouptime2tp", "bmi", "bmiIG","bmi2tp", "fd","fdIG"];
@@ -115,6 +115,7 @@ if param.PRESET == "standard"
     param.MASK = "brain";           
     param.EXCLFD = false;         
     param.INFERENCE_TYPE = ["cluster"];
[…]

### 2021-01-26 report/plot_glassbrain_abstract.py

@@ -4,16 +4,26 @@
 
 from nilearn import plotting
 from os import path
+import pandas as pd
 
 
 def main():
-    ROOT_DIR = '/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c02cl'
 
-    IN_FILE = path.join(ROOT_DIR, 'thresholded_cgnBMI_decrease.nii')
-    OUT_FILE = path.join(ROOT_DIR, 'thresholded_cgnBMI_decrease.pdf')
+    OUT_DIR = "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report"
+    INFO_FILE = "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD/result_report.txt"
 
-    print(f'Processing:\n- in:  {IN_FILE}\n- out: {OUT_FILE}')
-    plot(IN_FILE, OUT_FILE)
+    # dataframe with filenames, path list, cluster threshold
+    df = pd.read_table(INFO_FILE, header=None)
+    df[1] = df[1].str.replace('txt', 'nii') # root_paths  = df[1].tolist()
+
+    for i in range(len(df[1])):
+        IN_FILE = df[1][i]
+        OUT_FILE = path.join(OUT_DIR, (df[0][i] + ".pdf"))
+
+        # do for each result
+        print(f'Processing:\n- in:  {IN_FILE}\n- out: {OUT_FILE}')
+        plot(IN_FILE, OUT_FILE)
+
     print('done!')
 
 

### 2021-01-26 report/tables.R

@@ -2,6 +2,9 @@ DIR_ANALYSIS <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Anal
 
 # FC RESULTS -------------------------------------------------------------------
 mk_FCTables <- function() {
+  # previously csv files with SPM output tables need to be created, maps need
+  # to be inspected with the Anatomical Labelling Toolbox and its results saved 
+  # in a txt file.
   
 # ------------------------------------------------------------------------------
 # Directories
@@ -15,6 +18,7 @@ res_fc$BMImodel_dirs <-
 res_fc$FDBMImodel_dirs <- 
   c("noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c02cl", # deact
     "noExclFD/FD_total/brain/Nacc_cc_z/bmi-fd-age-sex_WB-c03cl", # act
+    "noExclFD/FD_total/brain/Nacc_gsr_z/bmi-fd-age-sex_WB-c02cl", # deact
     "noExclFD/FD_total/brain/Nacc_gsr_z/bmi-fd-age-sex_WB-c03cl", # act
     "noExclFD/FD_total/brain/PCC_cc_z/bmi-fd-age-sex_WB-c01cl") # deact
 res_fc$FDmodel_dirs <- 
@@ -27,26 +31,18 @@ res_fc$act_csv <- "results_act.csv"
 # ------------------------------------------------------------------------------
 # Anatomical Labeling
 # ------------------------------------------------------------------------------
+
+# get Anatomy tables
+# manually created list of all directories of result txt - better solution needed
 f <- file.path(DIR_ANALYSIS,"noExclFD/result_report.txt")
 result_report <- read.table(f, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)
 
 ResultFCList <- list()
 for (r in 1:nrow(result_report)) {
   f = result_report[r,"dir"]
-  txt <- read.table(trimws(f), sep="\t", fill = TRUE)
+  txt <- read.table(trimws(f),sep="\t",fill=TRUE,col.names=paste0("V", 1:13))
[…]

### 2021-01-27 manuscript/draft_rsfmri.Rmd
added introduction

--- 2021-01-07 manuscript/draft_rsfmri.Rmd
@@
 - Frauke Beyer* [1][2]
-- Kerstin Prehn
+- Kristin Prehn
+- Jürgen Ordemann 
 - Agnes Floel
-- Arno Villringer [1][2][3]
 - A. Veronica Witte [1][2][3]
@@
 
+source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/glmm_stability.r')
+source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')
+
 ```
@@
 setwd(knitr::opts_knit$get("root.dir"))
+```
 
-source("report/figures.R")
-source("report/tables.R")
-source("report/create_sample_df.R")
-```
+\newpage
+# Abstract
+
+\newpage
+# Introduction
+Obesity is a worldwide health issue, enchaining huge personal and societal costs.
+Excess amount of fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life.
+Conservative treatment options including dietary modification, physical activity and behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI (> 35 kg/m2).
+Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
+Roux-en-Y gastric bypass (RYGB) is the most effective surgical procedure [@Berthoud_2011].
+Here, a small pouch is formed from the proximal stomach and connected to the jejunum, disconnecting the larger part of the stomach and the proximal part of the small intestine intact from the digestive flow.
+Other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) reduce stomach volume, but leave the small intestine unchanged [@Mulla_2017].
+RYGB-induced weight loss is in part mediated by malabsorption of certain nutrients, especially fat, and reduced digestion efficiency [@Berthoud_2011] while the role of increased energy expenditure after RYGB is still unclear [@Yarmush_2017;@Lamarca_2019].
+Most importantly, RYGB alters the signaling cascades between the gut, fat tissue and the brain and thereby induces profound changes in the regulation of food intake and energy homeostasis [@Berthoud_2011].
+RYBG-patients frequently report altered smell and taste perception, i.e.
+increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items [@Mulla_2017, @Berthoud_2011].
+These changes in appetitive and reward signaling probably go back to altered gut-brain signaling, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin.
+Yet, precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
+Resting state functional magnetic resonance imaging (rsfMRI) is a powerful technique to investigate the dynamic organization of the brain via the correlation of spontaneous neural activity.
+Anatomically distinct brain regions show functional connectivity (FC), i.e.
+correlated neural activity over time, in the absence of a cognitive task.
+Crucially, these brain networks correspond to networks activated during specific cognitive processes, e.g.
+visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e.
+attention and executive control [@Smith_2011], and can thus be seen as default configuration of the brain.
+In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.
+The default mode network has been discovered as a “task-negative” network, and is involved in a variety of largely internally-directed cognitive functions including interoception, mind wandering and memory.
+Its core hubs are the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
+BA 39) [@Raichle_2015].
+Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2012; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
+Despite the large heterogenity and low median power, these findings might be summarized as a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
+salience and sensory networks.
+This has been interpreted as impaired interoceptive processing failing to adequately balance internal and external behavioral cues.
+Longitudinal studies largely support this view.
+Increased activation of the DMN in a food-viewing paradigm, i.e.
+insufficient down-regulation of DMN, has been shown to correlate with appetite measures after an overnight fast [@Tregellas_2011, @Wijngaarden_2015], an effect that could be partially reversed by an exercise intervention [@McFadden_2013].
+In a bariatric surgery study, [@Li_2018] reported a reduction of local FC in mid-line DMN regions in surgical patients compared to an obese control group.
+Simultaneously, PCC and medial prefrontal cortex were more strongly connected to dorsolateral prefrontal cortex, a region implicated in cognitive control.
+Another study in bariatric surgery patients showed that heightened connectivity of the PCC with anterior superior temporal gyrus (STG) before the surgery decreased over the course of one year [@Olivo_2017].
+Similarly, the profile of DMN connectivity with frontal regions, including ACC, frontal superior gyrus and the right OFC,  was similar in patients after bariatric surgery and normal weight controls, but dissimilar to that of obese controls [@Frank_2013].
+Another important brain network in the context of obesity is the reward network, comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
+These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
+Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced reward region activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020;].
+Food cue hyperactivation has been identified as a main neural vulnerability factor predicting future weight gain [@Stice_2015].
+Resting state fMRI studies have partly mirrored these results, showing increased local FC of reward network regions, i.e.
+Nacc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
+Imaging studies in bariatric surgery patients provided first evidence that hyperconnectivity between reward network regions (e.g.
+putamen, OFC) might be normalized to some extent by the surgery [@Duan_2020; @Wiemerslage_2016].
+Possible, the reconfiguration of the fronto-striatal brain networks might be related to altered gut signaling , e.g.
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013, but @Zoon_2018] .
+Thus, while there is some evidence hinting to a role for DMN and reward network connectivity in altered gut-brain communication after bariatric surgery, the existing evidence is inconclusive.
+Most studies have investigated small cohort of patients, often without adequate control group, and employed exploratory statistics [@George_2016].
+In this study, we aimed to extend the current literature and investigate differences of DMN and reward network connectivity before and after bariatric surgery in preregistered analyses and in comparison with a contemporaneously measured waiting-control group.
+Moreover, we explored whether the changes in functional connectivity were associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential dose-response relationship.
+Participants were measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
+Notably, we took into account possible non-linear time courses in increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
 
@@
 ```{r "Load data", include=FALSE}
+source("report/create_sample_df.R")
 final = create_sample_df(group = "all", tp = "all")
@@
 ```{r "tableSample", echo=FALSE}
+source("./report/tables.R")
 tab_sample <- mk_SampleTable(final)
@@
 
-<!-- STUDY DESIGN --> The ADIPOSITAS-Study conducted in the Charité Berlin is a observational, prospective and longitudinal design described in detail by @Prehn_2020.
-Data collection is still ongoing but only data available in April 2019 is used.
-The procedures followed were in accordance with the Helsinki Declaration.
-This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/yp42s](https://osf.io/yp42s).
-<!-- previous version: https://osf.io/f8tpn, update: https://osf.io/59bh7/ -->
-<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
-These were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
+<!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité Berlin.
+For more details see @Prehn_2020.
+We used all data acquired until April 2019.
+The study protocol was in accordance with the Helsinki Declaration and approved by the local ethics committee.
+The study was registered at clinicaltrials.gov as NCT01554228.
+We preregistered the present resting state analyses on the [OSF](https://osf.io/yp42s).
+We made [additional changes](https://osf.io/59bh7/) to the preregistration after preprocessing as we realized some aspects of the analysis were inadequately described in the first preregistration.
+<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> 
+Participants were recruited from the Center for Bariatric and Metabolic Surgery.
+Inclusion criteria were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tableDescr)).
-`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
-The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing only subjects that have received MRI.
+Participants were aged between $18$ and $70$ and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$).
+In total, $51$ participants out of the originally enrolled $69$ subjects received MRI, and we analyzed a total of $101$ data points.
+Five data points of three subjects had to be excluded due to bad anatomical image quality (see section \@ref{sec:Anatomical Quality control}).
+<!-- DESCRIPTIVE STATISTICS --> 
+The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`).
+Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment.
 Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tableSample)).
+In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
+In the intervention group, data is not missing completely at random, as the pre-surgery timepoint MRI could not be obtained in some patients due to the limited scanner bore.
 
-<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals 
-(n = `r nrow(subj_df)`, 
-`r nrow(filter(subj_df,Sex == "f"))` female; 
-`r round(mean(subj_df$Age_BL),2)` 
-$\pm$ `r round(sd(subj_df$Age_BL),2)` SD years old
-(range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), 
-that either underwent surgery 
-(n = `r cs["Sum","IG"]`, 
-`r cs["f","IG"]` female, 
-`r round(mean(subj_df$Age_BL[subj_df$condition == "IG"]),2)` 
-$\pm$ `r round(sd(subj_df$Age_BL[subj_df$condition == "IG"]),2)` SD years old) 
-or were waiting list controls 
-(n = `r cs["Sum","KG"]`, 
-`r cs["f","KG"]` female,
-`r round(mean(subj_df$Age_BL[subj_df$condition == "KG"]),2)`
-$\pm$ `r round(sd(subj_df$Age_BL[subj_df$condition == "KG"]),2)` SD years old) 
-The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
 ```{r tableBaselineCharacteristics, include=FALSE}
@@
 ```
-<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+<!-- BASELINE CHARACTERISTICS --> 
+Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
@@
 ```
+
 <!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
@@
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
+
 ```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
+source("report/figures.R")
+
 fig_BMIdescr <- mk_figBMIdescr(final)
 fig_BMIdescr
+
 ```
@@
 
-## Pre-proceessing
+## Preprocessing
 
@@
 Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rs preprocessing.
-The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@antsref].
+The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants].
 The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
-In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.
+*In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.*
 
@@
 We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
-As described in [@Pruim2015], the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+As described in [@pruim2015ica], the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
 Finally, the timeseries of the  nuisance components were regressed from the raw fMRI timeseries.
@@
 ### Global Signal regression   
-@Ciric_2017 showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+[@Ciric_2017] showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
 Yet, it introduces spatial dependency and spurious correlations.
@@
 To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed regions of interest (ROI), respectively.
+We chose these regions as they are core hubs of the networks.
+Because we knew about the large amount of motion in this dataset, we decided against Independent Component Analysis (ICA) to construct the networks, as ICA is recommended to work on minimally preprocessed data [GSR data](https://sourceforge.net/p/icatb/mailman/message/31466068/).
 Based on FreeSurfer's `aseg.mgz` and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
 Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
+We calculated the Pearson's correlation between them with `numpy.dot`, performed `r-to-z` transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
 Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
@@
 FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
-We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration .
+We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.
 <!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
@@
 #### Functional Quality control
-Resting state fMRI data quality control was performed along the protocol of @Ciric_2018.
-As the expected average head motion is high in this sample [@Beyer_2020], it is difficult to determine an appropriate exclusion threshold which does not lead to the exclusion of a high number of individuals.
-Thus, we did not exclude any participants based on mean framewise displacement estimates beyond those who failed the anatomical preprocessing.
+Resting state fMRI data quality control was performed according to the protocol by @Ciric_2018.
+As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020].
+Thus, we did not exclude any participants based on mean framewise displacement (FD) estimates beyond those who failed the anatomical preprocessing.
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
 In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
+```{r carpetplot, out.width="49%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap=c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
+carpets=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/carpetplotadi006fu2.jpeg","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/carpetplotadi16fu.jpeg")
+knitr::include_graphics(carpets)
+```
 On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
-Figure \@ref{carpet_plot} shows an exemplary carpet plot with (from top to bottom): DVARS (spatial root mean square of the data after temporal differencing), mFD, percent outliers defined by AFNI `OutlierCount`, voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data.
+Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum).
 All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
@@
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{example_FC}.
+Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
 We reported the mFD-FC correlation per preprocessing scheme.
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
-We used global signal regressed data as input and clusterwise bootstrapping with N=999.
+To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and timepoints, adjusted for age and sex.
+We used GSR regressed data as input and clusterwise bootstrapping with N=999.
 Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of p_FWE < 0.05.
-We extracted the average connectivity from these masks for all data points.
+We extracted the average connectivity from these masks for GSR regressed data.
 
@@
 Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
-During analyses an implicit mask and an explicit mask were used.
-For this, FSL's standard space the brain mask that corresponds th "152 nonlinear 6th generation" atlas was re-sampled to 3 x 3 x 3 mm$^3$.<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
--->
 
-<!-- MODEL SETUP --> As specified in the preregistration, we tested the interaction of time and group: $FC = group + time + group*time$.
-Unlike prespecified, we did not include the model without covariates (as this is deprecated \@ref{nuisance_paper}).
-Instead, we calculated a model adjusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2).
-We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
-The model was examined for the correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $8$ a priori planned analyses.
+### Design Specification
 
-Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$.
-As a result a total of $32$ different analyses were performed.
-Computation with the grey matter mask were exploratory and deviate from the preregistration.
-The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.
-
-<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
-It therefore accommodates any repeated measurement covariance structures.
-<!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
-Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.
--->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
-
+<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
+We did not perform the analysis as preregistered (i.e.
+without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
 <!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
---> For the first model, time was introduced as factor because we did not assume strict linearity for the increase in connectivity and therefore refrained from modeling time as continuous variable (see Figure \@ref(fig:figDesignmatrix)).
+--> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)).
 The resulting factorial design <!-- check: flexible factorial?
@@
 # fig.show="hold",  out.width="40%", fig.height = 8,
+source("report/figures.R")
 PlotList <- mk_figDesignMatrix()
@@
 ```
-For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
-With a model containing average BMI and longitudinal change per subject, we investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-<!-- NUISANCE COVARIATES --> Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
+As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
+This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
+We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3).
+Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
+<!-- Changed manuscript here
+investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
+  Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
 Age and sex were between-subject covariates while FD is time-varying.
 Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
-As change in FD and change in BMI share common variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
+ the inclusion of log mean FD may be disputed and done optionally.
 For that reason, we performed each analysis separately one with mean FD (three and without.
@@
 We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
-<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
-As these post-hoc analyses were not preregistered, they should be considered to be exploratory.
+<!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen.
+--> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
 
-<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+### Model input and brain mask
+For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC with GSR), resulting in $8$ preregistered analyses <!-- timexgroup (2 models) * 2* 2 --> , and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
+For all analyses, we used an explicit mask, which was either a whole brain or a GM mask derived from the MNI ICBM "152 nonlinear 6th generation" atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
+This resulted in a reduced number of tests (voxels of interest) from $52378$ to $29451$ for the GM mask.
+<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+-->
+Computation with the grey matter mask were exploratory and not planned in the preregistration.
+
+### SwE settings
+<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+It therefore accommodates any repeated measurement covariance structures.
+<!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
+Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.
+-->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
 Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual. 
-In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
+In contrast to the parametric estimation, wild bootstrap avoids any parametric distributional assumptions (e.g.
+random field theory), and unlike permutation testing it is also suitable <!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
+Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
@@
 
+### Aggregated FC analysis
 <!-- R analysis of aggregated values -->
 We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
-Following the preregistration, we first investigated a time-by-group interaction between baseline and follow-up.
-We used a model comparison of R1 and R0 models with `anova`:
-`R1= lmer(agg FC ~ timepoint*group + (1|subj))`
-`R0 = lmer(agg FC ~ timepoint + group + (1|subj))`
-We performed a sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI.
-Again, the time-by-group point interaction was considered significant if a model comparison between R1 and Null models using `anova` function yielded a $p < 0.05$.
-Finally, we estimated the above-mentioned interaction model for all three time points.
-Again, the interaction was considered significant if a model comparison between R1 and R0 using `anova` had $p < 0.05$.
+Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up.
+In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
+We performed model comparison between R1 and R0 models 
+`R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
+`R0 = lmer(agg FC ~ timepoint + group +  + age + (1|subj))`
+and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
+As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points.
+Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
+Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.
 
@@
 We performed different quality controls for the rsfMRI data.
-As described in @Beyer_2020, group and time show an interaction effect for mFD where the intervention group reduced in head motion in the follow-up assessments compared to the control group.
-We also inspected DVARS, a measure of whole-brain signal variation which is reflective of noise in the data.
-We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
-This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
+First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
+As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group. 
+Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points.
+We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
 ```{r tableQcfc, echo=FALSE}
@@
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
-Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency.
+Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
 Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
@@
 ```{r FDFC correlation, eval=FALSE, echo=FALSE}
-# do not evaluate if saved already
 source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
@@
 
-```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
-#setwd(knitr::opts_knit$get("root.dir"))
-networks = c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png",
-             "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png")
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
+networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
 knitr::include_graphics(networks)
@@
 
-Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC values.
+Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
 Both networks included the typically described brain regions (i.e.
 medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
+Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks).
+```{r prepare tsnr ROI plot, eval=FALSE}
+final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
 
-```{r prepare tsnr ROI plot, echo=FALSE}
-fig_tSNR <- mk_figtSNR(final)
+tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
+tsnr=tsnr[tsnr$V1!="subj",]
+colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+levels(tsnr$roi)=droplevels(tsnr$roi)
+levels(tsnr$roi)=c("NAcc", "PCC/precuneus")
+tsnr$mean=as.double(tsnr$mean)
+
+final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
+
+p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
+  scale_fill_manual(values=wes_palette("Darjeeling1",5,type="discrete")[3:4]) +
+  theme(axis.text=element_text(size=10),
+  axis.title=element_text(size=12),
+  strip.text = element_text(size=12),
+  legend.position="")
+ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", p)
 ```
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%", fig.align="center", echo=FALSE}
-
-fig_tSNR
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%",  eval=FALSE}
+p
 ```
-Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref{fig:figtSNR}.
 
@@
 
-#### with adjustment for age, sex, meanFD
-There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
-Nor was there a significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+#### Model 1a (with adjustment for age, sex)
 
-#### with adjustment for age and sex
-Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
-Results did not differ between analyses of connectivity maps from different preprocessing pipelines.
+<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
+There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc seeds against our initial hypothesis.
+There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+
+####  Model 1b (with adjustment for age and sex, meanFD)
+Similarly, there were no significant clusters when including log mean FD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+
 <!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
@@
 
-### DMN/reward network with GM mask (exploratory analysis)
-<!-- GM mask: Group*Time Interaction [OLD]-->
-<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
-
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
+There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models)
 ```{r load agg FC data and prepare, echo=FALSE}
@@
 final_FC$subj.ID=as.factor(final_FC$subj.ID)
+final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")
+
+#calculate mean BMI over timepoints
+final_FC.meanBMI=tapply(X=final_FC$BMI,
+                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
+final_FC$mean.BMI=
+  final_FC.meanBMI[as.numeric(final_FC$subj.ID)]
+final_FC$within.BMI=
+  final_FC$BMI-final_FC$mean.BMI
 
@@
 
-### DMN for two timepoints
-```{r vif for models, echo=FALSE}
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
+```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
+#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
+
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tp=anova(R1,R0)
+```
+
+```{qa check DMN model age,sex, echo=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
 vif_res=vif(test_vif)
@@
 
-```{r calc DMN two tp,results = 'asis', echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
+```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
+#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
+
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvDMN2tp=anova(R1,R0)
@@
 
-```{r DMN2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```{qa check DMN model age,sex, mFD, echo=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
+vif_res=vif(test_vif)
 ```
 
-### DMN for two timepoints, adjusted for age, sex, mean FD
-```{r calc DMN two tp adjusted, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvDMN2tpadj=anova(R1,R0)
-res=summary(R1)
-#tab_model(R0,R1)
+```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{r DMN2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","mean FD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-
-### DMN for three timepoints, adjusted for age, sex, mean FD
 ```{r calc DMN three tp, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
+#### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
 anvDMN3tp=anova(R1,R0)
@@
 
-```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
 #tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","mean FD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{r DMN conn over three timepoints, echo=FALSE}
+```{r figaggDMN, echo=FALSE}
 fig_DMNconn<-mk_figDMNdescr(final_FC)
@@
 
-### Reward network for two time points
-```{r calc rew two tp, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
+### Reward network for two time points, adjusted for age and sex (Model 1a)
+
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvRew2tp=anova(R1,R0)
 ```
+```{qa check Rew model age,sex, echo=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
 
-```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
 ```
 
-### Reward network for two timepoints, adjusted for age, sex, mean FD
+```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
 ```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
+#### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
+
 R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
@@
 ```
-```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","mean FD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+
+```{qa check Rew model age,sex,mFD, echo=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
+vif_res=vif(test_vif)
 ```
 
-### Reward network for three timepoints, adjusted for age, sex, mean FD
+```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
 ```{r calc rew three tp, eval = TRUE, echo = FALSE}
+#### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
+
 R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
@@
 ```
-```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
 #tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","mean FD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
-```{r Rew conn over three timepoints, echo=FALSE}
+```{r figaggRew, echo=FALSE}
 fig_Rewconn<-mk_figRewdescr(final_FC)
@@
 
-## Effect of mean and change in BMI (exploratory analysis)
-
+## Exploratory Analyses
+### Effect of mean and change in BMI 
+#### Whole brain analysis
 Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
 <!-- ### DMN -->
-<!-- #### DMN for three timepoints -->
-<!-- cc -->Using the images preprocessed with AROMA+CC, we did find a significant co-activation with the PCC in three clusters associated with average BMI.
-Omitting mean FD as nuisance covariate yielded similar results but four significant clusters.
-However, in both, there is significant peak activation within the largest cluster.
-<!-- gsr -->In analyses on images that were additionally preprocessed with GSR, no cluster survived statistical thresholds for either ROI.
-All results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
+<!-- cc -->
+Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate.
+When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
+NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes. 
+Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
+To disentangle effects of change in BMI and logmFD in this dataset, we also split mFD into average mFD and change in mFD. Age and sex remained as nuisance covariates.
+<!-- ### DMN -->
+<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+Test statistics of these clusters differed only slightly from the effect found for average BMI when using mFD as nuisance covariate.
+<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.
+
+<!-- ### Reward -->
+<!-- cc -->
+For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
+supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
+Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD.
+When running the model with GSR-preprocessed data, the associations between change in BMI and FC did not survive statistical thresholds.
+The relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
+
+#### Aggregated FC analysis
+```{r DMN model, mean, change BMI, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
+
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
+R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
+anvmeanBMI_adjmFD=anova(R1,R02a)
+```
+Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD. 
+```{r Rew model, mean, change BMI,, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+```
+
+```{r DMN model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvmeanFD=anova(R1,R03)
+anvwithinFD=anova(R1,R03)
+
+p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+```
+The association also remained significant when we split logmFD into average and change in logmFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in logmFD with DMN FC.
+
+#### Reward network, mean and within variability in BMI and logmFD
+```{r Rew model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvwithinFD=anova(R1,R03)
+anvmeanFD=anova(R1,R04)
+
+test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
+vif_res=vif(test_vif)
+``` 
+As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and logmFD.
+
+<!-- [OLD stuff] GM mask: Group*Time Interaction -->
 <!-- ### DMN for two timepoints
@@
  --> 
+<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
+Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
+While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
 
-## Effect of FD (exploratory analysis)
-
+### Effect of FD (exploratory analysis)
+#### Whole-brain analysis
 To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models. 
@@
 
-## Effect of BMI and FD (exploratory analysis)
+<!--### Effect on hypothalamic connectivity (exploratory analysis)-->
 
-Average BMI and change in BMI remain covariates of interest.
-While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as.
-Age and sex remain nuisance covariates.
-
-<!-- ### Reward -->
-<!-- cc -->There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
-Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
-<!-- gsr -->When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds.
-On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
-
-<!-- ### DMN -->
-<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
-There is a significant association between average BMI and the connecitivity between PCC and three other clusters.
-Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
-<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
-
-## Effect on hypothalamic connectivity (exploratory analysis)
-...
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
-
-```{r tableBMImodel}
+```{r tableBMImodel, echo=TRUE}
+source("report/tables.R")
 TableList <- mk_FCTables()
@@
 
-```{r tableFDmodel}
+```{r tableBMIFDmodel, eval=FALSE}
+tab_BMIFDmodel <- mk_FCTables()[[3]]
+tab_BMIFDmodel
+```
+
+```{r tableFDmodel, eval=FALSE}
 tab_FDmodel <- mk_FCTables()[[2]]
 tab_FDmodel
-```
-
-```{r tableBMIFDmodel}
-tab_BMIFDmodel <- mk_FCTables()[[3]]
-tab_BMIFDmodel
 ```

### 2021-01-27 manuscript/supplements_rsfmri.Rmd
added introduction

---
title: "Supplementary Material for Effects of bariatric surgery and weight loss on resting state functional connectivity"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Kristin Prehn
- Jürgen Ordemann 
- Agnes Floel
- A. Veronica Witte [1][2][3]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
  Day Clinic for Cognitive Neurology, University Clinic Leipzig
output:
  bookdown::pdf_book:
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: yes
    extra_dependencies: "subfig"
bibliography: "bibliography.bib"
editor_options:
  chunk_output_type: console
---

```{r "load packages", include=FALSE}

# load packages
library(car) # version 3.0.9
library(plyr) # version 1.8.6
library(dplyr) # version 1.0.2
library(tidyr) # version 1.1.2
library(kableExtra) # version 1.2.1
library(ggplot2) # version 3.3.2
library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
library(cowplot)
library(RColorBrewer) # version 1.1.2
library(psych) # version 2.0.8

#library(lsr)
library(wesanderson) # colour palette
#library(haven)

library(lme4)

library(sjPlot) #only works for HTML output
library(texreg) #works for Rmarkdown
library(knitr) # version 1.30

source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/glmm_stability.r')
source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')

```

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex") 
options(knitr.kable.NA = '') # hide NA in tables
# set working directory of knit (RMarkdown) to directory of Rproj
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(knitr::opts_knit$get("root.dir"))
```

## Interaction effect group by time on aggregated FC (preregistered analysis)
```{r load agg FC data and prepare, echo=FALSE}
agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
for (i in c(1:nrow(agg_FC_DMN_ID))){
  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
}

colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")

final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
final_FC$subj.ID=as.factor(final_FC$subj.ID)
final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")

#calculate mean BMI over timepoints
final_FC.meanBMI=tapply(X=final_FC$BMI,
                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
final_FC$mean.BMI=
  final_FC.meanBMI[as.numeric(final_FC$subj.ID)]
final_FC$within.BMI=
  final_FC$BMI-final_FC$mean.BMI

#calculate mean FD over timepoints
final_FC.meanFD=tapply(X=final_FC$logmFD,
                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
final_FC$mean.logmFD=
  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
final_FC$within.logmFD=
  final_FC$logmFD-final_FC$mean.logmFD
```

### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```

```{qa check DMN model age,sex, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvDMN2tp=anova(R1,R0)
```

```{qa check DMN model age,sex, mFD, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc DMN three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
anvDMN3tp=anova(R1,R0)
```

```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r DMN conn over three timepoints, echo=FALSE}
fig_DMNconn<-mk_figDMNdescr(final_FC)
fig_DMNconn
```

### Reward network for two time points, adjusted for age and sex (Model 1a)
```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tp=anova(R1,R0)
```
```{qa check Rew model age,sex, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary
```

```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
anvRew2tpadj=anova(R1,R0)
```

```{qa check Rew model age,sex,mFD, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

m.stab=glmm.model.stab(model.res = R0)
#to check whether there were any converge issues:
table(m.stab$detailed$warnings)
m.stab$summary

test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
vif_res=vif(test_vif)
```

```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
```{r calc rew three tp, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```
```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```
```{r Rew conn over three timepoints, echo=FALSE}
fig_Rewconn<-mk_figRewdescr(final_FC)
fig_Rewconn
```

### 2021-01-27 manuscript/draft_rsfmri.Rmd
modification of structure & figures after re-categorization of one participant

@@
 visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e.
-attention and executive control [@Smith_2011], and can thus be seen as default configuration of the brain.
+attention and executive control [@Smith_2009], and can thus be seen as default configuration of the brain.
 In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.
@@
 BA 39) [@Raichle_2015].
-Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2012; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
+Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
 Despite the large heterogenity and low median power, these findings might be summarized as a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
@@
 ```{r "Load data", include=FALSE}
-source("report/create_sample_df.R")
-final = create_sample_df(group = "all", tp = "all")
+final <- create_sample_df(group = "both", tp = "all")
 
@@
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-source("report/figures.R")
-
 fig_BMIdescr <- mk_figBMIdescr(final)
 fig_BMIdescr
-
 ```
@@
 ```{r FDFC correlation, eval=FALSE, echo=FALSE}
+# do not evaluate if saved already
 source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
 FDFC=calc_qc_fc(final)
-write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
-```
-
-```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
-#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
-networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
-knitr::include_graphics(networks)
-```
-
-Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
-Both networks included the typically described brain regions (i.e.
-medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
-Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks).
+write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
+```
+
 ```{r prepare tsnr ROI plot, eval=FALSE}
@@
 p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
-  scale_fill_manual(values=wes_palette("Darjeeling1",5,type="discrete")[3:4]) +
+  geom_violin(aes(fill=roi), alpha=0.5) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
+  scale_fill_manual(values=wes_palette("Darjeeling2",5,type="discrete")[2:3]) +
   theme(axis.text=element_text(size=10),
@@
   legend.position="")
-ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", p)
-```
+ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", height = 6.604, units="cm", p)
+```
+
+```{r figNetworks, echo=FALSE, out.width="33%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
+networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png")
+tsnr=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
+knitr::include_graphics(networks)
+knitr::include_graphics(tsnr)
+```
+
+Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
+Both networks included the typically described brain regions (i.e.
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
+Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks)).
 
@@
 <!-- Brain mask: Group*Time Interaction [CORRECTED] -->
-There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc seeds against our initial hypothesis.
+Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the PCC nor the NAcc seeds.
 There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
@@
 ####  Model 1b (with adjustment for age and sex, meanFD)
-Similarly, there were no significant clusters when including log mean FD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 
@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
-There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models)
 ```{r load agg FC data and prepare, echo=FALSE}
@@
 ```
+```{r figaggDMN, fig.cap="DMN connectivity per group over time", out.width="75%", echo=FALSE}
+fig_DMNconn<-mk_figDMNdescr(final_FC)
+fig_DMNconn
+```
+```{r figaggRew,fig.cap="Reward network connectivity per group over time", out.width="75%", echo=FALSE}
+fig_Rewconn<-mk_figRewdescr(final_FC)
+fig_Rewconn
+```
+
+```{r identify outliers in these plots, echo=FALSE}
+### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
+final_FC[final_FC$mean_DMN_conn<0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")] 
+#ADI003_fu2 & ADI041_fu -> do not have particularly high head motion 
+#plot(final_FC$mean_DMN_conn, final_FC$logmFD) -> see two left points
+#carpet plots look ok, brain maps of DMN look okish, though not pronounced
+#ADI003_fu2 & ADI041 fu: FS issues (Final_score:3, corrected)
+qa_info=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
+qa_info[qa_info$subj.ID=="ADI003"&qa_info$tp=="fu2",]
+qa_info[qa_info$subj.ID=="ADI041"&qa_info$tp=="fu",]
+
+final_FC[final_FC$mean_Rew_conn>0.3, c("mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
+plot(final_FC$mean_Rew_conn, final_FC$logmFD)
+#-> do not have particularly high head motion (two right points)
+#ADI069_bl
+#nifti looks ok, carpet plot shows lot of stripes.
+qa_info[qa_info$subj.ID=="ADI069"&qa_info$tp=="bl",]
+#FS issues, parts of skull included
+#ADI111_fu2
+#nifti looks ok, some residual carpet plot noise
+qa_info[qa_info$subj.ID=="ADI111"&qa_info$tp=="fu2",]
+#no FS issues
+```
+
+There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models).
 
@@
 
-```{r figaggDMN, echo=FALSE}
-fig_DMNconn<-mk_figDMNdescr(final_FC)
-fig_DMNconn
-```
-
 ```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
@@
 ```
-```{r figaggRew, echo=FALSE}
-fig_Rewconn<-mk_figRewdescr(final_FC)
-fig_Rewconn
-```
 
@@
 #### Whole brain analysis
+##### BMI-only model
 Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
@@
 <!-- cc -->
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate.
+Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate (FWE cluster p<0.05).
 When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap=c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
+knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
+```
+
 NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes. 
 Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
-To disentangle effects of change in BMI and logmFD in this dataset, we also split mFD into average mFD and change in mFD. Age and sex remained as nuisance covariates.
+
+##### BMI and FD model
+To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
 <!-- ### DMN -->
 <!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics of these clusters differed only slightly from the effect found for average BMI when using mFD as nuisance covariate.
+Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
 <!-- gsr -->As before, the effect did not survive when using GSR-corrected data.
+
+```{r "BMIFDagesexPCCccavgBMIdeact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "100%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
 
@@
 supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
-Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD.
-When running the model with GSR-preprocessed data, the associations between change in BMI and FC did not survive statistical thresholds.
-The relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction (p=$0.028$), also when evaluating the GSR corrected data (p=$0.038$).
+```{r "BMIFDagesexNACCcccgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI, average mFD and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
+```
+Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
+motor cortex, was significantly positively related to average mFD.
+This result remained similar when using GSR corrected data.
+```{r "BMIFDagesexNACCccavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
+```
 
@@
 ```
-The association also remained significant when we split logmFD into average and change in logmFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in logmFD with DMN FC.
+The association also remained significant when we split mFD into average and change in mFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in mFD with DMN FC.
 
@@
 ``` 
-As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and logmFD.
+As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and mFD.
 
@@
 In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
- --> 
+
 <!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
@@
 #### Whole-brain analysis
-To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models. 
+To investigate the unique contributions of the average value of and the longitudinal change in mFD, we constructed two additional models. 
 First, BMI measures are excluded from the design matrix.
-Covariance of the mFD and outcome meausres is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor.
+Covariance of the mFD and outcome measures is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor.
 Age and sex were nuisance covariates.
 <!-- #### Reward -->
-As shown in Table \@ref(tab:tableFDmodel), there is a significant co-activation of one cluster with the NAcc.
+As shown in Table \@ref(tab:tableFDmodel), there was a significant association between mFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas .
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
@@
 We did not find a any clusters for functional connectivity with the PCC.
-
-<!--### Effect on hypothalamic connectivity (exploratory analysis)-->
-
-```{r tableBMImodel, echo=TRUE}
-source("report/tables.R")
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap="Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCcc_avgFD_act.pdf")
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCgsr_avgFD_act.pdf")
+```
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+
+```{r tableBMImodel, echo=FALSE}
 TableList <- mk_FCTables()
@@
 
-```{r tableBMIFDmodel, eval=FALSE}
+```{r tableBMIFDmodel, echo=FALSE}
 tab_BMIFDmodel <- mk_FCTables()[[3]]
@@
 
-```{r tableFDmodel, eval=FALSE}
+```{r tableFDmodel, echo=FALSE}
 tab_FDmodel <- mk_FCTables()[[2]]
@@
 
-# Discussion
 In our sample, extend of micromovements of the head averaged across all measurements is associated with increased functional connectivity between reward and frontal regions.

### 2021-01-27 code/data_load.Rmd

---
title: "data_proprocessing"
author: "Hannah Sophie Heinrichs"
date: "1/20/2021"
output: 
  pdf_document: default
  html_document: default
---

# This is the summary of the MRI sample of ADI Study.

```{r "Package loading", include=FALSE}
library(lme4)
library(ggplot2)
library(haven)
library(plyr)
library(dplyr)
library(tidyr)
library(psych)
library(MuMIn)
```

### Loading Demographics
We get BMI from `ADI_MRT_Messueberblick_2019_04.csv` (this is an older version with correct columns (compared to _AW.xlsx). See README in `/data/p_02161/ADI_studie/metadata/`.

Furthermore, ADI061 was measured 4 times as bariatric surgery was performed after FU1. As data of measurement 4 is readily available, BL measure will be excluded and FU1 and FU2 will be recoded as BL and FU1, respectively.

```{r "Load data", include=FALSE}
bmi = read.csv(
  "/data/p_02161/ADI_studie/metadata/ADI_MRT_Messueberblick_2019_04_10.csv",
  na.strings = c("", "NA"),
  stringsAsFactors = FALSE
)
bmi_tmp = bmi[, c("X.1",
                  "BMI",
                  "X.2",
                  "BMI.1",
                  "X.3",
                  "BMI.2",
                  "Geschlecht",
                  "Alter")]

# collect subject IDs in one column
bmi_tmp$subj.ID = bmi_tmp$X.1  
bmi_tmp$subj.ID[!is.na(bmi_tmp$X.2)] = bmi_tmp$X.2[!is.na(bmi_tmp$X.2)]  
bmi_tmp$subj.ID[!is.na(bmi_tmp$X.3)] = bmi_tmp$X.3[!is.na(bmi_tmp$X.3)]

bmi_tmp = bmi_tmp[, c("subj.ID", "BMI", "BMI.1", "BMI.2", "Geschlecht", "Alter")]

# get BMI data from another ADI table which contains more subjects
adi_data = read_sav("/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav")
# select variables: only BMI
bmi_repl = adi_data[, c(
  "ADI_Code",
  "BMI_BL",
  "BMI_6m",
  "BMI_12m",
  "Sex",
  "Age_BL",
  "Age_6m",
  "Age_12m",
  "Educa_BL"
)]
bmi_repl$subj.ID = gsub("-", "", bmi_repl$ADI_Code)

# replace some missing BMI in the "Messueberblick" table with values from ADI_aktuell.sav
abmi = merge(bmi_tmp, bmi_repl, by = "subj.ID", all = TRUE)
abmi$BMI_BL[is.na(abmi$BMI_BL)] = abmi$BMI[is.na(abmi$BMI_BL)]
abmi$BMI_6m[is.na(abmi$BMI_6m)] = abmi$BMI.1[is.na(abmi$BMI_6m)]
abmi$BMI_12m[is.na(abmi$BMI_12m)] = abmi$BMI.2[is.na(abmi$BMI_12m)]
abmi$Sex[is.na(abmi$Sex)] = abmi$Geschlecht[is.na(abmi$Sex)]
abmi$Age_BL[is.na(abmi$Age_BL)] = abmi$Alter[is.na(abmi$Age_BL)]

# clean up data frame
lbmi <- abmi %>% gather(tp, BMI, BMI_BL, BMI_6m, BMI_12m)
lbmi = lbmi[, c("subj.ID", "Age_BL", "Sex", "tp",  "BMI")]
lbmi$tp = as.factor(lbmi$tp)
levels(lbmi$tp) = c("fu2", "fu", "bl")
lbmi$tp <- factor(lbmi$tp, levels = rev(levels(lbmi$tp)))
lbmi$Sex = as.factor(lbmi$Sex)
levels(lbmi$Sex) = c("male", "female", "male", "female")
```

```{r "Compare abmi with bmi from QOL list", include=FALSE }
# compare abmi (Messueberblick, ADI_aktuell.sav) with bmi from QOL list for those subjects who have it
QOL_sheet = read.csv(
  "/data/pt_02161/Results/Project1_headmotion/QOL_sheet_of_ADI_Messueberblick.csv"
)
QOL_sheet$subj.ID = gsub("-", "", QOL_sheet$X.1)

QOL_sheet$condition = "IG"
QOL_sheet[QOL_sheet$X == "KG", "condition"] = "KG"
adi_group = QOL_sheet[, c("subj.ID", "condition")]

bmi_compare = merge(QOL_sheet, abmi, by = "subj.ID", all = TRUE)
sum(bmi_compare$BMI.BL - bmi_compare$BMI_BL, na.rm = T)
sum(bmi_compare$BMI.FU - bmi_compare$BMI_6m, na.rm = T)
sum(bmi_compare$BMI.FU_2 - bmi_compare$BMI_12m, na.rm = T) # because of ADI061
```

### Load grouping infos

README: "Unfortunately, there are errors regarding BMI assignment (e.g. `/data/p_02161/ADI_studie/metadata/ADI_MRT_Messueberblick_AW.xlsx`, first sheet, where ADI002's FU2 weight is assigned to BL, for example). Similarly, group coding is wrong in the first two sheets (`ADI_MRT_Messueberblick_2019_04_` and `summary with SPSS data`). Correct group information, based on BMI trajectory, seems to be in sheet `QOL` (see ADI080, ADI036).
"

Take grouping infos first from QOL-sheet (accurate), and then add those that are not in QOL from "Messueberblick.csv".

```{r "Find grouping information", include=FALSE}
# merge data with group & find group for those that are not in QOL table
bmi_group = merge(lbmi, adi_group, by = "subj.ID", all.x = TRUE)

# this table has errors (contradictions for ADI080 and ADI036 with other table, but contains missing subjects)
group_coding_with_errors = read.csv(
  "/data/pt_02161/Results/Project1_headmotion/summary_with_SPSS_data_sheet_of_ADI_Messueberblick.csv"
)
group_coding_with_errors$subj.ID = gsub("-", "", group_coding_with_errors$Copy.SPSS)

group_for_replacement = merge(bmi_group[is.na(bmi_group$condition), ],
                              group_coding_with_errors,
                              all.x = TRUE,
                              by = "subj.ID")
bmi_group[is.na(bmi_group$condition), "condition"] = as.factor(group_for_replacement$Gruppe)
bmi_group$condition = as.factor(bmi_group$condition)
levels(bmi_group$condition) = c("IG", "KG", "IG", "KG")
```

Now, we loaded data for ALL participants (N=`r length(unique(bmi_group$subj.ID))`).

```{r "Write group coding for all subjects", echo=FALSE}
bmi_group_recoded <- bmi_group[!(bmi_group$subj.ID == "ADI061" &
                         bmi_group$tp == "bl"),]
# recode fu and fu2 into bl and fu
bmi_group_recoded$tp[(bmi_group_recoded$subj.ID == "ADI061" &
                         bmi_group_recoded$tp == "fu")] = "bl"
bmi_group_recoded$tp[(bmi_group_recoded$subj.ID == "ADI061" &
                         bmi_group_recoded$tp == "fu2")] = "fu"

write.csv(bmi_group_recoded[bmi_group_recoded$tp=="bl",], "/data/p_02161/ADI_studie/metadata/group_list_all.csv")
# as all other tables have the original coding for tp in ADI061, we have wo continue working with bmi_group and make the same recoding in the very end again.
```

### Add general MRI information

These infos were gathered during data copying, organizing and preprocessing, especially of the rs data. Yet, infos from other sequences are also included. See `/data/pt_02161/Analysis/Preprocessing/qa/README.md` for further information.
```{r "Loading rs QA infos from /data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv", echo=FALSE}
rs_QA=read.csv("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv")
for (i in 1:nrow(rs_QA)){
  #print(strsplit(toString(rs_QA[i,1]),'_')[[1]][1])
  rs_QA[i,"subj.ID"]=strsplit(toString(rs_QA[i,1]),'_')[[1]][1]
  rs_QA[i,"tp"]=strsplit(toString(rs_QA[i,1]),'_')[[1]][2]
}

# Don't use FS QA from this file.
bmi_group_MRI=merge(bmi_group, rs_QA[,c(2,3,11,12,13,14,15,16,17,18,19)],by=c("subj.ID","tp"))
```

### Add FS QA information based on Klapwijk et al. manual segmentation guidelines

See `/data/pt_02161/Analysis/Preprocessing/qa/README.md` for further information.
```{r "Loading FS infos from /data/p_02161/ADI_studie/BIDS/derivatives/freesurfer/QA/freesurfer_qa_qoala_11.7.csv", echo=FALSE}
FS_QA=read.csv("/data/p_02161/ADI_studie/BIDS/derivatives/freesurfer/QA/freesurfer_qa_qoala_11.7.csv")
for (i in 1:nrow(FS_QA)){
  #print(strsplit(toString(rs_QA[i,1]),'_')[[1]][1])
  FS_QA[i,"subj.ID"]=strsplit(toString(FS_QA[i,1]),'_')[[1]][1]
  FS_QA[i,"tp"]=strsplit(toString(FS_QA[i,1]),'_')[[1]][2]
}
bmi_group_MRI_QA=merge(bmi_group_MRI,FS_QA, by=c("subj.ID","tp"),all=TRUE)
```

### Loading mean/max FD values

These values describe the average and maximal motion during rsfMRI scans. Used for manuscript on head motion. See further information in `/data/pt_02161/Analysis/Project1_headmotion/`.
```{r "Loading mean/max FD values from /data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv", echo=FALSE}
rs_ID = read.csv("/data/pt_02161/Results/Project1_headmotion/subjects_rs.csv")
colnames(rs_ID)[1] = "subj"
res = read.csv(
  "/data/pt_02161/Results/Project1_headmotion/summary_FDvals_based_on_subjects_rs.csv",
  header = FALSE
)
colnames(res) = c("meanFD_bl",
                  "maxFD_bl",
                  "meanFD_fu",
                  "maxFD_fu",
                  "meanFD_fu2",
                  "maxFD_fu2")
rs_FD = data.frame(rs_ID$subj, res)
colnames(rs_FD)[1] = "subj.ID"
rs_FD[rs_FD == 0] = NA

lrs_meanFD <- rs_FD %>% gather(tp, meanFD, meanFD_bl, meanFD_fu, meanFD_fu2) %>% separate(tp, c("measure", "tp"))
lrs_maxFD <- rs_FD %>% gather(tp, maxFD, maxFD_bl, maxFD_fu, maxFD_fu2) %>% separate(tp, c("measure", "tp"))

lrs_FD = data.frame(lrs_meanFD$subj,
                    lrs_meanFD$tp,
                    lrs_meanFD$meanFD,
                    lrs_maxFD$maxFD)
colnames(lrs_FD) = c("subj.ID", "tp", "meanFD", "maxFD")
bmi_group_MRI_QA_FD = merge(bmi_group_MRI_QA,
                            lrs_FD,
                            by = c("subj.ID", "tp"),
                            all = TRUE)
```

However, because one participant had surgery after the first fu, baseline data will be disregarded and fu and fu2 will be recoded into bl and fu, respectively.

```{r "Reassignment of ADI061 data"}
# remove baseline measure (actually two time points before surgery)
bmi_group_MRI_QA_FD_recoded <- bmi_group_MRI_QA_FD[!(bmi_group_MRI_QA_FD$subj.ID == "ADI061" &
                         bmi_group_MRI_QA_FD$tp == "bl"),]
bmi_group_MRI_QA_FD_recoded$original_tp <- bmi_group_MRI_QA_FD_recoded$tp
# recode fu and fu2 into bl and fu
bmi_group_MRI_QA_FD_recoded$tp[(bmi_group_MRI_QA_FD_recoded$subj.ID == "ADI061" &
                         bmi_group_MRI_QA_FD_recoded$original_tp == "fu")] = "bl"
bmi_group_MRI_QA_FD_recoded$tp[(bmi_group_MRI_QA_FD_recoded$subj.ID == "ADI061" &
                         bmi_group_MRI_QA_FD_recoded$original_tp == "fu2")] = "fu"
```

### Final sample description.
```{r "Some characteristics of this sample", echo=FALSE}
final_sample=bmi_group_MRI_QA_FD_recoded
hist(final_sample$meanFD)
hist(final_sample$Final_Score)
```

We have `r length(unique(final_sample$subj.ID))` participants, where `r final_sample[final_sample$subj.ID=="ADI002"&final_sample$tp=="bl","subj.ID"]` has no data at all. So effectively, there are 50 participants with at least one measurement.

For anatomical imaging, we have `r sum(!is.na(final_sample$longtemplate.run))` longitudinal freesurfer runs.
`r sum(final_sample$Exclude, na.rm=TRUE)` scans have to be removed, due to **`r final_sample[!is.na(final_sample$Exclude),"Check_comments"][c(1,4,5,6)]`** and **`r final_sample[!is.na(final_sample$Exclude),"Correction_comments"][2:3]`**. This was the case for `r final_sample[!is.na(final_sample$Exclude),"subj.ID"]` at time points `r final_sample[!is.na(final_sample$Exclude),"tp"]`.

For functional imaging, we have `r sum(!is.na(final_sample$meanFD))` mean FD vals. For subj `r final_sample[!is.na(final_sample$meanFD)&is.na(final_sample$longtemplate.run),"subj.ID"]` at time point `r final_sample[!is.na(final_sample$meanFD)&is.na(final_sample$longtemplate.run),"tp"]` FS did not run through correctly.

```{r, include=FALSE}
write.csv(final_sample, "/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
```

README: "`group_list_all.csv` contains basic demographics and group coding for 77 participants. This table also includes participants without MRI.
A summary table was created from the 51 MRI subjects, including BMI trajectory, available sequences, and resting state and Freesurfer QA results. The resulting file is `final_sample_MRI_QA_info.csv`."

### Head motion correlates with quality score (the higher the head motion, the worse the quality of the T1 scan)
```{r, echo=FALSE}
plot(final_sample$meanFD, final_sample$Final_Score)
```

### 2021-01-27 manuscript/draft_rsfmri.Rmd

@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
+There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models)
 ```{r load agg FC data and prepare, echo=FALSE}
@@
 texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+
 ```
@@
 ```
+
 The association also remained significant when we split mFD into average and change in mFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in mFD with DMN FC.
@@
 ```
+
 As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and mFD.
@@
 Age and sex were nuisance covariates.
+
 <!-- #### Reward -->
@@
 ```
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 

### 2021-01-29 manuscript/draft_rsfmri.Rmd
create process.R to outsource computations from .Rmd file

--- 2021-01-27 manuscript/draft_rsfmri.Rmd
@@
 - Frauke Beyer* [1][2]
-- Kristin Prehn
-- Jürgen Ordemann 
+- Kerstin Prehn
 - Agnes Floel
+- Arno Villringer [1][2][3]
 - A. Veronica Witte [1][2][3]
@@
 
-source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/glmm_stability.r')
-source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')
-
 ```
@@
 setwd(knitr::opts_knit$get("root.dir"))
+
+source("code/figures.R")
+source("code/tables.R")
+source("code/create_sample_df.R")
 ```
-
-\newpage
-# Abstract
-
-\newpage
-# Introduction
-Obesity is a worldwide health issue, enchaining huge personal and societal costs.
-Excess amount of fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life.
-Conservative treatment options including dietary modification, physical activity and behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI (> 35 kg/m2).
-Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
-Roux-en-Y gastric bypass (RYGB) is the most effective surgical procedure [@Berthoud_2011].
-Here, a small pouch is formed from the proximal stomach and connected to the jejunum, disconnecting the larger part of the stomach and the proximal part of the small intestine intact from the digestive flow.
-Other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) reduce stomach volume, but leave the small intestine unchanged [@Mulla_2017].
-RYGB-induced weight loss is in part mediated by malabsorption of certain nutrients, especially fat, and reduced digestion efficiency [@Berthoud_2011] while the role of increased energy expenditure after RYGB is still unclear [@Yarmush_2017;@Lamarca_2019].
-Most importantly, RYGB alters the signaling cascades between the gut, fat tissue and the brain and thereby induces profound changes in the regulation of food intake and energy homeostasis [@Berthoud_2011].
-RYBG-patients frequently report altered smell and taste perception, i.e.
-increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items [@Mulla_2017, @Berthoud_2011].
-These changes in appetitive and reward signaling probably go back to altered gut-brain signaling, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin.
-Yet, precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
-Resting state functional magnetic resonance imaging (rsfMRI) is a powerful technique to investigate the dynamic organization of the brain via the correlation of spontaneous neural activity.
-Anatomically distinct brain regions show functional connectivity (FC), i.e.
-correlated neural activity over time, in the absence of a cognitive task.
-Crucially, these brain networks correspond to networks activated during specific cognitive processes, e.g.
-visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e.
-attention and executive control [@Smith_2011], and can thus be seen as default configuration of the brain.
-In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.
-The default mode network has been discovered as a “task-negative” network, and is involved in a variety of largely internally-directed cognitive functions including interoception, mind wandering and memory.
-Its core hubs are the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
-BA 39) [@Raichle_2015].
-Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2012; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
-Despite the large heterogenity and low median power, these findings might be summarized as a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
-salience and sensory networks.
-This has been interpreted as impaired interoceptive processing failing to adequately balance internal and external behavioral cues.
-Longitudinal studies largely support this view.
-Increased activation of the DMN in a food-viewing paradigm, i.e.
-insufficient down-regulation of DMN, has been shown to correlate with appetite measures after an overnight fast [@Tregellas_2011, @Wijngaarden_2015], an effect that could be partially reversed by an exercise intervention [@McFadden_2013].
-In a bariatric surgery study, [@Li_2018] reported a reduction of local FC in mid-line DMN regions in surgical patients compared to an obese control group.
-Simultaneously, PCC and medial prefrontal cortex were more strongly connected to dorsolateral prefrontal cortex, a region implicated in cognitive control.
-Another study in bariatric surgery patients showed that heightened connectivity of the PCC with anterior superior temporal gyrus (STG) before the surgery decreased over the course of one year [@Olivo_2017].
-Similarly, the profile of DMN connectivity with frontal regions, including ACC, frontal superior gyrus and the right OFC,  was similar in patients after bariatric surgery and normal weight controls, but dissimilar to that of obese controls [@Frank_2013].
-Another important brain network in the context of obesity is the reward network, comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
-These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
-Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced reward region activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020;].
-Food cue hyperactivation has been identified as a main neural vulnerability factor predicting future weight gain [@Stice_2015].
-Resting state fMRI studies have partly mirrored these results, showing increased local FC of reward network regions, i.e.
-Nacc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
-Imaging studies in bariatric surgery patients provided first evidence that hyperconnectivity between reward network regions (e.g.
-putamen, OFC) might be normalized to some extent by the surgery [@Duan_2020; @Wiemerslage_2016].
-Possible, the reconfiguration of the fronto-striatal brain networks might be related to altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013, but @Zoon_2018] .
-Thus, while there is some evidence hinting to a role for DMN and reward network connectivity in altered gut-brain communication after bariatric surgery, the existing evidence is inconclusive.
-Most studies have investigated small cohort of patients, often without adequate control group, and employed exploratory statistics [@George_2016].
-In this study, we aimed to extend the current literature and investigate differences of DMN and reward network connectivity before and after bariatric surgery in preregistered analyses and in comparison with a contemporaneously measured waiting-control group.
-Moreover, we explored whether the changes in functional connectivity were associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential dose-response relationship.
-Participants were measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
-Notably, we took into account possible non-linear time courses in increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
 
@@
 ```{r "Load data", include=FALSE}
-source("report/create_sample_df.R")
-final = create_sample_df(group = "all", tp = "all")
+#final <- create_sample_df(group = "both", tp = "all")
+final <- readRDS(file = "./report/tab/final.rds")
 
-# sample size
-#nrow(subj_df)
-subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each subject
-subj_df <- final[subj_idx,] # df with unique subjects
+# indices of first row of each subject
+subj_df <- final[match(unique(final$subj.ID), final$subj.ID),] # df with unique subjects
 
@@
 ```{r "tableSample", echo=FALSE}
-source("./report/tables.R")
-tab_sample <- mk_SampleTable(final)
+#tableSample <- mk_SampleTable(final)
+tableSample <- readRDS(file = "report/tab/tableSample.rds")
 
-tab_sample <-
+tab <-
   knitr::kable(
-    tab_sample,
+    tableSample,
     row.names = TRUE,
@@
     caption = "Distribution of data points at months after intervention")
-tab_sample
+tab
 ```
@@
 
-<!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité Berlin.
-For more details see @Prehn_2020.
-We used all data acquired until April 2019.
-The study protocol was in accordance with the Helsinki Declaration and approved by the local ethics committee.
-The study was registered at clinicaltrials.gov as NCT01554228.
-We preregistered the present resting state analyses on the [OSF](https://osf.io/yp42s).
-We made [additional changes](https://osf.io/59bh7/) to the preregistration after preprocessing as we realized some aspects of the analysis were inadequately described in the first preregistration.
-<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> 
-Participants were recruited from the Center for Bariatric and Metabolic Surgery.
-Inclusion criteria were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
+<!-- STUDY DESIGN --> The ADIPOSITAS-Study conducted in the Charité Berlin is a observational, prospective and longitudinal design described in detail by @Prehn_2020.
+Data collection is still ongoing but only data available in April 2019 is used.
+The procedures followed were in accordance with the Helsinki Declaration.
+This study was registered at clinicaltrials.gov as NCT01554228 and described in more detailed in the preregistration under [https://osf.io/yp42s](https://osf.io/yp42s).
+<!-- previous version: https://osf.io/f8tpn, update: https://osf.io/59bh7/ -->
+<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
+These were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$).
-In total, $51$ participants out of the originally enrolled $69$ subjects received MRI, and we analyzed a total of $101$ data points.
-Five data points of three subjects had to be excluded due to bad anatomical image quality (see section \@ref{sec:Anatomical Quality control}).
-<!-- DESCRIPTIVE STATISTICS --> 
-The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`).
-Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment.
+Participants with a history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$) were excluded from the study.
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment but not all participants have complete data for all time points (see Table \@ref(tab:tableDescr)).
+`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
+The current sample is the subsample of $69$ subjects that were originally enrolled in the study entailing only subjects that have received MRI.
 Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tableSample)).
-In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
-In the intervention group, data is not missing completely at random, as the pre-surgery timepoint MRI could not be obtained in some patients due to the limited scanner bore.
 
+<!-- DESCRIPTIVE STATISTICS --> The final sample entailed morbidly obese individuals 
+(n = `r nrow(subj_df)`, 
+`r nrow(filter(subj_df,Sex == "f"))` female; 
+`r round(mean(subj_df$Age_BL),2)` 
+$\pm$ `r round(sd(subj_df$Age_BL),2)` SD years old
+(range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`), 
+that either underwent surgery 
+(n = `r cs["Sum","IG"]`, 
+`r cs["f","IG"]` female, 
+`r round(mean(subj_df$Age_BL[subj_df$condition == "IG"]),2)` 
+$\pm$ `r round(sd(subj_df$Age_BL[subj_df$condition == "IG"]),2)` SD years old) 
+or were waiting list controls 
+(n = `r cs["Sum","KG"]`, 
+`r cs["f","KG"]` female,
+`r round(mean(subj_df$Age_BL[subj_df$condition == "KG"]),2)`
+$\pm$ `r round(sd(subj_df$Age_BL[subj_df$condition == "KG"]),2)` SD years old) 
+The study design was quasi-experimental, as patient in the waiting control list had also been recommended to undergo bariatric surgery, but were still waiting for their health insurance's approval.
 ```{r tableBaselineCharacteristics, include=FALSE}
-dfBL <- create_sample_df(group = "all", tp = "BL")
+#dfBL <- create_sample_df(group = "all", tp = "BL")
+dfBL <- readRDS(file = "./report/tab/dfBL.rds")
 
@@
 ```
-<!-- BASELINE CHARACTERISTICS --> 
-Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
@@
 # table for descriptive statistics
-descr <- psych::describeBy(dfw[,5:ncol(dfw)], group=dfw$condition,mat=T)
-descr$predictor <- c(rep("BMI",6),rep("meanFD",6))
-descr$tp <- car::recode(descr$vars, "c(1,4)='fu2';c(2,5)='fu';c(3,6)='bl'")
-descr$group <- descr$group1
-descr <- descr[,c("predictor","tp","group","n","mean","sd")]
-descr <- descr[with(descr, order(predictor, tp)),]
-descr[,c("mean","sd")] <- round(descr[,c("mean","sd")],2)
-descr[c(2:6,8:12),"predictor"] <- ""
-descr[c(2,4,6,8,10,12),"tp"] <- ""
-rownames(descr) <- NULL
 
-tab_descr <-
+tableDescr <- readRDS(file = "./report/tab/tableDescr.rds")
+
+tab <-
   knitr::kable(
-    descr,
+    tableDescr,
     row.names = FALSE,
@@
 
-tab_descr
+tab
 
 ```
-
 <!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
@@
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
-
 ```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-source("report/figures.R")
-
-fig_BMIdescr <- mk_figBMIdescr(final)
+fig_BMIdescr <- readRDS(file = "./report/fig/fig_BMIdescr.rds")
 fig_BMIdescr
-
 ```
@@
 
-## Preprocessing
+## Pre-proceessing
 
@@
 Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rs preprocessing.
-The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants].
+The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@antsref].
 The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
-*In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.*
+In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.
 
@@
 We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
-As described in [@pruim2015ica], the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+As described in [@Pruim2015], the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
 Finally, the timeseries of the  nuisance components were regressed from the raw fMRI timeseries.
@@
 To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed regions of interest (ROI), respectively.
-We chose these regions as they are core hubs of the networks.
-Because we knew about the large amount of motion in this dataset, we decided against Independent Component Analysis (ICA) to construct the networks, as ICA is recommended to work on minimally preprocessed data [GSR data](https://sourceforge.net/p/icatb/mailman/message/31466068/).
 Based on FreeSurfer's `aseg.mgz` and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
 Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson's correlation between them with `numpy.dot`, performed `r-to-z` transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
+We calculated the Pearson's correlation between them with `numpy.dot`, performed r-to-z transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
 Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
@@
 FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
-We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.
+We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration .
 <!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
@@
 #### Functional Quality control
-Resting state fMRI data quality control was performed according to the protocol by @Ciric_2018.
-As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020].
-Thus, we did not exclude any participants based on mean framewise displacement (FD) estimates beyond those who failed the anatomical preprocessing.
+Resting state fMRI data quality control was performed along the protocol of @Ciric_2018.
+As the expected average head motion is high in this sample [@Beyer_2020], it is difficult to determine an appropriate exclusion threshold which does not lead to the exclusion of a high number of individuals.
+Thus, we did not exclude any participants based on mean framewise displacement estimates beyond those who failed the anatomical preprocessing.
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
 In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
-```{r carpetplot, out.width="49%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap=c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
-carpets=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/carpetplotadi006fu2.jpeg","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/carpetplotadi16fu.jpeg")
-knitr::include_graphics(carpets)
-```
 On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
-Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum).
+Figure \@ref{carpet_plot} shows an exemplary carpet plot with (from top to bottom): DVARS (spatial root mean square of the data after temporal differencing), mFD, percent outliers defined by AFNI `OutlierCount`, voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data.
 All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
@@
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
+Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{example_FC}.
 We reported the mFD-FC correlation per preprocessing scheme.
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and timepoints, adjusted for age and sex.
-We used GSR regressed data as input and clusterwise bootstrapping with N=999.
+To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex.
+We used global signal regressed data as input and clusterwise bootstrapping with N=999.
 Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of p_FWE < 0.05.
-We extracted the average connectivity from these masks for GSR regressed data.
+We extracted the average connectivity from these masks for all data points.
 
@@
 Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+During analyses an implicit mask and an explicit mask were used.
+For this, FSL's standard space the brain mask that corresponds th "152 nonlinear 6th generation" atlas was re-sampled to 3 x 3 x 3 mm$^3$.<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+-->
 
-### Design Specification
+<!-- MODEL SETUP --> As specified in the preregistration, we tested the interaction of time and group: $FC = group + time + group*time$.
+Unlike prespecified, we did not include the model without covariates (as this is deprecated \@ref{nuisance_paper}).
+Instead, we calculated a model adjusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2).
+We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
+The model was examined for the correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $8$ a priori planned analyses.
 
-<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
-We did not perform the analysis as preregistered (i.e.
-without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
-<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
---> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)).
-The resulting factorial design <!-- check: flexible factorial?
---> contained one regressor for each time point per group.
-```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
-# fig.show="hold",  out.width="40%", fig.height = 8,
-source("report/figures.R")
-PlotList <- mk_figDesignMatrix()
+Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$.
+As a result a total of $32$ different analyses were performed.
+Computation with the grey matter mask were exploratory and deviate from the preregistration.
+The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.
 
-PlotList$plotA
-PlotList$plotB
-#gridExtra::grid.arrange(plotA, plotB, nrow=1)
-```
-We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
-As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
-This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3).
-Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-<!-- Changed manuscript here
-investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-  Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
-Age and sex were between-subject covariates while FD is time-varying.
-Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
- the inclusion of log mean FD may be disputed and done optionally.
-For that reason, we performed each analysis separately one with mean FD (three and without.
-Differences in results will be reported.
-We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
-<!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen.
---> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
-
-### Model input and brain mask
-For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC with GSR), resulting in $8$ preregistered analyses <!-- timexgroup (2 models) * 2* 2 --> , and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
-For all analyses, we used an explicit mask, which was either a whole brain or a GM mask derived from the MNI ICBM "152 nonlinear 6th generation" atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
-This resulted in a reduced number of tests (voxels of interest) from $52378$ to $29451$ for the GM mask.
-<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
--->
-Computation with the grey matter mask were exploratory and not planned in the preregistration.
-
-### SwE settings
 <!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
@@
 -->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
-<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+
+<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
+--> For the first model, time was introduced as factor because we did not assume strict linearity for the increase in connectivity and therefore refrained from modeling time as continuous variable (see Figure \@ref(fig:figDesignmatrix)).
+The resulting factorial design <!-- check: flexible factorial?
+--> contained one regressor for each time point per group.
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
+# fig.show="hold",  out.width="40%", fig.height = 8,
+
+DesignMatricesList <- readRDS(file = "./report/fig/DesignMatricesList.rds")
+DesignMatricesList$plotA
+DesignMatricesList$plotB
+#gridExtra::grid.arrange(plotA, plotB, nrow=1)
+```
+For the second and third model, we calculated the centered mean BMI and log mean FD per subject and the intra-subject-centered BMI as proposed by @Guillaume_2014.
+With a model containing average BMI and longitudinal change per subject, we investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
+<!-- NUISANCE COVARIATES --> Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
+Age and sex were between-subject covariates while FD is time-varying.
+Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
+As change in FD and change in BMI share common variance [@Beyer_2019], the inclusion of log mean FD may be disputed and done optionally.
+For that reason, we performed each analysis separately one with mean FD (three and without.
+Differences in results will be reported.
+We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
+<!-- POST HOC  --> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc model to disentangle the effects: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
+As these post-hoc analyses were not preregistered, they should be considered to be exploratory.
+
+<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using Wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
 Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
-In contrast to the parametric estimation, wild bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and unlike permutation testing it is also suitable <!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
+In contrast to the parametric estimation, Wild Bootstrap avoids any parametric distributional assumptions (e.g.
+random field theory), and is unlike permutation based framework appropriate<!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
+Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
@@
 
-### Aggregated FC analysis
 <!-- R analysis of aggregated values -->
 We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
-Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up.
-In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
-We performed model comparison between R1 and R0 models 
-`R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
-`R0 = lmer(agg FC ~ timepoint + group +  + age + (1|subj))`
-and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
-As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points.
-Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.
+Following the preregistration, we first investigated a time-by-group interaction between baseline and follow-up.
+We used a model comparison of R1 and R0 models with `anova`:
+`R1= lmer(agg FC ~ timepoint*group + (1|subj))`
+`R0 = lmer(agg FC ~ timepoint + group + (1|subj))`
+We performed a sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI.
+Again, the time-by-group point interaction was considered significant if a model comparison between R1 and Null models using `anova` function yielded a $p < 0.05$.
+Finally, we estimated the above-mentioned interaction model for all three time points.
+Again, the interaction was considered significant if a model comparison between R1 and R0 using `anova` had $p < 0.05$.
 
@@
 We performed different quality controls for the rsfMRI data.
-First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
-As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group. 
-Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points.
-We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
+As described in @Beyer_2020, group and time show an interaction effect for mFD where the intervention group reduced in head motion in the follow-up assessments compared to the control group.
+We also inspected DVARS, a measure of whole-brain signal variation which is reflective of noise in the data.
+We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
+This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
 ```{r tableQcfc, echo=FALSE}
 res <- read.csv("./report/FDFCcorrelations.csv")
-rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
-colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
-
-tab_qc_fc <- mk_tableQcfc(res)
-tab_qc_fc
+tab <- readRDS(file = "./report/fig/tableQcfc.rds")
+tab
 ```
@@
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
-Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
+Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency.
 Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
@@
 
-```{r calculate summary of QA, eval=FALSE, echo=FALSE}
-for (i in 1:nrow(final)) {
-  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
-  conf = read.csv(
-    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",
-      tmp, "/confounds.csv"))
-  tmp = cor.test(conf$DVARS, conf$FD)
-  final[i, "corr_DVARS_FD"] = tmp$estimate
-  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
-  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
-  final
-}
-```
-
 ```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
-final=read.csv(file = "./report/all_QA_measures.csv")
+final=read.csv(file = "report/all_QA_measures.csv")
 
@@
 ```{r FDFC correlation, eval=FALSE, echo=FALSE}
+# do not evaluate if saved already
 source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
 FDFC=calc_qc_fc(final)
-write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
+write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
 ```
 
-```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
-#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
-networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
+#setwd(knitr::opts_knit$get("root.dir"))
+networks = c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.pdf",
+             "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.pdf")
 knitr::include_graphics(networks)
@@
 
-Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
+Figure \@ref{fig:figNetworks} shows the DMN and reward network which we used for extracting the aggregated FC values.
 Both networks included the typically described brain regions (i.e.
 medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
-Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks).
-```{r prepare tsnr ROI plot, eval=FALSE}
-final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
 
-tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
-tsnr=tsnr[tsnr$V1!="subj",]
-colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
-levels(tsnr$roi)=droplevels(tsnr$roi)
-levels(tsnr$roi)=c("NAcc", "PCC/precuneus")
-tsnr$mean=as.double(tsnr$mean)
-
-final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
-
-p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-  geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") + 
-  scale_fill_manual(values=wes_palette("Darjeeling1",5,type="discrete")[3:4]) +
-  theme(axis.text=element_text(size=10),
-  axis.title=element_text(size=12),
-  strip.text = element_text(size=12),
-  legend.position="")
-ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", p)
+```{r prepare tsnr ROI plot, echo=FALSE}
+fig_tSNR <- mk_figtSNR(final)
 ```
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%",  eval=FALSE}
-p
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%", fig.align="center", echo=FALSE}
+
+fig_tSNR
 ```
+Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref{fig:figtSNR}.
 
@@
 
-#### Model 1a (with adjustment for age, sex)
+#### with adjustment for age, sex, meanFD
+There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc as was hypothesized.
+Nor was there a significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
 
-<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
-There was no interaction of group and time point regarding the resting state connectivity for neither the PCC nor the NAcc seeds against our initial hypothesis.
-There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
-
-####  Model 1b (with adjustment for age and sex, meanFD)
-Similarly, there were no significant clusters when including log mean FD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
-
+#### with adjustment for age and sex
+Similarly, there were only non-findings when reducing the number of nuisance covariates by leaving out mean FD.
+Results did not differ between analyses of connectivity maps from different preprocessing pipelines.
 <!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
@@
 
+### DMN/reward network with GM mask (exploratory analysis)
+<!-- GM mask: Group*Time Interaction [OLD]-->
+<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
+Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
+While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
+
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
-There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models)
 ```{r load agg FC data and prepare, echo=FALSE}
@@
 final_FC$subj.ID=as.factor(final_FC$subj.ID)
-final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")
-
-#calculate mean BMI over timepoints
-final_FC.meanBMI=tapply(X=final_FC$BMI,
-                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
-final_FC$mean.BMI=
-  final_FC.meanBMI[as.numeric(final_FC$subj.ID)]
-final_FC$within.BMI=
-  final_FC$BMI-final_FC$mean.BMI
 
@@
 
-```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
-#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
+### DMN for two timepoints
+```{r vif for models, echo=FALSE}
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
+vif_res=vif(test_vif)
+```
 
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+```{r calc DMN two tp,results = 'asis', echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvDMN2tp=anova(R1,R0)
@@
 
-```{qa check DMN model age,sex, echo=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
-vif_res=vif(test_vif)
+```{r DMN2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+### DMN for two timepoints, adjusted for age, sex, mean FD
+```{r calc DMN two tp adjusted, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tpadj=anova(R1,R0)
+res=summary(R1)
+#tab_model(R0,R1)
 ```
 
-```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
-#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
-
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvDMN2tp=anova(R1,R0)
+```{r DMN2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","mean FD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{qa check DMN model age,sex, mFD, echo=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
-vif_res=vif(test_vif)
-```
-
-```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-
+### DMN for three timepoints, adjusted for age, sex, mean FD
 ```{r calc DMN three tp, eval = TRUE, echo = FALSE}
-#### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD +(1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
 anvDMN3tp=anova(R1,R0)
@@
 
-```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
 #tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","mean FD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{r figaggDMN, echo=FALSE}
+```{r DMN conn over three timepoints, echo=FALSE}
 fig_DMNconn<-mk_figDMNdescr(final_FC)
@@
 
-```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
-### Reward network for two time points, adjusted for age and sex (Model 1a)
-
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+### Reward network for two time points
+```{r calc rew two tp, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvRew2tp=anova(R1,R0)
 ```
-```{qa check Rew model age,sex, echo=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
 
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
+```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-
+### Reward network for two timepoints, adjusted for age, sex, mean FD
 ```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
-#### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
-
 R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
@@
 ```
-
-```{qa check Rew model age,sex,mFD, echo=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
-vif_res=vif(test_vif)
+```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","mean FD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
 
-```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-
+### Reward network for three timepoints, adjusted for age, sex, mean FD
 ```{r calc rew three tp, eval = TRUE, echo = FALSE}
-#### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
-
 R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
@@
 ```
-```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
 #tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","mean FD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
 ```
-```{r figaggRew, echo=FALSE}
+```{r Rew conn over three timepoints, echo=FALSE}
 fig_Rewconn<-mk_figRewdescr(final_FC)
@@
 
-## Exploratory Analyses
-### Effect of mean and change in BMI 
-#### Whole brain analysis
+## Effect of mean and change in BMI (exploratory analysis)
+
 Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
 <!-- ### DMN -->
-<!-- cc -->
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate.
-When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
-NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes. 
-Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
-To disentangle effects of change in BMI and logmFD in this dataset, we also split mFD into average mFD and change in mFD. Age and sex remained as nuisance covariates.
-<!-- ### DMN -->
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics of these clusters differed only slightly from the effect found for average BMI when using mFD as nuisance covariate.
-<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.
-
-<!-- ### Reward -->
-<!-- cc -->
-For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
-supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
-Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD.
-When running the model with GSR-preprocessed data, the associations between change in BMI and FC did not survive statistical thresholds.
-The relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
-
-#### Aggregated FC analysis
-```{r DMN model, mean, change BMI, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
-
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
-R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
-anvmeanBMI_adjmFD=anova(R1,R02a)
-```
-Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD. 
-```{r Rew model, mean, change BMI,, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-```
-
-```{r DMN model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-anvmeanFD=anova(R1,R03)
-anvwithinFD=anova(R1,R03)
-
-p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
-```
-The association also remained significant when we split logmFD into average and change in logmFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in logmFD with DMN FC.
-
-#### Reward network, mean and within variability in BMI and logmFD
-```{r Rew model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-anvwithinFD=anova(R1,R03)
-anvmeanFD=anova(R1,R04)
-
-test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
-vif_res=vif(test_vif)
-``` 
-As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and logmFD.
-
-<!-- [OLD stuff] GM mask: Group*Time Interaction -->
+<!-- #### DMN for three timepoints -->
+<!-- cc -->Using the images preprocessed with AROMA+CC, we did find a significant co-activation with the PCC in three clusters associated with average BMI.
+Omitting mean FD as nuisance covariate yielded similar results but four significant clusters.
+However, in both, there is significant peak activation within the largest cluster.
+<!-- gsr -->In analyses on images that were additionally preprocessed with GSR, no cluster survived statistical thresholds for either ROI.
+All results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
 <!-- ### DMN for two timepoints
@@
  -->
-<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
+```{r "BMIagesexfd_PCCcc_avgBMI_deact", out.width = "50%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
+```
 
-### Effect of FD (exploratory analysis)
-#### Whole-brain analysis
+```{r "BMIagesex_PCCcc_avgBMI_deact", out.width = "85%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf")
+```
+
+## Effect of FD (exploratory analysis)
+
 To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models.
@@
 We did not find a any clusters for functional connectivity with the PCC.
+```{r "FDagesex_NACCcc_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCcc_avgFD_act.pdf")
+```
 
-<!--### Effect on hypothalamic connectivity (exploratory analysis)-->
+```{r "FDagesex_NACCgsr_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCgsr_avgFD_act.pdf")
+```
 
-```{r tableBMImodel, echo=TRUE}
-source("report/tables.R")
+## Effect of BMI and FD (exploratory analysis)
+
+Average BMI and change in BMI remain covariates of interest.
+While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as.
+Age and sex remain nuisance covariates.
+
+<!-- ### Reward -->
+<!-- cc -->There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
+Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
+<!-- gsr -->When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds.
+On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
+
+<!-- ### DMN -->
+<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
+There is a significant association between average BMI and the connecitivity between PCC and three other clusters.
+Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
+<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
+```{r "BMIFDagesex_NACCcc_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACCcc_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
+```
+
+```{r "BMIFDagesex_NACCgsr_avgFD_act", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACCgsr_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
+```
+
+```{r "BMIFDagesex_PCCcc_avgBMI_deact", out.width = "75%"}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
+
+## Effect on hypothalamic connectivity (exploratory analysis)
+...
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+
+```{r tableBMImodel}
 TableList <- mk_FCTables()
@@
 
-```{r tableBMIFDmodel, eval=FALSE}
+```{r tableFDmodel}
+tab_FDmodel <- mk_FCTables()[[2]]
+tab_FDmodel
+```
+
+```{r tableBMIFDmodel}
 tab_BMIFDmodel <- mk_FCTables()[[3]]
 tab_BMIFDmodel
-```
-
-```{r tableFDmodel, eval=FALSE}
-tab_FDmodel <- mk_FCTables()[[2]]
-tab_FDmodel
 ```

### 2021-01-29 code/process.R

@@ -1,0 +1,103 @@
+# processes data to report later in manuscript .Rmd file
+
+
+# Get functions -----------------------------------------------------------
+setwd("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/")
+source("figures.R")
+source("tables.R")
+source("create_sample_df.R")
+
+
+# Load data ---------------------------------------------------------------
+
+final <- create_sample_df(group = "both", tp = "all")
+
+# calculate summary of QA -------------------------------------------------
+
+for (i in 1:nrow(final)) {
+  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
+  #if (!stringr::str_detect(tmp, "^ADI\\d{3}_(bl|fu|fu2)$")){
+  #  next
+  #}
+  conf = read.csv(
+    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",
+           tmp, "/confounds.csv"))
+  tmp = cor.test(conf$DVARS, conf$FD)
+  final[i, "corr_DVARS_FD"] = tmp$estimate
+  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
+  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
+  final
+}
+saveRDS(object = final, file = "../report/tab/final.rds")
+
+
+saveRDS(object = final, file = "../report/tab/final.rds")
+
+# indices of first row of each subject
+subj_df <- final[match(unique(final$subj.ID), final$subj.ID),] # df with unique subjects
+subj_df$Sex <- plyr::mapvalues(subj_df$Sex, from = c("-1", "1"), to = c("f", "m"))
+saveRDS(object = subj_df, file = "../report/tab/subj_df.rds")
+
+# completed time points
+freqtable <- as.data.frame(table(final$subj.ID))
+tpfreq <- table(freqtable$Freq)
+
+# distribution of condition and sex
+cs <- addmargins(table(subj_df$Sex,subj_df$condition))
+
+dfw <- tidyr::pivot_wider(data = final,
+                          id_cols = c("subj.ID","condition","Sex","Age_BL"),
+                          names_from = "tp",
+                          values_from = c("BMI","meanFD"))
+saveRDS(object = dfw, file = "../report/tab/dfw.rds")
+
+
[…]

### 2021-01-29 manuscript/draft_rsfmri.Rmd
all computations transferred to process.R instead of .Rmn

@@
   chunk_output_type: console
+
+header-includes:
+- \usepackage[left]{lineno}
+- \linenumbers
 ---
@@
 # set working directory of knit (RMarkdown) to directory of Rproj
-opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-setwd(knitr::opts_knit$get("root.dir"))
+#opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+#setwd(knitr::opts_knit$get("root.dir"))
 
-source("code/figures.R")
-source("code/tables.R")
-source("code/create_sample_df.R")
 ```
@@
 #final <- create_sample_df(group = "both", tp = "all")
-final <- readRDS(file = "./report/tab/final.rds")
+final <- read.csv(file = "../report/final.csv")
 
@@
 #tableSample <- mk_SampleTable(final)
-tableSample <- readRDS(file = "report/tab/tableSample.rds")
+tableSample <- readRDS(file = "../report/tab/tableSample.rds")
 
@@
 <!-- previous version: https://osf.io/f8tpn, update: https://osf.io/59bh7/ -->
+```{r tableDescr, echo=FALSE}
+# table for descriptive statistics
+tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")
+
+knitr::kable(
+    tableDescr,
+    row.names = FALSE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = "BMI and mean FD of participants acquired in each condition and time point")
+```
 <!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA --> Participants were recruited from the Center for Bariatric and Metabolic Surgery if they were between age $18$ and $70$ and fulfilled the inclusion criteria.
@@
 #dfBL <- create_sample_df(group = "all", tp = "BL")
-dfBL <- readRDS(file = "./report/tab/dfBL.rds")
+dfBL <- readRDS(file = "../report/dfBL.rds")
 
@@
 ```
+
+```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
+# eval = TRUE, fig.asp = 0.7, fig.width = 6
+readRDS(file = "../report/fig/fig_BMIdescr.rds")
+```
 <!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
@@
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
-```{r tableDescr, echo=FALSE}
-
-# table_descr <- mk_tableDescr()
-# table for descriptive statistics
-
-tableDescr <- readRDS(file = "./report/tab/tableDescr.rds")
-
-tab <-
-  knitr::kable(
-    tableDescr,
-    row.names = FALSE,
-    format = "latex",
-    booktabs = T,
-    linesep = "",
-    caption = "BMI and mean FD of participants acquired in each condition and time point")
-
-tab
-
-```
 <!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for <!-- standadized??
+They underwent medical assessments including an interview, blood draw and anthropocentric measurements before having a one our break for <!-- standadized??
 -->breakfast.
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
-```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
-# eval = TRUE, fig.asp = 0.7, fig.width = 6
-fig_BMIdescr <- readRDS(file = "./report/fig/fig_BMIdescr.rds")
-fig_BMIdescr
-```
 
@@
 For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
-*This part of the analysis was described in the preregistration as "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated."*
+<!-- *This part of the analysis was described in the preregistration as "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated."*   -->
 
@@
 Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
-*In the preregistration we stated "Resting-state functional connectivity (FC) using seed-based connectivity analysis extracted with in-house pre-processing pipelines (see (Beyer et al., 2019)): Whole-brain FC maps for a priori defined seed regions (defined based on the results of the subcortical segmentation and Desikan-Killiany cortical parcellation in the Freesurfer version 6.0.0 longitudinal processing stream).
+<!-- *In the preregistration we stated "Resting-state functional connectivity (FC) using seed-based connectivity analysis extracted with in-house pre-processing pipelines (see (Beyer et al., 2019)): Whole-brain FC maps for a priori defined seed regions (defined based on the results of the subcortical segmentation and Desikan-Killiany cortical parcellation in the Freesurfer version 6.0.0 longitudinal processing stream).
 For the reward network a ROI combined from bilateral Nucleus accumbens from the subcortical segmentation (26, Left-Accumbens-area, 58, Right-Accumbens-area) was selected.
@@
 These steps will be performed using nilearn, nipype and python.
-Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).*
+Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).* -->
 
@@
 <!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
-*In the preregistration we stated that "Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary", and that "We will exclude participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
-caused Freesurfer to fail or perform very poorly (rated “Failed” according to (Klapwijk et al., 2019)".*
+<!-- *In the preregistration we stated that "Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary", and that "We will exclude participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
+caused Freesurfer to fail or perform very poorly (rated “Failed” according to (Klapwijk et al., 2019)".* -->
 
@@
 
-<!-- param.ROI_PREP = {roi_prep{[6,14]}}; % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {'Nacc_cc_z','Nacc_gsr_z','PCC_cc_z','PCC_gsr_z','LH_cc_z','LH_gsr_z','MH_cc_z','MH_gsr_z'} -->
-<!-- % Model definition -->
-<!-- % All three models have unique options for covariate definition, the -->
-<!-- % association to a model is indicated by the tens digit (GroupTime_: 1_;  -->
-<!-- % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit -->
-<!-- param.MODEL = {'alltp'}; % {'grouptime','grouptime2tp'} % {'bmi','bmiIG','bmi2tp'} % {'fd','fdIG'} % {'alltp'} % {'singletp}  -->
-<!-- param.COVARIATES = [41];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42]    % [41, 42, 43] -->
-
-<!-- % define masking and type of inference -->
-<!-- param.MASK = 'brain';               % 'brain' -->
-<!-- param.EXCLFD = false;               % false -->
-<!-- param.WILD_BOOT = false;            % false -->
-<!-- param.INFERENCE_TYPE = {'cluster'}; % {'voxel','cluster','tfce'}; -->
-
-*In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
-The mask will include all voxels surviving FWE-correction of p<0.05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"*
-
-<!-- MOTION CORRECTION We therefore apply best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitor the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-As quality checks statistical analysis were performed on both with and without global signal regression, considering the controversial discussion about GSR introducing negative correlations and spatial bias into connectivity data [@Ciric_2017; @Murphy_2017].
-Additionally, mean framewise displacement (FD) was extracted from 6 translational/rotational motion parameters of the whole time series according to @Power_2012, log-transformed and examined within and separate from the main models.
--->
+<!-- *In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
+The mask will include all voxels surviving FWE-correction of p<0.05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"* -->
 
@@
 
-DesignMatricesList <- readRDS(file = "./report/fig/DesignMatricesList.rds")
+DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
 DesignMatricesList$plotA
@@
 This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
+
+```{r mFDFC, echo=FALSE}
+res <- read.csv("../report/FDFCcorrelations.csv")
+```
+
 ```{r tableQcfc, echo=FALSE}
-res <- read.csv("./report/FDFCcorrelations.csv")
-tab <- readRDS(file = "./report/fig/tableQcfc.rds")
-tab
+tab_Qcfc <-
+  knitr::kable(
+    round(res, 3),
+    col.names = colnames(res),
+    row.names = TRUE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points"
+  ) %>%
+  column_spec(c(2:6), width = "1.8cm")
+tab_Qcfc
 ```
+
 Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018].
@@
 ```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
-final=read.csv(file = "report/all_QA_measures.csv")
 
-FigureList <- figDvarsmFD(final)
-(FigureList$fig1 | FigureList$fig2) / FigureList$fig3
-```
-
-```{r FDFC correlation, eval=FALSE, echo=FALSE}
-# do not evaluate if saved already
-source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
-FDFC=calc_qc_fc(final)
-write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
+figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
+(figList$fig1 | figList$fig2) / figList$fig3
 ```
@@
 #setwd(knitr::opts_knit$get("root.dir"))
-networks = c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.pdf",
-             "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.pdf")
+networks = c("../report/fig/DMN_thresholded_for_clusterFWE005.pdf",
+             "../report/fig/Rew_thresholded_for_clusterFWE005.pdf")
 knitr::include_graphics(networks)
@@
 
-```{r prepare tsnr ROI plot, echo=FALSE}
-fig_tSNR <- mk_figtSNR(final)
-```
-
 ```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="60%", fig.align="center", echo=FALSE}
-
-fig_tSNR
+readRDS(file = "../report/fig/fig_tSNR.rds")
 ```
@@
 
-### DMN/reward network with GM mask (exploratory analysis)
-<!-- GM mask: Group*Time Interaction [OLD]-->
-<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
-
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
 ```{r load agg FC data and prepare, echo=FALSE}
-agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
-agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
-agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
-agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
-for (i in c(1:nrow(agg_FC_DMN_ID))){
-  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
-  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
-}
-
-colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
-colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
-
-final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
-final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
-final_FC$subj.ID=as.factor(final_FC$subj.ID)
-
-#calculate mean FD over timepoints
-final_FC.meanFD=tapply(X=final_FC$logmFD,
-                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
-final_FC$mean.logmFD=
-  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
-final_FC$within.logmFD=
-  final_FC$logmFD-final_FC$mean.logmFD
+final_FC <- read.csv("../report/final_FC.csv")
 ```
@@
 
-```{r DMN conn over three timepoints, echo=FALSE}
-fig_DMNconn<-mk_figDMNdescr(final_FC)
-fig_DMNconn
+```{r figDMNconn, echo=FALSE}
+readRDS(file = "../report/fig/figDMNconn.rds")
 ```
@@
 ```
-```{r Rew conn over three timepoints, echo=FALSE}
-fig_Rewconn<-mk_figRewdescr(final_FC)
-fig_Rewconn
+
+```{r figRewconn, echo=FALSE}
+readRDS(file = "../report/fig/figRewconn.rds")
 ```
@@
 All results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
-<!-- ### DMN for two timepoints
-BMIONLYIG
-Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps).
-However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
-<!-- BMI2TP
-Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants.
-Omitting mean FD did not change these results.
--->
-<!-- GM mask
-Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
- -->
+
 ```{r "BMIagesexfd_PCCcc_avgBMI_deact", out.width = "50%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
+knitr::include_graphics("../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
 ```
@@
 ```{r "BMIagesex_PCCcc_avgBMI_deact", out.width = "85%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf")
+knitr::include_graphics("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf")
 ```
@@
 ```{r "FDagesex_NACCcc_avgFD_act", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCcc_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
 ```
@@
 ```{r "FDagesex_NACCgsr_avgFD_act", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCgsr_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/FDagesex_NACCgsr_avgFD_act.pdf")
 ```
@@
 <!-- ### DMN -->
+```{r tableBMImodel}
+# do not change chunk label as it corresponds to the manually defined (in table.R) table label
+tab_BMImodel <- readRDS(file = "../report/tab/FCTableList.rds")[[1]]
+tab_BMImodel
+```
+
+```{r tableFDmodel}
+# do not change chunk labelsas it corresponds to the manually defined (in table.R) table label
+tab_FDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[2]]
+tab_FDmodel
+```
+
+```{r tableBMIFDmodel}
+# do not change chunk label as it corresponds to the manually defined (in table.R) table label
+tab_BMIFDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[3]]
+tab_BMIFDmodel
+```
 <!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
-There is a significant association between average BMI and the connecitivity between PCC and three other clusters.
+There is a significant association between average BMI and the connectivity between PCC and three other clusters.
 Results are detailed in Table \@ref(tab:tableBMIFDmodel).
@@
 ```{r "BMIFDagesex_NACCcc_avgFD_act", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf")
 ```
@@
 ```{r "BMIFDagesex_NACCcc_cgnBMI_deact", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
 ```
@@
 ```{r "BMIFDagesex_NACCgsr_avgFD_act", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf")
 ```
@@
 ```{r "BMIFDagesex_NACCgsr_cgnBMI_deact", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
 ```
@@
 ```{r "BMIFDagesex_PCCcc_avgBMI_deact", out.width = "75%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
 ```
@@
 Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
-
-```{r tableBMImodel}
-TableList <- mk_FCTables()
-tab_BMImodel <- TableList[[1]]
-tab_BMImodel
-```
-
-```{r tableFDmodel}
-tab_FDmodel <- mk_FCTables()[[2]]
-tab_FDmodel
-```
-
-```{r tableBMIFDmodel}
-tab_BMIFDmodel <- mk_FCTables()[[3]]
-tab_BMIFDmodel
-```
 
@@
 # Acknowledgments
-# Author Contributions
-
-Stud conception and design
-Aqcuisiiton of data:
-Analysis andinterpretation of data:
-Frafting of manuscript:
-Ciritical revision:
 

### 2021-01-29 code/process.R

@@ -1,11 +1,13 @@
 # processes data to report later in manuscript .Rmd file
-
+ROOT_DIR = "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/"
 
 # Get functions -----------------------------------------------------------
-setwd("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/")
+setwd(ROOT_DIR)
 source("figures.R")
 source("tables.R")
 source("create_sample_df.R")
+source("../../../../../Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
+
 
 
 # Load data ---------------------------------------------------------------
@@ -28,10 +30,11 @@ for (i in 1:nrow(final)) {
   final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
   final
 }
-saveRDS(object = final, file = "../report/tab/final.rds")
 
+final <- within(final,  subj.ID_tp <- paste(subj.ID,tp, sep="_"))
 
-saveRDS(object = final, file = "../report/tab/final.rds")
+# save final data frame
+write.csv(final,"../report/final.csv", row.names = FALSE)
 
 # indices of first row of each subject
 subj_df <- final[match(unique(final$subj.ID), final$subj.ID),] # df with unique subjects
@@ -61,7 +64,7 @@ saveRDS(object = tableSample, file = "../report/tab/tableSample.rds")
 # tableBaselineCharacteristics --------------------------------------------
 
 dfBL <- create_sample_df(group = "all", tp = "BL")
-saveRDS(object = dfBL, file = "../report/tab/dfBL.rds")
+saveRDS(object = dfBL, file = "../report/dfBL.rds")
 
 
 # tableDescr --------------------------------------------------------------
@@ -98,6 +101,72 @@ res <- read.csv("../report/FDFCcorrelations.csv")
 rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
 colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
 
-tableQcfc <- mk_tableQcfc(res)
[…]

### 2021-01-29 code/tables.R

@@ -253,6 +253,7 @@ fn3 <- "To identify significant clusters, we applied a cluster size threshold wi
 fn4 <- "Connectivity with maximum three voxels that mark local maxima within the respective custer; more detailed description of anatomical regions that are assigned to overall clusters and and corresponding probability in Supplementary."
 title_vec <- c("BMI", "FD", "BMI-FD") 
 
+tabnames <- c("tableBMImodel","tableFDmodel","tableBMIFDmodel")
 TableList <- list()
 for (i in 1:length(ModelList)){
   TableList[[i]] <-
@@ -264,7 +265,8 @@ for (i in 1:length(ModelList)){
       format = "latex",
       booktabs = T,
       linesep = "",      # disable "\\addlinespace" at every 5th line
-      caption = sprintf("Changes in functional connectivity in whole brain analysis for %s model",title_vec[i])
+      label = tabnames[i],
+      caption = sprintf("Changes in functional connectivity in whole brain analysis for %s model",title_vec[i]),
     ) %>%
     kable_styling(latex_options = c('scale_down')) %>% # "scale_down"
     #column_spec(c(11), width = "3.2cm") %>%
@@ -276,7 +278,7 @@ for (i in 1:length(ModelList)){
     add_footnote(c(fn3, fn4), notation = "number") %>%
     kableExtra::landscape()
 }
-names(TableList) <- c("tab_BMImodel","tab_FDmodel","tab_BMIFDmodel")
+names(TableList) <- tabnames
 
 TableList[[1]] <- TableList[[1]] %>%
   group_rows("average BMI (decrease)", 1, nrow(df2)) %>%
@@ -319,21 +321,4 @@ mk_SampleTable <- function(final) {
   rownames(tab_sample) <- c("count: only 0","count: only 6","count: only 12","count: 0 and 6","count: 6 and 12","count: complete data","total number of subjects","total data points")
   colnames(tab_sample) <- c("BARS","NBARS")
   return(tab_sample)
-}
-
-mk_tableQcfc <- function(res) {
-  
-  tab_qc_fc <-
-    knitr::kable(
-      round(res,3),
-      col.names = colnames(res),
-      row.names = TRUE,
-      format = "latex",
-      booktabs = T,
-      linesep = "",
-      caption = "Summary of quality metrics for different denoising pipelines across conditions and time points") %>%
-    column_spec(c(2:6), width = "1.8cm") 
-  
-  
-  return(tab_qc_fc)
 }

### 2021-02-01 manuscript/draft_rsfmri.Rmd
added abstract glassbrain plots and changes to manuscript

--- 2021-01-27 manuscript/draft_rsfmri.Rmd
@@
 increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items [@Mulla_2017, @Berthoud_2011].
-These changes in appetitive and reward signaling probably go back to altered gut-brain signaling, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin.
+These changes in appetitive and reward signaling likely relate to altered gut-brain communication, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin [ @Brutman_2019].
 Yet, precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
@@
 We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
+```{r FDFC correlation, eval=FALSE, echo=FALSE}
+# do not evaluate if saved already
+source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
+FDFC=calc_qc_fc(final)
+write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
+```
+
 ```{r tableQcfc, echo=FALSE}
@@
 
-```{r calculate summary of QA, eval=FALSE, echo=FALSE}
+```{r calculate summary of QA, eval=TRUE, echo=FALSE}
 for (i in 1:nrow(final)) {
@@
 }
+write.csv(file = "report/all_QA_measures.csv", final)
 ```
@@
 
-```{r FDFC correlation, eval=FALSE, echo=FALSE}
-# do not evaluate if saved already
-source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
-FDFC=calc_qc_fc(final)
-write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
-```
-
-```{r prepare tsnr ROI plot, eval=FALSE}
+```{r prepare tsnr ROI plot, eval=TRUE}
 final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)

### 2021-02-01 manuscript/supplements_rsfmri.Rmd
added abstract glassbrain plots and changes to manuscript

@@
 source('/data/gh_gr_agingandobesity_share/literature/methods/statistics/linear_models_course_rogermundry_2018/functions/diagnostic_fcns.r')
-
 ```
@@
 setwd(knitr::opts_knit$get("root.dir"))
+source('report/figures.R')
+source("report/tables.R")
+source("report/create_sample_df.R")
 ```
@@
 ```{r load agg FC data and prepare, echo=FALSE}
+final=readRDS("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tab/final.rds")
+
 agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
@@
 for (i in c(1:nrow(agg_FC_DMN_ID))){
-  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
-  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
+  tmp=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
+  agg_FC_Rew[i,"subj.ID"]=strsplit(tmp,"_")[[1]][1]
+  agg_FC_Rew[i,"tp"]=strsplit(tmp,"_")[[1]][2]
+  tmp=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
+  agg_FC_DMN[i,"subj.ID"]=strsplit(tmp,"_")[[1]][1]
+  agg_FC_DMN[i,"tp"]=strsplit(tmp,"_")[[1]][2]
 }
 
-colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
-colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
+colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID", "tp")
+colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID","tp")
 
-final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
-final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
+final_FC=merge(final, agg_FC_Rew, by=c("subj.ID","tp"), all.x=TRUE)
+final_FC=merge(final_FC, agg_FC_DMN, by=c("subj.ID","tp"), all.x=TRUE)
 final_FC$subj.ID=as.factor(final_FC$subj.ID)

### 2021-02-01 report/figures.R

@@ -23,13 +23,8 @@ mk_figBMIdescr <- function(final) {
       size = 2,
       color = c(rep("#046C9A", 3), rep("#D69C4E", 3))
     ) +
-<<<<<<< HEAD
     facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) + xlab("time points")
   figBMIdescr= figBMIdescr +
-=======
-    facet_grid(. ~ condition, labeller = labeller(condition=condition.labs)) +
-    xlab("time points") +
->>>>>>> issuemergingwhannah
     scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
     scale_color_manual(
       values = c("#046C9A70", "#D69C4E70"),
@@ -47,10 +42,8 @@ mk_figBMIdescr <- function(final) {
   return(figBMIdescr)
 }
 
-<<<<<<< HEAD
-=======
 mk_figBMIdescr_for_OHBM <- function(final) {
-  
+
   condition.labs <- c("BS","NBS")
   names(condition.labs) <- c("IG","KG")
   # BMI over time each group
@@ -73,7 +66,7 @@ mk_figBMIdescr_for_OHBM <- function(final) {
     stat_summary(
       aes(group = condition),
       fun= mean,
-      fun.min = function(x) mean(x) - sd(x), 
+      fun.min = function(x) mean(x) - sd(x),
       fun.max = function(x) mean(x) + sd(x),
       geom  = "errorbar",
       size = 1, width = 0.1,
@@ -100,59 +93,6 @@ mk_figBMIdescr_for_OHBM <- function(final) {
   return(figBMIdescrOHBM)
 }
 
-mk_figFDdescr_for_OHBM <- function(final) {
-  
-  condition.labs <- c("BS","NBS")
-  names(condition.labs) <- c("IG","KG")
-  # BMI over time each group
-  figFDdescrOHBM <-
-    ggplot(final, aes(
-      x = tp,
-      y = meanFD,
-      group = subj.ID,
-      color = condition
-    )) +
-    geom_line(position=position_nudge(x = -0.02)) + geom_point(position=position_nudge(x = -0.02)) +
-    stat_summary(
-      aes(group = condition),
-      fun = mean,
-      geom  = "line",
-      size = 1,
-      position=position_nudge(x = 0.02, y = 0),
-      color = c(rep("#046C9A", 3), rep("#D69C4E", 3))
[…]

### 2021-02-01 manuscript/draft_rsfmri.Rmd
empty except for FDFCcorrelations

--- 2021-01-29 manuscript/draft_rsfmri.Rmd
@@
 editor_options:
+
   chunk_output_type: console
-
 header-includes:

### 2021-02-01 manuscript/draft_rsfmri.Rmd

--- 2021-02-01 manuscript/draft_rsfmri.Rmd
@@
 editor_options:
+
   chunk_output_type: console
+header-includes:
+- \usepackage[left]{lineno}
+- \linenumbers
 ---
@@
 # set working directory of knit (RMarkdown) to directory of Rproj
-opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-setwd(knitr::opts_knit$get("root.dir"))
-
-source("report/figures.R")
-source("report/tables.R")
-source("report/create_sample_df.R")
+#opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+#setwd(knitr::opts_knit$get("root.dir"))
+
 ```
@@
 ```{r "Load data", include=FALSE}
-final <- create_sample_df(group = "both", tp = "all")
-
-# sample size
-#nrow(subj_df)
-subj_idx <- match(unique(final$subj.ID), final$subj.ID) # indices of first row of each subject
-subj_df <- final[subj_idx,] # df with unique subjects
+#final <- create_sample_df(group = "both", tp = "all")
+final <- read.csv(file = "../report/final.csv")
+
+# indices of first row of each subject
+subj_df <- final[match(unique(final$subj.ID), final$subj.ID),] # df with unique subjects
 
@@
 ```{r "tableSample", echo=FALSE}
-tab_sample <- mk_SampleTable(final)
-
-tab_sample <-
+#tableSample <- mk_SampleTable(final)
+tableSample <- readRDS(file = "../report/tab/tableSample.rds")
+
+tab <-
   knitr::kable(
-    tab_sample,
+    tableSample,
     row.names = TRUE,
@@
 In the intervention group, data is not missing completely at random, as the pre-surgery timepoint MRI could not be obtained in some patients due to the limited scanner bore.
-
 ```{r tableBaselineCharacteristics, include=FALSE}
-dfBL <- create_sample_df(group = "all", tp = "BL")
+#dfBL <- create_sample_df(group = "all", tp = "BL")
+dfBL <- readRDS(file = "../report/dfBL.rds")
 
@@
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
-```{r tableDescr, echo=FALSE}
-
-# table_descr <- mk_tableDescr()
+
+```{r tableDescr, include=FALSE}
 # table for descriptive statistics
-descr <- psych::describeBy(dfw[,5:ncol(dfw)], group=dfw$condition,mat=T)
-descr$predictor <- c(rep("BMI",6),rep("meanFD",6))
-descr$tp <- car::recode(descr$vars, "c(1,4)='fu2';c(2,5)='fu';c(3,6)='bl'")
-descr$group <- descr$group1
-descr <- descr[,c("predictor","tp","group","n","mean","sd")]
-descr <- descr[with(descr, order(predictor, tp)),]
-descr[,c("mean","sd")] <- round(descr[,c("mean","sd")],2)
-descr[c(2:6,8:12),"predictor"] <- ""
-descr[c(2,4,6,8,10,12),"tp"] <- ""
-rownames(descr) <- NULL
-
-tab_descr <-
+tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")
+
 knitr::kable(
-    descr,
+    tableDescr,
     row.names = FALSE,
@@
     caption = "BMI and mean FD of participants acquired in each condition and time point")
-
-tab_descr
-
 ```
@@
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
-fig_BMIdescr <- mk_figBMIdescr(final)
-fig_BMIdescr
-```
+readRDS(file = "../report/fig/fig_BMIdescr.rds")
+```
+<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
+`r round(tab_age$mean[tab_age$group1=="IG"],2)`).
+Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
+<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
+<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+They underwent medical assessments including an interview, blood draw and anthropocentric measurements before having a one our break for <!-- standadized??
+-->breakfast.
+MRI scanning was done after further psychological tests (for details see @Prehn_2020).
 
@@
 without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
-<!-- DESIGN MATRIX, vielleicht beide auch als Abbildung kurz zeigen?
---> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)).
+<!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)).
 The resulting factorial design <!-- check: flexible factorial?
@@
 # fig.show="hold",  out.width="40%", fig.height = 8,
-PlotList <- mk_figDesignMatrix()
-
-PlotList$plotA
-PlotList$plotB
+
+DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
+DesignMatricesList$plotA
+DesignMatricesList$plotB
 #gridExtra::grid.arrange(plotA, plotB, nrow=1)
@@
 We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
-```{r FDFC correlation, eval=FALSE, echo=FALSE}
-# do not evaluate if saved already
-source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
-FDFC=calc_qc_fc(final)
-write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
-```
-
-```{r tableQcfc, echo=FALSE}
-res <- read.csv("report/FDFCcorrelations.csv")
-rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
-colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
-
-tab_qc_fc <- mk_tableQcfc(res)
-tab_qc_fc
+
+```{r mFDFC, echo=FALSE}
+res <- read.csv("../report/FDFCcorrelations.csv")
 ```
@@
 write.csv(file = "report/all_QA_measures.csv", final)
-```
-
-```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
-final=read.csv(file = "report/all_QA_measures.csv")
-
-FigureList <- figDvarsmFD(final)
-(FigureList$fig1 | FigureList$fig2) / FigureList$fig3
-```
+
+```{r tableQcfc, echo=FALSE}
+tab_Qcfc <-
+  knitr::kable(
+    round(res, 3),
+    col.names = colnames(res),
+    row.names = TRUE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points"
+  ) %>%
+  column_spec(c(2:6), width = "1.8cm")
+tab_Qcfc
+```
+
+Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018].
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`).
+This relationship was negatively associated with distance between nodes (mean Spearman's r = `r round(res[1,5],2)`).
+Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
+Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency.
+Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
+Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
 
@@
 ```{r figNetworks, echo=FALSE, out.width="33%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+
 #networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
@@
 tsnr=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
+```
+```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
+figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
+(figList$fig1 | figList$fig2) / figList$fig3
+```
+
+```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
+networks = c("../report/fig/DMN_thresholded_for_clusterFWE005.pdf",
+             "../report/fig/Rew_thresholded_for_clusterFWE005.pdf")
 knitr::include_graphics(networks)
@@
 ```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%",  eval=FALSE}
-p
+readRDS(file = "../report/fig/fig_tSNR.rds")
 ```
@@
 ```{r load agg FC data and prepare, echo=FALSE}
+
+final_FC <- read.csv("../report/final_FC.csv")
+
 agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
@@
   final_FC$logmFD-final_FC$mean.logmFD
+
 ```
@@
 
+```{r figDMNconn, echo=FALSE}
+readRDS(file = "../report/fig/figDMNconn.rds")
+```
+
 ```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
@@
 texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
+```{r figRewconn, echo=FALSE}
+readRDS(file = "../report/fig/figRewconn.rds")
 ```
@@
 <!-- ### DMN -->
-<!-- cc -->
 Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate (FWE cluster p<0.05).
-When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
+When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \@ref(tab:tableBMImodel).
 ```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap=c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
-knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
+knitr::include_graphics(c("../report/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
 ```
@@
 
-```{r "BMIFDagesexPCCccavgBMIdeact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "100%"}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```{r "BMIagesexfd_PCCcc_avgBMI_deact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "85%"}
+knitr::include_graphics("../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
+```
+
+```{r "BMIagesex_PCCcc_avgBMI_deact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "85%"}
+knitr::include_graphics("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf")
 ```
@@
 We did not find a any clusters for functional connectivity with the PCC.
+
 ```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap="Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCcc_avgFD_act.pdf")
-knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCgsr_avgFD_act.pdf")
-```
-
-```{r tableBMImodel, echo=FALSE}
-TableList <- mk_FCTables()
-tab_BMImodel <- TableList[[1]]
+knitr::include_graphics("../report/FDagesex_NACCcc_avgFD_act.pdf")
+knitr::include_graphics("../report/FDagesex_NACCgsr_avgFD_act.pdf")
+```
+
+## Effect of BMI and FD (exploratory analysis)
+
+Average BMI and change in BMI remain covariates of interest.
+While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as.
+Age and sex remain nuisance covariates.
+
+<!-- ### Reward -->
+<!-- cc -->There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
+Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
+<!-- gsr -->When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds.
+On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
+
+<!-- ### DMN -->
+```{r tableBMImodel}
+# do not change chunk label as it corresponds to the manually defined (in table.R) table label
+tab_BMImodel <- readRDS(file = "../report/tab/FCTableList.rds")[[1]]
 tab_BMImodel
@@
 
-```{r tableBMIFDmodel, echo=FALSE}
-tab_BMIFDmodel <- mk_FCTables()[[3]]
+```{r tableFDmodel}
+# do not change chunk labelsas it corresponds to the manually defined (in table.R) table label
+tab_FDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[2]]
+tab_FDmodel
+```
+
+```{r tableBMIFDmodel}
+# do not change chunk label as it corresponds to the manually defined (in table.R) table label
+tab_BMIFDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[3]]
 tab_BMIFDmodel
 ```
-
-```{r tableFDmodel, echo=FALSE}
-tab_FDmodel <- mk_FCTables()[[2]]
-tab_FDmodel
-```
+<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
+There is a significant association between average BMI and the connectivity between PCC and three other clusters.
+Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
+<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
+```{r "BMIFDagesex_NACCcc_avgFD_act", out.width = "75%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACCcc_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
+```
+
+```{r "BMIFDagesex_NACCgsr_avgFD_act", out.width = "75%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACCgsr_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
+```
+
+```{r "BMIFDagesex_PCCcc_avgBMI_deact", out.width = "75%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
+
+## Effect on hypothalamic connectivity (exploratory analysis)
+...
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 

### 2021-02-01 manuscript/draft_rsfmri.Rmd
merged into branch and solved conflicts

@@
 ```{r carpetplot, out.width="49%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap=c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
-carpets=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/carpetplotadi006fu2.jpeg","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/carpetplotadi16fu.jpeg")
+carpets=c("../report/carpetplotadi006fu2.jpeg","../report/carpetplotadi16fu.jpeg")
 knitr::include_graphics(carpets)
@@
 
-```{r calculate summary of QA, eval=TRUE, echo=FALSE}
-for (i in 1:nrow(final)) {
-  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
-  conf = read.csv(
-    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",
-      tmp, "/confounds.csv"))
-  tmp = cor.test(conf$DVARS, conf$FD)
-  final[i, "corr_DVARS_FD"] = tmp$estimate
-  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
-  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
-  final
-}
-write.csv(file = "report/all_QA_measures.csv", final)
-
 ```{r tableQcfc, echo=FALSE}
@@
 
-```{r prepare tsnr ROI plot, eval=TRUE}
-final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
-
-tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
-tsnr=tsnr[tsnr$V1!="subj",]
-colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
-levels(tsnr$roi)=droplevels(tsnr$roi)
-levels(tsnr$roi)=c("NAcc", "PCC/precuneus")
-tsnr$mean=as.double(tsnr$mean)
-
-final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
-
-p<-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-  geom_violin(aes(fill=roi), alpha=0.5) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") +
-  scale_fill_manual(values=wes_palette("Darjeeling2",5,type="discrete")[2:3]) +
-  theme(axis.text=element_text(size=10),
-  axis.title=element_text(size=12),
-  strip.text = element_text(size=12),
-  legend.position="")
-ggsave("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg", height = 6.604, units="cm", p)
-```
-
-```{r figNetworks, echo=FALSE, out.width="33%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
-
-#networks=c("./report/DMN_thresholded_for_clusterFWE005.png","./report/Rew_thresholded_for_clusterFWE005.png")
-networks=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE005.png", "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE005.png")
-tsnr=c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tsnr.jpeg")
-```
 ```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
@@
 
-```{r figNetworks, echo=FALSE, out.width="49%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens"),fig.show='hold',fig.align='center'}
+```{r figNetworks, echo=FALSE, out.width="33%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+
 networks = c("../report/fig/DMN_thresholded_for_clusterFWE005.pdf",
@@
 knitr::include_graphics(networks)
+tsnr <- "../report/tsnr.jpeg"
 knitr::include_graphics(tsnr)
@@
 ```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%",  eval=FALSE}
-readRDS(file = "../report/fig/fig_tSNR.rds")
+knitr::include_graphics(tsnr)
+#readRDS(file = "../report/fig/fig_tSNR.rds")
 ```
@@
 final_FC <- read.csv("../report/final_FC.csv")
-
-agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
-agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
-agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
-agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
-for (i in c(1:nrow(agg_FC_DMN_ID))){
-  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
-  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
-}
-
-colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")
-colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID_tp")
-
-final_FC=merge(final, agg_FC_Rew, by="subj.ID_tp", all.x=TRUE)
-final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
-final_FC$subj.ID=as.factor(final_FC$subj.ID)
-final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")
-
-#calculate mean BMI over timepoints
-final_FC.meanBMI=tapply(X=final_FC$BMI,
-                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
-final_FC$mean.BMI=
-  final_FC.meanBMI[as.numeric(final_FC$subj.ID)]
-final_FC$within.BMI=
-  final_FC$BMI-final_FC$mean.BMI
-
-#calculate mean FD over timepoints
-final_FC.meanFD=tapply(X=final_FC$logmFD,
-                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
-final_FC$mean.logmFD=
-  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
-final_FC$within.logmFD=
-  final_FC$logmFD-final_FC$mean.logmFD
-
-```
+```
+
 ```{r figaggDMN, fig.cap="DMN connectivity per group over time", out.width="75%", echo=FALSE}
-fig_DMNconn<-mk_figDMNdescr(final_FC)
+fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
 fig_DMNconn
 ```
+
 ```{r figaggRew,fig.cap="Reward network connectivity per group over time", out.width="75%", echo=FALSE}
-fig_Rewconn<-mk_figRewdescr(final_FC)
+fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
 fig_Rewconn
@@
 
-```{r figDMNconn, echo=FALSE}
-readRDS(file = "../report/fig/figDMNconn.rds")
-```
-
 ```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
@@
 ```
+
 ```{qa check Rew model age,sex, echo=FALSE}
@@
 texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-
-```{r figRewconn, echo=FALSE}
-readRDS(file = "../report/fig/figRewconn.rds")
 ```
@@
 ```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap=c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
-knitr::include_graphics(c("../report/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
+knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
 ```
@@
 
-```{r "BMIagesexfd_PCCcc_avgBMI_deact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "85%"}
-knitr::include_graphics("../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
-```
-
-```{r "BMIagesex_PCCcc_avgBMI_deact", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data, adjusted for BMI change, average mFD and change in mFD", fig.hold=TRUE, out.width = "85%"}
-knitr::include_graphics("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf")
-```
-
 <!-- ### Reward -->
@@
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction (p=$0.028$), also when evaluating the GSR corrected data (p=$0.038$).
-```{r "BMIFDagesexNACCcccgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI, average mFD and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
-knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
+```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
+                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
 ```
@@
 This result remained similar when using GSR corrected data.
-```{r "BMIFDagesexNACCccavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
-knitr::include_graphics(c("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf","/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
+```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
+                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
 ```
@@
 ```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap="Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
-knitr::include_graphics("../report/FDagesex_NACCcc_avgFD_act.pdf")
-knitr::include_graphics("../report/FDagesex_NACCgsr_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/FDagesex_NACCgsr_avgFD_act.pdf")
 ```
@@
 <!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
-```{r "BMIFDagesex_NACCcc_avgFD_act", out.width = "75%"}
+```{r "BMIFDagesex_NACC_avgFD_act", fig.show = "hold", out.width = "50%", fig.cap="Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf")
-```
-
-```{r "BMIFDagesex_NACCcc_cgnBMI_deact", out.width = "75%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf")
+```
+
+```{r "BMIFDagesex_NACC_cgnBMI_deact", fig.show = "hold", out.width = "50%", fig.cap="Positive association of change in BMI and FC of NAcc with cluster, adjusted for age, sex, average BMI, average mFD and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
-```
-
-```{r "BMIFDagesex_NACCgsr_avgFD_act", out.width = "75%"}
-knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf")
-```
-
-```{r "BMIFDagesex_NACCgsr_cgnBMI_deact", out.width = "75%"}
 knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")

### 2021-02-01 code/figures.R

@@ -1,3 +1,5 @@
+library(ggplot2)
+library(wesanderson)
 
 pink_blue = c("#00305E", "#C00045")
 blue_organge = c("#046C9A", "#D69C4E")
@@ -22,8 +24,7 @@ mk_figBMIdescr <- function(final) {
       geom  = "line",
       size = 2,
       color = c(rep("#046C9A", 3), rep("#D69C4E", 3))
-    ) +
-  figBMIdescr= figBMIdescr +
+    ) + 
     scale_x_discrete(labels = c("bl" = "0", "fu" = "6", "fu2" = "12")) +
     scale_color_manual(
       values = c("#046C9A70", "#D69C4E70"),
@@ -296,23 +297,23 @@ mk_figDesignMatrix <- function() {
 
 mk_figtSNR <- function (final) {
   final$subj.ID_tp=paste0(final$subj.ID,'_', final$tp)
-
+  
   tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
   tsnr=tsnr[tsnr$V1!="subj",]
   colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
   levels(tsnr$roi)=droplevels(tsnr$roi)
-  levels(tsnr$roi)=c("NAcc", "precuneus")
+  levels(tsnr$roi)=c("NAcc", "PCC/precuneus")
   tsnr$mean=as.double(tsnr$mean)
-
+  
   final_tsnr=merge(final, tsnr, by="subj.ID_tp", all.x=TRUE)
-
-  p <- ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
-    geom_violin(aes(fill=roi)) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") +
-    scale_fill_manual(values=wes_palette("Darjeeling2",4,type="discrete")[2:3]) +
-    theme_bw() + theme(axis.text=element_text(size=10),
-                       axis.title=element_text(size=12),
-                       strip.text = element_text(size=12),
-                       legend.position="")
+  
+  p <-ggplot(aes(as.factor(roi), mean),data=final_tsnr) +
+    geom_violin(aes(fill=roi), alpha=0.5) + geom_jitter(height = 0, width = 0.1) + xlab("Region of interest") + ylab("average tSNR") +
[…]

### 2021-02-01 code/process.R

@@ -1,3 +1,5 @@
+library(kableExtra)
+
 # processes data to report later in manuscript .Rmd file
 ROOT_DIR = "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/"
 
@@ -14,7 +16,6 @@ source("../../../../../Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/
 
 final <- create_sample_df(group = "both", tp = "all")
 
-# calculate summary of QA -------------------------------------------------
 
 for (i in 1:nrow(final)) {
   tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
@@ -30,6 +31,7 @@ for (i in 1:nrow(final)) {
   final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
   final
 }
+#write.csv(file = "../report/all_QA_measures.csv", final)
 
 final <- within(final,  subj.ID_tp <- paste(subj.ID,tp, sep="_"))
 
@@ -114,16 +116,16 @@ colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH'
 # prepare tsnr ROI plot ---------------------------------------------------
 
 fig_tSNR <- mk_figtSNR(final)
-saveRDS(object = fig_tSNR, file = "../report/fig/fig_tSNR.rds")
+ggsave("../report/tsnr.jpeg", height = 6.604, units="cm", p)
+#saveRDS(object = fig_tSNR, file = "../report/fig/fig_tSNR.rds")
 
 
 # load and prepare aggFC measures -----------------------------------------
 
 agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
 agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
-agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
[…]

### 2021-02-01 manuscript/draft_rsfmri.Rmd
problem with subfloat in float - does not knit

@@
 editor_options:
-
   chunk_output_type: console
@@
 options(knitr.kable.NA = '') # hide NA in tables
+knitr::knit_engines$get()
 # set working directory of knit (RMarkdown) to directory of Rproj
@@
 
-```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
+```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
@@
 
-```{qa check DMN model age,sex, echo=FALSE}
+```{r "qa check DMN model age,sex", echo=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
+```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
+```{r "calc DMN two tp, age and sex and mFD", results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 
-```{qa check DMN model age,sex, mFD, echo=FALSE}
+```{r, "qa check DMN model age,sex, mFD", echo=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
+```{r "DMN2tp results age sex mFD", caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r calc DMN three tp, eval = TRUE, echo = FALSE}
+```{r "calc DMN three tp", eval = TRUE, echo = FALSE}
 #### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 
-```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+```{r "DMN3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
+```{r "calc rew two tp, age, sex", eval = TRUE, echo = FALSE}
 ### Reward network for two time points, adjusted for age and sex (Model 1a)
@@
 
-```{qa check Rew model age,sex, echo=FALSE}
+```{r, "qa check Rew model age,sex", echo=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+```{r "Rew2tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
+```{r "calc rew two tp adjusted", eval = TRUE, echo = FALSE}
 #### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 
-```{qa check Rew model age,sex,mFD, echo=FALSE}
+```{r "qa check Rew model age, sex, mFD", echo=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+```{r "Rew2tpadj results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r calc rew three tp, eval = TRUE, echo = FALSE}
+```{r "calc rew three tp", eval = TRUE, echo = FALSE}
 #### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 ```
-```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
+
+```{r "Rew3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 #### Aggregated FC analysis
-```{r DMN model, mean, change BMI, eval = TRUE, echo = FALSE}
+```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
 R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
@@
 Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD.
-```{r Rew model, mean, change BMI,, eval = TRUE, echo = FALSE}
+```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
 R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
@@
 
-```{r DMN model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
+```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
 R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
@@
 #### Reward network, mean and within variability in BMI and logmFD
-```{r Rew model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
+```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
 R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)

### 2021-02-01 manuscript/draft_rsfmri.tex
problem with subfloat in float - does not knit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Effects of bariatric surgery and weight loss on resting state functional connectivity},
  pdfauthor={Hannah Sophie Heinrichs* {[}1{]}; Frauke Beyer* {[}1{]}{[}2{]}; Kristin Prehn; Jürgen Ordemann; Agnes Floel; A. Veronica Witte {[}1{]}{[}2{]}{[}3{]}},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage[left]{lineno}
\linenumbers
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Effects of bariatric surgery and weight loss on resting state functional connectivity}
\author{Hannah Sophie Heinrichs* {[}1{]} \and Frauke Beyer* {[}1{]}{[}2{]} \and Kristin Prehn \and Jürgen Ordemann \and Agnes Floel \and A. Veronica Witte {[}1{]}{[}2{]}{[}3{]}}
\date{1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline 2 CRC 1052 ``Obesity Mechanisms'', Subproject A1, University of Leipzig \newline 3 Day Clinic for Cognitive Neurology, University Clinic Leipzig}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\newpage

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Obesity is a worldwide health issue, enchaining huge personal and societal costs. Excess amount of fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life.
Conservative treatment options including dietary modification, physical activity and behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI (\textgreater{} 35 kg/m2). Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
Roux-en-Y gastric bypass (RYGB) is the most effective surgical procedure (Berthoud, Shin, and Zheng", n.d.). Here, a small pouch is formed from the proximal stomach and connected to the jejunum, disconnecting the larger part of the stomach and the proximal part of the small intestine intact from the digestive flow. Other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) reduce stomach volume, but leave the small intestine unchanged (Mulla, Middelbeek, and Patti 2017).
RYGB-induced weight loss is in part mediated by malabsorption of certain nutrients, especially fat, and reduced digestion efficiency (Berthoud, Shin, and Zheng", n.d.) while the role of increased energy expenditure after RYGB is still unclear (Yarmush, D'Alessandro, and Saeidi 2017; Lamarca et al. 2019). Most importantly, RYGB alters the signaling cascades between the gut, fat tissue and the brain and thereby induces profound changes in the regulation of food intake and energy homeostasis (Berthoud, Shin, and Zheng", n.d.). RYBG-patients frequently report altered smell and taste perception, i.e.~increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items (Mulla, Middelbeek, and Patti 2017, @Berthoud\_2011). These changes in appetitive and reward signaling likely relate to altered gut-brain communication, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin (Brutman, Sirohi, and Davis 2019).
Yet, precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
Resting state functional magnetic resonance imaging (rsfMRI) is a powerful technique to investigate the dynamic organization of the brain via the correlation of spontaneous neural activity. Anatomically distinct brain regions show functional connectivity (FC), i.e.~correlated neural activity over time, in the absence of a cognitive task. Crucially, these brain networks correspond to networks activated during specific cognitive processes, e.g.~visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e.~attention and executive control (Smith et al. 2009), and can thus be seen as default configuration of the brain. In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.
The default mode network has been discovered as a ``task-negative'' network, and is involved in a variety of largely internally-directed cognitive functions including interoception, mind wandering and memory. Its core hubs are the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx. BA 39) (Raichle 2015).
Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies (Beyer et al. 2017; Kullmann et al. 2011; Chao et al. 2018; Wijngaarden et al. 2015, @Borowitz\_2020; Ding et al. 2020; Doucet et al. 2017; Sadler, Shearrer, and Burger 2018).
Despite the large heterogenity and low median power, these findings might be summarized as a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.~salience and sensory networks. This has been interpreted as impaired interoceptive processing failing to adequately balance internal and external behavioral cues.
Longitudinal studies largely support this view. Increased activation of the DMN in a food-viewing paradigm, i.e.~insufficient down-regulation of DMN, has been shown to correlate with appetite measures after an overnight fast (Tregellas et al. 2011, @Wijngaarden\_2015), an effect that could be partially reversed by an exercise intervention (McFadden et al. 2013). In a bariatric surgery study, (Li et al. 2018) reported a reduction of local FC in mid-line DMN regions in surgical patients compared to an obese control group. Simultaneously, PCC and medial prefrontal cortex were more strongly connected to dorsolateral prefrontal cortex, a region implicated in cognitive control. Another study in bariatric surgery patients showed that heightened connectivity of the PCC with anterior superior temporal gyrus (STG) before the surgery decreased over the course of one year (Olivo et al. 2017). Similarly, the profile of DMN connectivity with frontal regions, including ACC, frontal superior gyrus and the right OFC, was similar in patients after bariatric surgery and normal weight controls, but dissimilar to that of obese controls (Frank et al. 2013).
Another important brain network in the context of obesity is the reward network, comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) (O'Doherty 2004; Liu et al. 2011). These brain regions have been found to guide food valuation processes and decision-making in humans (Bartra, McGuire, and Kable 2013; Hare, Malmaud, and Rangel 2011; Hutcherson et al. 2012; Schmidt et al. 2018). Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced reward region activation to actual taste of these foods {[}Devoto et al. (2018); García-García et al. (2014); Meng et al. (2020); Stoeckel et al. (2009); though see Morys, García-García, and Dagher (2020);{]}. Food cue hyperactivation has been identified as a main neural vulnerability factor predicting future weight gain (Stice, Burger, and Yokum 2015). Resting state fMRI studies have partly mirrored these results, showing increased local FC of reward network regions, i.e.~Nacc, vmPFC, putamen, insula (Contreras-Rodríguez et al. 2017; Coveleskie et al. 2015; Hogenkamp et al. 2016), and altered connectivity with salience, homeostatic and sensorimotor networks (Lips et al. 2014; Wijngaarden et al. 2015). Imaging studies in bariatric surgery patients provided first evidence that hyperconnectivity between reward network regions (e.g.~putamen, OFC) might be normalized to some extent by the surgery (Duan et al. 2020; Wiemerslage et al. 2016). Possible, the reconfiguration of the fronto-striatal brain networks might be related to altered gut signaling , e.g.~changes in ghrelin levels, via hypothalamic-striatal projections (Li et al. 2019; Karra et al. 2013, but @Zoon\_2018) .
Thus, while there is some evidence hinting to a role for DMN and reward network connectivity in altered gut-brain communication after bariatric surgery, the existing evidence is inconclusive. Most studies have investigated small cohort of patients, often without adequate control group, and employed exploratory statistics (George et al. 2016).
In this study, we aimed to extend the current literature and investigate differences of DMN and reward network connectivity before and after bariatric surgery in preregistered analyses and in comparison with a contemporaneously measured waiting-control group. Moreover, we explored whether the changes in functional connectivity were associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential dose-response relationship.
Participants were measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions (Shai et al. 2008). Notably, we took into account possible non-linear time courses in increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management (Olivo et al. 2017).

\newpage

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\begin{table}

\caption{\label{tab:tableSample}Distribution of data points at months after intervention}
\centering
\begin{tabular}[t]{lrr}
\toprule
  & BARS & NBARS\\
\midrule
count: only 0 & 7 & 2\\
count: only 6 & 3 & 0\\
count: only 12 & 2 & 0\\
count: 0 and 6 & 5 & 3\\
count: 6 and 12 & 7 & 1\\
count: complete data & 9 & 9\\
total number of subjects & 33 & 15\\
total data points & 63 & 37\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{sample-and-study-design}{%
\subsection{Sample and study design}\label{sample-and-study-design}}

The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité Berlin. For more details see Prehn et al. (2020). We used all data acquired until April 2019.\\
The study protocol was in accordance with the Helsinki Declaration and approved by the local ethics committee. The study was registered at clinicaltrials.gov as NCT01554228. We preregistered the present resting state analyses on the \href{https://osf.io/yp42s}{OSF}. We made \href{https://osf.io/59bh7/}{additional changes} to the preregistration after preprocessing as we realized some aspects of the analysis were inadequately described in the first preregistration.\\
Participants were recruited from the Center for Bariatric and Metabolic Surgery. Inclusion criteria were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI \(> 40\) km/m\(^2\) or (2) a BMI \(> 35\) kg/m\(^2\) and at least one typical co-morbidity (e.g.~type-2 diabetes, hypertension, non-alcoholic fatty liver disease)(Mechanick et al. 2013).
Participants were aged between \(18\) and \(70\) and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score \(< 34\)).
In total, \(51\) participants out of the originally enrolled \(69\) subjects received MRI, and we analyzed a total of \(101\) data points. Five data points of three subjects had to be excluded due to bad anatomical image quality (see section @ref\{sec:Anatomical Quality control\}).
The final sample entailed 48 morbidly obese individuals (37 females; aged 44.15 \(\pm\) 11.86 SD years (range 21-68). Participants either underwent surgery (n = 33, 26 female) or were waiting list controls (n = 15, 11 female), who waited for their health insurance's approval to undergo surgery. Measures were taken at baseline (BL), \(6\) (FU1) and \(12\) (FU2) months post-surgery or after the baseline appointment. Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \ref{tab:tableSample}). In more detail, 18 participants had complete data, 16 provided data for two time points and 14 for only one time point. In the intervention group, data is not missing completely at random, as the pre-surgery timepoint MRI could not be obtained in some patients due to the limited scanner bore.

Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = 4 vs.~n = 10) and a higher mean age (47.14 vs.~39.33). Change in BMI throughout the study is depicted in \ref{fig:figBMIdescr}.
As intervention, \(0\) participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and \(0\) surgery purely restricting energy intake (VSG: n = \(0\), GB: n = \(0\)).

Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast. They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for breakfast. MRI scanning was done after further psychological tests (for details see Prehn et al. (2020)).

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{draft_rsfmri_files/figure-latex/figBMIdescr-1}

}

\caption{BMI over time.}\label{fig:figBMIdescr}
\end{figure}

Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = 4 vs.~n = 10) and a higher mean age (47.14 vs.~39.33). Change in BMI throughout the study is depicted in \ref{fig:figBMIdescr}.
As intervention, \(0\) participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and \(0\) surgery purely restricting energy intake (VSG: n = \(0\), GB: n = \(0\)).
Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast. They underwent medical assessments including an interview, blood draw and anthropocentric measurements before having a one our break for breakfast. MRI scanning was done after further psychological tests (for details see Prehn et al. (2020)).

\hypertarget{fmri-acquisition}{%
\subsection{fMRI acquisition}\label{fmri-acquisition}}

MRI images were retrieved with a 12-channel head coil of a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
T1-weighted anatomical images were acquired as described in Prehn et al. (2020) (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = 1 x 1 x 1 mm\(^3\), 192 sagital slices).
Resting state echo-planar imaging was acquired with a TR of 2.3 s and TE of 30 ms. The image matrix was 64 x 64 with an in-plane resolution of 3 mm x 3 mm and 34 slices with a slice thickness of 4 mm. 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes. Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneties.
Participants were instructed to close their eyes but to remain awake during scanning.

\hypertarget{preprocessing}{%
\subsection{Preprocessing}\label{preprocessing}}

Imaging data was conducted using AFNI 19.1.05, ANTS `2.3.1', FSL `6.0.1' and FreeSurfer `6.0.0p1', wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15.
The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel. The workflow can be found on \href{https://github.com/fBeyer89/ADI_preproc/}{github}.

\hypertarget{anatomical-preprocessing}{%
\subsubsection{Anatomical preprocessing}\label{anatomical-preprocessing}}

Within the module for anatomical preprocessing, the T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline (Fischl 2012). Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs (Reuter and Fischl 2011). Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's \texttt{aseg.mgz} segmentation file for quality control of rs preprocessing.\\
The skull-stripped brain (\texttt{brainmask.mgz}) was then coregistered to the MNI152 1x1x1x mm template using ANTS (Avants et al. 2009). The following parameters were used for rigid {[}metric: mutual information, number of iterations: {[}1000, 500, 250, 100{]}{]}, affine {[}metric: mutual information, number of iterations: {[}1000, 500, 250, 100{]}{]} and SyN: {[}metric: cross-correlation, iterations: {[}100, 70, 50, 20{]}{]} registration steps with shrink factors {[}8, 4, 2, 1{]}, and smoothing sigmas={[}3, 2, 1, 0{]} for each step.
\emph{In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.}

\hypertarget{functional-preprocessing}{%
\subsection{Functional preprocessing}\label{functional-preprocessing}}

\hypertarget{minimal-functional-preprocessing}{%
\subsubsection{Minimal functional preprocessing}\label{minimal-functional-preprocessing}}

The function preprocessing module included the removal of first four volumes, motion correction (FSL's \texttt{MCFLIRT}), fieldmap distortion correction (FSL's \texttt{fsl\_prepare\_fieldmap} and \texttt{FUGUE}) and coregistration to the subject's individual longitudinal anatomical space (FreeSurfer's \texttt{bbregister}). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (\texttt{fslmaths\ -kernel\ gauss\ 2.548}).

\hypertarget{ica-aroma-preprocessing}{%
\subsubsection{ICA-AROMA preprocessing}\label{ica-aroma-preprocessing}}

We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relationship (Parkes et al. 2018). As described in (Pruim et al. 2015), the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's \texttt{melodic} and nuisance components were classified according to four spectral and spatial criteria. Finally, the timeseries of the nuisance components were regressed from the raw fMRI timeseries.
As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result. We used CompCor to estimate 5 variance components from a combined WM and CSF mask, which was further eroded using \texttt{fslmaths\ -nan\ -thr\ 0.99\ -ero\ -bin} (Behzadi et al. 2007; Muschelli et al. 2014). After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
\emph{This analysis was described in the preregistration as ``Then we will employ state-of-the art nuisance regression (i.e.~ICA-AROMA with compCor regression (CC) and, optionally global signal regression (GSR) (Ciric et al., 2017; Parkes et al., 2018) to reduce the effect of physiological noise and head motion on the connectivity estimates.''}

\hypertarget{global-signal-regression}{%
\subsubsection{Global Signal regression}\label{global-signal-regression}}

Ciric et al. (2017) showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone. Yet, it introduces spatial dependency and spurious correlations. Previously, we reported on head motion differences between groups over time in this study (Beyer et al. 2020), and thus it was very important to minimize the motion-FC correlation. The global signal was derived from the average of all voxels in the brain mask. Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA fMRI data and high-pass filtered the resulting file as above.
\emph{In the preregistration we stated " As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but introduces spatial dependency and spurious correlations, we perform GSR after ICA-AROMA in a separate sensitivity analysis."}

\hypertarget{functional-connectivity}{%
\subsubsection{Functional connectivity}\label{functional-connectivity}}

To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed regions of interest (ROI), respectively. We chose these regions as they are core hubs of the networks. Because we knew about the large amount of motion in this dataset, we decided against Independent Component Analysis (ICA) to construct the networks, as ICA is recommended to work on minimally preprocessed data \href{https://sourceforge.net/p/icatb/mailman/message/31466068/}{GSR data}. Based on FreeSurfer's \texttt{aseg.mgz} and Desikan-Killiany parcellation, we created seed masks using \texttt{mri\_binarize} (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres. Then, we used \texttt{NiftiLabelsMasker} and \texttt{NiftiMasker} to extract the standardized timeseries from the seed regions and the whole brain. We calculated the Pearson's correlation between them with \texttt{numpy.dot}, performed \texttt{r-to-z} transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr). Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
\emph{In the preregistration we stated "Resting-state functional connectivity (FC) using seed-based connectivity analysis extracted with in-house pre-processing pipelines (see (Beyer et al., 2019)): Whole-brain FC maps for a priori defined seed regions (defined based on the results of the subcortical segmentation and Desikan-Killiany cortical parcellation in the Freesurfer version 6.0.0 longitudinal processing stream). For the reward network a ROI combined from bilateral Nucleus accumbens from the subcortical segmentation (26, Left-Accumbens-area, 58, Right-Accumbens-area) was selected. Ventro-medialprefrontal cortex was not selected as seed region due to low signal-to-noise ratio in this part of the brain in many participants (noticed during quality checking). For the default mode network, the seed region was combined from bilateral precuneus from the Desikan-Killiany atlas (1025 ctx-lh-precuneus, 2025 ctx-rh-precuneus). We will calculate voxel-wise seed-based connectivity mapsin the individual subject's space (which is coregistered to the Freesurfer longitudinally preprocessed timepoint) and apply r-to-Z-transform to these images. These steps will be performed using nilearn, nipype and python. Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).}

\hypertarget{quality-assessment}{%
\subsubsection{Quality Assessment}\label{quality-assessment}}

\hypertarget{anatomical-quality-control}{%
\paragraph{Anatomical Quality control}\label{anatomical-quality-control}}

FreeSurfer cross-sectional and longitudinal segmentations were visually checked according to Klapwijk et al. (2019). We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.

\hypertarget{functional-quality-control}{%
\paragraph{Functional Quality control}\label{functional-quality-control}}

Resting state fMRI data quality control was performed according to the protocol by Ciric et al. (2018).
As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals (Beyer et al. 2020). Thus, we did not exclude any participants based on mean framewise displacement (FD) estimates beyond those who failed the anatomical preprocessing. We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations (Ciric et al. 2018). In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10\% data sets with highest mean FD (mFD).

\begin{figure}

{\centering \subfloat[Example 1 with strong amount of structured noise\label{fig:carpetplot-1}]{\includegraphics[width=0.49\linewidth,height=0.3\textheight]{../report/carpetplotadi006fu2} }\subfloat[Example 2 with less amount of structured noise\label{fig:carpetplot-2}]{\includegraphics[width=0.49\linewidth,height=0.3\textheight]{../report/carpetplotadi16fu} }

}

\caption{Carpet plot for quality control}\label{fig:carpetplot}
\end{figure}

On the single-subject level, we used \texttt{plot\_carpet} from \texttt{niworkflows.viz.plots} to generate carpet plots which depict denoising quality. Figure \ref{fig:carpetplot} shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, \texttt{OutlierCount}, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum). All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in \texttt{QA\ residuals}.
We evaluated mFD-FC relationships by computing a functional connectome based on (Power et al. 2012). In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in \texttt{NiftiSpheresMasker}, and calculated the Spearman rank correlation between these time series. Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure @ref\{exampleFC\}. We reported the mFD-FC correlation per preprocessing scheme.
Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.

\hypertarget{aggregated-fc-values}{%
\subsubsection{Aggregated FC values}\label{aggregated-fc-values}}

To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and timepoints, adjusted for age and sex. We used GSR regressed data as input and clusterwise bootstrapping with N=999. Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of p\_FWE \textless{} 0.05. We extracted the average connectivity from these masks for GSR regressed data.

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, MATLAB (2018)) using the SwE toolbox version 2.2.2 (Guillaume et al. 2014) as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al. (2014)), and R version 3.6.1 (Team and others 2013).

\hypertarget{design-specification}{%
\subsubsection{Design Specification}\label{design-specification}}

In a first group of models, we tested the interaction of time and group \(FC = group + time + group*time\) as specified in the preregistration (\emph{We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version \textgreater{} 9.0) to calculate the group-by-timepoint interaction})
We did not perform the analysis as preregistered (i.e.~without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \ref{fig:figDesignmatrix}). The resulting factorial design contained one regressor for each time point per group.

\begin{figure}

{\centering \subfloat[Design matrix with time as categorial factor.\label{fig:figDesignmatrix-1}]{\includegraphics[width=0.5\linewidth]{draft_rsfmri_files/figure-latex/figDesignmatrix-1} }\subfloat[Design matrix with time as continuous variable.\label{fig:figDesignmatrix-2}]{\includegraphics[width=0.5\linewidth]{draft_rsfmri_files/figure-latex/figDesignmatrix-2} }

}

\caption{Possible design matrices of time-by-group model.}\label{fig:figDesignmatrix}
\end{figure}

We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
As proposed by Guillaume et al. (2014), we calculated the between- and within subject centered values of BMI. This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC. We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
As we previously reported correlated change in BMI and mFD in this sample (Beyer et al. 2019), we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3). Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: \(FC = between-subject FD + within-subject FD\) with nuisance covariates age and sex.

\hypertarget{model-input-and-brain-mask}{%
\subsubsection{Model input and brain mask}\label{model-input-and-brain-mask}}

For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC with GSR), resulting in \(8\) preregistered analyses , and \(16\) exploratory whole brain analysis.
For all analyses, we used an explicit mask, which was either a whole brain or a GM mask derived from the MNI ICBM ``152 nonlinear 6th generation'' atlas (re-sampled to 3 x 3 x 3 mm\(^3\) and thresholded at 0.5 GM probability). This resulted in a reduced number of tests (voxels of interest) from \(52378\) to \(29451\) for the GM mask.
Computation with the grey matter mask were exploratory and not planned in the preregistration.

\hypertarget{swe-settings}{%
\subsubsection{SwE settings}\label{swe-settings}}

The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. It therefore accommodates any repeated measurement covariance structures. We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
In contrast to the parametric estimation, wild bootstrap avoids any parametric distributional assumptions (e.g.~random field theory), and unlike permutation testing it is also suitable for longitudinal data.
Unlike the description in the preregistration, we used a cluster forming threshold of \(p < 0.001\) for a more rigorous multiple comparison adjustment (instead of \(p < 0.01)\), and \(1000\) bootstraps due to required computation time (instead of 5000). Significant clusters are defined as family-wise error (FWE) corrected \(p < 0.05\). The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c (Eickhoff et al. 2005).

\hypertarget{aggregated-fc-analysis}{%
\subsubsection{Aggregated FC analysis}\label{aggregated-fc-analysis}}

We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 (Bates et al. 2007).
Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up. In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing \emph{sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI} (as stated in the preregistration).
We performed model comparison between R1 and R0 models
\texttt{R1=\ lmer(agg\ FC\ \textasciitilde{}\ timepoint*group\ +\ age\ +\ sex\ +\ (1\textbar{}subj))}
\texttt{R0\ =\ lmer(agg\ FC\ \textasciitilde{}\ timepoint\ +\ group\ +\ \ +\ age\ +\ (1\textbar{}subj))}
and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed \(p < 0.05\) compared with \texttt{anova}.
As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points. Again, the interaction was considered significant if a model comparison between R1 and R0 yielded \(p < 0.05\).
Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.

\newpage

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{quality-control}{%
\subsection{Quality control}\label{quality-control}}

We performed different quality controls for the rsfMRI data.
First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
As described in Beyer et al. (2020), there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group.
Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points.
We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.

Further, we evaluated the impact of our denoising pipelines on different quality metrics following (Ciric et al. 2018). Table \ref{tab:tableQcfc} shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = 0.08). This relationship was negatively associated with distance between nodes (mean Spearman's r = -0.3).\\
Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency. Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency (Power, Schlaggar, and Petersen 2015). Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = 0.01), and minimal positive distance dependence (mean Spearman's r = 0.07). Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.

\begin{table}

\caption{\label{tab:tableQcfc}Summary of quality metrics for different denoising pipelines across conditions and time points}
\centering
\begin{tabular}[t]{l>{\raggedleft\arraybackslash}p{1.8cm}>{\raggedleft\arraybackslash}p{1.8cm}>{\raggedleft\arraybackslash}p{1.8cm}>{\raggedleft\arraybackslash}p{1.8cm}>{\raggedleft\arraybackslash}p{1.8cm}r}
\toprule
  & V1 & V2 & V3 & V4 & V5 & V6\\
\midrule
1 & 0.076 & 0.075 & 7636 & 1814 & -0.296 & 0.000\\
2 & 0.064 & 0.067 & 5092 & 27 & -0.017 & 0.001\\
3 & 0.033 & 0.037 & 5840 & 227 & 0.061 & 0.000\\
4 & 0.005 & 0.011 & 5988 & 464 & 0.071 & 0.000\\
\bottomrule
\end{tabular}
\end{table}

Further, we evaluated the impact of our denoising pipelines on different quality metrics following (Ciric et al. 2018). Table \ref{tab:tableQcfc} shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = 0.08). This relationship was negatively associated with distance between nodes (mean Spearman's r = -0.3).\\
Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency. Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency. Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = 0.01), and minimal positive distance dependence (mean Spearman's r = 0.07). Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{draft_rsfmri_files/figure-latex/figDvarsmFD-1}

}

\caption{Summary of quality measures DVARS and mFD}\label{fig:figDvarsmFD}
\end{figure}

\begin{figure}

{\centering \subfloat[DMN seeded from precuneus\label{fig:figNetworks-1}]{\includegraphics[width=0.33\linewidth,height=0.3\textheight]{../report/fig/DMN_thresholded_for_clusterFWE005} }\subfloat[Reward network seeded from Nucleus Accumbens\label{fig:figNetworks-2}]{\includegraphics[width=0.33\linewidth,height=0.3\textheight]{../report/fig/Rew_thresholded_for_clusterFWE005} }\subfloat[tSNR differences between regions\label{fig:figNetworks-3}]{\includegraphics[width=0.33\linewidth,height=0.3\textheight]{../report/tsnr} }

}

\caption{Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)}\label{fig:figNetworks}
\end{figure}

Figure \ref{fig:figNetworks} shows the DMN and reward network average maps which we used for extracting aggregated FC values. Both networks included the typically described brain regions (i.e.~medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \ref{fig:figNetworks}).

\hypertarget{interaction-effect-of-intervention-and-time-preregistered-analysis}{%
\subsection{Interaction effect of intervention and time (preregistered analysis)}\label{interaction-effect-of-intervention-and-time-preregistered-analysis}}

\hypertarget{dmnreward-network-with-whole-brain-mask-preregistered-analysis}{%
\subsubsection{DMN/reward network with whole brain mask (preregistered analysis)}\label{dmnreward-network-with-whole-brain-mask-preregistered-analysis}}

\hypertarget{model-1a-with-adjustment-for-age-sex}{%
\paragraph{Model 1a (with adjustment for age, sex)}\label{model-1a-with-adjustment-for-age-sex}}

Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the PCC nor the NAcc seeds. There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.

\hypertarget{model-1b-with-adjustment-for-age-and-sex-meanfd}{%
\paragraph{Model 1b (with adjustment for age and sex, meanFD)}\label{model-1b-with-adjustment-for-age-and-sex-meanfd}}

Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.

Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on \href{insert\%20link\%20here}{NeuroVault}.

\hypertarget{interaction-effect-group-by-time-on-aggregated-fc-preregistered-analysis}{%
\subsection{Interaction effect group by time on aggregated FC (preregistered analysis)}\label{interaction-effect-group-by-time-on-aggregated-fc-preregistered-analysis}}

There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \ref{fig:figaggDMN} and \ref{fig:figaggRew} and supplements for a detailed summary of the models)

\begin{figure}
\includegraphics[width=0.75\linewidth]{draft_rsfmri_files/figure-latex/figaggDMN-1} \caption{DMN connectivity per group over time}\label{fig:figaggDMN}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\linewidth]{draft_rsfmri_files/figure-latex/figaggRew-1} \caption{Reward network connectivity per group over time}\label{fig:figaggRew}
\end{figure}

\begin{verbatim}
##    mean_DMN_conn mean_Rew_conn subj.ID_tp subj.ID  tp condition
## 2       0.063282      0.072570 ADI003_fu2  ADI003 fu2        IG
## 34      0.073654      0.038568  ADI041_fu  ADI041  fu        KG
\end{verbatim}

\begin{verbatim}
##   X subj.ID  tp Age_BL  Sex  BMI condition longtemplate.run long.runs
## 6 6  ADI003 fu2     53 male 40.9        IG                1         1
##   T1.comments DWI.comments rs.comments Preprocessed rs.QA.residuals rs.QA.GM
## 6                                                 1               1        0
##   rs.Qa.carpetplots.comment    Subject Exclude Final_Score Excessive_motion
## 6                        NA ADI003_fu2      NA           3                0
##   Temp_pole_missing_rh Temp_pole_miss_lh Dura_rh Dura_lh
## 6                    1                 0       1       1
##                                Check_comments Correction.done
## 6 dura: lh: sl 45-50; rh: 146:149, rh TL: 136               1
##                            Correction_comments Missing_Anterior_rh
## 6 corrected pial, wm rh temporal lobe manually                   0
##   Missing_Anterior_lh Missing_Superior_medial_rh Missing_Superior_Medial_lh
## 6                   0                          0                          0
##   Missing_Posterior_rh Missing_Posterior_lh long.timepoints_check    meanFD
## 6                    0                    0                    ok 0.1793491
##       maxFD original_tp
## 6 0.7365457         fu2
\end{verbatim}

\begin{verbatim}
##     X subj.ID tp Age_BL    Sex  BMI condition longtemplate.run long.runs
## 56 56  ADI041 fu     68 female 45.5        KG                1         1
##    T1.comments DWI.comments rs.comments Preprocessed rs.QA.residuals rs.QA.GM
## 56                                                 1               0        0
##    rs.Qa.carpetplots.comment   Subject Exclude Final_Score Excessive_motion
## 56                        NA ADI041_fu      NA           3                1
##    Temp_pole_missing_rh Temp_pole_miss_lh Dura_rh Dura_lh Check_comments
## 56                    0                 0       0       0               
##    Correction.done Correction_comments Missing_Anterior_rh Missing_Anterior_lh
## 56               1                                       0                   0
##    Missing_Superior_medial_rh Missing_Superior_Medial_lh Missing_Posterior_rh
## 56                          0                          0                    0
##    Missing_Posterior_lh long.timepoints_check    meanFD    maxFD original_tp
## 56                    0                    ok 0.5251626 1.153295          fu
\end{verbatim}

\begin{verbatim}
##    mean_Rew_conn subj.ID_tp subj.ID  tp condition
## 64      0.363538  ADI069_bl  ADI069  bl        IG
## 90      0.362268 ADI111_fu2  ADI111 fu2        KG
\end{verbatim}

\includegraphics{draft_rsfmri_files/figure-latex/identify outliers in these plots-1.pdf}

\begin{verbatim}
##     X subj.ID tp Age_BL    Sex  BMI condition longtemplate.run long.runs
## 93 94  ADI069 bl     52 female 45.6        IG                1         1
##    T1.comments DWI.comments rs.comments Preprocessed rs.QA.residuals rs.QA.GM
## 93                                                 1               1        1
##    rs.Qa.carpetplots.comment   Subject Exclude Final_Score Excessive_motion
## 93                        NA ADI069_bl      NA           2                0
##    Temp_pole_missing_rh Temp_pole_miss_lh Dura_rh Dura_lh
## 93                    0                 0       1       1
##                       Check_comments Correction.done Correction_comments
## 93 around slice 130 brain is missing               1                    
##    Missing_Anterior_rh Missing_Anterior_lh Missing_Superior_medial_rh
## 93                   0                   0                          0
##    Missing_Superior_Medial_lh Missing_Posterior_rh Missing_Posterior_lh
## 93                          0                    0                    0
##                                                                                                     long.timepoints_check
## 93 could be further corrected but no large parts of skull in brainmask, especially fu2 includes more skull but also other
##       meanFD    maxFD original_tp
## 93 0.3246258 1.250849          bl
\end{verbatim}

\begin{verbatim}
##       X subj.ID  tp Age_BL    Sex   BMI condition longtemplate.run long.runs
## 134 135  ADI111 fu2     45 female 38.86        KG                1         1
##     T1.comments DWI.comments rs.comments Preprocessed rs.QA.residuals rs.QA.GM
## 134                                                 1               1        1
##     rs.Qa.carpetplots.comment    Subject Exclude Final_Score Excessive_motion
## 134                        NA ADI111_fu2      NA           1                0
##     Temp_pole_missing_rh Temp_pole_miss_lh Dura_rh Dura_lh Check_comments
## 134                    0                 0       0       0               
##     Correction.done Correction_comments Missing_Anterior_rh Missing_Anterior_lh
## 134              NA                                       0                   0
##     Missing_Superior_medial_rh Missing_Superior_Medial_lh Missing_Posterior_rh
## 134                          0                          0                    0
##     Missing_Posterior_lh long.timepoints_check    meanFD     maxFD original_tp
## 134                    0                       0.2217564 0.6373868         fu2
\end{verbatim}

There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figures \ref{fig:figaggDMN} and \ref{fig:figaggRew} and supplements for a detailed summary of the models).

\includegraphics{draft_rsfmri_files/figure-latex/qa check DMN model age,sex-1.pdf} \includegraphics{draft_rsfmri_files/figure-latex/qa check DMN model age,sex-2.pdf}

\begin{verbatim}
## [1] "please carefully evaluate whether the result makes sense, and if not, please contact me"
\end{verbatim}

\begin{verbatim}
## 
## none 
##   46
\end{verbatim}

\begin{verbatim}
##                                          what          orig           min
## (Intercept)                       (Intercept)  0.3011366890  0.2882606175
## tpfu                                     tpfu -0.0042961373 -0.0075864874
## group_factorKG                 group_factorKG -0.0301407218 -0.0377386050
## Age_BL                                 Age_BL -0.0003086131 -0.0006174032
## Sex                                       Sex -0.0072143466 -0.0103655403
## subj.ID@(Intercept)@NA subj.ID@(Intercept)@NA  0.0102987605  0.0000000000
## Residual                             Residual  0.0589888164  0.0440983541
##                                  max
## (Intercept)             3.133646e-01
## tpfu                    4.638475e-03
## group_factorKG         -2.132058e-02
## Age_BL                 -6.008462e-05
## Sex                    -4.407144e-03
## subj.ID@(Intercept)@NA  2.819585e-02
## Residual                5.853199e-02
\end{verbatim}

\includegraphics{draft_rsfmri_files/figure-latex/qa check DMN model age,sex, mFD-1.pdf} \includegraphics{draft_rsfmri_files/figure-latex/qa check DMN model age,sex, mFD-2.pdf}

\begin{verbatim}
## [1] "please carefully evaluate whether the result makes sense, and if not, please contact me"
\end{verbatim}

\begin{verbatim}
## 
## none 
##   46
\end{verbatim}

\begin{verbatim}
##                                          what          orig           min
## (Intercept)                       (Intercept)  0.2882310534  0.2712944647
## tpfu                                     tpfu -0.0054314895 -0.0100474095
## group_factorKG                 group_factorKG -0.0288732597 -0.0371734358
## Age_BL                                 Age_BL -0.0001981959 -0.0004808581
## Sex                                       Sex -0.0062543208 -0.0094973527
## logmFD                                 logmFD -0.0133761940 -0.0311612579
## subj.ID@(Intercept)@NA subj.ID@(Intercept)@NA  0.0144570585  0.0000000000
## Residual                             Residual  0.0585590452  0.0437893877
##                                  max
## (Intercept)             3.043966e-01
## tpfu                    3.891318e-03
## group_factorKG         -1.942998e-02
## Age_BL                  3.457184e-05
## Sex                    -3.092665e-03
## logmFD                  1.596906e-03
## subj.ID@(Intercept)@NA  2.873301e-02
## Residual                5.837731e-02
\end{verbatim}

\includegraphics{draft_rsfmri_files/figure-latex/qa check Rew model age,sex-1.pdf} \includegraphics{draft_rsfmri_files/figure-latex/qa check Rew model age,sex-2.pdf}

\begin{verbatim}
## [1] "please carefully evaluate whether the result makes sense, and if not, please contact me"
\end{verbatim}

\begin{verbatim}
## 
## none 
##   46
\end{verbatim}

\begin{verbatim}
##                                          what          orig           min
## (Intercept)                       (Intercept)  0.1233716391  0.1027031052
## tpfu                                     tpfu -0.0035148229 -0.0070886207
## group_factorKG                 group_factorKG -0.0107355182 -0.0230637231
## Age_BL                                 Age_BL  0.0004245927  0.0001949297
## Sex                                       Sex -0.0058271841 -0.0107320399
## subj.ID@(Intercept)@NA subj.ID@(Intercept)@NA  0.0368043138  0.0285664602
## Residual                             Residual  0.0541471334  0.0428115665
##                                  max
## (Intercept)             0.1359575360
## tpfu                    0.0040817069
## group_factorKG         -0.0048054540
## Age_BL                  0.0008687551
## Sex                    -0.0021879107
## subj.ID@(Intercept)@NA  0.0387259006
## Residual                0.0541921574
\end{verbatim}

\includegraphics{draft_rsfmri_files/figure-latex/qa check Rew model age, sex, mFD-1.pdf} \includegraphics{draft_rsfmri_files/figure-latex/qa check Rew model age, sex, mFD-2.pdf}

\begin{verbatim}
## [1] "please carefully evaluate whether the result makes sense, and if not, please contact me"
\end{verbatim}

\begin{verbatim}
## 
## none 
##   46
\end{verbatim}

\begin{verbatim}
##                                          what          orig           min
## (Intercept)                       (Intercept)  0.0662061153  0.0446574064
## tpfu                                     tpfu -0.0084837422 -0.0108942561
## group_factorKG                 group_factorKG -0.0051835290 -0.0180592427
## Age_BL                                 Age_BL  0.0009275477  0.0005888916
## Sex                                       Sex -0.0014919066 -0.0066504423
## logmFD                                 logmFD -0.0586481745 -0.0700504675
## subj.ID@(Intercept)@NA subj.ID@(Intercept)@NA  0.0387074827  0.0306199614
## Residual                             Residual  0.0522837751  0.0415234010
##                                  max
## (Intercept)             0.0927661360
## tpfu                   -0.0007931101
## group_factorKG          0.0011426393
## Age_BL                  0.0012776894
## Sex                     0.0021175666
## logmFD                 -0.0414794258
## subj.ID@(Intercept)@NA  0.0384413597
## Residual                0.0522319920
\end{verbatim}

\hypertarget{exploratory-analyses}{%
\subsection{Exploratory Analyses}\label{exploratory-analyses}}

\hypertarget{effect-of-mean-and-change-in-bmi}{%
\subsubsection{Effect of mean and change in BMI}\label{effect-of-mean-and-change-in-bmi}}

\hypertarget{whole-brain-analysis}{%
\paragraph{Whole brain analysis}\label{whole-brain-analysis}}

\hypertarget{bmi-only-model}{%
\subparagraph{BMI-only model}\label{bmi-only-model}}

Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate (FWE cluster p\textless0.05). When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \ref{tab:tableBMImodel}.

\begin{figure}
\subfloat[adjusted for age and sex\label{fig:BMIagesexPCCccavgBMIdeact-1}]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIagesex_PCCcc_avgBMI_deact} }\subfloat[adjusted for age, sex and mFD\label{fig:BMIagesexPCCccavgBMIdeact-2}]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIagesexfd_PCCcc_avgBMI_deact} }\caption{Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data}\label{fig:BMIagesexPCCccavgBMIdeact}
\end{figure}

NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
Results with significant cluster co-activation are depicted in Table \ref{tab:tableBMImodel}.

\hypertarget{bmi-and-fd-model}{%
\subparagraph{BMI and FD model}\label{bmi-and-fd-model}}

To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \ref{tab:tableBMIFDmodel}.
Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
As before, the effect did not survive when using GSR-corrected data.

For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.~supplementary motor area (see Table \ref{tab:tableBMIFDmodel}). Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction (p=\(0.028\)), also when evaluating the GSR corrected data (p=\(0.038\)).

\begin{figure}
\subfloat[in AROMA+CC\label{fig:BMIFDagesexNACCcgnBMIdeact-1}]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCcc_cngBMI_deact} }\subfloat[in AROMA+CC+GSR data\label{fig:BMIFDagesexNACCcgnBMIdeact-2}]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact} }\caption{Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD}\label{fig:BMIFDagesexNACCcgnBMIdeact}
\end{figure}

Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.~motor cortex, was significantly positively related to average mFD. This result remained similar when using GSR corrected data.

\begin{figure}
\subfloat[in AROMA+CC\label{fig:BMIFDagesexNACCavgFDact-1}]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCcc_avgFD_act} }\subfloat[in AROMA+CC+GSR data\label{fig:BMIFDagesexNACCavgFDact-2}]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCgsr_avgFD_act} }\caption{Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD}\label{fig:BMIFDagesexNACCavgFDact}
\end{figure}

\hypertarget{aggregated-fc-analysis-1}{%
\paragraph{Aggregated FC analysis}\label{aggregated-fc-analysis-1}}

Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = 0.012), regardless of whether we adjusted for mFD.

The association also remained significant when we split mFD into average and change in mFD (p = NA, 0.0122) and there was no significant association of mean or change in mFD with DMN FC.

\hypertarget{reward-network-mean-and-within-variability-in-bmi-and-logmfd}{%
\paragraph{Reward network, mean and within variability in BMI and logmFD}\label{reward-network-mean-and-within-variability-in-bmi-and-logmfd}}

As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and mFD.

\hypertarget{effect-of-fd-exploratory-analysis}{%
\subsubsection{Effect of FD (exploratory analysis)}\label{effect-of-fd-exploratory-analysis}}

\hypertarget{whole-brain-analysis-1}{%
\paragraph{Whole-brain analysis}\label{whole-brain-analysis-1}}

To investigate the unique contributions of the average value of and the longitudinal change in mFD, we constructed two additional models.
First, BMI measures are excluded from the design matrix. Covariance of the mFD and outcome measures is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor. Age and sex were nuisance covariates.

As shown in Table \ref{tab:tableFDmodel}, there was a significant association between mFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas . Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
We did not find a any clusters for functional connectivity with the PCC.

\begin{figure}
\subfloat[in AROMA+CC\label{fig:FDagesexNACCccavgFDact-1}]{\includegraphics[width=0.5\linewidth]{../report/fig/FDagesex_NACCcc_avgFD_act} }\subfloat[in AROMA+CC+GSR data\label{fig:FDagesexNACCccavgFDact-2}]{\includegraphics[width=0.5\linewidth]{../report/fig/FDagesex_NACCgsr_avgFD_act} }\caption{Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD}\label{fig:FDagesexNACCccavgFDact}
\end{figure}

\hypertarget{effect-of-bmi-and-fd-exploratory-analysis}{%
\subsection{Effect of BMI and FD (exploratory analysis)}\label{effect-of-bmi-and-fd-exploratory-analysis}}

Average BMI and change in BMI remain covariates of interest. While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as. Age and sex remain nuisance covariates.

There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.
Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction. Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds.
On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.

\begin{landscape}\begin{table}

\caption{\label{tab:tableBMImodel}Changes in functional connectivity in whole brain analysis for BMI model}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrrrrrrrll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{3}{c}{clusterwise¹} & \multicolumn{6}{c}{voxelwise  at local maximum} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-11}
\multicolumn{6}{c}{ } & \multicolumn{3}{c}{MNI coordinates} \\
\cmidrule(l{3pt}r{3pt}){7-9}
Seed & Covariates & FWE-corr. P & cluster size & Z score & FWE-corr. P & X & Y & Z & Hem & Anatomical region²\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average BMI (decrease)}}\\
\hspace{1em}PCC (cc) & age, sex & 0.002 & 204 & 1.756 & 0.010 & -6 & -30 & -3 &  & -\\
\hspace{1em}NA & NA & NA & NA & NA & 0.061 & -3 & -45 & 6 &  & -\\
\hspace{1em}NA & NA & NA & NA & NA & 0.065 & 12 & -39 & 0 & R & Lingual Gyrus\\
\hspace{1em}NA & NA & 0.026 & 74 & 1.610 & 0.089 & -51 & -9 & -12 & R & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.369 & -54 & 0 & -21 & L & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.377 & -57 & -3 & -12 & L & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & 0.026 & 75 & 1.612 & 0.226 & 9 & 54 & -6 & L & Superior Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.268 & -3 & 57 & -12 & L & Middle Temporal Gyrus\\
\hspace{1em}NA & NA & 0.036 & 63 & 1.580 & 0.303 & 60 & -12 & -6 & R & Superior Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.643 & 66 & -21 & -9 & R & Middle Temporal Gyrus\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average BMI (decrease)}}\\
\hspace{1em}PCC (cc) & age, sex, log mFD & 0.001 & 233 & 1.765 & 0.011 & -6 & -30 & -3 &  & -\\
\hspace{1em}NA & NA & NA & NA & NA & 0.062 & -3 & -45 & 6 &  & -\\
\hspace{1em}NA & NA & NA & NA & NA & 0.063 & 12 & -39 & 0 & R & Lingual Gyrus\\
\hspace{1em}NA & NA & 0.039 & 73 & 1.603 & 0.074 & -51 & -9 & -12 & L & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.337 & -54 & -3 & -21 & R & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & 0.036 & 80 & 1.619 & 0.206 & 0 & 60 & -12 & L & Superior Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.241 & 9 & 54 & -6 & L & Middle Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.376 & -6 & 51 & -9 & R & Superior Temporal Gyrus\\
\hspace{1em}NA & NA & 0.049 & 57 & 1.556 & 0.279 & 60 & -12 & -6 & R & Superior Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.660 & 66 & -21 & -9 & R & Middle Temporal Gyrus\\
\bottomrule
\multicolumn{11}{l}{\textsuperscript{} Hem, hemisphere; L, left; R, right; FWE-corr., family-wise error corrected; MNI (Montreal Neurological Institute) coordinates of}\\
\multicolumn{11}{l}{primary peak location: X sagittal; Y, coronal; Z, axial; cc, preprocessing with AROMA-ICA + CompCor; gsr, preprocessing with}\\
\multicolumn{11}{l}{AROMA-ICA + CompCor + GSR.}\\
\multicolumn{11}{l}{\textsuperscript{1} To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000}\\
\multicolumn{11}{l}{samples.}\\
\multicolumn{11}{l}{\textsuperscript{2} Connectivity with maximum three voxels that mark local maxima within the respective custer; more detailed description of}\\
\multicolumn{11}{l}{anatomical regions that are assigned to overall clusters and and corresponding probability in Supplementary.}\\
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:tableFDmodel}Changes in functional connectivity in whole brain analysis for FD model}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrrrrrrrll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{3}{c}{clusterwise¹} & \multicolumn{6}{c}{voxelwise  at local maximum} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-11}
\multicolumn{6}{c}{ } & \multicolumn{3}{c}{MNI coordinates} \\
\cmidrule(l{3pt}r{3pt}){7-9}
Seed & Covariates & FWE-corr. P & cluster size & Z score & FWE-corr. P & X & Y & Z & Hem & Anatomical region²\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average log mean FD (increase)}}\\
\hspace{1em}NAcc (cc) & age, sex & 0.008 & 151 & 1.606 & 0.052 & 9 & -15 & 60 & R & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.265 & 12 & -27 & 63 & R & Paracentral Lobule\\
\hspace{1em}NA & NA & NA & NA & NA & 0.360 & -6 & -24 & 51 & L & Paracentral Lobule\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average log mean FD (increase)}}\\
\hspace{1em}NAcc (gsr) & age, sex & 0.017 & 57 & 1.812 & 0.131 & 9 & -15 & 60 & R & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.561 & 9 & -6 & 72 & R & Posterior-Medial Frontal\\
\bottomrule
\multicolumn{11}{l}{\textsuperscript{} Hem, hemisphere; L, left; R, right; FWE-corr., family-wise error corrected; MNI (Montreal Neurological Institute) coordinates of}\\
\multicolumn{11}{l}{primary peak location: X sagittal; Y, coronal; Z, axial; cc, preprocessing with AROMA-ICA + CompCor; gsr, preprocessing with}\\
\multicolumn{11}{l}{AROMA-ICA + CompCor + GSR.}\\
\multicolumn{11}{l}{\textsuperscript{1} To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000}\\
\multicolumn{11}{l}{samples.}\\
\multicolumn{11}{l}{\textsuperscript{2} Connectivity with maximum three voxels that mark local maxima within the respective custer; more detailed description of}\\
\multicolumn{11}{l}{anatomical regions that are assigned to overall clusters and and corresponding probability in Supplementary.}\\
\end{tabular}}
\end{table}
\end{landscape}

\begin{landscape}\begin{table}

\caption{\label{tab:tableBMIFDmodel}Changes in functional connectivity in whole brain analysis for BMI-FD model}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrrrrrrrll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{3}{c}{clusterwise¹} & \multicolumn{6}{c}{voxelwise  at local maximum} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-11}
\multicolumn{6}{c}{ } & \multicolumn{3}{c}{MNI coordinates} \\
\cmidrule(l{3pt}r{3pt}){7-9}
Seed & Covariates & FWE-corr. P & cluster size & Z score & FWE-corr. P & X & Y & Z & Hem & Anatomical region²\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{change in BMI (decrease)}}\\
\hspace{1em}NAcc (cc) & age, sex & 0.028 & 108 & 1.648 & 0.038 & 0 & 3 & 66 & L & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.649 & -3 & 12 & 66 & R & Posterior-Medial Frontal\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average log mean FD (increase)}}\\
\hspace{1em}NAcc (cc) & age, sex & 0.003 & 177 & 1.998 & 0.010 & 9 & -15 & 60 & R & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.076 & 12 & -27 & 63 & R & Paracentral Lobule\\
\hspace{1em}NA & NA & NA & NA & NA & 0.376 & -6 & -24 & 51 & L & Paracentral Lobule\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{change in BMI (decrease)}}\\
\hspace{1em}NAcc (gsr) & age, sex & 0.038 & 92 & 1.665 & 0.134 & 6 & 9 & 63 & R & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.468 & 0 & 3 & 66 & L & Posterior-Medial Frontal\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average log mean FD (increase)}}\\
\hspace{1em}NAcc (gsr) & age, sex & 0.007 & 71 & 1.807 & 0.091 & 9 & -15 & 60 & R & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.356 & 3 & -6 & 51 & R & Posterior-Medial Frontal\\
\hspace{1em}NA & NA & NA & NA & NA & 0.638 & 9 & -6 & 72 & R & Posterior-Medial Frontal\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{average BMI (decrease)}}\\
\hspace{1em}PCC (cc) & age, sex & 0.001 & 246 & 1.794 & 0.011 & -6 & -30 & -3 &  & -\\
\hspace{1em}NA & NA & NA & NA & NA & 0.047 & -3 & -45 & 6 &  & -\\
\hspace{1em}NA & NA & NA & NA & NA & 0.051 & 12 & -39 & 0 & R & Lingual Gyrus\\
\hspace{1em}NA & NA & 0.032 & 72 & 1.617 & 0.063 & -51 & -9 & -12 & R & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.340 & -54 & -3 & -21 & L & Mid Orbital Gyrus\\
\hspace{1em}NA & NA & 0.021 & 90 & 1.657 & 0.165 & 9 & 54 & -6 & L & Superior Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.187 & 0 & 60 & -12 & L & Middle Temporal Gyrus\\
\hspace{1em}NA & NA & NA & NA & NA & 0.294 & -6 & 51 & -9 &  & NA\\
\bottomrule
\multicolumn{11}{l}{\textsuperscript{} Hem, hemisphere; L, left; R, right; FWE-corr., family-wise error corrected; MNI (Montreal Neurological Institute) coordinates of}\\
\multicolumn{11}{l}{primary peak location: X sagittal; Y, coronal; Z, axial; cc, preprocessing with AROMA-ICA + CompCor; gsr, preprocessing with}\\
\multicolumn{11}{l}{AROMA-ICA + CompCor + GSR.}\\
\multicolumn{11}{l}{\textsuperscript{1} To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000}\\
\multicolumn{11}{l}{samples.}\\
\multicolumn{11}{l}{\textsuperscript{2} Connectivity with maximum three voxels that mark local maxima within the respective custer; more detailed description of}\\
\multicolumn{11}{l}{anatomical regions that are assigned to overall clusters and and corresponding probability in Supplementary.}\\
\end{tabular}}
\end{table}
\end{landscape}

The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed. There is a significant association between average BMI and the connectivity between PCC and three other clusters. Results are detailed in Table \ref{tab:tableBMIFDmodel}.
Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
As before, the effect did not survive when thresholding if the images had been GSR corrected.
\textbackslash begin\{figure\}
\subfloat[in AROMA+CC(\#fig:BMIFDagesex_NACC_avgFD_act-1)]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCcc_avgFD_act} }\subfloat[in AROMA+CC+GSR data(\#fig:BMIFDagesex_NACC_avgFD_act-2)]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCgsr_avgFD_act} }

\caption{Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change mFD}

(\#fig:BMIFDagesex\_NACC\_avgFD\_act)
\textbackslash end\{figure\}

\textbackslash begin\{figure\}
\subfloat[in AROMA+CC(\#fig:BMIFDagesex_NACC_cgnBMI_deact-1)]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCcc_cngBMI_deact} }\subfloat[in AROMA+CC+GSR data(\#fig:BMIFDagesex_NACC_cgnBMI_deact-2)]{\includegraphics[width=0.5\linewidth]{../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact} }

\caption{Positive association of change in BMI and FC of NAcc with cluster, adjusted for age, sex, average BMI, average mFD and change mFD}

(\#fig:BMIFDagesex\_NACC\_cgnBMI\_deact)
\textbackslash end\{figure\}

\includegraphics[width=0.75\linewidth]{../report/fig/BMIFDagesex_PCCcc_avgBMI_deact}

\hypertarget{effect-on-hypothalamic-connectivity-exploratory-analysis}{%
\subsection{Effect on hypothalamic connectivity (exploratory analysis)}\label{effect-on-hypothalamic-connectivity-exploratory-analysis}}

\ldots{}

Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{summary}{%
\paragraph{Summary}\label{summary}}

Utilizing resting-state functional MRI on morbidly obese individuals that either underwent or awaited bariatric surgery over the course of one year.
Building on the evidence for post-surgery changes in functional activity, this study explored the longitudinal relationship between surgery-induced weight loss and functional integrity in target networks indicated by resting-state functional connectivity. Over the course of one year we collected neuroimaging and anthropometric data from 33 morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of 15 patients on the waiting list.
Looking at time-by-group interactions, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline. Beyond that, main effects of the intervention and of the time were examined.
The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
To get a better understanding of our results, we conducted multiverse exploratory analyses representing different choices regarding preprocessing pipelines, inclusion of covariates and subsets of the data relevant for interpretation.

\hypertarget{support-of-original-hypotheses}{%
\paragraph{Support of original hypotheses}\label{support-of-original-hypotheses}}

To summarize, while our primary hypothesis on group time interactions could not be confirmed, there was a significant relationship between functional activation
and the continuous BMI measures for some but not all analyses.

\begin{itemize}
\tightlist
\item
  DMN: average BMI but only with lenient AROMA-CC but then for both covariate structures age/sex/mFD and age/sex; effect stable when separating avg mFD and cgn FD into two regressors
\item
  Reward: effect of change BMI only when separating mFD, bit high peak activation
  The individuals BMI averaged across measures taken throughout the one-year period negatively correlated with functional connectivity between PCC and {[}regions from clusters{]}. Results indicate a decrease a functional connectivity within the default mode network.
  \ldots{} not? in accordance with our hypothesis regarding network activity changes.
\end{itemize}

Regarding the reward network, we could not predict changes in functional connectivity with the preregistered models.
However, when investigating the unique contributions to variance of head motion measures and BMI measures, we detected a significant relationship between the individuals' change in BMI throughout the year.
This may be preliminary evidence for changes in the functional connectivity between reward and frontal regions induced by weight-loss.

In our sample, extend of micromovements of the head averaged across all measurements is associated with increased functional connectivity between reward and frontal regions.

Based on concerns expressed in the literature, we tried to closely inspect and rigorously control for the influence of head motion. The presence of multi-collinearity of our predictors has been established before (Beyer et al. 2019). As described by Monti (2011), in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
Surprisingly, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region is only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
However, such unexpected findings would be in accordance with Monti (2011) stating that estimates can behave erratic depending on which regressors are included in the model.

Notwithstanding the difficulties of quantitatively disentangling multicollinearity between head motion and BMI, preliminary conclusions of our analyses may be\ldots{}

Without an a priori power analysis it is not possible to determine if there is no effect or
To clarify, extended considerations should be made to use a priori power analysis (e.g.~with fMRIpower) or check evidence of absence with alternative statistical procedures.

\hypertarget{similarity-of-results-with-literature-random-notes}{%
\paragraph{Similarity of results with literature (random notes)}\label{similarity-of-results-with-literature-random-notes}}

Our results are not consistent with the present literature. Although there are few studies investigating the impact of bariatric surgery on functional connectivity at resting state, the present articles do significant changes in network activation {[}@;@;@{]}.

\hypertarget{interpretation-random-notes}{%
\paragraph{Interpretation (random notes)}\label{interpretation-random-notes}}

However, hypotheses about the plasticity functional networks following bariatric surgery could not be validated. Due to the shortcomings of the inferential paradigm used, our null findings may either put conclusions drawn from previous research into question or may neither provide evidence for the presence nor the absence of the effect.
A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus. After the intervention, functional connectivity was higher in bariatric patients. It decreased after a meal while dietary patients showed an increase (Lepping et al. 2015).
prandrial state is a relevant factor when investigating spontaneous brain activity (Wiemerslage et al. 2016). putamine, insula

\hypertarget{internal-validity}{%
\paragraph{Internal validity}\label{internal-validity}}

Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
Previous studies have compared the sample of bariatric patients to lean individuals, behavioral dieters, or obese individuals without a recommendation for bariatric surgery.
Despite the accumulated evidence, effects may still be attributable to baseline differences, contributing factors to the condition or other psychophysiological and dispositional difference between groups {[}@{]}.
In order to build upon and complement previous evidence, we aimed to isolate the impact of the surgical procedure on changes in resting-state connectivity to a maximum possible extent by conducting a prospective non-randomization trial study.
Such a study design is secondary to randomized control trials in regard to internal control by means of temporality and controllability.
We established temporality by contrasting pre- and post-surgery measures change while controlling for individual variance.
The waiting list control group is equivalent in BMI, treatment history and treatment recommendation and therefore allows controlling for concurrent factors (Thiese 2014).
Incorporating these consideration, our study aimed at a higher internal validity for conclusions about the changes in functional connectivity specific to surgery-induced changes.

\hypertarget{study-sample}{%
\paragraph{Study sample}\label{study-sample}}

Size and composition of our sample did not allow separate analyses for the sexes. However, the disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men (Fuchs et al. 2015; Chooi, Ding, and Magkos 2019).

\hypertarget{marginal-model}{%
\paragraph{Marginal model}\label{marginal-model}}

Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), it is often recommended to use modern models such as mixed-effect models and marginal models (Chen et al. 2013; Guillaume et al. 2014; McFarquhar 2019).
McFarquhar (2019) demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
Opposed to flexible factorial models, where each individual is modelled separately, marginal models are parsimonious and preserve more degrees of freedom.

\hypertarget{code-availability}{%
\section{Code availability}\label{code-availability}}

All code is available in the github repository \url{https://github.com/hsx1/adi2_rsfmri}. The code includes script for data preparation and analysis of different models.

\hypertarget{declaration-of-interest}{%
\section{Declaration of interest}\label{declaration-of-interest}}

\hypertarget{funding}{%
\section{Funding}\label{funding}}

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-spm12}{}%
Ashburner, John, Gareth Barnes, C Chen, Jean Daunizeau, Guillaume Flandin, Karl Friston, Stefan Kiebel, et al. 2014. ``SPM12 Manual.'' \emph{Wellcome Trust Centre for Neuroimaging, London, UK}, 2464.

\leavevmode\hypertarget{ref-avants2009ants}{}%
Avants, Brian B, Nicholas J Tustison, Gang Song, and James C Gee. 2009. ``Ants: Open-Source Tools for Normalization and Neuroanatomy.'' \emph{HeanetIe} 10: 1--11.

\leavevmode\hypertarget{ref-Bartra_2013}{}%
Bartra, Oscar, Joseph T McGuire, and Joseph W Kable. 2013. ``The Valuation System: A Coordinate-Based Meta-Analysis of Bold fMRI Experiments Examining Neural Correlates of Subjective Value.'' \emph{Neuroimage} 76 (August): 412--27. \url{https://doi.org/10.1016/j.neuroimage.2013.02.063}.

\leavevmode\hypertarget{ref-Bates_2007}{}%
Bates, Douglas, Deepayan Sarkar, Maintainer Douglas Bates, and L Matrix. 2007. ``The Lme4 Package.'' \emph{R Package Version} 2 (1): 74.

\leavevmode\hypertarget{ref-Behzadi_2007}{}%
Behzadi, Yashar, Khaled Restom, Joy Liau, and Thomas T Liu. 2007. ``A Component Based Noise Correction Method (Compcor) for Bold and Perfusion Based fMRI.'' \emph{Neuroimage} 37 (1): 90--101.

\leavevmode\hypertarget{ref-Berthoud_2011}{}%
Berthoud, Andrew C. Shin, and Huiyuan Zheng". n.d. ``"Obesity Surgery and Gut--Brain Communication".'' \emph{"Physiology \& Behavior"} "105" ("1"): "106--119". \url{https://doi.org/\%20\%22https://doi.org/10.1016/j.physbeh.2011.01.023\%22}.

\leavevmode\hypertarget{ref-Beyer_2019}{}%
Beyer, Frauke, Isabel García-García, Matthias Heinrich, Matthias L. Schroeter, Julia Sacher, Tobias Luck, Steffi G. Riedel-Heller, Michael Stumvoll, Arno Villringer, and A. Veronica Witte. 2019. ``Neuroanatomical Correlates of Food Addiction Symptoms and Body Mass Index in the General Population.'' \emph{Human Brain Mapping} 40 (9): 2747--58. \url{https://doi.org/10.1002/hbm.24557}.

\leavevmode\hypertarget{ref-Beyer_2017}{}%
Beyer, Frauke, Sharzhad Kharabian Masouleh, Julia M Huntenburg, Leonie Lampe, Tobias Luck, Steffi G Riedel-Heller, Markus Loeffler, et al. 2017. ``Higher Body Mass Index Is Associated with Reduced Posterior Default Mode Connectivity in Older Adults.'' \emph{Human Brain Mapping} 38 (7): 3502--15. \url{https://doi.org/10.1002/hbm.23605}.

\leavevmode\hypertarget{ref-Beyer_2020}{}%
Beyer, Frauke, Kristin Prehn, Katharina A. Wüsten, Arno Villringer, Jürgen Ordemann, Agnes Flöel, and A. Veronica Witte. 2020. ``Weight Loss Reduces Head Motion: Revisiting a Major Confound in Neuroimaging.'' \emph{Human Brain Mapping} 41 (9): 2490--4. \url{https://doi.org/10.1002/hbm.24959}.

\leavevmode\hypertarget{ref-Borowitz_2020}{}%
Borowitz, Michelle A., Sonja Yokum, Elizabeth R. Duval, and Ashley N. Gearhardt. 2020. ``Weight-Related Differences in Salience, Default~Mode, and Executive~Function Network Connectivity in Adolescents.'' \emph{Obesity} 28 (8): 1438--46. \url{https://doi.org/https://doi.org/10.1002/oby.22853}.

\leavevmode\hypertarget{ref-Brutman_2019}{}%
Brutman, Julianna N, Sunil Sirohi, and Jon F Davis. 2019. ``Recent Advances in the Neurobiology of Altered Motivation Following Bariatric Surgery.'' \emph{Current Psychiatry Reports} 21 (11): 1--10.

\leavevmode\hypertarget{ref-Chao_2018}{}%
Chao, Seh-Huang, Yin-To Liao, Vincent Chin-Hung Chen, Cheng-Jui Li, Roger S. McIntyre, Yena Lee, and Jun-Cheng Weng. 2018. ``Correlation Between Brain Circuit Segregation and Obesity.'' \emph{Behavioural Brain Research} 337: 218--27. \url{https://doi.org/https://doi.org/10.1016/j.bbr.2017.09.017}.

\leavevmode\hypertarget{ref-Chen_2013}{}%
Chen, Gang, Ziad S. Saad, Jennifer C. Britton, Daniel S. Pine, and Robert W. Cox. 2013. ``Linear Mixed-Effects Modeling Approach to FMRI Group Analysis.'' \emph{NeuroImage} 73 (June): 176--90. \url{https://doi.org/10.1016/j.neuroimage.2013.01.047}.

\leavevmode\hypertarget{ref-Chooi_2019}{}%
Chooi, Yu Chung, Cherlyn Ding, and Faidon Magkos. 2019. ``The Epidemiology of Obesity.'' \emph{Metabolism} 92 (March): 6--10. \url{https://doi.org/10.1016/j.metabol.2018.09.005}.

\leavevmode\hypertarget{ref-Ciric_2018}{}%
Ciric, Rastko, Adon FG Rosen, Guray Erus, Matthew Cieslak, Azeez Adebimpe, Philip A Cook, Danielle S Bassett, Christos Davatzikos, Daniel H Wolf, and Theodore D Satterthwaite. 2018. ``Mitigating Head Motion Artifact in Functional Connectivity Mri.'' \emph{Nature Protocols} 13 (12): 2801--26.

\leavevmode\hypertarget{ref-Ciric_2017}{}%
Ciric, Rastko, Daniel H Wolf, Jonathan D Power, David R Roalf, Graham L Baum, Kosha Ruparel, Russell T Shinohara, et al. 2017. ``Benchmarking of Participant-Level Confound Regression Strategies for the Control of Motion Artifact in Studies of Functional Connectivity.'' \emph{NeuroImage} 154 (July): 174--87. \url{https://doi.org/10.1016/j.neuroimage.2017.03.020}.

\leavevmode\hypertarget{ref-Contreras_2017}{}%
Contreras-Rodríguez, Oren, Cristina Martín-Pérez, Raquel Vilar-López, and Antonio Verdejo-Garcia. 2017. ``Ventral and Dorsal Striatum Networks in Obesity: Link to Food Craving and Weight Gain.'' \emph{Biological Psychiatry} 81 (9): 789--96. \url{https://doi.org/https://doi.org/10.1016/j.biopsych.2015.11.020}.

\leavevmode\hypertarget{ref-Coveleskie_2015}{}%
Coveleskie, K, A Gupta, LA Kilpatrick, ED Mayer, C Ashe-McNalley, J Stains, JS Labus, and EA Mayer. 2015. ``Altered Functional Connectivity Within the Central Reward Network in Overweight and Obese Women.'' \emph{Nutrition \& Diabetes} 5 (1): e148--e148.

\leavevmode\hypertarget{ref-Devoto_2018}{}%
Devoto, F., L. Zapparoli, R. Bonandrini, M. Berlingeri, A. Ferrulli, L. Luzi, G. Banfi, and E. Paulesu. 2018. ``Hungry Brains: A Meta-Analytical Review of Brain Activation Imaging Studies on Food Perception and Appetite in Obese Individuals.'' \emph{Neuroscience \& Biobehavioral Reviews} 94: 271--85. \url{https://doi.org/https://doi.org/10.1016/j.neubiorev.2018.07.017}.

\leavevmode\hypertarget{ref-Ding_2020}{}%
Ding, Yueyan, Gang Ji, Guanya Li, Wenchao Zhang, Yang Hu, Lei Liu, Yuanyuan Wang, et al. 2020. ``Altered Interactions Among Resting-State Networks in Individuals with Obesity.'' \emph{Obesity} 28 (3): 601--8. \url{https://doi.org/10.1002/oby.22731}.

\leavevmode\hypertarget{ref-Doucet_2017}{}%
Doucet, Gaelle E, Natalie Rasgon, Bruce S McEwen, Nadia Micali, and Sophia Frangou. 2017. ``Elevated Body Mass Index is Associated with Increased Integration and Reduced Cohesion of Sensory-Driven and Internally Guided Resting-State Functional Brain Networks.'' \emph{Cerebral Cortex} 28 (3): 988--97. \url{https://doi.org/10.1093/cercor/bhx008}.

\leavevmode\hypertarget{ref-Duan_2020}{}%
Duan, Shijun, Gang Ji, Guanya Li, Yang Hu, Wenchao Zhang, Jia Wang, Dardo Tomasi, et al. 2020. ``Bariatric Surgery Induces Alterations in Effective Connectivity Between the Orbitofrontal Cortex and Limbic Regions in Obese Patients.'' \emph{Science China Information Sciences} 63 (7). \url{https://doi.org/10.1007/s11432-019-2817-x}.

\leavevmode\hypertarget{ref-Eickhoff_2005}{}%
Eickhoff, Simon B., Klaas E. Stephan, Hartmut Mohlberg, Christian Grefkes, Gereon R. Fink, Katrin Amunts, and Karl Zilles. 2005. ``A New SPM Toolbox for Combining Probabilistic Cytoarchitectonic Maps and Functional Imaging Data.'' \emph{NeuroImage} 25 (4): 1325--35. \url{https://doi.org/10.1016/j.neuroimage.2004.12.034}.

\leavevmode\hypertarget{ref-fs_cross}{}%
Fischl, Bruce. 2012. ``FreeSurfer.'' \emph{NeuroImage} 62 (2): 774--81. \url{https://doi.org/https://doi.org/10.1016/j.neuroimage.2012.01.021}.

\leavevmode\hypertarget{ref-Frank_2013}{}%
Frank, S, B Wilms, R Veit, B Ernst, M Thurnheer, S Kullmann, A Fritsche, N Birbaumer, H Preissl, and B Schultes. 2013. ``Altered Brain Activity in Severely Obese Women May Recover After Roux-En Y Gastric Bypass Surgery.'' \emph{International Journal of Obesity} 38 (3): 341--48. \url{https://doi.org/10.1038/ijo.2013.60}.

\leavevmode\hypertarget{ref-Fuchs_2015}{}%
Fuchs, Hans F., Ryan C. Broderick, Cristina R. Harnsberger, David C. Chang, Bryan J. Sandler, Garth R. Jacobsen, and Santiago Horgan. 2015. ``Benefits of Bariatric Surgery Do Not Reach Obese Men.'' \emph{Journal of Laparoendoscopic \& Advanced Surgical Techniques} 25 (3): 196--201. \url{https://doi.org/10.1089/lap.2014.0639}.

\leavevmode\hypertarget{ref-Garcia_2014}{}%
García-García, I., A. Horstmann, M. A. Jurado, M. Garolera, S. J. Chaudhry, D. S. Margulies, A. Villringer, and J. Neumann. 2014. ``Reward Processing in Obesity, Substance Addiction and Non-Substance Addiction.'' \emph{Obesity Reviews} 15 (11): 853--69. \url{https://doi.org/https://doi.org/10.1111/obr.12221}.

\leavevmode\hypertarget{ref-George_2016}{}%
George, Brandon J., T. Mark Beasley, Andrew W. Brown, John Dawson, Rositsa Dimova, Jasmin Divers, TaShauna U. Goldsby, et al. 2016. ``Common Scientific and Statistical Errors in Obesity Research.'' \emph{Obesity} 24 (4): 781--90. \url{https://doi.org/10.1002/oby.21449}.

\leavevmode\hypertarget{ref-Guillaume_2014}{}%
Guillaume, Bryan, Xue Hua, Paul Thomson, Lourens Waldorp, and Thomas Nichols. 2014. ``The Swe Toolbox: A Toolbox for the Analysis of Longitudinal and Repeated Measures Neuroimaging Data.''

\leavevmode\hypertarget{ref-Hare_2011}{}%
Hare, Todd A, Jonathan Malmaud, and Antonio Rangel. 2011. ``Focusing Attention on the Health Aspects of Foods Changes Value Signals in vmPFC and Improves Dietary Choice.'' \emph{Journal of Neuroscience} 31 (30): 11077--87. \url{https://doi.org/10.1523/JNEUROSCI.6383-10.2011}.

\leavevmode\hypertarget{ref-Hogenkamp_2016}{}%
Hogenkamp, P S, W Zhou, L S Dahlberg, J Stark, A L Larsen, G Olivo, L Wiemerslage, et al. 2016. ``Higher Resting-State Activity in Reward-Related Brain Circuits in Obese Versus Normal-Weight Females Independent of Food Intake.'' \emph{International Journal of Obesity} 40 (11): 1687--92. \url{https://doi.org/10.1038/ijo.2016.105}.

\leavevmode\hypertarget{ref-Hutcherson_2012}{}%
Hutcherson, Cendri A, Hilke Plassmann, James J Gross, and Antonio Rangel. 2012. ``Cognitive Regulation During Decision Making Shifts Behavioral Control Between Ventromedial and Dorsolateral Prefrontal Value Systems.'' \emph{Journal of Neuroscience} 32 (39): 13543--54. \url{https://doi.org/10.1523/JNEUROSCI.6387-11.2012}.

\leavevmode\hypertarget{ref-Karra_2013}{}%
Karra, Efthimia, Owen G O'Daly, Agharul I Choudhury, Ahmed Yousseif, Steven Millership, Marianne T Neary, William R Scott, et al. 2013. ``A Link Between Fto, Ghrelin, and Impaired Brain Food-Cue Responsivity.'' \emph{The Journal of Clinical Investigation} 123 (8): 3539--51.

\leavevmode\hypertarget{ref-Klapwijk_2019}{}%
Klapwijk, Eduard T., Ferdi van de Kamp, Mara van der Meulen, Sabine Peters, and Lara M. Wierenga. 2019. ``Qoala-T: A Supervised-Learning Tool for Quality Control of FreeSurfer Segmented MRI Data.'' \emph{NeuroImage} 189 (April): 116--29. \url{https://doi.org/10.1016/j.neuroimage.2019.01.014}.

\leavevmode\hypertarget{ref-Kullmann_2011}{}%
Kullmann, Stephanie, Martin Heni, Ralf Veit, Caroline Ketterer, Fritz Schick, Hans-Ulrich Häring, Andreas Fritsche, and Hubert Preissl. 2011. ``The Obese Brain: Association of Body Mass Index and Insulin Sensitivity with Resting State Network Functional Connectivity.'' \emph{Human Brain Mapping} 33 (5): 1052--61. \url{https://doi.org/10.1002/hbm.21268}.

\leavevmode\hypertarget{ref-Lamarca_2019}{}%
Lamarca, Fernando, Mariana Silva Melendez-Araújo, Isabela Porto de Toledo, Eliane Said Dutra, and Kênia Mara Baiocchi de Carvalho. 2019. ``Relative Energy Expenditure Decreases During the First Year After Bariatric Surgery: A Systematic Review and Meta-Analysis.'' \emph{Obesity Surgery}, 1--12.

\leavevmode\hypertarget{ref-Lepping_2015}{}%
Lepping, Rebecca J., Amanda S. Bruce, Alex Francisco, Hung-Wen Yeh, Laura E. Martin, Joshua N. Powell, Laura Hancock, et al. 2015. ``Resting-State Brain Connectivity After Surgical and Behavioral Weight Loss.'' \emph{Obesity} 23 (7): 1422--8. \url{https://doi.org/10.1002/oby.21119}.

\leavevmode\hypertarget{ref-Li_2019}{}%
Li, Guanya, Gang Ji, Yang Hu, Li Liu, Qingchao Jin, Wenchao Zhang, Lei Liu, et al. 2019. ``Reduced Plasma Ghrelin Concentrations Are Associated with Decreased Brain Reactivity to Food Cues After Laparoscopic Sleeve Gastrectomy.'' \emph{Psychoneuroendocrinology} 100: 229--36. \url{https://doi.org/https://doi.org/10.1016/j.psyneuen.2018.10.022}.

\leavevmode\hypertarget{ref-Li_2018}{}%
Li, Guanya, Gang Ji, Yang Hu, Mingzhu Xu, Qingchao Jin, Li Liu, Karen M. Deneen, et al. 2018. ``Bariatric Surgery in Obese Patients Reduced Resting Connectivity of Brain Regions Involved with Self-Referential Processing.'' \emph{Human Brain Mapping} 39 (12): 4755--65. \url{https://doi.org/10.1002/hbm.24320}.

\leavevmode\hypertarget{ref-Lips_2014}{}%
Lips, Mirjam A, Marjolein A Wijngaarden, Jeroen van der Grond, Mark A van Buchem, Gerrit H de Groot, Serge ARB Rombouts, Hanno Pijl, and Ilya M Veer. 2014. ``Resting-State Functional Connectivity of Brain Regions Involved in Cognitive Control, Motivation, and Reward Is Enhanced in Obese Females.'' \emph{The American Journal of Clinical Nutrition} 100 (2): 524--31. \url{https://doi.org/10.3945/ajcn.113.080671}.

\leavevmode\hypertarget{ref-Liu_2011}{}%
Liu, Xun, Jacqueline Hairston, Madeleine Schrier, and Jin Fan. 2011. ``Common and Distinct Networks Underlying Reward Valence and Processing Stages: A Meta-Analysis of Functional Neuroimaging Studies.'' \emph{Neuroscience \& Biobehavioral Reviews} 35 (5): 1219--36. \url{https://doi.org/https://doi.org/10.1016/j.neubiorev.2010.12.012}.

\leavevmode\hypertarget{ref-MATLAB2018}{}%
MATLAB. 2018. \emph{9.7.0.1190202 (R2019b)}. Natick, Massachusetts: The MathWorks Inc.

\leavevmode\hypertarget{ref-McFadden_2013}{}%
McFadden, Kristina L, Marc-Andre Cornier, Edward L Melanson, Jamie L Bechtell, and Jason R Tregellas. 2013. ``Effects of Exercise on Resting-State Default Mode and Salience Network Activity in Overweight/Obese Adults.'' \emph{Neuroreport} 24 (15): 866.

\leavevmode\hypertarget{ref-McFarquhar_2019}{}%
McFarquhar, Martyn. 2019. ``Modeling Group-Level Repeated Measurements of Neuroimaging Data Using the Univariate General Linear Model.'' \emph{Frontiers in Neuroscience} 13 (April). \url{https://doi.org/10.3389/fnins.2019.00352}.

\leavevmode\hypertarget{ref-Mechanick_2013}{}%
Mechanick, Jeffrey I., Adrienne Youdim, Daniel B. Jones, W. Timothy Garvey, Daniel L. Hurley, M. Molly McMahon, Leslie J. Heinberg, et al. 2013. ``Clinical Practice Guidelines for the Perioperative Nutritional, Metabolic, and Nonsurgical Support of the Bariatric Surgery Patient2013 Update: Cosponsored by American Association of Clinical Endocrinologists, the Obesity Society, and American Society for Metabolic \& Bariatric Surgery.'' \emph{Surgery for Obesity and Related Diseases} 9 (2): 159--91. \url{https://doi.org/10.1016/j.soard.2012.12.010}.

\leavevmode\hypertarget{ref-Meng_2020}{}%
Meng, Xia, Duo Huang, Hua Ao, Xinyu Wang, and Xiao Gao. 2020. ``Food Cue Recruits Increased Reward Processing and Decreased Inhibitory Control Processing in the Obese/Overweight: An Activation Likelihood Estimation Meta-Analysis of fMRI Studies.'' \emph{Obesity Research \& Clinical Practice} 14 (2): 127--35. \url{https://doi.org/https://doi.org/10.1016/j.orcp.2020.02.004}.

\leavevmode\hypertarget{ref-Monti_2011}{}%
Monti, Martin. 2011. ``Statistical Analysis of fMRI Time-Series: A Critical Review of the GLM Approach.'' \emph{Frontiers in Human Neuroscience} 5. \url{https://doi.org/10.3389/fnhum.2011.00028}.

\leavevmode\hypertarget{ref-Morys_2020}{}%
Morys, Filip, Isabel García-García, and Alain Dagher. 2020. ``Is obesity related to enhanced neural reactivity to visual food cues? A review and meta-analysis.'' \emph{Social Cognitive and Affective Neuroscience}, August. \url{https://doi.org/10.1093/scan/nsaa113}.

\leavevmode\hypertarget{ref-Mulla_2017}{}%
Mulla, Christopher M., Roeland J. W. Middelbeek, and Mary-Elizabeth Patti. 2017. ``Mechanisms of Weight Loss and Improved Metabolism Following Bariatric Surgery.'' \emph{Annals of the New York Academy of Sciences} 1411 (1): 53--64. \url{https://doi.org/10.1111/nyas.13409}.

\leavevmode\hypertarget{ref-Muschelli_2014}{}%
Muschelli, John, Mary Beth Nebel, Brian S. Caffo, Anita D. Barber, James J. Pekar, and Stewart H. Mostofsky. 2014. ``Reduction of Motion-Related Artifacts in Resting State fMRI Using aCompCor.'' \emph{NeuroImage} 96: 22--35. \url{https://doi.org/https://doi.org/10.1016/j.neuroimage.2014.03.028}.

\leavevmode\hypertarget{ref-Odoherty_2004}{}%
O'Doherty, John P. 2004. ``Reward Representations and Reward-Related Learning in the Human Brain: Insights from Neuroimaging.'' \emph{Current Opinion in Neurobiology} 14 (6): 769--76. \url{https://doi.org/https://doi.org/10.1016/j.conb.2004.10.016}.

\leavevmode\hypertarget{ref-Olivo_2017}{}%
Olivo, Gaia, Wei Zhou, Magnus Sundbom, Christina Zhukovsky, Pleunie Hogenkamp, Lamia Nikontovic, Julia Stark, et al. 2017. ``Resting-State Brain Connectivity Changes in Obese Women After Roux-En-Y Gastric Bypass Surgery: A Longitudinal Study.'' \emph{Scientific Reports} 7 (1). \url{https://doi.org/10.1038/s41598-017-06663-5}.

\leavevmode\hypertarget{ref-Parkes_2018}{}%
Parkes, Linden, Ben Fulcher, Murat Yücel, and Alex Fornito. 2018. ``An Evaluation of the Efficacy, Reliability, and Sensitivity of Motion Correction Strategies for Resting-State Functional MRI.'' \emph{NeuroImage} 171 (May): 415--36. \url{https://doi.org/10.1016/j.neuroimage.2017.12.073}.

\leavevmode\hypertarget{ref-Power_2012}{}%
Power, Jonathan D., Kelly A. Barnes, Abraham Z. Snyder, Bradley L. Schlaggar, and Steven E. Petersen. 2012. ``Spurious but Systematic Correlations in Functional Connectivity MRI Networks Arise from Subject Motion.'' \emph{NeuroImage} 59 (3): 2142--54. \url{https://doi.org/10.1016/j.neuroimage.2011.10.018}.

\leavevmode\hypertarget{ref-Power_2015}{}%
Power, Jonathan D, Bradley L Schlaggar, and Steven E Petersen. 2015. ``Recent Progress and Outstanding Issues in Motion Correction in Resting State fMRI.'' \emph{Neuroimage} 105: 536--51.

\leavevmode\hypertarget{ref-Prehn_2020}{}%
Prehn, Kristin, Thorge Profitlich, Ida Rangus, Sebastian Heßler, A. Veronica Witte, Ulrike Grittner, Jürgen Ordemann, and Agnes Flöel. 2020. ``Bariatric Surgery and Brain HealthA Longitudinal Observational Study Investigating the Effect of Surgery on Cognitive Function and Gray Matter Volume.'' \emph{Nutrients} 12 (1): 127. \url{https://doi.org/10.3390/nu12010127}.

\leavevmode\hypertarget{ref-pruim2015ica}{}%
Pruim, Raimon HR, Maarten Mennes, Daan van Rooij, Alberto Llera, Jan K Buitelaar, and Christian F Beckmann. 2015. ``ICA-Aroma: A Robust Ica-Based Strategy for Removing Motion Artifacts from fMRI Data.'' \emph{Neuroimage} 112: 267--77.

\leavevmode\hypertarget{ref-Raichle_2015}{}%
Raichle, Marcus E. 2015. ``The Brains Default Mode Network.'' \emph{Annual Review of Neuroscience} 38 (1): 433--47. \url{https://doi.org/10.1146/annurev-neuro-071013-014030}.

\leavevmode\hypertarget{ref-fs_long}{}%
Reuter, Martin, and Bruce Fischl. 2011. ``Avoiding Asymmetry-Induced Bias in Longitudinal Image Processing.'' \emph{Neuroimage} 57 (1): 19--21.

\leavevmode\hypertarget{ref-Sadler_2018}{}%
Sadler, Jennifer R., Grace E. Shearrer, and Kyle S. Burger. 2018. ``Body Mass Variability Is Represented by Distinct Functional Connectivity Patterns.'' \emph{NeuroImage} 181: 55--63. \url{https://doi.org/https://doi.org/10.1016/j.neuroimage.2018.06.082}.

\leavevmode\hypertarget{ref-Schmidt_2018}{}%
Schmidt, Liane, Anita Tusche, Nicolas Manoharan, Cendri Hutcherson, Todd Hare, and Hilke Plassmann. 2018. ``Neuroanatomy of the vmPFC and dlPFC Predicts Individual Differences in Cognitive Regulation During Dietary Self-Control Across Regulation Strategies.'' \emph{The Journal of Neuroscience} 38 (25): 5799--5806. \url{https://doi.org/10.1523/JNEUROSCI.3402-17.2018}.

\leavevmode\hypertarget{ref-Shai_2008}{}%
Shai, Iris, Dan Schwarzfuchs, Yaakov Henkin, Danit R. Shahar, Shula Witkow, Ilana Greenberg, Rachel Golan, et al. 2008. ``Weight Loss with a Low-Carbohydrate, Mediterranean, or Low-Fat Diet.'' \emph{New England Journal of Medicine} 359 (3): 229--41. \url{https://doi.org/10.1056/NEJMoa0708681}.

\leavevmode\hypertarget{ref-Smith_2009}{}%
Smith, Stephen M, Peter T Fox, Karla L Miller, David C Glahn, P Mickle Fox, Clare E Mackay, Nicola Filippini, et al. 2009. ``Correspondence of the Brain's Functional Architecture During Activation and Rest.'' \emph{Proceedings of the National Academy of Sciences} 106 (31): 13040--5.

\leavevmode\hypertarget{ref-Stice_2015}{}%
Stice, Eric, Kyle S Burger, and Sonja Yokum. 2015. ``Reward Region Responsivity Predicts Future Weight Gain and Moderating Effects of the Taqia Allele.'' \emph{Journal of Neuroscience} 35 (28): 10316--24.

\leavevmode\hypertarget{ref-Stoeckel_2009}{}%
Stoeckel, Luke E., Jieun Kim, Rosalyn E. Weller, James E. Cox, Edwin W. Cook, and Barry Horwitz. 2009. ``Effective Connectivity of a Reward Network in Obese Women.'' \emph{Brain Research Bulletin} 79 (6): 388--95. \url{https://doi.org/https://doi.org/10.1016/j.brainresbull.2009.05.016}.

\leavevmode\hypertarget{ref-team2013r}{}%
Team, R Core, and others. 2013. ``R: A Language and Environment for Statistical Computing.'' Vienna, Austria.

\leavevmode\hypertarget{ref-Thiese_2014}{}%
Thiese, Matthew S. 2014. ``Observational and Interventional Study Design Types; an Overview.'' \emph{Biochemia Medica} 24 (2): 199--210. \url{https://doi.org/10.11613/BM.2014.022}.

\leavevmode\hypertarget{ref-Tregellas_2011}{}%
Tregellas, Jason R., Korey P. Wylie, Donald C. Rojas, Jody Tanabe, Jesse Martin, Eugene Kronberg, Dietmar Cordes, and Marc-Andre Cornier. 2011. ``Altered Default Network Activity in Obesity.'' \emph{Obesity} 19 (12): 2316--21. \url{https://doi.org/10.1038/oby.2011.119}.

\leavevmode\hypertarget{ref-Wiemerslage_2016}{}%
Wiemerslage, Lyle, Wei Zhou, Gaia Olivo, Julia Stark, Pleunie S. Hogenkamp, Elna-Marie Larsson, Magnus Sundbom, and Helgi B. Schiöth. 2016. ``A Resting-State fMRI Study of Obese Females Between Pre- and Postprandial States Before and After Bariatric Surgery.'' Edited by John Foxe. \emph{European Journal of Neuroscience} 45 (3): 333--41. \url{https://doi.org/10.1111/ejn.13428}.

\leavevmode\hypertarget{ref-Wijngaarden_2015}{}%
Wijngaarden, M. A., I. M. Veer, S. A. R. B. Rombouts, M. A. van Buchem, K. Willems van Dijk, H. Pijl, and J. van der Grond. 2015. ``Obesity Is Marked by Distinct Functional Connectivity in Brain Networks Involved in Food Reward and Salience.'' \emph{Behavioural Brain Research} 287: 127--34. \url{https://doi.org/https://doi.org/10.1016/j.bbr.2015.03.016}.

\leavevmode\hypertarget{ref-Yarmush_2017}{}%
Yarmush, Martin L, Matthew D'Alessandro, and Nima Saeidi. 2017. ``Regulation of Energy Homeostasis After Gastric Bypass Surgery.'' \emph{Annual Review of Biomedical Engineering} 19: 459--84.

\leavevmode\hypertarget{ref-Zoon_2018}{}%
Zoon, Harriët FA, Suzanne EM de Bruijn, Paul AM Smeets, Cees de Graaf, Ignace Janssen, Wendy Schijns, Edo O Aarts, Gerry Jager, and Sanne Boesveldt. 2018. ``Altered Neural Responsivity to Food Cues in Relation to Food Preferences, but Not Appetite-Related Hormone Concentrations After Rygb-Surgery.'' \emph{Behavioural Brain Research} 353: 194--202.

\end{document}

### 2021-02-02 code/data_load.Rmd

@@
 ---
-title: "data_proprocessing"
-author: "Hannah Sophie Heinrichs"
-date: "1/20/2021"
 output:
@@
 library(ggplot2)
-library(haven)
 library(plyr)
@@
 
+```{r All info files}
+# Messueberblick
+Messueberblick_path <- "/data/p_02161/ADI_studie/metadata/ADI_MRT_Messueberblick_2019_04_10.csv"
+
+# ADI_Aktuell  -  main table, BMI info is refilled with Messueberblick entries
+ADIAktuell_üath <- "/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav"
+
+# QOL_sheet  -  correct group information (unused, just for comparison)
+QOLsheet_path <- "/data/pt_02161/Results/Project1_headmotion/QOL_sheet_of_ADI_Messueberblick.csv"
+
+# summary_with_SPSS  - table has errors (contradictions for ADI080 and ADI036 with other table, but contains missing subjects)
+GroupCodingWitErrors_path <- "/data/pt_02161/Results/Project1_headmotion/summary_with_SPSS_data_sheet_of_ADI_Messueberblick.csv"
+
+# rs QA infos
+rsQA_path <- "/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv"
+
+# FS infos
+FSqa_path <- "/data/p_02161/ADI_studie/BIDS/derivatives/freesurfer/QA/freesurfer_qa_qoala_11.7.csv"
+
+# subjects_rs  -  Loading mean/max FD values
+rsID_path <- "/data/pt_02161/Results/Project1_headmotion/subjects_rs.csv"
+FDvals_path <- "/data/pt_02161/Results/Project1_headmotion/summary_FDvals_correctADI061.csv"
+
+# output
+out_path <- "/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv"
+```
 ### Loading Demographics
@@
 ```{r "Load data", include=FALSE}
+# read "Messüberblick"
+# collect subject IDs in one column "BMI"
 bmi = read.csv(
@@
   stringsAsFactors = FALSE
-)
-bmi_tmp = bmi[, c("X.1",
+)[, c("X.1",
       "BMI",
@@
 
-# collect subject IDs in one column
-bmi_tmp$subj.ID = bmi_tmp$X.1  
-bmi_tmp$subj.ID[!is.na(bmi_tmp$X.2)] = bmi_tmp$X.2[!is.na(bmi_tmp$X.2)]  
-bmi_tmp$subj.ID[!is.na(bmi_tmp$X.3)] = bmi_tmp$X.3[!is.na(bmi_tmp$X.3)]
-
-bmi_tmp = bmi_tmp[, c("subj.ID", "BMI", "BMI.1", "BMI.2", "Geschlecht", "Alter")]
+bmi$subj.ID = bmi$X.1  
+bmi$subj.ID[!is.na(bmi$X.2)] = bmi$X.2[!is.na(bmi$X.2)]  
+bmi$subj.ID[!is.na(bmi$X.3)] = bmi$X.3[!is.na(bmi$X.3)]  
+bmi = bmi[, c("subj.ID", "BMI", "BMI.1", "BMI.2", "Geschlecht", "Alter")]
 
 # get BMI data from another ADI table which contains more subjects
-adi_data = read_sav("/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav")
 # select variables: only BMI
-bmi_repl = adi_data[, c(
+bmi_repl = haven::read_sav("/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav")[, c(
   "ADI_Code",
@@
 
-# replace some missing BMI in the "Messueberblick" table with values from ADI_aktuell.sav
-abmi = merge(bmi_tmp, bmi_repl, by = "subj.ID", all = TRUE)
-abmi$BMI_BL[is.na(abmi$BMI_BL)] = abmi$BMI[is.na(abmi$BMI_BL)]
-abmi$BMI_6m[is.na(abmi$BMI_6m)] = abmi$BMI.1[is.na(abmi$BMI_6m)]
-abmi$BMI_12m[is.na(abmi$BMI_12m)] = abmi$BMI.2[is.na(abmi$BMI_12m)]
-abmi$Sex[is.na(abmi$Sex)] = abmi$Geschlecht[is.na(abmi$Sex)]
-abmi$Age_BL[is.na(abmi$Age_BL)] = abmi$Alter[is.na(abmi$Age_BL)]
+# replace some missing BMI in the "ADI_aktuell.sav" (bmi_repl) with values from "Messueberblick" (bmi)
+merged_bmi = merge(bmi, bmi_repl, by = "subj.ID", all = TRUE)
+merged_bmi$BMI_BL[is.na(merged_bmi$BMI_BL)] = merged_bmi$BMI[is.na(merged_bmi$BMI_BL)]
+merged_bmi$BMI_6m[is.na(merged_bmi$BMI_6m)] = merged_bmi$BMI.1[is.na(merged_bmi$BMI_6m)]
+merged_bmi$BMI_12m[is.na(merged_bmi$BMI_12m)] = merged_bmi$BMI.2[is.na(merged_bmi$BMI_12m)]
+merged_bmi$Sex[is.na(merged_bmi$Sex)] = merged_bmi$Geschlecht[is.na(merged_bmi$Sex)]
+merged_bmi$Age_BL[is.na(merged_bmi$Age_BL)] = merged_bmi$Alter[is.na(merged_bmi$Age_BL)]
+```
 
-# clean up data frame
-lbmi <- abmi %>% gather(tp, BMI, BMI_BL, BMI_6m, BMI_12m)
+```{r "ADI061 correction", include=FALSE}
+merged_bmi[merged_bmi$subj.ID == "ADI061",c("BMI_BL","BMI_6m","BMI_12m")]
+
+# indication on what is the correct BMI is in Messueberblick.csv
+merged_bmi[merged_bmi$subj.ID == "ADI061",c("BMI_BL","BMI_6m","BMI_12m")] <- 
+  c(bmi_repl$BMI_6m[bmi_repl$subj.ID == "ADI061"], # BL = FU value
+    bmi$BMI.2[bmi$subj.ID=="ADI061"],              # FU = FU2 value (from ADI_aktuell.sav)
+    bmi_repl$BMI_12m[bmi_repl$subj.ID == "ADI061"])# FU2 = FU2 value (from Messueberblick)
+merged_bmi[merged_bmi$subj.ID == "ADI061",c("BMI_BL","BMI_6m","BMI_12m")]
+```
+
+```{r "long format for data frame", include=FALSE}
+# clean up data frame - take refilled BMI of "ADI_aktuell.sav" (bmi_repl)
+lbmi <- merged_bmi %>% gather(tp, BMI, BMI_BL, BMI_6m, BMI_12m)
 lbmi = lbmi[, c("subj.ID", "Age_BL", "Sex", "tp",  "BMI")]
@@
 
-```{r "Compare abmi with bmi from QOL list", include=FALSE }
-# compare abmi (Messueberblick, ADI_aktuell.sav) with bmi from QOL list for those subjects who have it
+```{r "Compare merged_bmi with bmi from QOL list", include=FALSE }
+# does not use QOL list further on
+# compare merged_bmi (ADI_aktuell.sav/ Messueberblick) with bmi from QOL list for those subjects who have it
 QOL_sheet = read.csv(
@@
 
-bmi_compare = merge(QOL_sheet, abmi, by = "subj.ID", all = TRUE)
+bmi_compare = merge(QOL_sheet, merged_bmi, by = "subj.ID", all = TRUE)
 sum(bmi_compare$BMI.BL - bmi_compare$BMI_BL, na.rm = T)
 sum(bmi_compare$BMI.FU - bmi_compare$BMI_6m, na.rm = T)
-sum(bmi_compare$BMI.FU_2 - bmi_compare$BMI_12m, na.rm = T) # because of ADI061
+sum(bmi_compare$BMI.FU_2 - bmi_compare$BMI_12m, na.rm = T) 
+# differences because of  ADI061
 ```
@@
 Correct group information, based on BMI trajectory, seems to be in sheet `QOL` (see ADI080, ADI036).
-"
 
@@
 ```{r "Write group coding for all subjects", echo=FALSE}
-bmi_group_recoded <- bmi_group[!(bmi_group$subj.ID == "ADI061" &
-                         bmi_group$tp == "bl"),]
-# recode fu and fu2 into bl and fu
-bmi_group_recoded$tp[(bmi_group_recoded$subj.ID == "ADI061" &
-                         bmi_group_recoded$tp == "fu")] = "bl"
-bmi_group_recoded$tp[(bmi_group_recoded$subj.ID == "ADI061" &
-                         bmi_group_recoded$tp == "fu2")] = "fu"
-
-write.csv(bmi_group_recoded[bmi_group_recoded$tp=="bl",], "/data/p_02161/ADI_studie/metadata/group_list_all.csv")
-# as all other tables have the original coding for tp in ADI061, we have wo continue working with bmi_group and make the same recoding in the very end again.
+write.csv(bmi_group[bmi_group$tp=="bl",], "/data/p_02161/ADI_studie/metadata/group_list_all.csv")
 ```
@@
 See `/data/pt_02161/Analysis/Preprocessing/qa/README.md` for further information.
+
 ```{r "Loading rs QA infos from /data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv", echo=FALSE}
@@
 }
+
+```
+
+```{r "Exclude subjects with more than 3 time points"}
+# exclude if subject info does not have pattern ADI000_bl ADI000_fu or ADI000_fu2
+FS_QA <- FS_QA[stringr::str_detect(FS_QA$Subject, "^ADI\\d{3}_(bl|fu|fu2)$"),]
+
 bmi_group_MRI_QA=merge(bmi_group_MRI,FS_QA, by=c("subj.ID","tp"),all=TRUE)
@@
 res = read.csv(
-  "/data/pt_02161/Results/Project1_headmotion/summary_FDvals_based_on_subjects_rs.csv",
+  "/data/pt_02161/Results/Project1_headmotion/summary_FDvals_correctADI061.csv",
   header = FALSE
@@
 
-However, because one participant had surgery after the first fu, baseline data will be disregarded and fu and fu2 will be recoded into bl and fu, respectively.
-
-```{r "Reassignment of ADI061 data"}
-# remove baseline measure (actually two time points before surgery)
-bmi_group_MRI_QA_FD_recoded <- bmi_group_MRI_QA_FD[!(bmi_group_MRI_QA_FD$subj.ID == "ADI061" &
-                         bmi_group_MRI_QA_FD$tp == "bl"),]
-bmi_group_MRI_QA_FD_recoded$original_tp <- bmi_group_MRI_QA_FD_recoded$tp
-# recode fu and fu2 into bl and fu
-bmi_group_MRI_QA_FD_recoded$tp[(bmi_group_MRI_QA_FD_recoded$subj.ID == "ADI061" &
-                         bmi_group_MRI_QA_FD_recoded$original_tp == "fu")] = "bl"
-bmi_group_MRI_QA_FD_recoded$tp[(bmi_group_MRI_QA_FD_recoded$subj.ID == "ADI061" &
-                         bmi_group_MRI_QA_FD_recoded$original_tp == "fu2")] = "fu"
+```{r final sample}
+final_sample=bmi_group_MRI_QA_FD
 ```
@@
 ```{r "Some characteristics of this sample", echo=FALSE}
-final_sample=bmi_group_MRI_QA_FD_recoded
+final_sample=bmi_group_MRI_QA_FD
 hist(final_sample$meanFD)

### 2021-02-02 manuscript/draft_rsfmri.Rmd

@@
   bookdown::pdf_book:
+    extra_dependencies: "subfig"
     fig_caption: yes
-    always_allow_html: true
     number_sections: yes
     toc: yes
+  pdf_document:
+    toc: yes
     extra_dependencies: "subfig"
+    fig_caption: yes
+    number_sections: yes
 bibliography: "bibliography.bib"
@@
 ```{r "To do", include=FALSE}
-
-# [ ] Evaluate SPM model comparison (more activation in SPM..)
-# [ ] Discussion
-# [ ] brain images mit nilearn plotting tools einbinden
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] repeat analysis with exclusion of 10% of worst data ??
-# [ ] Network tests
 # [ ] Swap mFD with logmFD ??
-
 ```
@@
 #library(lsr)
-library(wesanderson) # colour palette
+library(wesanderson) # color palette
 #library(haven)
@@
 options(knitr.table.format = "latex")
-options(knitr.kable.NA = '') # hide NA in tables
-knitr::knit_engines$get()
+options(knitr.kable.NA = ' ') # hide NA in tables
 # set working directory of knit (RMarkdown) to directory of Rproj
@@
 # Abstract
+
+Obesity imposes serious health risks and is consistently associated with differences in brain activity, including resting-state functional connectivity which can be effectively treated by bariatric surgery.
+In this study, we investigated the effects of bariatric surgery and weight loss indicated by BMI on functional connectivity in the reward network and default mode network (DMN).
+Therefore, 33 bariatric surgery patients and 15 obese waiting-list control patients underwent MRI at baseline (before surgery/waiting period), after 6 and 12 months.
+To adequately control for motion artifacts shown to be an issue in obesity studies, we applied different denoising pipelines and tried to disentangle the effects of weight loss and head motion in the second level analyses.
+Although we did not find the proposed group-by-time interaction effect, head motion and BMI were significantly associated with functional connectivity.
+When adjusting for differences in head motion within and between individuals using mean framewise displacement, both, BMI decrease and higher average head motion, resulted in stronger functional connectivity in medial posterior frontal regions within the reward network.
+While these findings could imply that severe weight loss changes reward-related neural processing in obese patients, the overlapping effects of BMI change and head motion on FC in this brain area, together with previous findings showing a strong covariance between BMI and head motion renders definite conclusions difficult.
+Future studies with a more rigorous control of within-scanner head motion during data acquisition are warranted to further understand the effects of obesity surgery on functional connectivity measures.
 
@@
 These changes in appetitive and reward signaling likely relate to altered gut-brain communication, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin [ @Brutman_2019].
+
 Yet, precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
@@
 Similarly, the profile of DMN connectivity with frontal regions, including ACC, frontal superior gyrus and the right OFC,  was similar in patients after bariatric surgery and normal weight controls, but dissimilar to that of obese controls [@Frank_2013].
+
 Another important brain network in the context of obesity is the reward network, comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
@@
 Most studies have investigated small cohort of patients, often without adequate control group, and employed exploratory statistics [@George_2016].
+
 In this study, we aimed to extend the current literature and investigate differences of DMN and reward network connectivity before and after bariatric surgery in preregistered analyses and in comparison with a contemporaneously measured waiting-control group.
@@
 
-```{r "tableSample", echo=FALSE}
-#tableSample <- mk_SampleTable(final)
-tableSample <- readRDS(file = "../report/tab/tableSample.rds")
-
-tab <-
-  knitr::kable(
-    tableSample,
-    row.names = TRUE,
-    format = "latex",
-    booktabs = T,
-    linesep = "",
-    caption = "Distribution of data points at months after intervention")
-tab
-```
-
 ## Sample and study design
@@
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment.
-Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availability in \@ref(tab:tableSample)).
+Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availabile.
 In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
@@
 <!-- BASELINE CHARACTERISTICS -->
-Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mean framewise displacement (FD), and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
@@
 <!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
-
-```{r tableDescr, include=FALSE}
-# table for descriptive statistics
-tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")
-
-knitr::kable(
-    tableDescr,
-    row.names = FALSE,
-    format = "latex",
-    booktabs = T,
-    linesep = "",
-    caption = "BMI and mean FD of participants acquired in each condition and time point")
-```
 
@@
 
-```{r figBMIdescr, fig.align='center', out.width="90%", fig.cap="BMI over time.", echo=FALSE}
+```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
@@
 ```
-<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and log mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
 n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
@@
 As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020].
-Thus, we did not exclude any participants based on mean framewise displacement (FD) estimates beyond those who failed the anatomical preprocessing.
+Thus, we did not exclude any participants based on mean FD estimates beyond those who failed the anatomical preprocessing.
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD (mFD).
-```{r carpetplot, out.width="49%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap=c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
+In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD.
+```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
 carpets=c("../report/carpetplotadi006fu2.jpeg","../report/carpetplotadi16fu.jpeg")
@@
 We used GSR regressed data as input and clusterwise bootstrapping with N=999.
-Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of p_FWE < 0.05.
+Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
 We extracted the average connectivity from these masks for GSR regressed data.
-
-<!-- *In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
-The mask will include all voxels surviving FWE-correction of p<0.05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"* -->
+*In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
+The mask will include all voxels surviving FWE-correction of p<0.05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"*
 
@@
 without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
-<!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate (see Figure \@ref(fig:figDesignmatrix)).
+<!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
 The resulting factorial design <!-- check: flexible factorial?
 --> contained one regressor for each time point per group.
-```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap=c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
-# fig.show="hold",  out.width="40%", fig.height = 8,
 
-DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
-DesignMatricesList$plotA
-DesignMatricesList$plotB
-#gridExtra::grid.arrange(plotA, plotB, nrow=1)
-```
 We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
@@
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-<!-- Changed manuscript here
-investigated their differential effects on the complete sample and subsequently in a exploratory fashion for the intervention group only as the effect of longitudinal change in BMI was expected to be driven by this subsample.
-  Baseline age, sex, and log mean framewise displacement (FD) were entered into the first and second model as nuisance variables.
-Age and sex were between-subject covariates while FD is time-varying.
-Mean FD was log-transformed and age was demeaned by subtracting the overall mean across subjects and time points.
- the inclusion of log mean FD may be disputed and done optionally.
-For that reason, we performed each analysis separately one with mean FD (three and without.
-Differences in results will be reported.
-We consider this sensitivity analysis theoretically more meaningful than the preregistered comparison with the nuisance covariate free model and a model with four nuisance covariates (including baseline BMI).<!-- Prereg: We will perform sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI -->
 <!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen.
---> For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
+-->For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
 
@@
 res <- read.csv("../report/FDFCcorrelations.csv")
+rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
+colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
 ```
@@
 
-```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
+```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
 figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
-(figList$fig1 | figList$fig2) / figList$fig3
+#(figList$fig1 | figList$fig2) / figList$fig3
+
+lay <- rbind(c(1,1,2,2),
+             c(NA,3,3,NA))
+gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
 ```
 
-```{r figNetworks, echo=FALSE, out.width="33%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (p FWE<0.05)", fig.subcap=c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
 
@@
 knitr::include_graphics(networks)
-tsnr <- "../report/tsnr.jpeg"
-knitr::include_graphics(tsnr)
 ```
@@
 Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks)).
-
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%",  eval=FALSE}
-knitr::include_graphics(tsnr)
-#readRDS(file = "../report/fig/fig_tSNR.rds")
-```
 
@@
 ## Interaction effect group by time on aggregated FC (preregistered analysis)
-There was no significant time by group interaction for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models)
-```{r load agg FC data and prepare, echo=FALSE}
-
+```{r load agg FC data and prepare, echo = FALSE}
 final_FC <- read.csv("../report/final_FC.csv")
@@
 
-```{r figaggDMN, fig.cap="DMN connectivity per group over time", out.width="75%", echo=FALSE}
+```{r figaggFC, fig.cap="Mean network connectivity per group over time separately gor the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the Reward network"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
 fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
+fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
    fig_DMNconn
-```
-
-```{r figaggRew,fig.cap="Reward network connectivity per group over time", out.width="75%", echo=FALSE}
-fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
 fig_Rewconn
 ```
+There was no significant time by group interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models)
 
-```{r identify outliers in these plots, echo=FALSE}
+```{r identify outliers in these plots, eval=FALSE}
 ### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
@@
 
-There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figures \@ref(fig:figaggDMN) and \@ref(fig:figaggRew) and supplements for a detailed summary of the models).
+There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC)) and supplements for a detailed summary of the models).
 
@@
 
-```{r "qa check DMN model age,sex", echo=FALSE}
+```{r "qa check DMN model age,sex", eval = FALSE}
 diagnostics.plot(R0)
@@
 
-```{r, "qa check DMN model age,sex, mFD", echo=FALSE}
+```{r, "qa check DMN model age,sex, mFD", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r, "qa check Rew model age,sex", echo=FALSE}
+```{r, "qa check Rew model age,sex", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r "qa check Rew model age, sex, mFD", echo=FALSE}
+```{r "qa check Rew model age, sex, mFD", eval = FALSE}
 diagnostics.plot(R0)
@@
 When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \@ref(tab:tableBMImodel).
-```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap=c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
 knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
@@
 supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction (p=$0.028$), also when evaluating the GSR corrected data (p=$0.038$).
-```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
+```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 This result remained similar when using GSR corrected data.
-```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
                           "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
+```
+
+```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", out.width = "50%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
 ```
@@
 ```
-Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD.
+Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD.
 ```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 
-The association also remained significant when we split mFD into average and change in mFD (p = `r round(p_mean_BMI,4)`) and there was no significant association of mean or change in mFD with DMN FC.
+The association also remained significant when we split mFD into average and change in mFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in mFD with DMN FC.
 
@@
 
-<!-- [OLD stuff] GM mask: Group*Time Interaction -->
-<!-- ### DMN for two timepoints
-BMIONLYIG
-Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps).
-However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
-<!-- BMI2TP
-Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants.
-Omitting mean FD did not change these results.
--->
-<!-- GM mask
-Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
-In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
-
-<!-- Parametric and non-parametric significance test for the model testing for time-by-group interaction did not show any different results when using a grey matter mask.
-Again, parametric estimation and non-parametric estimation differed in regard to the time-by-group interaction of PCC and voxel ($18,42,30$) in the connectivity maps without GSR correction.
-While co-activation was significant in parametric estimation ($Z = 4.72$, $q_{FDRcorr} = 0.040$), it was non-significant when using wild bootstrap ($q_{\mathrm{FDRcorr}} = 0.998$)<!-- also FWE-corr n.s -->
-
 ### Effect of FD (exploratory analysis)
@@
 
-```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap="Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
@@
 
-## Effect of BMI and FD (exploratory analysis)
+<!-- ## Effect of BMI and FD (exploratory analysis) -->
 
-Average BMI and change in BMI remain covariates of interest.
+<!-- Average BMI and change in BMI remain covariates of interest.
 While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as.
 Age and sex remain nuisance covariates.
+-->
 
-<!-- ### Reward -->
-<!-- cc -->There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
+<!-- <!-- ### Reward -->
+<!-- cc There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.-->
+<!-- Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
 Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
-<!-- gsr -->When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds.
-On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank.
+-->
+<!-- <!-- gsr When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds. -->
+<!-- On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank. -->
 
@@
 ```
-<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
+<!-- cc The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
 There is a significant association between average BMI and the connectivity between PCC and three other clusters.
 Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
-<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
-```{r "BMIFDagesex_NACC_avgFD_act", fig.show = "hold", out.width = "50%", fig.cap="Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
-knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf")
-knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf")
-```
-
-```{r "BMIFDagesex_NACC_cgnBMI_deact", fig.show = "hold", out.width = "50%", fig.cap="Positive association of change in BMI and FC of NAcc with cluster, adjusted for age, sex, average BMI, average mFD and change mFD", fig.subcap=c("in AROMA+CC", "in AROMA+CC+GSR data")}
-knitr::include_graphics("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
-knitr::include_graphics("../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
-```
-
-```{r "BMIFDagesex_PCCcc_avgBMI_deact", out.width = "75%"}
-knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
-```
-
-## Effect on hypothalamic connectivity (exploratory analysis)
-...
+Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.-->
+<!-- gsr As before, the effect did not survive when thresholding if the images had been GSR corrected.-->
 
@@
 # Discussion
+In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) in group analysis.
+We found no interaction effect of group and timepoint or main effects on whole-brain FC of the PCC and Nacc, which we used to define the resting state networks DMN and reward network, respectively.
+Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
+In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme.
+When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI between the three time points and the connectivity of Nacc with a posterior-medial frontal cluster emerged.
+This result was significant in both denoising schemes.
+Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns.
+Finally, higher average head motion was associated with increased Nacc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
 
-#### Summary
-Utilizing resting-state functional MRI on morbidly obese individuals that either underwent or awaited bariatric surgery over the course of one year.
-<!-- general --> Building on the evidence for post-surgery changes in functional activity, this study explored the longitudinal relationship between surgery-induced weight loss and functional integrity in target networks indicated by resting-state functional connectivity.
-Over the course of one year we collected neuroimaging and anthropometric data from `r cs["Sum","IG"]` morbidly obese patients undergoing bariatric surgery and compared them to the longitudinal data of  `r cs["Sum","KG"]` patients on the waiting list.
-<!-- model specific --> Looking at time-by-group interactions, we primarily investigated whether changes in functional connectivity over time vary depending whether participants have received surgery or not, while controlling for individual variance in resting-state connectivity over time and at baseline.
-Beyond that, main effects of the intervention and of the time were examined.
-The second model further investigates if changes in connectivity are a function of an average BMI (controlled for individual variance in BMI) and the longitudinal change in BMI.
-<!-- sensitivity analyses -->To get a better understanding of our results, we conducted multiverse exploratory analyses representing different choices regarding preprocessing pipelines, inclusion of covariates and subsets of the data relevant for interpretation.
+Overall, we could not confirm our preregistered hypotheses of an interaction effect between group and time point on DMN and reward network connectivity.
+Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in lien with improved processing of visceral and bodily signals  after surgery [@Li_2018; @McFadden_2013].
+Further, we expected FC between reward network regions to decrease, which would support a reduction of reward hypersensitivity and hedonic motivation to eat following bariatric surgery [@Scholtz_2014; ].
+Yet, our analyses pointed to a possible effect of BMI change on the interplay of the reward network and a posterior frontal brain region.
+Here, against our expectation, we found stronger BMI decrease to predict increases in connectivity of the Nacc and posterior-medial frontal brain.
+This pattern could be interpreted as enhanced connectivity between regions governing salience attribution and reward processing, as expected in obesity.
+However, it might also indicate a more close monitoring of reward signaling, which would be in line with lower desire to eat high-calorie food items after surgery.
+Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity.
+Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2019], and therefore conducted careful analyses of the impact of head motion on our results.
+To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
+This result may be due to the presence of multi-collinearity between change in FD and BMI which appeared more pronounced in the split model, and lead to unreliable estimations of effects and standard errors.
+Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
+In favour of this argument speaks the survival of the cluster when using global signal regressed data.
+Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
+Thus, confounding of a positive effect of head motion on posterior-frontal connectivity and the inversely directed BMI effect seems unlikely.
 
-#### Support of original hypotheses
-<!-- false positive -->
-To summarize, while our primary hypothesis on group time interactions could not be confirmed, there was a significant relationship between functional activation
-and the continuous BMI measures for some but not all analyses.
+Head motion also played a role in the association of higher BMI and reduced DMN connectivity.
+While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN connectivity was negatively associated with BMI in both denoising schemes.
+Thereby, this results echos a previous finding from our group, and may be interpreted as accelerated DMN decline in relation to higher BMI.
+Yet, similar doubts regarding the complete removal of motion confounding remain in this study.
+The major strength of our study is prospective intervention controlled design.
+We compared BS patients to an obese control group who were scanned after the same time intervals and  equivalent in BMI, treatment history and  treatment recommendation to our intervention group [@Thiese_2014].
+Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
+In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended method to deal with longitudinal repeated measures @Guillaume_2014.
 
-- DMN: average BMI but only with lenient AROMA-CC but then for both covariate structures age/sex/mFD and age/sex; effect stable when separating avg mFD and cgn FD into two regressors
-- Reward: effect of change BMI only when separating mFD, bit high peak activation
-The individuals BMI averaged across measures taken throughout the one-year period negatively correlated with functional connectivity between PCC and [regions from clusters].
-Results indicate a decrease a functional connectivity within the default mode network.
-<!-- discuss consistency with previous literatur -->
-... not? in accordance with our hypothesis regarding network activity changes.
-
-Regarding the reward network, we could not predict changes in functional connectivity with the preregistered models.
-However, when investigating the unique contributions to variance of head motion measures and BMI measures, we detected a significant relationship between the individuals' change in BMI throughout the year.
-This may be preliminary evidence for changes in the functional connectivity between reward and frontal regions induced by weight-loss.
-<!-- discuss consistency with previous literatur -->
-
-In our sample, extend of micromovements of the head averaged across all measurements is associated with increased functional connectivity between reward and frontal regions.
-<!-- explanation?
-overlap with region from change in BMI?
--->
-
-<!-- Head motion as covariate, multicollinearity -->
-Based on concerns expressed in the literature, we tried to closely inspect and rigorously control for the influence of head motion.
-The presence of multi-collinearity of our predictors has been established before [@Beyer_2019].
-As described by @Monti_2011, in the presence of multicollinearity, the unique contribution of the regressor for the prediction cannot be disentangled.
-Surprisingly, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region is only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
-However, such unexpected findings would be in accordance with @Monti_2011 stating that <!-- not only will confidence in $\beta$ estimates suffer but  --> estimates can behave erratic depending on which regressors are included in the model.
-
-Notwithstanding the difficulties of quantitatively disentangling multicollinearity between head motion and BMI, preliminary conclusions of our analyses may be...
-<!-- Given the multiplicity of of models resulting from variations of inspected networks and sensitivity analysis for different preprocessing and correction approached, the obtained significant correlations between seed regions and voxels are likely to be false positive.
-Ventures to verify effects by re-calculating with non-parametric Wild bootstrap confirmed supported presumption as effects were not robust across estimation procedures.
--->
-
-<!-- limitations -->
-Without an a priori power analysis it is not possible to determine if there is no effect or
-To clarify, extended considerations should be made to use a priori power analysis (e.g. with fMRIpower) or check evidence of absence with alternative statistical procedures.
-
-#### Similarity of results with literature (random notes)
-Our results are not consistent with the present literature.
-Although there are few studies investigating the impact of bariatric surgery on functional connectivity at resting state, the present articles do significant changes in network activation [@;@;@].
-<!-- publication bias -->
-
-#### Interpretation (random notes)
-<!-- absence of evidence --> However, hypotheses about the plasticity functional networks following bariatric surgery could not be validated.
-Due to the shortcomings of the inferential paradigm used, our null findings may either put conclusions drawn from previous research into question or may neither provide evidence for the presence nor the absence of the effect.
-<!-- Control for satiety?
---> A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
+Opposed to flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
+Limitations of our study include the low number of patients who participated in all three time points.
+In total, only 19 participants contributed to the estimation of the longitudinal effects over three time points.
+Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI.
+While this sample size is comparable to previous resting state fMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
+We used seed-based connectivity to derive large-scale brain networks.
+While this approach yielded correct DMN and reward network maps, it is a univariate approach not taking into account the inter-relatedness of subnetworks or other regions.
+Furthermore, we did not explicitly investigate the interaction of these networks with attention, cognitive control or sensorimotor networks, which has been previously proposed to be altered by bariatric surgery.
+We did not control for hunger or satiety in our design.
+A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
 After the intervention, functional connectivity was higher in bariatric patients.
 It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
-prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016]. putamine, insula
+Prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016].
+putamine, insula
+Size and composition of our sample did not allow separate analyses for the sexes.
+However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
 
-<!-- GENERALIZABILITY -->
-#### Internal validity
-<!-- advantage: prospective intervention control study -->Common challenge in medical intervention studies is the difficulty to draw causal inference about the effect of interventions such as surgeries because randomization is often unethical or unfeasible as done in randomized control trials.
-Previous studies have compared the sample of bariatric patients to lean individuals, behavioral dieters, or obese individuals without a recommendation for bariatric surgery.
-Despite the accumulated evidence, effects may still be attributable to baseline differences, contributing factors to the condition or other psychophysiological and dispositional difference between groups [@].
-In order to build upon and complement previous evidence, we aimed to isolate the impact of the surgical procedure on changes in resting-state connectivity to a maximum possible extent by conducting a prospective non-randomization trial study.
-Such a study design is secondary to randomized control trials in regard to internal control by means of temporality and controllability.
-We established temporality by contrasting pre- and post-surgery measures change while controlling for individual variance.
-<!-- Such longitudinal analyses support the validity of conclusions about the effects of the surgery.
--->
-The waiting list control group is equivalent in BMI, treatment history and  treatment recommendation and therefore allows controlling for concurrent factors [@Thiese_2014].
-Incorporating these consideration, our study aimed at a higher internal validity for conclusions about the changes in functional connectivity specific to surgery-induced changes.
-
-#### Study sample
-<!-- advantage: representative sample -->
-Size and composition of our sample did not allow separate analyses for the sexes.
-However, the disproportionate sex distribution is reflective prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
-
-<!-- mention larger sample size -->
-
-#### Marginal model
-<!-- advantage: controlling for individual variance -->
-Due to the limitation of the traditional ANOVA framework (with respect to the spherity assumption, handling missing data, handling continuous data), it is often recommended to use modern models such as mixed-effect models and marginal models [@Chen_2013;@Guillaume_2014;@McFarquhar_2019].
-@McFarquhar_2019 demonstrated that specifying complex multi-level models and deriving the correct contrast weights with standard neuroimaging software packages such as SPM and FSL can be quite tedious and complicated if done appropriately.
-<!-- However, for the purpose of comparison, we re-ran the analysis for the time-by-group--interaction with age and sex as nuisance covariates.
-We opted for a simplified model investigating the time-by group interaction in ...
-because different nuisance covariates, time-invariant or time-varying, create additional complications.
-An advantage of the time-by-group interaction in our model is that the $F$-ratio does not require partitioning the error and includes only factors and time-invariant nuisance convariates.
-The model was set up as flexible factorial design with factors "subject", "group", and "time" assuming unequal group variances.
--->
-<!-- Differences could be attributed to different corrections for non-spherity and assumptions on voxel covariance structures.
-Detailed comparisons between SPM12, SwE, and other software packages have been described elsewhere [@McFarquhar_2016].
--->
-<!-- conclusion??? -->
-Opposed to flexible factorial models, where each individual is modelled separately, marginal models are parsimonious and preserve more degrees of freedom.
-
-<!-- IMPLICATIONS -->
-<!-- Relevance of rsfMRI as biomarker -->
-
-<!-- For all patients included, conservative obesity treatment had failed, e.g.
-behavioral diet intervention.
-As shown before, patients who were successful in behavioral diet increased connectivity in the default mode network.
-This might be indicative for self-referent information processing.
--->
+Taken together, this study showed that effects of bariatric surgery on resting state are subtle, and often confounded by head motion.
+Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to altered food preference and choice in patients after bariatric surgery.
+Further, larger, well-controlled and preregistered neuroimaging studies are necessary to understand the neural underpinnings of altered gut-brain communication after bariatric surgery.
 

### 2021-02-02 manuscript/supplements_rsfmri.Rmd

@@
 #library(lsr)
+#library(haven)
 library(wesanderson) # colour palette
-#library(haven)
 
@@
 options(knitr.table.format = "latex")
-options(knitr.kable.NA = '') # hide NA in tables
+options(knitr.kable.NA = ' ') # hide NA in tables
 # set working directory of knit (RMarkdown) to directory of Rproj
-opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-setwd(knitr::opts_knit$get("root.dir"))
-source('report/figures.R')
-source("report/tables.R")
-source("report/create_sample_df.R")
+```
+
+```{r "tableSample", echo=FALSE}
+
+tableSample <- readRDS(file = "../report/tab/tableSample.rds")
+
+tab <-
+  knitr::kable(
+    tableSample,
+    row.names = TRUE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = "Distribution of data points at months after intervention")
+tab
+```
+
+```{r tableDescr, include=FALSE}
+# table for descriptive statistics
+tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")
+
+knitr::kable(
+    tableDescr,
+    row.names = FALSE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = "BMI and mean FD of participants acquired in each condition and time point")
+```
+
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap = c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
+DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
+DesignMatricesList$plotA
+DesignMatricesList$plotB
+#gridExtra::grid.arrange(plotA, plotB, nrow=1)
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=FALSE}
+tsnr <- "../report/tsnr.pdf"
+knitr::include_graphics(tsnr)
 ```
@@
 ```{r load agg FC data and prepare, echo=FALSE}
-final=readRDS("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/tab/final.rds")
-
-agg_FC_DMN=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt")
-agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt")
-agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
-agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
-for (i in c(1:nrow(agg_FC_DMN_ID))){
-  tmp=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
-  agg_FC_Rew[i,"subj.ID"]=strsplit(tmp,"_")[[1]][1]
-  agg_FC_Rew[i,"tp"]=strsplit(tmp,"_")[[1]][2]
-  tmp=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
-  agg_FC_DMN[i,"subj.ID"]=strsplit(tmp,"_")[[1]][1]
-  agg_FC_DMN[i,"tp"]=strsplit(tmp,"_")[[1]][2]
-}
-
-colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID", "tp")
-colnames(agg_FC_DMN)=c("mean_DMN_conn","sd_DMN_conn","subj.ID","tp")
-
-final_FC=merge(final, agg_FC_Rew, by=c("subj.ID","tp"), all.x=TRUE)
-final_FC=merge(final_FC, agg_FC_DMN, by=c("subj.ID","tp"), all.x=TRUE)
-final_FC$subj.ID=as.factor(final_FC$subj.ID)
-final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")
-
-#calculate mean BMI over timepoints
-final_FC.meanBMI=tapply(X=final_FC$BMI,
-                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
-final_FC$mean.BMI=
-  final_FC.meanBMI[as.numeric(final_FC$subj.ID)]
-final_FC$within.BMI=
-  final_FC$BMI-final_FC$mean.BMI
-
-#calculate mean FD over timepoints
-final_FC.meanFD=tapply(X=final_FC$logmFD,
-                        INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
-final_FC$mean.logmFD=
-  final_FC.meanFD[as.numeric(final_FC$subj.ID)]
-final_FC$within.logmFD=
-  final_FC$logmFD-final_FC$mean.logmFD
+final_FC <- read.csv("../report/final_FC.csv")
 ```
@@
 
-```{qa check DMN model age,sex, echo=FALSE}
+```{r qa check DMN model age,sex, echo=FALSE}
 diagnostics.plot(R0)
@@
 
-```{qa check DMN model age,sex, mFD, echo=FALSE}
+```{r qa check DMN model age,sex, mFD, echo=FALSE}
 diagnostics.plot(R0)
@@
 ```{r DMN conn over three timepoints, echo=FALSE}
-fig_DMNconn<-mk_figDMNdescr(final_FC)
+fig_DMNconn <- readRDS(file = "../report/fig/figDMNconn.rds")
 fig_DMNconn
@@
 ```
-```{qa check Rew model age,sex, echo=FALSE}
+```{r qa check Rew model age,sex, echo=FALSE}
 diagnostics.plot(R0)
@@
 
-```{qa check Rew model age,sex,mFD, echo=FALSE}
+```{r qa check Rew model age,sex,mFD, echo=FALSE}
 diagnostics.plot(R0)
@@
 ```{r Rew conn over three timepoints, echo=FALSE}
-fig_Rewconn<-mk_figRewdescr(final_FC)
+fig_Rewconn <- readRDS(file = "../report/fig/figRewconn.rds")
 fig_Rewconn

### 2021-02-03 manuscript/draft_rsfmri.Rmd
added functional decoding and modified discussion

@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas.
+<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
 We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
@@
 Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.
+
+### Functinal decoding
+In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
+We uploaded the contrast images for change BMI and average mFD on [Neurovault](), and applied the decoding classifier.
+This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
+We reported the three top terms for both contrasts.
 
@@
 
-```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
+```{r "calc DMN two tp age and sex" ,results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
@@
 
-```{r "qa check DMN model age,sex", eval = FALSE}
+```{r "qa check DMN model age sex", eval = FALSE}
 diagnostics.plot(R0)
@@
 
-```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
+```{r "DMN2tp results age sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r "calc DMN two tp, age and sex and mFD", results = 'asis', echo = FALSE}
+```{r "calc DMN two tp age and sex and mFD", results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 
-```{r, "qa check DMN model age,sex, mFD", eval=FALSE}
+```{r "qa check DMN model age sex mFD", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r, "qa check Rew model age,sex", eval=FALSE}
+```{r "qa check Rew model age sex", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
+Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
 Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
@@
 <!-- cc -->
-For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
-supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
+For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region (AAL).
+The peak voxel was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
@@
 ```
-
-```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", out.width = "50%"}
-knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
-```
+###### Functional decoding
+Functional decoding of these two activation patterns showed different top association terms.
+For the BMI change cluster (frontal	0.246; anterior insula	0.23; inferior frontal	0.194) were the top terms, while for the average mFD () were the meta-analytic activation maps most similar to this contract.
+The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
 
 #### Aggregated FC analysis
+##### DMN (BMI and mFD model)
 ```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 
-#### Reward network, mean and within variability in BMI and logmFD
+##### Reward network (BMI and mFD model)
 ```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
@@
 # Discussion
+#### Summary
 In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
-Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) in group analysis.
-We found no interaction effect of group and timepoint or main effects on whole-brain FC of the PCC and Nacc, which we used to define the resting state networks DMN and reward network, respectively.
-Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) on the group level.
+We found no interaction effect of group and time point on whole-brain FC of the PCC and NAcc.
+We also did not observe main effects of group and time point on the FC of these DMN and reward network hubs, respectively  Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
 In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme.
@@
 
-Overall, we could not confirm our preregistered hypotheses of an interaction effect between group and time point on DMN and reward network connectivity.
-Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in lien with improved processing of visceral and bodily signals  after surgery [@Li_2018; @McFadden_2013].
-Further, we expected FC between reward network regions to decrease, which would support a reduction of reward hypersensitivity and hedonic motivation to eat following bariatric surgery [@Scholtz_2014; ].
-Yet, our analyses pointed to a possible effect of BMI change on the interplay of the reward network and a posterior frontal brain region.
-Here, against our expectation, we found stronger BMI decrease to predict increases in connectivity of the Nacc and posterior-medial frontal brain.
-This pattern could be interpreted as enhanced connectivity between regions governing salience attribution and reward processing, as expected in obesity.
-However, it might also indicate a more close monitoring of reward signaling, which would be in line with lower desire to eat high-calorie food items after surgery.
+In this study, we could not confirm these preregistered hypotheses and found no interaction effect between group and time point on DMN and reward network connectivity.   
+Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
+Further, we expected FC between reward network regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in reward network regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014;   @Cerit_2019].
+Yet, in an exporatory analysis,  stronger BMI decrease predicted higher connectivity of the Nacc and a cluster in a posterior-medial frontal brain region.
+Based on the Harvard-Oxford atlas and NeuroSynth decoding this region might be part of the salience network and involved in action preparation.
+AThe enhanced connectivity between the Nacc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
+Yet, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
+Previously,  a decrease in local (regional homogenity and frequency of low-amplitude oscillations) and global connectivity (degree centrality (DC)) measures in a similar region of the left SMA was reported after glucose administration  [@AlZubaidi_2018].
+Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
+Importantly, a global decrease in connectivity (DC) might go along with selectively increased connectivity, i.e.
+a shift from integration towards segregation and more specialized processing.
+However, this interpretation is highly speculative, and more studies are needed to corroborate these findings.
+
 Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity.
@@
 To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
-This result may be due to the presence of multi-collinearity between change in FD and BMI which appeared more pronounced in the split model, and lead to unreliable estimations of effects and standard errors.
+This result may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
 Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
-In favour of this argument speaks the survival of the cluster when using global signal regressed data.
+This argument is supported by the  survival of the cluster when using global signal regressed data, and the distinct results of the decoding analysis.
 Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
-Thus, confounding of a positive effect of head motion on posterior-frontal connectivity and the inversely directed BMI effect seems unlikely.
-
+Thus, confounding of the negative BMI change effect and the positive average head motion effect  on posterior-frontal connectivity  seems unlikely.
 Head motion also played a role in the association of higher BMI and reduced DMN connectivity.
 While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN connectivity was negatively associated with BMI in both denoising schemes.
-Thereby, this results echos a previous finding from our group, and may be interpreted as accelerated DMN decline in relation to higher BMI.
-Yet, similar doubts regarding the complete removal of motion confounding remain in this study.
+Thereby, this results echos a previous finding from our group, and may be interpreted as  accelerated age-related decline of the DMN  in relation to the cardiometabolic risk factor BMI.
+Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding also remain in this study [@Savalia_2017].
+
 The major strength of our study is prospective intervention controlled design.
-We compared BS patients to an obese control group who were scanned after the same time intervals and  equivalent in BMI, treatment history and  treatment recommendation to our intervention group [@Thiese_2014].
+We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals and , [@Thiese_2014].
 Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
-In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended method to deal with longitudinal repeated measures @Guillaume_2014.
-
+In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
 Opposed to flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
 Limitations of our study include the low number of patients who participated in all three time points.
-In total, only 19 participants contributed to the estimation of the longitudinal effects over three time points.
+In total, only X participants contributed to the estimation of the longitudinal effects with at least two  time points.
 Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI.
 While this sample size is comparable to previous resting state fMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
+
 We used seed-based connectivity to derive large-scale brain networks.
-While this approach yielded correct DMN and reward network maps, it is a univariate approach not taking into account the inter-relatedness of subnetworks or other regions.
-Furthermore, we did not explicitly investigate the interaction of these networks with attention, cognitive control or sensorimotor networks, which has been previously proposed to be altered by bariatric surgery.
-We did not control for hunger or satiety in our design.
-A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
-After the intervention, functional connectivity was higher in bariatric patients.
-It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
-Prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016].
-putamine, insula
+While this approach yielded reasonable DMN and reward network maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
+Furthermore, we did not explicitly investigate the connectivity of the DMN and reward network with other large-scale brain networks attention, e.g.
+cognitive control or sensorimotor networks, which has been previously proposed to be altered by obesity and bariatric surgery [@Lips_2014; @Olivo_2016; @Cerit_2019].
+
+We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized?
+breakfast following an overnight fast.
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2015; @Lepping_2015, @Wiemerslage_2016; Ochner_2012_prepostprandial], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow separate analyses for the sexes.
@@
 Taken together, this study showed that effects of bariatric surgery on resting state are subtle, and often confounded by head motion.
-Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to altered food preference and choice in patients after bariatric surgery.
+Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to observed changes in food wanting and choice in patients after bariatric surgery.
 Further, larger, well-controlled and preregistered neuroimaging studies are necessary to understand the neural underpinnings of altered gut-brain communication after bariatric surgery.

### 2021-02-03 manuscript/draft_rsfmri.Rmd
replaced mean FD/log mean FD with mFD

@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] Swap mFD with logmFD ??
 ```
@@
 In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
+We calculated mean framewise displacement (mFD) from the outputs of the motion correction procedure according to @Power_2012.
 For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
@@
 As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020].
-Thus, we did not exclude any participants based on mean FD estimates beyond those who failed the anatomical preprocessing.
+Thus, we did not exclude any participants based on mFD estimates beyond those who failed the anatomical preprocessing.
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD.
+In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mFD.
 ```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
@@
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
+Then, we correlated mFD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
 We reported the mFD-FC correlation per preprocessing scheme.
@@
 We did not perform the analysis as preregistered (i.e.
-without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
+without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mFD (Model 1b).
 <!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
@@
 This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3).
+We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for the logarithm of mFD (Model 2b).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and the logarithm of mFD and variability in both measures, along with baseline age and sex (Model 3).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
 <!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen.
--->For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
+-->For a better understanding of the unique contribution of average and longitudinal change mFD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject mFD + within-subject mFD$ with nuisance covariates age and sex.
 
@@
 Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up.
-In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
+In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and logarithm of mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
 We performed model comparison between R1 and R0 models
 `R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
-`R0 = lmer(agg FC ~ timepoint + group +  + age + (1|subj))`
+`R0 = lmer(agg FC ~ timepoint + group +  age + sex + (1|subj))`
 and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
@@
 Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.
+Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and change in mFD.
 
@@
 
-####  Model 1b (with adjustment for age and sex, meanFD)
+####  Model 1b (with adjustment for age, sex and mFD)
 Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
@@
 
-##### BMI and FD model
+##### BMI and mFD model
 To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
 <!-- ### DMN -->
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting mFD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
 Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
@@
 
-### Effect of FD (exploratory analysis)
+### Effect of mFD (exploratory analysis)
 #### Whole-brain analysis
@@
 <!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average mFD and change mFD will be are published on NeuroVault.
 

### 2021-02-03 code/data_load.Rmd
correct tables with complete data (N=101)

@@
 # ADI_Aktuell  -  main table, BMI info is refilled with Messueberblick entries
-ADIAktuell_üath <- "/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav"
+ADIAktuell_path <- "/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav"
 

### 2021-02-03 manuscript/draft_rsfmri.Rmd
correct tables with complete data (N=101)

--- 2021-02-02 manuscript/draft_rsfmri.Rmd
@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] Swap mFD with logmFD ??
+# [x] Swap mFD with logmFD ??
 ```
@@
 We did not perform the analysis as preregistered (i.e.
-without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
+without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (logmFD) (Model 1b).
 <!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
@@
 Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up.
-In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
+In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and logmFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
 We performed model comparison between R1 and R0 models
@@
 Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.
+Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average logmFD and variability in logmFD.
 
@@
 ####  Model 1b (with adjustment for age and sex, meanFD)
-Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+Similarly, there were no significant clusters when including logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 
@@
 
-```{r "calc DMN two tp, age and sex and mFD", results = 'asis', echo = FALSE}
+```{r "calc DMN two tp, age and sex and logmFD", results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 
-```{r, "qa check DMN model age,sex, mFD", eval=FALSE}
+```{r, "qa check DMN model age,sex, logmFD", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r "DMN2tp results age sex mFD", caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
+```{r "DMN2tp results age sex logmFD", caption="DMN connectivity compared bl to fu, adjusted for age, sex and logmFD", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r "qa check Rew model age, sex, mFD", eval = FALSE}
+```{r "qa check Rew model age, sex, logmFD", eval = FALSE}
 diagnostics.plot(R0)
@@
 <!-- ### DMN -->
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including mFD as nuisance covariate (FWE cluster p<0.05).
+Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE cluster p<0.05).
 When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \@ref(tab:tableBMImodel).
-```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and mFD"), fig.hold=TRUE, out.width = "50%"}
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and logmFD"), fig.hold=TRUE, out.width = "50%"}
 knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
@@
 ##### BMI and FD model
-To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
+To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
 <!-- ### DMN -->
 <!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
+Test statistics only differed slightly from the effect found when using logmFD as nuisance covariate.
 <!-- gsr -->As before, the effect did not survive when using GSR-corrected data.
@@
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
-```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
-motor cortex, was significantly positively related to average mFD.
+motor cortex, was significantly positively related to average logmFD.
 This result remained similar when using GSR corrected data.
-```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average logmFD on connectivity between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
@@
 
-```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", out.width = "50%"}
+```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
 knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
@@
 ```
-Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD.
+Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD.
 ```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 
-The association also remained significant when we split mFD into average and change in mFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in mFD with DMN FC.
+The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in logmFD with DMN FC.
 
@@
 
-As expected from the whole-brain findings, there was no association of mean BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and mFD.
+As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and logmFD.
 
@@
 #### Whole-brain analysis
-To investigate the unique contributions of the average value of and the longitudinal change in mFD, we constructed two additional models.
+To investigate the unique contributions of the average value of and the longitudinal change in logmFD, we constructed two additional models.
 First, BMI measures are excluded from the design matrix.
-Covariance of the mFD and outcome measures is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor.
+Covariance of the logmFD and outcome measures is investigated by splitting logmFD into average logmFD and longitudinal logmFD change and by inspecting associations with each predictor.
 Age and sex were nuisance covariates.
@@
 <!-- #### Reward -->
-As shown in Table \@ref(tab:tableFDmodel), there was a significant association between mFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas .
+As shown in Table \@ref(tab:tableFDmodel), there was a significant association between logmFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas .
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
@@
 
-```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of mean mFD and FC of NAcc with cluster, adjusted for age, sex, mean BMI, change BMI and change mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
@@
 <!-- Average BMI and change in BMI remain covariates of interest.
-While initially used as nuisance covariate, here, mFD is split into average mFD and change in mFD and used as.
+While initially used as nuisance covariate, here, logmFD is split into average logmFD and change in logmFD and used as.
 Age and sex remain nuisance covariates.
@@
 <!-- Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
-Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average mFD.
+Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average logmFD.
 -->
 <!-- <!-- gsr When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds. -->
-<!-- On the other hand, the relationship with average mFD was confirmed by analyses after GSR but cluster size shrank. -->
+<!-- On the other hand, the relationship with average logmFD was confirmed by analyses after GSR but cluster size shrank. -->
 
@@
 
+```{r tableBMIFDmodel}
+# do not change chunk label as it corresponds to the manually defined (in table.R) table label
+tab_BMIFDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[2]]
+tab_BMIFDmodel
+```
+
 ```{r tableFDmodel}
 # do not change chunk labelsas it corresponds to the manually defined (in table.R) table label
-tab_FDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[2]]
+tab_FDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[3]]
 tab_FDmodel
@@
 
-```{r tableBMIFDmodel}
-# do not change chunk label as it corresponds to the manually defined (in table.R) table label
-tab_BMIFDmodel <- readRDS(file = "../report/tab/FCTableList.rds")[[3]]
-tab_BMIFDmodel
-```
 <!-- cc The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
@@
 Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.-->
+Test statistics of these clusters differ only slightly from the effect found for average BMI when using logmFD as nuisance covariate.-->
 <!-- gsr As before, the effect did not survive when thresholding if the images had been GSR corrected.-->
@@
 Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2019], and therefore conducted careful analyses of the impact of head motion on our results.
-To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
-This result may be due to the presence of multi-collinearity between change in FD and BMI which appeared more pronounced in the split model, and lead to unreliable estimations of effects and standard errors.
-Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
+To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average logmFD and change in logmFD and introducing two instead of one regressors into the model.
+This result may be due to the presence of multi-collinearity between change in logmFD and BMI which appeared more pronounced in the split model, and lead to unreliable estimations of effects and standard errors.
+Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and logmFD the effect of change in BMI could be singled out.
 In favour of this argument speaks the survival of the cluster when using global signal regressed data.
-Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
+Further, average logmFD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
 Thus, confounding of a positive effect of head motion on posterior-frontal connectivity and the inversely directed BMI effect seems unlikely.

### 2021-02-03 code/df2SwEtxt.R

@@ -133,7 +133,7 @@ get_txt_for_swe <- function(group = "both", tp = "all",exclFD = FALSE){
   # Model specification incl. design matrix for different options
   output_dir <- parentdir
   if (exclFD == FALSE){
-    output_dir <- file.path(parentdir, "noExclFD")
+    output_dir <- file.path(parentdir, "noExclFD") 
     if (!dir.exists(output_dir)) {dir.create(output_dir)}
   }else if (exclFD == TRUE){
     output_dir <- file.path(parentdir, "ExclFD")
@@ -188,7 +188,7 @@ get_txt_for_swe <- function(group = "both", tp = "all",exclFD = FALSE){
     df <- df[df$tp != "fu2",]
   }
   cat("Data from",group,"group(s) and",tp,"measurements were selected to be kept in the data frame, therefore n =",nrow(df),".\n")
-  
+  cat("All txt files will be saved in:", output_dir)
   
   
   # order dataframe by group / subject / time point
@@ -251,7 +251,7 @@ get_txt_for_swe <- function(group = "both", tp = "all",exclFD = FALSE){
             file='KG_fu.txt')
   write.table(dfo$KG_bl, col.names=FALSE,row.names=FALSE,quote=FALSE,
             file='KG_bl.txt')
-
+  
   # end of computations --------------------------------------------------------
   
   # reset to original working directory

### 2021-02-03 code/extract_agg_FC.sh

@@ -5,6 +5,10 @@
 
 output_dir="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/"
 
+rm -rf $output_dir/DMN/agg_FC_DMN_ID.txt
+rm -rf $output_dir/DMN/agg_FC_DMN.txt
+rm -rf $output_dir/Rew/agg_FC_Rew.txt
+rm -rf $output_dir/Rew/agg_FC_Rew_ID.txt
 
 # DMN
 # threshold: log10(p_FWE=0.05)=-1.30103
@@ -13,9 +17,9 @@ in="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD
 
 fslmaths $in -thr 1.30103 -bin "$output_dir/DMN/DMN_swe_clusternorm_Tstat_lpFWE0.05-WB_c01_mask.nii"
 
-fslmerge -t "$output_dir/DMN/DMN_merged.nii.gz" `ls /data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/Individual_scans/PCC_gsr_z/*`
+fslmerge -t "$output_dir/DMN/DMN_merged.nii.gz" `ls /data/pt_02161/Results/Project2_resting_state/connectivity/calc_DMN_reward_seed_connectivity/*/PCC_gsr_z/pcc_cc_gsr_seed_correlation_z_trans.nii`
 
-ls /data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/Individual_scans/PCC_gsr_z/* >> $output_dir/DMN/agg_FC_DMN_ID.txt
+ls /data/pt_02161/Results/Project2_resting_state/connectivity/calc_DMN_reward_seed_connectivity/*/PCC_gsr_z/pcc_cc_gsr_seed_correlation_z_trans.nii >> $output_dir/DMN/agg_FC_DMN_ID.txt
 
 fslstats -t $output_dir/DMN/DMN_merged.nii.gz -k $output_dir/DMN/DMN_swe_clusternorm_Tstat_lpFWE0.05-WB_c01_mask.nii  -m -s >> $output_dir/DMN/agg_FC_DMN.txt
 

### 2021-02-03 code/figures.R

@@ -334,6 +334,7 @@ mk_figtSNR <- function (final) {
   tsnr=read.table("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/check_tsnr/tsnr_in_ROIs.txt")
   tsnr=tsnr[tsnr$V1!="subj",]
   colnames(tsnr)=c("subj.ID_tp","roi","median","mean","sd")
+  tsnr <- tsnr[stringr::str_detect(tsnr$subj.ID_tp, "^ADI\\d{3}_(bl|fu|fu2)$"),]
   levels(tsnr$roi)=droplevels(tsnr$roi)
   levels(tsnr$roi)=c("NAcc", "PCC/precuneus")
   tsnr$mean=as.double(tsnr$mean)

### 2021-02-03 code/plot_glasbrain_networks.py

@@ -1,13 +1,10 @@
 # Plot DMN and Rew networks on glass brain
 
-# use python 3 
+# use python 3
 
 from nilearn import plotting
 stat_img="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD/Network_0-6-12/brain/PCC_gsr_z/network-age-sex_WB-c01cl/thresholded.nii"
-plotting.plot_glass_brain(stat_img, output_file="/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/DMN_thresholded_for_clusterFWE0.05.pdf", display_mode='ortho', colorbar=True, figure=None, axes=None, title=None, threshold='auto', annotate=True, black_bg=False, cmap='plasma', alpha=0.7, vmin=None, vmax=None, plot_abs=True, symmetric_cbar='auto', resampling_interpolation='continuous')
+plotting.plot_glass_brain(stat_img, output_file="/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/fig/DMN_thresholded_for_clusterFWE0.05.pdf", display_mode='ortho', colorbar=True, figure=None, axes=None, title=None, threshold='auto', annotate=True, black_bg=False, cmap='plasma', alpha=0.7, vmin=None, vmax=None, plot_abs=True, symmetric_cbar='auto', resampling_interpolation='continuous')
 
 stat_img="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD/Network_0-6-12/brain/Nacc_gsr_z/network-age-sex_WB-c01cl/thresholded.nii"
-plotting.plot_glass_brain(stat_img, output_file="/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/Rew_thresholded_for_clusterFWE0.05.pdf", display_mode='ortho', colorbar=True, figure=None, axes=None, title=None, threshold='auto', annotate=True, black_bg=False, cmap='plasma', alpha=0.7, vmin=None, vmax=None, plot_abs=True, symmetric_cbar='auto', resampling_interpolation='continuous')
-
-
-
[…]

### 2021-02-03 code/plot_glassbrain_models.py

@@ -9,7 +9,7 @@ import pandas as pd
 
 def main():
 
-    OUT_DIR = "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report"
+    OUT_DIR = "/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/fig"
     INFO_FILE = "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD/result_report.txt"
 
     # dataframe with filenames, path list, cluster threshold

### 2021-02-03 code/run_or_display.m

@@ -127,7 +127,7 @@ elseif param.PRESET == "bmi"
 elseif param.PRESET == "gt"
     param.MODEL = ["grouptime"];
     param.ROI_PREP = roi_prep([4, 6, 12, 14]); 
-    param.COVARIATES = [11,12];  
+    param.COVARIATES = [11, 12];  
     param.MASK = "brain";           
     param.EXCLFD = false;          
     param.INFERENCE_TYPE = ["cluster"];
@@ -174,14 +174,14 @@ elseif param.PRESET == "test"
     param.VIEW = true;
 elseif param.PRESET == "manual"
     % define ROI
-    param.ROI_PREP = roi_prep([4]); % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {"Nacc_cc_z","Nacc_gsr_z","PCC_cc_z","PCC_gsr_z","LH_cc_z","LH_gsr_z","MH_cc_z","MH_gsr_z"}
+    param.ROI_PREP = roi_prep([12,14]); % {roi_prep{[4, 6, 12, 14, 20, 22, 28, 30]}} or {"Nacc_cc_z","Nacc_gsr_z","PCC_cc_z","PCC_gsr_z","LH_cc_z","LH_gsr_z","MH_cc_z","MH_gsr_z"}
 
     % Model definition
     % All three models have unique options for covariate definition, the
     % association to a model is indicated by the tens digit (GroupTime_: 1_; 
     % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
-    param.MODEL = ["alltp"]; % ["grouptime","grouptime2tp"] % ["bmi","bmiIG","bmi2tp"] % ["fd","fdIG"] % ["alltp"] % ["singletp"]
-    param.COVARIATES = [42];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
+    param.MODEL = ["fd"]; % ["grouptime","grouptime2tp"] % ["bmi","bmiIG","bmi2tp"] % ["fd","fdIG"] % ["alltp"] % ["singletp"]
+    param.COVARIATES = [31];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
 
     % define masking and type of inference
     param.MASK = "brain";               % "brain"
@@ -189,7 +189,7 @@ elseif param.PRESET == "manual"
[…]

### 2021-02-03 code/tables.R

@@ -134,11 +134,11 @@ df9$covariates[1] <- "age, sex"
 
 ## merge tables 
 BMImodel <- rbind(df2,df1)
-FDmodel <- rbind(df8,df9)
 BMIFDmodel <- rbind(df3,df4,df5,df6,df7)
+FDmodel <- rbind(df8,df9)
 
-ModelList <- list(BMImodel,FDmodel,BMIFDmodel)
-names(ModelList) <- c("BMImodel","FDmodel","BMIFDmodel")
+ModelList <- list(BMImodel,BMIFDmodel,FDmodel)
+names(ModelList) <- c("BMImodel","BMIFDmodel","FDmodel")
   
 for (i in 1:length(ModelList)) {
   ModelList[[i]]$anat <- NA
@@ -190,37 +190,44 @@ for (i in 1:length(ModelList)) {
 label1 <- data.table::rbindlist(
   list(
     # for cluster of df1, df2
-    # avgBMI
+    # avgBMI - age,sex
     data.frame(ResultFCList[[2]][1])[1:3, ] %>% select(last_col()), # cl1
-    data.frame(ResultFCList[[2]][2])[1:3, ] %>% select(last_col()),
+    data.frame(ResultFCList[[2]][2])[1:2, ] %>% select(last_col()),
     data.frame(ResultFCList[[2]][3])[1:2, ] %>% select(last_col()),
     data.frame(ResultFCList[[2]][4])[1:2, ] %>% select(last_col()),
+    # avgBMI - age,sex, mFD
     data.frame(ResultFCList[[1]][1])[1:3, ] %>% select(last_col()), # cl1
     data.frame(ResultFCList[[1]][2])[1:2, ] %>% select(last_col()), # cl2
-    data.frame(ResultFCList[[1]][3])[1:3, ] %>% select(last_col()), # ...
+    data.frame(ResultFCList[[1]][3])[1:2, ] %>% select(last_col()), # ...
     data.frame(ResultFCList[[1]][4])[1:2, ] %>% select(last_col())
   ),
   use.names = FALSE
 )
-# FD
+# BMI FD
 label2 <- data.table::rbindlist(
   list(
-    data.frame(ResultFCList[[8]][1])[1:3, ] %>% select(last_col()),
-    data.frame(ResultFCList[[9]][1])[1:2, ] %>% select(last_col())
-    ), 
-  use.names = FALSE
-  )
-# BMI FD
-label3 <- data.table::rbindlist(
-  list(
[…]

### 2021-02-05 manuscript/draft_rsfmri.Rmd
changed supplement structure and manuscript labels

--- 2021-02-03 manuscript/draft_rsfmri.Rmd
@@
 <!-- Neurovault -->
-Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](https://neurovault.org/collections/9426/).
 
@@
 Functional decoding of these two activation patterns showed different top association terms.
-For the BMI change cluster (frontal	0.246; anterior insula	0.23; inferior frontal	0.194) were the top terms, while for the average mFD () were the meta-analytic activation maps most similar to this contract.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor cortex and
+premotor were the meta-analytic activation maps most similar to this contract.
 The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.

### 2021-02-05 manuscript/supplements_rsfmri.Rmd
changed supplement structure and manuscript labels

@@
 - Frauke Beyer* [1][2]
-- Kristin Prehn
-- Jürgen Ordemann 
-- Agnes Floel
-- A. Veronica Witte [1][2][3]
+- Evelyn Medawar [1]
+- Kristin Prehn [3][4]
+- Jürgen Ordemann [8][9]
+- Agnes Floel [5][6][7]
+- A. Veronica Witte [1][2][10]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
-  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
-  Day Clinic for Cognitive Neurology, University Clinic Leipzig
+  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Stroke Research, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Zentrum für Adipositas und Metabolische Chirurgie, Vivantes Klinikum Spandau, Berlin \newline 10 Day Clinic for Cognitive Neurology, University Clinic Leipzig
 output:
@@
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=FALSE}
+```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="40%", eval=FALSE}
+tsnr <- "../report/tsnr.pdf"
+knitr::include_graphics(tsnr)
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=TRUE}
 tsnr <- "../report/tsnr.pdf"

### 2021-02-05 manuscript/draft_rsfmri.Rmd
added supplementary tables

--- 2021-02-03 manuscript/draft_rsfmri.Rmd
@@
 - Frauke Beyer* [1][2]
-- Kristin Prehn
-- Jürgen Ordemann
-- Agnes Floel
-- A. Veronica Witte [1][2][3]
+- Evelyn Medawar [1]
+- Kristin Prehn [3][4]
+- Jürgen Ordemann [8][9]
+- Agnes Floel [5][6][7]
+- A. Veronica Witte [1][2][10]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
-  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
-  Day Clinic for Cognitive Neurology, University Clinic Leipzig
+  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Stroke Research, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Zentrum für Adipositas und Metabolische Chirurgie, Vivantes Klinikum Spandau, Berlin \newline 10 Day Clinic for Cognitive Neurology, University Clinic Leipzig
 output:
@@
     toc: yes
-  pdf_document:
-    toc: yes
-    extra_dependencies: "subfig"
-    fig_caption: yes
-    number_sections: yes
 bibliography: "bibliography.bib"
@@
 - \linenumbers
+- \usepackage{setspace}
+- \onehalfspacing
+indent: true
 ---
@@
 ```{r "To do", include=FALSE}
-# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
+# [x] Upload auf Neurovolt, cf. cobidas reporting guidelines
 # [x] Swap mFD with logmFD ??
@@
 RYBG-patients frequently report altered smell and taste perception, i.e.
-increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items [@Mulla_2017, @Berthoud_2011].
-These changes in appetitive and reward signaling likely relate to altered gut-brain communication, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin [ @Brutman_2019].
+increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items [@Mulla_2017; @Berthoud_2011].
+These changes in appetitive and reward signaling likely relate to altered gut-brain communication, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin [@Brutman_2019].
 
@@
 BA 39) [@Raichle_2015].
-Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015, @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
+Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
 Despite the large heterogenity and low median power, these findings might be summarized as a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
@@
 Increased activation of the DMN in a food-viewing paradigm, i.e.
-insufficient down-regulation of DMN, has been shown to correlate with appetite measures after an overnight fast [@Tregellas_2011, @Wijngaarden_2015], an effect that could be partially reversed by an exercise intervention [@McFadden_2013].
+insufficient down-regulation of DMN, has been shown to correlate with appetite measures after an overnight fast [@Tregellas_2011; @Wijngaarden_2015], an effect that could be partially reversed by an exercise intervention [@McFadden_2013].
 In a bariatric surgery study, [@Li_2018] reported a reduction of local FC in mid-line DMN regions in surgical patients compared to an obese control group.
@@
 These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
-Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced reward region activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020;].
+Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced reward region activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
 Food cue hyperactivation has been identified as a main neural vulnerability factor predicting future weight gain [@Stice_2015].
@@
 Possible, the reconfiguration of the fronto-striatal brain networks might be related to altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013, but @Zoon_2018] .
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] .
 Thus, while there is some evidence hinting to a role for DMN and reward network connectivity in altered gut-brain communication after bariatric surgery, the existing evidence is inconclusive.
@@
 
-\newpage
 # Methods
-
 ```{r "Load data", include=FALSE}
@@
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment.
-Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data availabile.
+Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data available.
 In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
@@
 MRI scanning was done after further psychological tests (for details see @Prehn_2020).
-
 ```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
@@
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
+Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histograms (see supplements).
 We reported the mFD-FC correlation per preprocessing scheme.
@@
 *In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
-The mask will include all voxels surviving FWE-correction of p<0.05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"*
+The mask will include all voxels surviving FWE-correction of .05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"*
 
@@
 We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mFD and change in both measures, along with baseline age and sex (Model 2c).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
@@
 The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
-We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
-Significant clusters are defined as family-wise error - corrected p<0.05.
+We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p < 0.01).
+Significant clusters are defined as family-wise error - corrected p < 0.05.
 -->
@@
 
-\newpage
 # Results
@@
 ## Quality control
+
 We performed different quality controls for the rsfMRI data.
@@
 We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
-
 ```{r mFDFC, echo=FALSE}
@@
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
-
 ```{r tableQcfc, echo=FALSE}
@@
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
-
 ```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
@@
 ```
-
 ```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
@@
 
-### DMN/reward network with whole brain mask (preregistered analysis)
+### Whole-brain effects of bariatric surgery on DMN/reward network connectivity (preregistered analysis)
 
@@
 Similarly, there were no significant clusters when including logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+<!-- Neurovault -->Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
 
-<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
-
-## Interaction effect group by time on aggregated FC (preregistered analysis)
+### Effects of bariatric surgery on aggregated DMN/reward network FC (preregistered analysis)
+#### Model 1a and 1b (group-by-time interaction)
 ```{r load agg FC data and prepare, echo = FALSE}
@@
 ```
-There was no significant time by group interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models)
+There was no significant time by group interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless whether we adjusted for mFD.
 
@@
 ### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
-final_FC[final_FC$mean_DMN_conn<0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
+final_FC[final_FC$mean_DMN_conn< 0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
 #ADI003_fu2 & ADI041_fu -> do not have particularly high head motion
@@
 ```
-
-There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC)) and supplements for a detailed summary of the models).
-
 ```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
@@
 ```
-
 ```{r "qa check DMN model age,sex", eval = FALSE}
@@
 ```
-
 ```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
@@
 ```
-
 ```{r "calc DMN two tp, age and sex and logmFD", results = 'asis', echo = FALSE}
@@
 ```
-
 ```{r, "qa check DMN model age,sex, logmFD", eval=FALSE}
@@
 ```
-
 ```{r "DMN2tp results age sex logmFD", caption="DMN connectivity compared bl to fu, adjusted for age, sex and logmFD", results="asis", eval=FALSE}
@@
 ```
-
 ```{r "calc DMN three tp", eval = TRUE, echo = FALSE}
@@
 ```
-
 ```{r "DMN3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
@@
 ```
-
 ```{r "calc rew two tp, age, sex", eval = TRUE, echo = FALSE}
@@
 ```
-
 ```{r, "qa check Rew model age,sex", eval=FALSE}
@@
 ```
-
 ```{r "Rew2tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
@@
 ```
-
 ```{r "calc rew two tp adjusted", eval = TRUE, echo = FALSE}
@@
 ```
-
 ```{r "qa check Rew model age, sex, logmFD", eval = FALSE}
@@
 ```
-
 ```{r "Rew2tpadj results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
@@
 ```
-
 ```{r "calc rew three tp", eval = TRUE, echo = FALSE}
@@
 ```
-
 ```{r "Rew3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
@@
 ## Exploratory Analyses
-### Effect of mean and change in BMI
-#### Whole brain analysis
-##### BMI-only model
+
+### Whole-brain effects of mean and change in BMI on DMN/reward network connectivity
 Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
-<!-- ### DMN -->
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE cluster p<0.05).
-When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \@ref(tab:tableBMImodel).
+
+#### DMN   
+##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
+
+Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCavgBMIdeact)).
+When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \@ref(tab:tableBMImodel)).
 ```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and logmFD"), fig.hold=TRUE, out.width = "50%"}
@@
 ```
+##### Model 2c (average and change in BMI and mFD)
 
-NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
-Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
-
-##### BMI and FD model
-To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
+To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
 <!-- ### DMN -->
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+<!-- cc -->Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components.
+Results are detailed in Table \@ref(tab:tableBMIFDmodel) and Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2).
+```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
 Test statistics only differed slightly from the effect found when using logmFD as nuisance covariate.
 <!-- gsr -->As before, the effect did not survive when using GSR-corrected data.
+Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
 
-<!-- ### Reward -->
-<!-- cc -->
-For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
-supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
+#### Reward network
+##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
+NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for mFD.
+
+##### Model 2c (average and change in BMI and mFD)
+<!-- cc -->We found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
+supplementary motor area (see Table \@ref(tab:tableBMIFDmodel) and Figure \@ref(fig:BMIFDagesexNACCcgnBMIdeact)).
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.03$), also when evaluating the GSR corrected data ($p = 0.041$).
 ```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
@@
 Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
-motor cortex, was significantly positively related to average logmFD.
+motor cortex, was significantly positively related to average logmFD (see  Figure \@ref(fig:BMIFDagesexNACCavgFDact)).
 This result remained similar when using GSR corrected data.
@@
 
-```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
-knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
-```
-
-#### Aggregated FC analysis
+### Effects of average and change in BMI on aggregated DMN/reward network FC
+##### DMN
+##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
 ```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 ```
-Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD.
-```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-```
+Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD (see supplementary material for a detailed description of the models).
 
+##### Model 2c (average and change in BMI and mFD)
 ```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
@@
 ```
-
 The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in logmFD with DMN FC.
 
-#### Reward network, mean and within variability in BMI and logmFD
+##### Reward network
+```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+```
 ```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
@@
 ```
-
-As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and reward network connectivity in models adjusting for age, sex and logmFD.
+As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and reward network connectivity in models 2a, 2b or 2c adjusting for age, sex and logmFD.
 
 ### Effect of FD (exploratory analysis)
+
 #### Whole-brain analysis
+
 To investigate the unique contributions of the average value of and the longitudinal change in logmFD, we constructed two additional models.
@@
 <!-- #### Reward -->
-As shown in Table \@ref(tab:tableFDmodel), there was a significant association between logmFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas .
+As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), there was a significant positive association between mFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas .
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
-<!-- ### DMN -->
-We did not find a any clusters for functional connectivity with the PCC.
-
 ```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
@@
 ```
-
-<!-- ## Effect of BMI and FD (exploratory analysis) -->
-
-<!-- Average BMI and change in BMI remain covariates of interest.
-While initially used as nuisance covariate, here, logmFD is split into average logmFD and change in logmFD and used as.
-Age and sex remain nuisance covariates.
--->
-
-<!-- <!-- ### Reward -->
-<!-- cc There is a significant effect of change in BMI on functional connectivity between NAcc and one cluster in the posterior-medial frontal region.-->
-<!-- Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction.
-Furthermore, the functional connectivity between the NAcc and one cluster was significantly related to average logmFD.
--->
-<!-- <!-- gsr When running the model with images corrected with GSR, the associations between connectivity and change in BMI did not survive statistical thresholds. -->
-<!-- On the other hand, the relationship with average logmFD was confirmed by analyses after GSR but cluster size shrank. -->
+<!-- ### DMN -->
+We did not find a any clusters for functional connectivity with the PCC.
 
@@
 ```
-
 ```{r tableBMIFDmodel}
@@
 ```
-
 ```{r tableFDmodel}
@@
 ```
-
 <!-- cc The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
@@
 <!-- gsr As before, the effect did not survive when thresholding if the images had been GSR corrected.-->
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 
@@
 # Declaration of interest
+
 # Funding
+
 # Acknowledgments

### 2021-02-05 manuscript/supplements_rsfmri.Rmd
added supplementary tables

--- 2021-02-02 manuscript/supplements_rsfmri.Rmd
@@
     number_sections: yes
-    toc: yes
+    toc: no
     extra_dependencies: "subfig"
@@
 ```
+\newpage
 
+# Methods
+## Sample
 ```{r "tableSample", echo=FALSE}
@@
     linesep = "",
-    caption = "BMI and mean FD of participants acquired in each condition and time point")
+    caption = "BMI and mFD of participants acquired in each condition and time point")
 ```
 
+## Design
 ```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap = c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
@@
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=FALSE}
-tsnr <- "../report/tsnr.pdf"
+# Results
+## Quality Control
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=TRUE}
+tsnr <- "../report/tsnr.jpeg"
 knitr::include_graphics(tsnr)
@@
 
-```{r qa check DMN model age,sex, echo=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
+```{r qa check DMN model age,sex, eval=FALSE}
+diag=diagnostics.plot(R0)
+randefplot=ranef.diagn.plot(R0)
 
@@
 
-```{r qa check DMN model age,sex, mFD, echo=FALSE}
+```{r qa check DMN model age,sex, mFD, eval=FALSE}
 diagnostics.plot(R0)
@@
 ```
-```{r qa check Rew model age,sex, echo=FALSE}
+```{r qa check Rew model age,sex, eval=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r qa check Rew model age,sex,mFD, echo=FALSE}
+```{r qa check Rew model age,sex,mFD,eval=FALSE}
 diagnostics.plot(R0)
@@
 ```
+
+## Effects of average and change BMI on FC
+### DMN
+#### Model 2a and 2b (adjusting for mFD)
+```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
+
+R11 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
+R12a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
+anvmeanBMI_adjmFD=anova(R11,R12a)
+```
+
+```{r DMN_BMIFD results, caption="Model 2a and 2b for aggregated DMN FC", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R1,R11), custom.coef.names=c("Intercept", "average BMI", "change BMI", "Age", "Sex","logmFD"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+#### Model 2c (adjusting for average and change in mFD)
+```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvmeanFD=anova(R1,R03)
+anvwithinFD=anova(R1,R03)
+
+p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+```
+
+```{r DMN_BMI_splitFD results, caption="Model 2c for aggregated DMN FC", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+\newpage
+```{r, include = FALSE}
+library(dplyr)
+DIR_ANALYSIS <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
+df <- file.path(DIR_ANALYSIS,"noExclFD/result_report.txt")
+AnatomyResults <- read.table(df, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)
+
+txtList <- list()
+
+for (r in 1:nrow(AnatomyResults)) {
+
+  file = AnatomyResults[r,"dir"]
+  txt_cols <- c("VoxelCount","equals","PercClusterVolumeAssignedTo","in","Hem","Region","PercentOfArea","V8","V9","V10","V11","V12","V13")
+  txt <- read.table(trimws(f),sep="\t",fill=TRUE,col.names=txt_cols) # paste0("V", 1:13)
+
+  txt <- txt[sort(c(which(grepl("Cluster", txt$VoxelCount)),which(grepl("voxel", txt$equals)))),]
+  txt <- txt[,c("VoxelCount","PercClusterVolumeAssignedTo", "Hem","Region","PercentOfArea")]
+
+  idxClusterStart <- which(grepl("Cluster", txt$VoxelCount))
+
+  #txt$VoxelCount <- as.character(txt$VoxelCount)
+  txt <- cbind(descr = NA, txt)
+  txt$descr[idxClusterStart] <- stringr::str_remove(string = as.character(txt$VoxelCount[idxClusterStart]), pattern = "\\:[^:]*$")
+  txt$VoxelCount[idxClusterStart] <- NA
+
+  # attach model with  cluster list to list of models
+  name_model <- AnatomyResults$model[r]
+  txtList[[name_model]] <- txt
+}
+
+txt_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\n in cluster","\\% of cluster volume\nassigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))
+
+TableList <- list()
+for (i in 1:length(txtList)){
+  TableList[[i]] <-
+    txtList[[i]] %>%
+    knitr::kable(
+      escape = FALSE,     # use font spec of latex with kableExtra
+      col.names = txt_colname,
+      row.names = FALSE,
+      format = "latex",
+      booktabs = T,
+      linesep = "",      # disable "\\addlinespace" at every 5th line
+      #label = name_model[i],
+      caption = sprintf("Anatomical labelling for %s",name_model[i]),
+    ) %>%
+    kableExtra::kable_styling(latex_options = c('scale_down')) #%>% 
+    #kableExtra::landscape()
+}
+```
+
+### Anatomical Labelling for models
+
+```{r, echo=FALSE}
+
+AnatomyResults$model <- read.table(df, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)[1]
+l <- stringr:: str_split(AnatomyResults$model, '_')
+d <- data.frame(matrix(unlist(l), nrow=length(l), byrow=T))
+#stringr::str_replace(as.character(d[,2]),"^\\(PCC|NACC)cc$"," (cc)") # , "\\(PCC|NACC)(gsr)" = "(gsr)"
+
+TableList[[1]]
+TableList[[2]]
+TableList[[3]]
+TableList[[4]]
+TableList[[5]]
+TableList[[6]]
+TableList[[7]]
+TableList[[8]] 
+TableList[[9]]
+```

### 2021-02-05 code/extract_agg_FC.sh

@@ -30,9 +30,9 @@ in="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/noExclFD
 
 fslmaths $in -thr 1.30103 -bin "$output_dir/Rew/Rew_swe_clusternorm_Tstat_lpFWE0.05-WB_c01_mask.nii"
 
-fslmerge -t "$output_dir/Rew/Rew_merged.nii.gz" `ls /data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/Individual_scans/Nacc_gsr_z/*`
+fslmerge -t "$output_dir/Rew/Rew_merged.nii.gz" `ls /data/pt_02161/Results/Project2_resting_state/connectivity/calc_DMN_reward_seed_connectivity/*/Nacc_gsr_z/nacc_cc_gsr_seed_correlation_z_trans.nii`
 
-ls /data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/Individual_scans/Nacc_gsr_z/* >> $output_dir/Rew/agg_FC_Rew_ID.txt
+ls /data/pt_02161/Results/Project2_resting_state/connectivity/calc_DMN_reward_seed_connectivity/*/Nacc_gsr_z/nacc_cc_gsr_seed_correlation_z_trans.nii >> $output_dir/Rew/agg_FC_Rew_ID.txt
 
 fslstats -t $output_dir/Rew/Rew_merged.nii.gz -k $output_dir/Rew/Rew_swe_clusternorm_Tstat_lpFWE0.05-WB_c01_mask.nii  -m -s >> $output_dir/Rew/agg_FC_Rew.txt
 

### 2021-02-05 code/process.R

@@ -128,8 +128,8 @@ agg_FC_DMN_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connecti
 agg_FC_Rew_ID=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt")
 agg_FC_Rew=read.table("/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt")
 for (i in c(1:nrow(agg_FC_DMN_ID))){
-  agg_FC_Rew[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][10],'_nacc')[[1]][1]
-  agg_FC_DMN[i,"subj.ID_tp"]=strsplit(base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][10],'_pcc')[[1]][1]
+  agg_FC_Rew[i,"subj.ID_tp"]=base::strsplit(as.character(agg_FC_Rew_ID[i,"V1"]),'/')[[1]][8]
+  agg_FC_DMN[i,"subj.ID_tp"]=base::strsplit(as.character(agg_FC_DMN_ID[i,"V1"]),'/')[[1]][8]
 }
 
 colnames(agg_FC_Rew)=c("mean_Rew_conn","sd_Rew_conn","subj.ID_tp")

### 2021-02-05 code/tables.R

@@ -259,7 +259,7 @@ ModelList[[3]]$Hem <- LabelList[[3]]$Hem
 fn1 <- "Hem, hemisphere; L, left; R, right; FWE-corr., family-wise error corrected; MNI (Montreal Neurological Institute) coordinates of primary peak location: X sagittal; Y, coronal; Z, axial; cc, preprocessing with AROMA-ICA + CompCor; gsr, preprocessing with AROMA-ICA + CompCor + GSR."
 fn3 <- "To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000 samples."
 fn4 <- "Connectivity with maximum three voxels that mark local maxima within the respective custer; more detailed description of anatomical regions that are assigned to overall clusters and and corresponding probability in Supplementary."
-title_vec <- c("BMI", "FD", "BMI-FD") 
+title_vec <- c("BMI", "BMI-FD", "FD") 
 
 tabnames <- c("tableBMImodel","tableBMIFDmodel","tableFDmodel")
 TableList <- list()

### 2021-02-06 manuscript/draft_rsfmri.Rmd

@@
 <!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
-The resulting factorial design <!-- check: flexible factorial?
---> contained one regressor for each time point per group.
+The resulting factorial design contained one regressor for each time point per group.
 
@@
 `R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
-`R0 = lmer(agg FC ~ timepoint + group +  + age + (1|subj))`
+`R0 = lmer(agg FC ~ timepoint + group +  age + sex + (1|subj))`
 and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
@@
 Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average logmFD and variability in logmFD.
+Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and change in mFD.
+
+### Functinal decoding
+In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
+We uploaded the contrast images for change BMI and average mFD on [Neurovault](), and applied the decoding classifier.
+This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
+We reported the three top terms for both contrasts.
 
@@
 
-####  Model 1b (with adjustment for age and sex, meanFD)
-Similarly, there were no significant clusters when including logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
-<!-- Neurovault -->Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
+####  Model 1b (with adjustment for age, sex and mFD)
+Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 
-### Effects of bariatric surgery on aggregated DMN/reward network FC (preregistered analysis)
-#### Model 1a and 1b (group-by-time interaction)
+<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
+
+<!-- Neurovault -->
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](https://neurovault.org/collections/9426/).
+
+## Interaction effect group by time on aggregated FC (preregistered analysis)
 ```{r load agg FC data and prepare, echo = FALSE}
@@
 ```
-```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
+
+There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC)) and supplements for a detailed summary of the models).
+
+```{r "calc DMN two tp age and sex" ,results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
@@
 ```
-```{r "qa check DMN model age,sex", eval = FALSE}
+
+```{r "qa check DMN model age sex", eval = FALSE}
 diagnostics.plot(R0)
@@
 ```
-```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
+
+```{r "DMN2tp results age sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 ```
-```{r, "qa check Rew model age,sex", eval=FALSE}
+
+```{r "qa check Rew model age sex", eval=FALSE}
 diagnostics.plot(R0)
@@
 
+Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
+Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
+
+##### BMI and mFD model
 To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
 <!-- ### DMN -->
-<!-- cc -->Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components.
-Results are detailed in Table \@ref(tab:tableBMIFDmodel) and Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2).
+<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting mFD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
+
 ```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
@@
 
-##### Model 2c (average and change in BMI and mFD)
-<!-- cc -->We found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
-supplementary motor area (see Table \@ref(tab:tableBMIFDmodel) and Figure \@ref(fig:BMIFDagesexNACCcgnBMIdeact)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.03$), also when evaluating the GSR corrected data ($p = 0.041$).
-```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+<!-- ### Reward -->
+<!-- cc -->
+For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region (AAL).
+The peak voxel was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
+```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 ```
+###### Functional decoding
+Functional decoding of these two activation patterns showed different top association terms.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor cortex and
+premotor were the meta-analytic activation maps most similar to this contract.
+The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
 
-### Effects of average and change in BMI on aggregated DMN/reward network FC
-##### DMN
-##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
+#### Aggregated FC analysis
+##### DMN (BMI and mFD model)
 ```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 
-##### Reward network
-```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-```
+The association also remained significant when we split mFD into average and change in mFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in mFD with DMN FC.
+
+##### Reward network (BMI and mFD model)
 ```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
@@
 
-### Effect of FD (exploratory analysis)
-
+### Effect of mFD (exploratory analysis)
 #### Whole-brain analysis
@@
 
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average mFD and change mFD will be are published on NeuroVault.
+
 # Discussion
+#### Summary
 In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
-Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) in group analysis.
-We found no interaction effect of group and timepoint or main effects on whole-brain FC of the PCC and Nacc, which we used to define the resting state networks DMN and reward network, respectively.
-Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) on the group level.
+We found no interaction effect of group and time point on whole-brain FC of the PCC and NAcc.
+We also did not observe main effects of group and time point on the FC of these DMN and reward network hubs, respectively  Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
 In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme.
@@
 
-Overall, we could not confirm our preregistered hypotheses of an interaction effect between group and time point on DMN and reward network connectivity.
-Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in lien with improved processing of visceral and bodily signals  after surgery [@Li_2018; @McFadden_2013].
-Further, we expected FC between reward network regions to decrease, which would support a reduction of reward hypersensitivity and hedonic motivation to eat following bariatric surgery [@Scholtz_2014; ].
-Yet, our analyses pointed to a possible effect of BMI change on the interplay of the reward network and a posterior frontal brain region.
-Here, against our expectation, we found stronger BMI decrease to predict increases in connectivity of the Nacc and posterior-medial frontal brain.
-This pattern could be interpreted as enhanced connectivity between regions governing salience attribution and reward processing, as expected in obesity.
-However, it might also indicate a more close monitoring of reward signaling, which would be in line with lower desire to eat high-calorie food items after surgery.
+In this study, we could not confirm these preregistered hypotheses and found no interaction effect between group and time point on DMN and reward network connectivity.   
+Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
+Further, we expected FC between reward network regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in reward network regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014;   @Cerit_2019].
+Yet, in an exporatory analysis,  stronger BMI decrease predicted higher connectivity of the Nacc and a cluster in a posterior-medial frontal brain region.
+Based on the Harvard-Oxford atlas and NeuroSynth decoding this region might be part of the salience network and involved in action preparation.
+AThe enhanced connectivity between the Nacc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
+Yet, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
+Previously,  a decrease in local (regional homogenity and frequency of low-amplitude oscillations) and global connectivity (degree centrality (DC)) measures in a similar region of the left SMA was reported after glucose administration  [@AlZubaidi_2018].
+Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
+Importantly, a global decrease in connectivity (DC) might go along with selectively increased connectivity, i.e.
+a shift from integration towards segregation and more specialized processing.
+However, this interpretation is highly speculative, and more studies are needed to corroborate these findings.
+
 Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity.
 Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2019], and therefore conducted careful analyses of the impact of head motion on our results.
-To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average logmFD and change in logmFD and introducing two instead of one regressors into the model.
-This result may be due to the presence of multi-collinearity between change in logmFD and BMI which appeared more pronounced in the split model, and lead to unreliable estimations of effects and standard errors.
-Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and logmFD the effect of change in BMI could be singled out.
-In favour of this argument speaks the survival of the cluster when using global signal regressed data.
-Further, average logmFD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
-Thus, confounding of a positive effect of head motion on posterior-frontal connectivity and the inversely directed BMI effect seems unlikely.
-
+To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
+This result may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
+Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
+This argument is supported by the  survival of the cluster when using global signal regressed data, and the distinct results of the decoding analysis.
+Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
+Thus, confounding of the negative BMI change effect and the positive average head motion effect  on posterior-frontal connectivity  seems unlikely.
 Head motion also played a role in the association of higher BMI and reduced DMN connectivity.
 While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN connectivity was negatively associated with BMI in both denoising schemes.
-Thereby, this results echos a previous finding from our group, and may be interpreted as accelerated DMN decline in relation to higher BMI.
-Yet, similar doubts regarding the complete removal of motion confounding remain in this study.
+Thereby, this results echos a previous finding from our group, and may be interpreted as  accelerated age-related decline of the DMN  in relation to the cardiometabolic risk factor BMI.
+Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding also remain in this study [@Savalia_2017].
+
 The major strength of our study is prospective intervention controlled design.
-We compared BS patients to an obese control group who were scanned after the same time intervals and  equivalent in BMI, treatment history and  treatment recommendation to our intervention group [@Thiese_2014].
+We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals and , [@Thiese_2014].
 Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
-In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended method to deal with longitudinal repeated measures @Guillaume_2014.
-
+In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
 Opposed to flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
 Limitations of our study include the low number of patients who participated in all three time points.
-In total, only 19 participants contributed to the estimation of the longitudinal effects over three time points.
+In total, only X participants contributed to the estimation of the longitudinal effects with at least two  time points.
 Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI.
 While this sample size is comparable to previous resting state fMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
+
 We used seed-based connectivity to derive large-scale brain networks.
-While this approach yielded correct DMN and reward network maps, it is a univariate approach not taking into account the inter-relatedness of subnetworks or other regions.
-Furthermore, we did not explicitly investigate the interaction of these networks with attention, cognitive control or sensorimotor networks, which has been previously proposed to be altered by bariatric surgery.
-We did not control for hunger or satiety in our design.
-A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
-After the intervention, functional connectivity was higher in bariatric patients.
-It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
-Prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016].
-putamine, insula
+While this approach yielded reasonable DMN and reward network maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
+Furthermore, we did not explicitly investigate the connectivity of the DMN and reward network with other large-scale brain networks attention, e.g.
+cognitive control or sensorimotor networks, which has been previously proposed to be altered by obesity and bariatric surgery [@Lips_2014; @Olivo_2016; @Cerit_2019].
+
+We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized?
+breakfast following an overnight fast.
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2015; @Lepping_2015, @Wiemerslage_2016; Ochner_2012_prepostprandial], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow separate analyses for the sexes.

### 2021-02-06 manuscript/supplements_rsfmri.Rmd

@@
 
-# Results
-## Quality Control
+```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="40%", eval=FALSE}
+tsnr <- "../report/tsnr.pdf"
+knitr::include_graphics(tsnr)
+```
+
 ```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=TRUE}
-tsnr <- "../report/tsnr.jpeg"
+tsnr <- "../report/tsnr.pdf"
 knitr::include_graphics(tsnr)

### 2021-02-07 manuscript/draft_rsfmri.Rmd
merged changes + better supplemetary tables

--- 2021-02-05 manuscript/draft_rsfmri.Rmd
@@
 # [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
+# [ ] Swap mFD with logmFD ??
 ```
@@
 
-```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
+```{r figBMIdescr, fig.align='center', out.width="50%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
@@
 In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
-We calculated mean framewise displacement (mFD) from the outputs of the motion correction procedure according to @Power_2012.
 For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
@@
 As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020].
-Thus, we did not exclude any participants based on mFD estimates beyond those who failed the anatomical preprocessing.
+Thus, we did not exclude any participants based on mean FD estimates beyond those who failed the anatomical preprocessing.
 We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mFD.
+In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD.
 ```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
@@
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we correlated mFD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
+Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}.
 We reported the mFD-FC correlation per preprocessing scheme.
@@
 We did not perform the analysis as preregistered (i.e.
-without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mFD (Model 1b).
+without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
 <!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
@@
 This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for the logarithm of mFD (Model 2b).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and the logarithm of mFD and variability in both measures, along with baseline age and sex (Model 3).
+We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
 <!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen.
--->For a better understanding of the unique contribution of average and longitudinal change mFD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject mFD + within-subject mFD$ with nuisance covariates age and sex.
+-->For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
 
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas.
-<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
 We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01).
@@
 Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up.
-In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and logarithm of mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
+In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
 We performed model comparison between R1 and R0 models
 `R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
-`R0 = lmer(agg FC ~ timepoint + group +  age + sex + (1|subj))`
+`R0 = lmer(agg FC ~ timepoint + group +  + age + (1|subj))`
 and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
@@
 Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and change in mFD.
-
-### Functinal decoding
-In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
-We uploaded the contrast images for change BMI and average mFD on [Neurovault](), and applied the decoding classifier.
-This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
-We reported the three top terms for both contrasts.
+Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.
 
@@
 lay <- rbind(c(1,1,2,2),
+             c(1,1,2,2),
+             c(NA,3,3,NA),
              c(NA,3,3,NA))
 gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
+
 ```
@@
 
-####  Model 1b (with adjustment for age, sex and mFD)
+####  Model 1b (with adjustment for age and sex, meanFD)
 Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
@@
 <!-- Neurovault -->
-Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](https://neurovault.org/collections/9426/).
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
 
@@
 
-```{r "calc DMN two tp age and sex" ,results = 'asis', echo = FALSE}
+```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
@@
 
-```{r "qa check DMN model age sex", eval = FALSE}
+```{r "qa check DMN model age,sex", eval = FALSE}
 diagnostics.plot(R0)
@@
 
-```{r "DMN2tp results age sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
+```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r "calc DMN two tp age and sex and mFD", results = 'asis', echo = FALSE}
+```{r "calc DMN two tp, age and sex and mFD", results = 'asis', echo = FALSE}
 #### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
@@
 
-```{r "qa check DMN model age sex mFD", eval=FALSE}
+```{r, "qa check DMN model age,sex, mFD", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-```{r "qa check Rew model age sex", eval=FALSE}
+```{r, "qa check Rew model age,sex", eval=FALSE}
 diagnostics.plot(R0)
@@
 
-Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
+NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
 Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
 
-##### BMI and mFD model
+##### BMI and FD model
 To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
 <!-- ### DMN -->
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting mFD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
+<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
 Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
@@
 <!-- cc -->
-For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region (AAL).
-The peak voxel was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
+For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e.
+supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)).
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
@@
 ```
-Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
-motor cortex, was significantly positively related to average mFD.
-This result remained similar when using GSR corrected data.
 ```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
@@
 ```
-###### Functional decoding
-Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor cortex and
-premotor were the meta-analytic activation maps most similar to this contract.
-The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
+```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", out.width = "50%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
+Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
+motor cortex, was significantly positively related to average mFD.
+This result remained similar when using GSR corrected data.
 
 #### Aggregated FC analysis
-##### DMN (BMI and mFD model)
 ```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 
-##### Reward network (BMI and mFD model)
+#### Reward network, mean and within variability in BMI and logmFD
 ```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
@@
 
-### Effect of mFD (exploratory analysis)
+### Effect of FD (exploratory analysis)
 #### Whole-brain analysis
@@
 <!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average mFD and change mFD will be are published on NeuroVault.
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.
 
 # Discussion
-#### Summary
 In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
-Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) on the group level.
-We found no interaction effect of group and time point on whole-brain FC of the PCC and NAcc.
-We also did not observe main effects of group and time point on the FC of these DMN and reward network hubs, respectively  Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) in group analysis.
+We found no interaction effect of group and timepoint or main effects on whole-brain FC of the PCC and Nacc, which we used to define the resting state networks DMN and reward network, respectively.
+Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
 In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme.
@@
 
-In this study, we could not confirm these preregistered hypotheses and found no interaction effect between group and time point on DMN and reward network connectivity.   
-Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
-Further, we expected FC between reward network regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in reward network regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014;   @Cerit_2019].
-Yet, in an exporatory analysis,  stronger BMI decrease predicted higher connectivity of the Nacc and a cluster in a posterior-medial frontal brain region.
-Based on the Harvard-Oxford atlas and NeuroSynth decoding this region might be part of the salience network and involved in action preparation.
-AThe enhanced connectivity between the Nacc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
-Yet, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
-Previously,  a decrease in local (regional homogenity and frequency of low-amplitude oscillations) and global connectivity (degree centrality (DC)) measures in a similar region of the left SMA was reported after glucose administration  [@AlZubaidi_2018].
-Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
-Importantly, a global decrease in connectivity (DC) might go along with selectively increased connectivity, i.e.
-a shift from integration towards segregation and more specialized processing.
-However, this interpretation is highly speculative, and more studies are needed to corroborate these findings.
-
+Overall, we could not confirm our preregistered hypotheses of an interaction effect between group and time point on DMN and reward network connectivity.
+Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in lien with improved processing of visceral and bodily signals  after surgery [@Li_2018; @McFadden_2013].
+Further, we expected FC between reward network regions to decrease, which would support a reduction of reward hypersensitivity and hedonic motivation to eat following bariatric surgery [@Scholtz_2014; ].
+Yet, our analyses pointed to a possible effect of BMI change on the interplay of the reward network and a posterior frontal brain region.
+Here, against our expectation, we found stronger BMI decrease to predict increases in connectivity of the Nacc and posterior-medial frontal brain.
+This pattern could be interpreted as enhanced connectivity between regions governing salience attribution and reward processing, as expected in obesity.
+However, it might also indicate a more close monitoring of reward signaling, which would be in line with lower desire to eat high-calorie food items after surgery.
 Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity.
@@
 To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
-This result may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
+This result may be due to the presence of multi-collinearity between change in FD and BMI which appeared more pronounced in the split model, and lead to unreliable estimations of effects and standard errors.
 Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
-This argument is supported by the  survival of the cluster when using global signal regressed data, and the distinct results of the decoding analysis.
+In favour of this argument speaks the survival of the cluster when using global signal regressed data.
 Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
-Thus, confounding of the negative BMI change effect and the positive average head motion effect  on posterior-frontal connectivity  seems unlikely.
+Thus, confounding of a positive effect of head motion on posterior-frontal connectivity and the inversely directed BMI effect seems unlikely.
+
 Head motion also played a role in the association of higher BMI and reduced DMN connectivity.
 While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN connectivity was negatively associated with BMI in both denoising schemes.
-Thereby, this results echos a previous finding from our group, and may be interpreted as  accelerated age-related decline of the DMN  in relation to the cardiometabolic risk factor BMI.
-Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding also remain in this study [@Savalia_2017].
+Thereby, this results echos a previous finding from our group, and may be interpreted as accelerated DMN decline in relation to higher BMI.
+Yet, similar doubts regarding the complete removal of motion confounding remain in this study.
+The major strength of our study is prospective intervention controlled design.
+We compared BS patients to an obese control group who were scanned after the same time intervals and  equivalent in BMI, treatment history and  treatment recommendation to our intervention group [@Thiese_2014].
+Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
+In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended method to deal with longitudinal repeated measures @Guillaume_2014.
 
-The major strength of our study is prospective intervention controlled design.
-We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals and , [@Thiese_2014].
-Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
-In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
 Opposed to flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
 Limitations of our study include the low number of patients who participated in all three time points.
-In total, only X participants contributed to the estimation of the longitudinal effects with at least two  time points.
+In total, only 19 participants contributed to the estimation of the longitudinal effects over three time points.
 Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI.
 While this sample size is comparable to previous resting state fMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
-
 We used seed-based connectivity to derive large-scale brain networks.
-While this approach yielded reasonable DMN and reward network maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
-Furthermore, we did not explicitly investigate the connectivity of the DMN and reward network with other large-scale brain networks attention, e.g.
-cognitive control or sensorimotor networks, which has been previously proposed to be altered by obesity and bariatric surgery [@Lips_2014; @Olivo_2016; @Cerit_2019].
-
-We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized?
-breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2015; @Lepping_2015, @Wiemerslage_2016; Ochner_2012_prepostprandial], and might thus have confounded our results [@Li_2019].
+While this approach yielded correct DMN and reward network maps, it is a univariate approach not taking into account the inter-relatedness of subnetworks or other regions.
+Furthermore, we did not explicitly investigate the interaction of these networks with attention, cognitive control or sensorimotor networks, which has been previously proposed to be altered by bariatric surgery.
+We did not control for hunger or satiety in our design.
+A previous repeated measurement design with patients undergoing bariatric surgery and behavioral dietary intervention manipulating satiety showed a differential effect of satiety between intervention groups on the resting-state functional connectivity seeded in the precuneus.
+After the intervention, functional connectivity was higher in bariatric patients.
+It decreased after a meal while dietary patients showed an increase [@Lepping_2015].
+Prandrial state is a relevant factor when investigating spontaneous brain activity [@Wiemerslage_2016].
+putamine, insula
 Size and composition of our sample did not allow separate analyses for the sexes.
@@
 Taken together, this study showed that effects of bariatric surgery on resting state are subtle, and often confounded by head motion.
-Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to observed changes in food wanting and choice in patients after bariatric surgery.
+Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to altered food preference and choice in patients after bariatric surgery.
 Further, larger, well-controlled and preregistered neuroimaging studies are necessary to understand the neural underpinnings of altered gut-brain communication after bariatric surgery.

### 2021-02-07 code/figures.R

@@ -1,3 +1,7 @@
+# contact: heinrichs@cbs.mpg.de
+# info: creates figures
+
+# Load packages -----------------------------------------------------------
 library(ggplot2)
 
 pink_blue = c("#00305E", "#C00045")
@@ -16,13 +20,14 @@ mk_figBMIdescr <- function(final) {
       group = subj.ID,
       color = condition
     )) +
-    geom_line(position=position_nudge(x = -0.02)) + geom_point(position=position_nudge(x = -0.02)) +
+    geom_line(position=position_nudge(x = -0.02)) + 
+    geom_point(position=position_nudge(x = -0.02)) +
     stat_summary(
       aes(group = condition),
       fun = mean,
       geom  = "line",
       size = 2,
-      position=position_nudge(x = -0.02, y = 0),
+      position=position_nudge(x = 0.02, y = 0),
       color = c(rep("#046C9A", 3), rep("#D69C4E", 3))
     ) + 
     stat_summary(
@@ -167,7 +172,8 @@ mk_figRewdescr <- function(final_FC) {
       group = subj.ID,
       color = condition
     )) +
-    geom_line(position=position_nudge(x = -0.02)) + geom_point(position=position_nudge(x = -0.02)) +
+    geom_line(position=position_nudge(x = -0.02)) + 
+    geom_point(position=position_nudge(x = -0.02)) +
     stat_summary(
       aes(group = condition),
       fun = mean,
@@ -217,7 +223,8 @@ mk_figDMNdescr <- function(final_FC) {
       group = subj.ID,
       color = condition
     )) +    
-    geom_line(position=position_nudge(x = -0.02)) + geom_point(position=position_nudge(x = -0.02)) +
+    geom_line(position=position_nudge(x = -0.02)) + 
+    geom_point(position=position_nudge(x = -0.02)) +
     stat_summary(
       aes(group = condition),
       fun = mean,
@@ -526,12 +533,24 @@ figDvarsmFD <- function(final){
       group = subj.ID,
       color = condition
     )) +
-    geom_line() + geom_point() +
+    geom_line(size = 0.8, position=position_nudge(x = -0.02)) + 
+    geom_point(position=position_nudge(x = -0.02)) +
     stat_summary(
       aes(group = condition),
       fun = mean,
       geom  = "line",
-      size = 2,
+      size = 0.8,
+      position=position_nudge(x = 0.02, y = 0),
[…]

### 2021-02-07 code/tables.R

@@ -1,3 +1,9 @@
+# contact: heinrichs@cbs.mpg.de
+# info: creates tables for called upon by process.R, partly as R objects, partly
+# as LateX text obtained by knitr/kable. CAVE: especially FCTables requires manual
+# adaptation
+
+# Load packages -----------------------------------------------------------
 library(dplyr)
 library(kableExtra)
 
@@ -331,3 +337,124 @@ mk_SampleTable <- function(final) {
   return(tab_sample)
 }
 
+
+
+
+# Detailed FC Labels (supplement) -----------------------------------------
+
+mk_DetailedLabelTab <- function(){
+  # Contains two List, one with all Tables $EntryList, and one with all table 
+  # captions $CaptionList for alle Tables in the same order.
+  DIR_ANALYSIS <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
+  df <- file.path(DIR_ANALYSIS,"noExclFD/result_report.txt")
+  AnatomyResults <- read.table(df, sep="\t",col.names = c("label","dir","k"),stringsAsFactors = FALSE)
+  
+  
+  EntryList <- list()
+  
+  for (r in 1:nrow(AnatomyResults)) {
+    
+    file = AnatomyResults[r,"dir"]
+    txt_cols <- c("VoxelCount","equals","PercClusterVolumeAssignedTo","in","Hem","Region","PercentOfArea","V8","V9","V10","V11","V12","V13")
+    txt <- read.table(trimws(file),sep="\t",fill=TRUE,col.names=txt_cols) # paste0("V", 1:13)
+    
+    
+    
+    txt <- txt[sort(c(which(grepl("Cluster", txt$VoxelCount)),which(grepl("voxel", txt$equals)))),]
+    txt <- txt[,c("VoxelCount","PercClusterVolumeAssignedTo", "Hem","Region","PercentOfArea")]
+    
+    idxClusterStart <- which(grepl("Cluster", txt$VoxelCount))
+    
+    #txt$VoxelCount <- as.character(txt$VoxelCount)
+    txt <- cbind(descr = NA, txt)
+    txt$descr[idxClusterStart] <- stringr::str_remove(string = as.character(txt$VoxelCount[idxClusterStart]), pattern = "\\:[^:]*$")
+    txt$VoxelCount[idxClusterStart] <- NA
+    
+    # attach model with  cluster list to list of models
+    name_model <- AnatomyResults$label[r]
+    EntryList[[name_model]] <- txt
+  }
+  
+  
[…]

### 2021-02-08 manuscript/draft_rsfmri.Rmd
test

@@
 
-```{r "To do", include=FALSE}
-# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [ ] Swap mFD with logmFD ??
-```
-
 ```{r "load packages", include=FALSE}
-
-# load packages
 library(car) # version 3.0.9
@@
 options(knitr.kable.NA = ' ') # hide NA in tables
-# set working directory of knit (RMarkdown) to directory of Rproj
-#opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-#setwd(knitr::opts_knit$get("root.dir"))
-
 ```

### 2021-02-08 manuscript/supplements_rsfmri.Rmd
reverted deletion of Fraukes changes in draft_rsfmri.Rmd

@@
 ```{r, include = FALSE}
-library(dplyr)
-DIR_ANALYSIS <- "/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis"
-df <- file.path(DIR_ANALYSIS,"noExclFD/result_report.txt")
-AnatomyResults <- read.table(df, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)
 
-txtList <- list()
+source("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/tables.R")
 
-for (r in 1:nrow(AnatomyResults)) {
-
-  file = AnatomyResults[r,"dir"]
-  txt_cols <- c("VoxelCount","equals","PercClusterVolumeAssignedTo","in","Hem","Region","PercentOfArea","V8","V9","V10","V11","V12","V13")
-  txt <- read.table(trimws(f),sep="\t",fill=TRUE,col.names=txt_cols) # paste0("V", 1:13)
-
-  txt <- txt[sort(c(which(grepl("Cluster", txt$VoxelCount)),which(grepl("voxel", txt$equals)))),]
-  txt <- txt[,c("VoxelCount","PercClusterVolumeAssignedTo", "Hem","Region","PercentOfArea")]
-
-  idxClusterStart <- which(grepl("Cluster", txt$VoxelCount))
-
-  #txt$VoxelCount <- as.character(txt$VoxelCount)
-  txt <- cbind(descr = NA, txt)
-  txt$descr[idxClusterStart] <- stringr::str_remove(string = as.character(txt$VoxelCount[idxClusterStart]), pattern = "\\:[^:]*$")
-  txt$VoxelCount[idxClusterStart] <- NA
-
-  # attach model with  cluster list to list of models
-  name_model <- AnatomyResults$model[r]
-  txtList[[name_model]] <- txt
-}
-
-txt_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\n in cluster","\\% of cluster volume\nassigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))
+ToolboxOutputList <- mk_DetailedLabelTab()
+TableList_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\n in cluster","\\% of cluster volume\nassigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))
 
 TableList <- list()
-for (i in 1:length(txtList)){
+for (i in 1:length(ToolboxOutputList$EntryList)){
   TableList[[i]] <-
-    txtList[[i]] %>%
+    ToolboxOutputList$EntryList[[i]] %>%
     knitr::kable(
       escape = FALSE,     # use font spec of latex with kableExtra
-      col.names = txt_colname,
+      col.names = TableList_colname,
       row.names = FALSE,
@@
       #label = name_model[i],
-      caption = sprintf("Anatomical labelling for %s",name_model[i]),
+      caption = ToolboxOutputList$CaptionList[i],
     ) %>%
@@
 ```{r, echo=FALSE}
-
-AnatomyResults$model <- read.table(df, sep="\t",col.names = c("model","dir","k"),stringsAsFactors = FALSE)[1]
-l <- stringr:: str_split(AnatomyResults$model, '_')
-d <- data.frame(matrix(unlist(l), nrow=length(l), byrow=T))
-#stringr::str_replace(as.character(d[,2]),"^\\(PCC|NACC)cc$"," (cc)") # , "\\(PCC|NACC)(gsr)" = "(gsr)"
-
 TableList[[1]]

### 2021-02-08 code/process.R

@@ -1,3 +1,8 @@
+# contact: heinrichs@cbs.mpg.de
+# info: processes all computations, figure and tables required for the .Rmd file
+
+
+# Load packages -----------------------------------------------------------
 library(kableExtra)
 
 # processes data to report later in manuscript .Rmd file

### 2021-02-08 manuscript/draft_rsfmri.Rmd
restructuring result section, correcting typos.

--- 2021-02-06 manuscript/draft_rsfmri.Rmd
@@
 Resting state fMRI studies have partly mirrored these results, showing increased local FC of reward network regions, i.e.
-Nacc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
+NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
 Imaging studies in bariatric surgery patients provided first evidence that hyperconnectivity between reward network regions (e.g.
@@
 In total, $51$ participants out of the originally enrolled $69$ subjects received MRI, and we analyzed a total of $101$ data points.
-Five data points of three subjects had to be excluded due to bad anatomical image quality (see section \@ref{sec:Anatomical Quality control}).
+Five data points of three subjects had to be excluded due to bad anatomical image quality (see section [Anatomical Quality control]).
 <!-- DESCRIPTIVE STATISTICS -->
@@
 ### Functional connectivity
-To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (Nacc) and precuneus as seed regions of interest (ROI), respectively.
+To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (NAcc) and precuneus as seed regions of interest (ROI), respectively.
 We chose these regions as they are core hubs of the networks.
@@
 To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and timepoints, adjusted for age and sex.
-We used GSR regressed data as input and clusterwise bootstrapping with N=999.
+We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
 Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
@@
 
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, @spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
+-@spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
 
@@
 
-<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
+<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration.
+(*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
+
 We did not perform the analysis as preregistered (i.e.
 without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (logmFD) (Model 1b).
-<!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
+<!-- DESIGN MATRIX -->In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
 The resulting factorial design contained one regressor for each time point per group.
@@
 We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mFD and change in both measures, along with baseline age and sex (Model 2c).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model 2c).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-<!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen.
--->For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
+For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model, solely including the head motion measures as covariates of interest: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
 
 ### Model input and brain mask
-For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC with GSR), resulting in $8$ preregistered analyses <!-- timexgroup (2 models) * 2* 2 --> , and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
-For all analyses, we used an explicit mask, which was either a whole brain or a GM mask derived from the MNI ICBM "152 nonlinear 6th generation" atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
-This resulted in a reduced number of tests (voxels of interest) from $52378$ to $29451$ for the GM mask.
-<!-- MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC with GSR), resulting in $8$ preregistered analyses, and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
+For all analyses, we used an explicit mask, which was a whole brain derived from the MNI ICBM "152 nonlinear 6th generation" atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).<!--  This resulted in a number of tests (voxels of interest) from $52378$.
+MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
 -->
-Computation with the grey matter mask were exploratory and not planned in the preregistration.
 
@@
 It therefore accommodates any repeated measurement covariance structures.
-<!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
-Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.
--->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+*Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
+Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.*
+
 <!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
@@
 random field theory), and unlike permutation testing it is also suitable <!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
+Opposed to the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
@@
 In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and logmFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
-We performed model comparison between R1 and R0 models
-`R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
-`R0 = lmer(agg FC ~ timepoint + group +  age + sex + (1|subj))`
-and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
+We performed model comparison between R1 and R0 models, where `R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))`, and `R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))`, and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models using the `anova` command showed $p < 0.05$.
 As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points.
@@
 
-### Functinal decoding
+### Functional decoding
 In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
@@
 Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018].
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`).
-This relationship was negatively associated with distance between nodes (mean Spearman's r = `r round(res[1,5],2)`).
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`).
+This relationship was negatively associated with distance between nodes (mean Spearman's $r$ = `r round(res[1,5],2)`).
 Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
 Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
-Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
+Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r$ = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r$ = `r round(res[4,5],2)`).
 Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
-```{r tableQcfc, echo=FALSE}
+```{r tableQcfc, out.width="100%", echo=FALSE}
 tab_Qcfc <-
@@
 
-Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018].
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`).
-This relationship was negatively associated with distance between nodes (mean Spearman's r = `r round(res[1,5],2)`).
-Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
-Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency.
-Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`).
-Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
 ```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
@@
 
-### Whole-brain effects of bariatric surgery on DMN/reward network connectivity (preregistered analysis)
+### Whole-brain effects of bariatric surgery
 
-#### Model 1a (with adjustment for age, sex)
+#### Model 1a (adjusted for age, sex)
 
-<!-- Brain mask: Group*Time Interaction [CORRECTED] -->
 Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the PCC nor the NAcc seeds.
-There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE correction.
+There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE-correction.
 
-####  Model 1b (with adjustment for age, sex and mFD)
+####  Model 1b (adjusted for age, sex and mFD)
 Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+<!-- Neurovault -->Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are uploaded on [NeuroVault](https://neurovault.org/collections/9426/)).
 
-<!-- There was no other significant co-activation for any other region of interest for neither of the preprocessing variants. -->
-
-<!-- Neurovault -->
-Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](https://neurovault.org/collections/9426/).
-
-## Interaction effect group by time on aggregated FC (preregistered analysis)
+## Interaction effect group by time on aggregated FC
 ```{r load agg FC data and prepare, echo = FALSE}
@@
 ```
-There was no significant time by group interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless whether we adjusted for mFD.
+There was no significant group-by-time interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless whether we adjusted for mFD.
 
@@
 ```
-
-There was no significant time by group interaction effect for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC)) and supplements for a detailed summary of the models).
 
@@
 
-### Whole-brain effects of mean and change in BMI on DMN/reward network connectivity
-Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
+### Whole-brain effects of mean and change in BMI
+
+Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally on the DMN and the reward network connectivity across the three time points.
 
 #### DMN
+
 ##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
-
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCavgBMIdeact)).
-When using GSR preprocessed data, none of the clusters was significantly associated with average BMI (Table \@ref(tab:tableBMImodel)).
+Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCccavgBMIdeact)).
+When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
+Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
 ```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and logmFD"), fig.hold=TRUE, out.width = "50%"}
@@
 ```
-##### Model 2c (average and change in BMI and mFD)
 
-Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
-Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
-
-##### BMI and mFD model
+##### Model 2c (average and change in BMI and split mFD)
 To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
-<!-- ### DMN -->
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting mFD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
-
+<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
+<!-- gsr -->As before, the effect did not survive when using GSR-corrected data (see Table \@ref(tab:tableBMIFDmodel).
 ```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
@@
 ```
-Test statistics only differed slightly from the effect found when using logmFD as nuisance covariate.
-<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.
-Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
 
 #### Reward network
+
 ##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
@@
 
-<!-- ### Reward -->
-<!-- cc -->
-For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region (AAL).
+##### Model 2c (average and change in BMI and split mFD)
+<!-- cc -->For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region.
 The peak voxel was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), even if data was GSR ($p = 0.038$).
 ```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
@@
 ```
-Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
+<!-- gsr -->Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
 motor cortex, was significantly positively related to average logmFD (see  Figure \@ref(fig:BMIFDagesexNACCavgFDact)).
-This result remained similar when using GSR corrected data.
+This result was similar for GSR corrected data.
 ```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average logmFD on connectivity between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
@@
 ```
-###### Functional decoding
+<!-- Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
+Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
+-->
+
+##### Functional decoding
 Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor cortex and
-premotor were the meta-analytic activation maps most similar to this contract.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
+premotor cortex were the meta-analytic activation maps most similar to this contract.
 The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
 
-#### Aggregated FC analysis
-##### DMN (BMI and mFD model)
-```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
+### Aggregated FC analysis
+
+#### DMN 
+```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
 R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
@@
 ```
-Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD (see supplementary material for a detailed description of the models).
+Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD (model 2a and 2b).
+The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_meanBMI,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (modelc; see supplementary material for a detailed coefficients from all models).
 
-##### Model 2c (average and change in BMI and mFD)
 ```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
@@
 ```
-The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in logmFD with DMN FC.
 
-The association also remained significant when we split mFD into average and change in mFD ($p$ = `r round(p_mean_BMI,4)[2]`) and there was no significant association of mean or change in mFD with DMN FC.
-
-##### Reward network (BMI and mFD model)
+#### Reward network
 ```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
@@
 
-To investigate the unique contributions of the average value of and the longitudinal change in logmFD, we constructed two additional models.
-First, BMI measures are excluded from the design matrix.
+To investigate the unique contributions of the average value of and the longitudinal change in logmFD, we constructed an additional model, where BMI measures are excluded from the design matrix.
 Covariance of the logmFD and outcome measures is investigated by splitting logmFD into average logmFD and longitudinal logmFD change and by inspecting associations with each predictor.
@@
 
-<!-- #### Reward -->
-As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), there was a significant positive association between mFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas .
+<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), there was a significant positive association between average logmFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas.
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
+<!-- ### DMN -->We did not find a any clusters for functional connectivity with the PCC.
 ```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
@@
 ```
-<!-- ### DMN -->
-We did not find a any clusters for functional connectivity with the PCC.
 
-<!-- ### DMN -->
 ```{r tableBMImodel}
@@
 ```
-<!-- cc The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed.
-There is a significant association between average BMI and the connectivity between PCC and three other clusters.
-Results are detailed in Table \@ref(tab:tableBMIFDmodel).
-Test statistics of these clusters differ only slightly from the effect found for average BMI when using logmFD as nuisance covariate.-->
-<!-- gsr As before, the effect did not survive when thresholding if the images had been GSR corrected.-->
 
 <!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average mFD and change mFD will be are published on NeuroVault.
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault ([https://neurovault.org/collections/9426/](https://neurovault.org/collections/9426/)).
 
 # Discussion
-#### Summary
+
 In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
@@
 We found no interaction effect of group and time point on whole-brain FC of the PCC and NAcc.
-We also did not observe main effects of group and time point on the FC of these DMN and reward network hubs, respectively  Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
+Neither did we observer main effects of group and time point on the FC of these DMN and reward network hubs, respectively.
+Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
 In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme.
-When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI between the three time points and the connectivity of Nacc with a posterior-medial frontal cluster emerged.
+When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI across the three time points and the connectivity of NAcc with a posterior-medial frontal cluster emerged.
 This result was significant in both denoising schemes.
 Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns.
-Finally, higher average head motion was associated with increased Nacc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
+Finally, higher average head motion was associated with increased NAcc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
 
-In this study, we could not confirm these preregistered hypotheses and found no interaction effect between group and time point on DMN and reward network connectivity.   
+In this study, we could not confirm our preregistered hypotheses and found no interaction effect between group and time point on DMN and reward network connectivity.   
 Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
 Further, we expected FC between reward network regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in reward network regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019].
-Yet, in an exporatory analysis,  stronger BMI decrease predicted higher connectivity of the Nacc and a cluster in a posterior-medial frontal brain region.
+However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the NAcc and a cluster in a posterior-medial frontal brain region.
 Based on the Harvard-Oxford atlas and NeuroSynth decoding this region might be part of the salience network and involved in action preparation.
-AThe enhanced connectivity between the Nacc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
-Yet, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
-Previously,  a decrease in local (regional homogenity and frequency of low-amplitude oscillations) and global connectivity (degree centrality (DC)) measures in a similar region of the left SMA was reported after glucose administration  [@AlZubaidi_2018].
+The enhanced connectivity between the NAcc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
+One the other hand, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
+Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality (DC)) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi_2018].
 Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
@@
 Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2019], and therefore conducted careful analyses of the impact of head motion on our results.
-To our surprise, the effect of weight loss on the connectivity of the the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and introducing two instead of one regressors into the model.
-This result may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
+To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and thereby introducing two instead of one regressors into the model.
+These differential results may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
 Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
@@
 The major strength of our study is prospective intervention controlled design.
-We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals and , [@Thiese_2014].
+We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals [@Thiese_2014].
 Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
 In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
-Opposed to flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
+Opposed to the flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
 Limitations of our study include the low number of patients who participated in all three time points.
-In total, only X participants contributed to the estimation of the longitudinal effects with at least two  time points.
+In total, only $34$ participants contributed to the estimation of the longitudinal effects with at least two time points.
 Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI.
@@
 
-We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized?
-breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2015; @Lepping_2015, @Wiemerslage_2016; Ochner_2012_prepostprandial], and might thus have confounded our results [@Li_2019].
+We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized (?) breakfast following an overnight fast.
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2015; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow separate analyses for the sexes.

### 2021-02-08 manuscript/supplements_rsfmri.Rmd
restructuring result section, correcting typos.

@@
 ```{r "load packages", include=FALSE}
-
-# load packages
 library(car) # version 3.0.9
@@
 
-library(sjPlot) #only works for HTML output
-library(texreg) #works for Rmarkdown
+library(sjPlot) # only works for HTML output
+library(texreg) # for works for Rmarkdown
 library(knitr) # version 1.30
@@
 ```{r "tableSample", echo=FALSE}
-
 tableSample <- readRDS(file = "../report/tab/tableSample.rds")
-
 tab <-
@@
 tableDescr <- readRDS(file = "../report/tab/tableDescr.rds")
-
 knitr::kable(
@@
 
-```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="40%", eval=FALSE}
+```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="100%", fig.align = "center", eval=FALSE}
 tsnr <- "../report/tsnr.pdf"
@@
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="40%", eval=TRUE}
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="100%", fig.align = "center", eval=TRUE}
 tsnr <- "../report/tsnr.pdf"
@@
 #tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(
+  list(R0, R1),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "Timepoint x Group"
+  ),
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
 ```
@@
 #tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(
+  list(R0, R1),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "logmFD",
+    "Timepoint x Group"
+  ),
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
 ```
@@
 #tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 4, 8, 9, 5, 6, 7),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Timepoint FU2",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "logmFD",
+    "Timepoint FU x Group",
+    "Timepoint FU2 x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
 ```
@@
 #tab_model(R1,R0)
-texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(
+  list(R0, R1),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "Timepoint x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
 ```
@@
 #tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 7, 4, 5, 6),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "logmFD",
+    "Timepoint x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
 ```
@@
 #tab_model(R1,R0)
-texreg(list(R1,R11), custom.coef.names=c("Intercept", "average BMI", "change BMI", "Age", "Sex","logmFD"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(
+  list(R1, R11),
+  custom.coef.names = c("Intercept", "average BMI", "change BMI", "Age", "Sex", "logmFD"),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
 ```
@@
 ```
+
 \newpage
+### Anatomical Labelling for models
+
+Find in Table \@ref(tab:firstTab) to Table \@ref(tab:lastTab) with the detailed Output of the SPM Anatomy toolbox, version 2.2c.
+For Clusters without any labels, the toolbox was unable to assign any grey matter area.
+
 ```{r, include = FALSE}
-
-source("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/tables.R")
-
-ToolboxOutputList <- mk_DetailedLabelTab()
+ToolboxOutputList <- readRDS(file = "../report/tab/DetailedFCTableList.rds")
 TableList_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\n in cluster","\\% of cluster volume\nassigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))
@@
 
-### Anatomical Labelling for models
+```{r "firstTab", echo=FALSE}
+i = 1
+TableList[[i]] <-
+    ToolboxOutputList$EntryList[[i]] %>%
+    knitr::kable(
+      escape = FALSE,     # use font spec of latex with kableExtra
+      col.names = TableList_colname,
+      row.names = FALSE,
+      format = "latex",
+      booktabs = T,
+      linesep = "",      # disable "\\addlinespace" at every 5th line
+      #label = name_model[i],
+      caption = ToolboxOutputList$CaptionList[i],
+    ) %>%
+    kableExtra::kable_styling(latex_options = c('scale_down')) #%>%
+    #kableExtra::landscape()
+TableList[[1]] 
+```
 
 ```{r, echo=FALSE}
-TableList[[1]]
 TableList[[2]]
@@
 TableList[[8]]
+```
+
+```{r "lastTab", echo=FALSE}
+i = 9
+TableList[[i]] <-
+    ToolboxOutputList$EntryList[[i]] %>%
+    knitr::kable(
+      escape = FALSE,     # use font spec of latex with kableExtra
+      col.names = TableList_colname,
+      row.names = FALSE,
+      format = "latex",
+      booktabs = T,
+      linesep = "",      # disable "\\addlinespace" at every 5th line
+      #label = name_model[i],
+      caption = ToolboxOutputList$CaptionList[i],
+    ) %>%
+    kableExtra::kable_styling(latex_options = c('scale_down')) #%>%
+    #kableExtra::landscape()
 TableList[[9]] 

### 2021-02-08 code/process.R

@@ -145,7 +145,7 @@ final_FC=merge(final_FC, agg_FC_DMN, by="subj.ID_tp", all.x=TRUE)
 final_FC$subj.ID=as.factor(final_FC$subj.ID)
 final_FC$group_factor=relevel(final_FC$group_factor, ref = "KG")
 
-# alculate mean BMI over timepoints
+# calculate mean BMI over timepoints
 final_FC.meanBMI=tapply(X=final_FC$BMI,
                         INDEX=final_FC$subj.ID, FUN=mean, na.rm=TRUE)
 final_FC$mean.BMI=
@@ -188,3 +188,7 @@ saveRDS(object = fig_Rewconn, file = "../report/fig/figRewconn.rds")
 FCTableList <- mk_FCTables()
 saveRDS(object = FCTableList, file = "../report/tab/FCTableList.rds")
 
+
+# DetailedFCTables ----------------------------------------------------------------
+ToolboxOutputList <- mk_DetailedLabelTab()
+saveRDS(object = ToolboxOutputList, file = "../report/tab/DetailedFCTableList.rds")

### 2021-02-08 manuscript/draft_rsfmri.Rmd
typos

@@
 In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
-In the intervention group, data is not missing completely at random, as the pre-surgery timepoint MRI could not be obtained in some patients due to the limited scanner bore.
+In the intervention group, data is not missing completely at random, as the pre-surgery time point MRI could not be obtained in some patients due to the limited scanner bore.
 ```{r tableBaselineCharacteristics, include=FALSE}
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and timepoints, adjusted for age and sex.
+To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and NAccs, adjusted for age and sex.
 We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
@@
 ```
-```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all timepoints and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all NAccs and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
 
@@
 ```{r "calc DMN two tp age and sex" ,results = 'asis', echo = FALSE}
-#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
+#### Aggregated DMN for two NAccs, adjusted for age and sex (Model 1a)
 
@@
 ```{r "calc DMN two tp, age and sex and logmFD", results = 'asis', echo = FALSE}
-#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
+#### Aggregated DMN for two NAccs, adjusted for age, sex, mean FD (Model 1b)
 
@@
 ```{r "calc DMN three tp", eval = TRUE, echo = FALSE}
-#### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
+#### Aggregated DMN for three NAccs, adjusted for age, sex, mean FD (Model 1b)
 R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
@@
 ```{r "calc rew two tp adjusted", eval = TRUE, echo = FALSE}
-#### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
+#### Reward network for two NAccs, adjusted for age, sex, mean FD (Model 1b)
 
@@
 ```{r "calc rew three tp", eval = TRUE, echo = FALSE}
-#### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
+#### Reward network for three NAccs, adjusted for age, sex, mean FD (Model 1b)
 
@@
 In total, only $34$ participants contributed to the estimation of the longitudinal effects with at least two time points.
-Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI.
+Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI scanner.
 While this sample size is comparable to previous resting state fMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.

### 2021-02-09 manuscript/supplements_rsfmri.Rmd
updated tables

@@
 options(knitr.kable.NA = ' ') # hide NA in tables
+knitr::opts_chunk$set(fig.pos = "!ht", out.extra = "")
 # set working directory of knit (RMarkdown) to directory of Rproj
@@
 
-Find in Table \@ref(tab:firstTab) to Table \@ref(tab:lastTab) with the detailed Output of the SPM Anatomy toolbox, version 2.2c.
+Find in the detailed Output of the SPM Anatomy toolbox, version 2.2c.
 For Clusters without any labels, the toolbox was unable to assign any grey matter area.
+Models evaluated were model 2a estimating the effects of average BMI and change in BMI, adjusting for age and sex (Table \@ref(tab:Tab2a)), then model 2b where we additionally adjusted for logmFD (Table \@ref(tab:Tab2b)), and model 2c estimating the effects of average BMI and BMI variability as well as average logmFD and change in logmFD, adjusting for age and sex (Table \@ref(tab:Tab2cfirst) to Table \@ref(tab:Tab2clast)).
+Last, the FD model estimating the unique effects of average logmFD and change in logmFD when adjusting for age and sex was evaluated (Table \@ref(tab:TabFDfirst) and Table \@ref(tab:TabFDlast)).
 
@@
 ToolboxOutputList <- readRDS(file = "../report/tab/DetailedFCTableList.rds")
-TableList_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\n in cluster","\\% of cluster volume\nassigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))
+TableList_colname <- kableExtra::linebreak(c("Cluster","Number of voxels\nin cluster","\\% of cluster\nvolume assigned", "Hemisphere","Area","\\% of area overlap\nwith cluster"))
 
@@
     ) %>%
-    kableExtra::kable_styling(latex_options = c('scale_down')) #%>%
-    #kableExtra::landscape()
+    kableExtra::kable_styling(latex_options = c('scale_down'))
 }
@@
 
-```{r "firstTab", echo=FALSE}
-i = 1
-TableList[[i]] <-
+```{r Tab2a, echo=FALSE}
+i = 2
 ToolboxOutputList$EntryList[[i]] %>%
   knitr::kable(
-      escape = FALSE,     # use font spec of latex with kableExtra
+    escape = FALSE,
     col.names = TableList_colname,
@@
     booktabs = T,
-      linesep = "",      # disable "\\addlinespace" at every 5th line
-      #label = name_model[i],
+    linesep = "",
     caption = ToolboxOutputList$CaptionList[i],
   ) %>%
-    kableExtra::kable_styling(latex_options = c('scale_down')) #%>%
-    #kableExtra::landscape()
-TableList[[1]] 
+  kableExtra::kable_styling(latex_options = c('scale_down')) 
 ```
 
-```{r, echo=FALSE}
-TableList[[2]]
-TableList[[3]]
-TableList[[4]]
-TableList[[5]]
-TableList[[6]]
-TableList[[7]]
-TableList[[8]]
-```
-
-```{r "lastTab", echo=FALSE}
-i = 9
-TableList[[i]] <-
+```{r "Tab2b", echo=FALSE}
+i = 1
 ToolboxOutputList$EntryList[[i]] %>%
   knitr::kable(
-      escape = FALSE,     # use font spec of latex with kableExtra
+    escape = FALSE,
     col.names = TableList_colname,
@@
     booktabs = T,
-      linesep = "",      # disable "\\addlinespace" at every 5th line
-      #label = name_model[i],
+    linesep = "",
     caption = ToolboxOutputList$CaptionList[i],
   ) %>%
-    kableExtra::kable_styling(latex_options = c('scale_down')) #%>%
-    #kableExtra::landscape()
-TableList[[9]] 
+  kableExtra::kable_styling(latex_options = c('scale_down')) 
 ```
+
+```{r "Tab2cfirst", echo=FALSE}
+i = 3
+ToolboxOutputList$EntryList[[i]] %>%
+  knitr::kable(
+    escape = FALSE,
+    col.names = TableList_colname,
+    row.names = FALSE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = ToolboxOutputList$CaptionList[i],
+  ) %>%
+  kableExtra::kable_styling(latex_options = c('scale_down')) 
+```
+
+```{r, echo=FALSE}
+TableList[[4]]
+TableList[[5]]
+TableList[[6]]
+```
+
+```{r "Tab2clast", echo=FALSE}
+i = 7
+ToolboxOutputList$EntryList[[i]] %>%
+  knitr::kable(
+    escape = FALSE,
+    col.names = TableList_colname,
+    row.names = FALSE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = ToolboxOutputList$CaptionList[i],
+  ) %>%
+  kableExtra::kable_styling(latex_options = c('scale_down')) 
+```
+
+```{r "TabFDfirst", echo=FALSE}
+i = 8
+ToolboxOutputList$EntryList[[i]] %>%
+  knitr::kable(
+    escape = FALSE,
+    col.names = TableList_colname,
+    row.names = FALSE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = ToolboxOutputList$CaptionList[i],
+  ) %>%
+  kableExtra::kable_styling(latex_options = c('scale_down')) 
+```
+
+```{r "TabFDlast", echo=FALSE}
+i = 8
+ToolboxOutputList$EntryList[[i]] %>%
+  knitr::kable(
+    escape = FALSE,
+    col.names = TableList_colname,
+    row.names = FALSE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = ToolboxOutputList$CaptionList[i],
+  ) %>%
+  kableExtra::kable_styling(latex_options = c('scale_down'))
+
+```

### 2021-02-09 code/tables.R

@@ -358,17 +358,20 @@ mk_DetailedLabelTab <- function(){
     txt_cols <- c("VoxelCount","equals","PercClusterVolumeAssignedTo","in","Hem","Region","PercentOfArea","V8","V9","V10","V11","V12","V13")
     txt <- read.table(trimws(file),sep="\t",fill=TRUE,col.names=txt_cols) # paste0("V", 1:13)
     
-    
-    
     txt <- txt[sort(c(which(grepl("Cluster", txt$VoxelCount)),which(grepl("voxel", txt$equals)))),]
     txt <- txt[,c("VoxelCount","PercClusterVolumeAssignedTo", "Hem","Region","PercentOfArea")]
     
     idxClusterStart <- which(grepl("Cluster", txt$VoxelCount))
     
-    #txt$VoxelCount <- as.character(txt$VoxelCount)
     txt <- cbind(descr = NA, txt)
     txt$descr[idxClusterStart] <- stringr::str_remove(string = as.character(txt$VoxelCount[idxClusterStart]), pattern = "\\:[^:]*$")
+    txt$descr <- stringr::str_replace_all(txt$descr,'vox', 'voxel')
     txt$VoxelCount[idxClusterStart] <- NA
+    
+    txt$VoxelCount <- as.numeric(as.character(txt$VoxelCount))
+    txt$PercClusterVolumeAssignedTo <- as.numeric(as.character(txt$PercClusterVolumeAssignedTo))
+    txt$Region <- as.character(txt$Region)
+    txt$Region[(txt$Region == "" & is.na(txt$descr))] = "assigned in total" # "cluster volume assigned in total"
     
     # attach model with  cluster list to list of models
     name_model <- AnatomyResults$label[r]
@@ -411,22 +414,22 @@ mk_DetailedLabelTab <- function(){
     stringr::str_replace_all('avg', 'average ') %>%
     stringr::str_replace_all('FD', 'mFD')
   
-  AnatomyResults <-
-    tidyr::extract(
-      data = AnatomyResults,
-      col = "label",
-      into = "adjust",
-      regex = "([a-z]+)",
-      remove = FALSE
-    ) %>%
-    tidyr::extract(col = "label", into = "model", regex = "([A-Z]+)")
-  
-  AnatomyResults$model <- AnatomyResults$model %>%
-    stringr::str_replace_all('BMIFD', 'BMI-mFD')
-  
-  AnatomyResults$adjust <- AnatomyResults$adjust %>%
[…]

### 2021-02-10 manuscript/draft_rsfmri.Rmd
missing model in supplement and typos

@@
 <!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité Berlin.
-For more details see @Prehn_2020.
+For more details see Prehn et al.
+-@Prehn_2020.
 We used all data acquired until April 2019.
@@
 `r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
-<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
+Change in BMI throughout the study is depicted in Figure \@ref(fig:figBMIdescr).
+<!-- SURGICAL PROCEDURE --> As intervention, $?$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $?$ surgery purely restricting energy intake (VSG: n = $?$, GB: n = $?$).
 
@@
 -->breakfast.
-MRI scanning was done after further psychological tests (for details see @Prehn_2020).
+MRI scanning was done after further psychological tests (for details see Prehn et al.
+-@Prehn_2020).
 ```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
@@
 ```
-<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
-n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
-`r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
-<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
-<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropocentric measurements before having a one our break for <!-- standadized??
--->breakfast.
-MRI scanning was done after further psychological tests (for details see @Prehn_2020).
 
@@
 We did not perform the analysis as preregistered (i.e.
-without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (logmFD) (Model 1b).
+without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (called logmFD in the following) (Model 1b).
 <!-- DESIGN MATRIX -->In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
@@
 ### Model input and brain mask
-For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC with GSR), resulting in $8$ preregistered analyses, and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
+For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC+GSR), resulting in $8$ preregistered analyses, and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
 For all analyses, we used an explicit mask, which was a whole brain derived from the MNI ICBM "152 nonlinear 6th generation" atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).<!--  This resulted in a number of tests (voxels of interest) from $52378$.
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
 We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p < 0.01).
@@
 Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and change in mFD.
+Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average logmFD and change in logmFD.
 
@@
 In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
-We uploaded the contrast images for change BMI and average mFD on [Neurovault](), and applied the decoding classifier.
+We uploaded the contrast images for change BMI and average logmFD on [Neurovault](), and applied the decoding classifier.
 This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
@@
 ```
-```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all NAccs and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all time points and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
 
@@
 medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
-Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see Figure \@ref(fig:figNetworks)).
+Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see supplements).
 
@@
 
-####  Model 1b (adjusted for age, sex and mFD)
-Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+####  Model 1b (adjusted for age, sex and logmFD)
+Similarly, there were no significant clusters when including logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 <!-- Neurovault -->Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are uploaded on [NeuroVault](https://neurovault.org/collections/9426/)).
@@
 
-```{r figaggFC, fig.cap="Mean network connectivity per group over time separately gor the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the Reward network"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
+```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the Reward network"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
 fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
@@
 ```
-There was no significant group-by-time interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless whether we adjusted for mFD.
+There was no significant group-by-time interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless whether we adjusted for logmFD.
+Thus, we did not repeat the analysis without the 10% of highest motion subjects, as preregistered.
 
@@
 
-##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
+##### Model 2a and 2b (average and change in BMI, adjusted for logmFD)
 Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCccavgBMIdeact)).
@@
 
-##### Model 2c (average and change in BMI and split mFD)
-To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
+##### Model 2c (average and change in BMI and split logmFD)
+To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
 <!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
@@
 
-##### Model 2a and 2b (average and change in BMI, adjusted for mFD)
-NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for mFD.
+##### Model 2a and 2b (average and change in BMI, adjusted for logmFD)
+NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for logmFD.
 
-##### Model 2c (average and change in BMI and split mFD)
+##### Model 2c (average and change in BMI and split logmFD)
 <!-- cc -->For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region. 
-The peak voxel was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
+The peak voxel derived from investigating the effect of average logmFD was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), even if data was GSR ($p = 0.038$).
-```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
 premotor cortex were the meta-analytic activation maps most similar to this contract.
@@
 
-### Effect of mFD (exploratory analysis)
+### Effect of logmFD (exploratory analysis)
 #### Whole-brain analysis
@@
 We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized (?) breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2015; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow separate analyses for the sexes.

### 2021-02-10 manuscript/supplements_rsfmri.Rmd
missing model in supplement and typos

@@
 ```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvDMN2tp=anova(R1,R0)
@@
 
+### Reward network
+#### Model 2a and 2b (adjusting for mFD)
+```{r "reward model, mean, change BMI", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
+
+R11 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
+R12a = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
+anvmeanBMI_adjmFD=anova(R11,R12a)
+```
+
+```{r reward model_BMIFD results, caption="Model 2a and 2b for aggregated DMN FC", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R1, R11),
+  custom.coef.names = c("Intercept", "average BMI", "change BMI", "Age", "Sex", "logmFD"),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = FALSE
+)
+```
+#### Model 2c (adjusting for average and change in mFD)
+```{r "reward model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvmeanFD=anova(R1,R03)
+anvwithinFD=anova(R1,R03)
+
+p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+```
+
+```{r reward_BMI_splitFD results, caption="Model 2c for aggregated DMN FC", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+```
+
 \newpage

### 2021-02-15 manuscript/draft_rsfmri.Rmd
remove line numbers

@@
 header-includes:
-- \usepackage[left]{lineno}
-- \linenumbers
 - \usepackage{setspace}

### 2021-02-15 manuscript/draft_rsfmri.Rmd
changed affiliation

@@
 - Kristin Prehn [3][4]
-- Jürgen Ordemann [8][9]
-- Agnes Floel [5][6][7]
-- A. Veronica Witte [1][2][10]
+- Jürgen Ordemann [7][8]
+- Agnes Floel [5][6]
+- A. Veronica Witte [1][2][9]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
-  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Stroke Research, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Zentrum für Adipositas und Metabolische Chirurgie, Vivantes Klinikum Spandau, Berlin \newline 10 Day Clinic for Cognitive Neurology, University Clinic Leipzig
+  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin \newline 9 Day Clinic for Cognitive Neurology, University Clinic Leipzig
 output:

### 2021-03-04 manuscript/draft_rsfmri.Rmd
includes imputed baseline BMI for one participant

@@
 ---
-title: "Effects of bariatric surgery and weight loss on resting state functional connectivity"
+title: "Effects of bariatric surgery loss on resting state functional connectivity of the reward and default mode network: a pre-registered analysis"
 author:
@@
 
-Obesity imposes serious health risks and is consistently associated with differences in brain activity, including resting-state functional connectivity which can be effectively treated by bariatric surgery.
-In this study, we investigated the effects of bariatric surgery and weight loss indicated by BMI on functional connectivity in the reward network and default mode network (DMN).
-Therefore, 33 bariatric surgery patients and 15 obese waiting-list control patients underwent MRI at baseline (before surgery/waiting period), after 6 and 12 months.
-To adequately control for motion artifacts shown to be an issue in obesity studies, we applied different denoising pipelines and tried to disentangle the effects of weight loss and head motion in the second level analyses.
-Although we did not find the proposed group-by-time interaction effect, head motion and BMI were significantly associated with functional connectivity.
-When adjusting for differences in head motion within and between individuals using mean framewise displacement, both, BMI decrease and higher average head motion, resulted in stronger functional connectivity in medial posterior frontal regions within the reward network.
-While these findings could imply that severe weight loss changes reward-related neural processing in obese patients, the overlapping effects of BMI change and head motion on FC in this brain area, together with previous findings showing a strong covariance between BMI and head motion renders definite conclusions difficult.
+Obesity imposes serious health risks and is consistently associated with differences in brain activity, including resting-state functional connectivity.
+Effective treatment can be achieved by bariatric surgery, yet its effects on functional connectivity of brain networks involved in reward processingand intereoception are still under debate.
+In this pre-registered analysis, we aimed to determine the effects of bari-atric surgery and weight loss on functional connectivity in the reward network and default mode network (DMN.
+Therefore, 33 bariatric surgery patients (males/females, BMI, age etc) and 15 obese waiting-list control patients (males/females, BMI, age etc) underwent MRI at baseline (before surgery/waiting period), and after 6 and 12 months.
+We pre-registered (https://osf.io/f8tpn and https://osf.io/59bh7/) the whole-brain time-by-group interaction on network connectivity, assessed between baseline and followup and for two denoising pipelines.
+We also pre-registered a time-by-group interaction analysis of aggregated FC within networks.
+In exploratory analyses, we investigated all three time points and assessed the effects of weight loss and head motion.
+According to whole-brain marginal models, bariatric surgery compared to waiting did not significantly affect functional connectivity of the the reward and DMN networks (FWE p > 0.05).
+There also was no interaction effect on within network FC in pre-registered analysis of aggregated FC.
+Exploratory analyses also did not show an interaction effect for three time points, but when adding differences in head motion within and between individuals using mean framewise displacement, surgery-related BMI decrease and higher average head motion resulted in stronger functional connectivity in medial posterior frontal regions within the reward network.
+This pre-registered analysis of longitudinal neuroimaging data did not support a strong effect of bariatric surgery, compared to waiting, on reward and DMN network functional connectivity after 6 months in obese patients.
+Exploratory analyses indicated that differences in head motion might have confounded effects of weight loss on reward-related functional connectivity, yet definite conclusions remain difficult.
 Future studies with a more rigorous control of within-scanner head motion during data acquisition are warranted to further understand the effects of obesity surgery on functional connectivity measures.
@@
 # Introduction
-Obesity is a worldwide health issue, enchaining huge personal and societal costs.
-Excess amount of fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life.
-Conservative treatment options including dietary modification, physical activity and behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI (> 35 kg/m2).
+Obesity is a worldwide health issue, entailing huge personal and societal costs.
+Ex-cess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life (hier eine Ref).
+Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI (> 35 kg/m2).
 Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
-Roux-en-Y gastric bypass (RYGB) is the most effective surgical procedure [@Berthoud_2011].
-Here, a small pouch is formed from the proximal stomach and connected to the jejunum, disconnecting the larger part of the stomach and the proximal part of the small intestine intact from the digestive flow.
-Other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) reduce stomach volume, but leave the small intestine unchanged [@Mulla_2017].
-RYGB-induced weight loss is in part mediated by malabsorption of certain nutrients, especially fat, and reduced digestion efficiency [@Berthoud_2011] while the role of increased energy expenditure after RYGB is still unclear [@Yarmush_2017;@Lamarca_2019].
-Most importantly, RYGB alters the signaling cascades between the gut, fat tissue and the brain and thereby induces profound changes in the regulation of food intake and energy homeostasis [@Berthoud_2011].
-RYBG-patients frequently report altered smell and taste perception, i.e.
-increased sensitivity to sweet taste, show less desire to consume high-calorie and high-fat foods and actually reduce intake of such food items [@Mulla_2017; @Berthoud_2011].
-These changes in appetitive and reward signaling likely relate to altered gut-brain communication, mediated by afferent vagal nerve signaling and gut hormones like GLP-1, PYY and ghrelin [@Brutman_2019].
-
-Yet, precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
-Resting state functional magnetic resonance imaging (rsfMRI) is a powerful technique to investigate the dynamic organization of the brain via the correlation of spontaneous neural activity.
-Anatomically distinct brain regions show functional connectivity (FC), i.e.
-correlated neural activity over time, in the absence of a cognitive task.
-Crucially, these brain networks correspond to networks activated during specific cognitive processes, e.g.
-visual, auditory and somatosensory processing, or higher-order cognitive functions, i.e.
-attention and executive control [@Smith_2009], and can thus be seen as default configuration of the brain.
-In the context of bariatric surgery, the default mode network, a higher-order network, involved in interoception and governing shifts between external-internal processes, and the reward network, processing hedonic value and internal motivation, are promising candidates to mediate altered gut-brain signaling.
-The default mode network has been discovered as a “task-negative” network, and is involved in a variety of largely internally-directed cognitive functions including interoception, mind wandering and memory.
-Its core hubs are the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
+RYBG is the most invasive, yet most effective surgical pro-cedure, but other techniques like vertical sleeve gastrectomy (VSG) and gadstric banding (GB)are also common [@Berthoud_2011].
+Bariatric surgery induces weight loss  in part by malabsorption of nutrients, especially fat, and reduced digestion efficiency.
+More importantly, weight loss following the surgical procedures, and especially RYGB, is related to profound alterations in appetite, food perception and the central regulation of food intake [@Mulla_2017; @Brutman_2019].
+Precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
+Resting-state functional magnetic resonance imaging (rsfMRI) is a technique to investigate the dynamic organization of the brain.
+Functional connectivity (FC) networks, i.e.
+brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009].
+The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
+The reward network (RN), comprises the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
+These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
+Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
+Resting state fMRI studies also shown increased local FC of re-ward network regions, i.e.
+NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
+In bariatric surgery patients, connectivity  within the RN (e.g.
+putamen, OFC) might be normalized by the surgery, however the evidence is limited due to a lack of longitudinal obese control groups  [@Duan_2020; @Wiemerslage_2016].
+Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling , e.g.
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] ..
+ The default mode network includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
 BA 39) [@Raichle_2015].
-Higher BMI and obesity have been associated with differences in DMN FC in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
-Despite the large heterogenity and low median power, these findings might be summarized as a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
-salience and sensory networks.
-This has been interpreted as impaired interoceptive processing failing to adequately balance internal and external behavioral cues.
-Longitudinal studies largely support this view.
-Increased activation of the DMN in a food-viewing paradigm, i.e.
-insufficient down-regulation of DMN, has been shown to correlate with appetite measures after an overnight fast [@Tregellas_2011; @Wijngaarden_2015], an effect that could be partially reversed by an exercise intervention [@McFadden_2013].
-In a bariatric surgery study, [@Li_2018] reported a reduction of local FC in mid-line DMN regions in surgical patients compared to an obese control group.
-Simultaneously, PCC and medial prefrontal cortex were more strongly connected to dorsolateral prefrontal cortex, a region implicated in cognitive control.
-Another study in bariatric surgery patients showed that heightened connectivity of the PCC with anterior superior temporal gyrus (STG) before the surgery decreased over the course of one year [@Olivo_2017].
-Similarly, the profile of DMN connectivity with frontal regions, including ACC, frontal superior gyrus and the right OFC,  was similar in patients after bariatric surgery and normal weight controls, but dissimilar to that of obese controls [@Frank_2013].
-
-Another important brain network in the context of obesity is the reward network, comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
-These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
-Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced reward region activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
-Food cue hyperactivation has been identified as a main neural vulnerability factor predicting future weight gain [@Stice_2015].
-Resting state fMRI studies have partly mirrored these results, showing increased local FC of reward network regions, i.e.
-NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
-Imaging studies in bariatric surgery patients provided first evidence that hyperconnectivity between reward network regions (e.g.
-putamen, OFC) might be normalized to some extent by the surgery [@Duan_2020; @Wiemerslage_2016].
-Possible, the reconfiguration of the fronto-striatal brain networks might be related to altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] .
-Thus, while there is some evidence hinting to a role for DMN and reward network connectivity in altered gut-brain communication after bariatric surgery, the existing evidence is inconclusive.
-Most studies have investigated small cohort of patients, often without adequate control group, and employed exploratory statistics [@George_2016].
-
-In this study, we aimed to extend the current literature and investigate differences of DMN and reward network connectivity before and after bariatric surgery in preregistered analyses and in comparison with a contemporaneously measured waiting-control group.
-Moreover, we explored whether the changes in functional connectivity were associated with the amount of weight lost after bariatric surgery treatment in obesity to explore a potential dose-response relationship.
-Participants were measured at baseline, at 6 months and at 24 months to capture both phases of rapid weight-loss and maintenance as expected from non-surgical interventions [@Shai_2008].
-Notably, we took into account possible non-linear time courses in increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
+Higher BMI and obesity have been associated with a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
+salience and sensory networks in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
+Studies in bariatric surgery patients have pointed to a normalization of the connectivity between  DMN and cognitive control and salience brain regions after the procedure, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
+In sum, while there is some evidence hinting to a role for DMN and RN connectivity in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
+Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].
+Most resting-state fMRI studies in obesity have reported an association of higher BMI and higher head motion during MRI [@Beyer_2017;@Hodgson_2016].
+Head motion is a major confounder in resting-state fMRI as it increases short-range and disruptes long-range connectivity FC.
+In this sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
+Here, we thus aimed to rigorously control for motion-related variance.
+In this pre-registered study, our hypotheses were:
+1. Whole-brain FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to waiting-list control group
+2. Within-network FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to waiting-list control group
+We tested hypothesis 1 by investigating the interaction of bariatric surgery and time on DMN and RN whole-brain FC.
+We pre-registed two denoising pipelines, and three covariate schemes.
+For hypothesis 2, we performed a planned analysis of the group-by-time interaction on aggregated, within-network FC, for two and three time points and the same covariate schemes.
+In exploratory analyses, we examined the whole-brain interaction effect for three time points, the effects of head motion on FC and whether weight loss, a proxy of treatment success, predicted changes in FC.
 
@@
 
-<!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité Berlin.
+<!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité University Medicine Berlin, Germany.
 For more details see Prehn et al.
--@Prehn_2020.
+(2020).
 We used all data acquired until April 2019.
-The study protocol was in accordance with the Helsinki Declaration and approved by the local ethics committee.
+The study protocol was in accordance with the Helsinki Declaration and approved by the ethics committee of the Charité University Medicine Berlin.
 The study was registered at clinicaltrials.gov as NCT01554228.
-We preregistered the present resting state analyses on the [OSF](https://osf.io/yp42s).
-We made [additional changes](https://osf.io/59bh7/) to the preregistration after preprocessing as we realized some aspects of the analysis were inadequately described in the first preregistration.
+We preregistered the present resting state analyses on the OSF https://osf.io/yp42s.
+We made additional changes https://osf.io/59bh7/ to the preregistration after preprocessing as we realized some aspects of the analysis were inadequately described in the initial preregistration.
 <!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA -->
-Participants were recruited from the Center for Bariatric and Metabolic Surgery.
-Inclusion criteria were, in accordance with BARS guidelines, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
+Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin.
+Inclusion criteria were, in accordance with German guidelines for bariatric surgery, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score $< 34$).
+Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score < 34).
  In total, $51$ participants out of the originally enrolled $69$ subjects received MRI, and we analyzed a total of $101$ data points.
-Five data points of three subjects had to be excluded due to bad anatomical image quality (see section [Anatomical Quality control]).
+Five data points of three subjects had to be excluded due to bad anatomical image quality (see below).
 <!-- DESCRIPTIVE STATISTICS -->
@@
 Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery or after the baseline appointment.
-Analyses were performed on all participants that provided at least one data point of fMRI data (detailed information on data available.
-In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for only one time point.
-In the intervention group, data is not missing completely at random, as the pre-surgery time point MRI could not be obtained in some patients due to the limited scanner bore.
-```{r tableBaselineCharacteristics, include=FALSE}
-#dfBL <- create_sample_df(group = "all", tp = "BL")
-dfBL <- readRDS(file = "../report/dfBL.rds")
-
-# distributions per group
-# BMI
-p <- ggplot(data=dfBL, aes(group = condition, x=BMI, fill=condition)) +
-  geom_density(alpha=0.6) +
-  scale_fill_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")) +
-  theme_bw()
-# log mean FD
-p <- ggplot(data=dfBL, aes(group = condition, x=logmFD, fill=condition)) +
-  geom_density(alpha=0.6) +
-  scale_fill_manual(
-    values = c("#00305E55", "#C0004555"),
-    labels = c("intervention", "control")) +
-  theme_bw()
-# Age
-p <- ggplot(data=dfBL, aes(group = condition, x=Age_BL, fill=condition)) +
-  geom_density(alpha=0.6) +
-  scale_fill_manual(
-    values = c("#00305E77", "#C0004577"),
-    labels = c("intervention", "control")) +
-  theme_bw()
-
-tab_age <- psych::describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
-
-# Sex
-p <- ggplot(data = dfBL, aes(x = condition, fill = Sex)) +
-  geom_bar(position = "dodge") +
-  scale_fill_manual(labels = c("female", "male"),
-                    values = c("#00305E55", "#C0004555")
-  )
-tab_sex <- table(dfBL$condition, dfBL$Sex)
-```
-<!-- BASELINE CHARACTERISTICS -->
-Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mean framewise displacement (FD), and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
-n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
-`r round(tab_age$mean[tab_age$group1=="IG"],2)`).
-Change in BMI throughout the study is depicted in Figure \@ref(fig:figBMIdescr).
-<!-- SURGICAL PROCEDURE --> As intervention, $?$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $?$ surgery purely restricting energy intake (VSG: n = $?$, GB: n = $?$).
-
-<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast.
-They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one our break for <!-- standadized??
--->breakfast.
-MRI scanning was done after further psychological tests (for details see Prehn et al.
--@Prehn_2020).
-```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
-# eval = TRUE, fig.asp = 0.7, fig.width = 6
-readRDS(file = "../report/fig/fig_BMIdescr.rds")
-```
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016].
+Analyses were performed on all participants that provided at least one data point of fMRI data.
+In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point
+**N patients underwent RYGB, N underwent VSG and N GB.** 
+Patients arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
+MRI scanning was done after further psychological tests (for details see @Prehn_2020)).
 
@@
 Imaging data was conducted using AFNI 19.1.05, ANTS '2.3.1', FSL '6.0.1' and FreeSurfer '6.0.0p1', wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15.
-The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel.
-The workflow can be found on [github](https://github.com/fBeyer89/ADI_preproc/).
+The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel. The workflow can be found on https://github.com/fBeyer89/ADI_preproc/.
 
@@
 Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.
-
-### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and NAccs, adjusted for age and sex.
-We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
-Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
-We extracted the average connectivity from these masks for GSR regressed data.
-*In the preregistration we wrote: "We will use the average of connectivity values within masks of the respective networks as aggregate network connectivity measure.
-The mask will include all voxels surviving FWE-correction of .05 in a one-sample t-test of the individual connectivity maps across all participants and time points, calculated with the modified SwE option in the SwE toolbox"*
-
-## Analysis
-
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
--@spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
-
-### Design Specification
-
-<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration.
-(*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
-
-We did not perform the analysis as preregistered (i.e.
-without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (called logmFD in the following) (Model 1b).
-<!-- DESIGN MATRIX -->In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate.
-The resulting factorial design contained one regressor for each time point per group.
-
-We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
-As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
-This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model 2c).
-Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model, solely including the head motion measures as covariates of interest: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.
-
-### Model input and brain mask
-For each brain network, we examined the models for the two preprocessing pipelines (AROMA+CC / AROMA+CC+GSR), resulting in $8$ preregistered analyses, and $16$ <!-- bmi change (2 models), bmi + mfd change, fd change=4 * 2* 2 --> exploratory whole brain analysis.
-For all analyses, we used an explicit mask, which was a whole brain derived from the MNI ICBM "152 nonlinear 6th generation" atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).<!--  This resulted in a number of tests (voxels of interest) from $52378$.
-MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
--->
-
-### SwE settings
-<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
-It therefore accommodates any repeated measurement covariance structures.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
-*Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach.
-Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images.*
-
-<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
-Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
-In contrast to the parametric estimation, wild bootstrap avoids any parametric distributional assumptions (e.g.
-random field theory), and unlike permutation testing it is also suitable <!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
-Opposed to the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000).
-Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom.
-We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p < 0.01).
-Significant clusters are defined as family-wise error - corrected p < 0.05.
--->
-
-### Aggregated FC analysis
-<!-- R analysis of aggregated values -->
-We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
-Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up.
-In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and logmFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
-We performed model comparison between R1 and R0 models, where `R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))`, and `R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))`, and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models using the `anova` command showed $p < 0.05$.
-As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points.
-Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
-Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average logmFD and change in logmFD.
-
-### Functional decoding
-In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
-We uploaded the contrast images for change BMI and average logmFD on [Neurovault](), and applied the decoding classifier.
-This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
-We reported the three top terms for both contrasts.
-
-# Results
-
 ## Quality control
@@
 ```
-```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all time points and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
 
-networks = c("../report/fig/DMN_thresholded_for_clusterFWE005.pdf",
-             "../report/fig/Rew_thresholded_for_clusterFWE005.pdf")
-knitr::include_graphics(networks)
+### Aggregated FC values
+To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and NAccs, adjusted for age and sex.
+We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
+Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
+We extracted the average connectivity from these masks for GSR regressed data.
+
+## Analysis
+
+Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
+-@spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
+
+### Design Specification
+
+<!-- MODEL SETUP --> 
+#### Confirmatory Analysis
+We tested the preregistered hypothesis of a  time by group interaction:  $FC=group + time + group * time$. 
+As preregistered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average mFD of both timepoints (CA2) and age,sex, average mFD and baseline BMI (CA3).
+ In this model, time was represented as factor because we took into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management (Olivo et al.
+2017).
+The resulting factorial design contained one regressor for each time point per group (see supplements for depiction of design matrix).
+
+#### Exploratory analysis
+##### Group-by-time interaction in three time points
+Then, we extended our whole-brain analysis of the group-by-time point interaction to the data from all three timepoints.
+This analysis had not been preregistered.
+Here, we used individual mFD values (not averaged) as covariate to capture variance in mFD change over time points.We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and mFD).
+##### Average and change BMI
+We further  explored the relationship of average and change in BMI in a separate group of models.
+As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
+This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
+We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for log mean mFD (Model EA 2.2).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 2.3c).
+Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
+For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: $FC = between-subject FD + within-subject FD$ age and sex as nuisance covariates.
+
+### Functional decoding
+In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
+We uploaded the contrast images for change BMI and average logmFD on [Neurovault](), and applied the decoding classifier.
+This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
+We reported the three top terms for both contrasts.
+
+### Model input and brain mask
+For each brain network, we examined the models for the two denoising pipelines (AROMA+CC / AROMA+CC+GSR)as preregistered.
+For all analyses, we used an explicit brain mask, which was a whole brain derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).<!--  This resulted in a number of tests (voxels of interest) from $52378$.
+MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
+-->
+
+### SwE settings
+<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+It therefore accommodates any repeated measurement covariance structures.
+We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
+
+<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+Deviating from the preregistration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual.
+Opposed to the description in the preregistration, we used a cluster forming threshold of p<0.001 for a more rigorous multiple comparison adjustment (instead of p<0.01, and 1000 bootstraps due to required computation time (instead of 5000)).
+Significant clusters are defined as family-wise error (FWE) corrected p<0.05.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.
+
+### Aggregated FC analysis
+#### Confirmatory analyses
+<!-- R analysis of aggregated values -->
+We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
+We deviated from the pre-registration by only investigating data from the GSR denoising pipeline.
+First, we investigated the time-by-group interaction for baseline and follow-up time points only.
+We either performed no correction (CA1), adjusted for baseline age, sex, average of mean FD of both time points (CA2) or additionally adjusted for baseline BM (CA3)I .
+We performed model comparison between R1 and R0 models, where R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj)), and R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj)).
+As planned in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.
+The interaction effect was considered significant if the model comparison between R1 and R0 models using the anova command showed p<0.05.
+
+#### Exploratory analysis
+We performed an exploratory analysis with average BMI and variability in BMI as predictors of the aggregated FC from GSR data of both networks, in line with our exploratory whole brain analysis.
+We calculated three models with average and change in BMI as predictors of inter-est and adjusted for baseline age and sex (Model EA2.1), mFD (Model EA 2.2) and average and change mFD (Model EA 2.3).
+
+# Results
+<!-- BASELINE CHARACTERISTICS -->
+```{r tableBaselineCharacteristics, include=FALSE}
+#dfBL <- create_sample_df(group = "all", tp = "BL")
+dfBL <- readRDS(file = "../report/dfBL.rds")
+
+# distributions per group
+# BMI
+p <- ggplot(data=dfBL, aes(group = condition, x=BMI, fill=condition)) +
+  geom_density(alpha=0.6) +
+  scale_fill_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")) +
+  theme_bw()
+# log mean FD
+p <- ggplot(data=dfBL, aes(group = condition, x=logmFD, fill=condition)) +
+  geom_density(alpha=0.6) +
+  scale_fill_manual(
+    values = c("#00305E55", "#C0004555"),
+    labels = c("intervention", "control")) +
+  theme_bw()
+# Age
+p <- ggplot(data=dfBL, aes(group = condition, x=Age_BL, fill=condition)) +
+  geom_density(alpha=0.6) +
+  scale_fill_manual(
+    values = c("#00305E77", "#C0004577"),
+    labels = c("intervention", "control")) +
+  theme_bw()
+
+tab_age <- psych::describeBy(data=dfBL, x=dfBL$Age_BL, group=dfBL$condition, mat=T)
+
+# Sex
+p <- ggplot(data = dfBL, aes(x = condition, fill = Sex)) +
+  geom_bar(position = "dodge") +
+  scale_fill_manual(labels = c("female", "male"),
+                    values = c("#00305E55", "#C0004555")
+  )
+tab_sex <- table(dfBL$condition, dfBL$Sex)
+```
+Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mFD.
+There were slight differences in the distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs.
+n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs.
+`r round(tab_age$mean[tab_age$group1=="IG"],2)`).
+In the intervention group, baseline MRI data is not missing completely at random, due to the limited scanner bore.
+Similarly, we observed a higher drop-out in the control than in the surgery group (baseline timepoint only, n control = ; n surgery = ).
+Change in BMI throughout the study is depicted in Figure \@ref(fig:figBMIdescr).
+
+```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
+# eval = TRUE, fig.asp = 0.7, fig.width = 6
+readRDS(file = "../report/fig/fig_BMIdescr.rds")
 ```
 
-Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
-Both networks included the typically described brain regions (i.e.
-medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
-Yet, the reward network was less pronounced (overall lower T-values) and symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see supplements).
-
-## Interaction effect of intervention and time (preregistered analysis)
-
-### Whole-brain effects of bariatric surgery
-
-#### Model 1a (adjusted for age, sex)
-
-Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the PCC nor the NAcc seeds.
+## Whole-brain analyses: Confirmatory analysis 
+First, we performed the preregistered analysis in participants with baseline and followup data, using two denoising pipelines and three covariate adjustements.
+### Model CA1 (without adjustments)
+Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the NAcc nor the PCC seeds.
 There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE-correction.
 
-####  Model 1b (adjusted for age, sex and logmFD)
-Similarly, there were no significant clusters when including logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
-<!-- Neurovault -->Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are uploaded on [NeuroVault](https://neurovault.org/collections/9426/)).
+###  Model CA2 (adjusted for age, sex and average of mFD)
+Similarly, there were no significant clusters when adjusting for age, sex and average of logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 
-## Interaction effect group by time on aggregated FC
-```{r load agg FC data and prepare, echo = FALSE}
-final_FC <- read.csv("../report/final_FC.csv")
-```
+###  Model CA3 (adjusted for age, sex, average of mFD and baseline BMI)
+There was no significant interacion when adding baseline BMI.
 
-```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the Reward network"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
-fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
-fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
-   fig_DMNconn
-fig_Rewconn
-```
-There was no significant group-by-time interaction for aggregated DMN and reward network FC (see Figure \@ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless whether we adjusted for logmFD.
-Thus, we did not repeat the analysis without the 10% of highest motion subjects, as preregistered.
+<!-- Neurovault -->The unthresholded contrast maps of group, time and group-by-time interaction for the unadjusted model are uploaded on [NeuroVault](https://neurovault.org/collections/9426/)).
 
-```{r identify outliers in these plots, eval=FALSE}
-### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
-final_FC[final_FC$mean_DMN_conn< 0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
-#ADI003_fu2 & ADI041_fu -> do not have particularly high head motion
-#plot(final_FC$mean_DMN_conn, final_FC$logmFD) -> see two left points
-#carpet plots look ok, brain maps of DMN look okish, though not pronounced
-#ADI003_fu2 & ADI041 fu: FS issues (Final_score:3, corrected)
-qa_info=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
-qa_info[qa_info$subj.ID=="ADI003"&qa_info$tp=="fu2",]
-qa_info[qa_info$subj.ID=="ADI041"&qa_info$tp=="fu",]
-
-final_FC[final_FC$mean_Rew_conn>0.3, c("mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
-plot(final_FC$mean_Rew_conn, final_FC$logmFD)
-#-> do not have particularly high head motion (two right points)
-#ADI069_bl
-#nifti looks ok, carpet plot shows lot of stripes.
-qa_info[qa_info$subj.ID=="ADI069"&qa_info$tp=="bl",]
-#FS issues, parts of skull included
-#ADI111_fu2
-#nifti looks ok, some residual carpet plot noise
-qa_info[qa_info$subj.ID=="ADI111"&qa_info$tp=="fu2",]
-#no FS issues
-```
-
-```{r "calc DMN two tp age and sex" ,results = 'asis', echo = FALSE}
-#### Aggregated DMN for two NAccs, adjusted for age and sex (Model 1a)
-
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvDMN2tp=anova(R1,R0)
-```
-
-```{r "qa check DMN model age sex", eval = FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-   #to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
-vif_res=vif(test_vif)
-```
-
-```{r "DMN2tp results age sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-
-```
-```{r "calc DMN two tp, age and sex and logmFD", results = 'asis', echo = FALSE}
-#### Aggregated DMN for two NAccs, adjusted for age, sex, mean FD (Model 1b)
-
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvDMN2tp=anova(R1,R0)
-```
-```{r, "qa check DMN model age,sex, logmFD", eval=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
-vif_res=vif(test_vif)
-```
-```{r "DMN2tp results age sex logmFD", caption="DMN connectivity compared bl to fu, adjusted for age, sex and logmFD", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-```{r "calc DMN three tp", eval = TRUE, echo = FALSE}
-#### Aggregated DMN for three NAccs, adjusted for age, sex, mean FD (Model 1b)
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-anvDMN3tp=anova(R1,R0)
-```
-```{r "DMN3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-```{r "calc rew two tp, age, sex", eval = TRUE, echo = FALSE}
-### Reward network for two time points, adjusted for age and sex (Model 1a)
-
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvRew2tp=anova(R1,R0)
-```
-
-```{r "qa check Rew model age sex", eval=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-```
-```{r "Rew2tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-```{r "calc rew two tp adjusted", eval = TRUE, echo = FALSE}
-#### Reward network for two NAccs, adjusted for age, sex, mean FD (Model 1b)
-
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvRew2tpadj=anova(R1,R0)
-```
-```{r "qa check Rew model age, sex, logmFD", eval = FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
-vif_res=vif(test_vif)
-```
-```{r "Rew2tpadj results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-```{r "calc rew three tp", eval = TRUE, echo = FALSE}
-#### Reward network for three NAccs, adjusted for age, sex, mean FD (Model 1b)
-
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
-anvRew3tp=anova(R1,R0)
-```
-```{r "Rew3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
-#tab_model(R1,R0)
-texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
-```
-
-## Exploratory Analyses
+## Whole-brain: Exploratory Analyses
+### Group-by-time interaction for three time points (EA 1.1 and EA 1.2)
+Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA 1.1 and EA 1.2 in the full data set for neither reward network nor DMN.
 
 ### Whole-brain effects of mean and change in BMI
+##### Reward network
 
-Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally on the DMN and the reward network connectivity across the three time points.
-
-#### DMN
-
-##### Model 2a and 2b (average and change in BMI, adjusted for logmFD)
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCccavgBMIdeact)).
-When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
-Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
-```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and logmFD"), fig.hold=TRUE, out.width = "50%"}
-knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
-```
-
-##### Model 2c (average and change in BMI and split logmFD)
-To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
-<!-- gsr -->As before, the effect did not survive when using GSR-corrected data (see Table \@ref(tab:tableBMIFDmodel).
-```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
-knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
-```
-
-#### Reward network
-
-##### Model 2a and 2b (average and change in BMI, adjusted for logmFD)
+###### Average and change in BMI, adjusted for logmFD (EA 2.1 and EA 2.2)
 NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for logmFD.
 
-##### Model 2c (average and change in BMI and split logmFD)
+###### Average and change in BMI, adjusted for split mFD (EA 2.3)
 <!-- cc -->For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region. 
@@
 
-##### Functional decoding
-Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
-premotor cortex were the meta-analytic activation maps most similar to this contract.
-The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
+##### DMN
 
-### Aggregated FC analysis
-
-#### DMN 
-```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
-
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
-R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
-anvmeanBMI_adjmFD=anova(R1,R02a)
-```
-Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD (model 2a and 2b).
-The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_meanBMI,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (modelc; see supplementary material for a detailed coefficients from all models).
-
-```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-anvmeanFD=anova(R1,R03)
-anvwithinFD=anova(R1,R03)
-
-p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+###### Average and change in BMI, adjusted for mFD (EA 2.1 and EA 2.2)
+Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCccavgBMIdeact)).
+When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
+Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and logmFD"), fig.hold=TRUE, out.width = "50%"}
+knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
 ```
 
-#### Reward network
-```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-anvwithinFD=anova(R1,R03)
-anvmeanFD=anova(R1,R04)
+###### Average and change in BMI, adjusted for split mFD (EA 2.3)
+To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
+<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
+<!-- gsr -->As before, the effect did not survive when using GSR-corrected data (see Table \@ref(tab:tableBMIFDmodel).
+```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
+knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
+```
+<!-- Neurovault -->
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault https://neurovault.org/collections/9426/.
 
-test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
-vif_res=vif(test_vif)
-```
-As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and reward network connectivity in models 2a, 2b or 2c adjusting for age, sex and logmFD.
+##### Effect of mFD (EA 2.4)
 
-### Effect of logmFD (exploratory analysis)
-#### Whole-brain analysis
-
-To investigate the unique contributions of the average value of and the longitudinal change in logmFD, we constructed an additional model, where BMI measures are excluded from the design matrix.
+To explore the unique contributions of the average value of and the longitudinal change in logmFD, we constructed an additional model, where BMI measures are excluded from the design matrix.
 Covariance of the logmFD and outcome measures is investigated by splitting logmFD into average logmFD and longitudinal logmFD change and by inspecting associations with each predictor.
@@
 ```
+
+##### Functional decoding
+Functional decoding of these two activation patterns showed different top association terms.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
+premotor cortex were the meta-analytic activation maps most similar to this contract.
+The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
 
@@
 
-<!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault ([https://neurovault.org/collections/9426/](https://neurovault.org/collections/9426/)).
+## Aggregated FC analysis
+### Confirmatory analysis of group-by-time interaction (CA1, CA2, CA3)
+```{r load agg FC data and prepare, echo = FALSE}
+final_FC <- read.csv("../report/final_FC.csv")
+final_FC=merge(final_FC,subset(final_FC, tp=="bl",select=c("subj.ID", "BMI")), by='subj.ID')
+colnames(final_FC)[28]="BMI.bl"
+```
+
+```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the Reward network"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
+fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
+fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
+   fig_DMNconn
+fig_Rewconn
+```
+There was no significant group-by-time interaction for aggregated DMN and reward network FC (see Figure @ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless of the adjustments (CA1: without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
+Thus, we did not repeat the analysis with-out the 10% of highest motion subjects, as preregistered.
+
+### Exploratory analysis: effects of mean and change in BMI on aggregated FC
+#### Reward network
+##### Models EA 2.1, 2.2. and 2.3
+```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvwithinFD=anova(R1,R03)
+anvmeanFD=anova(R1,R04)
+
+test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
+vif_res=vif(test_vif)
+```
+As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and within reward network connectivity in models 2a, 2b or 2c adjusting for age, sex and logmFD.
+#### DMN
+##### Models EA 2.1, 2.2. and 2.3
+```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
+
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
+R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
+anvmeanBMI_adjmFD=anova(R1,R02a)
+```
+Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD (model EA 2.1 and EA 2.2).
+The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_meanBMI,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (modelc; see supplementary material for a detailed coefficients from all models).
+
+```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvmeanFD=anova(R1,R03)
+anvwithinFD=anova(R1,R03)
+
+p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+```
 
@@
 
-In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
+In this pre-registered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
 Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) on the group level.

### 2021-03-04 manuscript/supplements_rsfmri.Rmd
includes imputed baseline BMI for one participant

--- 2021-02-09 manuscript/supplements_rsfmri.Rmd
@@
     toc: no
-    extra_dependencies: "subfig"
+    extra_dependencies: ["float","subfig"]
 bibliography: "bibliography.bib"
@@
 
+## Quality Control
+
+```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="100%", fig.align = "center", eval=FALSE}
+tsnr <- "../report/tsnr.pdf"
+knitr::include_graphics(tsnr)
+```
+
+## Networks
+Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
+Both networks included the typically described brain regions (i.e.
+medial prefrontal cortex, parietal lobules and posterior hippocampus for the DMN, and anterior cingulate, right amygdala and ventromedial-prefrontal cortex for the reward network).
+Yet, the reward network was less pronounced (overall lower T-values) and less symmetric (no significant connectivity with left amygdala), which might be due to lower SNR in the seeded brain region (see \@ref(fig:figtSNR)).
+Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) are on Neurovault.
+```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all time points and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+
+networks = c("../report/fig/DMN_thresholded_for_clusterFWE005.pdf",
+             "../report/fig/Rew_thresholded_for_clusterFWE005.pdf")
+knitr::include_graphics(networks)
+```
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="100%", fig.align = "center", eval=TRUE}
+tsnr <- "../report/tsnr.pdf"
+knitr::include_graphics(tsnr)
+```
+
 ## Design
@@
 
-```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="100%", fig.align = "center", eval=FALSE}
-tsnr <- "../report/tsnr.pdf"
-knitr::include_graphics(tsnr)
+## Aggregated FC: confirmatory analysis of an interaction effect of group and time 
+```{r load agg FC data and prepare, echo=FALSE}
+final_FC <- read.csv("../report/final_FC.csv")
+final_FC$group_factor=relevel(final_FC$group_factor, "KG")
+final_FC=merge(final_FC,subset(final_FC, tp=="bl",select=c("subj.ID", "BMI")), by='subj.ID')
+colnames(final_FC)[27]="BMI.bl"
 ```
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="100%", fig.align = "center", eval=TRUE}
-tsnr <- "../report/tsnr.pdf"
-knitr::include_graphics(tsnr)
+```{r identify outliers in these plots, eval=FALSE}
+### SUMMARY: all in all there is no reason to exclude the subjects with outlying values in mean Rew/DMN connectivity
+final_FC[final_FC$mean_DMN_conn< 0.1, c("mean_DMN_conn", "mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
+#ADI003_fu2 & ADI041_fu -> do not have particularly high head motion
+#plot(final_FC$mean_DMN_conn, final_FC$logmFD) -> see two left points
+#carpet plots look ok, brain maps of DMN look okish, though not pronounced
+#ADI003_fu2 & ADI041 fu: FS issues (Final_score:3, corrected)
+qa_info=read.csv("/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
+qa_info[qa_info$subj.ID=="ADI003"&qa_info$tp=="fu2",]
+qa_info[qa_info$subj.ID=="ADI041"&qa_info$tp=="fu",]
+
+final_FC[final_FC$mean_Rew_conn>0.3, c("mean_Rew_conn", "subj.ID_tp", "subj.ID", "tp", "condition")]
+plot(final_FC$mean_Rew_conn, final_FC$logmFD)
+#-> do not have particularly high head motion (two right points)
+#ADI069_bl
+#nifti looks ok, carpet plot shows lot of stripes.
+qa_info[qa_info$subj.ID=="ADI069"&qa_info$tp=="bl",]
+#FS issues, parts of skull included
+#ADI111_fu2
+#nifti looks ok, some residual carpet plot noise
+qa_info[qa_info$subj.ID=="ADI111"&qa_info$tp=="fu2",]
+#no FS issues
 ```
 
-## Interaction effect group by time on aggregated FC (preregistered analysis)
-```{r load agg FC data and prepare, echo=FALSE}
-final_FC <- read.csv("../report/final_FC.csv")
+### Reward network 
+#### Two time points, without adjustments (Model CA1)
+
+<b>
+
+```{r calc rew two tp, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvRew2tp=anova(R1,R0)
 ```
 
-### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)
-```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvDMN2tp=anova(R1,R0)
-```
-
-```{r qa check DMN model age,sex, eval=FALSE}
-diag=diagnostics.plot(R0)
-randefplot=ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex, data=final_FC)
-vif_res=vif(test_vif)
-```
-
-```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(
-  list(R0, R1),
-  custom.model.names = c("R0", "R1"),
-  float.pos = "h",
-  custom.coef.names = c(
-    "Intercept (BL, NBARS)",
-    "Timepoint FU",
-    "Group BARS",
-    "Age",
-    "Sex",
-    "Timepoint x Group"
-  ),
-  center = TRUE,
-  digits = 3,
-  leading.zero = FALSE,
-  single.row = TRUE,
-  include.adjrs = FALSE,
-  include.bic = FALSE,
-  table = FALSE,
-  use.packages = FALSE
-)
-```
-
-### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)
-
-```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvDMN2tp=anova(R1,R0)
-```
-
-```{r qa check DMN model age,sex, mFD, eval=FALSE}
-diagnostics.plot(R0)
-ranef.diagn.plot(R0)
-
-m.stab=glmm.model.stab(model.res = R0)
-#to check whether there were any converge issues:
-table(m.stab$detailed$warnings)
-m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
-vif_res=vif(test_vif)
-```
-
-```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(
-  list(R0, R1),
-  custom.model.names = c("R0", "R1"),
-  float.pos = "h",
-  custom.coef.names = c(
-    "Intercept (BL, NBARS)",
-    "Timepoint FU",
-    "Group BARS",
-    "Age",
-    "Sex",
-    "logmFD",
-    "Timepoint x Group"
-  ),
-  center = TRUE,
-  digits = 3,
-  leading.zero = FALSE,
-  single.row = TRUE,
-  include.adjrs = FALSE,
-  include.bic = FALSE,
-  table = FALSE,
-  use.packages = FALSE
-)
-```
-
-### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
-```{r calc DMN three tp, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-anvDMN3tp=anova(R1,R0)
-```
-
-```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
-#tab_model(R1,R0)
-texreg(
-  list(R0, R1),
-  reorder.coef = c(1, 2, 3, 4, 8, 9, 5, 6, 7),
-  custom.coef.names = c(
-    "Intercept (BL, NBARS)",
-    "Timepoint FU",
-    "Timepoint FU2",
-    "Group BARS",
-    "Age",
-    "Sex",
-    "logmFD",
-    "Timepoint FU x Group",
-    "Timepoint FU2 x Group"
-  ),
-  custom.model.names = c("R0", "R1"),
-  float.pos = "h",
-  center = TRUE,
-  digits = 3,
-  leading.zero = FALSE,
-  single.row = TRUE,
-  include.adjrs = FALSE,
-  include.bic = FALSE,
-  table = FALSE,
-  use.packages = FALSE
-)
-```
-
-```{r DMN conn over three timepoints, echo=FALSE}
-fig_DMNconn <- readRDS(file = "../report/fig/figDMNconn.rds")
-fig_DMNconn
-```
-
-### Reward network for two time points, adjusted for age and sex (Model 1a)
-```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-anvRew2tp=anova(R1,R0)
-```
 ```{r qa check Rew model age,sex, eval=FALSE}
@@
 
-```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+```{r Rew2tp results, caption="Interaction of group and time on reward network connectivity for two time points", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
     "Group BARS",
-    "Age",
-    "Sex",
     "Timepoint x Group"
@@
   custom.model.names = c("R0", "R1"),
-  float.pos = "h",
+  float.pos = "h!",
   center = TRUE,
@@
   table = FALSE,
-  use.packages = FALSE
+  use.packages = TRUE
 )
@@
 
-### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)
-```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+#### Two timepoints, adjusted for age, sex and average mFD (Model CA2)
+
+```{r calc rew two tp adjusted 2, fig.cap="Two timepoints, adjusted for age, sex and average mFD (Model CA2)", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvRew2tpadj=anova(R1,R0)
@@
 
-```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+```{r Rew2tpadj2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
     "Sex",
-    "logmFD",
+    "av mFD",
     "Timepoint x Group"
@@
 
-### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)
+#### Two timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+
+```{r calc rew two tp adjusted 3, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvRew2tpadj=anova(R1,R0)
+```
+
+```{r qa check Rew model age,sex,mFD,baselineBMI,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r Rew2tpadj3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 8, 4, 5, 6,7),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "av mFD",
+    "baseline BMI",
+    "Timepoint x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Three time points, without adjustments (Model CA1)
+
 ```{r calc rew three tp, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
-anvRew3tp=anova(R1,R0)
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC)
+anvRew2tp=anova(R1,R0)
 ```
-```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", echo=FALSE}
+```{r qa check Rew model age,sex three timepoints, eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+```
+
+```{r Rew3tp results, caption="Interaction of group and time on reward network connectivity for two time points", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Timepoint FU2",
+    "Group BARS",
+    "Timepoint FU x Group", 
+    "Timepoint FU2 x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Three timepoints, adjusted for age, sex and average mFD (Model CA2)
+
+```{r calc rew three tp adjusted 2, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
+anvRew2tpadj=anova(R1,R0)
+```
+
+```{r qa check Rew model age,sex,mFD, three timepoints,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r Rew3tpadj 2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
     "Sex",
-    "logmFD",
+    "av mFD",
     "Timepoint FU x Group", 
@@
   table = FALSE,
-  use.packages = FALSE
+  use.packages = TRUE
 )
 ```
+
+#### Three timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+
+```{r calc rew three tp adjusted 3, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC)
+anvRew2tpadj=anova(R1,R0)
+```
+
+```{r qa check Rew model age,sex,mFD,baselineBMI, three timepoints,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r Rew3tpadj 3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 4,9,10, 5, 6, 7,8),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Timepoint FU2",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "av mFD",
+    "baseline BMI",
+    "Timepoint FU x Group", 
+    "Timepoint FU2 x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
 ```{r Rew conn over three timepoints, echo=FALSE}
@@
 
-## Effects of average and change BMI on FC
 ### DMN 
-#### Model 2a and 2b (adjusting for mFD)
+#### Two time points, without adjustments (Model CA1)
+
+```{r calc rew two tp DMN, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tp=anova(R1,R0)
+```
+```{r qa check DMN model age,sex, eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+```
+
+```{r DMN2tp results, caption="Interaction of group and time on DMN connectivity for two time points", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Timepoint x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Two timepoints, adjusted for age, sex and average mFD (Model CA2)
+
+```{r calc DMN two tp adjusted 2, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tpadj=anova(R1,R0)
+```
+
+```{r qa check DMN model age,sex,mFD,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r DMN2tpadj2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 7, 4, 5, 6),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "av mFD",
+    "Timepoint x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Two timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+
+```{r calc DMN two tp adjusted 3, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+anvDMN2tpadj=anova(R1,R0)
+```
+
+```{r qa check DMN model age,sex,mFD,baselineBMI,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r DMN2tpadj3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 8, 4, 5, 6,7),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "av mFD",
+    "baseline BMI",
+    "Timepoint x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Three time points, without adjustments (Model CA1)
+
+```{r calc DMN three tp, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor +  (1|subj.ID), data=final_FC)
+anvDMN3tp=anova(R1,R0)
+```
+```{r qa check DMN model age,sex three timepoints, eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+```
+
+```{r DMN3tp results, caption="Interaction of group and time on DMN connectivity for two time points", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Timepoint FU2",
+    "Group BARS",
+    "Timepoint FU x Group", 
+    "Timepoint FU2 x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Three timepoints, adjusted for age, sex and average mFD (Model CA2)
+
+```{r calc DMN three tp adjusted 2, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC)
+anvDMN3tpadj=anova(R1,R0)
+```
+
+```{r qa check DMN model age,sex,mFD, three timepoints,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r DMN3tpadj 2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 4, 8, 9, 5, 6, 7),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Timepoint FU2",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "av mFD",
+    "Timepoint FU x Group", 
+    "Timepoint FU2 x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+#### Three timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+```{r calc DMN three tp adjusted 3, eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC)
+anvDMN3tpadj=anova(R1,R0)
+```
+
+```{r qa check DMN model age,sex,mFD,baselineBMI, three timepoints,eval=FALSE}
+diagnostics.plot(R0)
+ranef.diagn.plot(R0)
+
+m.stab=glmm.model.stab(model.res = R0)
+#to check whether there were any converge issues:
+table(m.stab$detailed$warnings)
+m.stab$summary
+
+test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
+vif_res=vif(test_vif)
+```
+
+```{r DMN3tpadj 3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R0, R1),
+  reorder.coef = c(1, 2, 3, 4, 9, 10, 5, 6, 7,8),
+  custom.coef.names = c(
+    "Intercept (BL, NBARS)",
+    "Timepoint FU",
+    "Timepoint FU2",
+    "Group BARS",
+    "Age",
+    "Sex",
+    "av mFD",
+    "baseline BMI",
+    "Timepoint FU x Group", 
+    "Timepoint FU2 x Group"
+  ),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+
+```{r DMN conn over three timepoints, echo=FALSE}
+fig_DMNconn <- readRDS(file = "../report/fig/figDMNconn.rds")
+fig_DMNconn
+```
+
+## Exploratory Analyses: Effects of average and change BMI on FC
+### Reward network
+#### Model 2.1 and 2.2 (adjusting for mFD)
+```{r "reward model, mean, change BMI", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
+
+R11 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
+R12a = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
+anvmeanBMI_adjmFD=anova(R11,R12a)
+```
+
+```{r reward model_BMIFD results, caption="Model 2a and 2b for aggregated reward network FC", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(
+  list(R1, R11),
+  custom.coef.names = c("Intercept", "average BMI", "change BMI", "Age", "Sex", "logmFD"),
+  custom.model.names = c("R0", "R1"),
+  float.pos = "h!",
+  center = TRUE,
+  digits = 3,
+  leading.zero = FALSE,
+  single.row = TRUE,
+  include.adjrs = FALSE,
+  include.bic = FALSE,
+  table = FALSE,
+  use.packages = TRUE
+)
+```
+#### Model EA 2.3 (adjusting for average and change in mFD)
+```{r "reward model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvmeanFD=anova(R1,R03)
+anvwithinFD=anova(R1,R03)
+
+p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+```
+
+```{r reward_BMI_splitFD results, caption="Model EA 2.3 for aggregated reward network FC", results="asis", echo=FALSE}
+#tab_model(R1,R0)
+texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h!",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = TRUE)
+```
+
+### DMN
+#### Model 2.1 and 2.2 (adjusting for mFD)
 ```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 
-```{r DMN_BMIFD results, caption="Model 2a and 2b for aggregated DMN FC", results="asis", echo=FALSE}
+```{r DMN_BMIFD results, caption="Model 2.1 and 2.2 for aggregated DMN FC", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
   table = FALSE,
-  use.packages = FALSE
+  use.packages = TRUE
 )
 ```
-#### Model 2c (adjusting for average and change in mFD)
+#### Model 2.3 (adjusting for average and change in mFD)
 ```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
@@
 
-```{r DMN_BMI_splitFD results, caption="Model 2c for aggregated DMN FC", results="asis", echo=FALSE}
+```{r DMN_BMI_splitFD results, caption="Model 2.3 for aggregated DMN FC", results="asis", echo=FALSE}
 #tab_model(R1,R0)
-texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
+texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h!",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = TRUE)
 ```

### 2021-03-04 code/create_sample_df.R

@@ -2,7 +2,7 @@ library(dplyr) # version 1.0.2
 library(tidyr) # version 1.1.2
 
 ConnectMapsPath <- "/data/pt_02161/Results/Project2_resting_state/connectivity/calc_DMN_reward_seed_connectivity"
-DataframePath <- "/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv"
+dfPath <- "/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv"
 
 create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "final"){
   # group = IG/KG/both
@@ -40,6 +40,7 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
           "meanFD",
           "BMI",
           "BMI_BL",
+          "BMI_BLi",
           "exclude_prep"
         )], by = c("subj.ID", "tp"))
       cat("n =", nrow(mri_sample), "fMRI data points, ")
@@ -110,7 +111,7 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
   
   # load info from full sample -------------------------------------------------
   
-  full = read.csv(DataframePath) # CAVE: info file elsewhere !!!
+  full = read.csv(dfPath) # CAVE: info file elsewhere !!!
   full = full[, c(
     "subj.ID",
     "condition",
@@ -140,6 +141,67 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
     }
   }
   
+  
+  # missingness -------------------------------------------------------------
+  
+  VIM::barMiss(full[ ,c("condition","BMI")])
+  VIM::histMiss(full[,c("BMI_BL","BMI")])
+  
+  
+  # mice for imputation in multilevel data ---------------------------------------
+  
+  # see: https://stackoverflow.com/questions/47950304/random-effects-in-longitudinal-multilevel-imputation-models-using-mice
+  # imputation based on following variables
+  tmp <- full[,c("subj.ID","condition", "tp", "Age_BL", "BMI")]
+  tmp$subj.ID <- as.integer(tmp$subj.ID)
+  
+  # set imputation method
+  impmethod <- c("", "","", "", "2l.pan") # why 2l.pan?
+  names(impmethod) <- colnames(tmp)
+  
[…]

### 2021-03-04 code/process.R

@@ -32,68 +32,6 @@ final <- create_sample_df(group = "both", tp = "all")
 full <- create_sample_df(group = "both", tp = "all", out = "full")
 
 
-# missingness -------------------------------------------------------------
-
-VIM::barMiss(full[ ,c("condition","BMI")])
-VIM::histMiss(full[,c("BMI_BL","BMI")])
-
-
-# mice for imputation in multilevel data ---------------------------------------
-
-# see: https://stackoverflow.com/questions/47950304/random-effects-in-longitudinal-multilevel-imputation-models-using-mice
-# imputation based on following variables
-tmp <- full[,c("subj.ID","condition", "tp", "Age_BL", "BMI")]
-tmp$subj.ID <- as.integer(tmp$subj.ID)
-
-# set imputation method
-impmethod <- c("", "","", "", "2l.pan") # why 2l.pan?
-names(impmethod) <- colnames(tmp)
-
-# set default predictor matrix
-ini <- mice::mice(tmp, maxit=0)
-
-# create imputation model for BMI, predictor matrix (following 'type')
-pred <- ini$predictorMatrix
-pred["BMI",] <- c(-2, 1, 2, 1, 0) 
-# (cluster variable, fixef with randint, fixef with randint + randslo, dependent)
-
-dfs = 50
-iter = 10
-imp <- mice::mice(
-  tmp,
-  method = impmethod,
-  pred = pred,
-  maxit = iter,
-  m = dfs,
-  seed = 8745
-)
-imp$predictorMatrix
-summary(imp)
-
-# compute mean from imputed values in complete datasets 
-com <- complete(imp, "repeated")
-com <- cbind(tmp,com[,(ncol(com)-(dfs-1)):ncol(com)])
-meanval <- apply(com[,(ncol(tmp)+1):ncol(com)],1,mean)
-sdval <- apply(com[,(ncol(tmp)+1):ncol(com)],1,sd)
-
-# impute data only for baseline (bl) and construct BMI_BLi
-full$BMI_BLi <- meanval # imputed data for all tp
-full$BMI_BLi[full$tp != "bl"] <- NA
-full$BMI_BLi[full$tp == "bl"] <-
-  meanval[full$tp == "bl"]
-for (i in 1:nrow(full)) {
-  # check if bl exist for subject
-  if (any(full$tp[full$subj.ID == full$subj.ID[i]] == "bl")) {
-    # replace with value of same subject at bl
-    full$BMI_BLi[i] <-
[…]

### 2021-03-04 manuscript/draft_rsfmri.Rmd
Fraukes change in draft + supplement

@@
 ---
-title: "Effects of bariatric surgery loss on resting state functional connectivity of the reward and default mode network: a pre-registered analysis"
+title: "Effects of bariatric surgery loss on resting-state functional connectivity of the reward and default mode network: a pre-registered analysis"
 author:
@@
 Precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
-Resting-state functional magnetic resonance imaging (rsfMRI) is a technique to investigate the dynamic organization of the brain.
+Resting-state functional magnetic resonance imaging (rsrsfMRI) is a technique to investigate the dynamic organization of the brain.
 Functional connectivity (FC) networks, i.e.
@@
 Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
-Resting state fMRI studies also shown increased local FC of re-ward network regions, i.e.
+resting-state rsfMRI studies also shown increased local FC of re-ward network regions, i.e.
 NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
@@
 Higher BMI and obesity have been associated with a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
-salience and sensory networks in several resting state and task-based fMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
+salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
 Studies in bariatric surgery patients have pointed to a normalization of the connectivity between  DMN and cognitive control and salience brain regions after the procedure, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
@@
 Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].
-Most resting-state fMRI studies in obesity have reported an association of higher BMI and higher head motion during MRI [@Beyer_2017;@Hodgson_2016].
-Head motion is a major confounder in resting-state fMRI as it increases short-range and disruptes long-range connectivity FC.
+Most resting-state rsfMRI studies in obesity have reported an association of higher BMI and higher head motion during MRI [@Beyer_2017;@Hodgson_2016].
+Head motion is a major confounder in resting-state rsfMRI as it increases short-range and disruptes long-range connectivity FC.
 In this sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
@@
 The study was registered at clinicaltrials.gov as NCT01554228.
-We preregistered the present resting state analyses on the OSF https://osf.io/yp42s.
-We made additional changes https://osf.io/59bh7/ to the preregistration after preprocessing as we realized some aspects of the analysis were inadequately described in the initial preregistration.
+We preregistered the present resting-state analyses on the OSF https://osf.io/yp42s.
+We made additional changes https://osf.io/59bh7/ to the pre-registration after preprocessing as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
 <!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA -->
+
 Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin.
@@
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score < 34).
+Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score < $34$).
  In total, $51$ participants out of the originally enrolled $69$ subjects received MRI, and we analyzed a total of $101$ data points.
@@
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016].
-Analyses were performed on all participants that provided at least one data point of fMRI data.
-In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point
+Analyses were performed on all participants that provided at least one data point of rsfMRI data.
+In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point.
+
 **N patients underwent RYGB, N underwent VSG and N GB.**
-Patients arrived in the morning (between 07:00 and 12:00) after an overnight fast.
+
+Patients arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast.
 They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
@@
 
-## fMRI acquisition
+## RsrsfMRI acquisition
 
@@
 <!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
-<!-- rsfMRI --> Resting state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms.
+<!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms.
 The image matrix was 64 x 64 with an in-plane resolution of 3 mm x 3 mm and 34 slices with a slice thickness of 4 mm.
 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes.
-Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneties.
+Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneities.
 Participants were instructed to close their eyes but to remain awake during scanning.
@@
 The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
-*In the preregistration it was only mentioned that longitudinal FreeSurfer would be used, but no details regarding the exact analysis.*
 
@@
 In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
-We calculated mean framewise displacement (mFD) from the outputs of the motion correction procedure according to @Power_2012.
 For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
-<!-- *This part of the analysis was described in the preregistration as "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated."*   -->
+
+## Denoising pipelines
+Among several denoising pipelines, ICA-AROMA provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relation-ship [@Parkes_2018].
+[@Ciric_2017] showed that GSR is very efficient in re-moving the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+Yet, it introduces spatial dependency and spurious correlations.
+Thus, we pre-registered the main analyses to be performed with these two denoising pipelines.
 
 ### ICA-AROMA preprocessing   
-We used ICA-AROMA for fMRI denoising as it provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relationship [@Parkes_2018].
-As described in [@pruim2015ica], the minimally preprocessed data was processed with independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
-Finally, the timeseries of the  nuisance components were regressed from the raw fMRI timeseries.
+As described in [@pruim2015ica], the minimally preprocessed data underwent independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+Finally, the timeseries of the  nuisance components were regressed from the raw rsfMRI timeseries non-aggressively.
 As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result.
@@
 After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
-*This analysis was described in the preregistration as "Then we will employ state-of-the art nuisance regression (i.e.
-ICA-AROMA with compCor regression (CC) and, optionally global signal regression (GSR) (Ciric et al., 2017; Parkes et al., 2018) to reduce the effect of physiological noise and head motion on the connectivity estimates."*
 
-### Global Signal regression   
-@Ciric_2017 showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
-Yet, it introduces spatial dependency and spurious correlations.
-Previously, we reported on head motion differences between groups over time in this study [@Beyer_2020], and thus it was very important to minimize the motion-FC correlation.
+### Global Signal Regression (GSR) 
 The global signal was derived from the average of all voxels in the brain mask.
-Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA fMRI data and high-pass filtered the resulting file as above.
-*In the preregistration we stated " As global signal regression is very efficient in removing the correlation of head motion and connectivity dependence, but introduces spatial dependency and spurious correlations, we perform GSR after ICA-AROMA in a separate sensitivity analysis."*
+Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA rsfMRI data and high-pass filtered the resulting file as above.
 
-### Functional connectivity
-To derive reward and DMN functional connectivity maps, we used Nucleus Accumbens (NAcc) and precuneus as seed regions of interest (ROI), respectively.
-We chose these regions as they are core hubs of the networks.
-Because we knew about the large amount of motion in this dataset, we decided against Independent Component Analysis (ICA) to construct the networks, as ICA is recommended to work on minimally preprocessed data [GSR data](https://sourceforge.net/p/icatb/mailman/message/31466068/).
-Based on FreeSurfer's `aseg.mgz` and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
-Then, we used `NiftiLabelsMasker` and `NiftiMasker` to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson's correlation between them with `numpy.dot`, performed `r-to-z` transform and saved the resulting correlation maps for each processing mode (minimally preprocessed, aroma+cc, aroma+gsr).
-Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
-*In the preregistration we stated "Resting-state functional connectivity (FC) using seed-based connectivity analysis extracted with in-house pre-processing pipelines (see (Beyer et al., 2019)): Whole-brain FC maps for a priori defined seed regions (defined based on the results of the subcortical segmentation and Desikan-Killiany cortical parcellation in the Freesurfer version 6.0.0 longitudinal processing stream).
-For the reward network a ROI combined from bilateral Nucleus accumbens from the subcortical segmentation (26, Left-Accumbens-area, 58, Right-Accumbens-area) was selected.
-Ventro-medialprefrontal cortex was not selected as seed region due to low signal-to-noise ratio in this part of the brain in many participants (noticed during quality checking).
-For the default mode network, the seed region was combined from bilateral precuneus from the Desikan-Killiany atlas (1025 ctx-lh-precuneus, 2025 ctx-rh-precuneus).
-We will calculate  voxel-wise seed-based connectivity mapsin the individual subject’s space (which is coregistered to the Freesurfer longitudinally preprocessed timepoint) and apply r-to-Z-transform to these images.
-These steps will be performed using nilearn, nipype and python.
-Then, we will transform the connectivity maps into MNI space with a non-linear warp derived from anatomical preprocessing of the longitudinally preprocessed timepoints (using ANTS).*
+### Functional connectivity (FC)
+To derive RN and DMN FC maps, we used nucleus accumbens (NAcc) and precu-neus as seed regions of interest (ROI), respectively.
+We did not select vmPFC for the RN due to low SNR.
+Because we knew about the large amount of motion in this dataset beforehand, we decided against ICA to construct the networks, as ICA is recommended to work on minimally preprocessed data.
+Based on FreeSurfer’s aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
+Then, we used NiftiLa-belsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
+We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each processing mode (minimally preprocessed, aro-ma+cc, aroma+gsr).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS dur-ing anatomical preprocessing.
 
@@
 #### Anatomical Quality control
-
 FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
@@
 <!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
-<!-- *In the preregistration we stated that "Freesurfer segmentations and surfaces will be visually controlled and corrected according to published recommendations, if necessary", and that "We will exclude participants in which head motion artifacts affected the T1-weighted image preprocessing (e.g.
-caused Freesurfer to fail or perform very poorly (rated “Failed” according to (Klapwijk et al., 2019)".* -->
 
-#### Functional Quality control
-Resting state fMRI data quality control was performed according to the protocol by @Ciric_2018.
-As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020].
-Thus, we did not exclude any participants based on mFD estimates beyond those who failed the anatomical preprocessing.
-We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
-In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mFD.
+#### Resting-state fMRI quality control
+Resting-state rsfMRI quality control was performed according to the protocol by @Ciric_2018. 
+Head motion is a major confounder in rsfMRI. We calculated mean framewise displacement (mFD) from the outputs of `MCFLIRT` according to @Power_2012.
+As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
+Instead, we applied best-practice motion correction (AROMA+CC and AROMA+CC+GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
+
+**Individual level resting-state rsfMRI QC**
 ```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
@@
 ```
-On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
+On the individual level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
 Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum).
 All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
-We evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012].
+
+**Group level resting-state rsfMRI QC**
+
+First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
+```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
+figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
+#(figList$fig1 | figList$fig2) / figList$fig3
+
+lay <- rbind(c(1,1,2,2),
+             c(NA,3,3,NA))
+gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
+```
+
+As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group.
+Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points (see Figure \@ref(fig:figDvarsmFD).
+Following the protocol by [@Ciric_2018], we evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012].
 In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histograms (see supplements).
-We reported the mFD-FC correlation per preprocessing scheme.
 Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.
-## Quality control
-
-We performed different quality controls for the rsfMRI data.
-First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
-As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group.
-Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points.
-We concluded that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period in both groups.
 ```{r mFDFC, echo=FALSE}
@@
 ```
-Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018].
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`).
-This relationship was negatively associated with distance between nodes (mean Spearman's $r$ = `r round(res[1,5],2)`).
-Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
-Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
-Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r$ = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r$ = `r round(res[4,5],2)`).
-Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.
 ```{r tableQcfc, out.width="100%", echo=FALSE}
@@
 ```
-
-```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
-figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
-#(figList$fig1 | figList$fig2) / figList$fig3
-
-lay <- rbind(c(1,1,2,2),
-             c(NA,3,3,NA))
-gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
-```
+Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes were significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`).
+This relationship was negatively associated with distance between nodes (mean Spearman's $r$ = `r round(res[1,5],2)`).
+Denoising pipelines (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
+Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
+Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r$ = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r$ = `r round(res[4,5],2)`)
+In addition, for models showing a significant time-by-group interaction, we planned to perform a sensitivity analysis excluding the 10% data sets with highest mFD.
 
@@
 As preregistered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average mFD of both timepoints (CA2) and age,sex, average mFD and baseline BMI (CA3).
- In this model, time was represented as factor because we took into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management (Olivo et al.
-2017).
+ In this model, time was represented as factor because we took into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
 The resulting factorial design contained one regressor for each time point per group (see supplements for depiction of design matrix).
@@
 Here, we used individual mFD values (not averaged) as covariate to capture variance in mFD change over time points.We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and mFD).
+
 ##### Average and change BMI
@@
 For each brain network, we examined the models for the two denoising pipelines (AROMA+CC / AROMA+CC+GSR)as preregistered.
-For all analyses, we used an explicit brain mask, which was a whole brain derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).<!--  This resulted in a number of tests (voxels of interest) from $52378$.
-MNI ICBM 2009a Nonlinear Symmetric grey matter mask ([http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009](http://www.bic.mni.mcgill.ca/ServicesAtlases/ICBM152NLin2009)) were re-sampled to 3 x 3 x 3 mm$^3$ and binarized at threshold $0.5$.
--->
+For all analyses, we used an explicit brain mask, which was a whole brain derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
 
@@
 <!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
-Deviating from the preregistration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual.
-Opposed to the description in the preregistration, we used a cluster forming threshold of p<0.001 for a more rigorous multiple comparison adjustment (instead of p<0.01, and 1000 bootstraps due to required computation time (instead of 5000)).
+Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual.
+Opposed to the description in the pre-registration, we used a cluster forming threshold of p<0.001 for a more rigorous multiple comparison adjustment (instead of p<0.01, and 1000 bootstraps due to required computation time (instead of 5000)).
 Significant clusters are defined as family-wise error (FWE) corrected p<0.05.
@@
 
-In this pre-registered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design.
+In this pre-registered study, we investigated the effects of bariatric surgery on the FC of major resting-state brain networks in a longitudinal controlled design.
 Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) on the group level.
@@
 Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI scanner.
-While this sample size is comparable to previous resting state fMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
+While this sample size is comparable to previous resting-state rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
 
@@
 We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized (?) breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow separate analyses for the sexes.
@@
 
-Taken together, this study showed that effects of bariatric surgery on resting state are subtle, and often confounded by head motion.
+Taken together, this study showed that effects of bariatric surgery on resting-state are subtle, and often confounded by head motion.
 Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to observed changes in food wanting and choice in patients after bariatric surgery.
@@
 
-All code is available in the github repository [https://github.com/hsx1/adi2_rsfmri](https://github.com/hsx1/adi2_rsfmri).
+All code is available in the github repository [https://github.com/hsx1/adi2_rsrsfMRI](https://github.com/hsx1/adi2_rsrsfMRI).
 The code includes script for data preparation and analysis of different models.

### 2021-03-04 manuscript/supplements_rsfmri.Rmd
Fraukes change in draft + supplement

@@
 final_FC$group_factor=relevel(final_FC$group_factor, "KG")
-final_FC=merge(final_FC,subset(final_FC, tp=="bl",select=c("subj.ID", "BMI")), by='subj.ID')
-colnames(final_FC)[27]="BMI.bl"
 ```
@@
 #### Two time points, without adjustments (Model CA1)
-
-<b>
 

### 2021-03-04 code/run_or_display.m

@@ -107,7 +107,7 @@ param.PRESET = "manual";
 param.ONLY_DISPLAY = false;      
 param.OVERWRITE = true; 
 param.WILD_BOOT = true;   
-param.parallel = true;
+param.parallel = false;
 
 if param.PRESET == "standard"
     param.MODEL = ["grouptime","grouptime2tp", "bmi", "bmiIG","bmi2tp", "fd","fdIG"];
@@ -168,7 +168,7 @@ elseif param.PRESET == "full"
 elseif param.PRESET == "test"
     param.MODEL = ["grouptime"];
     param.ROI_PREP = roi_prep([4, 6, 12, 14]); 
-    param.COVARIATES = [13];  
+    param.COVARIATES = [14];  
     param.MASK = "brain";           
     param.EXCLFD = false;        
     param.INFERENCE_TYPE = ["cluster"];

### 2021-03-05 manuscript/draft_rsfmri.Rmd
changes of VW and EM added by FB

@@
 Effective treatment can be achieved by bariatric surgery, yet its effects on functional connectivity of brain networks involved in reward processingand intereoception are still under debate.
-In this pre-registered analysis, we aimed to determine the effects of bari-atric surgery and weight loss on functional connectivity in the reward network and default mode network (DMN.
+In this pre-registered analysis, we aimed to determine the effects of bari-atric surgery and weight loss on functional connectivity in the RN and default mode network (DMN.
 Therefore, 33 bariatric surgery patients (males/females, BMI, age etc) and 15 obese waiting-list control patients (males/females, BMI, age etc) underwent MRI at baseline (before surgery/waiting period), and after 6 and 12 months.
@@
 There also was no interaction effect on within network FC in pre-registered analysis of aggregated FC.
-Exploratory analyses also did not show an interaction effect for three time points, but when adding differences in head motion within and between individuals using mean framewise displacement, surgery-related BMI decrease and higher average head motion resulted in stronger functional connectivity in medial posterior frontal regions within the reward network.
+Exploratory analyses also did not show an interaction effect for three time points, but when adding differences in head motion within and between individuals using mean framewise displacement, surgery-related BMI decrease and higher average head motion resulted in stronger functional connectivity in medial posterior frontal regions within the RN.
 This pre-registered analysis of longitudinal neuroimaging data did not support a strong effect of bariatric surgery, compared to waiting, on reward and DMN network functional connectivity after 6 months in obese patients.
@@
 Precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
-Resting-state functional magnetic resonance imaging (rsrsfMRI) is a technique to investigate the dynamic organization of the brain.
+Resting-state functional magnetic resonance imaging (rsfMRI) is a technique to investigate the dynamic organization of the brain.
 Functional connectivity (FC) networks, i.e.
 brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009].
-The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
-The reward network (RN), comprises the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
+The RN (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
+The RN (RN), comprises the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
 These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
-Frequently, obesity has been associated with hyperactivation of reward network regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
+Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
 resting-state rsfMRI studies also shown increased local FC of re-ward network regions, i.e.
@@
 Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] ..
- The default mode network includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] .
+ The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
 BA 39) [@Raichle_2015].
@@
 Studies in bariatric surgery patients have pointed to a normalization of the connectivity between DMN and cognitive control and salience brain regions after the procedure, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
-In sum, while there is some evidence hinting to a role for DMN and RN connectivity in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
+In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
 Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].
-Most resting-state rsfMRI studies in obesity have reported an association of higher BMI and higher head motion during MRI [@Beyer_2017;@Hodgson_2016].
-Head motion is a major confounder in resting-state rsfMRI as it increases short-range and disruptes long-range connectivity FC.
+Most rsfMRI studies in obesity have reported an association of higher BMI and higher head motion during MRI [@Beyer_2017;@Hodgson_2016].
+Head motion is a major confounder in rsfMRI as it increases short-range and disrupts long-range connectivity FC.
 In this sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
 Here, we thus aimed to rigorously control for motion-related variance.
-In this pre-registered study, our hypotheses were:
-1. Whole-brain FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to waiting-list control group
-2. Within-network FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to waiting-list control group
+In this pre-registered study, our confirmatory hypotheses were:
+1. Whole-brain FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to a waiting-list control group
+2. Within-network FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to a waiting-list control group
 We tested hypothesis 1 by investigating the interaction of bariatric surgery and time on DMN and RN whole-brain FC.
@@
 <!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité University Medicine Berlin, Germany.
-For more details see Prehn et al.
-(2020).
+For more details see [@Prehn_2020].
 We used all data acquired until April 2019.
@@
 The study was registered at clinicaltrials.gov as NCT01554228.
-We preregistered the present resting-state analyses on the OSF https://osf.io/yp42s.
+We pre-registered the present resting-state analyses on the OSF https://osf.io/yp42s.
 We made additional changes https://osf.io/59bh7/ to the pre-registration after preprocessing as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
@@
 Analyses were performed on all participants that provided at least one data point of rsfMRI data.
-In more detail, `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point.
+`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point (for more details see Table S1).
+The pre-planned analysis of changes from baseline to follow-up included $24$ participants with both time points, in total $72$ data points.
 
@@
 
-Patients arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast.
+Participants arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast.
 They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
@@
 
-## RsrsfMRI acquisition
+## MRI acquisition
 
@@
 Among several denoising pipelines, ICA-AROMA provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relation-ship [@Parkes_2018].
-[@Ciric_2017] showed that GSR is very efficient in re-moving the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+@Ciric_2017 showed that GSR is very efficient in re-moving the correlation of head motion and FC, and outperforms ICA-AROMA alone.
 Yet, it introduces spatial dependency and spurious correlations.
@@
 
-### ICA-AROMA preprocessing   
-As described in [@pruim2015ica], the minimally preprocessed data underwent independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+### ICA-AROMA denoising pipeline   
+As described in @pruim2015ica, the minimally preprocessed data underwent independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
 Finally, the timeseries of the  nuisance components were regressed from the raw rsfMRI timeseries non-aggressively.
@@
 We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each processing mode (minimally preprocessed, aro-ma+cc, aroma+gsr).
-Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS dur-ing anatomical preprocessing.
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
 
@@
 
-#### Resting-state fMRI quality control
-Resting-state rsfMRI quality control was performed according to the protocol by @Ciric_2018. 
-Head motion is a major confounder in rsfMRI. We calculated mean framewise displacement (mFD) from the outputs of `MCFLIRT` according to @Power_2012.
+#### RsfMRI quality control
+RsfMRI quality control was performed according to the protocol by @Ciric_2018. 
+Head motion is a major confounder in rsfMRI. We calculated mean framewise displacement (mFD) according to @Power_2012, and log-transformed it for further analysis (logmFD).
 As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
@@
 
-**Individual level resting-state rsfMRI QC**
+##### Individual level resting-state rsfMRI QC
 ```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
@@
 Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum).
-All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
+All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`, and checked again visually.
+No participant was excluded based on this visual inspection.
 
-**Group level resting-state rsfMRI QC**
+##### Group level resting-state rsfMRI QC
 
-First, we inspected parameters of head motion (mFD) and a measure of whole-brain signal variation which is reflective of noise in the data (DVARS).
+As described in @Beyer_2020, there was a group-by-time interaction on mFD (see Figure \@ref(fig:figDvarsmFD), upper right panel).
+The intervention group reduced in head motion in the follow-up assessments compared to the control group.
+For this study, we further explored mFD and its relation to DVARS, the root mean square of the temporal change of the voxel-wise signal at each time point.
 ```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
@@
 ```
-
-As described in @Beyer_2020, there was a group-by-time interaction on mFD. The intervention group reduced in head motion in the follow-up assessments compared to the control group.
-Further, DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points (see Figure \@ref(fig:figDvarsmFD).
-Following the protocol by [@Ciric_2018], we evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012].
-In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.
+DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points (see Figure \@ref(fig:figDvarsmFD).
+To further assess the confounding of head motion and FC, we evaluated mFD-FC relationships by computing a functional connectome based on the protocol by [@Ciric_2018].
+We used the MNI coordinates of 246 cortical and subcortical spherical seed regions from @Power_2012 with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
+Then, we calculated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation.
 ```{r mFDFC, echo=FALSE}
@@
 ```
-Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes were significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`).
+In the minimally preprocessed data, mFD and FC between nodes were significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`)(see Table \@ref(tab:tableQcfc)).
 This relationship was negatively associated with distance between nodes (mean Spearman's $r$ = `r round(res[1,5],2)`).
-Denoising pipelines (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency.
+Denoising pipelines (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and FC, and reduced the distance dependency.
 Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and reward network over all participants and NAccs, adjusted for age and sex.
+To extract aggregated FC values, we first calculated the mean DMN and RN over all participants and NAccs, adjusted for age and sex.
 We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
 Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
-We extracted the average connectivity from these masks for GSR regressed data.
+We extracted the average FC from these masks for GSR-data.
 
@@
 #### Confirmatory Analysis
-We tested the preregistered hypothesis of a  time by group interaction:  $FC=group + time + group * time$. 
-As preregistered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average mFD of both timepoints (CA2) and age,sex, average mFD and baseline BMI (CA3).
+We tested the pre-registered hypothesis of a  time by group interaction:  $FC=group + time + group * time$. 
+As pre-registered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average logmFD of both timepoints (CA2) and age,sex, average logmFD and baseline BMI (CA3).
  In this model, time was represented as factor because we took into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
@@
 Then, we extended our whole-brain analysis of the group-by-time point interaction to the data from all three timepoints.
-This analysis had not been preregistered.
-Here, we used individual mFD values (not averaged) as covariate to capture variance in mFD change over time points.We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and mFD).
+This analysis had not been pre-registered.
+Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points.We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and logmFD).
 
@@
 This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for log mean mFD (Model EA 2.2).
+We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for log mFD (Model EA 2.2).
 As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 2.3c).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: $FC = between-subject FD + within-subject FD$ age and sex as nuisance covariates.
+For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: $FC = between-subject logmFD + within-subject logmFD$ age and sex as nuisance covariates.
 
@@
 ### Model input and brain mask
-For each brain network, we examined the models for the two denoising pipelines (AROMA+CC / AROMA+CC+GSR)as preregistered.
+For each brain network, we examined the models for the two denoising pipelines (AROMA+CC/ AROMA+CC+GSR)as pre-registered.
 For all analyses, we used an explicit brain mask, which was a whole brain derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
@@
 <!-- R analysis of aggregated values -->
-We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
+We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007].
 We deviated from the pre-registration by only investigating data from the GSR denoising pipeline.
 First, we investigated the time-by-group interaction for baseline and follow-up time points only.
-We either performed no correction (CA1), adjusted for baseline age, sex, average of mean FD of both time points (CA2) or additionally adjusted for baseline BM (CA3)I .
+We either performed no correction (CA1), adjusted for baseline age, sex, average of logmFD of both time points (CA2) or additionally adjusted for baseline BM (CA3)I .
 We performed model comparison between R1 and R0 models, where R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj)), and R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj)).
@@
 We performed an exploratory analysis with average BMI and variability in BMI as predictors of the aggregated FC from GSR data of both networks, in line with our exploratory whole brain analysis.
-We calculated three models with average and change in BMI as predictors of inter-est and adjusted for baseline age and sex (Model EA2.1), mFD (Model EA 2.2) and average and change mFD (Model EA 2.3).
+We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (Model EA2.1), logmFD (Model EA 2.2) and average and change logmFD (Model EA 2.3).
 
@@
 ## Whole-brain analyses: Confirmatory analysis 
-First, we performed the preregistered analysis in participants with baseline and followup data, using two denoising pipelines and three covariate adjustements.
+First, we performed the pre-registered analysis in participants with baseline and followup data, using two denoising pipelines and three covariate adjustements.
+
 ### Model CA1 (without adjustments)
@@
 
-###  Model CA2 (adjusted for age, sex and average of mFD)
+###  Model CA2 (adjusted for age, sex and average of logmFD)
 Similarly, there were no significant clusters when adjusting for age, sex and average of logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
 
-###  Model CA3 (adjusted for age, sex, average of mFD and baseline BMI)
-There was no significant interacion when adding baseline BMI.
+###  Model CA3 (adjusted for age, sex, average of logmFD and baseline BMI)
+There was no significant interaction when adding baseline BMI.
 
@@
 ### Group-by-time interaction for three time points (EA 1.1 and EA 1.2)
-Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA 1.1 and EA 1.2 in the full data set for neither reward network nor DMN.
+Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA 1.1 and EA 1.2 in the full data set for neither RN nor DMN.
 
 ### Whole-brain effects of mean and change in BMI
-##### Reward network
+#### RN
 
-###### Average and change in BMI, adjusted for logmFD (EA 2.1 and EA 2.2)
+##### Average and change in BMI, adjusted for logmFD (EA 2.1 and EA 2.2)
 NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for logmFD.
 
-###### Average and change in BMI, adjusted for split mFD (EA 2.3)
-<!-- cc -->For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region. 
+##### Average and change in BMI, adjusted for split logmFD (EA 2.3)
+<!-- cc -->For the RN, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region. 
 The peak voxel derived from investigating the effect of average logmFD was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), even if data was GSR ($p = 0.038$).
-```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Cluster showing a significant negative effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 This result was similar for GSR corrected data.
-```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average logmFD on connectivity between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+```{r "BMIFDagesexNACCavgFDact", fig.cap="Cluster showing a significant positive Effect of average logmFD on FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
@@
 ```
-<!-- Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
+<!-- RN connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
 Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
@@
 
-##### DMN
+#### DMN
 
-###### Average and change in BMI, adjusted for mFD (EA 2.1 and EA 2.2)
+##### Average and change in BMI, adjusted for mFD (EA 2.1 and EA 2.2)
 Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCccavgBMIdeact)).
-When using GSR preprocessed data, none of the clusters was significantly associated with average BMI.
+When using GSR-data, none of the clusters was significantly associated with average BMI.
 Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
@@
 
-###### Average and change in BMI, adjusted for split mFD (EA 2.3)
+##### Average and change in BMI, adjusted for split logmFD (EA 2.3)
+
 To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
 <!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
-<!-- gsr -->As before, the effect did not survive when using GSR-corrected data (see Table \@ref(tab:tableBMIFDmodel).
+<!-- gsr -->As before, the effect did not survive when using GSR-data (see Table \@ref(tab:tableBMIFDmodel).
 ```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
@@
 
-##### Effect of mFD (EA 2.4)
+##### Functional decoding
+Functional decoding of these two activation patterns showed different top association terms.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
+premotor cortex were the meta-analytic activation maps most similar to this contract.
+The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
+
+##### Effect of logmFD (EA 2.4)
 
@@
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
-<!-- ### DMN -->We did not find a any clusters for functional connectivity with the PCC.
+<!-- ### DMN -->We did not find any clusters with a significant association of logmFD and PCC connectivity.
 ```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
@@
 ```
-
-##### Functional decoding
-Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
-premotor cortex were the meta-analytic activation maps most similar to this contract.
-The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
 
@@
 final_FC <- read.csv("../report/final_FC.csv")
-final_FC=merge(final_FC,subset(final_FC, tp=="bl",select=c("subj.ID", "BMI")), by='subj.ID')
-colnames(final_FC)[28]="BMI.bl"
 ```
 
-```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the Reward network"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
+```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the RN"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
 fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
@@
 ```
-There was no significant group-by-time interaction for aggregated DMN and reward network FC (see Figure @ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless of the adjustments (CA1: without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
-Thus, we did not repeat the analysis with-out the 10% of highest motion subjects, as preregistered.
+There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure @ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless of the adjustments (CA1: without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
+Thus, we did not repeat the analysis with-out the 10% of highest motion subjects, as pre-registered.
 
 ### Exploratory analysis: effects of mean and change in BMI on aggregated FC
-#### Reward network
+#### RN
 ##### Models EA 2.1, 2.2. and 2.3
@@
 ```
-As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and within reward network connectivity in models 2a, 2b or 2c adjusting for age, sex and logmFD.
+
+As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and within RN connectivity in models EA 2.1, 2.2 or 2.3 adjusting for age, sex and logmFD.
+
 #### DMN
+
 ##### Models EA 2.1, 2.2. and 2.3
+
 ```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
@@
 anvmeanBMI_adjmFD=anova(R1,R02a)
+p_meanBMI_adjmFD=anvmeanBMI_adjmFD$`Pr(>Chisq)`[2]
 ```
-Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity ($p$ = `r round(p_meanBMI,3)`), regardless of whether we adjusted for logmFD (model EA 2.1 and EA 2.2).
-The association also remained significant when we split logmFD into average and change in logmFD ($p$ = `r round(p_meanBMI,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (modelc; see supplementary material for a detailed coefficients from all models).
 
@@
 
-p_mean_BMI=anvmeanBMI$`Pr(>Chisq)`
+p_mean_BMI_splitmFD=anvmeanBMI$`Pr(>Chisq)`
 ```
+Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity, regardless of whether we adjusted for logmFD (model EA 2.1: $p$ = `r round(p_meanBMI,3)` and EA 2.2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
+The association also remained significant when we split logmFD into average and change in logmFD (EA 2.3) ($p$ = `r round(p_mean_BMI_splitmFD,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (see supplements material for overview of all models).
 
@@
 In this pre-registered study, we investigated the effects of bariatric surgery on the FC of major resting-state brain networks in a longitudinal controlled design.
-Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) on the group level.
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising pipelines and controlling for mFD on the group level.
 We found no interaction effect of group and time point on whole-brain FC of the PCC and NAcc.
-Neither did we observer main effects of group and time point on the FC of these DMN and reward network hubs, respectively.
+Neither did we observer main effects of group and time point on the FC of these DMN and RN hubs, respectively.
 Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
-In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme.
+In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising pipeline.
 When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI across the three time points and the connectivity of NAcc with a posterior-medial frontal cluster emerged.
-This result was significant in both denoising schemes.
+This result was significant in both denoising pipelines.
 Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns.
@@
 
-In this study, we could not confirm our preregistered hypotheses and found no interaction effect between group and time point on DMN and reward network connectivity.   
+In this study, we could not confirm our pre-registered hypotheses and found no interaction effect between group and time point on DMN and RN connectivity.   
 Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
-Further, we expected FC between reward network regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in reward network regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019].
+Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019].
 However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the NAcc and a cluster in a posterior-medial frontal brain region.
@@
 Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
-Importantly, a global decrease in connectivity (DC) might go along with selectively increased connectivity, i.e.
+Importantly, a global decrease in connectivity might go along with selectively increased connectivity, i.e.
 a shift from integration towards segregation and more specialized processing.
@@
 Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity.
-Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2019], and therefore conducted careful analyses of the impact of head motion on our results.
-To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and thereby introducing two instead of one regressors into the model.
+Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.
+To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and thereby introducing two instead of one regressor into the model.
 These differential results may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
 Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
-This argument is supported by the survival of the cluster when using global signal regressed data, and the distinct results of the decoding analysis.
+This argument is supported by the survival of the cluster when using GSR-data, and the distinct results of the decoding analysis.
 Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
@@
 Head motion also played a role in the association of higher BMI and reduced DMN connectivity.
-While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN connectivity was negatively associated with BMI in both denoising schemes.
+While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN F was negatively associated with BMI in both denoising schemes.
 Thereby, this results echos a previous finding from our group, and may be interpreted as accelerated age-related decline of the DMN  in relation to the cardiometabolic risk factor BMI.
@@
 We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals [@Thiese_2014].
-Another strength of our study was the preregistered analysis plan, which was corrected after we realized some major flaws in the first version.
+Another strength of our study was the pre-registered analysis plan, which was corrected after we realized some major flaws in the first version.
 In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
@@
 We used seed-based connectivity to derive large-scale brain networks.
-While this approach yielded reasonable DMN and reward network maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
-Furthermore, we did not explicitly investigate the connectivity of the DMN and reward network with other large-scale brain networks attention, e.g.
+While this approach yielded reasonable DMN and RN maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
+Furthermore, we did not explicitly investigate the connectivity of the DMN and RN with other large-scale brain networks attention, e.g.
 cognitive control or sensorimotor networks, which has been previously proposed to be altered by obesity and bariatric surgery [@Lips_2014; @Olivo_2016; @Cerit_2019].
@@
 We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized (?) breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict reward network responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow separate analyses for the sexes.
@@
 Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to observed changes in food wanting and choice in patients after bariatric surgery.
-Further, larger, well-controlled and preregistered neuroimaging studies are necessary to understand the neural underpinnings of altered gut-brain communication after bariatric surgery.
+Further, larger, well-controlled and pre-registered neuroimaging studies are necessary to understand the neural underpinnings of altered gut-brain communication after bariatric surgery.
 
@@
 
-All code is available in the github repository [https://github.com/hsx1/adi2_rsrsfMRI](https://github.com/hsx1/adi2_rsrsfMRI).
+All code is available in the github repository [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
 The code includes script for data preparation and analysis of different models.

### 2021-03-05 manuscript/supplements_rsfmri.Rmd
changes of VW and EM added by FB

@@
 anvRew2tp=anova(R1,R0)
+anvRew2tp
 ```
@@
 
-test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + logmFD, data=final_FC)
+test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex + mean.logmFD, data=final_FC)
 vif_res=vif(test_vif)
@@
 ```{r calc rew two tp adjusted 3, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI_BLi + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI_BLi + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvRew2tpadj=anova(R1,R0)
@@
 
-test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
+test_vif=lm(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI_BLi, data=final_FC)
 vif_res=vif(test_vif)
@@
 ```{r calc rew three tp adjusted 3, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC)
-anvRew2tpadj=anova(R1,R0)
+R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI_BLi + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI_BLi + (1|subj.ID), data=final_FC)
+anvRew3tpadj=anova(R1,R0)
 ```
@@
 ```{r calc DMN two tp adjusted 3, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI_BLi + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI_BLi + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvDMN2tpadj=anova(R1,R0)
@@
 ```{r calc DMN three tp adjusted 3, eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI.bl + (1|subj.ID), data=final_FC)
-R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI.bl + (1|subj.ID), data=final_FC)
+R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + BMI_BLi + (1|subj.ID), data=final_FC)
+R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +BMI_BLi + (1|subj.ID), data=final_FC)
 anvDMN3tpadj=anova(R1,R0)
@@
 m.stab$summary
-
-test_vif=lm(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +  BMI.bl, data=final_FC)
-vif_res=vif(test_vif)
 ```

### 2021-03-08 manuscript/draft_rsfmri.Rmd
count surgeries

@@
 ```{r "To do", include=FALSE}
-# [x] Upload auf Neurovolt, cf. cobidas reporting guidelines
-# [x] Swap mFD with logmFD ??
+# [ ] add grouptime2tp results
+# [ ] check stringency
+# [ ] check model names
+# [ ] add dropout to discussion
+# [ ] add comorbidities to methods
+# [ ] add imputations to methods
 ```
@@
 Obesity is a worldwide health issue, entailing huge personal and societal costs.
-Ex-cess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life (hier eine Ref).
-Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI (> 35 kg/m2).
+Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life (hier eine Ref).
+Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI ($> 35$ kg/m$^2$).
 Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
-RYBG is the most invasive, yet most effective surgical pro-cedure, but other techniques like vertical sleeve gastrectomy (VSG) and gadstric banding (GB)are also common [@Berthoud_2011].
+RYBG is the most invasive, yet most effective surgical pro-cedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].
 Bariatric surgery induces weight loss  in part by malabsorption of nutrients, especially fat, and reduced digestion efficiency.
@@
 brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009].
-The RN (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
-The RN (RN), comprises the ventromedial prefrontal cortex (vmPFC), the striatum (STR), the amygdala (AMY) and the anterior insula (antINS) [@Odoherty_2004;@Liu_2011].
+The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
+The RN (RN), comprises the ventromedial prefrontal cortex (vmPFC), the striatum, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
 These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
@@
 We tested hypothesis 1 by investigating the interaction of bariatric surgery and time on DMN and RN whole-brain FC.
-We pre-registed two denoising pipelines, and three covariate schemes.
+We pre-registered two denoising pipelines, and three covariate schemes.
 For hypothesis 2, we performed a planned analysis of the group-by-time interaction on aggregated, within-network FC, for two and three time points and the same covariate schemes.
@@
                           values_from = c("BMI","meanFD"))
+# patients at baseline
+dfBL <- readRDS(file = "../report/dfBL.rds")
+
+# surgeries
+data.frame(table(subj_df$interv, subj_df$condition, exclude=NULL))
+# why NA ?
+subj_df$intervention_de[is.na(subj_df$interv)]
+sum(subj_df$comorbidities[subj_df$condition=="IG"])
+sum(subj_df$comorbidities[subj_df$condition=="KG"])
+
+subj_df$comorbidities
 ```
@@
 
-**N patients underwent RYGB, N underwent VSG and N GB.**
+`r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the BARS group information was not available.
 
@@
 We tested the pre-registered hypothesis of a  time by group interaction:  $FC=group + time + group * time$. 
-As pre-registered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average logmFD of both timepoints (CA2) and age,sex, average logmFD and baseline BMI (CA3).
+As pre-registered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
 In this model, time was represented as factor because we took into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
@@
 ```{r tableBaselineCharacteristics, include=FALSE}
-#dfBL <- create_sample_df(group = "all", tp = "BL")
-dfBL <- readRDS(file = "../report/dfBL.rds")
-
 # distributions per group

### 2021-03-08 code/create_sample_df.R

@@ -42,6 +42,9 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
           "BMI",
           "BMI_BL",
           "BMI_BLi",
+          "comorbidities",
+          "interv",
+          "intervention_de",
           "exclude_prep"
         )], by = c("subj.ID", "tp"))
       cat("n =", nrow(mri_sample), "fMRI data points, ")
@@ -235,17 +238,27 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
   addit_fu2$tp <- "fu2"
   addit_info <- rbind(addit_bl,addit_fu,addit_fu2)
   rm(addit_bl,addit_fu,addit_fu2)
-  addit_info$intervention <- addit_info$Operationsverfahren
+  addit_info$intervention_de <- addit_info$Operationsverfahren
   addit_info$tp <- as.factor(addit_info$tp)
   addit_info$subj.ID <- as.factor(addit_info$subj.ID)
+
   
-  blub <- merge(full, addit_info, by=c("subj.ID","tp"), all = TRUE)
-  
+  full_surg_comorbid <- merge(full, addit_info, by=c("subj.ID","tp"), all = TRUE)
+  rm(addit_info)
+  # check if KG have 
+  t(table(full_surg_comorbid$condition, full_surg_comorbid$intervention_de))
+  full_surg_comorbid$interv <- NA
+  full_surg_comorbid$interv[stringr::str_detect(full_surg_comorbid$intervention_de, stringr::regex("resektion|schlauch",ignore_case = T)) & !is.na(full_surg_comorbid$intervention_de)] <- "VSG"
+  full_surg_comorbid$interv[stringr::str_detect(full_surg_comorbid$intervention_de, stringr::regex("band",ignore_case = T)) & !is.na(full_surg_comorbid$intervention_de)] <- "GB"
+  full_surg_comorbid$interv[stringr::str_detect(full_surg_comorbid$intervention_de, stringr::regex("bypass",ignore_case = T)) & !is.na(full_surg_comorbid$intervention_de)] <- "RYGB"
+  full_surg_comorbid$interv[stringr::str_detect(full_surg_comorbid$intervention_de, "KG") & !is.na(full_surg_comorbid$intervention_de)] <- "no"
+  full_surg_comorbid[,c("interv", "intervention_de")]
+  full_surg_comorbid$intervention 
   
   
[…]

### 2021-03-08 code/process.R

@@ -20,7 +20,7 @@ library(VIM)
 ROOT_DIR ="/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/code/"
 aggFCDMN_TAB_PATH="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN.txt"
 aggFCDMNID_TAB_PATH="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/DMN/agg_FC_DMN_ID.txt"
-aggFCRewID_TAB_PATH="data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt"
+aggFCRewID_TAB_PATH="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew_ID.txt"
 aggFCRew_TAB_PATH="/data/pt_02161/Results/Project2_resting_state/connectivity/Analysis/aggFC/Rew/agg_FC_Rew.txt"
 
 # Get functions -----------------------------------------------------------
@@ -110,6 +110,10 @@ tableDescr[c(2:6,8:12),"predictor"] <- ""
 tableDescr[c(2,4,6,8,10,12),"tp"] <- ""
 rownames(tableDescr) <- NULL
 saveRDS(object = tableDescr, file = "../report/tab/tableDescr.rds")
+
+
+# Surgery+Comorbidities ---------------------------------------------------
+
 
 
 # figBMIdescr -------------------------------------------------------------

### 2021-03-09 manuscript/draft_rsfmri.Rmd
minor changes

@@
 ---
-title: "Effects of bariatric surgery loss on resting-state functional connectivity of the reward and default mode network: a pre-registered analysis"
+title: "Effects of bariatric surgery on resting-state functional connectivity of the reward and default mode network: a pre-registered analysis"
 author:
@@
 
-Obesity imposes serious health risks and is consistently associated with differences in brain activity, including resting-state functional connectivity.
-Effective treatment can be achieved by bariatric surgery, yet its effects on functional connectivity of brain networks involved in reward processingand intereoception are still under debate.
-In this pre-registered analysis, we aimed to determine the effects of bari-atric surgery and weight loss on functional connectivity in the RN and default mode network (DMN.
-Therefore, 33 bariatric surgery patients (males/females, BMI, age etc) and 15 obese waiting-list control patients (males/females, BMI, age etc) underwent MRI at baseline (before surgery/waiting period), and after 6 and 12 months.
-We pre-registered (https://osf.io/f8tpn and https://osf.io/59bh7/) the whole-brain time-by-group interaction on network connectivity, assessed between baseline and followup and for two denoising pipelines.
-We also pre-registered a time-by-group interaction analysis of aggregated FC within networks.
+Obesity imposes serious health risks and relates to differences in resting-state functional connectivity (FC) of brain networks involved in eating behavior.
+Bariatric surgery is an effective treatment, but its effects on FC are still under debate.
+In this pre-registered analysis, we aimed to determine the effects of bariatric surgery on FC in the reward (RN) and default mode network (DMN).
+33 bariatric surgery patients and 15 obese waiting-list control patients (37 females; aged 44.15 ± 11.86 SD years (range 21-68)) underwent magnetic resonance imaging at baseline, after 6 and 12 months.
+We pre-registered (https://osf.io/f8tpn/https://osf.io/59bh7/) a whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network FC.
 In exploratory analyses, we investigated all three time points and assessed the effects of weight loss and head motion.
-According to whole-brain marginal models, bariatric surgery compared to waiting did not significantly affect functional connectivity of the the reward and DMN networks (FWE p > 0.05).
-There also was no interaction effect on within network FC in pre-registered analysis of aggregated FC.
-Exploratory analyses also did not show an interaction effect for three time points, but when adding differences in head motion within and between individuals using mean framewise displacement, surgery-related BMI decrease and higher average head motion resulted in stronger functional connectivity in medial posterior frontal regions within the RN.
-This pre-registered analysis of longitudinal neuroimaging data did not support a strong effect of bariatric surgery, compared to waiting, on reward and DMN network functional connectivity after 6 months in obese patients.
-Exploratory analyses indicated that differences in head motion might have confounded effects of weight loss on reward-related functional connectivity, yet definite conclusions remain difficult.
-Future studies with a more rigorous control of within-scanner head motion during data acquisition are warranted to further understand the effects of obesity surgery on functional connectivity measures.
+Bariatric surgery compared to waiting did not significantly affect FC of the RN and DMN (FWE p > .05).
+There also was no interaction effect on within-network FC.
+In exploratory analyses, surgery-related BMI decrease (FWE p= .041) and higher average head motion (FWE p= .021) resulted in significantly stronger FC in medial posterior frontal regions within the RN.
+This pre-registered analysis of longitudinal neuroimaging data did not support a strong effect of bariatric surgery, compared to waiting, on RN and DMN FC after 6 months.
+Exploratory analyses indicated that differences in head motion might have confounded the effects, thus a more rigorous control of within-scanner head motion during data acquisition is needed to better understand the effects of bariatric surgery on brain organization.
 
@@
 Obesity is a worldwide health issue, entailing huge personal and societal costs.
-Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life (hier eine Ref).
+Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life [@Albanese_2017].
 Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI ($> 35$ kg/m$^2$).
@@
 RYBG is the most invasive, yet most effective surgical pro-cedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].
-Bariatric surgery induces weight loss  in part by malabsorption of nutrients, especially fat, and reduced digestion efficiency.
-More importantly, weight loss following the surgical procedures, and especially RYGB, is related to profound alterations in appetite, food perception and the central regulation of food intake [@Mulla_2017; @Brutman_2019].
-Precise mechanisms how bariatric surgery leads to altered appetitive and reward signaling are yet to be elucidated.
+Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla_2017; @Brutman_2019].
+Precise mechanisms how bariatric surgery leads to altered appetitive signaling are yet to be elucidated.
 Resting-state functional magnetic resonance imaging (rsfMRI) is a technique to investigate the dynamic organization of the brain.
@@
 The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
-The RN (RN), comprises the ventromedial prefrontal cortex (vmPFC), the striatum, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
+The RN (RN), comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (Nacc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
 These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
 Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
-resting-state rsfMRI studies also shown increased local FC of re-ward network regions, i.e.
-NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
+RsfMRI studies also showed increased local FC of reward network regions, i.e.
+Nacc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
 In bariatric surgery patients, connectivity  within the RN (e.g.
@@
 changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] .
-The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex (approx.
-BA 39) [@Raichle_2015].
+The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle_2015].
 Higher BMI and obesity have been associated with a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
 salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
-Studies in bariatric surgery patients have pointed to a normalization of the connectivity between DMN and cognitive control and salience brain regions after the procedure, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
+After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
 In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
 Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].
-Most rsfMRI studies in obesity have reported an association of higher BMI and higher head motion during MRI [@Beyer_2017;@Hodgson_2016].
-Head motion is a major confounder in rsfMRI as it increases short-range and disrupts long-range connectivity FC.
-In this sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
-Here, we thus aimed to rigorously control for motion-related variance.
-In this pre-registered study, our confirmatory hypotheses were:
+Further, while higher BMI has been consistently associated with more head motion during rsfMRI, previous studies in bariatric surgery patients have not taken this important confounder of FC into account [@Beyer_2017;@Hodgson_2016].
+In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
+Thus, we aimed to rigorously control for motion-related variance in our analyses.
+We had the following confirmatory hypotheses:
 1. Whole-brain FC of the RN and DMN changes differently from baseline to followup in the bariatric surgery compared to a waiting-list control group
@@
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (MMSE score < $34$).
- In total, $51$ participants out of the originally enrolled $69$ subjects received MRI, and we analyzed a total of $101$ data points.
-Five data points of three subjects had to be excluded due to bad anatomical image quality (see below).
+Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).
+ In total, $51$ participants out of the originally enrolled $69$ subjects received MRI.
+Five data points of three subjects had to be excluded due to bad anatomical image quality (see below), which led to a final data set with $101$ rsfMRI sessions.
 <!-- DESCRIPTIVE STATISTICS -->
-The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`).
+The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`)).
+$X$% of participants had clinically diagnosed hypertension, $x$% had type-2 or type-1 diabetes, and $X$% smoked.
 Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016].
-Analyses were performed on all participants that provided at least one data point of rsfMRI data.
+Analyses were performed on all participants who provided at least one data point of rsfMRI data.
 `r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point (for more details see Table S1).
@@
 They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
-MRI scanning was done after further psychological tests (for details see @Prehn_2020)).
+MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020)).
 
@@
 
-MRI images were retrieved with a 12-channel head coil of a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
+MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
 <!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
@@
 Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long].
-Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rs preprocessing.
+Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.
 The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants].
@@
 ### Minimal functional preprocessing
-The function preprocessing module included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
+The functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
 In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
@@
 Yet, it introduces spatial dependency and spurious correlations.
-Thus, we pre-registered the main analyses to be performed with these two denoising pipelines.
+Thus, we pre-registered the main analyses to be performed with both denoising pipelines.
 
@@
 ### Functional connectivity (FC)
-To derive RN and DMN FC maps, we used nucleus accumbens (NAcc) and precu-neus as seed regions of interest (ROI), respectively.
+To derive RN and DMN FC maps, we used Nacc and precuneus as seed regions of interest (ROI), respectively.
 We did not select vmPFC for the RN due to low SNR.
 Because we knew about the large amount of motion in this dataset beforehand, we decided against ICA to construct the networks, as ICA is recommended to work on minimally preprocessed data.
-Based on FreeSurfer’s aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
-Then, we used NiftiLa-belsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each processing mode (minimally preprocessed, aro-ma+cc, aroma+gsr).
+Based on FreeSurfer’s aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for Nacc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
+Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
+We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+GSR).
 Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
@@
 
-#### RsfMRI quality control
+#### RsfMRI quality control (QC)
 RsfMRI quality control was performed according to the protocol by @Ciric_2018. 
-Head motion is a major confounder in rsfMRI. We calculated mean framewise displacement (mFD) according to @Power_2012, and log-transformed it for further analysis (logmFD).
+We calculated mean framewise displacement (mFD) according to @Power_2012, and log-transformed it for further analysis (logmFD).
 As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
@@
 
-##### Group level resting-state rsfMRI QC
+##### Group level rsfMRI QC
 
@@
 For this study, we further explored mFD and its relation to DVARS, the root mean square of the temporal change of the voxel-wise signal at each time point.
+DVARS measures volume-to-volume signal variation.
 ```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
@@
 ### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and RN over all participants and NAccs, adjusted for age and sex.
+To extract aggregated FC values, we first calculated the mean DMN and RN over all participants and Naccs, adjusted for age and sex.
 We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
@@
 #### Confirmatory Analysis
-We tested the pre-registered hypothesis of a  time by group interaction:  $FC=group + time + group * time$. 
-As pre-registered, we performed the analysis for baseline and followup time points only, in both denoising pipelines and adjusting for no confounders (CA 1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
-In this model, time was represented as factor because we took into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
+We tested the pre-registered hypothesis of a time-by-group interaction:  $FC=group + time + group * time$. 
+As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (CA 1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
+In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
 The resulting factorial design contained one regressor for each time point per group (see supplements for depiction of design matrix).
@@
 ##### Group-by-time interaction in three time points
-Then, we extended our whole-brain analysis of the group-by-time point interaction to the data from all three timepoints.
+Then, we extended our whole-brain analysis of the group-by-time point interaction to the data from all three time points.
 This analysis had not been pre-registered.
-Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points.We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and logmFD).
+Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points.
+We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and logmFD).
 
@@
 As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
-This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for log mFD (Model EA 2.2).
+This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC.
+We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for logmFD (Model EA 2.2).
 As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 2.3c).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: $FC = between-subject logmFD + within-subject logmFD$ age and sex as nuisance covariates.
+For a better understanding of the unique contribution of average and longitudinal change in logmFD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: $FC = between-subject logmFD + within-subject logmFD$ with age and sex as nuisance covariates.
 
@@
 ### Model input and brain mask
-For each brain network, we examined the models for the two denoising pipelines (AROMA+CC/ AROMA+CC+GSR)as pre-registered.
+For each brain network, we examined the models for the two denoising pipelines (AROMA+CC/AROMA+CC+GSR) as pre-registered.
 For all analyses, we used an explicit brain mask, which was a whole brain derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
@@
 We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007].
-We deviated from the pre-registration by only investigating data from the GSR denoising pipeline.
+We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline.
 First, we investigated the time-by-group interaction for baseline and follow-up time points only.
-We either performed no correction (CA1), adjusted for baseline age, sex, average of logmFD of both time points (CA2) or additionally adjusted for baseline BM (CA3)I .
-We performed model comparison between R1 and R0 models, where R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj)), and R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj)).
+We either performed no correction (CA1), adjusted for baseline age, sex, average of logmFD of both time points (CA2) or additionally adjusted for baseline BMI (CA3).
+We performed model comparison between R1 and R0 models, where $R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))$, and $R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))$.
 As planned in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.
@@
 #### Exploratory analysis
-We performed an exploratory analysis with average BMI and variability in BMI as predictors of the aggregated FC from GSR data of both networks, in line with our exploratory whole brain analysis.
-We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (Model EA2.1), logmFD (Model EA 2.2) and average and change logmFD (Model EA 2.3).
+We performed an exploratory analysis with average BMI and change in BMI as predictors of the aggregated FC from AROMA+CC+GSR denoised data of both networks, in line with our exploratory whole brain analysis.
+We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (Model EA2.1), logmFD (Model EA 2.2) and average and change in logmFD (Model EA 2.3).
 
@@
 ### Model CA1 (without adjustments)
-Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the NAcc nor the PCC seeds.
+Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the Nacc nor the PCC seeds.
 There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE-correction.
@@
 There was no significant interaction when adding baseline BMI.
+Yet, we found a significant main effect of time in this model.
+For AROMA+CC+GSR preprocessed data, we found higher FC of the Nacc with the lateral occipital cortex at baseline compared to followup (FWE $p=0.030$), and higher FC of the PCC within the DMN to the medial ACC (FWE $p=0.046$)(see supplements).
 
@@
 ##### Average and change in BMI, adjusted for logmFD (EA 2.1 and EA 2.2)
-NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for logmFD.
+Nacc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for logmFD.
 
 ##### Average and change in BMI, adjusted for split logmFD (EA 2.3)
-<!-- cc -->For the RN, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region. 
+<!-- cc -->For the RN, we found a significant negative association of change in BMI with FC between Nacc and a cluster in the posterior-medial frontal region. 
 The peak voxel derived from investigating the effect of average logmFD was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), even if data was GSR ($p = 0.038$).
-```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Cluster showing a significant negative effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
-knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
-                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
+```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Cluster showing a significant negative effect of change BMI on Nacc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+knitr::include_graphics(c("../report/fig/BMIFDagesex_Nacccc_cngBMI_deact.pdf",
+                          "../report/fig/BMIFDagesex_Naccgsr_cngBMI_deact.pdf"))
 ```
-<!-- gsr -->Moreover, FC between the NAcc and one cluster in a more posterior region, i.e.
-motor cortex, was significantly positively related to average logmFD (see  Figure \@ref(fig:BMIFDagesexNACCavgFDact)).
-This result was similar for GSR corrected data.
-```{r "BMIFDagesexNACCavgFDact", fig.cap="Cluster showing a significant positive Effect of average logmFD on FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
-knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
-                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
+<!-- gsr -->Moreover, FC between the Nacc and one cluster in a more posterior region, i.e.
+motor cortex, was significantly positively related to average logmFD (see  Figure \@ref(fig:BMIFDagesexNaccavgFDact)).
+This result was similar for AROMA+CC+GSR preprocessed data.
+```{r "BMIFDagesexNaccavgFDact", fig.cap="Cluster showing a significant positive Effect of average logmFD on FC between Nacc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+knitr::include_graphics(c("../report/fig/BMIFDagesex_Nacccc_avgFD_act.pdf",
+                          "../report/fig/BMIFDagesex_Naccgsr_avgFD_act.pdf"))
 ```
@@
 <!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
-<!-- gsr -->As before, the effect did not survive when using GSR-data (see Table \@ref(tab:tableBMIFDmodel).
+<!-- gsr -->As before, the effect did not survive when using AROMA+CC+GSR preprocessed (see Table \@ref(tab:tableBMIFDmodel).
 ```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
@@
 Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and
-premotor cortex were the meta-analytic activation maps most similar to this contract.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract.
 The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
@@
 
-To explore the unique contributions of the average value of and the longitudinal change in logmFD, we constructed an additional model, where BMI measures are excluded from the design matrix.
-Covariance of the logmFD and outcome measures is investigated by splitting logmFD into average logmFD and longitudinal logmFD change and by inspecting associations with each predictor.
-Age and sex were nuisance covariates.
+To explore the unique contributions of the average value of and the longitudinal change in logmFD, we constructed an additional model, where BMI measures were excluded from the design matrix.
+Covariance of the logmFD and outcome measures was investigated by splitting logmFD into average logmFD and longitudinal logmFD change and by inspecting associations with each predictor.
+Age and sex were used as nuisance covariates.
 
-<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), there was a significant positive association between average logmFD and co-activation of NAcc with one cluster located in proximity to the central sulcus and motor areas.
+<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNaccccavgFDact), there was a significant positive association between average logmFD and co-activation of Nacc with one cluster located in proximity to the central sulcus and motor areas.
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
-<!-- ### DMN -->We did not find any clusters with a significant association of logmFD and PCC connectivity.
-```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
-knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
-knitr::include_graphics("../report/fig/FDagesex_NACCgsr_avgFD_act.pdf")
+<!-- ### DMN -->We did not find any clusters with a significant association of logmFD and PCC FC.
+```{r "FDagesexNaccccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of Nacc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
+knitr::include_graphics("../report/fig/FDagesex_Nacccc_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/FDagesex_Naccgsr_avgFD_act.pdf")
 ```
@@
 ```
-Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN connectivity, regardless of whether we adjusted for logmFD (model EA 2.1: $p$ = `r round(p_meanBMI,3)` and EA 2.2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
+Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN FC, regardless of whether we adjusted for logmFD (model EA 2.1: $p$ = `r round(p_meanBMI,3)` and EA 2.2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
 The association also remained significant when we split logmFD into average and change in logmFD (EA 2.3) ($p$ = `r round(p_mean_BMI_splitmFD,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (see supplements material for overview of all models).
@@
 Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising pipelines and controlling for mFD on the group level.
-We found no interaction effect of group and time point on whole-brain FC of the PCC and NAcc.
-Neither did we observer main effects of group and time point on the FC of these DMN and RN hubs, respectively.
+We found no interaction effect of group and time point on whole-brain FC of the PCC and Nacc.
+We observed a main effect of time when adjusting for age, sex, average logmFD and BMI, and which pointed to higher FC at baseline compared to followup for both DMN and RN hubs.
 Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
-In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising pipeline.
-When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI across the three time points and the connectivity of NAcc with a posterior-medial frontal cluster emerged.
+In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN FC for the more lenient denoising pipeline.
+When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI across the three time points and the connectivity of Nacc with a posterior-medial frontal cluster emerged.
 This result was significant in both denoising pipelines.
 Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns.
-Finally, higher average head motion was associated with increased NAcc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
+Finally, higher average head motion was associated with increased Nacc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
 
-In this study, we could not confirm our pre-registered hypotheses and found no interaction effect between group and time point on DMN and RN connectivity.   
-Based on previous studies in bariatric surgery patients, we expected within-DMN connectivity to increase, and DMN connectivity to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
+In this study, we could not confirm our pre-registered hypotheses and found no interaction effect between group and time point on DMN and RN FC.   
+Based on previous studies in bariatric surgery patients, we expected within-DMN FC to increase, and DMN FC to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
 Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat  [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019].
-However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the NAcc and a cluster in a posterior-medial frontal brain region.
+However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the Nacc and a cluster in a posterior-medial frontal brain region.
 Based on the Harvard-Oxford atlas and NeuroSynth decoding this region might be part of the salience network and involved in action preparation.
-The enhanced connectivity between the NAcc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
+The enhanced FC between the Nacc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
 One the other hand, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
-Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality (DC)) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi_2018].
+Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi_2018].
 Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
@@
 Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.
-To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and thereby introducing two instead of one regressor into the model.
-These differential results may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
+To our surprise, the effect of weight loss on the connectivity of the Nacc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and thereby introducing two instead of one regressor into the model.
+These results may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
 Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
-This argument is supported by the survival of the cluster when using GSR-data, and the distinct results of the decoding analysis.
+This argument is supported by the survival of the cluster when using AROMA+CC+GSR preprocessed data, and the distinct results of the decoding analysis.
 Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
-Thus, confounding of the negative BMI change effect and the positive average head motion effect on posterior-frontal connectivity  seems unlikely.
-Head motion also played a role in the association of higher BMI and reduced DMN connectivity.
-While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN F was negatively associated with BMI in both denoising schemes.
-Thereby, this results echos a previous finding from our group, and may be interpreted as accelerated age-related decline of the DMN  in relation to the cardiometabolic risk factor BMI.
-Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding also remain in this study [@Savalia_2017].
+Thus, confounding of the negative BMI change effect and the positive average head motion effect on posterior-frontal FC  seems unlikely.
+Head motion also played a role in the association of higher BMI and reduced DMN FC.
+While this result was no longer significant on a whole-brain level when using stringent denoising, aggregated within-DMN FC was negatively associated with BMI in both denoising schemes.
+Thereby, this results echoes a previous finding from our group, and may be interpreted as accelerated age-related decline of the DMN in relation to the cardiometabolic risk related to BMI.
+Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia_2017].
 
 The major strength of our study is prospective intervention controlled design.
-We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities (?), treatment history and  treatment recommendation and were scanned after the same time intervals [@Thiese_2014].
+We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history and  treatment recommendation and were scanned after the same time intervals [@Thiese_2014].
 Another strength of our study was the pre-registered analysis plan, which was corrected after we realized some major flaws in the first version.
@@
 Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI scanner.
-While this sample size is comparable to previous resting-state rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
+While this sample size is comparable to previous rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
 
@@
 While this approach yielded reasonable DMN and RN maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
-Furthermore, we did not explicitly investigate the connectivity of the DMN and RN with other large-scale brain networks attention, e.g.
+Furthermore, we did not explicitly investigate the FC of the DMN and RN with other large-scale brain networks attention, e.g.
 cognitive control or sensorimotor networks, which has been previously proposed to be altered by obesity and bariatric surgery [@Lips_2014; @Olivo_2016; @Cerit_2019].
 
-We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a standardized (?) breakfast following an overnight fast.
+We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a breakfast following an overnight fast.
 Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
@@
 
-Taken together, this study showed that effects of bariatric surgery on resting-state are subtle, and often confounded by head motion.
+Taken together, this study showed that effects of bariatric surgery on resting-state brain organization are, if at all, subtle, and often confounded by head motion.
 Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to observed changes in food wanting and choice in patients after bariatric surgery.

### 2021-03-09 code/create_sample_df.R

@@ -32,21 +32,7 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
       full$exclude_prep <- full$Exclude
       full$exclude_prep[is.na(full$exclude_prep)] = FALSE
       mri_sample <-
-        merge(mri_files, full[!is.na(full$condition), c(
-          "subj.ID",
-          "condition",
-          "tp",
-          "Age_BL",
-          "Sex",
-          "meanFD",
-          "BMI",
-          "BMI_BL",
-          "BMI_BLi",
-          "comorbidities",
-          "interv",
-          "intervention_de",
-          "exclude_prep"
-        )], by = c("subj.ID", "tp"))
+        merge(mri_files, full[!is.na(full$condition),], by = c("subj.ID", "tp"))
       cat("n =", nrow(mri_sample), "fMRI data points, ")
       
       mri_sample[mri_sample$exclude_prep == TRUE, ] # ADI063_bl ADI063_fu ADI116_bl ADI116_fu ADI116_fu2
@@ -238,13 +224,17 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
   addit_fu2$tp <- "fu2"
   addit_info <- rbind(addit_bl,addit_fu,addit_fu2)
   rm(addit_bl,addit_fu,addit_fu2)
-  addit_info$intervention_de <- addit_info$Operationsverfahren
+  
+  addit_info$diabetes_type_I <- addit_info$`Typ-1-Diabetes mellitus`
+  addit_info$diabetes_type_II <- addit_info$`Typ-1-Diabetes mellitus`
+  addit_info$arterial_hypotension <- addit_info$`Arterielle Hypertonie`
+  addit_info$intervention <- addit_info$Operationsverfahren
   addit_info$tp <- as.factor(addit_info$tp)
   addit_info$subj.ID <- as.factor(addit_info$subj.ID)
-
-  
-  full_surg_comorbid <- merge(full, addit_info, by=c("subj.ID","tp"), all = TRUE)
+  surg_comorbid <- addit_info[,c("intervention", "diabetes_type_I", "diabetes_type_II", "arterial_hypotension", "subj.ID", "tp")]
   rm(addit_info)
+  
+  full_surg_comorbid <- merge(full, surg_comorbid, by=c("subj.ID","tp"), all = TRUE)
   # check if KG have 
   t(table(full_surg_comorbid$condition, full_surg_comorbid$intervention_de))
   full_surg_comorbid$interv <- NA

### 2021-03-09 manuscript/draft_rsfmri.Rmd
report comorbidities

@@
 
-subj_df$comorbidities
 ```
@@
 The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`)).
-$X$% of participants had clinically diagnosed hypertension, $x$% had type-2 or type-1 diabetes, and $X$% smoked.
+`r round(sum(subj_df$arterial_hypotension == 1, na.rm = T) / nrow(subj_df), 3)*100`% of participants had clinically diagnosed hypertension, `r round((sum(subj_df$diabetes_type_II == 1, na.rm = T) + sum(subj_df$diabetes_type_I == 1, na.rm = T))  / nrow(subj_df), 3)*100`% had type-2 or type-1 diabetes, and `r round(sum(subj_df$smoking_status == 1, na.rm = T) / nrow(subj_df), 3)*100`% reported to smoke.
 Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
@@
 
-`r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the BARS group information was not available.
+`r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the intervention group information was not available.
 
@@
 We tested the pre-registered hypothesis of a time-by-group interaction:  $FC=group + time + group * time$. 
-As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (CA 1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
+As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (CA1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
 In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
 The resulting factorial design contained one regressor for each time point per group (see supplements for depiction of design matrix).
+Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package mice based on subject identification, condition, baseline age, and available BMI values.
+This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.
 
@@
 We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for logmFD (Model EA 2.2).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 2.3c).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 2.3).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-For a better understanding of the unique contribution of average and longitudinal change in logmFD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: $FC = between-subject logmFD + within-subject logmFD$ with age and sex as nuisance covariates.
+For a better understanding of the unique contribution of average and longitudinal change in logmFD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates.
 
@@
 ```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Cluster showing a significant negative effect of change BMI on Nacc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
-knitr::include_graphics(c("../report/fig/BMIFDagesex_Nacccc_cngBMI_deact.pdf",
-                          "../report/fig/BMIFDagesex_Naccgsr_cngBMI_deact.pdf"))
+knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
+                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
 ```
@@
 ```{r "BMIFDagesexNaccavgFDact", fig.cap="Cluster showing a significant positive Effect of average logmFD on FC between Nacc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
-knitr::include_graphics(c("../report/fig/BMIFDagesex_Nacccc_avgFD_act.pdf",
-                          "../report/fig/BMIFDagesex_Naccgsr_avgFD_act.pdf"))
+knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
+                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
 ```
@@
 
-<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNaccccavgFDact), there was a significant positive association between average logmFD and co-activation of Nacc with one cluster located in proximity to the central sulcus and motor areas.
+<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), there was a significant positive association between average logmFD and co-activation of Nacc with one cluster located in proximity to the central sulcus and motor areas.
 Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
 <!-- ### DMN -->We did not find any clusters with a significant association of logmFD and PCC FC.
-```{r "FDagesexNaccccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of Nacc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
-knitr::include_graphics("../report/fig/FDagesex_Nacccc_avgFD_act.pdf")
-knitr::include_graphics("../report/fig/FDagesex_Naccgsr_avgFD_act.pdf")
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of Nacc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
+knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
+knitr::include_graphics("../report/fig/FDagesex_NACCgsr_avgFD_act.pdf")
 ```

### 2021-03-09 code/create_sample_df.R

@@ -204,7 +204,7 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
                            stringsAsFactors = FALSE
                            )
   colnames(addit_info) <- addit_info[1,]
-  addit_info <- addit_info[2:nrow(addit_info),c(1:38)]
+  addit_info <- addit_info[2:nrow(addit_info),c(1:38,48)]
   addit_info$subj.ID <- gsub("-", "", addit_info$Code)
   comorbidities <- addit_info %>%
     dplyr::select(c("Adipositas durch Kalorienzufuhr, [BMI] 40 und mehr":"Migräne")) %>%
@@ -228,21 +228,24 @@ create_sample_df <- function(group = "both", tp = "all", exclFD = FALSE, out = "
   addit_info$diabetes_type_I <- addit_info$`Typ-1-Diabetes mellitus`
   addit_info$diabetes_type_II <- addit_info$`Typ-1-Diabetes mellitus`
   addit_info$arterial_hypotension <- addit_info$`Arterielle Hypertonie`
+  addit_info$smoking_status <- NA
+  addit_info$smoking_status[addit_info$`BL - Rauchen aktuell (ja/nein)` == "ja"] = 1
+  addit_info$smoking_status[addit_info$`BL - Rauchen aktuell (ja/nein)` == "nein"] = 0
   addit_info$intervention <- addit_info$Operationsverfahren
   addit_info$tp <- as.factor(addit_info$tp)
   addit_info$subj.ID <- as.factor(addit_info$subj.ID)
-  surg_comorbid <- addit_info[,c("intervention", "diabetes_type_I", "diabetes_type_II", "arterial_hypotension", "subj.ID", "tp")]
+  surg_comorbid <- addit_info[,c("intervention", "diabetes_type_I", "diabetes_type_II", "arterial_hypotension", "smoking_status", "subj.ID", "tp")]
   rm(addit_info)
   
   full_surg_comorbid <- merge(full, surg_comorbid, by=c("subj.ID","tp"), all = TRUE)
   # check if KG have 
-  t(table(full_surg_comorbid$condition, full_surg_comorbid$intervention_de))
+  t(table(full_surg_comorbid$condition, full_surg_comorbid$intervention))
   full_surg_comorbid$interv <- NA
[…]

### 2021-03-19 manuscript/supplements_rsfmri.Rmd
recent changes after 2nd round co-author review

@@
 - Evelyn Medawar [1]
-- Kristin Prehn [3][4]
+- Kristin Prehn [4][5]
 - Jürgen Ordemann [8][9]
-- Agnes Floel [5][6][7]
-- A. Veronica Witte [1][2][10]
+- Agnes Floel [4][5]
+- A. Veronica Witte [3][1][2]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
-  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Stroke Research, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Zentrum für Adipositas und Metabolische Chirurgie, Vivantes Klinikum Spandau, Berlin \newline 10 Day Clinic for Cognitive Neurology, University Clinic Leipzig
+  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Clinic of Cognitive Neurology, University Clinic Leipzig \newline 4 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 5 Department of Psychology, Medical School Hamburg, Hamburg \newline 6 Department of Neurology, University of Greifswald, Greifswald \newline 7 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin
 output:
@@
 # Methods
+Table \@ref(tab:tableSample) shows the distribution of data points in the present rsfMRI study, for the bariatric surgery group (BARS) and the waiting-list control group (NBARS).
+
 ## Sample
+
 ```{r "tableSample", echo=FALSE}
@@
 
+## Denoising Pipelines
+We pre-registered two distinct denoising pipelines which were previously shown to efficiently remove motion artifacts in fMRI.
+In @Parkes_2018,  ICA-AROMA provided an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relationship.
+@Ciric_2017 showed that GSR is very efficient in removing the correlation of head motion and FC, and outperforms ICA-AROMA alone.
+Yet, it introduces spatial dependency and spurious correlations.
+
+### ICA-AROMA denoising pipeline   
+As described in @pruim2015ica, the minimally preprocessed data underwent independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
+Finally, the timeseries of the  nuisance components were regressed from the raw rsfMRI timeseries non-aggressively.
+As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result.
+We used CompCor to estimate 5 variance components from a combined WM and CSF mask, which was further eroded using `fslmaths -nan -thr 0.99 -ero -bin` [@Behzadi_2007;@Muschelli_2014].
+After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
+
+### Global Signal Regression (GSR) 
+The global signal was derived from the average of all voxels in the brain mask.
+Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA rsfMRI data and high-pass filtered the resulting file as above.
+
 ## Quality Control
+### Individual level resting-state rsfMRI QC
+```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
+carpets=c("../report/carpetplotadi006fu2.jpeg","../report/carpetplotadi16fu.jpeg")
+knitr::include_graphics(carpets)
+```
+On the individual level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
+Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum).
+All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`, and checked again visually.
+No participant was excluded based on this visual inspection.
 
-```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different preprocessing schemes)", out.width="100%", fig.align = "center", eval=FALSE}
+### Group level rsfMRI QC
+
+As described in @Beyer_2020, there was a group-by-time interaction on mFD (see Figure \@ref(fig:figDvarsmFD), upper right panel).
+The intervention group reduced in head motion in the followup assessments compared to the control group.
+For this study, we further explored mFD and its relation to DVARS, the root mean square of the temporal change of the voxel-wise signal at each time point.
+DVARS measures volume-to-volume signal variation.
+DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points (see Figure S2 in suppplements).
+To further assess the confounding of head motion and FC, we evaluated mFD-FC relationships by computing a functional connectome based on the protocol by [@Ciric_2018].
+We used the MNI coordinates of 246 cortical and subcortical spherical seed regions from @Power_2012 with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
+Then, we calculated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation.
+```{r mFDFC, echo=FALSE}
+res <- read.csv("../report/FDFCcorrelations.csv")
+rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
+colnames(res) <- c('mean mFD-FC', 'median mFD-FC', 'sig. conn.', 'sig. conn. BH', 'mean distance-mFDFC', 'p distance')
+```
+```{r tableQcfc, out.width="100%", echo=FALSE}
+tab_Qcfc <-
+  knitr::kable(
+    round(res, 3),
+    col.names = colnames(res),
+    row.names = TRUE,
+    format = "latex",
+    booktabs = T,
+    linesep = "",
+    caption = "Summary of quality metrics mean mFD-FC (mean Spearman's correlation of mFD and FC),median mFD-FC (median Spearman's correlation of mFD and FC), sig.
+conn.
+(number of significant connections), sig.
+conn.
+BH (number of FDR-corrected, significant connections), mean distance-mFDFC (mean Spearman's correlation of distance and mFD-FC correlation), p distance (p-value of distance dependency),  for minimal preprocessing,AROMA, AROMA+CC and AROMA+CC+GSR denoising pipelines over all data points"
+  ) %>%
+  column_spec(c(2:6), width = "1.8cm")
+tab_Qcfc
+```
+In the minimally preprocessed data, mFD and FC between nodes were significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`)(see Table \@ref(tab:tableQcfc)).
+This relationship was negatively associated with distance between nodes (mean Spearman's $r$ = `r round(res[1,5],2)`).
+Denoising pipelines (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and FC, and reduced the distance dependency.
+Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
+Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r$ = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r$ = `r round(res[4,5],2)`)
+In addition, for models showing a significant time-by-group interaction, we planned to perform a sensitivity analysis excluding the 10% data sets with Fhighest mFD.
+
+```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS, mFD and their correlation', fig.show='hold',fig.align='center', echo=FALSE}
+figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
+#(figList$fig1 | figList$fig2) / figList$fig3
+
+lay <- rbind(c(1,1,2,2),
+             c(NA,3,3,NA))
+gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
+```
+
+```{r fighistograms,fig.cap="Histograms of FC-QC correlations for different denoising pipelines)", out.width="100%", fig.align = "default", echo=FALSE}
+knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/dp_aranged.jpeg")
+```
+Figure \@ref(fig:fighistograms) shows the distribution of FC-QC correlations for minimally preprocessed, AROMA, AROMA+CC and AROMA+CC+GSR preprocessed data.
+With more denoising, there is a decreasing number of high correlations between FC and the measure of head motion, supporting the positive effect of denoising.
+
+```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="100%", fig.align = "center", eval=TRUE}
 tsnr <- "../report/tsnr.pdf"
@@
 
-## Networks
+## DMN and RN Networks
 Figure \@ref(fig:figNetworks) shows the DMN and reward network average maps which we used for extracting aggregated FC values.
@@
 
-```{r figtSNR,fig.cap="Lower tSNR in the region of interest NAcc versus precuneus used for seed-based connectivity (shown are all 101 timepoints)", out.width="100%", fig.align = "center", eval=TRUE}
-tsnr <- "../report/tsnr.pdf"
-knitr::include_graphics(tsnr)
-```
-
 ## Design
@@
 #gridExtra::grid.arrange(plotA, plotB, nrow=1)
+```
+
+# Results
+
+## Confirmatory Analysis
+
+```{r "GT2tpagesexmFDBMIblNaccgsrTimeDeact", fig.cap="Main effect of time: Negative Association of FC between NAcc and this cluster in lateral parietal cortex, adjusted for age, sex, avgFD, and baseline BMI in AROMA+CC+GSR preprocessed data", fig.hold=TRUE, out.width = "50%", fig.align="center"}
+knitr::include_graphics(c("../report/fig/GroupTime2tp_NACCgsr_time_deact.pdf"))
+```
+
+```{r "GT2tpagesexmFDBMIblPCCgsrTimeDeact", fig.cap="Main effect of time: Negative Association of FC between PCC/precuneus and this cluster in ACC, adjusted for age, sex, avgFD, and baseline BMI in AROMA+CC+GSR preprocessed data", fig.hold=TRUE, out.width = "50%", fig.align="center"}
+knitr::include_graphics(c("../report/fig/GroupTime2tp_PCCgsr_time_deact.pdf"))
 ```
@@
 anvRew2tp=anova(R1,R0)
-anvRew2tp
 ```
@@
 R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
-R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD +(1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
+R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
 anvRew2tpadj=anova(R1,R0)

### 2021-03-19 code/figures.R

@@ -199,7 +199,7 @@ mk_figRewdescr <- function(final_FC) {
       values = c("#046C9A70", "#D69C4E70"),
       labels = c("intervention", "control")
     ) +
-    xlab("Time after intervention [months]") + ylab ("mean reward connectivity") +
+    xlab("Time after intervention [months]") + ylab ("mean RN FC [Z-score]") +
     theme_bw() +  theme(
       axis.text = element_text(size = 10),
       axis.title = element_text(size = 12),
@@ -250,7 +250,7 @@ mk_figDMNdescr <- function(final_FC) {
       values = c("#046C9A70", "#D69C4E70"),
       labels = c("intervention", "control")
     ) +
-    xlab("Time after intervention [months]") + ylab ("mean DMN connectivity") +
+    xlab("Time after intervention [months]") + ylab ("mean DMN FC [Z-score]") +
     theme_bw() +  theme(
       axis.text = element_text(size = 10),
       axis.title = element_text(size = 12),
@@ -560,14 +560,16 @@ figDvarsmFD <- function(final){
       values = c("#046C9A70", "#D69C4E70"),
       labels = c("intervention", "control")
     ) +
-    xlab("Time [months]") + ylab ("Mean DVARS") +
+    xlab("Time [months]") + ylab (expression(paste("Mean DVARS in\n 10% BOLD change"))) +
     theme_bw() +  theme(
       axis.text = element_text(size = 10),
       axis.title = element_text(size = 12),
       strip.text = element_text(size = 12),
       legend.text = element_text(size = 10),
-      legend.position = "none"
+      legend.position = "none",
+      plot.margin=unit(c(5.5, 5.5, 5.5, 15), "points")
     )
+  
   # mFD over time each group
   fig_mFDdescr <-
     ggplot(final, aes(
@@ -646,7 +648,7 @@ figDvarsmFD <- function(final){
       values = c("#046C9A70", "#D69C4E70"),
       labels = c("intervention", "control")
     ) +
-    xlab("Time [months]") + ylab ("DVARS-FD correlation") +
+    xlab("Time [months]") + ylab ("DVARS-FD correlation r") +
     theme_bw() +  theme(
       axis.text = element_text(size = 10),
       axis.title = element_text(size = 12),

### 2021-03-19 code/process.R

@@ -124,8 +124,8 @@ saveRDS(object = fig_BMIdescr, file = "../report/fig/fig_BMIdescr.rds")
 
 # figFDdescr -------------------------------------------------------------
 
-fig_BMIdescr <- mk_figFDdescr_for_OHBM(final)
-saveRDS(object = fig_BMIdescr, file = "../report/fig/fig_BMIdescr.rds")
+fig_FDdescr <- mk_figFDdescr_for_OHBM(final)
+saveRDS(object = fig_FDdescr, file = "../report/fig/fig_FDdescr.rds")
 
 
 # figDesignmatrix ---------------------------------------------------------

### 2021-03-19 code/run_or_display.m

@@ -104,10 +104,10 @@ roi_prep = convertCharsToStrings(readcell(fullfile(param.INFO_DIR,"ROIs.txt"), "
 %% ------------------------------------------------------------------------
 
 param.PRESET = "manual";
-param.ONLY_DISPLAY = false;
+param.ONLY_DISPLAY = true;
 param.OVERWRITE = false; 
 param.WILD_BOOT = true;   
-param.parallel = true;
+param.parallel = false;
 
 if param.PRESET == "standard"
     param.MODEL = ["grouptime","grouptime2tp", "bmi", "bmiIG","bmi2tp", "fd","fdIG"];
@@ -182,7 +182,7 @@ elseif param.PRESET == "manual"
     % association to a model is indicated by the tens digit (GroupTime_: 1_; 
     % BMI_ = 2_; FD_ = 3_) the specific covariate combination by the ones digit
     param.MODEL = ["grouptime2tp"]; % ["grouptime","grouptime2tp"] % ["bmi","bmiIG","bmi2tp"] % ["fd","fdIG"] % ["alltp"] % ["singletp"]
-    param.COVARIATES = [11,12,13,14];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
+    param.COVARIATES = [14];     % [11, 12];                    % [21, 22];                % [31, 32];  % [41, 42, 43]    % [41, 42, 43]
 
     % define masking and type of inference
     param.MASK = "brain";               % "brain"

### 2021-03-19 code/tables.R

@@ -88,57 +88,57 @@ for (r in 1:nrow(result_report)) {
 df1 <- read.csv(file.path(res_fc$result_dir,res_fc$BMImodel_dirs[1],res_fc$deact_csv))
 colnr <- ncol(df1)
 df1$seed <- NULL
-df1[1,"seed"] <- c("PCC (cc)")
+df1[1,"seed"] <- c("PCC (CC)")
 df1$covariates <- NA
 df1$covariates[1] <- "age, sex, log mFD"
 # BMIagesex_PCCcc_avgBMI_deact 
 df2 <- read.csv(file.path(res_fc$result_dir,res_fc$BMImodel_dirs[2],res_fc$deact_csv))
 df2$seed <- NULL
-df2[1,"seed"] <- c("PCC (cc)")
+df2[1,"seed"] <- c("PCC (CC)")
 df2$covariates <- NA
 df2$covariates[1] <- "age, sex"
 
 # BMIFDagesex_NACCcc_cggBMI_deact 
 df3 <- read.csv(file.path(res_fc$result_dir,res_fc$FDBMImodel_dirs[1],res_fc$deact_csv))
 df3$seed <- NULL
-df3[1,"seed"] <- c("NAcc (cc)")
+df3[1,"seed"] <- c("NAcc (CC)")
 df3$covariates <- NA
 df3$covariates[1] <- c("age, sex")
 # BMIFDagesex_NACCcc_avgFD_act
 df4 <- read.csv(file.path(res_fc$result_dir,res_fc$FDBMImodel_dirs[2],res_fc$act_csv))
 df4$seed <- NULL
-df4[1,"seed"] <- c("NAcc (cc)")
+df4[1,"seed"] <- c("NAcc (CC)")
 df4$covariates <- NA
 df4$covariates[1] <- "age, sex"
 # BMIFDagesex_NACCgsr_cggBMI_deact 
 df5 <- read.csv(file.path(res_fc$result_dir,res_fc$FDBMImodel_dirs[3],res_fc$deact_csv))
 df5$seed <- NULL
-df5[1,"seed"] <- c("NAcc (gsr)")
+df5[1,"seed"] <- c("NAcc (GSR)")
 df5$covariates <- NA
 df5$covariates[1] <- "age, sex"
 # BMIFDagesex_NACCgsr_avgFD_act 
[…]

### 2021-03-22 manuscript/draft_rsfmri.Rmd
changes

@@
 ---
-title: "Effects of bariatric surgery on resting-state functional connectivity of the reward and default mode network: a pre-registered analysis"
+title: "Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis"
 author:
@@
 - Evelyn Medawar [1]
-- Kristin Prehn [3][4]
-- Jürgen Ordemann [7][8]
-- Agnes Floel [5][6]
-- A. Veronica Witte [1][2][9]
+- Kristin Prehn [4][5]
+- Jürgen Ordemann [8][9]
+- Agnes Floel [4][5]
+- A. Veronica Witte [3][1][2]
 date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
-  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 4 Department of Psychology, Medical School Hamburg, Hamburg \newline 5 Department of Neurology, University of Greifswald, Greifswald \newline 6 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 7 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 8 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin \newline 9 Day Clinic for Cognitive Neurology, University Clinic Leipzig
+  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Clinic of Cognitive Neurology, University Clinic Leipzig \newline 4 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 5 Department of Psychology, Medical School Hamburg, Hamburg \newline 6 Department of Neurology, University of Greifswald, Greifswald \newline 7 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin
 output:
@@
 
-Obesity imposes serious health risks and relates to differences in resting-state functional connectivity (FC) of brain networks involved in eating behavior.
-Bariatric surgery is an effective treatment, but its effects on FC are still under debate.
-In this pre-registered analysis, we aimed to determine the effects of bariatric surgery on FC in the reward (RN) and default mode network (DMN).
+Obesity imposes serious health risks and involves alterations in resting-state functional connectivity  of brain networks involved in eating behavior.
+Bariatric surgery is an effective treatment, but its effects on functional connectivity are still under debate.
+In this pre-registered study, we investigated the effects of bariatric surgery on major resting-state brain networks (reward and default mode network) in a longitudinal controlled design.
 33 bariatric surgery patients and 15 obese waiting-list control patients (37 females; aged 44.15 ± 11.86 SD years (range 21-68)) underwent magnetic resonance imaging at baseline, after 6 and 12 months.
-We pre-registered (https://osf.io/f8tpn/https://osf.io/59bh7/) a whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network FC.
-In exploratory analyses, we investigated all three time points and assessed the effects of weight loss and head motion.
-Bariatric surgery compared to waiting did not significantly affect FC of the RN and DMN (FWE p > .05).
-There also was no interaction effect on within-network FC.
-In exploratory analyses, surgery-related BMI decrease (FWE p= .041) and higher average head motion (FWE p= .021) resulted in significantly stronger FC in medial posterior frontal regions within the RN.
-This pre-registered analysis of longitudinal neuroimaging data did not support a strong effect of bariatric surgery, compared to waiting, on RN and DMN FC after 6 months.
-Exploratory analyses indicated that differences in head motion might have confounded the effects, thus a more rigorous control of within-scanner head motion during data acquisition is needed to better understand the effects of bariatric surgery on brain organization.
+We conducted a pre-registered whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network connectivity (https://osf.io/f8tpn/, https://osf.io/59bh7/).
+In exploratory analyses, we investigated the effects of weight loss and head motion.
+Bariatric surgery compared to waiting did not significantly affect functional connectivity (FWE-corrected $p > .05$), neither whole-brain nor within-network.
+In exploratory analyses, surgery-related BMI decrease (FWE-corrected $p = .041$) and higher average head motion (FWE-corrected $p = .021$) resulted in significantly stronger connectivity of the reward network with medial posterior frontal regions.
+This pre-registered well-controlled study did not support a strong effect of bariatric surgery, compared to waiting, on major resting-state brain networks after 6 months.
+Exploratory analyses indicated that head motion might have confounded the effects.
+Data pooling and more rigorous control of within-scanner head motion during data acquisition are needed to substantiate effects of bariatric surgery on brain organization.
 
@@
 Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
-RYBG is the most invasive, yet most effective surgical pro-cedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].
+RYBG is the most invasive, yet most effective surgical procedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].\
 Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla_2017; @Brutman_2019].
 Precise mechanisms how bariatric surgery leads to altered appetitive signaling are yet to be elucidated.
-Resting-state functional magnetic resonance imaging (rsfMRI) is a technique to investigate the dynamic organization of the brain.
+One option to adress these questions on brain-behavior relationships is to use resting-state functional magnetic resonance imaging (rsfMRI), a technique capturing the dynamic organization of the brain.
 Functional connectivity (FC) networks, i.e.
 brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009].
-The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
-The RN (RN), comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (Nacc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
-These brain regions have been found to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
+The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.\
+The RN comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
+These brain regions have been suggested to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
 Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
 RsfMRI studies also showed increased local FC of reward network regions, i.e.
-Nacc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
+NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
 In bariatric surgery patients, connectivity  within the RN (e.g.
@@
 Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018] .
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018].\
 The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle_2015].
@@
 salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
-After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
+After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].\
 In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
-Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].
-Further, while higher BMI has been consistently associated with more head motion during rsfMRI, previous studies in bariatric surgery patients have not taken this important confounder of FC into account [@Beyer_2017;@Hodgson_2016].
+Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].\
+Further, while higher BMI has been consistently associated with more head motion during rsfMRI [@Beyer_2017;@Hodgson_2016], previous studies in bariatric surgery patients have not taken this important confounder of FC into account.
 In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
-Thus, we aimed to rigorously control for motion-related variance in our analyses.
+Thus, we aimed to rigorously control for motion-related variance in our analyses.\
 We had the following confirmatory hypotheses:
@@
 We pre-registered two denoising pipelines, and three covariate schemes.
-For hypothesis 2, we performed a planned analysis of the group-by-time interaction on aggregated, within-network FC, for two and three time points and the same covariate schemes.
+For hypothesis 2, we performed a confirmatory analysis of the group-by-time interaction on aggregated, within-network FC, for two and three time points and the same covariate schemes.
 In exploratory analyses, we examined the whole-brain interaction effect for three time points, the effects of head motion on FC and whether weight loss, a proxy of treatment success, predicted changes in FC.
@@
 For more details see [@Prehn_2020].
-We used all data acquired until April 2019.
+We used all data acquired until April 2019.\
 The study protocol was in accordance with the Helsinki Declaration and approved by the ethics committee of the Charité University Medicine Berlin.
 The study was registered at clinicaltrials.gov as NCT01554228.
-We pre-registered the present resting-state analyses on the OSF https://osf.io/yp42s.
-We made additional changes https://osf.io/59bh7/ to the pre-registration after preprocessing as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
-<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA -->
-
+We pre-registered the present resting-state analyses on the OSF (https://osf.io/yp42s).
+We made additional changes (see https://osf.io/59bh7/) to the pre-registration after preprocessing the rsfMRI data as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
+For a comparison of the pre-registration and the manuscript, please visit https://osf.io/45n9f/.\
 Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin.
@@
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).
+Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).\
 In total, $51$ participants out of the originally enrolled $69$ subjects received MRI.
-Five data points of three subjects had to be excluded due to bad anatomical image quality (see below), which led to a final data set with $101$ rsfMRI sessions.
-<!-- DESCRIPTIVE STATISTICS -->
+Five data points of three subjects had to be excluded due to bad anatomical image quality (see below), which led to a final data set with $101$ rsfMRI sessions.\
 The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`)).
-`r round(sum(subj_df$arterial_hypotension == 1, na.rm = T) / nrow(subj_df), 3)*100`% of participants had clinically diagnosed hypertension, `r round((sum(subj_df$diabetes_type_II == 1, na.rm = T) + sum(subj_df$diabetes_type_I == 1, na.rm = T))  / nrow(subj_df), 3)*100`% had type-2 or type-1 diabetes, and `r round(sum(subj_df$smoking_status == 1, na.rm = T) / nrow(subj_df), 3)*100`% reported to smoke.
-Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
+`r round(sum(subj_df$arterial_hypotension == 1, na.rm = T) / nrow(subj_df), 3)*100`% of participants had clinically diagnosed hypertension, `r round((sum(subj_df$diabetes_type_II == 1, na.rm = T) + sum(subj_df$diabetes_type_I == 1, na.rm = T))  / nrow(subj_df), 3)*100`% had type-2 or type-1 diabetes, and `r round(sum(subj_df$smoking_status == 1, na.rm = T) / nrow(subj_df), 3)*100`% reported to smoke.\
+Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` females) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` females), who waited for their health insurance's approval to undergo surgery.
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016].
 Analyses were performed on all participants who provided at least one data point of rsfMRI data.
-`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point (for more details see Table S1).
-The pre-planned analysis of changes from baseline to follow-up included $24$ participants with both time points, in total $72$ data points.
-
+`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point (for more details see Figure 1 in supplementary information (SI)).
+The pre-registered analysis of changes from baseline to followup included $24$ participants with both time points, in total $72$ data points.\
 `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the intervention group information was not available.
-
+\
 Participants arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast.
 They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
-MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020)).
+MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020).
 
@@
 
-Imaging data was conducted using AFNI 19.1.05, ANTS '2.3.1', FSL '6.0.1' and FreeSurfer '6.0.0p1', wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15.
-The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel. The workflow can be found on https://github.com/fBeyer89/ADI_preproc/.
-
-### Anatomical preprocessing
-Within the module for anatomical preprocessing, the T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross].
+### Minimal preprocessing
+Imaging data analysis was conducted using AFNI 19.1.05, ANTS 2.3.1, FSL 6.0.1 and FreeSurfer 6.0.0p1, wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15 which can be found on https://github.com/fBeyer89/ADI_preproc/.
+T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross].
 Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long].
@@
 The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants]. 
-The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
-
-## Functional preprocessing
-
-### Minimal functional preprocessing
-The functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
+Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
 In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
@@
 
-## Denoising pipelines
-Among several denoising pipelines, ICA-AROMA provides an acceptable trade-off between the mitigation of motion-FC correlation and the introduction of a distance-dependence on motion-FC relation-ship [@Parkes_2018].
-@Ciric_2017 showed that GSR is very efficient in re-moving the correlation of head motion and FC, and outperforms ICA-AROMA alone.
-Yet, it introduces spatial dependency and spurious correlations.
-Thus, we pre-registered the main analyses to be performed with both denoising pipelines.
-
-### ICA-AROMA denoising pipeline   
-As described in @pruim2015ica, the minimally preprocessed data underwent independent component analysis (ICA) using FSL's `melodic` and nuisance components were classified according to four spectral and spatial criteria.
-Finally, the timeseries of the  nuisance components were regressed from the raw rsfMRI timeseries non-aggressively.
-As recommended in the original paper, we performed regression of WM and CSF signal on the ICA-AROMA result.
-We used CompCor to estimate 5 variance components from a combined WM and CSF mask, which was further eroded using `fslmaths -nan -thr 0.99 -ero -bin` [@Behzadi_2007;@Muschelli_2014].
-After regression of these nuisance components from the data, we performed high-pass filtering at 0.01 Hz.
-
-### Global Signal Regression (GSR) 
-The global signal was derived from the average of all voxels in the brain mask.
-Then, we regressed WM and CSF CompCor components along with the GS from the ICA-AROMA rsfMRI data and high-pass filtered the resulting file as above.
-
-### Functional connectivity (FC)
-To derive RN and DMN FC maps, we used Nacc and precuneus as seed regions of interest (ROI), respectively.
-We did not select vmPFC for the RN due to low SNR.
-Because we knew about the large amount of motion in this dataset beforehand, we decided against ICA to construct the networks, as ICA is recommended to work on minimally preprocessed data.
-Based on FreeSurfer’s aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for Nacc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
-Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+GSR).
-Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
+### Denoising pipelines
+In the pre-registration, we specified two denoising pipelines (AROMA+CC including ICA-AROMA + CompCor and AROMA+CC+GSR including ICA-AROMA + CompCor + global signal regression (GSR)) [@Parkes_2018; @Ciric_2017, for details see SI].
 
@@
 
-#### Anatomical Quality control
+**Anatomical quality control**
 FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
-We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.
+We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.\
 <!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
+**RsfMRI quality control**
+RsfMRI quality control was performed according to the protocol by @Ciric_2018 (see SI for more details). 
+Head motion was quantified using mean framewise displacement (mFD) according to @Power_2012, and log-transformed for further analysis (logmFD).
+As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
 
-#### RsfMRI quality control (QC)
-RsfMRI quality control was performed according to the protocol by @Ciric_2018. 
-We calculated mean framewise displacement (mFD) according to @Power_2012, and log-transformed it for further analysis (logmFD).
-As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
-Instead, we applied best-practice motion correction (AROMA+CC and AROMA+CC+GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018].
+## Functional connectivity (FC)
+To derive RN and DMN FC maps, we used NAcc and PCC/precuneus as seed regions of interest (ROI), respectively.
+We did not select vmPFC for the RN due to low SNR.
+Based on FreeSurfer's aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
+Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
+We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+CC+GSR).
+Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
 
-##### Individual level resting-state rsfMRI QC
-```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
-carpets=c("../report/carpetplotadi006fu2.jpeg","../report/carpetplotadi16fu.jpeg")
-knitr::include_graphics(carpets)
-```
-On the individual level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality.
-Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum).
-All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`, and checked again visually.
-No participant was excluded based on this visual inspection.
+## Aggregated within-network FC 
+To extract aggregated within-network FC, we first calculated the mean DMN and RN over all participants and time points, adjusted for age and sex.
+We used GSR-denoised data as input and clusterwise bootstrapping with $N = 1000$.
+Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
+We extracted the average GSR-denoised FC from these masks.
 
-##### Group level rsfMRI QC
+## Statistical Analysis
 
-As described in @Beyer_2020, there was a group-by-time interaction on mFD (see Figure \@ref(fig:figDvarsmFD), upper right panel).
-The intervention group reduced in head motion in the follow-up assessments compared to the control group.
-For this study, we further explored mFD and its relation to DVARS, the root mean square of the temporal change of the voxel-wise signal at each time point.
-DVARS measures volume-to-volume signal variation.
-```{r figDvarsmFD, out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', fig.show='hold',fig.align='center', echo=FALSE}
-figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
-#(figList$fig1 | figList$fig2) / figList$fig3
-
-lay <- rbind(c(1,1,2,2),
-             c(NA,3,3,NA))
-gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
-```
-DVARS did not qualitatively differ between groups and time points, and the correlation of DVARS and mFD remained similar over time points (see Figure \@ref(fig:figDvarsmFD).
-To further assess the confounding of head motion and FC, we evaluated mFD-FC relationships by computing a functional connectome based on the protocol by [@Ciric_2018].
-We used the MNI coordinates of 246 cortical and subcortical spherical seed regions from @Power_2012 with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series.
-Then, we calculated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation.
-```{r mFDFC, echo=FALSE}
-res <- read.csv("../report/FDFCcorrelations.csv")
-rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
-colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')
-```
-```{r tableQcfc, out.width="100%", echo=FALSE}
-tab_Qcfc <-
-  knitr::kable(
-    round(res, 3),
-    col.names = colnames(res),
-    row.names = TRUE,
-    format = "latex",
-    booktabs = T,
-    linesep = "",
-    caption = "Summary of quality metrics for different denoising pipelines across conditions and time points"
-  ) %>%
-  column_spec(c(2:6), width = "1.8cm")
-tab_Qcfc
-```
-In the minimally preprocessed data, mFD and FC between nodes were significantly positively correlated (mean Spearman's $r$ = `r round(res[1,1],2)`)(see Table \@ref(tab:tableQcfc)).
-This relationship was negatively associated with distance between nodes (mean Spearman's $r$ = `r round(res[1,5],2)`).
-Denoising pipelines (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and FC, and reduced the distance dependency.
-Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency [@Power_2015].
-Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r$ = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r$ = `r round(res[4,5],2)`)
-In addition, for models showing a significant time-by-group interaction, we planned to perform a sensitivity analysis excluding the 10% data sets with highest mFD.
-
-### Aggregated FC values
-To extract aggregated FC values, we first calculated the mean DMN and RN over all participants and Naccs, adjusted for age and sex.
-We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
-Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
-We extracted the average FC from these masks for GSR-data.
-
-## Analysis
-
-Statistical analysis were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
--@spm12)<!-- spm_check_installation('rev') -->, and R version 3.6.1 [@team2013r].
-
-### Design Specification
+Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
+-@spm12).
+The marginal model implemented in the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
+We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size.
+We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability) for all analyses.
+Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].
 
 <!-- MODEL SETUP --> 
-#### Confirmatory Analysis
-We tested the pre-registered hypothesis of a time-by-group interaction:  $FC=group + time + group * time$. 
-As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (CA1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
+### Confirmatory analyses
+**Whole-brain analysis**
+We tested the pre-registered hypothesis of a time-by-group interaction:  `FC=group + time + group * time`.\
+As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both timepoints (model CA2) and age, sex, average logmFD and baseline BMI (model CA3).
+Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values.
+This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.\
+**Aggregated FC analysis**
+We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007].
+We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline.
+First, we investigated the time-by-group interaction for baseline and followup time points only.
+We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3).
+We performed model comparison between R1 and R0 models, where $R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))$, and $R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))$.
+As specified in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.
+
+### Exploratory analysis (pre-registered)
+**Whole-brain analysis**
+As described in the pre-registration, we performed an exploratory analysis of the relationship of average and change in BMI in a separate group of models.
+As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
+This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC.
+We first estimated their effects in a model adjusting for baseline age and sex (model EA1) and then additionally controlling for logmFD (model EA2).
+As we previously reported correlated change in BMI and head motion in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (model EA3).
+Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.\
+**Aggregated FC analysis**
+We also performed an exploratory analysis with average BMI and change in BMI as predictors of the aggregated FC from AROMA+CC+GSR denoised data of both networks.
+We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (model EA1), logmFD (model EA2) and average and change in logmFD (model EA3).
+
+### Exploratory analysis (not pre-registered)
+**Whole-brain analysis**
+**Group-by-time interaction in three time points**
+We explored our whole-brain group-by-time point interaction models for the data of all three time points.
 In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
-The resulting factorial design contained one regressor for each time point per group (see supplements for depiction of design matrix).
-Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package mice based on subject identification, condition, baseline age, and available BMI values.
-This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.
-
-#### Exploratory analysis
-##### Group-by-time interaction in three time points
-Then, we extended our whole-brain analysis of the group-by-time point interaction to the data from all three time points.
+The resulting factorial design contained one regressor for each time point per group (see SI for depiction of design matrix).
 This analysis had not been pre-registered.
 Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points.
-We investigated two models (EA1.1: group x time, age, sex and EA1.2: group x time, age, sex and logmFD).
+We investigated two models adjusting for age and sex (model EA4) and age, sex and logmFD (model EA5).\
+**Average and change in logmFD**
+Further, we tested the association of head motion and FC in another exploratory model (model EA6): `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates.
+With this model, we aimed for a better understanding of the unique contribution of average and longitudinal change in logmFD measures.
 
-##### Average and change BMI
-We further  explored the relationship of average and change in BMI in a separate group of models.
-As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
-This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (Model EA2.1) and then additionally controlling for logmFD (Model EA 2.2).
-As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 2.3).
-Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
-For a better understanding of the unique contribution of average and longitudinal change in logmFD measures, we separately employed a third post-hoc, exploratory model (EA 2.4), solely including the head motion measures as covariates of interest: `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates.
+### Statistical Inference
+
+**Whole-brain analysis**
+<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
+Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual.
+Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$, and 1000 bootstraps due to required computation time (instead of 5000)).
+Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
+**Aggregated FC analysis**
+The interaction effect of group and time point in the models of aggregated FC was considered significant if the model comparison between R1 and R0 models using the anova command showed $p < 0.05$.
+In all exploratory models, we considered all coefficients with $p < 0.05$ as significant.
 
 ### Functional decoding
-In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
-We uploaded the contrast images for change BMI and average logmFD on [Neurovault](), and applied the decoding classifier.
+In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the NeuroSynth (https://www.neurosynth.org/) database [@Yarkoni_2011].
+We uploaded the contrast images for change BMI and average logmFD on Neurovault (https://neurovault.org/collections/9426/), and applied the decoding classifier.
 This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
 We reported the three top terms for both contrasts.
-
-### Model input and brain mask
-For each brain network, we examined the models for the two denoising pipelines (AROMA+CC/AROMA+CC+GSR) as pre-registered.
-For all analyses, we used an explicit brain mask, which was a whole brain derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability).
-
-### SwE settings
-<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
-It therefore accommodates any repeated measurement covariance structures.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
-
-<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
-Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual.
-Opposed to the description in the pre-registration, we used a cluster forming threshold of p<0.001 for a more rigorous multiple comparison adjustment (instead of p<0.01, and 1000 bootstraps due to required computation time (instead of 5000)).
-Significant clusters are defined as family-wise error (FWE) corrected p<0.05.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.
-
-### Aggregated FC analysis
-#### Confirmatory analyses
-<!-- R analysis of aggregated values -->
-We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007].
-We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline.
-First, we investigated the time-by-group interaction for baseline and follow-up time points only.
-We either performed no correction (CA1), adjusted for baseline age, sex, average of logmFD of both time points (CA2) or additionally adjusted for baseline BMI (CA3).
-We performed model comparison between R1 and R0 models, where $R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))$, and $R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))$.
-As planned in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.
-The interaction effect was considered significant if the model comparison between R1 and R0 models using the anova command showed p<0.05.
-
-#### Exploratory analysis
-We performed an exploratory analysis with average BMI and change in BMI as predictors of the aggregated FC from AROMA+CC+GSR denoised data of both networks, in line with our exploratory whole brain analysis.
-We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (Model EA2.1), logmFD (Model EA 2.2) and average and change in logmFD (Model EA 2.3).
 
@@
 In the intervention group, baseline MRI data is not missing completely at random, due to the limited scanner bore.
-Similarly, we observed a higher drop-out in the control than in the surgery group (baseline timepoint only, n control = ; n surgery = ).
 Change in BMI throughout the study is depicted in Figure \@ref(fig:figBMIdescr).
 
-```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
+```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectory of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
 # eval = TRUE, fig.asp = 0.7, fig.width = 6
@@
 
-## Whole-brain analyses: Confirmatory analysis 
-First, we performed the pre-registered analysis in participants with baseline and followup data, using two denoising pipelines and three covariate adjustements.
+## Confirmatory analysis 
+### Whole-brain analysis
+Against our initial hypothesis, there was no interaction effect of group and time point on neither RN nor DMN FC in model CA 1 (no adjustments).
+There also was no significant main effect for any of the effects of interest (time, group) in clusterwise inference with FWE-correction.
+The same was true for models CA2 (adjusting for age, sex and average of logmFD). Results did not differ between AROMA+CC and AROMA+CC+GSR denoising pipelines.
+In model CA3 (adjusting for age, sex, average of logmFD and baseline BMI) there also was no significant interaction when adding baseline BMI.
+Yet, we found a significant main effect of time in this model.
+For AROMA+CC+GSR denoised data, we found decreased FC of the NAcc with the lateral occipital cortex at baseline compared to followup (FWE-corrected $p = 0.030$), and decreased FC of the PCC within the DMN to the medial anterior cingulate cortex (FWE-corrected $p = 0.046$) (see SI).
+<!-- Neurovault -->The unthresholded contrast maps of group, time and group-by-time interaction for the unadjusted model for AROMA+CC+GSR are uploaded on NeuroVault [https://neurovault.org/collections/9426/](https://neurovault.org/collections/9426/).
 
-### Model CA1 (without adjustments)
-Against our initial hypothesis, there was no interaction effect of group and time point for FC for neither the Nacc nor the PCC seeds.
-There also was no significant main effect for any of the covariates of interest (time, group) in clusterwise inference with FWE-correction.
+### Aggregated FC analysis
+```{r load agg FC data and prepare, echo = FALSE}
+final_FC <- read.csv("../report/final_FC.csv")
+```
 
-###  Model CA2 (adjusted for age, sex and average of logmFD)
-Similarly, there were no significant clusters when adjusting for age, sex and average of logmFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.
+```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the RN"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
+fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
+fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
+   fig_DMNconn
+fig_Rewconn
+```
+There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure \@ref(fig:figaggFC) and SI for a detailed summary of the models), regardless of the adjustments (CA1 without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
 
-###  Model CA3 (adjusted for age, sex, average of logmFD and baseline BMI)
-There was no significant interaction when adding baseline BMI.
-Yet, we found a significant main effect of time in this model.
-For AROMA+CC+GSR preprocessed data, we found higher FC of the Nacc with the lateral occipital cortex at baseline compared to followup (FWE $p=0.030$), and higher FC of the PCC within the DMN to the medial ACC (FWE $p=0.046$)(see supplements).
-
-<!-- Neurovault -->The unthresholded contrast maps of group, time and group-by-time interaction for the unadjusted model are uploaded on [NeuroVault](https://neurovault.org/collections/9426/)).
-
-## Whole-brain: Exploratory Analyses
-### Group-by-time interaction for three time points (EA 1.1 and EA 1.2)
-Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA 1.1 and EA 1.2 in the full data set for neither RN nor DMN.
-
-### Whole-brain effects of mean and change in BMI
-#### RN
-
-##### Average and change in BMI, adjusted for logmFD (EA 2.1 and EA 2.2)
-Nacc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes and regardless whether we adjusted for logmFD.
-
-##### Average and change in BMI, adjusted for split logmFD (EA 2.3)
-<!-- cc -->For the RN, we found a significant negative association of change in BMI with FC between Nacc and a cluster in the posterior-medial frontal region. 
-The peak voxel derived from investigating the effect of average logmFD was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), even if data was GSR ($p = 0.038$).
-```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Cluster showing a significant negative effect of change BMI on Nacc to posterior-medial frontal region, adjusted for age, sex, mean BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+## Exploratory Analyses (pre-registered)
+### Whole-brain analysis
+RN FC was not significantly related to neither average nor change in BMI, for either of the denoising pipelines and regardless whether we adjusted for logmFD in models EA1 and EA2.
+Only in model EA3 (adjusting for average and change in both BMI and logmFD), we found that more BMI decrease (e.g.
+weight loss) predicted higher FC between NAcc and a cluster in the posterior-medial frontal region (see Figure \@ref(fig:BMIFDagesexNacccgnBMIdeact) and Table \@ref(tab:tableBMIFDmodel)).\
+The peak voxel was classified as belonging to superior frontal gyrus ($45$% probability) and supplementary motor area (SMA) ($37$%  probability) in the Harvard-Oxford atlas.
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.030$), also in AROMA+CC+GSR-denoised data ($p = 0.041$).\
+```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 ```
-<!-- gsr -->Moreover, FC between the Nacc and one cluster in a more posterior region, i.e.
-motor cortex, was significantly positively related to average logmFD (see  Figure \@ref(fig:BMIFDagesexNaccavgFDact)).
-This result was similar for AROMA+CC+GSR preprocessed data.
-```{r "BMIFDagesexNaccavgFDact", fig.cap="Cluster showing a significant positive Effect of average logmFD on FC between Nacc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
@@
 ```
-<!-- RN connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
-Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
--->
-
-#### DMN
-
-##### Average and change in BMI, adjusted for mFD (EA 2.1 and EA 2.2)
-Using the images preprocessed with AROMA+CC, we found a significant negative association of average BMI and DMN connectivity in the lingual gyrus, mid orbital gyrus and temporal gyrus, which remained stable after including logmFD as nuisance covariate (FWE-cluster p < 0.05, see Figure \@ref(fig:BMIagesexPCCccavgBMIdeact)).
-When using GSR-data, none of the clusters was significantly associated with average BMI.
-Results with significant cluster activation are shown in Table \@ref(tab:tableBMImodel).
-```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", fig.subcap = c("adjusted for age and sex", "adjusted for age, sex and logmFD"), fig.hold=TRUE, out.width = "50%"}
+Moreover, average logmFD was positively associated with FC between NAcc and motor cortex in model EA3 for both denoising pipelines (see Figure \@ref(fig:BMIFDagesexNaccavgFDact) and Table \@ref(tab:tableBMIFDmodel)).\
+For the DMN, we found that higher average BMI predicted lower FC of the precuneus/PCC with the lingual gyrus, mid orbital gyrus and temporal gyrus in the images denoised with AROMA+CC.
+This finding was significant in models EA1, EA2 and EA3 (see Figures \@ref(fig:BMIagesexPCCccavgBMIdeact) and \@ref(fig:BMIFDagesexPCCccavgBMIdeactt) and Table \@ref(tab:tableBMImodel)).
+Yet, none of the clusters survived when using AROMA+CC+GSR denoised data (see Table \@ref(tab:tableBMIFDmodel).\
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (Model EA1)", "adjusted for age, sex and logmFD (EA2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
 knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
 ```
-
-##### Average and change in BMI, adjusted for split logmFD (EA 2.3)
-
-To disentangle effects of change in BMI and logmFD in this dataset, we split logmFD into average logmFD and change in logmFD. Age and sex were kept as nuisance covariates.
-<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting logmFD into its components (Figure \@ref(fig:BMIFDagesexPCCccavgBMIdeact2)).
-<!-- gsr -->As before, the effect did not survive when using AROMA+CC+GSR preprocessed (see Table \@ref(tab:tableBMIFDmodel).
-```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC of PCC/precuneus with a cluster, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%"}
+```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data (Model EA3)", out.width = "50%", fig.align="center"}
 knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
@@
 <!-- Neurovault -->
-Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault https://neurovault.org/collections/9426/.
+Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault [https://neurovault.org/collections/9426/](https://neurovault.org/collections/9426/).
 
-##### Functional decoding
-Functional decoding of these two activation patterns showed different top association terms.
-For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract.
-The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
+### Aggregated FC analysis
+As expected, there was no association of average BMI or within-subject BMI change and within RN FC in models EA1, EA2 or EA3 adjusting for age, sex and logmFD (for details see SI).\
+```{r "Rew model average&change of BMI&logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvwithinFD=anova(R1,R03)
+anvmeanFD=anova(R1,R04)
 
-##### Effect of logmFD (EA 2.4)
+test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
+vif_res=vif(test_vif)
+```
+```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
 
-To explore the unique contributions of the average value of and the longitudinal change in logmFD, we constructed an additional model, where BMI measures were excluded from the design matrix.
-Covariance of the logmFD and outcome measures was investigated by splitting logmFD into average logmFD and longitudinal logmFD change and by inspecting associations with each predictor.
-Age and sex were used as nuisance covariates.
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
+R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
+anvmeanBMI_adjmFD=anova(R1,R02a)
+p_meanBMI_adjmFD=anvmeanBMI_adjmFD$`Pr(>Chisq)`[2]
+```
+```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
+R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
+R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
+R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
+anvwithinBMI=anova(R1,R01)
+anvmeanBMI=anova(R1,R02)
+anvmeanFD=anova(R1,R03)
+anvwithinFD=anova(R1,R03)
 
-<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), there was a significant positive association between average logmFD and co-activation of Nacc with one cluster located in proximity to the central sulcus and motor areas.
-Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
-<!-- ### DMN -->We did not find any clusters with a significant association of logmFD and PCC FC.
-```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of Nacc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
+p_mean_BMI_splitmFD=anvmeanBMI$`Pr(>Chisq)`
+```
+Like in the whole-brain analysis, higher average BMI was associated with reduced aggregated DMN FC, regardless of whether we adjusted for logmFD (model EA1: $p$ = `r round(p_meanBMI,3)` and EA2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
+The association also remained significant when we split logmFD into average and change in logmFD (EA3) ($p$ = `r round(p_mean_BMI_splitmFD,3)[2]`) and there was no significant association of average or change in logmFD with DMN FC (see SI for overview of all models).
+
+## Exploratory analsis (not pre-registered)
+### Group-by-time interaction in three time points 
+Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA4 and EA5 in the full data set for neither RN nor DMN.
+
+### Average and change in logmFD 
+<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), higher average logmFD was associated with stronger FC between the NAcc and a cluster located in proximity to the central sulcus and motor areas.
+This Cluster only differed in size between denoising pipelines.\
+We did not find any clusters with a significant association of logmFD and DMN FC.
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with a cluster in motor cortex, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
@@
 ```
+### Functional decoding
+Functional decoding of the two activation patterns from the pre-registered, exploratory analysis EA3 showed different top association terms.
+For the [https://neurosynth.org/decode/?neurovault=441783](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula and inferior frontal were the top terms, while for the [https://neurosynth.org/decode/?neurovault=441784](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract.
+The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
 
@@
 
-## Aggregated FC analysis
-### Confirmatory analysis of group-by-time interaction (CA1, CA2, CA3)
-```{r load agg FC data and prepare, echo = FALSE}
-final_FC <- read.csv("../report/final_FC.csv")
-```
-
-```{r figaggFC, fig.cap="Mean network connectivity per group over time separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", fig.subcap = c("for the DMN", "for the RN"), fig.hold=TRUE, out.width = "50%", fig.align="center", echo = FALSE}
-fig_DMNconn<-readRDS(file = "../report/fig/figDMNconn.rds")
-fig_Rewconn<-readRDS(file = "../report/fig/figRewconn.rds")
-   fig_DMNconn
-fig_Rewconn
-```
-There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure @ref(fig:figaggFC) and supplements for a detailed summary of the models), regardless of the adjustments (CA1: without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
-Thus, we did not repeat the analysis with-out the 10% of highest motion subjects, as pre-registered.
-
-### Exploratory analysis: effects of mean and change in BMI on aggregated FC
-#### RN
-##### Models EA 2.1, 2.2. and 2.3
-```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-anvwithinFD=anova(R1,R03)
-anvmeanFD=anova(R1,R04)
-
-test_vif=lm(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex, data=final_FC)
-vif_res=vif(test_vif)
-```
-
-As expected from the whole-brain findings, there was no association of average BMI or within-subject BMI change and within RN connectivity in models EA 2.1, 2.2 or 2.3 adjusting for age, sex and logmFD.
-
-#### DMN
-
-##### Models EA 2.1, 2.2. and 2.3
-
-```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]
-
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
-R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
-anvmeanBMI_adjmFD=anova(R1,R02a)
-p_meanBMI_adjmFD=anvmeanBMI_adjmFD$`Pr(>Chisq)`[2]
-```
-
-```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
-R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
-R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
-R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
-anvwithinBMI=anova(R1,R01)
-anvmeanBMI=anova(R1,R02)
-anvmeanFD=anova(R1,R03)
-anvwithinFD=anova(R1,R03)
-
-p_mean_BMI_splitmFD=anvmeanBMI$`Pr(>Chisq)`
-```
-Corresponding to the whole brain results, there was a negative association of average BMI and aggregated DMN FC, regardless of whether we adjusted for logmFD (model EA 2.1: $p$ = `r round(p_meanBMI,3)` and EA 2.2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
-The association also remained significant when we split logmFD into average and change in logmFD (EA 2.3) ($p$ = `r round(p_mean_BMI_splitmFD,3)[2]`) and there was no significant association of mean or change in logmFD with DMN FC (see supplements material for overview of all models).
-
 # Discussion
@@
 In this pre-registered study, we investigated the effects of bariatric surgery on the FC of major resting-state brain networks in a longitudinal controlled design.
-Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising pipelines and controlling for mFD on the group level.
-We found no interaction effect of group and time point on whole-brain FC of the PCC and Nacc.
-We observed a main effect of time when adjusting for age, sex, average logmFD and BMI, and which pointed to higher FC at baseline compared to followup for both DMN and RN hubs.
-Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC.
-In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN FC for the more lenient denoising pipeline.
-When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI across the three time points and the connectivity of Nacc with a posterior-medial frontal cluster emerged.
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising pipelines and controlling for head motion on the group level.\
+We did not detect significant effects of bariatric surgery compared to waiting on whole-brain FC of the PCC and NAcc, core hubs of the RN and DMN, according to pre-registered whole-brain analyses.
+This was regardless of whether we adjusted for age, sex and individual head motion.
+DMN and RN FC was lower at baseline compared to followup for the whole group only when adjusting for age, sex, average logmFD and baseline BMI.
+In an exploratory model disentangling the effects of average and change in BMI, higher BMI was associated with lower DMN FC for the more lenient denoising pipeline.
+When we additionally adjusted for both average and change in head motion, decreases in BMI between the three time points were associated with increased connectivity of the NAcc with a posterior-medial frontal cluster.
 This result was significant in both denoising pipelines.
 Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns.
-Finally, higher average head motion was associated with increased Nacc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
-
-In this study, we could not confirm our pre-registered hypotheses and found no interaction effect between group and time point on DMN and RN FC.   
+Finally, higher average head motion was associated with increased NAcc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the change BMI cluster.\
+In this study, we could not confirm our pre-registered hypotheses.
 Based on previous studies in bariatric surgery patients, we expected within-DMN FC to increase, and DMN FC to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
 Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019].
-However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the Nacc and a cluster in a posterior-medial frontal brain region.
+These studies, notably, did not include adequate longitudinal obese control groups, making false-positive findings possible.
+We thus conclude that surgery-induced heavy weight loss does not strongly affect DMN and RN FC based on the current results.\
+However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the NAcc and a cluster in a posterior-medial frontal brain region.
 Based on the Harvard-Oxford atlas and NeuroSynth decoding this region might be part of the salience network and involved in action preparation.
-The enhanced FC between the Nacc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
+The enhanced FC between the NAcc and this regions seems at odds with our expectation of reduced hedonic drive to eat after bariatric surgery.
 One the other hand, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
@@
 a shift from integration towards segregation and more specialized processing.
-However, this interpretation is highly speculative, and more studies are needed to corroborate these findings.
-
-Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity.
-Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.
-To our surprise, the effect of weight loss on the connectivity of the Nacc with the posterior-medial frontal region was only present when separating average mFD and change in mFC and thereby introducing two instead of one regressor into the model.
+However, this interpretation is highly speculative, and more studies are needed to corroborate these considerations.\
+Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity, challenging definite conclusions on the relationship between weight loss and FC changes.
+Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.\
+To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFD and thereby introducing two instead of one regressor into the model.
 These results may be due to the presence of multi-collinearity between change in FD and BMI which might appear more pronounced in the split model, and thus lead to unreliable estimations of effects and standard errors.
 Contrarily, one could argue that only with the careful disentanglement of average and change in BMI and FD the effect of change in BMI could be singled out.
-This argument is supported by the survival of the cluster when using AROMA+CC+GSR preprocessed data, and the distinct results of the decoding analysis.
+This argument is supported by the survival of the cluster when using AROMA+CC+GSR denoised data, and the distinct results of the decoding analysis.
 Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
-Thus, confounding of the negative BMI change effect and the positive average head motion effect on posterior-frontal FC  seems unlikely.
+Thus, confounding of the negative BMI change effect and the positive average head motion effect on posterior-frontal FC seems unlikely.\
 Head motion also played a role in the association of higher BMI and reduced DMN FC.
@@
 Thereby, this results echoes a previous finding from our group, and may be interpreted as accelerated age-related decline of the DMN in relation to the cardiometabolic risk related to BMI.
-Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia_2017].
-
-The major strength of our study is prospective intervention controlled design.
-We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history and  treatment recommendation and were scanned after the same time intervals [@Thiese_2014].
-Another strength of our study was the pre-registered analysis plan, which was corrected after we realized some major flaws in the first version.
-In particular, we fixed the number of preprocessing schemes and models and determined that we would use SwE toolbox, a recommended statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
-Opposed to the flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
+Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia_2017].\
+The major strength of our study is the prospective intervention controlled design.
+We compared bariatric surgery patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history or recommendation and were scanned after the same time intervals [@Thiese_2014].
+Another strength of our study was the pre-registered analysis plan, which was corrected prior to statistical analysis after we noticed flaws in the first version.
+In particular, we gave more details on denoising pipelins and models and determined that we would use SwE toolbox, an advanced statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
+Opposed to the flexible factorial models which is the standard in SPM, marginal models use less degrees of freedom, and thus allow for the inclusion of covariates and higher power.\
 Limitations of our study include the low number of patients who participated in all three time points.
 In total, only $34$ participants contributed to the estimation of the longitudinal effects with at least two time points.
-Furthermore, patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI scanner.
-While this sample size is comparable to previous rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
-
+Patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI scanner.
+While this sample size is comparable to previous rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.\
 We used seed-based connectivity to derive large-scale brain networks.
 While this approach yielded reasonable DMN and RN maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
-Furthermore, we did not explicitly investigate the FC of the DMN and RN with other large-scale brain networks attention, e.g.
-cognitive control or sensorimotor networks, which has been previously proposed to be altered by obesity and bariatric surgery [@Lips_2014; @Olivo_2016; @Cerit_2019].
-
-We did not control for hunger or satiety in our design, although all participants were scanned after the intake of a breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
-Size and composition of our sample did not allow separate analyses for the sexes.
-However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
-
-Taken together, this study showed that effects of bariatric surgery on resting-state brain organization are, if at all, subtle, and often confounded by head motion.
-Yet, changes in BMI might relate to altered reward and salience processing, which might contribute to observed changes in food wanting and choice in patients after bariatric surgery.
-Further, larger, well-controlled and pre-registered neuroimaging studies are necessary to understand the neural underpinnings of altered gut-brain communication after bariatric surgery.
+Furthermore, our rsfMRI was relatively short, which might have further reduced our power.\
+We did not monitor hunger or satiety in our design, although all participants were scanned after the intake of a breakfast following an overnight fast.
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].\
+Size and composition of our sample did not allow sex-stratified analyses.
+However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].\
+Taken together, this prospective well-controlled study did not confirm previous findings claiming strong effects of bariatric surgery on FC of the RN and DMN in obese patients.
+In our cohort, differential changes in head motion adjustment strongly altered rsfMRI neuroimaging results.
+We thus recommend to rigorously control head motion at acquisition through online monitoring or prospective motion correction, to acquire longer rsfMRI scans or to investigate brain organization with less motion-prone techniques such as task-based fMRI.
+Furthermore, to increase validity and power, future studies should include obese control groups, and increase efforts to share and pool valuable patient data into meta-analysis.
+Finally, pre-registration of concrete and testable hypotheses would help to increase replicability of the field and truely enhance our understanding of the neural underpinnings of altered gut-brain communication after bariatric surgery.
 
 # Code availability
-
-All code is available in the github repository [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
-The code includes script for data preparation and analysis of different models.
+All code is available on github.
+Preprocessing scripts are here [https://github.com/fBeyer89/ADI_preproc](https://github.com/fBeyer89/ADI_preproc), and analyses scripts are here: [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
 
 # Declaration of interest
+The authors declare no competing interests.
 
 # Funding
+This work was funded by the German Research Foundation (Grant: WI 3342/3-1) and supported by the Max-Planck Society.
 

### 2021-03-22 manuscript/supplements_rsfmri.Rmd
changes

@@
 ---
-title: "Supplementary Material for Effects of bariatric surgery and weight loss on resting state functional connectivity"
+title: "Supplementary Information for 'Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis'"
 author:
@@
 # Methods
+
+## Sample
 Table \@ref(tab:tableSample) shows the distribution of data points in the present rsfMRI study, for the bariatric surgery group (BARS) and the waiting-list control group (NBARS).
-
-## Sample
 
@@
     linesep = "",
-    caption = "BMI and mFD of participants acquired in each condition and time point")
+    caption = "BMI and logmFD of participants acquired in each condition and time point")
 ```
@@
 ## Quality Control
-### Individual level resting-state rsfMRI QC
+### Individual level rsfMRI QC
 ```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
@@
 Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's $r$ = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's $r$ = `r round(res[4,5],2)`)
-In addition, for models showing a significant time-by-group interaction, we planned to perform a sensitivity analysis excluding the 10% data sets with Fhighest mFD.
+In addition, for models showing a significant time-by-group interaction, we planned to perform a sensitivity analysis excluding the 10% data sets with highest mFD.
 
@@
 ### Reward network 
-#### Two time points, without adjustments (Model CA1)
+**Two time points, without adjustments (Model CA1) **
 
@@
 
-#### Two timepoints, adjusted for age, sex and average mFD (Model CA2)
-
-```{r calc rew two tp adjusted 2, fig.cap="Two timepoints, adjusted for age, sex and average mFD (Model CA2)", eval = TRUE, echo = FALSE}
+**Two timepoints, adjusted for age, sex and average logmFD (Model CA2)**
+
+```{r calc rew two tp adjusted 2, fig.cap="Two timepoints, adjusted for age, sex and average logmFD (Model CA2)", eval = TRUE, echo = FALSE}
 R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + mean.logmFD + (1|subj.ID), data=final_FC[final_FC$tp!="fu2",])
@@
 
-```{r Rew2tpadj2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
+```{r Rew2tpadj2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of logmFD", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
     "Sex",
-    "av mFD",
+    "av logmFD",
     "Timepoint x Group"
@@
 
-#### Two timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+**Two timepoints, adjusted for age, sex, average of logmFD and baseline BMI (Model CA3)**
 
@@
 
-```{r Rew2tpadj3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+```{r Rew2tpadj3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of logmFD and baseline BMI", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
     "Sex",
-    "av mFD",
+    "av logmFD",
     "baseline BMI",
@@
 
-#### Three time points, without adjustments (Model CA1)
+**Three time points, without adjustments (Model CA1)**
 
@@
 
-#### Three timepoints, adjusted for age, sex and average mFD (Model CA2)
+**Three timepoints, adjusted for age, sex and average logmFD (Model CA2)**
 
@@
 
-```{r Rew3tpadj 2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
+```{r Rew3tpadj 2 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex and average of logmFD", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
     "Sex",
-    "av mFD",
+    "av logmFD",
     "Timepoint FU x Group", 
@@
 
-#### Three timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+**Three timepoints, adjusted for age, sex, average of logmFD and baseline BMI (Model CA3)**
 
@@
 
-```{r Rew3tpadj 3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+```{r Rew3tpadj 3 results, caption="Interaction of group and time on reward network connectivity for two time points, adjusted for age, sex, average of logmFD and baseline BMI", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
 ### DMN 
-#### Two time points, without adjustments (Model CA1)
+**Two time points, without adjustments (Model CA1)**
 
@@
 
-#### Two timepoints, adjusted for age, sex and average mFD (Model CA2)
+**Two timepoints, adjusted for age, sex and average logmFD (Model CA2)**
 
@@
 
-```{r DMN2tpadj2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
+```{r DMN2tpadj2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of logmFD", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r DMN2tpadj3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+```{r DMN2tpadj3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of logmFD and baseline BMI", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
 
-```{r DMN3tpadj 2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of mFD", results="asis", echo=FALSE}
+```{r DMN3tpadj 2 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex and average of logmFD", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
 
-#### Three timepoints, adjusted for age, sex, average of mFD and baseline BMI (Model CA3)
+**Three timepoints, adjusted for age, sex, average of logmFD and baseline BMI (Model CA3)**
 ```{r calc DMN three tp adjusted 3, eval = TRUE, echo = FALSE}
@@
 
-```{r DMN3tpadj 3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of mFD and baseline BMI", results="asis", echo=FALSE}
+```{r DMN3tpadj 3 results, caption="Interaction of group and time on DMN connectivity for two time points, adjusted for age, sex, average of logmFD and baseline BMI", results="asis", echo=FALSE}
 #tab_model(R1,R0)
@@
 ### Reward network
-#### Model 2.1 and 2.2 (adjusting for mFD)
+
+**Model EA1 (without adjustment) and EA2 (adjusting for logmFD)**
 ```{r "reward model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 ```
-#### Model EA 2.3 (adjusting for average and change in mFD)
+
+**Model EA 3 (adjusting for average and change in logmFD)**
 ```{r "reward model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
@@
 #tab_model(R1,R0)
-texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average mFD", "change mFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h!",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = TRUE)
+texreg(list(R1), custom.coef.names=c("Intercept", "average BMI", "change BMI", "average logmFD", "change logmFD", "Age", "Sex"), custom.model.names = "R1", float.pos = "h!",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = TRUE)
 ```
@@
 ### DMN
-#### Model 2.1 and 2.2 (adjusting for mFD)
+
+**Model EA1 (without adjustment) and EA2 (adjusting for logmFD)**
 ```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
@@
 ```
-#### Model 2.3 (adjusting for average and change in mFD)
+
+**Model EA3 (adjusting for average and change in logmFD)**
 ```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
@@
 ```
+\newpage
+### References

### 2021-03-22 code/run_or_display.m

@@ -103,14 +103,14 @@ param.MASK_B = "MNI_resampled_brain_mask.nii,1";
 roi_prep = convertCharsToStrings(readcell(fullfile(param.INFO_DIR,"ROIs.txt"), "Delimiter"," ","Whitespace","'"));
 %% ------------------------------------------------------------------------
 
-param.PRESET = "manual";
+param.PRESET = "test";
 param.ONLY_DISPLAY = true;
 param.OVERWRITE = false; 
 param.WILD_BOOT = true;   
 param.parallel = false;
 
 if param.PRESET == "standard"
-    param.MODEL = ["grouptime","grouptime2tp", "bmi", "bmiIG","bmi2tp", "fd","fdIG"];
+    param.MODEL = ["grouptime", "grouptime2tp", "bmi", "fd"];
     param.ROI_PREP = roi_prep([4, 6, 12, 14]); 
     param.COVARIATES = [11, 12, 13, 14, 21, 22, 31, 32];  
     param.MASK = "brain";           
@@ -166,9 +166,9 @@ elseif param.PRESET == "full"
     param.INFERENCE_TYPE = ["voxel","cluster","tfce"];
     param.VIEW = true;
 elseif param.PRESET == "test"
-    param.MODEL = ["grouptime"];
-    param.ROI_PREP = roi_prep([4, 6, 12, 14]); 
-    param.COVARIATES = [14];  
+    param.MODEL = ["grouptime2tp"];
+    param.ROI_PREP = roi_prep([4]); 
+    param.COVARIATES = [12];  
     param.MASK = "brain";           
     param.EXCLFD = false;        
     param.INFERENCE_TYPE = ["cluster"];

### 2021-03-22 code/tables.R

@@ -269,7 +269,7 @@ ModelList[[3]]$Hem <- LabelList[[3]]$Hem
 fn1 <- "Hem, hemisphere; L, left; R, right; FWE-corr., family-wise error corrected; MNI (Montreal Neurological Institute) coordinates of primary peak location: X, sagittal; Y, coronal; Z, axial; CC, preprocessing with AROMA+CC; GSR, preprocessing with AROMA+CC+GSR."
 fn3 <- "To identify significant clusters, we applied a cluster size threshold with p < 0.001 determined by Wild Bootstrap of 1000 samples."
 fn4 <- "Connectivity with maximum three voxels that mark local maxima within the respective custer; more detailed description of anatomical regions that are assigned to overall clusters and and corresponding probability in Supplementary."
-title_vec <- c("Models EA 2.1 and EA 2.2", "Model EA 2.3", "Model EA 2.4") 
+title_vec <- c("Models EA1 and EA2", "Model EA3", "Model EA6") 
 
 tabnames <- c("tableBMImodel","tableBMIFDmodel","tableFDmodel")
 TableList <- list()
@@ -304,14 +304,14 @@ TableList[[1]] <- TableList[[1]] %>%
 
 TableList[[2]] <- TableList[[2]] %>%
   kableExtra::group_rows("change in BMI (decrease)", 1, nrow(df3)) %>%
-  kableExtra::group_rows("average log mean FD (increase)", nrow(df3)+1, nrow(df3)+nrow(df4)) %>%
+  kableExtra::group_rows("average logmFD (increase)", nrow(df3)+1, nrow(df3)+nrow(df4)) %>%
   kableExtra::group_rows("change in BMI (decrease)", nrow(df3)+nrow(df4)+1, nrow(df3)+nrow(df4)+nrow(df5)) %>%
-  kableExtra::group_rows("average log mean FD (increase)", nrow(df3)+nrow(df4)+nrow(df5)+1, nrow(df3)+nrow(df4)+nrow(df5)+nrow(df6)) %>%
+  kableExtra::group_rows("average logmFD (increase)", nrow(df3)+nrow(df4)+nrow(df5)+1, nrow(df3)+nrow(df4)+nrow(df5)+nrow(df6)) %>%
[…]

### 2021-03-22 code/data_load.Rmd

@@
 Messueberblick_path <- "/data/p_02161/ADI_studie/metadata/ADI_MRT_Messueberblick_2019_04_10.csv"
-
 # ADI_Aktuell  -  main table, BMI info is refilled with Messueberblick entries
 ADIAktuell_path <- "/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav"
-
 # QOL_sheet  -  correct group information (unused, just for comparison)
 QOLsheet_path <- "/data/pt_02161/Results/Project1_headmotion/QOL_sheet_of_ADI_Messueberblick.csv"
-
-# summary_with_SPSS  - table has errors (contradictions for ADI080 and ADI036 with other table, but contains missing subjects)
+# summary_with_SPSS  - table has errors (contradictions for ADI080 and ADI036 with other table, but contains missing subjects group allocation)
 GroupCodingWitErrors_path <- "/data/pt_02161/Results/Project1_headmotion/summary_with_SPSS_data_sheet_of_ADI_Messueberblick.csv"
-
 # rs QA infos
 rsQA_path <- "/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv"
-
-# FS infos
+# Freesurfer infos
 FSqa_path <- "/data/p_02161/ADI_studie/BIDS/derivatives/freesurfer/QA/freesurfer_qa_qoala_11.7.csv"
-
 # subjects_rs  -  Loading mean/max FD values
 rsID_path <- "/data/pt_02161/Results/Project1_headmotion/subjects_rs.csv"
-FDvals_path <- "/data/pt_02161/Results/Project1_headmotion/summary_FDvals_correctADI061.csv"
-
+FDvals_path <- FDvals_path
 # output
@@
 bmi = read.csv(
-  "/data/p_02161/ADI_studie/metadata/ADI_MRT_Messueberblick_2019_04_10.csv",
+  Messueberblick_path,
   na.strings = c("", "NA"),
@@
 # select variables: only BMI
-bmi_repl = haven::read_sav("/data/p_02161/ADI_studie/metadata/Daten ADI aktuell.sav")[, c(
+bmi_repl = haven::read_sav(ADIAktuell_üath)[, c(
   "ADI_Code",
@@
 QOL_sheet = read.csv(
-  "/data/pt_02161/Results/Project1_headmotion/QOL_sheet_of_ADI_Messueberblick.csv"
+  QOLsheet_path
 )
@@
 group_coding_with_errors = read.csv(
-  "/data/pt_02161/Results/Project1_headmotion/summary_with_SPSS_data_sheet_of_ADI_Messueberblick.csv"
+  GroupCodingWitErrors_path
 )
@@
 ```{r "Loading rs QA infos from /data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv", echo=FALSE}
-rs_QA=read.csv("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv")
+rs_QA=read.csv(rsQA_path)
 for (i in 1:nrow(rs_QA)){
@@
 ```{r "Loading FS infos from /data/p_02161/ADI_studie/BIDS/derivatives/freesurfer/QA/freesurfer_qa_qoala_11.7.csv", echo=FALSE}
-FS_QA=read.csv("/data/p_02161/ADI_studie/BIDS/derivatives/freesurfer/QA/freesurfer_qa_qoala_11.7.csv")
+FS_QA=read.csv(FSqa_path)
 for (i in 1:nrow(FS_QA)){
@@
 ```{r "Loading mean/max FD values from /data/pt_02161/Analysis/Preprocessing/qa/rs_qa/ADI_rs_qa.csv", echo=FALSE}
-rs_ID = read.csv("/data/pt_02161/Results/Project1_headmotion/subjects_rs.csv")
+rs_ID = read.csv(rsID_path)
 colnames(rs_ID)[1] = "subj"
 res = read.csv(
-  "/data/pt_02161/Results/Project1_headmotion/summary_FDvals_correctADI061.csv",
+  FDvals_path,
   header = FALSE
@@
 ```{r, include=FALSE}
-write.csv(final_sample, "/data/p_02161/ADI_studie/metadata/final_sample_MRI_QA_info.csv")
+write.csv(final_sample, out_path)
 ```

### 2021-03-22 manuscript/draft_rsfmri.Rmd

@@
 ---
-title: "Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis"
+title: "Effects of bariatric surgery on resting-state functional connectivity of the reward and default mode network: a pre-registered analysis"
 author:
@@
 
-Obesity imposes serious health risks and involves alterations in resting-state functional connectivity  of brain networks involved in eating behavior.
-Bariatric surgery is an effective treatment, but its effects on functional connectivity are still under debate.
-In this pre-registered study, we investigated the effects of bariatric surgery on major resting-state brain networks (reward and default mode network) in a longitudinal controlled design.
+Obesity imposes serious health risks and relates to differences in resting-state functional connectivity (FC) of brain networks involved in eating behavior.
+Bariatric surgery is an effective treatment, but its effects on FC are still under debate.
+In this pre-registered analysis, we aimed to determine the effects of bariatric surgery on FC in the reward (RN) and default mode network (DMN).
 33 bariatric surgery patients and 15 obese waiting-list control patients (37 females; aged 44.15 ± 11.86 SD years (range 21-68)) underwent magnetic resonance imaging at baseline, after 6 and 12 months.
-We conducted a pre-registered whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network connectivity (https://osf.io/f8tpn/, https://osf.io/59bh7/).
-In exploratory analyses, we investigated the effects of weight loss and head motion.
-Bariatric surgery compared to waiting did not significantly affect functional connectivity (FWE-corrected $p > .05$), neither whole-brain nor within-network.
-In exploratory analyses, surgery-related BMI decrease (FWE-corrected $p = .041$) and higher average head motion (FWE-corrected $p = .021$) resulted in significantly stronger connectivity of the reward network with medial posterior frontal regions.
-This pre-registered well-controlled study did not support a strong effect of bariatric surgery, compared to waiting, on major resting-state brain networks after 6 months.
-Exploratory analyses indicated that head motion might have confounded the effects.
-Data pooling and more rigorous control of within-scanner head motion during data acquisition are needed to substantiate effects of bariatric surgery on brain organization.
+We pre-registered (https://osf.io/f8tpn/https://osf.io/59bh7/) a whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network FC.
+In exploratory analyses, we investigated all three time points and assessed the effects of weight loss and head motion.
+Bariatric surgery compared to waiting did not significantly affect FC of the RN and DMN (FWE p > .05).
+There also was no interaction effect on within-network FC.
+In exploratory analyses, surgery-related BMI decrease (FWE p= .041) and higher average head motion (FWE p= .021) resulted in significantly stronger FC in medial posterior frontal regions within the RN.
+This pre-registered analysis of longitudinal neuroimaging data did not support a strong effect of bariatric surgery, compared to waiting, on RN and DMN FC after 6 months.
+Exploratory analyses indicated that differences in head motion might have confounded the effects, thus a more rigorous control of within-scanner head motion during data acquisition is needed to better understand the effects of bariatric surgery on brain organization.
 
@@
 Here, bariatric surgery is a viable option to rapidly induce weight loss and improve glycemic status.
-RYBG is the most invasive, yet most effective surgical procedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].\
+RYBG is the most invasive, yet most effective surgical procedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].
 Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla_2017; @Brutman_2019].
@@
 brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009].
-The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.\
-The RN comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
+The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.
+The RN (RN) comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
 These brain regions have been suggested to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
@@
 Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018].\
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018].
 The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle_2015].
@@
 salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
-After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].\
+After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].
 In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
-Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].\
+Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].
 Further, while higher BMI has been consistently associated with more head motion during rsfMRI [@Beyer_2017;@Hodgson_2016], previous studies in bariatric surgery patients have not taken this important confounder of FC into account.
 In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
-Thus, we aimed to rigorously control for motion-related variance in our analyses.\
+Thus, we aimed to rigorously control for motion-related variance in our analyses.
 We had the following confirmatory hypotheses:
@@
 For more details see [@Prehn_2020].
-We used all data acquired until April 2019.\
+We used all data acquired until April 2019.
 The study protocol was in accordance with the Helsinki Declaration and approved by the ethics committee of the Charité University Medicine Berlin.
 The study was registered at clinicaltrials.gov as NCT01554228.
-We pre-registered the present resting-state analyses on the OSF (https://osf.io/yp42s).
-We made additional changes (see https://osf.io/59bh7/) to the pre-registration after preprocessing the rsfMRI data as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
-For a comparison of the pre-registration and the manuscript, please visit https://osf.io/45n9f/.\
+We pre-registered the present resting-state analyses on the OSF https://osf.io/yp42s.
+We made additional changes https://osf.io/59bh7/ to the pre-registration after denoising as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
+<!-- RECRUITEMENT, INCLUSION/EXCLUSION CRITERIA -->
+
 Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin.
@@
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).\
+Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).
  In total, $51$ participants out of the originally enrolled $69$ subjects received MRI.
-Five data points of three subjects had to be excluded due to bad anatomical image quality (see below), which led to a final data set with $101$ rsfMRI sessions.\
+Five data points of three subjects had to be excluded due to bad anatomical image quality (see below), which led to a final data set with $101$ rsfMRI sessions.
+<!-- DESCRIPTIVE STATISTICS -->
 The final sample entailed `r nrow(subj_df)` morbidly obese individuals (`r nrow(filter(subj_df,Sex == "f"))` females; aged `r round(mean(subj_df$Age_BL),2)` $\pm$ `r round(sd(subj_df$Age_BL),2)` SD years (range `r min(subj_df$Age_BL)`-`r max(subj_df$Age_BL)`)).
-`r round(sum(subj_df$arterial_hypotension == 1, na.rm = T) / nrow(subj_df), 3)*100`% of participants had clinically diagnosed hypertension, `r round((sum(subj_df$diabetes_type_II == 1, na.rm = T) + sum(subj_df$diabetes_type_I == 1, na.rm = T))  / nrow(subj_df), 3)*100`% had type-2 or type-1 diabetes, and `r round(sum(subj_df$smoking_status == 1, na.rm = T) / nrow(subj_df), 3)*100`% reported to smoke.\
-Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` females) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` females), who waited for their health insurance's approval to undergo surgery.
+`r round(sum(subj_df$arterial_hypotension == 1, na.rm = T) / nrow(subj_df), 3)*100`% of participants had clinically diagnosed hypertension, `r round((sum(subj_df$diabetes_type_II == 1, na.rm = T) + sum(subj_df$diabetes_type_I == 1, na.rm = T))  / nrow(subj_df), 3)*100`% had type-2 or type-1 diabetes, and `r round(sum(subj_df$smoking_status == 1, na.rm = T) / nrow(subj_df), 3)*100`% reported to smoke.
+Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` female) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` female), who waited for their health insurance's approval to undergo surgery.
 Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016].
 Analyses were performed on all participants who provided at least one data point of rsfMRI data.
-`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point (for more details see Figure 1 in supplementary information (SI)).
-The pre-registered analysis of changes from baseline to followup included $24$ participants with both time points, in total $72$ data points.\
+`r sum(freqtable$Freq == 3)` participants had complete data, `r sum(freqtable$Freq == 2)` provided data for two time points and `r sum(freqtable$Freq == 1)` for one time point (for more details see Table S1).
+The pre-registered analysis of changes from baseline to followup included $24$ participants with both time points, in total $72$ data points.
+
 `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the intervention group information was not available.
-\
+
 Participants arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast.
 They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
-MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020).
+MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020)).
 
@@
 ### Minimal preprocessing
-Imaging data analysis was conducted using AFNI 19.1.05, ANTS 2.3.1, FSL 6.0.1 and FreeSurfer 6.0.0p1, wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15 which can be found on https://github.com/fBeyer89/ADI_preproc/.
+Imaging data was conducted using AFNI 19.1.05, ANTS '2.3.1', FSL '6.0.1' and FreeSurfer '6.0.0p1', wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15.
+The preprocessing workflow included modules to process anatomical, functional and diffusion-weighted imaging in parallel. The workflow can be found on https://github.com/fBeyer89/ADI_preproc/.
 T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross].
@@
 The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants].
+The following parameters were used for rigid [metric: mutual information, number of iterations: [1000, 500, 250, 100]], affine [metric: mutual information, number of iterations: [1000, 500, 250, 100]] and SyN: [metric: cross-correlation, iterations: [100, 70, 50, 20]] registration steps with shrink factors [8, 4, 2, 1], and smoothing sigmas=[3, 2, 1, 0] for each step.
 Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
@@
 ### Denoising pipelines
-In the pre-registration, we specified two denoising pipelines (AROMA+CC including ICA-AROMA + CompCor and AROMA+CC+GSR including ICA-AROMA + CompCor + global signal regression (GSR)) [@Parkes_2018; @Ciric_2017, for details see SI].
+We specified two denoising pipelines (ICA-AROMA + CompCor and ICA-AROMA + CompCor + global signal regression (GSR) in the pre-registration.
+AROMA+CC and AROMA+CC+GSR [@Parkes_2018; @Ciric_2017, for details see SI].
 
@@
 FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
-We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.\
+We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.
 <!--ADI009_fu2, ADI063_bl, ADI063_fuADI116_bl, ADI116_fu, ADI116_fu2-->
+
 **RsfMRI quality control**
 RsfMRI quality control was performed according to the protocol by @Ciric_2018 (see SI for more details). 
-Head motion was quantified using mean framewise displacement (mFD) according to @Power_2012, and log-transformed for further analysis (logmFD).
+We calculated mean framewise displacement (mFD) according to @Power_2012, and log-transformed it for further analysis (logmFD).
 As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
@@
 Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+CC+GSR).
+We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+GSR).
 Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
 
-## Aggregated within-network FC 
-To extract aggregated within-network FC, we first calculated the mean DMN and RN over all participants and time points, adjusted for age and sex.
-We used GSR-denoised data as input and clusterwise bootstrapping with $N = 1000$.
+## Aggregated FC values
+To extract aggregated FC values, we first calculated the mean DMN and RN over all participants and NAccs, adjusted for age and sex.
+We used GSR regressed data as input and clusterwise bootstrapping with $N = 1000$.
 Network masks were formed from all voxels within clusters which survived a clusterwise multiple comparison correction of FWE-corrected $p < 0.05$.
-We extracted the average GSR-denoised FC from these masks.
+We extracted the average FC from these masks for GSR-data.
 
@@
 The marginal model implemented in the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term.
-We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size.
+It therefore accommodates any repeated measurement covariance structures.
+We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
 We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability) for all analyses.
-Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].
+Statistical analyses on the aggregated FC and imputation were performed in R version 3.6.1 [@team2013r].
 
@@
 ### Confirmatory analyses
-**Whole-brain analysis**
-We tested the pre-registered hypothesis of a time-by-group interaction:  `FC=group + time + group * time`.\
-As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both timepoints (model CA2) and age, sex, average logmFD and baseline BMI (model CA3).
-Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values.
-This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.\
-**Aggregated FC analysis**
+#### Whole-brain analysis
+We tested the pre-registered hypothesis of a time-by-group interaction:  `FC=group + time + group * time`. 
+As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (CA1), age, sex and average logmFD of both timepoints (CA2) and age, sex, average logmFD and baseline BMI (CA3).
+Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package mice based on subject identification, condition, baseline age, and available BMI values.
+This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.
+
+#### Aggregated FC analysis
 We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007].
@@
 First, we investigated the time-by-group interaction for baseline and followup time points only.
-We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3).
+We either performed no correction (CA1), adjusted for baseline age, sex, average of logmFD of both time points (CA2) or additionally adjusted for baseline BMI (CA3).
 We performed model comparison between R1 and R0 models, where $R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))$, and $R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))$.
@@
 ### Exploratory analysis (pre-registered)
-**Whole-brain analysis**
+#### Whole-brain analysis
 As described in the pre-registration, we performed an exploratory analysis of the relationship of average and change in BMI in a separate group of models.
@@
 This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC.
-We first estimated their effects in a model adjusting for baseline age and sex (model EA1) and then additionally controlling for logmFD (model EA2).
-As we previously reported correlated change in BMI and head motion in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (model EA3).
-Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.\
-**Aggregated FC analysis**
+We first estimated their effects in a model adjusting for baseline age and sex (Model EA 1) and then additionally controlling for logmFD (Model EA 2).
+As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (Model EA 3).
+Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
+
+#### Aggregated FC analysis
 We also performed an exploratory analysis with average BMI and change in BMI as predictors of the aggregated FC from AROMA+CC+GSR denoised data of both networks.
-We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (model EA1), logmFD (model EA2) and average and change in logmFD (model EA3).
+We calculated three models with average and change in BMI as predictors of interest and adjusted for baseline age and sex (Model EA 1), logmFD (Model EA 2) and average and change in logmFD (Model EA 3).
 
 ### Exploratory analysis (not pre-registered)
-**Whole-brain analysis**
-**Group-by-time interaction in three time points**
+#### Whole-brain analysis
+##### Group-by-time interaction in three time points
 We explored our whole-brain group-by-time point interaction models for the data of all three time points.
 In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
-The resulting factorial design contained one regressor for each time point per group (see SI for depiction of design matrix).
+The resulting factorial design contained one regressor for each time point per group (see supplements for depiction of design matrix).
 This analysis had not been pre-registered.
 Here, we used individual logmFD values (not averaged) as covariate to capture variance in logmFD change over time points.
-We investigated two models adjusting for age and sex (model EA4) and age, sex and logmFD (model EA5).\
-**Average and change in logmFD**
-Further, we tested the association of head motion and FC in another exploratory model (model EA6): `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates.
+We investigated two models (EA 4: group x time, age, sex and EA 5: group x time, age, sex and logmFD).
+##### Average and change in logmFD
+Further, we tested the association of head motion and FC in another exploratory model (EA 6): `FC = between-subject logmFD + within-subject logmFD` with age and sex as nuisance covariates.
 With this model, we aimed for a better understanding of the unique contribution of average and longitudinal change in logmFD measures.
@@
 
-**Whole-brain analysis**
+#### Whole-brain analysis
 <!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
-**Aggregated FC analysis**
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.
+
+#### Aggregated FC analysis
 The interaction effect of group and time point in the models of aggregated FC was considered significant if the model comparison between R1 and R0 models using the anova command showed $p < 0.05$.
@@
 ### Functional decoding
-In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the NeuroSynth (https://www.neurosynth.org/) database [@Yarkoni_2011].
-We uploaded the contrast images for change BMI and average logmFD on Neurovault (https://neurovault.org/collections/9426/), and applied the decoding classifier.
+In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011].
+We uploaded the contrast images for change BMI and average logmFD on [Neurovault](), and applied the decoding classifier.
 This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps.
@@
 In the intervention group, baseline MRI data is not missing completely at random, due to the limited scanner bore.
+Similarly, we observed a higher drop-out in the control than in the surgery group (baseline timepoint only, n control = ; n surgery = ).
 Change in BMI throughout the study is depicted in Figure \@ref(fig:figBMIdescr).
@@
 The same was true for models CA2 (adjusting for age, sex and average of logmFD). Results did not differ between AROMA+CC and AROMA+CC+GSR denoising pipelines.
-In model CA3 (adjusting for age, sex, average of logmFD and baseline BMI) there also was no significant interaction when adding baseline BMI.
+In Model CA3 (adjusting for age, sex, average of logmFD and baseline BMI) there also was no significant interaction when adding baseline BMI.
 Yet, we found a significant main effect of time in this model.
@@
 ```
-There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure \@ref(fig:figaggFC) and SI for a detailed summary of the models), regardless of the adjustments (CA1 without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
+There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure \@ref(fig:figaggFC) and SI for a detailed summary of the models), regardless of the adjustments (CA1: without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
 
@@
 ### Whole-brain analysis
-RN FC was not significantly related to neither average nor change in BMI, for either of the denoising pipelines and regardless whether we adjusted for logmFD in models EA1 and EA2.
-Only in model EA3 (adjusting for average and change in both BMI and logmFD), we found that more BMI decrease (e.g.
-weight loss) predicted higher FC between NAcc and a cluster in the posterior-medial frontal region (see Figure \@ref(fig:BMIFDagesexNacccgnBMIdeact) and Table \@ref(tab:tableBMIFDmodel)).\
-The peak voxel was classified as belonging to superior frontal gyrus ($45$% probability) and supplementary motor area (SMA) ($37$%  probability) in the Harvard-Oxford atlas.
-Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.030$), also in AROMA+CC+GSR-denoised data ($p = 0.041$).\
-```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+RN FC was not significantly related to neither average nor change in BMI, for either of the denoising pipelines and regardless whether we adjusted for logmFD (EA 1 and EA 2).
+Only in model EA 3 (adjusting for average and change in both BMI and logmFD), we found that more BMI decrease (e.g.
+weight loss) predicted higher FC between NAcc and a cluster in the posterior-medial frontal region (see Table \@ref(tab:tableBMImodel)).
+The peak voxel derived from investigating the effect of average logmFD was classified $45$% superior frontal gyrus and $37$% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)).
+Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also in AROMA+CC+GSR-denoised data ($p = 0.038$).
+
+```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 ```
-```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
@@
 ```
-Moreover, average logmFD was positively associated with FC between NAcc and motor cortex in model EA3 for both denoising pipelines (see Figure \@ref(fig:BMIFDagesexNaccavgFDact) and Table \@ref(tab:tableBMIFDmodel)).\
+Moreover, average logmFD was positively associated with FC between NAcc and motor cortex in model EA 3 for both denoising pipelines (see Figure \@ref(fig:BMIFDagesexNaccavgFDact)).
+
 For the DMN, we found that higher average BMI predicted lower FC of the precuneus/PCC with the lingual gyrus, mid orbital gyrus and temporal gyrus in the images denoised with AROMA+CC.
-This finding was significant in models EA1, EA2 and EA3 (see Figures \@ref(fig:BMIagesexPCCccavgBMIdeact) and \@ref(fig:BMIFDagesexPCCccavgBMIdeactt) and Table \@ref(tab:tableBMImodel)).
-Yet, none of the clusters survived when using AROMA+CC+GSR denoised data (see Table \@ref(tab:tableBMIFDmodel).\
-```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (Model EA1)", "adjusted for age, sex and logmFD (EA2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
+This finding was significant in models EA 1, EA 2 and EA 3 (see Figures \@ref(fig:BMIagesexPCCccavgBMIdeact) and \@ref(fig:BMIFDagesexPCCccavgBMIdeactt) and Table \@ref(tab:tableBMImodel)).
+Yet, none of the clusters survived when using AROMA+CC+GSR denoised data (see Table \@ref(tab:tableBMIFDmodel).
+
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (EA 1)", "andjusted for age, sex and logmFD (EA 2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
 knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
 ```
-```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data (Model EA3)", out.width = "50%", fig.align="center"}
+```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data", out.width = "50%", fig.align="center"}
 knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
@@
 ### Aggregated FC analysis
-As expected, there was no association of average BMI or within-subject BMI change and within RN FC in models EA1, EA2 or EA3 adjusting for age, sex and logmFD (for details see SI).\
+As expected, there was no association of average BMI or within-subject BMI change and within RN FC in models EA 1, EA 2 or EA 3 adjusting for age, sex and logmFD (for details see SI).
 ```{r "Rew model average&change of BMI&logmFD", eval = TRUE, echo = FALSE}
@@
 ```
+
 ```{r "DMN model, mean, change BMI", eval = TRUE, include = FALSE}
@@
 ```
+
 ```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
@@
 ```
-Like in the whole-brain analysis, higher average BMI was associated with reduced aggregated DMN FC, regardless of whether we adjusted for logmFD (model EA1: $p$ = `r round(p_meanBMI,3)` and EA2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
-The association also remained significant when we split logmFD into average and change in logmFD (EA3) ($p$ = `r round(p_mean_BMI_splitmFD,3)[2]`) and there was no significant association of average or change in logmFD with DMN FC (see SI for overview of all models).
+Like in the whole-brain analysis, higher average BMI was associated with reduced aggregated DMN FC, regardless of whether we adjusted for logmFD (model EA 1: $p$ = `r round(p_meanBMI,3)` and EA 2: $p$ = `r round(p_meanBMI_adjmFD,3)`).
+The association also remained significant when we split logmFD into average and change in logmFD (EA 3) ($p$ = `r round(p_mean_BMI_splitmFD,3)[2]`) and there was no significant association of average or change in logmFD with DMN FC (see SI for overview of all models).
 
@@
 ### Group-by-time interaction in three time points 
-Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA4 and EA5 in the full data set for neither RN nor DMN.
+Similarly to the analysis with two time points, there was no significant interaction or main effect when analyzing models EA 4 and EA 5 in the full data set for neither RN nor DMN.
 
 ### Average and change in logmFD 
-<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), higher average logmFD was associated with stronger FC between the NAcc and a cluster located in proximity to the central sulcus and motor areas.
-This Cluster only differed in size between denoising pipelines.\
-We did not find any clusters with a significant association of logmFD and DMN FC.
-```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with a cluster in motor cortex, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
+<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), higher average logmFD was positively associated with FC between NAcc and cluster located in proximity to the central sulcus and motor areas.
+Cluster co-activation appeared irrespective of the denoising pipeline, differing mainly in cluster size.
+<!-- ### DMN -->We did not find any clusters with a significant association of logmFD and DMN FC.
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of Nacc with cluster, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
@@
 ```
-### Functional decoding
-Functional decoding of the two activation patterns from the pre-registered, exploratory analysis EA3 showed different top association terms.
-For the [https://neurosynth.org/decode/?neurovault=441783](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula and inferior frontal were the top terms, while for the [https://neurosynth.org/decode/?neurovault=441784](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract.
+##### Functional decoding
+Functional decoding of the two activation patterns from the pre-registered, exploratory analysis EA 3 showed different top association terms.
+For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula and inferior frontal were the top terms, while for the [average logmFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract.
 The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
@@
 In this pre-registered study, we investigated the effects of bariatric surgery on the FC of major resting-state brain networks in a longitudinal controlled design.
-Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising pipelines and controlling for head motion on the group level.\
+Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising pipelines and controlling for head motion on the group level.
 We did not detect significant effects of bariatric surgery compared to waiting on whole-brain FC of the PCC and NAcc, core hubs of the RN and DMN, according to pre-registered whole-brain analyses.
@@
 Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns.
-Finally, higher average head motion was associated with increased NAcc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the change BMI cluster.\
+Finally, higher average head motion was associated with increased NAcc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the change BMI cluster.
+
 In this study, we could not confirm our pre-registered hypotheses.
@@
 These studies, notably, did not include adequate longitudinal obese control groups, making false-positive findings possible.
-We thus conclude that surgery-induced heavy weight loss does not strongly affect DMN and RN FC based on the current results.\
+We thus conclude that surgery-induced heavy weight loss does not strongly affect DMN and RN FC based on the current results.
 However, in an exploratory analysis, stronger BMI decrease predicted higher connectivity of the NAcc and a cluster in a posterior-medial frontal brain region.
@@
 a shift from integration towards segregation and more specialized processing.
-However, this interpretation is highly speculative, and more studies are needed to corroborate these considerations.\
+However, this interpretation is highly speculative, and more studies are needed to corroborate these considerations.
+
 Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity, challenging definite conclusions on the relationship between weight loss and FC changes.
-Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.\
+Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.
 To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFD and thereby introducing two instead of one regressor into the model.
@@
 Further, average FD was associated with a cluster in a similar, yet not identical region, and crucially, this association was positive.
-Thus, confounding of the negative BMI change effect and the positive average head motion effect on posterior-frontal FC seems unlikely.\
+Thus, confounding of the negative BMI change effect and the positive average head motion effect on posterior-frontal FC  seems unlikely.
 Head motion also played a role in the association of higher BMI and reduced DMN FC.
@@
 Thereby, this results echoes a previous finding from our group, and may be interpreted as accelerated age-related decline of the DMN in relation to the cardiometabolic risk related to BMI.
-Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia_2017].\
+Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia_2017].
+
 The major strength of our study is the prospective intervention controlled design.
-We compared bariatric surgery patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history or recommendation and were scanned after the same time intervals [@Thiese_2014].
+We compared BS patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history and  treatment recommendation and were scanned after the same time intervals [@Thiese_2014].
 Another strength of our study was the pre-registered analysis plan, which was corrected prior to statistical analysis after we noticed flaws in the first version.
-In particular, we gave more details on denoising pipelins and models and determined that we would use SwE toolbox, an advanced statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
-Opposed to the flexible factorial models which is the standard in SPM, marginal models use less degrees of freedom, and thus allow for the inclusion of covariates and higher power.\
+In particular, we fixed the number of denoising schemes and models and determined that we would use SwE toolbox, an advanced statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
+Opposed to the flexible factorial models which is the standard in SPM, marginal models are parsimonious and preserve the degrees of freedom, allowing for the inclusion of covariates and higher power.
 Limitations of our study include the low number of patients who participated in all three time points.
@@
 Patients in the intervention group were not missing at random over time points, as often, before surgery, they did not fit into the MRI scanner.
-While this sample size is comparable to previous rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.\
+While this sample size is comparable to previous rsfMRI studies in bariatric surgery, it seems unlikely that our power was high enough to detect small effect sizes.
+
 We used seed-based connectivity to derive large-scale brain networks.
 While this approach yielded reasonable DMN and RN maps, it is a univariate approach not taking into account the interrelatedness of subnetworks and assuming that the connectivity of a central hub reflects the connectivity of the network as a whole.
-Furthermore, our rsfMRI was relatively short, which might have further reduced our power.\
+Furthermore, our rsfMRI was relatively short, which might have further reduced our power.
+
 We did not monitor hunger or satiety in our design, although all participants were scanned after the intake of a breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].\
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].
 Size and composition of our sample did not allow sex-stratified analyses.
-However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].\
+However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].
+
 Taken together, this prospective well-controlled study did not confirm previous findings claiming strong effects of bariatric surgery on FC of the RN and DMN in obese patients.
@@
 # Code availability
-All code is available on github.
-Preprocessing scripts are here [https://github.com/fBeyer89/ADI_preproc](https://github.com/fBeyer89/ADI_preproc), and analyses scripts are here: [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
+
+All code is available in the github repository [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
+The code includes script for data preparation and analysis of different models.
 
 # Declaration of interest
-The authors declare no competing interests.
 
 # Funding
-This work was funded by the German Research Foundation (Grant: WI 3342/3-1) and supported by the Max-Planck Society.
 

### 2021-03-22 manuscript/draft_rsfmri.Rmd

--- 2021-03-22 manuscript/draft_rsfmri.Rmd
@@
 In exploratory analyses, we investigated the effects of weight loss and head motion.
-Bariatric surgery compared to waiting did not significantly affect functional connectivity (FWE-corrected $p > .05$), neither whole-brain nor within-network.
-In exploratory analyses, surgery-related BMI decrease (FWE-corrected $p = .041$) and higher average head motion (FWE-corrected $p = .021$) resulted in significantly stronger connectivity of the reward network with medial posterior frontal regions.
+Bariatric surgery compared to waiting did not significantly affect functional connectivity (FWE-corrected $p > 0.05$), neither whole-brain nor within-network.
+In exploratory analyses, surgery-related BMI decrease (FWE-corrected $p = 0.041$) and higher average head motion (FWE-corrected $p = 0.021$) resulted in significantly stronger connectivity of the reward network with medial posterior frontal regions.
 This pre-registered well-controlled study did not support a strong effect of bariatric surgery, compared to waiting, on major resting-state brain networks after 6 months.
@@
 We made additional changes (see https://osf.io/59bh7/) to the pre-registration after preprocessing the rsfMRI data as we realized some aspects of the analysis were inadequately described in the initial pre-registration.
-For a comparison of the pre-registration and the manuscript, please visit https://osf.io/45n9f/.\
+For a comparison of the pre-registration and the manuscript, please visit https://osf.io/45n9f/.
+\
 Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin.
@@
 type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
-Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).\
+Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score $< 34$).\
 In total, $51$ participants out of the originally enrolled $69$ subjects received MRI.
@@
 `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the intervention group information was not available.
-\
 Participants arrived in the morning (between 07:00 and 12:00 a.m.) after an overnight fast.
@@
 MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
-<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
+<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = $1 \times 1 \times 1$ mm$^3$, 192 sagital slices).
 <!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms.
-The image matrix was 64 x 64 with an in-plane resolution of 3 mm x 3 mm and 34 slices with a slice thickness of 4 mm.
+The image matrix was $64 \times 64$ with an in-plane resolution of 3 mm $\times$ 3 mm and 34 slices with a slice thickness of 4 mm.
 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes.
@@
 Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.
-The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants]. 
+The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 $1 \times 1 \times 1$ mm template using ANTS [@avants2009ants]. 
 Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`).
@@
 We did not select vmPFC for the RN due to low SNR.
-Based on FreeSurfer's aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
+Based on FreeSurfer's aseg.mgz and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres.
 Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain.
-We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+CC+GSR).
+We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed, AROMA, AROMA+CC, AROMA+CC+GSR).
 Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.
@@
 We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size.
-We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability) for all analyses.
+We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to $3 \times 3 \times 3$ mm$^3$ and thresholded at 0.5 GM probability) for all analyses.
 Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].
@@
 **Whole-brain analysis**
-We tested the pre-registered hypothesis of a time-by-group interaction:  `FC=group + time + group * time`.\
+We tested the pre-registered hypothesis of a time-by-group interaction:  `FC = group + time + group * time`.\
 As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both timepoints (model CA2) and age, sex, average logmFD and baseline BMI (model CA3).
@@
 We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3).
-We performed model comparison between R1 and R0 models, where $R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))$, and $R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))$.
+We performed model comparison between R1 and R0 models, where `R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))`, and `R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))`.
 As specified in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.
@@
 Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual.
-Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$, and 1000 bootstraps due to required computation time (instead of 5000)).
+Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$), and 1000 bootstraps due to required computation time (instead of 5000).
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
@@
 Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.030$), also in AROMA+CC+GSR-denoised data ($p = 0.041$).\
-```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
+```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD (Model EA3).
+Legend denotes empirical Z-values.", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
@@
 ```
-```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
+```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD (Model EA3).
+Legend denotes empirical Z-values.", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
 knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
@@
 Yet, none of the clusters survived when using AROMA+CC+GSR denoised data (see Table \@ref(tab:tableBMIFDmodel).\
-```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (Model EA1)", "adjusted for age, sex and logmFD (EA2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
+```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (Model EA1).
+Legend denotes empirical Z-values.", "adjusted for age, sex and logmFD (EA2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
 knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
 ```
-```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data (Model EA3)", out.width = "50%", fig.align="center"}
+```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data (Model EA3).
+Legend denotes empirical Z-values.", out.width = "50%", fig.align="center"}
 knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
@@
 We did not find any clusters with a significant association of logmFD and DMN FC.
-```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with a cluster in motor cortex, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
+```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with a cluster in motor cortex, adjusted for age, sex, average BMI, change BMI and change logmFD.
+Legend denotes empirical Z-values.", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
 knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")

### 2021-03-22 manuscript/supplements_rsfmri.Rmd

@@
 figList <- readRDS(file = "../report/fig/figDvarsmFDList.rds")
-#(figList$fig1 | figList$fig2) / figList$fig3
-
 lay <- rbind(c(1,1,2,2),
              c(NA,3,3,NA))
-gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
+figDvarsmFD <- gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
+figDvarsmFD
+
 ```
@@
 Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) are on Neurovault.
-```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all time points and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$)", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
+```{r figNetworks, echo=FALSE, out.width="50%", out.height="30%", fig.cap="Resting state networks based on AROMA+CC+GSR input data, one-sample t-tests adjusting for age and sex over all time points and participants with bootstrapped clusterwise inference (FWE-corrected $p < 0.05$).
+Legend denotes empirical Z-values.", fig.subcap = c("DMN seeded from precuneus", "Reward network seeded from Nucleus Accumbens", "tSNR differences between regions"),fig.show='hold',fig.align='center'}
 
@@
 ## Design
-```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "50%", fig.subcap = c('Design matrix with time as categorial factor.','Design matrix with time as continuous variable.'), fig.cap="Possible design matrices of time-by-group model.", echo=FALSE}
+```{r figDesignmatrix, fig.align = 'center', fig.ncol = 2, out.width = "100%",  fig.cap="Design matrices of time-by-group model with time as categorial factor.
+N, NBARS; B, BARS; numbers indicate month of measurement since beginning, 0 indicates baseline(pre-surgery measurement, 6 and 12 indicate months after baseline/surgery.", echo=FALSE}
 DesignMatricesList <- readRDS(file = "../report/fig/DesignMatricesList.rds")
 DesignMatricesList$plotA
-DesignMatricesList$plotB
 #gridExtra::grid.arrange(plotA, plotB, nrow=1)
@@
 
-```{r "GT2tpagesexmFDBMIblNaccgsrTimeDeact", fig.cap="Main effect of time: Negative Association of FC between NAcc and this cluster in lateral parietal cortex, adjusted for age, sex, avgFD, and baseline BMI in AROMA+CC+GSR preprocessed data", fig.hold=TRUE, out.width = "50%", fig.align="center"}
+```{r "GT2tpagesexmFDBMIblNaccgsrTimeDeact", fig.cap="Main effect of time: Negative Association of FC between NAcc and this cluster in lateral parietal cortex, adjusted for age, sex, avgFD, and baseline BMI in AROMA+CC+GSR preprocessed data.
+Legend denotes empirical Z-values.", fig.hold=TRUE, out.width = "50%", fig.align="center"}
 knitr::include_graphics(c("../report/fig/GroupTime2tp_NACCgsr_time_deact.pdf"))
@@
 
-```{r "GT2tpagesexmFDBMIblPCCgsrTimeDeact", fig.cap="Main effect of time: Negative Association of FC between PCC/precuneus and this cluster in ACC, adjusted for age, sex, avgFD, and baseline BMI in AROMA+CC+GSR preprocessed data", fig.hold=TRUE, out.width = "50%", fig.align="center"}
+```{r "GT2tpagesexmFDBMIblPCCgsrTimeDeact", fig.cap="Main effect of time: Negative Association of FC between PCC/precuneus and this cluster in ACC, adjusted for age, sex, avgFD, and baseline BMI in AROMA+CC+GSR preprocessed data.
+Legend denotes empirical Z-values.", fig.hold=TRUE, out.width = "50%", fig.align="center"}
 knitr::include_graphics(c("../report/fig/GroupTime2tp_PCCgsr_time_deact.pdf"))

### 2021-03-22 manuscript/draft_rsfmri.Rmd
final bib + corrections

@@
 We tested the pre-registered hypothesis of a time-by-group interaction:  `FC = group + time + group * time`.\
-As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both timepoints (model CA2) and age, sex, average logmFD and baseline BMI (model CA3).
+As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both time points (model CA2) and age, sex, average logmFD and baseline BMI (model CA3).
 Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values.
@@
 ```
-There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure \@ref(fig:figaggFC) and SI for a detailed summary of the models), regardless of the adjustments (CA1 without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three timepoints.
+There was no significant group-by-time interaction for aggregated DMN and RN FC (see Figure \@ref(fig:figaggFC) and SI for a detailed summary of the models), regardless of the adjustments (CA1 without adjustment, CA2 adjusting for age,sex and average of mFD or CA3 age,sex, average of mFD and baseline BMI), and whether the analyses were performed in two and three time points.
 
@@
 ```
+
 ### Functional decoding
@@
 Furthermore, to increase validity and power, future studies should include obese control groups, and increase efforts to share and pool valuable patient data into meta-analysis.
-Finally, pre-registration of concrete and testable hypotheses would help to increase replicability of the field and truely enhance our understanding of the neural underpinnings of altered gut-brain communication after bariatric surgery.
+Finally, pre-registration of concrete and testable hypotheses would help to increase replicability of the field and truly enhance our understanding of the neural underpinnings of altered gut-brain communication after bariatric surgery.
 
@@
 All code is available on github.
-Preprocessing scripts are here [https://github.com/fBeyer89/ADI_preproc](https://github.com/fBeyer89/ADI_preproc), and analyses scripts are here: [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
+Preprocessing scripts can be found under [https://github.com/fBeyer89/ADI_preproc](https://github.com/fBeyer89/ADI_preproc), and analyses scripts can be found under [https://github.com/hsx1/adi2_rsfMRI](https://github.com/hsx1/adi2_rsfMRI).
 

### 2021-03-23 manuscript/draft_rsfmri.Rmd
changed reference to latex compatible

@@
 author:
-- Hannah Sophie Heinrichs* [1]
-- Frauke Beyer* [1][2]
-- Evelyn Medawar [1]
-- Kristin Prehn [4][5]
-- Jürgen Ordemann [8][9]
-- Agnes Floel [4][5]
-- A. Veronica Witte [3][1][2]
-date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
-  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Clinic of Cognitive Neurology, University Clinic Leipzig \newline 4 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 5 Department of Psychology, Medical School Hamburg, Hamburg \newline 6 Department of Neurology, University of Greifswald, Greifswald \newline 7 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin
+- Hannah Sophie Heinrichs* 
+- Frauke Beyer* 
+- Evelyn Medawar
+- Kristin Prehn
+- Jürgen Ordemann
+- Agnes Floel
+- A. Veronica Witte
+date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline 2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3 Clinic of Cognitive Neurology, University Clinic Leipzig \newline 4 Department of Neurology and NeuroCure Clinical Research Center, Charité University Medicine, Berlin \newline 5 Department of Psychology, Medical School Hamburg, Hamburg \newline 6 Department of Neurology, University of Greifswald, Greifswald \newline 7 German Center for Neurodegenerative Diseases, Standort Rostock/Greifswald, Greifswald \newline 8 Center for Bariatric and Metabolic Surgery, Charité University Medicine, Berlin \newline 9 Center for Bariatric and Metabolic Surgery, Vivantes Clinic Spandau, Berlin
+
+bookdocumentclass: article
 output:
@@
     toc: yes
+    #toc_depth: 4
+    keep_tex: yes
+    #citation_package: none
+    pandoc_args: [ "--csl","apa.csl" ]
 bibliography: "bibliography.bib"
+link-citations: true
+
 editor_options:
@@
 - \onehalfspacing
-indent: true
+#- \usepackage{bibtex}
+#indent: true
 ---
@@
 library(knitr) # version 1.30
+library(distill)
 
@@
 Obesity is a worldwide health issue, entailing huge personal and societal costs.
-Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life [@Albanese_2017].
+Excess amount of body fat not only affects cardiovascular and metabolic health, but also increases the risk for cognitive decline and dementia later in life [@Albanese2017].
 Conservative treatment options including behavioral therapy often do not yield the desired weight loss, especially in patients with very high BMI ($> 35$ kg/m$^2$).
@@
 RYBG is the most invasive, yet most effective surgical procedure, but other techniques like vertical sleeve gastrectomy (VSG) and gastric banding (GB) are also common [@Berthoud_2011].\
-Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla_2017; @Brutman_2019].
+Bariatric surgery leads to weight loss due to malabsorption of nutrients, reduced digestion efficiency and altered food perception, appetite and central regulation of food intake [@Mulla2017; @Brutman2019].
 Precise mechanisms how bariatric surgery leads to altered appetitive signaling are yet to be elucidated.
-One option to adress these questions on brain-behavior relationships is to use resting-state functional magnetic resonance imaging (rsfMRI), a technique capturing the dynamic organization of the brain.
+One option to address these questions on brain-behavior relationships is to use resting-state functional magnetic resonance imaging (rsfMRI), a technique capturing the dynamic organization of the brain.
 Functional connectivity (FC) networks, i.e.
-brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith_2009].
+brain regions with correlated neural activity over time, are in anatomical correspondence with specific brain networks involved in cognitive processes, including attention and executive control [@Smith2009].
 The reward network (RN), processing hedonic value and internal motivation, and the default mode network (DMN), a higher-order network, involved in interoception and governing shifts between external-internal processes,  are promising candidates to mediate altered central regulation of food intake after bariatric surgery.\
-The RN comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty_2004;@Liu_2011].
-These brain regions have been suggested to guide food valuation processes and decision-making in humans [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018].
-Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto_2018; @Garcia_2014; @Meng_2020; @Stoeckel_2009; though see @Morys_2020].
+The RN comprises the ventromedial prefrontal cortex (vmPFC), the nucleus accumbens (NAcc), the putamen, the amygdala and the anterior insula  [@Odoherty2004;@Liu2011].
+These brain regions have been suggested to guide food valuation processes and decision-making in humans [@Bartra2013; @Hare2011; @Hutcherson2012; @Schmidt2018].
+Frequently, obesity has been associated with hyperactivation of RN regions during anticipation of (high-caloric) food cues, and in contrast, reduced activation to actual taste of these foods [@Devoto2018; @Garcia2014; @Meng2020; @Stoeckel2009; though see @Morys2020].
 RsfMRI studies also showed increased local FC of reward network regions, i.e.
-NAcc, vmPFC, putamen, insula [@Contreras_2017; @Coveleskie_2015;@Hogenkamp_2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips_2014;  @Wijngaarden_2015].
+NAcc, vmPFC, putamen, insula [@Contreras2017; @Coveleskie2015;@Hogenkamp2016], and altered connectivity with salience, homeostatic and sensorimotor networks [@Lips2014;  @Wijngaarden2015].
 In bariatric surgery patients, connectivity  within the RN (e.g.
-putamen, OFC) might be normalized by the surgery, however the evidence is limited due to a lack of longitudinal obese control groups  [@Duan_2020; @Wiemerslage_2016].
-Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling , e.g.
-changes in ghrelin levels, via hypothalamic-striatal projections [@Li_2019; @Karra_2013; but @Zoon_2018].\
-The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle_2015].
+putamen, OFC) might be normalized by the surgery, however the evidence is limited due to a lack of longitudinal obese control groups  [@Duan2020; @Wiemerslage2016].
+Possibly, a reconfiguration of the fronto-striatal brain networks could emerge from altered gut signaling, e.g.
+changes in ghrelin levels, via hypothalamic-striatal projections [@Li2019; @Karra2013; but @Zoon2018].\
+The DMN includes the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and the inferior and lateral parietal cortex [@Raichle2015].
 Higher BMI and obesity have been associated with a pattern of decreased within DMN FC and increased FC of DMN regions to other networks, i.e.
-salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer_2017; @Kullmann_2011; @Chao_2018; @Wijngaarden_2015; @Borowitz_2020; @Ding_2020; @Doucet_2017; @Sadler_2018].
-After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li_2018; @Olivo_2017; @Frank_2013].\
+salience and sensory networks in several resting-state and task-based rsfMRI studies [@Beyer2017; @Kullmann2011; @Chao2018; @Wijngaarden2015; @Borowitz2020; @Ding2020; @Doucet2017; @Sadler2018].
+After bariatric surgery, a normalization of the connectivity between DMN and cognitive control and salience brain regions might occur, yet no study has included a longitudinal control group [@Li2018; @Olivo2017; @Frank2013].\
 In sum, while there is some evidence hinting to a role for DMN and RN FC in altered regulation of food intake after bariatric surgery, the existing evidence is inconclusive.
-Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George_2016].\
-Further, while higher BMI has been consistently associated with more head motion during rsfMRI [@Beyer_2017;@Hodgson_2016], previous studies in bariatric surgery patients have not taken this important confounder of FC into account.
-In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer_2020].
+Most studies have investigated small cohorts of patients, without adequate obese control groups, and did not rigorously separate confirmatory from exploratory analyses [@George2016].\
+Further, while higher BMI has been consistently associated with more head motion during rsfMRI [@Beyer2017;@Hodgson2016], previous studies in bariatric surgery patients have not taken this important confounder of FC into account.
+In the present sample, we previously reported a group-by-time interaction on head motion [@Beyer2020].
 Thus, we aimed to rigorously control for motion-related variance in our analyses.\
@@
 <!-- STUDY DESIGN --> The ADIPOSITAS-Study investigated the effects of bariatric surgery on brain structure and function in a prospective design at the Charité University Medicine Berlin, Germany.
-For more details see [@Prehn_2020].
+For more details see [@Prehn2020].
 We used all data acquired until April 2019.\
@@
 Inclusion criteria were, in accordance with German guidelines for bariatric surgery, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g.
-type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013].
+type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick2013].
 Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score $< 34$).\
@@
 Participants either underwent surgery (n = `r cs["Sum","IG"]`, `r cs["f","IG"]` females) or were waiting list controls (n = `r cs["Sum","KG"]`, `r cs["f","KG"]` females), who waited for their health insurance's approval to undergo surgery.
-Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski_2016].
+Measures were taken at baseline (BL), $6$ (FU1) and $12$ (FU2) months post-surgery/baseline appointment to capture both phases of rapid weight-loss and maintenance [@Maciejewski2016].
 Analyses were performed on all participants who provided at least one data point of rsfMRI data.
@@
 They underwent medical assessments including an interview, blood draw and anthropometric measurements before having a one hour break for breakfast.
-MRI scanning was done after performing a psychological test battery (for details see @Prehn_2020).
+MRI scanning was done after performing a psychological test battery (for details see @Prehn2020).
 
@@
 MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
-<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = $1 \times 1 \times 1$ mm$^3$, 192 sagital slices).
+<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = $1 \times 1 \times 1$ mm$^3$, 192 sagital slices).
 <!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms.
@@
 T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross].
-Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long].
+Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fslong].
 Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.
@@
 ### Denoising pipelines
-In the pre-registration, we specified two denoising pipelines (AROMA+CC including ICA-AROMA + CompCor and AROMA+CC+GSR including ICA-AROMA + CompCor + global signal regression (GSR)) [@Parkes_2018; @Ciric_2017, for details see SI].
+In the pre-registration, we specified two denoising pipelines (AROMA+CC including ICA-AROMA + CompCor and AROMA+CC+GSR including ICA-AROMA + CompCor + global signal regression (GSR)) [@Parkes2018; @Ciric2017, for details see SI].
 
@@
 **Anatomical quality control**
-FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk_2019.
+FreeSurfer cross-sectional and longitudinal  segmentations were visually checked according to @Klapwijk2019.
 We excluded 5 datasets from 3 participants because of excessive head motion leading to failed pial reconstruction and anatomical-functional coregistration.\
@@
 **RsfMRI quality control**
-RsfMRI quality control was performed according to the protocol by @Ciric_2018 (see SI for more details). 
-Head motion was quantified using mean framewise displacement (mFD) according to @Power_2012, and log-transformed for further analysis (logmFD).
-As pre-registered, we did not exclude anybody based on high average head motion [@Beyer_2020].
+RsfMRI quality control was performed according to the protocol by @Ciric2018 (see SI for more details). 
+Head motion was quantified using mean framewise displacement (mFD) according to @Power2012, and log-transformed for further analysis (logmFD).
+As pre-registered, we did not exclude anybody based on high average head motion [@Beyer2020].
 
@@
 
-Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
+Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al.
 -@spm12).
@@
 **Aggregated FC analysis**
-We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007].
+We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates2007].
 We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline.
@@
 As described in the pre-registration, we performed an exploratory analysis of the relationship of average and change in BMI in a separate group of models.
-As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI.
+As proposed by @Guillaume2014, we calculated the between- and within subject centered values of BMI.
 This model, containing average BMI and BMI change, allowed us to disentangle the differential effects of these variables on FC.
 We first estimated their effects in a model adjusting for baseline age and sex (model EA1) and then additionally controlling for logmFD (model EA2).
-As we previously reported correlated change in BMI and head motion in this sample [@Beyer_2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (model EA3).
+As we previously reported correlated change in BMI and head motion in this sample [@Beyer2019], we explored a refined model including average BMI and logmFD and change in both measures, along with baseline age and sex (model EA3).
 Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.\
@@
 We explored our whole-brain group-by-time point interaction models for the data of all three time points.
-In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo_2017].
+In this model, time was represented as factor taking into account possible non-linear time courses in the increase and decrease of FC over the course of one year, which may occur depending on the phase of weight management [@Olivo2017].
 The resulting factorial design contained one regressor for each time point per group (see SI for depiction of design matrix).
@@
 Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$.
-The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
+The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
 **Aggregated FC analysis**
@@
 ### Functional decoding
-In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the NeuroSynth (https://www.neurosynth.org/) database [@Yarkoni_2011].
+In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the NeuroSynth (https://www.neurosynth.org/) database [@Yarkoni2011].
 We uploaded the contrast images for change BMI and average logmFD on Neurovault (https://neurovault.org/collections/9426/), and applied the decoding classifier.
@@
 In this study, we could not confirm our pre-registered hypotheses.
-Based on previous studies in bariatric surgery patients, we expected within-DMN FC to increase, and DMN FC to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li_2018; @McFadden_2013; @Frank_2013].
-Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat [@Ochner2012_wanting; @Scholtz_2014; @Cerit_2019].
+Based on previous studies in bariatric surgery patients, we expected within-DMN FC to increase, and DMN FC to other somatosensory and attention networks to decrease, in line with more efficient processing of visceral and bodily signals after surgery [@Li2018; @McFadden2013; @Frank2013].
+Further, we expected FC between RN regions to decrease, as bariatric surgery has been previously shown to reduce hyperactivation in RN regions and hedonic motivation to eat [@Ochner2012_wanting; @Scholtz2014; @Cerit2019].
 These studies, notably, did not include adequate longitudinal obese control groups, making false-positive findings possible.
@@
 One the other hand, higher connectivity might also indicate a better integration of hedonic drive and salience processing in action planning.
-Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi_2018].
+Previously, a decrease in local (regional homogeneity and frequency of low-amplitude oscillations) and global connectivity (degree centrality) measures in a similar region of the left SMA was reported after glucose administration [@AlZubaidi2018].
 Reduced connectivity in this region was interpreted as an inhibition of action planning or initiation because of fulfilled energy requirements and reduced need for foraging.
@@
 Overall, our results point to the importance of head motion as a confounder in neuroimaging studies in obesity, challenging definite conclusions on the relationship between weight loss and FC changes.
-Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer_2020], and therefore conducted careful analyses of the impact of head motion on our results.\
+Previously, we reported multi-collinearity between BMI and head motion in this sample [@Beyer2020], and therefore conducted careful analyses of the impact of head motion on our results.\
 To our surprise, the effect of weight loss on the connectivity of the NAcc with the posterior-medial frontal region was only present when separating average mFD and change in mFD and thereby introducing two instead of one regressor into the model.
@@
 Thereby, this results echoes a previous finding from our group, and may be interpreted as accelerated age-related decline of the DMN in relation to the cardiometabolic risk related to BMI.
-Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia_2017].\
+Yet, midline regions are prone to motion artifacts and doubts regarding the complete removal of motion confounding remain [@Savalia2017].\
 The major strength of our study is the prospective intervention controlled design.
-We compared bariatric surgery patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history or recommendation and were scanned after the same time intervals [@Thiese_2014].
+We compared bariatric surgery patients to an obese control group who did not differ in baseline BMI, comorbidities, treatment history or recommendation and were scanned after the same time intervals [@Thiese2014].
 Another strength of our study was the pre-registered analysis plan, which was corrected prior to statistical analysis after we noticed flaws in the first version.
-In particular, we gave more details on denoising pipelins and models and determined that we would use SwE toolbox, an advanced statistical toolbox to deal with longitudinal repeated measures [@Guillaume_2014].
+In particular, we gave more details on denoising pipelins and models and determined that we would use SwE toolbox, an advanced statistical toolbox to deal with longitudinal repeated measures [@Guillaume2014].
 Opposed to the flexible factorial models which is the standard in SPM, marginal models use less degrees of freedom, and thus allow for the inclusion of covariates and higher power.\
@@
 We did not monitor hunger or satiety in our design, although all participants were scanned after the intake of a breakfast following an overnight fast.
-Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer_2013; @Lepping_2015; @Wiemerslage_2016; @Ochner_2012], and might thus have confounded our results [@Li_2019].\
+Hunger feelings and levels of appetite regulating hormones such as insulin and ghrelin have been shown to predict RN responsivity to food cues, as well as resting-state brain organization [@Kroemer2013; @Lepping2015; @Wiemerslage2016; @Ochner2012], and might thus have confounded our results [@Li2019].\
 Size and composition of our sample did not allow sex-stratified analyses.
-However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs_2015; @Chooi_2019].\
+However, the disproportionate sex distribution is reflective of the prevalence differences and under-utilization of bariatric surgery by men [@Fuchs2015; @Chooi2019].\
 Taken together, this prospective well-controlled study did not confirm previous findings claiming strong effects of bariatric surgery on FC of the RN and DMN in obese patients.
